### 2015-06-26 article.tex
first import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,10pt]{llncs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[utf8x]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
%\usepackage{llncsdoc}
\usepackage[english]{babel}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{todonotes}
\usepackage{subfigure}
%\usepackage[mathcal]{euscript}
\usepackage{color}
\newcommand{\rt}{\textcolor{red}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{matrix}
\input{math}
\input{complexity}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \renewcommand{\floatpagefraction}{.9}
% \renewcommand{\textfraction}{.1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\PET}{pet}
\newcommand{\SBL}{sbl}
\newcommand{\horiz}{{\centering\rule{350pt}{.4pt}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\makeatletter
\newcommand{\pushright}[1]{\ifmeasuring@#1\else\omit\hfill$\displaystyle#1$\fi\ignorespaces}
\newcommand{\pushleft}[1]{\ifmeasuring@#1\else\omit$\displaystyle#1$\hfill\fi\ignorespaces}
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{%
Obtaining a Triangular Matrix by Independent Row-Column Permutations
}%

\author{%
  Guillaume Fertin\inst{1} \and
  Irena Rusu\inst{1} \and
  St\'ephane Vialette\inst{2}
}% end author
\institute{%
  LINA UMR CNRS 6241, Universit\'e de Nantes, France \\
  \email{\{guillaume.fertin,irena.rusu\}@univ-nantes.fr}
  \and
  Universit\'e Paris-Est, LIGM (UMR 8049), CNRS, UPEM, ESIEE Paris, ENPC, 
  F-77454, Marne-la-Vallée, France\\
  \email{vialette@univ-mlv.fr}
}% end institute
\date{\today}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
Given a square $(0,1)$-matrix $A$,
we consider the problem of deciding whether there exists a permutation
of the rows and a permutation of the columns of 
$A$ such that after carrying out these permutations,
the resulting matrix is triangular. The complexity of the problem was
posed as an open question by Wilf~\cite{Wilf:1997} in 1997.
%(\emph{i.e.}, do there exist permutation matrices $P$ and $Q$ such that
%$PAQ$ is triangular)?
In 1998, DasGupta et al.~\cite{DasGuptaJiangKannanLiSweedyk:1998} seemingly answered the
question, proving it is \NPC. However, we show here that their result is flawed, which 
leaves the question still open. Therefore, we give a definite answer
to this question by proving that the problem is \NPC. We finally present an
exponential-time algorithm for solving the problem.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%%
%% ---- Introduction ----
%%
\section{Introduction}
\label{section:Introduction}

In his contribution in the tribute to the late Professor
Erd\"os~\cite{Wilf:1997}, Wilf posed the following question:
\emph{
  ``Let $A$ be a $m \times n$ matrix of $0$'s and $1$'s. Consider the
  computational problem: do there exist permutations $P$ of the rows
  of $A$, and $Q$, of the columns of $A$ such that after carrying out
  these permutations, $A$ is triangular? The question we ask concerns
  the complexity of the problem.
  Is this problem \NPC?
  Or, does there exist a polynomial-time algorithm for doing it?''}
As noted by Wilf,
this problem is strongly related to job scheduling with precedence
constraints, a well-known problem in theoretical computer science.
The present paper is devoted to giving an answer to this question.

A square matrix is called {\em lower triangular} if all the entries above the main diagonal are zero.
Similarly, a square matrix is called {\em upper triangular} if all the entries below the main diagonal
are zero.
A {\em triangular matrix} is one that is either lower triangular or upper triangular.
Because matrix equations with triangular matrices are easier to solve,
they are very important in  linear algebra and numerical analysis.
We refer the reader to \cite{MatrixComputations:1996} for a further discussion.

For an arbitrary square matrix $A$, it is well-known in linear algebra
that there exists an invertible matrix $S$ such that $S^{-1}AS$ is
upper triangular.
We focus here, however, on permutation matrices.
Recall that a permutation matrix is a square matrix obtained from the same
size identity matrix by a permutation of rows. 
A product of permutation matrices is again a permutation matrix and
the inverse of a permutation matrix is again a permutation matrix.
In fact, for any permutation matrix $P$, $P^{-1} = P^T$.

% Nicolson [5] shows that for a nonnegative matrix $A$ the following four conditions
% are equivalent:
% (i) there exists a permutation matrix $P$ such that
% $P A P^T$ is strictly upper triangular;
% (ii) there is no positive cycle in $M$
% (i.e. in the weighted digraph there is no cycle whose edges all have positive weight);
% (iii) the permanent of $M + I$ is $1$; and
% (iv) $A$ is nilpotent.

This paper is organized as follows.
In Section~\ref{section:Notations}, we provide the basic material
needed for this paper. %with a
%strong emphasis on the associated combinatorial structures: configurations of subsets,
%bipartite graphs and directed graphs.
Section~\ref{section:Hardness result} is devoted to proving hardness of
determining whether a square $(0,1)$-matrix is permutation equivalent
triangular, i.e. whether it can be transformed into a triangular matrix by independent
row and column permutations.
In Section~\ref{section:Exponential-time algorithm}, we give some properties of
permutation equivalent triangular matrices (or \PET\ matrices, for short) and present 
an exponential-time algorithm to determine whether a matrix is a \PET\ matrix.
The paper concludes with suggestions for further research directions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Preliminaries ----
%%
\section{Notations}
\label{section:Notations}

%\subsection{$(0,1)$-matrices}
%\subsubsection{$(0,1)$-matrices} 
For any positive integer $n$, denote $[n]=\{1, 2, \ldots, n\}$. 
Let $A = [a_{i,j}]$, $1 \leq i \leq m$ and $1 \leq j \leq n$,
be a matrix of $m$ rows and $n$ columns.
% We say that $A$ is of size $m$ by $n$ and we also refer to $A$
% as an $m$ by $n$ matrix.
In the case that $m = n$ then the matrix is \emph{square} of \emph{order} $n$.
It is always assumed that the entries of the matrix are elements
of some underlying field $F$.
It is convenient to refer to either a row or a column of the matrix as a
\emph{line} of the matrix.
We use the notation $A^T$ for the \emph{transpose} of matrix $A$.
We always designate a zero matrix by $\zeros$, a matrix with every entry equal to
$1$ by $J$, and the identity matrix of order $n$ by $I$.
In order to emphasize the size of these matrices we sometimes include subscripts.
Thus $J_{m,n}$ denotes the all $1$'s matrix of size $m$ by $n$,
and this is abbreviated to $J_n$ if $m = n$. 
Notations $\zeros_{m,n}$, $\zeros_n$ and $I_n$ are similarly defined.
In displaying a matrix we often use $*$ to designate a submatrix of no particular structure.
Two matrices $A$ and $B$ are said to be \emph{permutation equivalent},
denoted by $A \sim B$, if there exist permutation matrices $P$ and $Q$ of
suitable sizes such that $B = PAQ$.

We will be greatly concerned with matrices whose entries consist exclusively of
the integers $0$ and $1$.
Such matrices are referred to as \emph{$(0,1)$-matrices}.
For a $(0,1)$-matrix $A$, we let $\omega(A)$ stand for the number of $1$'s in $A$.
A square matrix $A = [a_{i,j}]$ of order $n$ is said to be
\emph{lower left triangular} if it has only $0$'s above the main diagonal
(\emph{i.e.} $a_{i,j} = 0$ for $1 \leq i < j \leq n$.
We write $\lltm_{\,n}$ for the lower left triangular $(0,1)$-matrix 
whose $0$'s are exclusively above the main diagonal.
For two matrices $A = [a_{i,j}]$ and $B = [b_{i,j}]$ of size $m$ by $n$,
we write $A \leq B$ if $a_{i,j} \leq b_{i,j}$ for $1 \leq i \leq m$
and $1 \leq j \leq n$, so that a square matrix $A$ of order $n$ is
lower left triangular if $A \leq \lltm_{\,n}$.
In the context of permutation equivalent matrices,
we will sometimes not be interested in any particular orientation of a
triangular matrix and forget about any
specific orientation such as ``\emph{lower left}''.
Furthermore, for readability, a matrix which is permutation equivalent
to a triangular matrix is said to be a \emph{\PET\ matrix}.
The \emph{row sum vector}
$\RSV(A) = \begin{bmatrix} r_1 & r_2 & \ldots & r_m \end{bmatrix}$
and the  \emph{column sum vector}
$\CSV(A) = \begin{bmatrix} c_1 & c_2 & \ldots & c_n \end{bmatrix}$
of $A$ are defined by
$r_i = \sum_{1 \leq j \leq n} a_{i,j}$ for $1 \leq i \leq m$
and
$c_j = \sum_{1 \leq i \leq m} a_{i,j}$ for $1 \leq j \leq n$.
The row sum vector $\RSV(A)$
(resp. column sum vector $\CSV(A)$)
is \emph{stepwise bounded}
if  $|\{i : r_i \leq k \}| \geq k$
(resp. $|\{j : c_j \leq k \}| \geq k$)
for $1 \leq k \leq n$.
It is clear that if a $(0,1)$-matrix $A$ is a \PET\ matrix then
both $\RSV(A)$ and $\CSV(A)$ are stepwise bounded.

% The definition of the \emph{permanent} of $A$ follows the classical definition of the
% determinant but with a simplification. Ironically, this simplification in the formula makes
% it more difficult to compute the permanent.
The permanent of $A = [a_{i,j}]$ is defined as the number given by the formula
$\PER(A) = \sum_{(j_1, j_2, \ldots, j_n) \in S_n}
a_{1,j_1}\,a_{2,j_2}\, \ldots a_{n,j_n}$,
where the summation is over all permutations
$(j_1,j_2, \ldots, j_n)$ of $[n]$.
Observe that, unlike the determinant, we do not put a minus sign in front of some of the terms in the
summation.
Of particular importance,
the permanent does not change if we permute the rows of $A$ and
permute the columns of $A$.

Let $A = [a_{i,j}]$ be an $m$ by $n$ matrix.
For convenience, for a set $K \subseteq [m]$ we will
write $\overline{K}$ for the set
$[m] \setminus K$.
Let $K = \{i_1, i_2, \ldots, i_k\}$ be a set of $k$ elements with
$K \subseteq [m]$,
and let $L = \{j_1, j_2, \ldots, j_l\}$ be a set of
$l$ elements with $L \subseteq [n]$.
The sets $K$ and $L$ designate a collection of row indices and column indices,
respectively, of the matrix $A$, and the $k$ by $l$ submatrix determined by them is
denoted $A[K,L]$.
% $$
% A[K,L]=
% \begin{bmatrix}
% a_{i_1,j_1} & a_{i_1,j_2} & \dots  & a_{i_1,j_l} \\
% a_{i_2,j_1} & a_{i_2,j_2} & \dots  & a_{i_2,j_l} \\
% \vdots      & \vdots      & \ddots & \vdots      \\
% a_{i_k,j_1} & a_{i_k,j_2} & \dots  & a_{i_k,j_l}
% \end{bmatrix}\text{.}
% $$
% If $L = K$, then $A[K, K]$ is a \emph{principal} submatrix of $A$,
% sometimes denoted more simply as $A[K]$.

%\subsection{Related combinatorial objects}
%\subsubsection{Configuration of subsets}

% Combinatorial matrix theory \cite{Brualdi:Ryser:1991} includes a wide variety of 
% combinatorial areas, e.g. graph theory \cite{Diestel:2000}, transversal theory
% \cite{Mirsky:1971}, combinatorial configurations \cite{Ryser:1969}.
% We shall make intensive use of configuration of subsets.

Let $X = \{x_i : 1 \leq i \leq n\}$ be a non-empty set of $n$
elements, that we call an \emph{$n$-set}.
Let $\mathcal{S} = (S_i : 1 \leq i \leq m)$ be $m$ not necessarily
distinct
subsets of the $n$-set $X$. We refer to this collection of subsets of an
$n$-set as a \emph{configuration of subsets}.
We set $a_{i,j} = 1$ if $x_j \in S_i$, and $a_{i,j} = 0$
if $x_i \notin S_i$.
The resulting $(0,1)$-matrix $A = [a_{i,j}]$, $1 \leq i \leq m$ and
$1 \leq j \leq n$ of size $m$ by $n$ is the \emph{incidence matrix}
for the configuration of subsets $\mathcal{S}$ of the $n$-set $X$.
The $1$s in row $\alpha_i$ of $A$ display the elements in the subset
$S_i$, and the $1$'s in column $\beta_j$ display the occurrences of $x_j$
among the subsets. 
Let $\mathcal{S} = (S_i : 1 \leq i \leq n)$ be a configuration of subsets of some 
ground $n$-set $X$. A bijective mapping $\varphi : \mathcal{S} \to [n]$ is said to be a
\emph{stepwise bounded labeling} (or \SBL\ for short) of $\mathcal{S}$ if
$\left|\bigcup_{\varphi(S_j) \leq i}S_j\right| \leq i$
for $1 \leq i \leq n$.

%  A \emph{system of distinct representatives} (SDR) for the family $\mathcal{S}$
% is a family $(x_1, x_2, \ldots, x_n)$ of elements of $X$ satisfying
% the conditions
% (i) $x_i \in S_i$ for $1 \leq i \leq n$, and
% (ii) $x_i \neq xj$ for $i \neq j$.
% (The first condition asserts that the elements are representatives of the sets,
% and the second that they are distinct.)
% The collection $\mathcal{S}$ satisfies the \emph{marriage} condition
% if and only if for each subcollection $\mathcal{S}' \subseteq \mathcal{S}$, we have
% $|\mathcal{S}'| \leq \left|\bigcup_{S \in \mathcal{S}'} S\right|$.
% In other words, the number of sets in each subcollection $\mathcal{S}'$ is less
% than or equal to the number of distinct elements in the union over the
% subcollection $\mathcal{S}'$.
% Hall's theorem states that $\mathcal{S}$ has a SDR if and only if
% $\mathcal{S}$ satisfies the marriage condition.

% \paragraph{Bipartite graphs.}

% A graph $G=(V,E)$ is \emph{bipartite} if 
% $V$ can be partitioned into two sets $U$ and $W$ such that each edge
% joins a vertex in $U$ to a vertex in $W$, and the pair $(U, W)$ is called a
% \emph{bipartition} of $V$ (or of $G$).
% The \emph{reduced adjacency matrix} $B = [b_{i,j}]$ is an $m \times n$
% $(0,1)$-matrix defined by $b_{i, j} = 1$ if and only if
% there is an edge beteween vertex $i$ of $U$ and vertex $j$ of $W$.
% A matching $M$
% in $G$ is a collection of edges no two of which have a vertex in
% common. If $v$ is a vertex and there is an edge of $M$ containing $v$,
% then $v$ \emph{meets} the matching $M$ and the matching $M$ meets the vertex
% $v$. A \emph{perfect matching} of $G$ is a matching
% that meets all vertices of $G$.

% \paragraph{Directed graphs.}

% A \emph{directed graph} (for short, a \emph{digraph}) $G$
% consists of a finite set $V$ of elements called vertices and a set $E$ of
% ordered pairs of vertices called arcs.
% The \emph{adjacency matrix} associated to $G$ is the $|V| \times |V|$
% $(0,1)$-matrix $A = [a_{i,j}]$ defined by $a_{i,j} = 1$ if and only if
% there is an arc from vertex $i$ to vertex $j$ in $G$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Properties ----
%%
\section{Answering Wilf's question}%Hardness result}
\label{section:Hardness result}

We prove in this section that, given a square $(0,1)$-matrix $A$, deciding whether there 
exists a permutation matrix $P$ and a permutation matrix $Q$ of suitable size
such that $PAQ$ is triangular is \NPC.

\subsection{Disproving a previous related result}

Before giving our proof, it is worth mentioning that the following problem
(called LBQIS$(n,k)$ and rephrased to fit the context of this paper) is
claimed to be \NPC in~\cite{DasGuptaJiangKannanLiSweedyk:1998}:
Given a $(0,1)$-matrix of order $n$ and positive integer $k \leq n$,
do there exist permutation matrices $P$ and $Q$ such that
$
PAQ =
\begin{bmatrix}
A_{1,1} & A_{1,2} \\
A_{2,1} & A_{2,2}
\end{bmatrix}
$
with $A_{1,2}$ a square lower triangular matrix of size $k$ by $k$?
It is not very difficult to find a polynomial transformation from
LBQIS to Wilf's question, which would prove the \NPcompleteness of the
latter. Just add $n-k$ empty rows and $n-k$ empty columns 
to matrix $A$ to obtain a new matrix $A'$. Now, notice that each 
submatrix $A_{1,2}$ in a solution for LBQIS may be completed with the 
$n-k$ empty rows put before row 1 of $A_{1,2}$ and with the $n-k$ empty 
columns put after column $k$ of $A_{1,2}$ to yield a solution for the 
instance $A'$ in Wilf's question, and viceversa.

Unfortunately, paper~\cite{DasGuptaJiangKannanLiSweedyk:1998} 
contains a serious flaw in the proof. To fix things, note that
in~\cite{DasGuptaJiangKannanLiSweedyk:1998} LBQIS is stated in
terms of bipartite graphs, for which matrix $A$ is the reduced adjacency
matrix. Then, LBQIS$(n,k)$ is proved \NPC by reduction 
from another problem on bipartite graphs called LBIS$(n,k)$, using the so-called 
Rearrangement Lemma (Lemma 3.5 in ~\cite{DasGuptaJiangKannanLiSweedyk:1998}).
Two affirmations in the proof of this lemma are contradicted by the following
example. Let $G$ be the graph (input for LBIS) with vertices $U=\{i\, |\, 1\leq i\leq 4\}$
and $V=\{i\, |\, 1\leq i\leq 4\}$, whose edges are $(1,1), (2,1), (2,2),
(3,2), (3,4), (4,3)$ and $(4,4)$. Thus, $n=4$. Define $k=1$.  
Let $G'$ be the input graph for LBQIS built as in~\cite{DasGuptaJiangKannanLiSweedyk:1998},
and $k'=k^2+k=2$. Then the vertex subset $U'\cup V'$ of $G'$, with $U'=\{[2,4], [1,2]\}$
and $V'=\{[1,3], [2,2]\}$ is a solution of LBQIS of size $k'$ for which the
assumption on the first line of the Rearrangement Lemma's proof is false.
Also, the vertex subset $U'\cup V'$ of $G'$, with $U'=\{[1,1], [2,1]\}$
and $V'=\{[1,2], [1,3]\}$ is a solution of LBQIS of size $k'$ for which
the second affirmation in the same lemma (``clearly $q_1\leq p_1$'') is also false.

\subsection{Our \NPcompleteness proof for Wilf's question}

We present our results in terms of \SBL\ for configurations
of subsets. The rationale for considering \SBL\ for configurations
of subsets stems from the following lemma.

\begin{lemma}
    \label{lemma:bounded stepwise property <=> PAQ triangular}
    Let $\mathcal{S} = (S_i : 1 \leq i \leq n)$ be a configuration of subsets of
    some ground $n$-set, and let $A$ be the corresponding incidence matrix.
    There exist permutation matrices $P$ and $Q$ of order $n$
    such that $PAQ \leq \lltm_{\,n}$ if and only if
    there exists an \SBL\ of $\mathcal{S}$.
\end{lemma}

We need to focus our attention on a special type of \SBL. Call a bijective mapping
$\varphi : \mathcal{S} \to [n]$ \emph{normalized} if $\varphi$
maps the identical subsets of elements of $\mathcal{S}$
to a set of consecutive integers.
Most of the interest in normalized bijective labelings
stems from the following intuitive lemma.

\begin{lemma}
  \label{lemma:normalized stepwise bounded}
  Let $\mathcal{S} = (S_i : 1 \leq i \leq n)$ be a configuration of subsets of
    some ground $n$-set.
    If there exists an \SBL\ of $\mathcal{S}$
    then there exists a normalized \SBL\ of $\mathcal{S}$.
\end{lemma}

We are now ready to prove that
deciding whether there exists an \SBL\ of some
configuration of subsets is \NPC thereby proving that 
deciding whether a square $(0,1)$-matrix is a \PET\ matrix is \NPcomplete as well.
The proof proceeds by a reduction from the \PB{3Sat} problem - which is a known
\NPcomplete problem \cite{Cook:1971}.
% (see also
% \cite{Garey:Johnson:1979} and \cite{Papadimitriou:1994} for variants).
Let an arbitrary
instance of the \PB{3Sat} problem be given by a $3$CNF formula 
$\phi = c_1 \vee c_2 \vee \ldots \vee c_m$ over variables 
$x_1, x_2, \ldots, x_n$. 
Our construction is divided into two steps: 
(1) construction of a (polynomial size) ground set $\mathbf{X}$ and 
(2) construction of a configuration of subsets $C$ of the ground set
$\mathbf{X}$. 
Throughout the proof, parts of the ground set
$\mathbf{X}$ are written as capital bold letters 
($\mathbf{V}, \mathbf{T}, \mathbf{F}, \ldots$)
and subsets of the configuration are written with capital calligraphic
letters ($\mathcal{V}_i, \mathcal{T}_i, \mathcal{F}_i, \ldots$).

  \begin{figure}[H]
    \centering
    \begin{minipage}[c]{\textwidth}
      \centering
      \input{r.pstex_t}
      \caption{\label{fig:stepwise bounded labeling construction}
          Illustration of the construction 
          % used in the proof of 
          % Proposition~\ref{proposition:Stepwise Bounded Labeling is NPC} 
          for the $3$CNF formula 
          $\phi = (x_1 \vee x_2 \vee x_3) \wedge 
          (\overline{x}_1 \vee \overline{x}_2 \vee x_3) \wedge 
          (\overline{x}_1 \vee x_2 \vee \overline{x}_3)$. 
          Identical subsets are not distinguishable
          in our representation. A satisfying truth assignment is
          given by $f(x_1) = \mbox{TRUE}$, $f(x_2) = \mbox{FALSE}$ and
          $f(x_3) = \mbox{FALSE}$. For sake of clarity, neither
          the ground set $\mathbf{X}$ nor the collection of subsets
          $C$ is fully represented. 
          % Note that we have here dropped the
          % assumption that each clause is satisfied by its first
          % literal
          } 
    \end{minipage}%
  \end{figure}

To begin with, define 
$p_i = 3(n + m + 1 - i) + 2$,
$q_i = 3(n + m + 1 - i) + 1$ and
$r_i = 3(n + m + 1 - i)$ for $1 \leq i \leq n + m$.
Furthermore, define $p_{n+m+1}=1$,
$K = \sum_{i=1}^{n} q_i + 2 \sum_{i=n+1}^{n+m} q_i$ and
$L = \sum_{i=1}^{n+m} (p_{i+1} + r_i)$.
Let us now define the ground set $\mathbf{X}$. 
Consider the pairwise disjoint sets defined as follows:
$\mathbf{V}_i = \{v_{i, j} \mid 1 \leq j \leq p_i\}$,
$\mathbf{V}'_i  = \{v'_{i, j} \mid 1 \leq j \leq r_i\}$,
$\mathbf{T}_i = \{t_{i, j} \mid 1 \leq j \leq q_i\}$,
$\mathbf{F}_i = \{f_{i, j} \mid 1 \leq j \leq q_i\}$
for $1 \leq i \leq n$. Furthermore, define
$\mathbf{C}_i = \{c_{i, j} \mid 1 \leq j \leq p_{n + i}\}$
$\mathbf{C}'_i = \{c'_{i, j} \mid 1 \leq j \leq r_{n + i}\}$
for $1 \leq i \leq m$, and
$\mathbf{L}_{i, k} = \{\ell_{i, k, j} \mid 1 \leq j \leq q_{n + i}\}$
for $1 \leq i \leq m$ and $1 \leq k \leq 3$.
Finally, define $\mathbf{S} = \{s\}$.
For simplicity of notation, write
$\mathbf{V} = \bigcup_{1 \leq i \leq n} \mathbf{V}_i$,
$\mathbf{V}'= \bigcup_{1 \leq i \leq n} \mathbf{V}'_i$,
$\mathbf{T} = \bigcup_{1 \leq i \leq n} \mathbf{T}_i$,
$\mathbf{F} = \bigcup_{1 \leq i \leq n} \mathbf{F}_i$,
$\mathbf{C} = \bigcup_{1 \leq i \leq m} \mathbf{C}_i$,
$\mathbf{C}' = \bigcup_{1 \leq i \leq m} \mathbf{C}'_i$,
and
$\mathbf{L}_i = \bigcup_{1 \leq k \leq 3} \mathbf{L}_{i, k}$
for $1 \leq i \leq m$
and
$\mathbf{L} = \bigcup_{1 \leq i \leq m} \mathbf{L}_i$.
Informally, elements of $\mathbf{V} \cup \mathbf{V}'$ are associated to variables, 
elements of $\mathbf{T} \cup \mathbf{F}$ are associated to literals, elements of 
$\mathbf{C} \cup \mathbf{C}'$ are associated to clauses, elements of $\mathbf{L}$ 
are associated to literals in clauses and $\mathbf{S}$ is a  
\emph{separator set}. The ground set $\mathbf{X}$ of our construction is defined to be
$\mathbf{X} = 
\mathbf{V} \cup \mathbf{V}' \cup \mathbf{T} \cup 
\mathbf{F} \cup \mathbf{C} \cup \mathbf{C}' \cup 
\mathbf{L} \cup \mathbf{S}$.

Having defined the ground set $\mathbf{X}$, 
we now turn to the detailed construction of a configuration of subsets $C$ of $\mathbf{X}$. 
For sake of clarity, this will be divided into several steps. 
First, each variable $x_i$, $1 \leq i \leq n$, is associated to identical subsets 
$\mathcal{V}_{i, j}$, $1 \leq j \leq q_i$, in $C$. These subsets are defined as follows:
$\mathcal{V}_{i, j} = 
\left( \bigcup_{1 \leq k \leq i} \mathbf{V}_k \right) \cup 
\left( \bigcup_{1 \leq k \leq i-1} \mathbf{V}'_k \right)$
for $1 \leq i \leq n$ and $1 \leq j \leq q_i$.
Let us denote by $\mathcal{V}_i$, $1 \leq i \leq n$, 
the collection $(\mathcal{V}_{i, j} \mid 1 \leq j \leq q_i)$.
Next, each (positive) literal $x_i$, $1 \leq i \leq n$, 
is associated to identical subsets $\mathcal{T}_{i, j}$, $1 \leq j \leq r_i$, 
and to identical subset $\mathcal{T}'_{i, j}$, $1 \leq j \leq p_{i + 1}$. 
These subsets are defined as follows:
$\mathcal{T}_{i, j} = \mathbf{T}_i \cup 
\left( \bigcup_{1 \leq k \leq i} \mathbf{V}_k \right) \cup 
\left( \bigcup_{1 \leq k \leq i-1} \mathbf{V}'_k \right)$
for $1 \leq i \leq n$ and $1 \leq j \leq r_i$, and
$\mathcal{T}'_{i, j} = \mathbf{T}_i \cup \left( 
\bigcup_{1 \leq k \leq i} \mathbf{V}_k \right) \cup 
\left( \bigcup_{1 \leq k \leq i} \mathbf{V}'_k \right)$
for $1 \leq i \leq n$ and $1 \leq j \leq p_{i+1}$.
Of course, a similar construction of subsets applies for the negation 
$\overline{x}_i$ of each variable $x_i$, \emph{i.e.},
$\mathcal{F}_{i, j} = 
\mathbf{F}_i \cup \left( \bigcup_{1 \leq k \leq i} 
\mathbf{V}_k \right) \cup \left( \bigcup_{1 \leq k \leq i-1} 
\mathbf{V}'_k \right)$ for $1 \leq i \leq n$ and $1 \leq j \leq r_i$, and
$\mathcal{F}'_{i, j} = \mathbf{F}_i \cup \left( 
\bigcup_{1 \leq k \leq i} \mathbf{V}_k \right) \cup 
\left( \bigcup_{1 \leq k \leq i} \mathbf{V}'_k \right)$ 
for $1 \leq i \leq n$ and $1 \leq j \leq p_{i+1}$.
For readability, write 
$\mathcal{T}_i = (\mathcal{T}_{i, j} \mid 1 \leq j \leq r_i)$, 
$\mathcal{T}'_i = (\mathcal{T}'_{i, j} \mid 1 \leq j \leq p_{i+1})$, 
$\mathcal{F}_i = (\mathcal{F}_{i, j} \mid 1 \leq j \leq r_i)$ and 
$\mathcal{F}'_i = (\mathcal{F}'_{i, j} \mid 1 \leq j \leq p_{i+1})$ for $1 \leq i \leq n$. 
Note that the following (strict) inclusions hold for all $1 \leq i \leq n$, 
$1 \leq j_1 \leq q_i$, $1 \leq j_2 \leq r_i$ and $1 \leq j_3 \leq p_{i+1}$: 
(i) $\mathcal{V}_{i, j_1} \subset \mathcal{T}_{i, j_2} \subset \mathcal{T}'_{i, j_3}$ and 
(ii) $\mathcal{V}_{i, j_1} \subset \mathcal{F}_{i, j_2} \subset \mathcal{F}'_{i, j_3}$.
We now turn to the $m$ clauses of the $3$CNF formula. 
Each clause $c_i$, $1 \leq i \leq m$, is associated to identical subsets 
$\mathcal{C}_{i, j}$, $1 \leq j \leq q_{n+i}$. These subsets are defined as follows:
$\mathcal{C}_{i, j} = \mathbf{V} \cup \mathbf{V}' \cup 
\left( \bigcup_{1 \leq k \leq i} \mathbf{C}_k \right) 
\cup \left( \bigcup_{1 \leq k \leq i-1} \mathbf{C}'_k \right)$
for $1 \leq i \leq m$ and $1 \leq j \leq q_{n+i}$.
Let us denote by $\mathcal{C}_i$, $1 \leq i \leq m$, 
the collection $(\mathcal{C}_{i, j} \mid 1 \leq j \leq q_{n+i})$. 
It is easily seen that $\mathcal{V}_{i, j_1} \subset \mathcal{C}_{k, j_2}$ 
for all $1 \leq i \leq n$, $1 \leq j_1 \leq q_i$, $1 \leq k \leq m$ and 
$1 \leq j_2 \leq q_{n+k}$.

Now, we consider the only part of the construction that depends on which 
literal occurs in which clauses. Denote by $\lambda_{i, k}$ the $k$-th literal 
of clause $c_i$, 
that is write $c_i = \lambda_{i, 1} \vee \lambda_{i, 2} \vee \lambda_{i, 3}$ 
for $1 \leq i \leq m$,
where each $\lambda_{i, k}$ is a variable or its negation. 
The $k$-th literal, $1 \leq k \leq 3$, of each clause $c_i$, $1 \leq i \leq m$, 
is associated to identical subsets $\mathcal{L}_{i, k, j}$, $1 \leq j \leq r_{n+i}$, 
and to identical subsets $\mathcal{L}'_{i, k, j}$, $1 \leq j \leq p_{n+i+1}$. 
These subsets are defined as follows:
$\mathcal{L}_{i, k, j} = \mathbf{V} \cup \mathbf{V}' \cup 
\mathbf{A}_k \cup \mathbf{L}_{i, k} \cup 
\left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}_\ell \right) \cup 
\left( \bigcup_{1 \leq \ell \leq i-1} \mathbf{C}'_\ell \right)$
for $1 \leq i \leq m$, $1 \leq j \leq r_{n+i}$ and $1 \leq k \leq 3$ and
$\mathcal{L}'_{i, k, j} = \mathbf{V} \cup \mathbf{V}' \cup \mathbf{A}_k 
\cup \mathbf{L}_{i, k} \cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}_\ell \right) 
\cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}'_\ell \right)$ 
for $1 \leq i \leq m$, $1 \leq j \leq p_{n+i+1}$ and $1 \leq k \leq 3$, 
where $\mathbf{A}_k = \mathbf{T}_\ell$ if $\lambda_{i, k} = x_\ell$ 
and $\mathbf{A}_k = \mathbf{F}_\ell$ if $\lambda_{i, k} = \overline{x}_\ell$. 
For the sake of clarity, write 
$\mathcal{L}_{i, k} = (\mathcal{L}_{i, k, j} \mid 1 \leq j \leq r_{n+i})$ and 
$\mathcal{L}'_{i, k} = (\mathcal{L}'_{i, k, j} \mid 1 \leq j \leq p_{n+i+1})$ 
for $1 \leq i \leq m$ and $1 \leq k \leq 3$. 
Again, observe that 
$\mathcal{C}_{i, j_1} \subset \mathcal{L}_{i, k, j_2} \subset \mathcal{L}_{i, k, j_2}$ 
for all $1 \leq i \leq m$, $1 \leq j_1 \leq q_{n+i}$, $1 \leq j_2 \leq r_{n+i}$, 
$1 \leq j_3 \leq p_{n+i+1}$ and $1 \leq k \leq 3$.

Our construction ends with $p_1 + K - 1$ \emph{utility subsets}. 
These subsets will be partitioned into two separate classes according to their 
indented function: \emph{bootstrap subsets} and \emph{separator subsets}. 
First, $C$ contains identical bootstrap subsets 
$\mathcal{B}_i$, $1 \leq i \leq p_1 - 1$, defined as follows:
$\mathcal{B}_i = \emptyset$ for $1 \leq i \leq p_1 - 1$. The idea is to force any \SBL\ to map the $p_1 - 1$ 
empty sets of $\mathcal{B}$ to the first $p_1  - 1 = 3(n + m) + 1$ integers. 
Indeed, it is easily seen that all the above defined subsets of the configuration 
of subsets $C$ but those of $\mathcal{B}$ contain at least $p_1$ elements and 
hence cannot be mapped to an integer $i \leq p_1 - 1$ in any \SBL\ of 
$C$. 
Second, $C$ contains identical separator subsets $\mathcal{S}_i$, $1 \leq i \leq K$, 
defined by:
$\mathcal{S}_i = 
\mathbf{V} \cup \mathbf{V}' \cup \mathbf{C} \cup \mathbf{C}' \cup \mathbf{S}$
for $1 \leq i \leq K$.
The rationale of these subsets is that we need a separator between subsets in $C$ 
corresponding to a satisfying truth assignment $f$ for the $3$CNF formula $\phi$ 
and garbage subsets of $C$, that is subsets not involved in the satisfying truth 
assignment $f$. For simplicity, let us denote by $\mathcal{B}$ the collection 
$(\mathcal{B}_i \mid 1 \leq i \leq p_1 - 1)$ and by $\mathcal{S}$ 
the collection $(\mathcal{S}_i \mid 1 \leq i \leq K)$. 
Clearly our construction can be carried on in polynomial time: indeed, we have
$|\mathbf{X}| = O(m^2 + n^2)$ and
$|C| = O(m^2 + n^2)$.

\begin{lemma}
  \label{lemma:Stepwise Bounded Labeling is NPC}
  There exists a satisfying truth assignment $f$ for $\phi$ if and only if 
  there exists an \SBL\ of the configuration of subsets 
  $C$ of the ground set $\mathbf{X}$.
\end{lemma}

The key elements of the proof are as follows.
First, it is crucial to focus on solutions that map identical subsets 
of elements of $\mathcal{S}$ to a set of consecutive elements
(see Lemma~\ref{lemma:normalized stepwise bounded}).
Second,
the general shape of the solution is largely guided by the construction.
Indeed,
the empty subsets have to be placed first, followed by subsets corresponding to literals 
(either the positive or the negative literal of each variable has been
chosen)
and next by subsets corresponding to clauses
(one satisfying literal of each clause is chose).
Finally the separator subsets have to be placed, with the result that 
(thanks to the large polynomial number of such subsets)
the remaining subsets can be placed in any order without violating
the sought \SBL\ property.
The reader is invited to consider 
Figure~\ref{fig:stepwise bounded labeling construction}
for a schematic illustration of the reduction.
We now briefly discuss, in an informal way, the two key arguments that are used in 
the proof.
First, the whole procedure is, to some extent, similar to 
the accounting method used in amortized complexity analysis. 
Indeed, one might view the operation of placing a set (one after the other)
as the process of charging some customer, the cost being the number of
new elements that are introduced.
With this metaphor in mind, notice that 
we do not charge when a subset does not introduce any new element,
so that the leftover amount can be stored as "credit". 
When we place a new subset that does introduce some new elements, 
we can use the "credit" stored to pay for the cost of the operation.
Second, when a subset uses the "credit" stored to pay the cost
of introducing new elements, the following invariants can be shown 
to hold true:
(i) it uses all the available credit and
(ii) it does not allow to accumulate 
(it should be now clear that consecutive identical subsets do allow
for accumulating credit)
as much credit as it has consumed,
thereby proving that subsets introduce less and less new elements as we progress
adding subsets one after the other.

The main result of this paper can now be stated.

\begin{theorem}
  \label{corollary:PAQ triangular is NPC}
  Let $A$ be a $(0,1)$-matrix.
  Deciding whether $A$ is a 
  \PET\ matrix is \NPC.
\end{theorem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Exponential-time algorithm ----
%%
\section{Exponential-time algorithm}
\label{section:Exponential-time algorithm}

We present here an exponential-time algorithm for deciding whether a
given a $(0,1)$-matrix $A$ of order $n$ is a \PET\ matrix.
We start by presenting some basic properties of square $(0,1)$-matrices
that can be transformed into some triangular matrix by row and
column independent permutations to help solving involved 
algorithmic issues. We of course focus of polynomial-time checkable
properties.

We first focus on the permanent of a square $(0,1)$-matrix.
A well-known result (see e.g. \cite{Brualdi:Ryser:1991})
states that for a $(0,1)$-matrix $A$ of order $n$,
one has $\PER(A)=1$ if and only if the lines of $A$
may be permuted to yield a triangular matrix with $1$'s
in the $n$ main diagonal positions and $0$'s above the mail diagonal.
This theorem amounts to saying that
$\PER(A)=1$ if and only if there exist permutation matrices $P$ and $Q$
such that $I \leq PAQ \leq \lltm$.
As shown in the following lemma, $\PER(A)=1$ is certainly 
a threshold value in our context.

\begin{lemma}
    \label{lemma:permanent}
Let $A$ be $(0,1)$-matrix.
If $A$ is a \PET\  matrix
then $\PER(A) \leq 1$.
\end{lemma}

Notice that deciding $\PER(A) \leq 1$ for $(0,1)$-matrices of order $n$ 
reduces to computing at most
$n+1$ perfect matchings in bipartite graphs 
\cite{Brualdi:Ryser:1991},
and hence the above test is 
$O(n^3\,\sqrt{n})$ time 
as the Hopcroft–Karp algorithm for computing a maximum matching
in a bipartite graph $B = (V,E)$ runs in 
$O(|E|\,\sqrt{|V|})$~\cite{HopcroftKarp:1975}.

Next, it is a simple matter to check that 
if a $(0,1)$-matrix $A$ of order $n$ is a \PET\ matrix, 
then it contains at most $\frac{1}{2}n(n+1)$ $1$'s 
(\emph{i.e.}, $\omega(A) \leq \frac{1}{2}n(n+1)$).
The following lemma gives a lower bound.

\begin{lemma}
    \label{lemma:just a few 1s}
Let $A$ be $(0,1)$-matrix of order $n$, $n \geq 2$.
If $A$ contains at most $n+1$ $1$'s, then $A$ is a \PET\ matrix.
\end{lemma}

Notice that, albeit not very impressive, 
Lemma~\ref{lemma:just a few 1s} is tight as the square matrix
$\begin{bmatrix}
I_{n-2} & \zeros_{n-2,2} \\
\zeros_{2,n-2} & J_{2}
\end{bmatrix}$
of order $n$ has $n-2+4 = n + 2$ $1$'s and is not a \PET\ matrix.

Finally, the following trivial lemma gives another condition
that helps improving the running time of the algorithm in practice.

\begin{lemma}
  \label{lemma:acyclic}
  Let $A$ be $(0,1)$-matrix of order $n$
  and $D$ the directed graph associated to $A$
  (\emph{i.e.}, the adjacency matrix of $D$ is $A$).
  % the $(0,1)$-matrix obtained from $A$
  % by replacing every non-zero diagonal matrix element by zero).
  If the digraph $D$ is acyclic (regardless self-loops),
  then $A$ is a \PET\ matrix.
\end{lemma}

We now turn to presenting the exponential-time algorithm.
The simplest exhaustive algorithm considers every possible pairs
of permutation matrices $(P, Q)$ yielding a $O((n!)^2 \;\cdot\; \POLY(n))$ time
algorithm.
% , where $f(n)$ is the time needed to compute $PAQ$ and check
% $PAQ \leq \lltm_{\,n}$.
However, according to 
Lemma~\ref{lemma:bounded stepwise property <=> PAQ triangular},
it is enough to consider every permutation matrix $P$ of order $n$
and check whether the first $i$, $1 \leq i \leq n$, rows of $PA$
have $1$'s in at most $i$ columns.
This observation yields a $O(n! \,\cdot\, \POLY(n))$ time algorithm.
% where $g(n)$ is the time needed to
% compute $PA$ and check that for the first $i$, $1 \leq i \leq n$,
% rows of $PA$ have $1$'s in at most $i$ columns.
We propose here another exhaustive algorithm that improves on the
$O(n! \,\cdot\, \POLY(n))$ time algorithm. 
The basic idea is to recursively split into smaller submatrices, 
instead of enumerating all permutations.
For a $(0,1)$-matrix $A$ of order $n$, we consider every possible set $R$
of $\left\lceil n/2 \right\rceil$ rows of $A$ and 
every possible set of $\left\lceil n/2 \right\rceil$ columns $C$ of $A$,
and check whether these lines induce a zero matrix
(or a matrix with at most one $1$ in case the matrix has odd order; 
details follow).
%Let $A' = A[R, C]$, that is $A'$ is the submatrix of $A$ with rows $R$ and columns $C$.

\begin{figure}[!t]
    \centering
    \subfigure[Even]{%
$
PAQ =
  \begin{bmatrix}
    A_1 & \zeros \\
    * & A_2
  \end{bmatrix}
$
    \label{fig:mat:even}}
    \qquad
    \subfigure[Odd and one $1$]{%
$
PAQ =
  \begin{bmatrix}
    A_1 & \zeros & \zeros \\
    *       & 1      & \zeros \\
    *       & *      & A_2
  \end{bmatrix}
$
    \label{fig:mat:odd 1}}
    \qquad
    \subfigure[Odd and zero $1$]{%
$
PAQ =
  \begin{bmatrix}
    A_1 & \zeros & \zeros \\
    *       & 0      & \zeros \\
    *       & *      & A_2
  \end{bmatrix}
$
    \label{fig:mat:odd 0}}
    \caption{\label{fig:matrices 1 and 2}}
\end{figure}
If $n$ is even, we let $P$ and $Q$ be two permutation matrices that put the rows in $R$
at the first $\left\lceil n/2 \right\rceil$ positions and the columns in $C$ 
at the last $\left\lceil n/2 \right\rceil$ positions.
The key element for the improvement
is that no specific order is required 
for the rows in $R$ nor for the columns in $C$.
The algorithm rejects the matrix $A$ for the subsets $R$ and $C$ 
if $\omega(A[R, C]) > 1$, otherwise we can write $PAQ$ as in
Figure~\ref{fig:mat:even},
where $A_1$ and $A_2$ are matrices of order $\left\lceil n/2 \right\rceil = n/2$,
and we proceed to recursively check that
both $A_1$ and $A_2$ are \PET\  matrices.
% The algorithm reject the matrix $A$ for the subsets $R$ and $C$ 
% if any of the above-mentioned checks fails, and accept otherwise.

The case $n$ is odd is a little bit more complicated.
%Recall that $A' = A[R,C]$.
First, the algorithm rejects matrix $A$ for the subsets $R$ and $C$ 
if $\omega(A[R, C]) > 1$.
Otherwise,
we need to consider two (possibly positive) cases:
(i) $\omega(A[R, C]) = 1$ 
%(\emph{i.e.}, the submatrix $A[R, C]$ contains exactly one $1$) 
or 
(ii) $\omega(A[R, C]) = 0$.
% (\emph{i.e.}, $A[R, C] = \zeros$).
If $\omega(A[R, C]) = 1$, 
we let $P$ and $Q$ be two permutation matrices that put the rows in $R$
at the first $\left\lceil n/2 \right\rceil$ positions and the columns of $C$ 
at the last $\left\lceil n/2 \right\rceil$ positions
(no specific order for the rows in $R$ nor for the columns in $C$, except that the $1$ of
$A$ is at row index $\left\lceil n/2 \right\rceil$ and at column index $\left\lceil n/2 \right\rceil$
in $PAQ$).
We can write $PAQ$ as in Figure~\ref{fig:mat:odd 1},
where $A_1$ and $A_2$ are matrices of order $\left\lceil n/2 \right\rceil - 1 = n/2$,
and we proceed to recursively check that
both $A_1$ and $A_2$ are \PET\  matrices.
Finally, if $\omega(A[R, C]) = 0$, for every row index $i \in R$ and
every column index $j \in C$,
we let $P$ and $Q$ be two permutation matrices that put the rows in $R$
at the first $\left\lceil n/2 \right\rceil$ positions and the columns of $C$ 
at the last $\left\lceil n/2 \right\rceil$ positions
(no specific order for the rows in $R$ nor for the columns in $C$ except that row $i$ in
$A$ is at row index $\left\lceil n/2 \right\rceil$ and column index $j$ is at
column index $\left\lceil n/2 \right\rceil$ in $PAQ$).
We can write $PAQ$ as in Figure~\ref{fig:mat:odd 0},
where $A_1$ and $A_2$ are matrices of order $\left\lceil n/2 \right\rceil - 1 = n/2$,
and we proceed to recursively check that
both $A_1$ and $A_2$ are \PET\ matrices.

\begin{figure}[!t]
  \begin{algorithm}[H]
    \TitleOfAlgo{\ALGO{permTriangular}}
    \DontPrintSemicolon
    \KwData{A square matrix $A =[a_{i,j}]$ of order $n$}
    \KwResult{true if $A$ is a \PET\ matrix,
    false otherwise}

    \lIf{$(\omega(A) \leq n+1)$ or $(\PER(A) = 1)$ or ($A$ is stepwise bounded) or ($D(A)$ is acyclic)}{
        \Return true
    }

    \lIf{$(\omega(A) > \frac{n(n+1)}{2})$ or $(\PER(A) > 1)$ or ($\RSV(A)$ or $\CSV(A)$ is not stepwise bounded)}{
        \Return false
    }

    \For{every subset $R \subset [n]$ of size $\left\lceil\frac{n}{2}\right\rceil$ and
    every subset $C \subset [n]$ of size $\left\lceil\frac{n}{2}\right\rceil$}{
        \eIf{$n$ is even}{
            \Return $\ALGO{permTriangularEven}(A, R, C)$
        }{
            \Return $\ALGO{permTriangularOdd}(A, R, C)$
        }
    }
    \Return false
    \caption{\label{algo:permTriangular}%
    Recognizing \PET\ matrices.}
  \end{algorithm}
\end{figure}

\begin{figure}[!t]
    \begin{algorithm}[H]
    \TitleOfAlgo{\ALGO{permTriangularEven}}
    \DontPrintSemicolon
    \KwData{A square matrix $A =[a_{i,j}]$ of even order $n$, and non-empty subsets
    $R \subset [n]$ and $C \subset [n]$, both of size $\frac{n}{2}$}
    \KwResult{true if $A$ is a \PET\ matrix
    with $A[R,C]$ as the upper right submatrix, false otherwise}
    \lIf{$\omega(A[R, C]) > 0$}{
        % \tcp{$R$ and $C$ do not induce the zero matrix}
        \Return false
    }

    Let $A_{\text{ul}} = A[R,\overline{C}]$ and $A_{\text{lr}} = A[\overline{R},C]$ \\
    \Return $\ALGO{permTriangular}(A_{\text{ul}}) \;\&\&\; \ALGO{permTriangular}(A_{\text{lr}})$
    \caption{\label{algo:permTriangularEven}%
    Subprocedure for recognizing \PET\ matrices of even order.}
    \end{algorithm}
\end{figure}

\begin{figure}[!t]
    \begin{algorithm}[H]
    \TitleOfAlgo{\ALGO{permTriangularOdd}}
    \DontPrintSemicolon
    \KwData{A square matrix $A =[a_{i,j}]$ of odd order $n$, and non-empty subsets
     $R \subset [n]$ and $C \subset [n]$, both of size $\left\lceil\frac{n}{2}\right\rceil$}
    \KwResult{true if $A$ is a \PET\ matrix
    with $A[R,C]$ as the upper right submatrix}
    \;
    \lIf{$\omega(A[R, C]) > 1$}{
        % \tcp{$R$ and $C$ do not induce a matrix with at most one $1$}
        \Return false
    }

    \eIf{$\omega(A) = 0$}{
        % \tcp{$R$ and $C$ induce a zero matrix}
        \For{every $i \in R$ and every $j \in C$}{
            Let $A_{\text{ul}} = A[R\setminus\{i\},\overline{C}]$ and
            $A_{\text{lr}} = A[\overline{R},C\setminus\{j\}]$ \;
             \lIf{$\ALGO{permTriangular}(A_{\text{ul}}) \;\&\&\;  \ALGO{permTriangular}(A_{\text{lr}})$}{
 %           \If{both $\ALGO{permTriangular}(A_{\text{ul}})$ and $\ALGO{permTriangular}(A_{\text{lr}})$ return true}{
               \Return true
            }
        }
        \Return false
    }{
        % \tcp{$R$ and $C$ induce a matrix with exactly one $1$}
        Let $i$ and $j$ be the row and column indices of the unique $1$ in $A[R, C]$\;
        Let $A_{\text{ul}} = A[R\setminus\{i\},\overline{C}]$ and
        $A_{\text{lr}} = A[\overline{R},C\setminus\{j\}]$ \;
        \Return{$\ALGO{permTriangular}(A_{\text{ul}})) \;\&\&\; \ALGO{permTriangular}(A_{\text{lr}})$}
    }
    \caption{\label{algo:permTriangularOdd}%
    Subprocedure for recognizing \PET\ matrices of odd order.}
    \end{algorithm}
\end{figure}

A detailed description is given in
Algorithms~\ref{algo:permTriangular},
\ref{algo:permTriangularEven} and
\ref{algo:permTriangularOdd}.
We now turn to evaluating the time complexity of this algorithm
and we write $T(n)$ for the time complexity of calling
$\ALGO{permTriangular}(A)$ for some $(0,1)$-matrix $A$ or order $n$.
We have
$$
T(n) \leq
\begin{cases} 
\left(\left\lceil n/2 \right\rceil\right)^2 \binom{n}{\left\lceil n/2 \right\rceil}^2
\left(2T\left(\left\lfloor n/2 \right\rfloor\right)+1\right) 
+ O(n^3\sqrt{n}) & \text{if $n$ is odd} \\
2\,\left(\left\lceil n/2 \right\rceil\right)^2 \binom{n}{\left\lceil n/2 \right\rceil}^2\;
T\left(\left\lfloor n/2 \right\rfloor\right) 
+ O(n^3\sqrt{n}) & \text{if $n$ is even}
\end{cases}
$$
with $T(1) = O(1)$.
The $O(n^3\sqrt{n})$ term
is the time complexity for lines $2$ and $3$ in 
Algorithm~\ref{algo:permTriangular}.
We also observe that the worst case occurs when $n = 2^m - 1$ as 
$\left\lceil n/2 \right\rceil,
\left\lceil n/4 \right\rceil,
\ldots$
are odd integers.
Looking for an asymptotic solution of the worst case, we thus write
the following simplified recurrence:
$
T(2^m)
=
2^{2m - 2} \,
\binom{2^m}{2^{m-1}}^2 \,
\left(2T(2^{m-1}) + 1\right) + 2^{7m/6}
$,
with $T(1) = 1$. 
Now, write $\alpha(2^m) = 2^{2m - 2} \, \binom{2^m}{2^{m-1}}^2$.
Clearly, $\alpha(2^m) \geq 2^{7m/6}$, and hence we focus for now on on
the recurrence
$T(2^m) = 2\,\alpha(2^m)\,\left(T(2^{m-1}) + 1\right)$.
A convenient non-recursive form of $T(2^m)$ is given in the following lemma.

\begin{lemma}
\label{lemma:T}
$
T(2^m)
=
\left(
2^m \prod_{i=1}^{m} \alpha(2^{i})
\right)
+
\left(
\sum_{i=1}^{m}
2^{m-i+1} \prod_{j=i}^{m} \alpha(2^{j})
\right)
$.
\end{lemma}

We now need the following lemma, in order to give an asymptotic
solution for $T(n)$ in Proposition~\ref{proposition:running time}.

% \begin{lemma}
% \label{lemma:prod alpha}
% $$
% \prod_{i=1}^{m} \alpha(2^i)
% =
% O\left(
% \frac{2^{2^{m+2}-3}}{\pi^m}
% \right)
% \text{.}
% $$
% \end{lemma}

\begin{lemma}
\label{lemma:sum prod alpha}
$
\sum_{i=1}^{m}
2^{m-i} \prod_{j=i}^{m} \alpha(2^{j})
=
O\left(m \, 2^{2^{m+2}+m+1}\right)
$.
\end{lemma}

\begin{proposition}
\label{proposition:running time}
Algorithm~\ALGO{permTriangular} runs in 
$O\left(
n\,2^{4n}\,\pi^{-\log(n)}
\right)$ time.
\end{proposition}

\begin{proof}%[of Proposition~\ref{proposition:running time}]
We have already observed that the worst case occurs for $n=2^m-1$.
According to Lemma~\ref{lemma:sum prod alpha},
we have
$
T(2^m) =
O\left(
2^{2^{m+2}+m-3}\,\pi^{-m}
\right)$
and hence
$
T(n) = 
O\left(
2^{2^{\log(n)+2}+\log(n)-3}\,\pi^{-\log(n)}
\right)
=
O\left(
n\,2^{4n}\,\pi^{-\log(n)}
\right)$.
\qed
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Conclusion ----
%%
\section{Conclusion}
\label{section:Conclusion}

We suggest for further research directions regarding 
the hardness of recognizing \PET\ $(0,1)$-matrices.
(i)
Suppose a $(0,1)$-matrix $A$ of order $n$ has $n+k$ $1$'s with
$n+2 \leq n+k \leq  \frac{1}{2}n(n+1)$. 
Can one decide in $f(k)\,n^{O(1)}$ time whether 
$A$ is a \PET\ matrix, where $f$ is an arbitrary function depending 
only on $k$?
(ii)
What is the average running time of Algorithm~\ALGO{permTriangular}
for \PET\ matrices?
(iii)
A graph labeling strongly related to symmetric \PET\ $(0,1)$-matrices
can be defined as follows: Given a graph $G=(V,E)$ or order $n$,
decide whether there exists a bijective mapping $f : V \to [n]$ such that
$f(u) + f(v) > n$ for every edge $\{u,v\} \in E$
(\emph{i.e.}, $PAP^T \leq \lrtm_{\,n}$).
Investigating the relationships between the two combinatorial problems
is expected to yield fruitful results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Bibliography
%%

\bibliographystyle{plain}
\bibliography{biblio}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{Appendix (Reviewers' version only)}

\begin{proof}[of Lemma~\ref{lemma:bounded stepwise property <=> PAQ triangular}]
The forward direction if obvious.
For the reverse direction,
let $\mathcal{S} = (S_i : 1 \leq i \leq n)$ be a configuration of subsets of
some ground set of cardinality $n$
and $\varphi : \mathcal{S} \to [n]$ be a bijective mapping
such that $\left|\bigcup_{S \in \mathcal{S} : \varphi(S) \leq i} S\right| \leq i$
for all $1\leq i \leq n$.
Let $A = [a_{i,j}]$ be the incidence matrix of $\mathcal{S}$.
Now let $P$ be the permutation matrix of order $n$ that permutes
the rows of $A$ so that row $i$ of $PA$ correspond to subset
$\varphi^{-1}(i)$ for all $1 \leq i \leq n$.
We denote this row-permuted matrix by $A' = [a'_{i,j}]$.
Define the function $\tau : [n] \to [n+1]$ that indicates for each column index
$j$ the minimum row index $i$ so that $a'_{i,j} = 1$, and we adhere to the convention
that $\tau(j) = n+1$ if row $j$ does not contain a $1$.
Define a permutation matrix $Q$ or order $n$ that permutes the
columns of $A'$ by ascending $\tau$ values,
breaking ties arbitrarily.
We designate this row-permuted matrix by $A'' = [a''_{i,j}]$.
We claim that $A'' = PAQ \leq \lltm_{\,n}$.
Indeed, suppose, aiming at a contradiction, that
$A'' \not\leq \lltm_{\,n}$.
Let $i_{\min}$ be the smallest row index such that
$a''_{i_{\min},j_{\min}} = 1$
for some $j_{\min} > i_{\min}$.
Then it follows that $\tau(j) \leq \tau(j_{\min})$ for $j < j_{\min}$
and hence
$\left|\bigcup_{S \in \mathcal{S} : \varphi(S) \leq i_{\min}} S\right| > i_{\min}$.
This is the desired contradiction, and hence
$A'' = PAQ \leq \lltm_{\,n}$.
\qed
\end{proof}

\horiz

\begin{proof}[of Lemma~\ref{lemma:normalized stepwise bounded}]
  The proof is by contradiction. Denote by $\Phi(\mathcal{S})$ the set
  of all \SBL\ s of the configuration of subsets $\mathcal{S}$. We claim that
  there exists a normalized \SBL\ $\varphi \in \Phi(\mathcal{S})$
  which maps the identical subsets of $\mathcal{S}$ to sets of consecutive
  integers. For each $\varphi \in \Phi(\mathcal{S})$, define
  $$
  \mathcal{M}(\varphi) = \{S_j \mid \exists S_i, S_k \text{ s.t. }
  \varphi(S_i) < \varphi(S_k) < \varphi(S_j) \text{ and } S_i = S_j \neq S_k\}
  $$
  Then, there exists a mapping $\varphi^* \in \Phi(\mathcal{S})$ such that
  $|\mathcal{M}(\varphi^*)| \leq |\mathcal{M}(\varphi')|$ for all $\varphi' \in
  \Phi(\mathcal{S})$. We show that $|\mathcal{M}(\varphi^*)| = 0$, and hence that
  $\varphi^*$ is our desired mapping. Suppose, for the sake of contradiction,
  that $|\mathcal{M}(\varphi^*)| > 0$. Let $S_j \in \mathcal{M}(\varphi^*)$ be
  such that $\varphi^*(S_j) \leq \varphi^*(S)$ for all
  $S \in \mathcal{M}(\varphi^*)$. Then, there exists two subsets $S_i, S_k \in \mathcal{S}$
  such that $S_i = S_j \neq S_k$ and
  $\varphi^*(S_i) + 1 = \varphi^*(S_k) < \varphi^*(S_j)$.
  Consider a new labeling $\varphi'$ defined by
  \begin{alignat*}{2}
    &\varphi'(S) = \varphi^*(S)         &\qquad& \text{for all $S \in \mathcal{S}$ such that
      $1 \leq \varphi^*(S) \leq \varphi^*(S_i)$} \\
    &\varphi'(S) = \varphi^*(S) + 1     &\qquad& \text{for all $S \in \mathcal{S}$ such that
      $\varphi^*(S_i) < \varphi^*(S) < \varphi^*(S_j)$} \\
    &\varphi'(S_j) = \varphi^*(S_k)               &\qquad& \\
    &\varphi'(S) = \varphi^*(S) + 1     &\qquad& \text{for all $S \in \mathcal{S}$ such that
      $\varphi^*(S_j) < \varphi^*(S) \leq m$}
  \end{alignat*}

%   \begin{minipage}[c]{\textwidth}
%     \bigskip
%     \centering
%     \input{fig/fig-normalized-layout.pstex_t}
%     \bigskip
%   \end{minipage}

  For simplicity of notation, we write $i$ instead of $\varphi^*(S_i)$,
  $j$ instead of $\varphi^*(S_j)$ and $k$ instead of
  $\varphi^*(S_k)$. Observe that $i+1 = k$. We claim that $\varphi'$ is
  an \SBL\ for $\mathcal{S}$. It is sufficient to show that 
  \begin{align*}
    \bigcup_{\varphi'(S) \leq \ell} S &\subseteq \bigcup_{\varphi^*(S) \leq \ell} S \\
    \intertext{for all $k \leq \ell \leq j$. We check at once that}
    \bigcup_{1 \leq \varphi'(S) \leq \ell} S &= \left( \bigcup_{1 \leq \varphi'(S) \leq i} S \right) \;\cup\; S_j \;\cup\; \left( \bigcup_{k+1 \leq \varphi'(S) \leq \ell} S \right) \\
    &= \left( \bigcup_{1 \leq \varphi^*(S) \leq i} S \right) \;\cup\; S_j \;\cup\; \left( \bigcup_{k \leq \varphi^*(S) \leq \ell-1} S \right) \\
    &= \left( \bigcup_{1 \leq \varphi^*(S) \leq \ell-1} S \right) \;\cup\; S_j \\
    &= \bigcup_{1 \leq \varphi^*(S) \leq \ell-1} S \qquad\qquad \text{(since $S_i = S_j$)} \\
    &\subseteq \bigcup_{1 \leq \varphi^*(S) \leq \ell} S
  \end{align*}
  It follows immediately that $S_j \notin \mathcal{M}(\varphi')$. Indeed, $\varphi^*(S_j) \leq \varphi^*(S)$ for all $S \in \mathcal{M}(\varphi^*)$.

  We proceed to show that $|\mathcal{M}(\varphi^*)| >
  |\mathcal{M}(\varphi')|$. If $|\mathcal{M}(\varphi')| = 0$, we are done, so
  that we may assume $|\mathcal{M}(\varphi')| > 0$. Let $S' \in
  \mathcal{M}(\varphi')$. Then there exist subsets $S_a$ and $S_b$ such that
  $\varphi'(S_a) < \varphi'(S_b) < \varphi'(S')$ and $S_a = S' \neq S_b$. Now,
  observe that we must have $\varphi^*(S_j) < \varphi^*(S')$. Again, this follows
  from the fact that $\varphi^*(S_j) \leq \varphi^*(S)$ for all $S \in
  \mathcal{M}(\varphi^*)$. Therefore, by construction of the labeling
  $\varphi'$, either $\varphi^*(S_a) < \varphi^*(S_b) < \varphi^*(S')$ or $\varphi^*(S_b) <
  \varphi^*(S_a) < \varphi^*(S')$. We claim that $S' \in \mathcal{M}(\varphi^*)$. The
  result is certainly valid in case $\varphi^*(S_a) < \varphi^*(S_b)$. Suppose now
  that $\varphi^*(S_b) < \varphi^*(S_a)$. Then it follows that we must have $S_a =
  S_j$. But $S_i = S_j$. Therefore $\varphi^*(S_i) < \varphi^*(S_b) < \varphi^*(S')$
  and $S_i = S' \neq S_b$, and hence $S' \in
  \mathcal{M}(\varphi^*)$. Combining this with the fact that $S_j \notin
  \mathcal{M}(\varphi')$ yields  $|\mathcal{M}(\varphi^*)| >
  |\mathcal{M}(\varphi')|$. This contradicts the choice of
  $\varphi^*$. Therefore, we must have $|\mathcal{M}(\varphi^*)| = 0$ and hence
  $\varphi^*$ is a normalized \SBL\ of $\mathcal{S}$.
  \qed
\end{proof}

\horiz

\begin{proof}[of Lemma~\ref{lemma:Stepwise Bounded Labeling is NPC}]
We claim that there exists a satisfying truth assignment $f$ for
$\phi$ if and only if there exists an \SBL\ of the configuration of subsets $C$ of the ground set $\mathbf{X}$.

  Suppose that there exists a satisfying truth assignment $f$ for the 
  $3$CNF formula $\phi$. 
  Due to symmetry, there is no loss of generality in assuming that each clause is 
  satisfied by its first literal. 
  Define an labeling $\tau$ of the configuration of subsets $C$ as follows. 
  First, we begin our labeling construction by placing all the empty subsets of $C$, that is
  $$
  1 \leq \tau(\mathcal{B}_i) \leq p_1 - 1
  $$
  for all $\mathcal{B}_i \in B$. 
  Next, all the subsets corresponding to the variables are ordered as follows: 
  for all $\mathcal{V}_{i, j} \in \mathcal{V}_{i}$, $1 \leq i \leq n$,
  $$
  p_i + \sum_{k = 1}^{i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{V}_{i, j}) \leq p_i + q_i + \sum_{k = 1}^{i-1} (p_k + q_k + r_k) - 1
  $$
  The satisfying truth assignment $f$ for $\phi$ is coded in our construction as follows: 
  for all $\mathcal{A}_{i, j} \in \mathcal{A}_i$, $1 \leq i \leq n$, 
  where $\mathcal{A}_i = \mathcal{T}_i$ if $f(x_i) = \mbox{TRUE}$ and 
  $\mathcal{A}_i = \mathcal{F}_i$ if $f(x_i) = \mbox{FALSE}$,
  $$
  p_i + q_i + \sum_{k = 1}^{i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{A}_{i, j}) \leq \sum_{k = 1}^{i} (p_k + q_k + r_k) - 1
  $$
  and for all $\mathcal{A}'_{i, j} \in \mathcal{A}'_i$, $1 \leq i \leq n$, 
  where $\mathcal{A}'_i = \mathcal{T}'_i$ if $f(x_i) = \mbox{TRUE}$ and $\mathcal{A}'_i = \mathcal{F}'_i$ if $f(x_i) = \mbox{FALSE}$,
  $$
  \sum_{k = 1}^{i} (p_k + q_k + r_k) \leq \tau(\mathcal{A}'_{i, j}) \leq p_{i+1} + \sum_{k = 1}^{i} (p_k + q_k + r_k) - 1
  $$
  It easily follows that
  $$
  \tau(\mathcal{V}_i) < \tau(\mathcal{A}_i) < \tau(\mathcal{A}'_i) < \tau(\mathcal{V}_j) < \tau(\mathcal{A}_j) < \tau(\mathcal{A}'_j)
  $$
  for all $1 \leq i < j \leq n$.

  Having disposed of all those subsets corresponding to the variables and to the satisfying truth assignment $f$ for $\phi$, we now turn to the clauses. This will be divided into two parts. First, the subsets of $\mathcal{C}_i$, $1 \leq i \leq m$, are ordered as follows: for all $\mathcal{C}_{i, j} \in \mathcal{C}_i$, $1 \leq i \leq m$,
  $$
  p_{n+i} + \sum_{k = 1}^{n+i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{C}_{i, j}) \leq p_{n+i} + q_{n+i} + \sum_{k = 1}^{n+i-1} (p_k + q_k + r_k) - 1
  $$
  Second, all the subsets of $\mathcal{L}_{i, 1}$ and $\mathcal{L}'_{i, 1}$, $1 \leq i \leq m$, are ordered as follows\footnote{Recall that we assume that each clause is satisfied by its first literal.}: for all $\mathcal{L}_{i, 1, j} \in \mathcal{L}_{i, 1}$, $1 \leq i \leq m$,
  $$
  p_{n+i} + q_{n+i} + \sum_{k = 1}^{n+i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{L}_{i, 1, j}) \leq \sum_{k = 1}^{n+i} (p_k + q_k + r_k) - 1
  $$
  and for all $\mathcal{L}'_{i, 1, j} \in \mathcal{L}'_{i, 1}$, $1 \leq i \leq m$,
  $$
  \sum_{k = 1}^{n+i} (p_k + q_k + r_k) \leq \tau(\mathcal{L}'_{i, 1, j}) \leq p_{n+i+1} + \sum_{k = 1}^{n+i} (p_k + q_k + r_k) - 1
  $$
  A trivial verification shows that
  $$
  \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i,1}) < \tau(\mathcal{L}'_{i, 1}) < \tau(\mathcal{C}_j) < \tau(\mathcal{L}_{j,1}) < \tau(\mathcal{L}'_{j, 1})
  $$
  for all $1 \leq i < j \leq m$.

  Here come the $K$ separator subsets of $\mathcal{S}$, that is for all $\mathcal{S}_i \in \mathcal{S}$
  $$
  p_{n+m+1} + \sum_{k = 1}^{n+m} (p_k + q_k + r_k) \leq \tau(\mathcal{S}_i) \leq p_{n+m+1} + K + \sum_{k = 1}^{n+m} (p_k + q_k + r_k) - 1
  $$
  Up to now, all the subsets of $C$ but $L$ of them have been used to define $\tau$. Our labeling construction ends with those subsets by packing them in the $L$ last places:
  $$
  p_{n+m+1} + K + \sum_{k = 1}^{n+m} (p_k + q_k + r_k) \leq \tau(\mathcal{X}_j) \leq p_{n+m+1} + K + L + \sum_{k = 1}^{n+m} (p_k + q_k + r_k) - 1
  $$
  A careful examination of $\tau$ shows that $\left| \bigcup_{\tau(X)
    \leq i} X \right| \leq i$ for $1 \leq i \leq |C|$ and hence that
  $\tau$ is an \SBL\footnote{Observe that the construction of $\tau$
    may be specialized to yield a normalized \SBL. Indeed, all the subsets of $C$ but the $L$ last ones are
    mapped by $\tau$ to a set of consecutive integers. But there is no
    less of generality in assuming that these subsets are mapped by
    $\tau$ to a set of consecutive integers as well.} of the
  configuration of subsets $C$ of the ground set $\mathbf{X}$. An
  (partial) illustration of the construction of the \SBL\ $\tau$ is shown in Figure~\ref{fig:stepwise bounded
    labeling construction}.

  % \begin{figure}[ht]
  %   \centering
  %   \bigskip
  %   \begin{minipage}[c]{\textwidth}
  %     \centering
  %     \input{r.pstex_t}
  %     \caption{\label{fig:stepwise bounded labeling
  %         construction}Illustration of the construction used in the
  %       proof of proposition~\ref{proposition:Stepwise Bounded
  %         Labeling is NPC} for the $3$CNF formula $\phi = (x_1 \vee
  %       x_2 \vee x_3) \wedge (\overline{x}_1 \vee \overline{x}_2
  %       \vee x_3) \wedge (\overline{x}_1 \vee x_2 \vee
  %       \overline{x}_3)$. Identical subsets are not distinguishable
  %       in our representation. A satisfying truth assignment is
  %       given by $f(x_1) = \mbox{TRUE}$, $f(x_2) = \mbox{FALSE}$ and
  %       $f(x_3) = \mbox{FALSE}$. For the sake of clarity, neither
  %       the ground set $\mathbf{X}$ nor the collection of subsets
  %       $C$ is fully represented. Note that we have here dropped the
  %       assumption that each clause is satisfied by its first
  %       literal.} 
  %   \end{minipage}%

  %   \bigskip
  % \end{figure}

  For the converse, suppose that there exists an \SBL\ $\tau$ for $C$, that is an labeling such that $\left|
    \bigcup_{\tau(X_j) \leq i} X_j \right| \leq i$ for $1 \leq i \leq
  |C|$. According to lemma~\ref{lemma:normalized stepwise bounded}, there is
  no loss of generality in assuming that $\tau$ is a normalized \SBL\, \emph{i.e.}, the identical subsets of the
  configuration $C$ are mapped by $\tau$ to a set of consecutive
  integers. This property is crucial in our  proof.

  Let us start by proving that $1 \leq \tau(\mathcal{B}_i) \leq p_1 - 1$
  for all $\mathcal{B}_i \in \mathcal{B}$. Indeed, as mentioned earlier
  in the proof, all the above defined subsets of the configuration of
  subsets $C$ but those of $\mathcal{B}$ contain at least $p_1$ elements
  and hence can not be mapped to an integer $i \leq p_1 - 1$ in any
  \SBL\ $\tau$ for $C$.

  Define the subcollection $C^* \subset C$ as follows:
  $$
  C^* = (X \in C \mid \max \{\tau(\mathcal{B}_i) \mid \mathcal{B}_i \in
  \mathcal{B}\} < \tau(X) < \min \{\tau(\mathcal{S}_i) \mid
  \mathcal{S}_i \in \mathcal{S}\}) 
  $$
  In other words, the subcollection $C^*$ contains all those subsets
  that go after the last bootstrap subset $\mathcal{B}_i \in
  \mathcal{B}$ and before the first separator subset $\mathcal{S}_i \in
  \mathcal{S}$. We claim that $C^*$ contains (1) the subsets of
  $\mathcal{V}_i$ for $1 \leq i \leq n$, (2) either the subsets of
  $\mathcal{T}_i$ and $\mathcal{T}'_i$ or the subsets of $\mathcal{F}_i$
  and $\mathcal{F}'_i$ for $1 \leq i \leq n$, (3) the subsets of
  $\mathcal{C}_i$ for $1 \leq i \leq m$ and (4) the subsets of
  $\mathcal{L}_{i,1}$ and $\mathcal{L}'_{i,1}$ or the subsets of
  $\mathcal{L}_{i,2}$ and $\mathcal{L}'_{i,2}$ or the subsets of
  $\mathcal{L}_{i,3}$ and $\mathcal{L}'_{i,3}$ for $1 \leq i \leq
  m$. For simplicity, this will be divided into several steps.

  Having disposed of the bootstrap subsets, we now turn to the non-empty
  subsets of the configuration of subsets $C$. Roughly speaking, we show
  that our construction implies a \emph{force placement} of the subsets
  of $C$ together with choices corresponding to a satisfying truth
  assignment for the $3$CNF formula $\phi$. First, we must have: 
  $$
  p_1 \leq \tau(\mathcal{V}_{1, j}) \leq p_1 + q_1 - 1
  $$
  for all $1 \leq j \leq q_1$. Indeed, all other subsets have more than
  $p_1$ elements and hence can not be mapped to $p_1$ by
  $\tau$. Moreover, $\tau$ is a normalized \SBL\ for
  $C$, \emph{i.e.}, the subsets $\mathcal{V}_{1, j}$, $1 \leq j \leq
  q_1$, are mapped by $\tau$ to a set of consecutive integers. Now, the
  key point is that the subset which is mapped to $p_1 + q_1$ by $\tau$
  must have cardinality at most $p_1 + q_1$. But a careful examination
  of the configuration of subsets $C$ shows that all the remaining
  subsets contain $\mathbf{V}_1$, and hence the subset which is mapped
  to $p_1 + q_1$ by $\tau$ may introduce at most $q_1$ new elements of
  the ground set. Then it follows from our construction that this subset
  is either a subset of the collection $\mathcal{T}_1$ or a subset of
  the collection $\mathcal{F}_1$. Indeed, it is sufficient to note that
  $r_1 + p_2 > q_1$, and hence that no subset $\mathcal{V}_{2, j} \in
  \mathcal{V}_2$ can satisfy $\tau(\mathcal{V}_{2, j}) = p_1 + q_1$. As
  an immediate result, exactly one of the following two statements is
  true: 
  \begin{align}
    &p_1 + q_1 \leq \tau(\mathcal{T}_{1, j}) \leq p_1 + q_1 + r_1 - 1
    \label{eq:stepwise bounded labeling 1-T} \\ 
    &p_1 + q_1 \leq \tau(\mathcal{F}_{1, j}) \leq p_1 + q_1 + r_1 - 1
    \label{eq:stepwise bounded labeling 1-F} 
  \end{align}
  for all $1 \leq j \leq r_1$. We can now proceed analogously to the
  above to obtain: 
  \begin{align*}
    \eqref{eq:stepwise bounded labeling 1-T} &\quad\Rightarrow\quad p_1 +
    q_1 + r_1 \leq \tau(\mathcal{T}'_{1, j}) \leq p_1 + q_1 + r_1 +p_2 - 1
    \\ 
    \eqref{eq:stepwise bounded labeling 1-F} &\quad\Rightarrow\quad p_1 +
    q_1 + r_1 \leq \tau(\mathcal{F}'_{1, j}) \leq p_1 + q_1 + r_1 +p_2 - 1 
  \end{align*}
  for all $1 \leq j \leq p_2$. Summarizing, exactly one of the following
  two statements is true: 
  \begin{align*}
    p_1 \leq \tau(\mathcal{V}_1) < \tau(\mathcal{T}_1) < \tau(\mathcal{T}'_1) \leq p_1 + q_1 + r_1 +p_2 - 1 \\
    p_1 \leq \tau(\mathcal{V}_1) < \tau(\mathcal{F}_1) < \tau(\mathcal{F}'_1) \leq p_1 + q_1 + r_1 +p_2 - 1
  \end{align*}
  We continue in this fashion obtaining that exactly one of the following two statements is true:
  \begin{align*}
    &p_i + \sum_{k=1}^{i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{V}_i) < \tau(\mathcal{T}_i) < \tau(\mathcal{T}'_i) \leq p_{i+1} + \sum_{k=1}^{i} (p_k + q_k + r_k) - 1 \\
    &p_i + \sum_{k=1}^{i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{V}_i) < \tau(\mathcal{F}_i) < \tau(\mathcal{F}'_i) \leq p_{i+1} + \sum_{k=1}^{i} (p_k + q_k + r_k) - 1 
  \end{align*}
  for all $1 \leq i \leq n$. This result is crucial as it allows us to construct a truth assignment $f$ for the $3$CNF formula $\phi$. Indeed, subsets of $\mathcal{T}_i$ and $\mathcal{T}'_i$ are associated in our construction to literal $x_i$ while subsets of $\mathcal{F}_i$ and $\mathcal{F}'_i$ are associated in our construction to literal $\overline{x}_i$.

  Having disposed of the variables we now turn to the clauses. For the sake of clarity, let us first introduce the temporary notations
  \begin{align*}
    l_i &= p_{n+i} + \sum_{k=1}^{n+i-1} (p_k + q_k + r_k) \\
    h_i  &= p_{n+i+1} + \sum_{k=1}^{n+i} (p_k + q_k + r_k) - 1
  \end{align*}
  for $1 \leq i \leq m$. We may now proceed analogously to the above to obtain that 
  exactly one of the following three statements is true:
  \begin{align*}
    &l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i, 1}) < \tau(\mathcal{L}'_{i, 1}) \leq h_i \\
    &l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i, 2}) < \tau(\mathcal{L}'_{i, 2}) \leq h_i \\
    &l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i, 3}) < \tau(\mathcal{L}'_{i, 3}) \leq h_i 
  \end{align*}
  for all $1 \leq i \leq m$.

  According to the above, 
  the subcollection $C^*$ contains either the subsets of $\mathcal{T}_i$ and $\mathcal{T}'_i$ 
  or the subsets of $\mathcal{F}_i$ and $\mathcal{F}'_i$ for $1 \leq i \leq n$. 
  Therefore we can define a truth assignment $f$ for the $3$CNF formula $\phi$ as follows: 
  $f(x_i) = \mbox{TRUE}$ if $\mathcal{T}_i \subset C^*$ and $f(x_i) = \mbox{FALSE}$ 
  if $\mathcal{F}_i \subset C^*$ for $1 \leq i \leq n$. 
  We claim that $f$ is a satisfying truth assignment for $\phi$. 
  Indeed, for each $i$, $1 \leq i \leq m$, consider the true statement
  $$
  l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i, k}) < \tau(\mathcal{L}'_{i, k}) \leq h_i
  $$
  where $k$ is either $1$, $2$ or $3$. By construction we have
  \begin{align*}
    \mathcal{L}_{i, k, j}  &= \mathbf{V} \cup \mathbf{V}' \cup \mathbf{A}_k \cup \mathbf{L}_{i, k} \cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}_\ell \right) \cup \left( \bigcup_{1 \leq \ell \leq i-1} \mathbf{C}'_\ell \right) \\
    \intertext{for $1 \leq j \leq r_{n+i}$ and}
    \mathcal{L}'_{i, k, j} &= \mathbf{V} \cup \mathbf{V}' \cup \mathbf{A}_k \cup \mathbf{L}_{i, k} \cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}_\ell \right) \cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}'_\ell \right) 
  \end{align*}
  for $1 \leq j \leq p_{n+i+1}$, where $\mathbf{A}_k = \mathbf{T}_\ell$
  if $\lambda_{i, k} = x_\ell$ and $\mathbf{A}_k = \mathbf{F}_\ell$ if
  $\lambda_{i, k} = \overline{x}_\ell$. But a careful examination of the
  configuration of subsets $C$ and of the \SBL\
  $\tau$ shows that none of the previous subsets contain the elements of
  $\mathbf{L}_{i, k}$. Since $|\mathcal{C}_i| = q_{n+i}$ and
  $|\mathbf{L}_{i, k}| = q_{n+i}$, then it follows that all the elements
  of $\mathbf{A}_k$ must have been introduced by previous
  subsets. Indeed, $l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i,
    k}) < \tau(\mathcal{L}'_{i, k}) \leq h_i$ and hence $\mathcal{C}_i$
  and $\mathcal{L}_{i, k}$ are mapped by $\tau$ to a set of consecutive
  integers. Therefore, if $\mathbf{A}_k = \mathbf{T}_\ell$
  (resp. $\mathbf{A}_k = \mathbf{F}_\ell$) then we must have chosen
  $\mathcal{T}_\ell$ (resp. $\mathcal{F}_\ell$) in a previous step, that
  is $\mathcal{T}_\ell, \mathcal{T}'_\ell \in C^*$
  (resp. $\mathcal{F}_\ell, \mathcal{F}'_\ell \in C^*$). Then it follows
  that clause $c_i$ is satisfied by its $k$-th literal. Hence, $f$ is a
  satisfying truth assignment for $\phi$ and the lemma is proved. 
\qed
\end{proof}

\horiz

\begin{proof}[of Lemma~\ref{lemma:permanent}]
The proof is by induction.
The result is obvious for $n=1$.
For the inductive step, assume the statement holds for some natural number $n$
and let $A$ be a $(0,1)$-matrix of order $n+1$.
Since $A$ is assumed to be a \PET\
matrix there exist permutation matrices $P$ and $Q$ such that
$PAQ \leq \lltm_{\,n}$.
Let $A' = PAQ$.
As $\PER(A)$ remains invariant under arbitrary permutation of the lines of $A$
we have $\PER(A) = \PER(A')$.
If $\PER(A') = 0$, we are done.
Otherwise $\PER(A') \geq 1$ and hence
$n$ $1$'s appear on the main diagonal of $A'$.
We designate the matrix obtained from $A'$ by deleting
the first row and the first column permuted matrix by $A''$.
As the first row of $A'$ contains exactly one $1$,
we have $\PER(A'') = \PER(A') = \PER(A) \leq 1$, thereby proving the lemma.
\qed
\end{proof}

\horiz

\begin{proof}[of Lemma \ref{lemma:just a few 1s}]
The proof is by induction.
The assertion is certainly valid for $n = 2$ as
all $(0,1)$-matrix of order $2$ are \PET\ matrices but $J_2$.
For the inductive step, assume the statement holds for some natural number $n$ and let
$A$ be $(0,1)$-matrix of order $n+1$.
If $A = \zeros$ we are done.
Otherwise, let $\alpha_i$ be a row of $A$ with minimum row sum $r_i$.
Since $A$ contains at most $n+1$ $1$'s, we certainly have $r_i \leq 1$.
We need to consider two cases:
\begin{itemize}
\item
If $r_i = 0$,
let $\beta_j$ be a column of $A$ with maximum column sum $c_j$.
Since $A \neq \zeros$, we certainly have $c_j > 0$.
We now permute the lines of $A$ so that row $\alpha_i$ and
column $\beta_j$ are the first row and column of the permuted matrix.
We now delete the first row and column of this permuted matrix and
apply the induction hypothesis to this submatrix of order $n$.
\item
If $r_i = 1$,
we may permute the rows of $A$ so that row $1$ of the permuted matrix
contains a $1$ in the $(1,1)$ position and $O$'s elsewhere.
We now delete the first row and column of this matrix and apply the induction
hypothesis to this submatrix of order $n$.
\end{itemize}
\qed
\end{proof}

\horiz

\begin{proof}[of Lemma~\ref{lemma:T}]
The proof is by induction on $m$.
First, the assertion is certainly valid for $m = 1$ as
\begin{align*}
2\,\alpha(2^1)\,(T(2^0) + 1)
&= 
2\times 4 \times (1 + 1) 
\\
&=
16
\\
\intertext{and}
\left(
2^1 \prod_{i=1}^{1} \alpha(2^{i})
\right)
+
\left(
\sum_{i=1}^{1}
2^{1-i+1} \prod_{j=i}^{1} \alpha(2^{j})
\right)
&= (2^1 \times \alpha(2^1)) + (2^1 \times \alpha(2^1))
\\
&=
(2 \times 4) + (2 \times 4) 
\\
&=
16\text{.}
\end{align*}

Suppose now that the assertion is true for $2^m$.
Then
\begin{align*}
T(2^{m+1})
&=
2\,\alpha(2^{m+1})\,T(2^{m}) + 2\,\alpha(2^{m+1})
\\
&=
2\,\alpha(2^{m+1})\,
\left(
\left(
2^m \prod_{i=1}^{m} \alpha(2^{i})
\right)
+
\left(
\sum_{i=1}^{m}
2^{m-i+1} \prod_{j=i}^{m} \alpha(2^{j})
\right)
\right)
 + 2\,\alpha(2^{m+1})
 \\
&=
\left(
2^{m+1} \prod_{i=1}^{m+1} \alpha(2^{i})
\right)
+
\left(
\sum_{i=1}^{m}
2^{m+1-i+1} \prod_{j=i}^{m+1} \alpha(2^{j})
\right)
 + 2\,\alpha(2^{m+1})
 \\ 
&=
\left(
2^{m+1} \prod_{i=1}^{m+1} \alpha(2^{i})
\right)
+
\left(
\sum_{i=1}^{m+1}
2^{m+1-i+1} \prod_{j=i}^{m+1} \alpha(2^{j})
\right)
- 2\,\alpha(2^{m+1}) 
+ 2\,\alpha(2^{m+1})
 \\
&=
\left(
2^{m+1} \prod_{i=1}^{m+1} \alpha(2^{i})
\right)
+
\left(
\sum_{i=1}^{m+1}
2^{m+1-i+1} \prod_{j=i}^{m+1} \alpha(2^{j})
\right)\text{.} 
\end{align*}
\qed
\end{proof}

\horiz

\begin{proof}[of Lemma~\ref{lemma:sum prod alpha}]
We need the following claims.

\begin{claim}
\label{claim:inner:1}
$$
\alpha(2^i)
=
\frac{2^{2^{m+1} + m - 1}}{\pi}\left(1 + O\left(2^{1-m}\right)\right)
\text{.}
$$
\end{claim}
\begin{proof}
We use the following well-known asymptotic (see e.g. \cite{ConcreteMathematics:1994}):
\begin{align*}
\binom{2n}{n} &= 
\frac{4^n}{\sqrt{\pi n}} \left(1 + O\left(n^{-1}\right)\right)\text{.}
\\
\intertext{Therefore,}
\alpha(2^m)
&=
2^{2m-2} \, \binom{2^m}{2^{m-1}}^2
\\
&=
2^{2m-2}\;
\left(\frac{4^{2^{m-1}}}{\sqrt{\pi 2^{m-1}}} \left(1 + O\left(2^{1-m}\right)\right)\right)^2
\\
&=
2^{2m-2}\;
\left(\frac{4^{2^{m-1}}}{\sqrt{\pi 2^{m-1}}}\right)^2 \left(1 + O\left(2^{1-m}\right)\right)^2
\\
&=
2^{2m-2}\;
\frac{4^{2^m}}{\pi 2^{m-1}} \left(1 + O\left(2^{1-m}\right)\right)^2
\\
&=
\frac{2^{2^{m+1} + m - 1}}{\pi}\left(1 + O\left(2^{1-m}\right)\right)^2
\text{.}
\\
\intertext{We now develop the error term to obtain}
\alpha(2^m)
&=
\frac{2^{2^{m+1} + m - 1}}{\pi}\,
\left(1 + O\left(2^{1-m}\right)\right)
\,
\left(1 + O\left(2^{1-m}\right)\right)
\\
&=
\frac{2^{2^{m+1} + m - 1}}{\pi}\,
\left(
1 + O\left(2^{1-m}\right) + O\left(2^{1-m}\right) + O\left(2^{2-2m}\right)
\right)
\\
&=
\frac{2^{2^{m+1} + m - 1}}{\pi}\left(1 + O\left(2^{1-m}\right)\right)
\text{.}
\end{align*}
\qed
\end{proof}

\begin{claim}
\label{claim:inner:2}
$$
\prod_{i=1}^{m} \alpha(2^i)
=
O\left(
\frac{2^{2^{m+2}-3}}{\pi^m}
\right)
\text{.}
$$
\end{claim}
\begin{proof}
According to Claim~\ref{claim:inner:1}, we may write
\begin{align*}
\prod_{i=1}^{m} \alpha(2^i)
&=
\prod_{i=1}^{m} \frac{2^{2^{i+1} + i - 1}}{\pi}\left(1 + O\left(2^{1-i}\right)\right)
\\
&=
\left(
\frac{1}{2\pi}
\right)^{m}
\,
\left(
\prod_{i=1}^{m} 2^{2^{i+1}}
\right)
\,
\left(
\prod_{i=1}^{m} 2^i
\right)
\,
\left(
\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right)
\right)
\\
&=
\left(
\frac{1}{2\pi}
\right)^{m}
\,
\left(2^{2^{m+2}-4}\right)
\,
\left(2^{m+1}-2\right)
\,
\left(
\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right)
\right)
\\
&=
\frac{2^{2^{m+2}-3}}{\pi^m}
\;
\left(\frac{2^m-1}{2^m}\right)
\,
\left(
\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right)
\right)
\\
&=
\frac{2^{2^{m+2}-3}}{\pi^m}
\;
\left(1 - 2^{-m}\right)
\,
\left(
\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right)
\right)\text{.}
\\
\intertext{But
$\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right) = O(1)$ and hence}
\prod_{i=1}^{m} \alpha(2^i)
&=
\frac{2^{2^{m+2}-3}}{\pi^m}
\;
O\left(1 - 2^{-m}\right)
\\
&=
O\left(
\frac{2^{2^{m+2}-3}}{\pi^m}
\right)
\text{.}
\end{align*}
\qed
\end{proof}

Having disposed of the preliminary steps,
write $\beta(i) = 2^{m-i} \prod_{j=i}^{m} \alpha(2^j)$ so that our goal
reduces to evaluating $\sum_{i=1}^{m} \beta(i)$.
For $1 \leq i < m$, we first observe that
\begin{align*}
2\;\beta(i+1)
&=
2 \; 2^{m-i-1} \prod_{j=i+1}^{m} \alpha(2^{j}) 
\\
&= 
2^{m-i} \prod_{j=i+1}^{m} \alpha(2^{j}) 
\\
&\leq 
2^{m-i} \prod_{j=i}^{m} \alpha(2^{j}) \qquad\text{(since $\alpha(i) \geq 1$)}
\\
&= 
\beta(i)\text{.}
\\
\intertext{Therefore}
\sum_{i=1}^{m} \beta(i)
&\leq 
\sum_{i=1}^{m} \frac{\beta(1)}{2^{i-1}}
\\
&=
2\,\beta(1) \; \sum_{i=1}^{m} {2^{-i}}
\\
&=
2\beta(1) \, \left(1 - 2^{-m}\right)
\\
&\leq 
2\beta(1)\text{.}
\\
\intertext{Then it follows that}
\sum_{i=1}^{m}
2^{m-i} \prod_{j=i}^{m} \alpha(2^{j})
&\leq 
2\,\beta(1)
\\
&=
2 \; 2^{m-1} \; \prod_{j=1}^{m} \alpha(2^{j})
\\
&=
2^m \; O\left(
\frac{2^{2^{m+2}-3}}{\pi^m}
\right)
\\
&=
O\left(
\frac{2^{2^{m+2}+m-3}}{\pi^m}
\right)\text{.}
\end{align*}
\qed
\end{proof}

% \newpage

% \newpage

% Let $X_{j,j}$ the event that $a_{i,j} = 1$:

% $$
% P[X_{i,j}] =
% \begin{cases}
% p & \text{if $j \leq i$} \\
% 0 & \text{if $j > i$.}
% \end{cases}
% $$

% \medskip
% Let $B=PA$.

% For $1 \leq i \leq n$ and $1 \leq j \leq n$,
% let $\mathcal{X}_{i,j}$ be event $b_{i,j} = 1$.
% We have

% $$
% \PROBABILITY[\mathcal{X}_{i,j}] = \frac{p(n-j+1)}{n}
% $$

% \medskip

% For $1 \leq i \leq n$ and $1 \leq j \leq n$,
% let $\mathcal{Y}_{i,j}$ be the event
% $\sum_{k=1}^{i} b_{k,j} > 0$,
% and let $\INDICATOR_{\mathcal{Y}_{i,j}}$ the indicator variable for event
% $\mathcal{Y}_{i,j}$.

% \begin{align*}
% \PROBABILITY\left[\mathcal{Y}_{i,j}\right]
% &=
% 1 - \PROBABILITY\left[\,\overline{\mathcal{Y}_{i,j}}\,\right] \\
% &=
% 1 - \PROBABILITY\left[ \bigcap_{1\leq k \leq i} \overline{\mathcal{X}_{k,j}}\right] \\
% &=
% 1 - \prod_{k=1}^{i} \left(1 - \PROBABILITY\left[\mathcal{X}_{k,j}\right]\right) \\
% &=
% 1 - \prod_{k=1}^{i} \left( 1 - \frac{p(n-j+1)}{n}\right) \\
% &=
% 1 - \left( 1 - \frac{p(n-j+1)}{n} \right)^i
% \end{align*}

% For $1 \leq i \leq n$, define the random variable
% $Z_i = \sum_{j=1}^{n} \INDICATOR_{\mathcal{Y}_{i,j}}$

% \begin{align*}
% P[Z_i = k]
% &\geq
% \binom{n}{k} \; \left(1-\PROBABILITY\left[\mathcal{Y}_{i,1}\right]\right)^{n-k} \;
% \left(\PROBABILITY\left[\mathcal{Y}_{i,n}\right]\right)^{k} \\
% \intertext{and hence}
% P[Z_i < i+1]
% &\geq
% \sum_{k=0}^{i}
% \binom{n}{k} \;
% \left(1-\PROBABILITY\left[\mathcal{Y}_{i,1}\right]\right)^{n-k} \;
% \left(\PROBABILITY\left[\mathcal{Y}_{i,n}\right]\right)^{k} \\
% &=
% \sum_{k=0}^{i}
% \binom{n}{k} \;
% \left(1 - \left(1 - \left(1 - p\right)^{i}\right)\right)^{n-k}
% \left(1-\left(1 - \frac{p}{n}\right)^{i}\right)^{k} \\
% &=
% \sum_{k=0}^{i}
% \binom{n}{k} \;
% \left(\left(1-p\right)^{i}\right)^{n-k}
% \left(1-\left(1 - \frac{p}{n}\right)^{i}\right)^{k} \\
% &\geq
% \sum_{k=0}^{i}
% \binom{i}{k} \;
% \left(\left(1-p\right)^{i}\right)^{i-k}
% \left(1-\left(1 - \frac{p}{n}\right)^{i}\right)^{k} \\
% &=
% \left(1 + \left(1-p\right)^{i} - \left(1 - \frac{p}{n}\right)^{i}\right)^{i}
% \text{.}
% \end{align*}

% Let $\mathcal{E}(B)$ be the event that $B$ does not have
% the stepwise bounded property.

% \begin{align*}
% \PROBABILITY\left[\mathcal{E}(B)\right]
% &=
% \PROBABILITY\left[\mathcal{E}(B)\right] \\
% &\leq
% \sum_{i=1}^{n} \PROBABILITY\left[Z_i \geq i+1\right] \\
% &=
% \sum_{i=1}^{n} \left(1 - \PROBABILITY\left[Z_i < i+1 \right]\right) \\
% \end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-06-26 article.tex
figure

@@
 \usepackage{subfigure}
-%\usepackage[mathcal]{euscript}
-\usepackage{color}
-\newcommand{\rt}{\textcolor{red}}
+
+\usepackage{tikz}
+\usetikzlibrary{matrix, calc, arrows}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\input{matrix}
-\input{math}
 \input{complexity}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\PET}{pet}
-\newcommand{\SBL}{sbl}
-\newcommand{\horiz}{{\centering\rule{350pt}{.4pt}}}
 
@@
 \title{%
-Obtaining a Triangular Matrix by Independent Row-Column Permutations
-}%
+Unshuffling Permutations}%
 
 \author{%
-  Guillaume Fertin\inst{1} \and
-  Irena Rusu\inst{1} \and
-  St\'ephane Vialette\inst{2}
+  Samuele Giraudo\inst{1} \and
+  St\'ephane Vialette\inst{1}
 }% end author
 \institute{%
-  LINA UMR CNRS 6241, Universit\'e de Nantes, France \\
-  \email{\{guillaume.fertin,irena.rusu\}@univ-nantes.fr}
-  \and
   Universit\'e Paris-Est, LIGM (UMR 8049), CNRS, UPEM, ESIEE Paris, ENPC, 
@@
 \begin{abstract}
-Given a square $(0,1)$-matrix $A$,
-we consider the problem of deciding whether there exists a permutation
-of the rows and a permutation of the columns of 
-$A$ such that after carrying out these permutations,
-the resulting matrix is triangular. The complexity of the problem was
-posed as an open question by Wilf~\cite{Wilf:1997} in 1997.
-%(\emph{i.e.}, do there exist permutation matrices $P$ and $Q$ such that
-%$PAQ$ is triangular)?
-In 1998, DasGupta et al.~\cite{DasGuptaJiangKannanLiSweedyk:1998} seemingly answered the
-question, proving it is \NPC. However, we show here that their result is flawed, which 
-leaves the question still open. Therefore, we give a definite answer
-to this question by proving that the problem is \NPC. We finally present an
-exponential-time algorithm for solving the problem.
+bla bla bla \ldots
 \end{abstract}
@@
 
-In his contribution in the tribute to the late Professor
-Erd\"os~\cite{Wilf:1997}, Wilf posed the following question:
-\emph{
-  ``Let $A$ be a $m \times n$ matrix of $0$'s and $1$'s. Consider the
-  computational problem: do there exist permutations $P$ of the rows
-  of $A$, and $Q$, of the columns of $A$ such that after carrying out
-  these permutations, $A$ is triangular? The question we ask concerns
-  the complexity of the problem.
-  Is this problem \NPC?
-  Or, does there exist a polynomial-time algorithm for doing it?''}
-As noted by Wilf,
-this problem is strongly related to job scheduling with precedence
-constraints, a well-known problem in theoretical computer science.
-The present paper is devoted to giving an answer to this question.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-A square matrix is called {\em lower triangular} if all the entries above the main diagonal are zero.
-Similarly, a square matrix is called {\em upper triangular} if all the entries below the main diagonal
-are zero.
-A {\em triangular matrix} is one that is either lower triangular or upper triangular.
-Because matrix equations with triangular matrices are easier to solve,
-they are very important in  linear algebra and numerical analysis.
-We refer the reader to \cite{MatrixComputations:1996} for a further discussion.
-
-For an arbitrary square matrix $A$, it is well-known in linear algebra
-that there exists an invertible matrix $S$ such that $S^{-1}AS$ is
-upper triangular.
-We focus here, however, on permutation matrices.
-Recall that a permutation matrix is a square matrix obtained from the same
-size identity matrix by a permutation of rows. 
-A product of permutation matrices is again a permutation matrix and
-the inverse of a permutation matrix is again a permutation matrix.
-In fact, for any permutation matrix $P$, $P^{-1} = P^T$.
-
-% Nicolson [5] shows that for a nonnegative matrix $A$ the following four conditions
-% are equivalent:
-% (i) there exists a permutation matrix $P$ such that
-% $P A P^T$ is strictly upper triangular;
-% (ii) there is no positive cycle in $M$
-% (i.e. in the weighted digraph there is no cycle whose edges all have positive weight);
-% (iii) the permanent of $M + I$ is $1$; and
-% (iv) $A$ is nilpotent.
-
-This paper is organized as follows.
-In Section~\ref{section:Notations}, we provide the basic material
-needed for this paper. %with a
-%strong emphasis on the associated combinatorial structures: configurations of subsets,
-%bipartite graphs and directed graphs.
-Section~\ref{section:Hardness result} is devoted to proving hardness of
-determining whether a square $(0,1)$-matrix is permutation equivalent
-triangular, i.e. whether it can be transformed into a triangular matrix by independent
-row and column permutations.
-In Section~\ref{section:Exponential-time algorithm}, we give some properties of
-permutation equivalent triangular matrices (or \PET\ matrices, for short) and present 
-an exponential-time algorithm to determine whether a matrix is a \PET\ matrix.
-The paper concludes with suggestions for further research directions.
+%%
+%% ---- Notations ----
+%%
+\section{Notations}
+\label{section:Notations}
 
@@
 %%
-%% ---- Preliminaries ----
+%% ---- Algorithmic issues ----
 %%
-\section{Notations}
-\label{section:Notations}
+\section{Algorithmic issues}
+\label{section:Algorithmic issues}
 
-%\subsection{$(0,1)$-matrices}
-%\subsubsection{$(0,1)$-matrices} 
-For any positive integer $n$, denote $[n]=\{1, 2, \ldots, n\}$. 
-Let $A = [a_{i,j}]$, $1 \leq i \leq m$ and $1 \leq j \leq n$,
-be a matrix of $m$ rows and $n$ columns.
-% We say that $A$ is of size $m$ by $n$ and we also refer to $A$
-% as an $m$ by $n$ matrix.
-In the case that $m = n$ then the matrix is \emph{square} of \emph{order} $n$.
-It is always assumed that the entries of the matrix are elements
-of some underlying field $F$.
-It is convenient to refer to either a row or a column of the matrix as a
-\emph{line} of the matrix.
-We use the notation $A^T$ for the \emph{transpose} of matrix $A$.
-We always designate a zero matrix by $\zeros$, a matrix with every entry equal to
-$1$ by $J$, and the identity matrix of order $n$ by $I$.
-In order to emphasize the size of these matrices we sometimes include subscripts.
-Thus $J_{m,n}$ denotes the all $1$'s matrix of size $m$ by $n$,
-and this is abbreviated to $J_n$ if $m = n$. 
-Notations $\zeros_{m,n}$, $\zeros_n$ and $I_n$ are similarly defined.
-In displaying a matrix we often use $*$ to designate a submatrix of no particular structure.
-Two matrices $A$ and $B$ are said to be \emph{permutation equivalent},
-denoted by $A \sim B$, if there exist permutation matrices $P$ and $Q$ of
-suitable sizes such that $B = PAQ$.
-
-We will be greatly concerned with matrices whose entries consist exclusively of
-the integers $0$ and $1$.
-Such matrices are referred to as \emph{$(0,1)$-matrices}.
-For a $(0,1)$-matrix $A$, we let $\omega(A)$ stand for the number of $1$'s in $A$.
-A square matrix $A = [a_{i,j}]$ of order $n$ is said to be
-\emph{lower left triangular} if it has only $0$'s above the main diagonal
-(\emph{i.e.} $a_{i,j} = 0$ for $1 \leq i < j \leq n$.
-We write $\lltm_{\,n}$ for the lower left triangular $(0,1)$-matrix 
-whose $0$'s are exclusively above the main diagonal.
-For two matrices $A = [a_{i,j}]$ and $B = [b_{i,j}]$ of size $m$ by $n$,
-we write $A \leq B$ if $a_{i,j} \leq b_{i,j}$ for $1 \leq i \leq m$
-and $1 \leq j \leq n$, so that a square matrix $A$ of order $n$ is
-lower left triangular if $A \leq \lltm_{\,n}$.
-In the context of permutation equivalent matrices,
-we will sometimes not be interested in any particular orientation of a
-triangular matrix and forget about any
-specific orientation such as ``\emph{lower left}''.
-Furthermore, for readability, a matrix which is permutation equivalent
-to a triangular matrix is said to be a \emph{\PET\ matrix}.
-The \emph{row sum vector}
-$\RSV(A) = \begin{bmatrix} r_1 & r_2 & \ldots & r_m \end{bmatrix}$
-and the  \emph{column sum vector}
-$\CSV(A) = \begin{bmatrix} c_1 & c_2 & \ldots & c_n \end{bmatrix}$
-of $A$ are defined by
-$r_i = \sum_{1 \leq j \leq n} a_{i,j}$ for $1 \leq i \leq m$
-and
-$c_j = \sum_{1 \leq i \leq m} a_{i,j}$ for $1 \leq j \leq n$.
-The row sum vector $\RSV(A)$
-(resp. column sum vector $\CSV(A)$)
-is \emph{stepwise bounded}
-if  $|\{i : r_i \leq k \}| \geq k$
-(resp. $|\{j : c_j \leq k \}| \geq k$)
-for $1 \leq k \leq n$.
-It is clear that if a $(0,1)$-matrix $A$ is a \PET\ matrix then
-both $\RSV(A)$ and $\CSV(A)$ are stepwise bounded.
-
-% The definition of the \emph{permanent} of $A$ follows the classical definition of the
-% determinant but with a simplification. Ironically, this simplification in the formula makes
-% it more difficult to compute the permanent.
-The permanent of $A = [a_{i,j}]$ is defined as the number given by the formula
-$\PER(A) = \sum_{(j_1, j_2, \ldots, j_n) \in S_n}
-a_{1,j_1}\,a_{2,j_2}\, \ldots a_{n,j_n}$,
-where the summation is over all permutations
-$(j_1,j_2, \ldots, j_n)$ of $[n]$.
-Observe that, unlike the determinant, we do not put a minus sign in front of some of the terms in the
-summation.
-Of particular importance,
-the permanent does not change if we permute the rows of $A$ and
-permute the columns of $A$.
-
-Let $A = [a_{i,j}]$ be an $m$ by $n$ matrix.
-For convenience, for a set $K \subseteq [m]$ we will
-write $\overline{K}$ for the set
-$[m] \setminus K$.
-Let $K = \{i_1, i_2, \ldots, i_k\}$ be a set of $k$ elements with
-$K \subseteq [m]$,
-and let $L = \{j_1, j_2, \ldots, j_l\}$ be a set of
-$l$ elements with $L \subseteq [n]$.
-The sets $K$ and $L$ designate a collection of row indices and column indices,
-respectively, of the matrix $A$, and the $k$ by $l$ submatrix determined by them is
-denoted $A[K,L]$.
-% $$
-% A[K,L]=
-% \begin{bmatrix}
-% a_{i_1,j_1} & a_{i_1,j_2} & \dots  & a_{i_1,j_l} \\
-% a_{i_2,j_1} & a_{i_2,j_2} & \dots  & a_{i_2,j_l} \\
-% \vdots      & \vdots      & \ddots & \vdots      \\
-% a_{i_k,j_1} & a_{i_k,j_2} & \dots  & a_{i_k,j_l}
-% \end{bmatrix}\text{.}
-% $$
-% If $L = K$, then $A[K, K]$ is a \emph{principal} submatrix of $A$,
-% sometimes denoted more simply as $A[K]$.
-
-%\subsection{Related combinatorial objects}
-%\subsubsection{Configuration of subsets}
-
-% Combinatorial matrix theory \cite{Brualdi:Ryser:1991} includes a wide variety of 
-% combinatorial areas, e.g. graph theory \cite{Diestel:2000}, transversal theory
-% \cite{Mirsky:1971}, combinatorial configurations \cite{Ryser:1969}.
-% We shall make intensive use of configuration of subsets.
-
-Let $X = \{x_i : 1 \leq i \leq n\}$ be a non-empty set of $n$
-elements, that we call an \emph{$n$-set}.
-Let $\mathcal{S} = (S_i : 1 \leq i \leq m)$ be $m$ not necessarily
-distinct
-subsets of the $n$-set $X$. We refer to this collection of subsets of an
-$n$-set as a \emph{configuration of subsets}.
-We set $a_{i,j} = 1$ if $x_j \in S_i$, and $a_{i,j} = 0$
-if $x_i \notin S_i$.
-The resulting $(0,1)$-matrix $A = [a_{i,j}]$, $1 \leq i \leq m$ and
-$1 \leq j \leq n$ of size $m$ by $n$ is the \emph{incidence matrix}
-for the configuration of subsets $\mathcal{S}$ of the $n$-set $X$.
-The $1$s in row $\alpha_i$ of $A$ display the elements in the subset
-$S_i$, and the $1$'s in column $\beta_j$ display the occurrences of $x_j$
-among the subsets. 
-Let $\mathcal{S} = (S_i : 1 \leq i \leq n)$ be a configuration of subsets of some 
-ground $n$-set $X$. A bijective mapping $\varphi : \mathcal{S} \to [n]$ is said to be a
-\emph{stepwise bounded labeling} (or \SBL\ for short) of $\mathcal{S}$ if
-$\left|\bigcup_{\varphi(S_j) \leq i}S_j\right| \leq i$
-for $1 \leq i \leq n$.
-
-%  A \emph{system of distinct representatives} (SDR) for the family $\mathcal{S}$
-% is a family $(x_1, x_2, \ldots, x_n)$ of elements of $X$ satisfying
-% the conditions
-% (i) $x_i \in S_i$ for $1 \leq i \leq n$, and
-% (ii) $x_i \neq xj$ for $i \neq j$.
-% (The first condition asserts that the elements are representatives of the sets,
-% and the second that they are distinct.)
-% The collection $\mathcal{S}$ satisfies the \emph{marriage} condition
-% if and only if for each subcollection $\mathcal{S}' \subseteq \mathcal{S}$, we have
-% $|\mathcal{S}'| \leq \left|\bigcup_{S \in \mathcal{S}'} S\right|$.
-% In other words, the number of sets in each subcollection $\mathcal{S}'$ is less
-% than or equal to the number of distinct elements in the union over the
-% subcollection $\mathcal{S}'$.
-% Hall's theorem states that $\mathcal{S}$ has a SDR if and only if
-% $\mathcal{S}$ satisfies the marriage condition.
-
-% \paragraph{Bipartite graphs.}
-
-% A graph $G=(V,E)$ is \emph{bipartite} if 
-% $V$ can be partitioned into two sets $U$ and $W$ such that each edge
-% joins a vertex in $U$ to a vertex in $W$, and the pair $(U, W)$ is called a
-% \emph{bipartition} of $V$ (or of $G$).
-% The \emph{reduced adjacency matrix} $B = [b_{i,j}]$ is an $m \times n$
-% $(0,1)$-matrix defined by $b_{i, j} = 1$ if and only if
-% there is an edge beteween vertex $i$ of $U$ and vertex $j$ of $W$.
-% A matching $M$
-% in $G$ is a collection of edges no two of which have a vertex in
-% common. If $v$ is a vertex and there is an edge of $M$ containing $v$,
-% then $v$ \emph{meets} the matching $M$ and the matching $M$ meets the vertex
-% $v$. A \emph{perfect matching} of $G$ is a matching
-% that meets all vertices of $G$.
-
-% \paragraph{Directed graphs.}
-
-% A \emph{directed graph} (for short, a \emph{digraph}) $G$
-% consists of a finite set $V$ of elements called vertices and a set $E$ of
-% ordered pairs of vertices called arcs.
-% The \emph{adjacency matrix} associated to $G$ is the $|V| \times |V|$
-% $(0,1)$-matrix $A = [a_{i,j}]$ defined by $a_{i,j} = 1$ if and only if
-% there is an arc from vertex $i$ to vertex $j$ in $G$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Properties ----
-%%
-\section{Answering Wilf's question}%Hardness result}
-\label{section:Hardness result}
-
-We prove in this section that, given a square $(0,1)$-matrix $A$, deciding whether there 
-exists a permutation matrix $P$ and a permutation matrix $Q$ of suitable size
-such that $PAQ$ is triangular is \NPC.
-
-\subsection{Disproving a previous related result}
-
-Before giving our proof, it is worth mentioning that the following problem
-(called LBQIS$(n,k)$ and rephrased to fit the context of this paper) is
-claimed to be \NPC in~\cite{DasGuptaJiangKannanLiSweedyk:1998}:
-Given a $(0,1)$-matrix of order $n$ and positive integer $k \leq n$,
-do there exist permutation matrices $P$ and $Q$ such that
-$
-PAQ =
-\begin{bmatrix}
-A_{1,1} & A_{1,2} \\
-A_{2,1} & A_{2,2}
-\end{bmatrix}
-$
-with $A_{1,2}$ a square lower triangular matrix of size $k$ by $k$?
-It is not very difficult to find a polynomial transformation from
-LBQIS to Wilf's question, which would prove the \NPcompleteness of the
-latter. Just add $n-k$ empty rows and $n-k$ empty columns 
-to matrix $A$ to obtain a new matrix $A'$. Now, notice that each 
-submatrix $A_{1,2}$ in a solution for LBQIS may be completed with the 
-$n-k$ empty rows put before row 1 of $A_{1,2}$ and with the $n-k$ empty 
-columns put after column $k$ of $A_{1,2}$ to yield a solution for the 
-instance $A'$ in Wilf's question, and viceversa.
-
-Unfortunately, paper~\cite{DasGuptaJiangKannanLiSweedyk:1998} 
-contains a serious flaw in the proof. To fix things, note that
-in~\cite{DasGuptaJiangKannanLiSweedyk:1998} LBQIS is stated in
-terms of bipartite graphs, for which matrix $A$ is the reduced adjacency
-matrix. Then, LBQIS$(n,k)$ is proved \NPC by reduction 
-from another problem on bipartite graphs called LBIS$(n,k)$, using the so-called 
-Rearrangement Lemma (Lemma 3.5 in ~\cite{DasGuptaJiangKannanLiSweedyk:1998}).
-Two affirmations in the proof of this lemma are contradicted by the following
-example. Let $G$ be the graph (input for LBIS) with vertices $U=\{i\, |\, 1\leq i\leq 4\}$
-and $V=\{i\, |\, 1\leq i\leq 4\}$, whose edges are $(1,1), (2,1), (2,2),
-(3,2), (3,4), (4,3)$ and $(4,4)$. Thus, $n=4$. Define $k=1$.  
-Let $G'$ be the input graph for LBQIS built as in~\cite{DasGuptaJiangKannanLiSweedyk:1998},
-and $k'=k^2+k=2$. Then the vertex subset $U'\cup V'$ of $G'$, with $U'=\{[2,4], [1,2]\}$
-and $V'=\{[1,3], [2,2]\}$ is a solution of LBQIS of size $k'$ for which the
-assumption on the first line of the Rearrangement Lemma's proof is false.
-Also, the vertex subset $U'\cup V'$ of $G'$, with $U'=\{[1,1], [2,1]\}$
-and $V'=\{[1,2], [1,3]\}$ is a solution of LBQIS of size $k'$ for which
-the second affirmation in the same lemma (``clearly $q_1\leq p_1$'') is also false.
-
-\subsection{Our \NPcompleteness proof for Wilf's question}
-
-We present our results in terms of \SBL\ for configurations
-of subsets. The rationale for considering \SBL\ for configurations
-of subsets stems from the following lemma.
-
-\begin{lemma}
-    \label{lemma:bounded stepwise property <=> PAQ triangular}
-    Let $\mathcal{S} = (S_i : 1 \leq i \leq n)$ be a configuration of subsets of
-    some ground $n$-set, and let $A$ be the corresponding incidence matrix.
-    There exist permutation matrices $P$ and $Q$ of order $n$
-    such that $PAQ \leq \lltm_{\,n}$ if and only if
-    there exists an \SBL\ of $\mathcal{S}$.
-\end{lemma}
-
-We need to focus our attention on a special type of \SBL. Call a bijective mapping
-$\varphi : \mathcal{S} \to [n]$ \emph{normalized} if $\varphi$
-maps the identical subsets of elements of $\mathcal{S}$
-to a set of consecutive integers.
-Most of the interest in normalized bijective labelings
-stems from the following intuitive lemma.
-
-\begin{lemma}
-  \label{lemma:normalized stepwise bounded}
-  Let $\mathcal{S} = (S_i : 1 \leq i \leq n)$ be a configuration of subsets of
-    some ground $n$-set.
-    If there exists an \SBL\ of $\mathcal{S}$
-    then there exists a normalized \SBL\ of $\mathcal{S}$.
-\end{lemma}
-
-We are now ready to prove that
-deciding whether there exists an \SBL\ of some
-configuration of subsets is \NPC thereby proving that 
-deciding whether a square $(0,1)$-matrix is a \PET\ matrix is \NPcomplete as well.
-The proof proceeds by a reduction from the \PB{3Sat} problem - which is a known
-\NPcomplete problem \cite{Cook:1971}.
-% (see also
-% \cite{Garey:Johnson:1979} and \cite{Papadimitriou:1994} for variants).
-Let an arbitrary
-instance of the \PB{3Sat} problem be given by a $3$CNF formula 
-$\phi = c_1 \vee c_2 \vee \ldots \vee c_m$ over variables 
-$x_1, x_2, \ldots, x_n$. 
-Our construction is divided into two steps: 
-(1) construction of a (polynomial size) ground set $\mathbf{X}$ and 
-(2) construction of a configuration of subsets $C$ of the ground set
-$\mathbf{X}$. 
-Throughout the proof, parts of the ground set
-$\mathbf{X}$ are written as capital bold letters 
-($\mathbf{V}, \mathbf{T}, \mathbf{F}, \ldots$)
-and subsets of the configuration are written with capital calligraphic
-letters ($\mathcal{V}_i, \mathcal{T}_i, \mathcal{F}_i, \ldots$).
-
-  \begin{figure}[H]
-    \centering
-    \begin{minipage}[c]{\textwidth}
-      \centering
-      \input{r.pstex_t}
-      \caption{\label{fig:stepwise bounded labeling construction}
-          Illustration of the construction 
-          % used in the proof of 
-          % Proposition~\ref{proposition:Stepwise Bounded Labeling is NPC} 
-          for the $3$CNF formula 
-          $\phi = (x_1 \vee x_2 \vee x_3) \wedge 
-          (\overline{x}_1 \vee \overline{x}_2 \vee x_3) \wedge 
-          (\overline{x}_1 \vee x_2 \vee \overline{x}_3)$. 
-          Identical subsets are not distinguishable
-          in our representation. A satisfying truth assignment is
-          given by $f(x_1) = \mbox{TRUE}$, $f(x_2) = \mbox{FALSE}$ and
-          $f(x_3) = \mbox{FALSE}$. For sake of clarity, neither
-          the ground set $\mathbf{X}$ nor the collection of subsets
-          $C$ is fully represented. 
-          % Note that we have here dropped the
-          % assumption that each clause is satisfied by its first
-          % literal
-          } 
-    \end{minipage}%
-  \end{figure}
-
-To begin with, define 
-$p_i = 3(n + m + 1 - i) + 2$,
-$q_i = 3(n + m + 1 - i) + 1$ and
-$r_i = 3(n + m + 1 - i)$ for $1 \leq i \leq n + m$.
-Furthermore, define $p_{n+m+1}=1$,
-$K = \sum_{i=1}^{n} q_i + 2 \sum_{i=n+1}^{n+m} q_i$ and
-$L = \sum_{i=1}^{n+m} (p_{i+1} + r_i)$.
-Let us now define the ground set $\mathbf{X}$. 
-Consider the pairwise disjoint sets defined as follows:
-$\mathbf{V}_i = \{v_{i, j} \mid 1 \leq j \leq p_i\}$,
-$\mathbf{V}'_i  = \{v'_{i, j} \mid 1 \leq j \leq r_i\}$,
-$\mathbf{T}_i = \{t_{i, j} \mid 1 \leq j \leq q_i\}$,
-$\mathbf{F}_i = \{f_{i, j} \mid 1 \leq j \leq q_i\}$
-for $1 \leq i \leq n$. Furthermore, define
-$\mathbf{C}_i = \{c_{i, j} \mid 1 \leq j \leq p_{n + i}\}$
-$\mathbf{C}'_i = \{c'_{i, j} \mid 1 \leq j \leq r_{n + i}\}$
-for $1 \leq i \leq m$, and
-$\mathbf{L}_{i, k} = \{\ell_{i, k, j} \mid 1 \leq j \leq q_{n + i}\}$
-for $1 \leq i \leq m$ and $1 \leq k \leq 3$.
-Finally, define $\mathbf{S} = \{s\}$.
-For simplicity of notation, write
-$\mathbf{V} = \bigcup_{1 \leq i \leq n} \mathbf{V}_i$,
-$\mathbf{V}'= \bigcup_{1 \leq i \leq n} \mathbf{V}'_i$,
-$\mathbf{T} = \bigcup_{1 \leq i \leq n} \mathbf{T}_i$,
-$\mathbf{F} = \bigcup_{1 \leq i \leq n} \mathbf{F}_i$,
-$\mathbf{C} = \bigcup_{1 \leq i \leq m} \mathbf{C}_i$,
-$\mathbf{C}' = \bigcup_{1 \leq i \leq m} \mathbf{C}'_i$,
-and
-$\mathbf{L}_i = \bigcup_{1 \leq k \leq 3} \mathbf{L}_{i, k}$
-for $1 \leq i \leq m$
-and
-$\mathbf{L} = \bigcup_{1 \leq i \leq m} \mathbf{L}_i$.
-Informally, elements of $\mathbf{V} \cup \mathbf{V}'$ are associated to variables, 
-elements of $\mathbf{T} \cup \mathbf{F}$ are associated to literals, elements of 
-$\mathbf{C} \cup \mathbf{C}'$ are associated to clauses, elements of $\mathbf{L}$ 
-are associated to literals in clauses and $\mathbf{S}$ is a  
-\emph{separator set}. The ground set $\mathbf{X}$ of our construction is defined to be
-$\mathbf{X} = 
-\mathbf{V} \cup \mathbf{V}' \cup \mathbf{T} \cup 
-\mathbf{F} \cup \mathbf{C} \cup \mathbf{C}' \cup 
-\mathbf{L} \cup \mathbf{S}$.
-
-Having defined the ground set $\mathbf{X}$, 
-we now turn to the detailed construction of a configuration of subsets $C$ of $\mathbf{X}$. 
-For sake of clarity, this will be divided into several steps. 
-First, each variable $x_i$, $1 \leq i \leq n$, is associated to identical subsets 
-$\mathcal{V}_{i, j}$, $1 \leq j \leq q_i$, in $C$. These subsets are defined as follows:
-$\mathcal{V}_{i, j} = 
-\left( \bigcup_{1 \leq k \leq i} \mathbf{V}_k \right) \cup 
-\left( \bigcup_{1 \leq k \leq i-1} \mathbf{V}'_k \right)$
-for $1 \leq i \leq n$ and $1 \leq j \leq q_i$.
-Let us denote by $\mathcal{V}_i$, $1 \leq i \leq n$, 
-the collection $(\mathcal{V}_{i, j} \mid 1 \leq j \leq q_i)$.
-Next, each (positive) literal $x_i$, $1 \leq i \leq n$, 
-is associated to identical subsets $\mathcal{T}_{i, j}$, $1 \leq j \leq r_i$, 
-and to identical subset $\mathcal{T}'_{i, j}$, $1 \leq j \leq p_{i + 1}$. 
-These subsets are defined as follows:
-$\mathcal{T}_{i, j} = \mathbf{T}_i \cup 
-\left( \bigcup_{1 \leq k \leq i} \mathbf{V}_k \right) \cup 
-\left( \bigcup_{1 \leq k \leq i-1} \mathbf{V}'_k \right)$
-for $1 \leq i \leq n$ and $1 \leq j \leq r_i$, and
-$\mathcal{T}'_{i, j} = \mathbf{T}_i \cup \left( 
-\bigcup_{1 \leq k \leq i} \mathbf{V}_k \right) \cup 
-\left( \bigcup_{1 \leq k \leq i} \mathbf{V}'_k \right)$
-for $1 \leq i \leq n$ and $1 \leq j \leq p_{i+1}$.
-Of course, a similar construction of subsets applies for the negation 
-$\overline{x}_i$ of each variable $x_i$, \emph{i.e.},
-$\mathcal{F}_{i, j} = 
-\mathbf{F}_i \cup \left( \bigcup_{1 \leq k \leq i} 
-\mathbf{V}_k \right) \cup \left( \bigcup_{1 \leq k \leq i-1} 
-\mathbf{V}'_k \right)$ for $1 \leq i \leq n$ and $1 \leq j \leq r_i$, and
-$\mathcal{F}'_{i, j} = \mathbf{F}_i \cup \left( 
-\bigcup_{1 \leq k \leq i} \mathbf{V}_k \right) \cup 
-\left( \bigcup_{1 \leq k \leq i} \mathbf{V}'_k \right)$ 
-for $1 \leq i \leq n$ and $1 \leq j \leq p_{i+1}$.
-For readability, write 
-$\mathcal{T}_i = (\mathcal{T}_{i, j} \mid 1 \leq j \leq r_i)$, 
-$\mathcal{T}'_i = (\mathcal{T}'_{i, j} \mid 1 \leq j \leq p_{i+1})$, 
-$\mathcal{F}_i = (\mathcal{F}_{i, j} \mid 1 \leq j \leq r_i)$ and 
-$\mathcal{F}'_i = (\mathcal{F}'_{i, j} \mid 1 \leq j \leq p_{i+1})$ for $1 \leq i \leq n$. 
-Note that the following (strict) inclusions hold for all $1 \leq i \leq n$, 
-$1 \leq j_1 \leq q_i$, $1 \leq j_2 \leq r_i$ and $1 \leq j_3 \leq p_{i+1}$: 
-(i) $\mathcal{V}_{i, j_1} \subset \mathcal{T}_{i, j_2} \subset \mathcal{T}'_{i, j_3}$ and 
-(ii) $\mathcal{V}_{i, j_1} \subset \mathcal{F}_{i, j_2} \subset \mathcal{F}'_{i, j_3}$.
-We now turn to the $m$ clauses of the $3$CNF formula. 
-Each clause $c_i$, $1 \leq i \leq m$, is associated to identical subsets 
-$\mathcal{C}_{i, j}$, $1 \leq j \leq q_{n+i}$. These subsets are defined as follows:
-$\mathcal{C}_{i, j} = \mathbf{V} \cup \mathbf{V}' \cup 
-\left( \bigcup_{1 \leq k \leq i} \mathbf{C}_k \right) 
-\cup \left( \bigcup_{1 \leq k \leq i-1} \mathbf{C}'_k \right)$
-for $1 \leq i \leq m$ and $1 \leq j \leq q_{n+i}$.
-Let us denote by $\mathcal{C}_i$, $1 \leq i \leq m$, 
-the collection $(\mathcal{C}_{i, j} \mid 1 \leq j \leq q_{n+i})$. 
-It is easily seen that $\mathcal{V}_{i, j_1} \subset \mathcal{C}_{k, j_2}$ 
-for all $1 \leq i \leq n$, $1 \leq j_1 \leq q_i$, $1 \leq k \leq m$ and 
-$1 \leq j_2 \leq q_{n+k}$.
-
-Now, we consider the only part of the construction that depends on which 
-literal occurs in which clauses. Denote by $\lambda_{i, k}$ the $k$-th literal 
-of clause $c_i$, 
-that is write $c_i = \lambda_{i, 1} \vee \lambda_{i, 2} \vee \lambda_{i, 3}$ 
-for $1 \leq i \leq m$,
-where each $\lambda_{i, k}$ is a variable or its negation. 
-The $k$-th literal, $1 \leq k \leq 3$, of each clause $c_i$, $1 \leq i \leq m$, 
-is associated to identical subsets $\mathcal{L}_{i, k, j}$, $1 \leq j \leq r_{n+i}$, 
-and to identical subsets $\mathcal{L}'_{i, k, j}$, $1 \leq j \leq p_{n+i+1}$. 
-These subsets are defined as follows:
-$\mathcal{L}_{i, k, j} = \mathbf{V} \cup \mathbf{V}' \cup 
-\mathbf{A}_k \cup \mathbf{L}_{i, k} \cup 
-\left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}_\ell \right) \cup 
-\left( \bigcup_{1 \leq \ell \leq i-1} \mathbf{C}'_\ell \right)$
-for $1 \leq i \leq m$, $1 \leq j \leq r_{n+i}$ and $1 \leq k \leq 3$ and
-$\mathcal{L}'_{i, k, j} = \mathbf{V} \cup \mathbf{V}' \cup \mathbf{A}_k 
-\cup \mathbf{L}_{i, k} \cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}_\ell \right) 
-\cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}'_\ell \right)$ 
-for $1 \leq i \leq m$, $1 \leq j \leq p_{n+i+1}$ and $1 \leq k \leq 3$, 
-where $\mathbf{A}_k = \mathbf{T}_\ell$ if $\lambda_{i, k} = x_\ell$ 
-and $\mathbf{A}_k = \mathbf{F}_\ell$ if $\lambda_{i, k} = \overline{x}_\ell$. 
-For the sake of clarity, write 
-$\mathcal{L}_{i, k} = (\mathcal{L}_{i, k, j} \mid 1 \leq j \leq r_{n+i})$ and 
-$\mathcal{L}'_{i, k} = (\mathcal{L}'_{i, k, j} \mid 1 \leq j \leq p_{n+i+1})$ 
-for $1 \leq i \leq m$ and $1 \leq k \leq 3$. 
-Again, observe that 
-$\mathcal{C}_{i, j_1} \subset \mathcal{L}_{i, k, j_2} \subset \mathcal{L}_{i, k, j_2}$ 
-for all $1 \leq i \leq m$, $1 \leq j_1 \leq q_{n+i}$, $1 \leq j_2 \leq r_{n+i}$, 
-$1 \leq j_3 \leq p_{n+i+1}$ and $1 \leq k \leq 3$.
-
-Our construction ends with $p_1 + K - 1$ \emph{utility subsets}. 
-These subsets will be partitioned into two separate classes according to their 
-indented function: \emph{bootstrap subsets} and \emph{separator subsets}. 
-First, $C$ contains identical bootstrap subsets 
-$\mathcal{B}_i$, $1 \leq i \leq p_1 - 1$, defined as follows:
-$\mathcal{B}_i = \emptyset$ for $1 \leq i \leq p_1 - 1$. The idea is to force any \SBL\ to map the $p_1 - 1$ 
-empty sets of $\mathcal{B}$ to the first $p_1  - 1 = 3(n + m) + 1$ integers. 
-Indeed, it is easily seen that all the above defined subsets of the configuration 
-of subsets $C$ but those of $\mathcal{B}$ contain at least $p_1$ elements and 
-hence cannot be mapped to an integer $i \leq p_1 - 1$ in any \SBL\ of 
-$C$. 
-Second, $C$ contains identical separator subsets $\mathcal{S}_i$, $1 \leq i \leq K$, 
-defined by:
-$\mathcal{S}_i = 
-\mathbf{V} \cup \mathbf{V}' \cup \mathbf{C} \cup \mathbf{C}' \cup \mathbf{S}$
-for $1 \leq i \leq K$.
-The rationale of these subsets is that we need a separator between subsets in $C$ 
-corresponding to a satisfying truth assignment $f$ for the $3$CNF formula $\phi$ 
-and garbage subsets of $C$, that is subsets not involved in the satisfying truth 
-assignment $f$. For simplicity, let us denote by $\mathcal{B}$ the collection 
-$(\mathcal{B}_i \mid 1 \leq i \leq p_1 - 1)$ and by $\mathcal{S}$ 
-the collection $(\mathcal{S}_i \mid 1 \leq i \leq K)$. 
-Clearly our construction can be carried on in polynomial time: indeed, we have
-$|\mathbf{X}| = O(m^2 + n^2)$ and
-$|C| = O(m^2 + n^2)$.
-
-\begin{lemma}
-  \label{lemma:Stepwise Bounded Labeling is NPC}
-  There exists a satisfying truth assignment $f$ for $\phi$ if and only if 
-  there exists an \SBL\ of the configuration of subsets 
-  $C$ of the ground set $\mathbf{X}$.
-\end{lemma}
-
-The key elements of the proof are as follows.
-First, it is crucial to focus on solutions that map identical subsets 
-of elements of $\mathcal{S}$ to a set of consecutive elements
-(see Lemma~\ref{lemma:normalized stepwise bounded}).
-Second,
-the general shape of the solution is largely guided by the construction.
-Indeed,
-the empty subsets have to be placed first, followed by subsets corresponding to literals 
-(either the positive or the negative literal of each variable has been
-chosen)
-and next by subsets corresponding to clauses
-(one satisfying literal of each clause is chose).
-Finally the separator subsets have to be placed, with the result that 
-(thanks to the large polynomial number of such subsets)
-the remaining subsets can be placed in any order without violating
-the sought \SBL\ property.
-The reader is invited to consider 
-Figure~\ref{fig:stepwise bounded labeling construction}
-for a schematic illustration of the reduction.
-We now briefly discuss, in an informal way, the two key arguments that are used in 
-the proof.
-First, the whole procedure is, to some extent, similar to 
-the accounting method used in amortized complexity analysis. 
-Indeed, one might view the operation of placing a set (one after the other)
-as the process of charging some customer, the cost being the number of
-new elements that are introduced.
-With this metaphor in mind, notice that 
-we do not charge when a subset does not introduce any new element,
-so that the leftover amount can be stored as "credit". 
-When we place a new subset that does introduce some new elements, 
-we can use the "credit" stored to pay for the cost of the operation.
-Second, when a subset uses the "credit" stored to pay the cost
-of introducing new elements, the following invariants can be shown 
-to hold true:
-(i) it uses all the available credit and
-(ii) it does not allow to accumulate 
-(it should be now clear that consecutive identical subsets do allow
-for accumulating credit)
-as much credit as it has consumed,
-thereby proving that subsets introduce less and less new elements as we progress
-adding subsets one after the other.
-
-The main result of this paper can now be stated.
-
-\begin{theorem}
-  \label{corollary:PAQ triangular is NPC}
-  Let $A$ be a $(0,1)$-matrix.
-  Deciding whether $A$ is a 
-  \PET\ matrix is \NPC.
-\end{theorem}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Exponential-time algorithm ----
-%%
-\section{Exponential-time algorithm}
-\label{section:Exponential-time algorithm}
-
-We present here an exponential-time algorithm for deciding whether a
-given a $(0,1)$-matrix $A$ of order $n$ is a \PET\ matrix.
-We start by presenting some basic properties of square $(0,1)$-matrices
-that can be transformed into some triangular matrix by row and
-column independent permutations to help solving involved 
-algorithmic issues. We of course focus of polynomial-time checkable
-properties.
-
-We first focus on the permanent of a square $(0,1)$-matrix.
-A well-known result (see e.g. \cite{Brualdi:Ryser:1991})
-states that for a $(0,1)$-matrix $A$ of order $n$,
-one has $\PER(A)=1$ if and only if the lines of $A$
-may be permuted to yield a triangular matrix with $1$'s
-in the $n$ main diagonal positions and $0$'s above the mail diagonal.
-This theorem amounts to saying that
-$\PER(A)=1$ if and only if there exist permutation matrices $P$ and $Q$
-such that $I \leq PAQ \leq \lltm$.
-As shown in the following lemma, $\PER(A)=1$ is certainly 
-a threshold value in our context.
-
-\begin{lemma}
-    \label{lemma:permanent}
-Let $A$ be $(0,1)$-matrix.
-If $A$ is a \PET\  matrix
-then $\PER(A) \leq 1$.
-\end{lemma}
-
-Notice that deciding $\PER(A) \leq 1$ for $(0,1)$-matrices of order $n$ 
-reduces to computing at most
-$n+1$ perfect matchings in bipartite graphs 
-\cite{Brualdi:Ryser:1991},
-and hence the above test is 
-$O(n^3\,\sqrt{n})$ time 
-as the Hopcroft–Karp algorithm for computing a maximum matching
-in a bipartite graph $B = (V,E)$ runs in 
-$O(|E|\,\sqrt{|V|})$~\cite{HopcroftKarp:1975}.
-
-Next, it is a simple matter to check that 
-if a $(0,1)$-matrix $A$ of order $n$ is a \PET\ matrix, 
-then it contains at most $\frac{1}{2}n(n+1)$ $1$'s 
-(\emph{i.e.}, $\omega(A) \leq \frac{1}{2}n(n+1)$).
-The following lemma gives a lower bound.
-
-\begin{lemma}
-    \label{lemma:just a few 1s}
-Let $A$ be $(0,1)$-matrix of order $n$, $n \geq 2$.
-If $A$ contains at most $n+1$ $1$'s, then $A$ is a \PET\ matrix.
-\end{lemma}
-
-Notice that, albeit not very impressive, 
-Lemma~\ref{lemma:just a few 1s} is tight as the square matrix
-$\begin{bmatrix}
-I_{n-2} & \zeros_{n-2,2} \\
-\zeros_{2,n-2} & J_{2}
-\end{bmatrix}$
-of order $n$ has $n-2+4 = n + 2$ $1$'s and is not a \PET\ matrix.
-
-Finally, the following trivial lemma gives another condition
-that helps improving the running time of the algorithm in practice.
-
-\begin{lemma}
-  \label{lemma:acyclic}
-  Let $A$ be $(0,1)$-matrix of order $n$
-  and $D$ the directed graph associated to $A$
-  (\emph{i.e.}, the adjacency matrix of $D$ is $A$).
-  % the $(0,1)$-matrix obtained from $A$
-  % by replacing every non-zero diagonal matrix element by zero).
-  If the digraph $D$ is acyclic (regardless self-loops),
-  then $A$ is a \PET\ matrix.
-\end{lemma}
-
-We now turn to presenting the exponential-time algorithm.
-The simplest exhaustive algorithm considers every possible pairs
-of permutation matrices $(P, Q)$ yielding a $O((n!)^2 \;\cdot\; \POLY(n))$ time
-algorithm.
-% , where $f(n)$ is the time needed to compute $PAQ$ and check
-% $PAQ \leq \lltm_{\,n}$.
-However, according to 
-Lemma~\ref{lemma:bounded stepwise property <=> PAQ triangular},
-it is enough to consider every permutation matrix $P$ of order $n$
-and check whether the first $i$, $1 \leq i \leq n$, rows of $PA$
-have $1$'s in at most $i$ columns.
-This observation yields a $O(n! \,\cdot\, \POLY(n))$ time algorithm.
-% where $g(n)$ is the time needed to
-% compute $PA$ and check that for the first $i$, $1 \leq i \leq n$,
-% rows of $PA$ have $1$'s in at most $i$ columns.
-We propose here another exhaustive algorithm that improves on the
-$O(n! \,\cdot\, \POLY(n))$ time algorithm. 
-The basic idea is to recursively split into smaller submatrices, 
-instead of enumerating all permutations.
-For a $(0,1)$-matrix $A$ of order $n$, we consider every possible set $R$
-of $\left\lceil n/2 \right\rceil$ rows of $A$ and 
-every possible set of $\left\lceil n/2 \right\rceil$ columns $C$ of $A$,
-and check whether these lines induce a zero matrix
-(or a matrix with at most one $1$ in case the matrix has odd order; 
-details follow).
-%Let $A' = A[R, C]$, that is $A'$ is the submatrix of $A$ with rows $R$ and columns $C$.
-
-\begin{figure}[!t]
-    \centering
-    \subfigure[Even]{%
-$
-PAQ =
-  \begin{bmatrix}
-    A_1 & \zeros \\
-    * & A_2
-  \end{bmatrix}
-$
-    \label{fig:mat:even}}
-    \qquad
-    \subfigure[Odd and one $1$]{%
-$
-PAQ =
-  \begin{bmatrix}
-    A_1 & \zeros & \zeros \\
-    *       & 1      & \zeros \\
-    *       & *      & A_2
-  \end{bmatrix}
-$
-    \label{fig:mat:odd 1}}
-    \qquad
-    \subfigure[Odd and zero $1$]{%
-$
-PAQ =
-  \begin{bmatrix}
-    A_1 & \zeros & \zeros \\
-    *       & 0      & \zeros \\
-    *       & *      & A_2
-  \end{bmatrix}
-$
-    \label{fig:mat:odd 0}}
-    \caption{\label{fig:matrices 1 and 2}}
-\end{figure}
-If $n$ is even, we let $P$ and $Q$ be two permutation matrices that put the rows in $R$
-at the first $\left\lceil n/2 \right\rceil$ positions and the columns in $C$ 
-at the last $\left\lceil n/2 \right\rceil$ positions.
-The key element for the improvement
-is that no specific order is required 
-for the rows in $R$ nor for the columns in $C$.
-The algorithm rejects the matrix $A$ for the subsets $R$ and $C$ 
-if $\omega(A[R, C]) > 1$, otherwise we can write $PAQ$ as in
-Figure~\ref{fig:mat:even},
-where $A_1$ and $A_2$ are matrices of order $\left\lceil n/2 \right\rceil = n/2$,
-and we proceed to recursively check that
-both $A_1$ and $A_2$ are \PET\  matrices.
-% The algorithm reject the matrix $A$ for the subsets $R$ and $C$ 
-% if any of the above-mentioned checks fails, and accept otherwise.
-
-The case $n$ is odd is a little bit more complicated.
-%Recall that $A' = A[R,C]$.
-First, the algorithm rejects matrix $A$ for the subsets $R$ and $C$ 
-if $\omega(A[R, C]) > 1$.
-Otherwise,
-we need to consider two (possibly positive) cases:
-(i) $\omega(A[R, C]) = 1$ 
-%(\emph{i.e.}, the submatrix $A[R, C]$ contains exactly one $1$) 
-or 
-(ii) $\omega(A[R, C]) = 0$.
-% (\emph{i.e.}, $A[R, C] = \zeros$).
-If $\omega(A[R, C]) = 1$, 
-we let $P$ and $Q$ be two permutation matrices that put the rows in $R$
-at the first $\left\lceil n/2 \right\rceil$ positions and the columns of $C$ 
-at the last $\left\lceil n/2 \right\rceil$ positions
-(no specific order for the rows in $R$ nor for the columns in $C$, except that the $1$ of
-$A$ is at row index $\left\lceil n/2 \right\rceil$ and at column index $\left\lceil n/2 \right\rceil$
-in $PAQ$).
-We can write $PAQ$ as in Figure~\ref{fig:mat:odd 1},
-where $A_1$ and $A_2$ are matrices of order $\left\lceil n/2 \right\rceil - 1 = n/2$,
-and we proceed to recursively check that
-both $A_1$ and $A_2$ are \PET\  matrices.
-Finally, if $\omega(A[R, C]) = 0$, for every row index $i \in R$ and
-every column index $j \in C$,
-we let $P$ and $Q$ be two permutation matrices that put the rows in $R$
-at the first $\left\lceil n/2 \right\rceil$ positions and the columns of $C$ 
-at the last $\left\lceil n/2 \right\rceil$ positions
-(no specific order for the rows in $R$ nor for the columns in $C$ except that row $i$ in
-$A$ is at row index $\left\lceil n/2 \right\rceil$ and column index $j$ is at
-column index $\left\lceil n/2 \right\rceil$ in $PAQ$).
-We can write $PAQ$ as in Figure~\ref{fig:mat:odd 0},
-where $A_1$ and $A_2$ are matrices of order $\left\lceil n/2 \right\rceil - 1 = n/2$,
-and we proceed to recursively check that
-both $A_1$ and $A_2$ are \PET\ matrices.
-
-\begin{figure}[!t]
-  \begin{algorithm}[H]
-    \TitleOfAlgo{\ALGO{permTriangular}}
-    \DontPrintSemicolon
-    \KwData{A square matrix $A =[a_{i,j}]$ of order $n$}
-    \KwResult{true if $A$ is a \PET\ matrix,
-    false otherwise}
-
-    \lIf{$(\omega(A) \leq n+1)$ or $(\PER(A) = 1)$ or ($A$ is stepwise bounded) or ($D(A)$ is acyclic)}{
-        \Return true
+\newcommand\INCREASING[1]{%
+  \begin{tikzpicture}[>=latex]
+  \draw [->] (0,0) -- node[midway,below] {$#1$} -- ++ (1,1);
+  \end{tikzpicture}
 }
 
-    \lIf{$(\omega(A) > \frac{n(n+1)}{2})$ or $(\PER(A) > 1)$ or ($\RSV(A)$ or $\CSV(A)$ is not stepwise bounded)}{
-        \Return false
-    }
-
-    \For{every subset $R \subset [n]$ of size $\left\lceil\frac{n}{2}\right\rceil$ and
-    every subset $C \subset [n]$ of size $\left\lceil\frac{n}{2}\right\rceil$}{
-        \eIf{$n$ is even}{
-            \Return $\ALGO{permTriangularEven}(A, R, C)$
-        }{
-            \Return $\ALGO{permTriangularOdd}(A, R, C)$
-        }
-    }
-    \Return false
-    \caption{\label{algo:permTriangular}%
-    Recognizing \PET\ matrices.}
-  \end{algorithm}
+\newpage
+\begin{figure}[t!]
+  \centering
+  \begin{tikzpicture}[
+    >=stealth',
+    shorten >=1pt,
+    main node/.style={align=center}
+    %cell node/.style={fill=black!5}
+    ]
+    % nodes
+    % 1
+    \draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
+    \node [main node] (1a) at (0.5,10.5) {$\nu_1$};
+    \draw [draw,ultra thick,fill=black!5] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
+    \node [main node] (1b) at (2.5,11.5) {$\nu'_1$};
+    % 2
+    \draw [draw,ultra thick,fill=black!5] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
+    \node [main node] (2a) at (1.5,1.5) {$\nu_2$};
+    \draw [draw,ultra thick,fill=black!5] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
+    \node [main node] (2b) at (6.5,0.5) {$\nu'_2$};
+    % 3
+    \draw [draw,ultra thick,fill=black!5] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
+    \node [main node] (3a) at (3.5,8.5) {$\nu_3$};
+    \draw [draw,ultra thick,fill=black!5] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
+    \node [main node] (3b) at (7.5,9.5) {$\nu'_3$};
+    % 4
+    \draw [draw,ultra thick,fill=black!5] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
+    \node [main node] (3a) at (5.5,7.5) {$\nu_4$};
+    \draw [draw,ultra thick,fill=black!5] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
+    \node [main node] (3b) at (9.5,2.5) {$\nu'_4$};
+    % sigma 1
+    \draw [draw,ultra thick,fill=black!5] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
+    \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
+    % pi 1
+    \draw [draw,ultra thick,fill=black!5] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
+    \node [main node] (pi1) at (8.5,5.5) {$\pi'$};    
+    % pi 2
+    \draw [draw,ultra thick,fill=black!5] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
+    \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
+    % sigma 2
+    \draw [draw,ultra thick,fill=black!5] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
+    \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};    
+    % edges
+    % 1
+    \draw[line width=1.5pt] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
+    \node (claim edge 1) at (1.5,13) {Claim\ref{}};
+    % 2
+    \draw[line width=1.5pt] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
+    \node (claim edge 1) at (4,3.5) {Claim\ref{}};
+    % 3
+    \draw[line width=1.5pt] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
+    \node (claim edge 1) at (5.5,11.5) {Claim\ref{}};
+    % 4
+    \draw[line width=1.5pt] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
+    \node (claim edge 1) at (8.5,8.5) {Claim\ref{}};
+    % sigma 1 - pi 1
+    \draw[line width=1.5pt] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
+    \node (claim edge 1) at (6.5,4) {Claim\ref{}};
+    % pi 1 - pi 2
+    \draw[line width=1.5pt,black!20] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
+    % pi 2 - sigma 2
+    \draw[line width=1.5pt,black!20] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
+    % grid
+    \draw[step=1cm,gray,ultra thin] (0,0) grid (12,12);
+  \end{tikzpicture}
 \end{figure}
-
-\begin{figure}[!t]
-    \begin{algorithm}[H]
-    \TitleOfAlgo{\ALGO{permTriangularEven}}
-    \DontPrintSemicolon
-    \KwData{A square matrix $A =[a_{i,j}]$ of even order $n$, and non-empty subsets
-    $R \subset [n]$ and $C \subset [n]$, both of size $\frac{n}{2}$}
-    \KwResult{true if $A$ is a \PET\ matrix
-    with $A[R,C]$ as the upper right submatrix, false otherwise}
-    \lIf{$\omega(A[R, C]) > 0$}{
-        % \tcp{$R$ and $C$ do not induce the zero matrix}
-        \Return false
-    }
-
-    Let $A_{\text{ul}} = A[R,\overline{C}]$ and $A_{\text{lr}} = A[\overline{R},C]$ \\
-    \Return $\ALGO{permTriangular}(A_{\text{ul}}) \;\&\&\; \ALGO{permTriangular}(A_{\text{lr}})$
-    \caption{\label{algo:permTriangularEven}%
-    Subprocedure for recognizing \PET\ matrices of even order.}
-    \end{algorithm}
-\end{figure}
-
-\begin{figure}[!t]
-    \begin{algorithm}[H]
-    \TitleOfAlgo{\ALGO{permTriangularOdd}}
-    \DontPrintSemicolon
-    \KwData{A square matrix $A =[a_{i,j}]$ of odd order $n$, and non-empty subsets
-     $R \subset [n]$ and $C \subset [n]$, both of size $\left\lceil\frac{n}{2}\right\rceil$}
-    \KwResult{true if $A$ is a \PET\ matrix
-    with $A[R,C]$ as the upper right submatrix}
-    \;
-    \lIf{$\omega(A[R, C]) > 1$}{
-        % \tcp{$R$ and $C$ do not induce a matrix with at most one $1$}
-        \Return false
-    }
-
-    \eIf{$\omega(A) = 0$}{
-        % \tcp{$R$ and $C$ induce a zero matrix}
-        \For{every $i \in R$ and every $j \in C$}{
-            Let $A_{\text{ul}} = A[R\setminus\{i\},\overline{C}]$ and
-            $A_{\text{lr}} = A[\overline{R},C\setminus\{j\}]$ \;
-             \lIf{$\ALGO{permTriangular}(A_{\text{ul}}) \;\&\&\;  \ALGO{permTriangular}(A_{\text{lr}})$}{
- %           \If{both $\ALGO{permTriangular}(A_{\text{ul}})$ and $\ALGO{permTriangular}(A_{\text{lr}})$ return true}{
-               \Return true
-            }
-        }
-        \Return false
-    }{
-        % \tcp{$R$ and $C$ induce a matrix with exactly one $1$}
-        Let $i$ and $j$ be the row and column indices of the unique $1$ in $A[R, C]$\;
-        Let $A_{\text{ul}} = A[R\setminus\{i\},\overline{C}]$ and
-        $A_{\text{lr}} = A[\overline{R},C\setminus\{j\}]$ \;
-        \Return{$\ALGO{permTriangular}(A_{\text{ul}})) \;\&\&\; \ALGO{permTriangular}(A_{\text{lr}})$}
-    }
-    \caption{\label{algo:permTriangularOdd}%
-    Subprocedure for recognizing \PET\ matrices of odd order.}
-    \end{algorithm}
-\end{figure}
-
-A detailed description is given in
-Algorithms~\ref{algo:permTriangular},
-\ref{algo:permTriangularEven} and
-\ref{algo:permTriangularOdd}.
-We now turn to evaluating the time complexity of this algorithm
-and we write $T(n)$ for the time complexity of calling
-$\ALGO{permTriangular}(A)$ for some $(0,1)$-matrix $A$ or order $n$.
-We have
-$$
-T(n) \leq
-\begin{cases} 
-\left(\left\lceil n/2 \right\rceil\right)^2 \binom{n}{\left\lceil n/2 \right\rceil}^2
-\left(2T\left(\left\lfloor n/2 \right\rfloor\right)+1\right) 
-+ O(n^3\sqrt{n}) & \text{if $n$ is odd} \\
-2\,\left(\left\lceil n/2 \right\rceil\right)^2 \binom{n}{\left\lceil n/2 \right\rceil}^2\;
-T\left(\left\lfloor n/2 \right\rfloor\right) 
-+ O(n^3\sqrt{n}) & \text{if $n$ is even}
-\end{cases}
-$$
-with $T(1) = O(1)$.
-The $O(n^3\sqrt{n})$ term
-is the time complexity for lines $2$ and $3$ in 
-Algorithm~\ref{algo:permTriangular}.
-We also observe that the worst case occurs when $n = 2^m - 1$ as 
-$\left\lceil n/2 \right\rceil,
-\left\lceil n/4 \right\rceil,
-\ldots$
-are odd integers.
-Looking for an asymptotic solution of the worst case, we thus write
-the following simplified recurrence:
-$
-T(2^m)
-=
-2^{2m - 2} \,
-\binom{2^m}{2^{m-1}}^2 \,
-\left(2T(2^{m-1}) + 1\right) + 2^{7m/6}
-$,
-with $T(1) = 1$. 
-Now, write $\alpha(2^m) = 2^{2m - 2} \, \binom{2^m}{2^{m-1}}^2$.
-Clearly, $\alpha(2^m) \geq 2^{7m/6}$, and hence we focus for now on on
-the recurrence
-$T(2^m) = 2\,\alpha(2^m)\,\left(T(2^{m-1}) + 1\right)$.
-A convenient non-recursive form of $T(2^m)$ is given in the following lemma.
-
-\begin{lemma}
-\label{lemma:T}
-$
-T(2^m)
-=
-\left(
-2^m \prod_{i=1}^{m} \alpha(2^{i})
-\right)
-+
-\left(
-\sum_{i=1}^{m}
-2^{m-i+1} \prod_{j=i}^{m} \alpha(2^{j})
-\right)
-$.
-\end{lemma}
-
-We now need the following lemma, in order to give an asymptotic
-solution for $T(n)$ in Proposition~\ref{proposition:running time}.
-
-% \begin{lemma}
-% \label{lemma:prod alpha}
-% $$
-% \prod_{i=1}^{m} \alpha(2^i)
-% =
-% O\left(
-% \frac{2^{2^{m+2}-3}}{\pi^m}
-% \right)
-% \text{.}
-% $$
-% \end{lemma}
-
-\begin{lemma}
-\label{lemma:sum prod alpha}
-$
-\sum_{i=1}^{m}
-2^{m-i} \prod_{j=i}^{m} \alpha(2^{j})
-=
-O\left(m \, 2^{2^{m+2}+m+1}\right)
-$.
-\end{lemma}
-
-\begin{proposition}
-\label{proposition:running time}
-Algorithm~\ALGO{permTriangular} runs in 
-$O\left(
-n\,2^{4n}\,\pi^{-\log(n)}
-\right)$ time.
-\end{proposition}
-
-\begin{proof}%[of Proposition~\ref{proposition:running time}]
-We have already observed that the worst case occurs for $n=2^m-1$.
-According to Lemma~\ref{lemma:sum prod alpha},
-we have
-$
-T(2^m) =
-O\left(
-2^{2^{m+2}+m-3}\,\pi^{-m}
-\right)$
-and hence
-$
-T(n) = 
-O\left(
-2^{2^{\log(n)+2}+\log(n)-3}\,\pi^{-\log(n)}
-\right)
-=
-O\left(
-n\,2^{4n}\,\pi^{-\log(n)}
-\right)$.
-\qed
-\end{proof}
 
@@
 
-We suggest for further research directions regarding 
-the hardness of recognizing \PET\ $(0,1)$-matrices.
-(i)
-Suppose a $(0,1)$-matrix $A$ of order $n$ has $n+k$ $1$'s with
-$n+2 \leq n+k \leq  \frac{1}{2}n(n+1)$. 
-Can one decide in $f(k)\,n^{O(1)}$ time whether 
-$A$ is a \PET\ matrix, where $f$ is an arbitrary function depending 
-only on $k$?
-(ii)
-What is the average running time of Algorithm~\ALGO{permTriangular}
-for \PET\ matrices?
-(iii)
-A graph labeling strongly related to symmetric \PET\ $(0,1)$-matrices
-can be defined as follows: Given a graph $G=(V,E)$ or order $n$,
-decide whether there exists a bijective mapping $f : V \to [n]$ such that
-$f(u) + f(v) > n$ for every edge $\{u,v\} \in E$
-(\emph{i.e.}, $PAP^T \leq \lrtm_{\,n}$).
-Investigating the relationships between the two combinatorial problems
-is expected to yield fruitful results.
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla 
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla 
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla 
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla 
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla .
 
@@
 
-\begin{proof}[of Lemma~\ref{lemma:bounded stepwise property <=> PAQ triangular}]
-The forward direction if obvious.
-For the reverse direction,
-let $\mathcal{S} = (S_i : 1 \leq i \leq n)$ be a configuration of subsets of
-some ground set of cardinality $n$
-and $\varphi : \mathcal{S} \to [n]$ be a bijective mapping
-such that $\left|\bigcup_{S \in \mathcal{S} : \varphi(S) \leq i} S\right| \leq i$
-for all $1\leq i \leq n$.
-Let $A = [a_{i,j}]$ be the incidence matrix of $\mathcal{S}$.
-Now let $P$ be the permutation matrix of order $n$ that permutes
-the rows of $A$ so that row $i$ of $PA$ correspond to subset
-$\varphi^{-1}(i)$ for all $1 \leq i \leq n$.
-We denote this row-permuted matrix by $A' = [a'_{i,j}]$.
-Define the function $\tau : [n] \to [n+1]$ that indicates for each column index
-$j$ the minimum row index $i$ so that $a'_{i,j} = 1$, and we adhere to the convention
-that $\tau(j) = n+1$ if row $j$ does not contain a $1$.
-Define a permutation matrix $Q$ or order $n$ that permutes the
-columns of $A'$ by ascending $\tau$ values,
-breaking ties arbitrarily.
-We designate this row-permuted matrix by $A'' = [a''_{i,j}]$.
-We claim that $A'' = PAQ \leq \lltm_{\,n}$.
-Indeed, suppose, aiming at a contradiction, that
-$A'' \not\leq \lltm_{\,n}$.
-Let $i_{\min}$ be the smallest row index such that
-$a''_{i_{\min},j_{\min}} = 1$
-for some $j_{\min} > i_{\min}$.
-Then it follows that $\tau(j) \leq \tau(j_{\min})$ for $j < j_{\min}$
-and hence
-$\left|\bigcup_{S \in \mathcal{S} : \varphi(S) \leq i_{\min}} S\right| > i_{\min}$.
-This is the desired contradiction, and hence
-$A'' = PAQ \leq \lltm_{\,n}$.
-\qed
-\end{proof}
-
-\horiz
-
-\begin{proof}[of Lemma~\ref{lemma:normalized stepwise bounded}]
-  The proof is by contradiction. Denote by $\Phi(\mathcal{S})$ the set
-  of all \SBL\ s of the configuration of subsets $\mathcal{S}$. We claim that
-  there exists a normalized \SBL\ $\varphi \in \Phi(\mathcal{S})$
-  which maps the identical subsets of $\mathcal{S}$ to sets of consecutive
-  integers. For each $\varphi \in \Phi(\mathcal{S})$, define
-  $$
-  \mathcal{M}(\varphi) = \{S_j \mid \exists S_i, S_k \text{ s.t. }
-  \varphi(S_i) < \varphi(S_k) < \varphi(S_j) \text{ and } S_i = S_j \neq S_k\}
-  $$
-  Then, there exists a mapping $\varphi^* \in \Phi(\mathcal{S})$ such that
-  $|\mathcal{M}(\varphi^*)| \leq |\mathcal{M}(\varphi')|$ for all $\varphi' \in
-  \Phi(\mathcal{S})$. We show that $|\mathcal{M}(\varphi^*)| = 0$, and hence that
-  $\varphi^*$ is our desired mapping. Suppose, for the sake of contradiction,
-  that $|\mathcal{M}(\varphi^*)| > 0$. Let $S_j \in \mathcal{M}(\varphi^*)$ be
-  such that $\varphi^*(S_j) \leq \varphi^*(S)$ for all
-  $S \in \mathcal{M}(\varphi^*)$. Then, there exists two subsets $S_i, S_k \in \mathcal{S}$
-  such that $S_i = S_j \neq S_k$ and
-  $\varphi^*(S_i) + 1 = \varphi^*(S_k) < \varphi^*(S_j)$.
-  Consider a new labeling $\varphi'$ defined by
-  \begin{alignat*}{2}
-    &\varphi'(S) = \varphi^*(S)         &\qquad& \text{for all $S \in \mathcal{S}$ such that
-      $1 \leq \varphi^*(S) \leq \varphi^*(S_i)$} \\
-    &\varphi'(S) = \varphi^*(S) + 1     &\qquad& \text{for all $S \in \mathcal{S}$ such that
-      $\varphi^*(S_i) < \varphi^*(S) < \varphi^*(S_j)$} \\
-    &\varphi'(S_j) = \varphi^*(S_k)               &\qquad& \\
-    &\varphi'(S) = \varphi^*(S) + 1     &\qquad& \text{for all $S \in \mathcal{S}$ such that
-      $\varphi^*(S_j) < \varphi^*(S) \leq m$}
-  \end{alignat*}
-
-%   \begin{minipage}[c]{\textwidth}
-%     \bigskip
-%     \centering
-%     \input{fig/fig-normalized-layout.pstex_t}
-%     \bigskip
-%   \end{minipage}
-
-  For simplicity of notation, we write $i$ instead of $\varphi^*(S_i)$,
-  $j$ instead of $\varphi^*(S_j)$ and $k$ instead of
-  $\varphi^*(S_k)$. Observe that $i+1 = k$. We claim that $\varphi'$ is
-  an \SBL\ for $\mathcal{S}$. It is sufficient to show that 
-  \begin{align*}
-    \bigcup_{\varphi'(S) \leq \ell} S &\subseteq \bigcup_{\varphi^*(S) \leq \ell} S \\
-    \intertext{for all $k \leq \ell \leq j$. We check at once that}
-    \bigcup_{1 \leq \varphi'(S) \leq \ell} S &= \left( \bigcup_{1 \leq \varphi'(S) \leq i} S \right) \;\cup\; S_j \;\cup\; \left( \bigcup_{k+1 \leq \varphi'(S) \leq \ell} S \right) \\
-    &= \left( \bigcup_{1 \leq \varphi^*(S) \leq i} S \right) \;\cup\; S_j \;\cup\; \left( \bigcup_{k \leq \varphi^*(S) \leq \ell-1} S \right) \\
-    &= \left( \bigcup_{1 \leq \varphi^*(S) \leq \ell-1} S \right) \;\cup\; S_j \\
-    &= \bigcup_{1 \leq \varphi^*(S) \leq \ell-1} S \qquad\qquad \text{(since $S_i = S_j$)} \\
-    &\subseteq \bigcup_{1 \leq \varphi^*(S) \leq \ell} S
-  \end{align*}
-  It follows immediately that $S_j \notin \mathcal{M}(\varphi')$. Indeed, $\varphi^*(S_j) \leq \varphi^*(S)$ for all $S \in \mathcal{M}(\varphi^*)$.
-
-  We proceed to show that $|\mathcal{M}(\varphi^*)| >
-  |\mathcal{M}(\varphi')|$. If $|\mathcal{M}(\varphi')| = 0$, we are done, so
-  that we may assume $|\mathcal{M}(\varphi')| > 0$. Let $S' \in
-  \mathcal{M}(\varphi')$. Then there exist subsets $S_a$ and $S_b$ such that
-  $\varphi'(S_a) < \varphi'(S_b) < \varphi'(S')$ and $S_a = S' \neq S_b$. Now,
-  observe that we must have $\varphi^*(S_j) < \varphi^*(S')$. Again, this follows
-  from the fact that $\varphi^*(S_j) \leq \varphi^*(S)$ for all $S \in
-  \mathcal{M}(\varphi^*)$. Therefore, by construction of the labeling
-  $\varphi'$, either $\varphi^*(S_a) < \varphi^*(S_b) < \varphi^*(S')$ or $\varphi^*(S_b) <
-  \varphi^*(S_a) < \varphi^*(S')$. We claim that $S' \in \mathcal{M}(\varphi^*)$. The
-  result is certainly valid in case $\varphi^*(S_a) < \varphi^*(S_b)$. Suppose now
-  that $\varphi^*(S_b) < \varphi^*(S_a)$. Then it follows that we must have $S_a =
-  S_j$. But $S_i = S_j$. Therefore $\varphi^*(S_i) < \varphi^*(S_b) < \varphi^*(S')$
-  and $S_i = S' \neq S_b$, and hence $S' \in
-  \mathcal{M}(\varphi^*)$. Combining this with the fact that $S_j \notin
-  \mathcal{M}(\varphi')$ yields  $|\mathcal{M}(\varphi^*)| >
-  |\mathcal{M}(\varphi')|$. This contradicts the choice of
-  $\varphi^*$. Therefore, we must have $|\mathcal{M}(\varphi^*)| = 0$ and hence
-  $\varphi^*$ is a normalized \SBL\ of $\mathcal{S}$.
-  \qed
-\end{proof}
-
-\horiz
-
-\begin{proof}[of Lemma~\ref{lemma:Stepwise Bounded Labeling is NPC}]
-We claim that there exists a satisfying truth assignment $f$ for
-$\phi$ if and only if there exists an \SBL\ of the configuration of subsets $C$ of the ground set $\mathbf{X}$.
-
-  Suppose that there exists a satisfying truth assignment $f$ for the 
-  $3$CNF formula $\phi$. 
-  Due to symmetry, there is no loss of generality in assuming that each clause is 
-  satisfied by its first literal. 
-  Define an labeling $\tau$ of the configuration of subsets $C$ as follows. 
-  First, we begin our labeling construction by placing all the empty subsets of $C$, that is
-  $$
-  1 \leq \tau(\mathcal{B}_i) \leq p_1 - 1
-  $$
-  for all $\mathcal{B}_i \in B$. 
-  Next, all the subsets corresponding to the variables are ordered as follows: 
-  for all $\mathcal{V}_{i, j} \in \mathcal{V}_{i}$, $1 \leq i \leq n$,
-  $$
-  p_i + \sum_{k = 1}^{i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{V}_{i, j}) \leq p_i + q_i + \sum_{k = 1}^{i-1} (p_k + q_k + r_k) - 1
-  $$
-  The satisfying truth assignment $f$ for $\phi$ is coded in our construction as follows: 
-  for all $\mathcal{A}_{i, j} \in \mathcal{A}_i$, $1 \leq i \leq n$, 
-  where $\mathcal{A}_i = \mathcal{T}_i$ if $f(x_i) = \mbox{TRUE}$ and 
-  $\mathcal{A}_i = \mathcal{F}_i$ if $f(x_i) = \mbox{FALSE}$,
-  $$
-  p_i + q_i + \sum_{k = 1}^{i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{A}_{i, j}) \leq \sum_{k = 1}^{i} (p_k + q_k + r_k) - 1
-  $$
-  and for all $\mathcal{A}'_{i, j} \in \mathcal{A}'_i$, $1 \leq i \leq n$, 
-  where $\mathcal{A}'_i = \mathcal{T}'_i$ if $f(x_i) = \mbox{TRUE}$ and $\mathcal{A}'_i = \mathcal{F}'_i$ if $f(x_i) = \mbox{FALSE}$,
-  $$
-  \sum_{k = 1}^{i} (p_k + q_k + r_k) \leq \tau(\mathcal{A}'_{i, j}) \leq p_{i+1} + \sum_{k = 1}^{i} (p_k + q_k + r_k) - 1
-  $$
-  It easily follows that
-  $$
-  \tau(\mathcal{V}_i) < \tau(\mathcal{A}_i) < \tau(\mathcal{A}'_i) < \tau(\mathcal{V}_j) < \tau(\mathcal{A}_j) < \tau(\mathcal{A}'_j)
-  $$
-  for all $1 \leq i < j \leq n$.
-
-  Having disposed of all those subsets corresponding to the variables and to the satisfying truth assignment $f$ for $\phi$, we now turn to the clauses.
-This will be divided into two parts.
-First, the subsets of $\mathcal{C}_i$, $1 \leq i \leq m$, are ordered as follows: for all $\mathcal{C}_{i, j} \in \mathcal{C}_i$, $1 \leq i \leq m$,
-  $$
-  p_{n+i} + \sum_{k = 1}^{n+i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{C}_{i, j}) \leq p_{n+i} + q_{n+i} + \sum_{k = 1}^{n+i-1} (p_k + q_k + r_k) - 1
-  $$
-  Second, all the subsets of $\mathcal{L}_{i, 1}$ and $\mathcal{L}'_{i, 1}$, $1 \leq i \leq m$, are ordered as follows\footnote{Recall that we assume that each clause is satisfied by its first literal.}: for all $\mathcal{L}_{i, 1, j} \in \mathcal{L}_{i, 1}$, $1 \leq i \leq m$,
-  $$
-  p_{n+i} + q_{n+i} + \sum_{k = 1}^{n+i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{L}_{i, 1, j}) \leq \sum_{k = 1}^{n+i} (p_k + q_k + r_k) - 1
-  $$
-  and for all $\mathcal{L}'_{i, 1, j} \in \mathcal{L}'_{i, 1}$, $1 \leq i \leq m$,
-  $$
-  \sum_{k = 1}^{n+i} (p_k + q_k + r_k) \leq \tau(\mathcal{L}'_{i, 1, j}) \leq p_{n+i+1} + \sum_{k = 1}^{n+i} (p_k + q_k + r_k) - 1
-  $$
-  A trivial verification shows that
-  $$
-  \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i,1}) < \tau(\mathcal{L}'_{i, 1}) < \tau(\mathcal{C}_j) < \tau(\mathcal{L}_{j,1}) < \tau(\mathcal{L}'_{j, 1})
-  $$
-  for all $1 \leq i < j \leq m$.
-
-  Here come the $K$ separator subsets of $\mathcal{S}$, that is for all $\mathcal{S}_i \in \mathcal{S}$
-  $$
-  p_{n+m+1} + \sum_{k = 1}^{n+m} (p_k + q_k + r_k) \leq \tau(\mathcal{S}_i) \leq p_{n+m+1} + K + \sum_{k = 1}^{n+m} (p_k + q_k + r_k) - 1
-  $$
-  Up to now, all the subsets of $C$ but $L$ of them have been used to define $\tau$. Our labeling construction ends with those subsets by packing them in the $L$ last places:
-  $$
-  p_{n+m+1} + K + \sum_{k = 1}^{n+m} (p_k + q_k + r_k) \leq \tau(\mathcal{X}_j) \leq p_{n+m+1} + K + L + \sum_{k = 1}^{n+m} (p_k + q_k + r_k) - 1
-  $$
-  A careful examination of $\tau$ shows that $\left| \bigcup_{\tau(X)
-    \leq i} X \right| \leq i$ for $1 \leq i \leq |C|$ and hence that
-  $\tau$ is an \SBL\footnote{Observe that the construction of $\tau$
-    may be specialized to yield a normalized \SBL. Indeed, all the subsets of $C$ but the $L$ last ones are
-    mapped by $\tau$ to a set of consecutive integers. But there is no
-    less of generality in assuming that these subsets are mapped by
-    $\tau$ to a set of consecutive integers as well.} of the
-  configuration of subsets $C$ of the ground set $\mathbf{X}$. An
-  (partial) illustration of the construction of the \SBL\ $\tau$ is shown in Figure~\ref{fig:stepwise bounded
-    labeling construction}.
-
-  % \begin{figure}[ht]
-  %   \centering
-  %   \bigskip
-  %   \begin{minipage}[c]{\textwidth}
-  %     \centering
-  %     \input{r.pstex_t}
-  %     \caption{\label{fig:stepwise bounded labeling
-  %         construction}Illustration of the construction used in the
-  %       proof of proposition~\ref{proposition:Stepwise Bounded
-  %         Labeling is NPC} for the $3$CNF formula $\phi = (x_1 \vee
-  %       x_2 \vee x_3) \wedge (\overline{x}_1 \vee \overline{x}_2
-  %       \vee x_3) \wedge (\overline{x}_1 \vee x_2 \vee
-  %       \overline{x}_3)$. Identical subsets are not distinguishable
-  %       in our representation. A satisfying truth assignment is
-  %       given by $f(x_1) = \mbox{TRUE}$, $f(x_2) = \mbox{FALSE}$ and
-  %       $f(x_3) = \mbox{FALSE}$. For the sake of clarity, neither
-  %       the ground set $\mathbf{X}$ nor the collection of subsets
-  %       $C$ is fully represented. Note that we have here dropped the
-  %       assumption that each clause is satisfied by its first
-  %       literal.} 
-  %   \end{minipage}%
-
-  %   \bigskip
-  % \end{figure}
-
-  For the converse, suppose that there exists an \SBL\ $\tau$ for $C$, that is an labeling such that $\left|
-    \bigcup_{\tau(X_j) \leq i} X_j \right| \leq i$ for $1 \leq i \leq
-  |C|$. According to lemma~\ref{lemma:normalized stepwise bounded}, there is
-  no loss of generality in assuming that $\tau$ is a normalized \SBL\, \emph{i.e.}, the identical subsets of the
-  configuration $C$ are mapped by $\tau$ to a set of consecutive
-  integers. This property is crucial in our  proof.
-
-  Let us start by proving that $1 \leq \tau(\mathcal{B}_i) \leq p_1 - 1$
-  for all $\mathcal{B}_i \in \mathcal{B}$. Indeed, as mentioned earlier
-  in the proof, all the above defined subsets of the configuration of
-  subsets $C$ but those of $\mathcal{B}$ contain at least $p_1$ elements
-  and hence can not be mapped to an integer $i \leq p_1 - 1$ in any
-  \SBL\ $\tau$ for $C$.
-
-  Define the subcollection $C^* \subset C$ as follows:
-  $$
-  C^* = (X \in C \mid \max \{\tau(\mathcal{B}_i) \mid \mathcal{B}_i \in
-  \mathcal{B}\} < \tau(X) < \min \{\tau(\mathcal{S}_i) \mid
-  \mathcal{S}_i \in \mathcal{S}\}) 
-  $$
-  In other words, the subcollection $C^*$ contains all those subsets
-  that go after the last bootstrap subset $\mathcal{B}_i \in
-  \mathcal{B}$ and before the first separator subset $\mathcal{S}_i \in
-  \mathcal{S}$. We claim that $C^*$ contains (1) the subsets of
-  $\mathcal{V}_i$ for $1 \leq i \leq n$, (2) either the subsets of
-  $\mathcal{T}_i$ and $\mathcal{T}'_i$ or the subsets of $\mathcal{F}_i$
-  and $\mathcal{F}'_i$ for $1 \leq i \leq n$, (3) the subsets of
-  $\mathcal{C}_i$ for $1 \leq i \leq m$ and (4) the subsets of
-  $\mathcal{L}_{i,1}$ and $\mathcal{L}'_{i,1}$ or the subsets of
-  $\mathcal{L}_{i,2}$ and $\mathcal{L}'_{i,2}$ or the subsets of
-  $\mathcal{L}_{i,3}$ and $\mathcal{L}'_{i,3}$ for $1 \leq i \leq
-  m$. For simplicity, this will be divided into several steps.
-
-  Having disposed of the bootstrap subsets, we now turn to the non-empty
-  subsets of the configuration of subsets $C$. Roughly speaking, we show
-  that our construction implies a \emph{force placement} of the subsets
-  of $C$ together with choices corresponding to a satisfying truth
-  assignment for the $3$CNF formula $\phi$. First, we must have: 
-  $$
-  p_1 \leq \tau(\mathcal{V}_{1, j}) \leq p_1 + q_1 - 1
-  $$
-  for all $1 \leq j \leq q_1$. Indeed, all other subsets have more than
-  $p_1$ elements and hence can not be mapped to $p_1$ by
-  $\tau$. Moreover, $\tau$ is a normalized \SBL\ for
-  $C$, \emph{i.e.}, the subsets $\mathcal{V}_{1, j}$, $1 \leq j \leq
-  q_1$, are mapped by $\tau$ to a set of consecutive integers. Now, the
-  key point is that the subset which is mapped to $p_1 + q_1$ by $\tau$
-  must have cardinality at most $p_1 + q_1$. But a careful examination
-  of the configuration of subsets $C$ shows that all the remaining
-  subsets contain $\mathbf{V}_1$, and hence the subset which is mapped
-  to $p_1 + q_1$ by $\tau$ may introduce at most $q_1$ new elements of
-  the ground set. Then it follows from our construction that this subset
-  is either a subset of the collection $\mathcal{T}_1$ or a subset of
-  the collection $\mathcal{F}_1$. Indeed, it is sufficient to note that
-  $r_1 + p_2 > q_1$, and hence that no subset $\mathcal{V}_{2, j} \in
-  \mathcal{V}_2$ can satisfy $\tau(\mathcal{V}_{2, j}) = p_1 + q_1$. As
-  an immediate result, exactly one of the following two statements is
-  true: 
-  \begin{align}
-    &p_1 + q_1 \leq \tau(\mathcal{T}_{1, j}) \leq p_1 + q_1 + r_1 - 1
-    \label{eq:stepwise bounded labeling 1-T} \\ 
-    &p_1 + q_1 \leq \tau(\mathcal{F}_{1, j}) \leq p_1 + q_1 + r_1 - 1
-    \label{eq:stepwise bounded labeling 1-F} 
-  \end{align}
-  for all $1 \leq j \leq r_1$. We can now proceed analogously to the
-  above to obtain: 
-  \begin{align*}
-    \eqref{eq:stepwise bounded labeling 1-T} &\quad\Rightarrow\quad p_1 +
-    q_1 + r_1 \leq \tau(\mathcal{T}'_{1, j}) \leq p_1 + q_1 + r_1 +p_2 - 1
-    \\ 
-    \eqref{eq:stepwise bounded labeling 1-F} &\quad\Rightarrow\quad p_1 +
-    q_1 + r_1 \leq \tau(\mathcal{F}'_{1, j}) \leq p_1 + q_1 + r_1 +p_2 - 1 
-  \end{align*}
-  for all $1 \leq j \leq p_2$. Summarizing, exactly one of the following
-  two statements is true: 
-  \begin{align*}
-    p_1 \leq \tau(\mathcal{V}_1) < \tau(\mathcal{T}_1) < \tau(\mathcal{T}'_1) \leq p_1 + q_1 + r_1 +p_2 - 1 \\
-    p_1 \leq \tau(\mathcal{V}_1) < \tau(\mathcal{F}_1) < \tau(\mathcal{F}'_1) \leq p_1 + q_1 + r_1 +p_2 - 1
-  \end{align*}
-  We continue in this fashion obtaining that exactly one of the following two statements is true:
-  \begin{align*}
-    &p_i + \sum_{k=1}^{i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{V}_i) < \tau(\mathcal{T}_i) < \tau(\mathcal{T}'_i) \leq p_{i+1} + \sum_{k=1}^{i} (p_k + q_k + r_k) - 1 \\
-    &p_i + \sum_{k=1}^{i-1} (p_k + q_k + r_k) \leq \tau(\mathcal{V}_i) < \tau(\mathcal{F}_i) < \tau(\mathcal{F}'_i) \leq p_{i+1} + \sum_{k=1}^{i} (p_k + q_k + r_k) - 1 
-  \end{align*}
-  for all $1 \leq i \leq n$.
-This result is crucial as it allows us to construct a truth assignment $f$ for the $3$CNF formula $\phi$.
-Indeed, subsets of $\mathcal{T}_i$ and $\mathcal{T}'_i$ are associated in our construction to literal $x_i$ while subsets of $\mathcal{F}_i$ and $\mathcal{F}'_i$ are associated in our construction to literal $\overline{x}_i$.
-
-  Having disposed of the variables we now turn to the clauses. For the sake of clarity, let us first introduce the temporary notations
-  \begin{align*}
-    l_i &= p_{n+i} + \sum_{k=1}^{n+i-1} (p_k + q_k + r_k) \\
-    h_i  &= p_{n+i+1} + \sum_{k=1}^{n+i} (p_k + q_k + r_k) - 1
-  \end{align*}
-  for $1 \leq i \leq m$. We may now proceed analogously to the above to obtain that 
-  exactly one of the following three statements is true:
-  \begin{align*}
-    &l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i, 1}) < \tau(\mathcal{L}'_{i, 1}) \leq h_i \\
-    &l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i, 2}) < \tau(\mathcal{L}'_{i, 2}) \leq h_i \\
-    &l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i, 3}) < \tau(\mathcal{L}'_{i, 3}) \leq h_i 
-  \end{align*}
-  for all $1 \leq i \leq m$.
-
-  According to the above, 
-  the subcollection $C^*$ contains either the subsets of $\mathcal{T}_i$ and $\mathcal{T}'_i$ 
-  or the subsets of $\mathcal{F}_i$ and $\mathcal{F}'_i$ for $1 \leq i \leq n$. 
-  Therefore we can define a truth assignment $f$ for the $3$CNF formula $\phi$ as follows: 
-  $f(x_i) = \mbox{TRUE}$ if $\mathcal{T}_i \subset C^*$ and $f(x_i) = \mbox{FALSE}$ 
-  if $\mathcal{F}_i \subset C^*$ for $1 \leq i \leq n$. 
-  We claim that $f$ is a satisfying truth assignment for $\phi$. 
-  Indeed, for each $i$, $1 \leq i \leq m$, consider the true statement
-  $$
-  l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i, k}) < \tau(\mathcal{L}'_{i, k}) \leq h_i
-  $$
-  where $k$ is either $1$, $2$ or $3$. By construction we have
-  \begin{align*}
-    \mathcal{L}_{i, k, j}  &= \mathbf{V} \cup \mathbf{V}' \cup \mathbf{A}_k \cup \mathbf{L}_{i, k} \cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}_\ell \right) \cup \left( \bigcup_{1 \leq \ell \leq i-1} \mathbf{C}'_\ell \right) \\
-    \intertext{for $1 \leq j \leq r_{n+i}$ and}
-    \mathcal{L}'_{i, k, j} &= \mathbf{V} \cup \mathbf{V}' \cup \mathbf{A}_k \cup \mathbf{L}_{i, k} \cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}_\ell \right) \cup \left( \bigcup_{1 \leq \ell \leq i} \mathbf{C}'_\ell \right)
-  \end{align*}
-  for $1 \leq j \leq p_{n+i+1}$, where $\mathbf{A}_k = \mathbf{T}_\ell$
-  if $\lambda_{i, k} = x_\ell$ and $\mathbf{A}_k = \mathbf{F}_\ell$ if
-  $\lambda_{i, k} = \overline{x}_\ell$. But a careful examination of the
-  configuration of subsets $C$ and of the \SBL\
-  $\tau$ shows that none of the previous subsets contain the elements of
-  $\mathbf{L}_{i, k}$. Since $|\mathcal{C}_i| = q_{n+i}$ and
-  $|\mathbf{L}_{i, k}| = q_{n+i}$, then it follows that all the elements
-  of $\mathbf{A}_k$ must have been introduced by previous
-  subsets. Indeed, $l_i \leq \tau(\mathcal{C}_i) < \tau(\mathcal{L}_{i,
-    k}) < \tau(\mathcal{L}'_{i, k}) \leq h_i$ and hence $\mathcal{C}_i$
-  and $\mathcal{L}_{i, k}$ are mapped by $\tau$ to a set of consecutive
-  integers. Therefore, if $\mathbf{A}_k = \mathbf{T}_\ell$
-  (resp. $\mathbf{A}_k = \mathbf{F}_\ell$) then we must have chosen
-  $\mathcal{T}_\ell$ (resp. $\mathcal{F}_\ell$) in a previous step, that
-  is $\mathcal{T}_\ell, \mathcal{T}'_\ell \in C^*$
-  (resp. $\mathcal{F}_\ell, \mathcal{F}'_\ell \in C^*$). Then it follows
-  that clause $c_i$ is satisfied by its $k$-th literal. Hence, $f$ is a
-  satisfying truth assignment for $\phi$ and the lemma is proved. 
-\qed
-\end{proof}
-
-\horiz
-
-\begin{proof}[of Lemma~\ref{lemma:permanent}]
-The proof is by induction.
-The result is obvious for $n=1$.
-For the inductive step, assume the statement holds for some natural number $n$
-and let $A$ be a $(0,1)$-matrix of order $n+1$.
-Since $A$ is assumed to be a \PET\
-matrix there exist permutation matrices $P$ and $Q$ such that
-$PAQ \leq \lltm_{\,n}$.
-Let $A' = PAQ$.
-As $\PER(A)$ remains invariant under arbitrary permutation of the lines of $A$
-we have $\PER(A) = \PER(A')$.
-If $\PER(A') = 0$, we are done.
-Otherwise $\PER(A') \geq 1$ and hence
-$n$ $1$'s appear on the main diagonal of $A'$.
-We designate the matrix obtained from $A'$ by deleting
-the first row and the first column permuted matrix by $A''$.
-As the first row of $A'$ contains exactly one $1$,
-we have $\PER(A'') = \PER(A') = \PER(A) \leq 1$, thereby proving the lemma.
-\qed
-\end{proof}
-
-\horiz
-
-\begin{proof}[of Lemma \ref{lemma:just a few 1s}]
-The proof is by induction.
-The assertion is certainly valid for $n = 2$ as
-all $(0,1)$-matrix of order $2$ are \PET\ matrices but $J_2$.
-For the inductive step, assume the statement holds for some natural number $n$ and let
-$A$ be $(0,1)$-matrix of order $n+1$.
-If $A = \zeros$ we are done.
-Otherwise, let $\alpha_i$ be a row of $A$ with minimum row sum $r_i$.
-Since $A$ contains at most $n+1$ $1$'s, we certainly have $r_i \leq 1$.
-We need to consider two cases:
-\begin{itemize}
-\item
-If $r_i = 0$,
-let $\beta_j$ be a column of $A$ with maximum column sum $c_j$.
-Since $A \neq \zeros$, we certainly have $c_j > 0$.
-We now permute the lines of $A$ so that row $\alpha_i$ and
-column $\beta_j$ are the first row and column of the permuted matrix.
-We now delete the first row and column of this permuted matrix and
-apply the induction hypothesis to this submatrix of order $n$.
-\item
-If $r_i = 1$,
-we may permute the rows of $A$ so that row $1$ of the permuted matrix
-contains a $1$ in the $(1,1)$ position and $O$'s elsewhere.
-We now delete the first row and column of this matrix and apply the induction
-hypothesis to this submatrix of order $n$.
-\end{itemize}
-\qed
-\end{proof}
-
-\horiz
-
-\begin{proof}[of Lemma~\ref{lemma:T}]
-The proof is by induction on $m$.
-First, the assertion is certainly valid for $m = 1$ as
-\begin{align*}
-2\,\alpha(2^1)\,(T(2^0) + 1)
-&= 
-2\times 4 \times (1 + 1) 
-\\
-&=
-16
-\\
-\intertext{and}
-\left(
-2^1 \prod_{i=1}^{1} \alpha(2^{i})
-\right)
-+
-\left(
-\sum_{i=1}^{1}
-2^{1-i+1} \prod_{j=i}^{1} \alpha(2^{j})
-\right)
-&= (2^1 \times \alpha(2^1)) + (2^1 \times \alpha(2^1))
-\\
-&=
-(2 \times 4) + (2 \times 4) 
-\\
-&=
-16\text{.}
-\end{align*}
-
-Suppose now that the assertion is true for $2^m$.
-Then
-\begin{align*}
-T(2^{m+1})
-&=
-2\,\alpha(2^{m+1})\,T(2^{m}) + 2\,\alpha(2^{m+1})
-\\
-&=
-2\,\alpha(2^{m+1})\,
-\left(
-\left(
-2^m \prod_{i=1}^{m} \alpha(2^{i})
-\right)
-+
-\left(
-\sum_{i=1}^{m}
-2^{m-i+1} \prod_{j=i}^{m} \alpha(2^{j})
-\right)
-\right)
- + 2\,\alpha(2^{m+1})
- \\
-&=
-\left(
-2^{m+1} \prod_{i=1}^{m+1} \alpha(2^{i})
-\right)
-+
-\left(
-\sum_{i=1}^{m}
-2^{m+1-i+1} \prod_{j=i}^{m+1} \alpha(2^{j})
-\right)
- + 2\,\alpha(2^{m+1})
- \\ 
-&=
-\left(
-2^{m+1} \prod_{i=1}^{m+1} \alpha(2^{i})
-\right)
-+
-\left(
-\sum_{i=1}^{m+1}
-2^{m+1-i+1} \prod_{j=i}^{m+1} \alpha(2^{j})
-\right)
-- 2\,\alpha(2^{m+1}) 
-+ 2\,\alpha(2^{m+1})
- \\
-&=
-\left(
-2^{m+1} \prod_{i=1}^{m+1} \alpha(2^{i})
-\right)
-+
-\left(
-\sum_{i=1}^{m+1}
-2^{m+1-i+1} \prod_{j=i}^{m+1} \alpha(2^{j})
-\right)\text{.} 
-\end{align*}
-\qed
-\end{proof}
-
-\horiz
-
-\begin{proof}[of Lemma~\ref{lemma:sum prod alpha}]
-We need the following claims.
-
-\begin{claim}
-\label{claim:inner:1}
-$$
-\alpha(2^i)
-=
-\frac{2^{2^{m+1} + m - 1}}{\pi}\left(1 + O\left(2^{1-m}\right)\right)
-\text{.}
-$$
-\end{claim}
-\begin{proof}
-We use the following well-known asymptotic (see e.g. \cite{ConcreteMathematics:1994}):
-\begin{align*}
-\binom{2n}{n} &= 
-\frac{4^n}{\sqrt{\pi n}} \left(1 + O\left(n^{-1}\right)\right)\text{.}
-\\
-\intertext{Therefore,}
-\alpha(2^m)
-&=
-2^{2m-2} \, \binom{2^m}{2^{m-1}}^2
-\\
-&=
-2^{2m-2}\;
-\left(\frac{4^{2^{m-1}}}{\sqrt{\pi 2^{m-1}}} \left(1 + O\left(2^{1-m}\right)\right)\right)^2
-\\
-&=
-2^{2m-2}\;
-\left(\frac{4^{2^{m-1}}}{\sqrt{\pi 2^{m-1}}}\right)^2 \left(1 + O\left(2^{1-m}\right)\right)^2
-\\
-&=
-2^{2m-2}\;
-\frac{4^{2^m}}{\pi 2^{m-1}} \left(1 + O\left(2^{1-m}\right)\right)^2
-\\
-&=
-\frac{2^{2^{m+1} + m - 1}}{\pi}\left(1 + O\left(2^{1-m}\right)\right)^2
-\text{.}
-\\
-\intertext{We now develop the error term to obtain}
-\alpha(2^m)
-&=
-\frac{2^{2^{m+1} + m - 1}}{\pi}\,
-\left(1 + O\left(2^{1-m}\right)\right)
-\,
-\left(1 + O\left(2^{1-m}\right)\right)
-\\
-&=
-\frac{2^{2^{m+1} + m - 1}}{\pi}\,
-\left(
-1 + O\left(2^{1-m}\right) + O\left(2^{1-m}\right) + O\left(2^{2-2m}\right)
-\right)
-\\
-&=
-\frac{2^{2^{m+1} + m - 1}}{\pi}\left(1 + O\left(2^{1-m}\right)\right)
-\text{.}
-\end{align*}
-\qed
-\end{proof}
-
-\begin{claim}
-\label{claim:inner:2}
-$$
-\prod_{i=1}^{m} \alpha(2^i)
-=
-O\left(
-\frac{2^{2^{m+2}-3}}{\pi^m}
-\right)
-\text{.}
-$$
-\end{claim}
-\begin{proof}
-According to Claim~\ref{claim:inner:1}, we may write
-\begin{align*}
-\prod_{i=1}^{m} \alpha(2^i)
-&=
-\prod_{i=1}^{m} \frac{2^{2^{i+1} + i - 1}}{\pi}\left(1 + O\left(2^{1-i}\right)\right)
-\\
-&=
-\left(
-\frac{1}{2\pi}
-\right)^{m}
-\,
-\left(
-\prod_{i=1}^{m} 2^{2^{i+1}}
-\right)
-\,
-\left(
-\prod_{i=1}^{m} 2^i
-\right)
-\,
-\left(
-\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right)
-\right)
-\\
-&=
-\left(
-\frac{1}{2\pi}
-\right)^{m}
-\,
-\left(2^{2^{m+2}-4}\right)
-\,
-\left(2^{m+1}-2\right)
-\,
-\left(
-\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right)
-\right)
-\\
-&=
-\frac{2^{2^{m+2}-3}}{\pi^m}
-\;
-\left(\frac{2^m-1}{2^m}\right)
-\,
-\left(
-\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right)
-\right)
-\\
-&=
-\frac{2^{2^{m+2}-3}}{\pi^m}
-\;
-\left(1 - 2^{-m}\right)
-\,
-\left(
-\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right)
-\right)\text{.}
-\\
-\intertext{But
-$\prod_{i=1}^{m} \left(1 + O\left(2^{1-i}\right)\right) = O(1)$ and hence}
-\prod_{i=1}^{m} \alpha(2^i)
-&=
-\frac{2^{2^{m+2}-3}}{\pi^m}
-\;
-O\left(1 - 2^{-m}\right)
-\\
-&=
-O\left(
-\frac{2^{2^{m+2}-3}}{\pi^m}
-\right)
-\text{.}
-\end{align*}
-\qed
-\end{proof}
-
-Having disposed of the preliminary steps,
-write $\beta(i) = 2^{m-i} \prod_{j=i}^{m} \alpha(2^j)$ so that our goal
-reduces to evaluating $\sum_{i=1}^{m} \beta(i)$.
-For $1 \leq i < m$, we first observe that
-\begin{align*}
-2\;\beta(i+1)
-&=
-2 \; 2^{m-i-1} \prod_{j=i+1}^{m} \alpha(2^{j}) 
-\\
-&= 
-2^{m-i} \prod_{j=i+1}^{m} \alpha(2^{j}) 
-\\
-&\leq 
-2^{m-i} \prod_{j=i}^{m} \alpha(2^{j}) \qquad\text{(since $\alpha(i) \geq 1$)}
-\\
-&= 
-\beta(i)\text{.}
-\\
-\intertext{Therefore}
-\sum_{i=1}^{m} \beta(i)
-&\leq 
-\sum_{i=1}^{m} \frac{\beta(1)}{2^{i-1}}
-\\
-&=
-2\,\beta(1) \; \sum_{i=1}^{m} {2^{-i}}
-\\
-&=
-2\beta(1) \, \left(1 - 2^{-m}\right)
-\\
-&\leq 
-2\beta(1)\text{.}
-\\
-\intertext{Then it follows that}
-\sum_{i=1}^{m}
-2^{m-i} \prod_{j=i}^{m} \alpha(2^{j})
-&\leq 
-2\,\beta(1)
-\\
-&=
-2 \; 2^{m-1} \; \prod_{j=1}^{m} \alpha(2^{j})
-\\
-&=
-2^m \; O\left(
-\frac{2^{2^{m+2}-3}}{\pi^m}
-\right)
-\\
-&=
-O\left(
-\frac{2^{2^{m+2}+m-3}}{\pi^m}
-\right)\text{.}
-\end{align*}
-\qed
-\end{proof}
-
-% \newpage
-
-% \newpage
-
-% Let $X_{j,j}$ the event that $a_{i,j} = 1$:
-
-% $$
-% P[X_{i,j}] =
-% \begin{cases}
-% p & \text{if $j \leq i$} \\
-% 0 & \text{if $j > i$.}
-% \end{cases}
-% $$
-
-% \medskip
-% Let $B=PA$.
-
-% For $1 \leq i \leq n$ and $1 \leq j \leq n$,
-% let $\mathcal{X}_{i,j}$ be event $b_{i,j} = 1$.
-% We have
-
-% $$
-% \PROBABILITY[\mathcal{X}_{i,j}] = \frac{p(n-j+1)}{n}
-% $$
-
-% \medskip
-
-% For $1 \leq i \leq n$ and $1 \leq j \leq n$,
-% let $\mathcal{Y}_{i,j}$ be the event
-% $\sum_{k=1}^{i} b_{k,j} > 0$,
-% and let $\INDICATOR_{\mathcal{Y}_{i,j}}$ the indicator variable for event
-% $\mathcal{Y}_{i,j}$.
-
-% \begin{align*}
-% \PROBABILITY\left[\mathcal{Y}_{i,j}\right]
-% &=
-% 1 - \PROBABILITY\left[\,\overline{\mathcal{Y}_{i,j}}\,\right] \\
-% &=
-% 1 - \PROBABILITY\left[ \bigcap_{1\leq k \leq i} \overline{\mathcal{X}_{k,j}}\right] \\
-% &=
-% 1 - \prod_{k=1}^{i} \left(1 - \PROBABILITY\left[\mathcal{X}_{k,j}\right]\right) \\
-% &=
-% 1 - \prod_{k=1}^{i} \left( 1 - \frac{p(n-j+1)}{n}\right) \\
-% &=
-% 1 - \left( 1 - \frac{p(n-j+1)}{n} \right)^i
-% \end{align*}
-
-% For $1 \leq i \leq n$, define the random variable
-% $Z_i = \sum_{j=1}^{n} \INDICATOR_{\mathcal{Y}_{i,j}}$
-
-% \begin{align*}
-% P[Z_i = k]
-% &\geq
-% \binom{n}{k} \; \left(1-\PROBABILITY\left[\mathcal{Y}_{i,1}\right]\right)^{n-k} \;
-% \left(\PROBABILITY\left[\mathcal{Y}_{i,n}\right]\right)^{k} \\
-% \intertext{and hence}
-% P[Z_i < i+1]
-% &\geq
-% \sum_{k=0}^{i}
-% \binom{n}{k} \;
-% \left(1-\PROBABILITY\left[\mathcal{Y}_{i,1}\right]\right)^{n-k} \;
-% \left(\PROBABILITY\left[\mathcal{Y}_{i,n}\right]\right)^{k} \\
-% &=
-% \sum_{k=0}^{i}
-% \binom{n}{k} \;
-% \left(1 - \left(1 - \left(1 - p\right)^{i}\right)\right)^{n-k}
-% \left(1-\left(1 - \frac{p}{n}\right)^{i}\right)^{k} \\
-% &=
-% \sum_{k=0}^{i}
-% \binom{n}{k} \;
-% \left(\left(1-p\right)^{i}\right)^{n-k}
-% \left(1-\left(1 - \frac{p}{n}\right)^{i}\right)^{k} \\
-% &\geq
-% \sum_{k=0}^{i}
-% \binom{i}{k} \;
-% \left(\left(1-p\right)^{i}\right)^{i-k}
-% \left(1-\left(1 - \frac{p}{n}\right)^{i}\right)^{k} \\
-% &=
-% \left(1 + \left(1-p\right)^{i} - \left(1 - \frac{p}{n}\right)^{i}\right)^{i}
-% \text{.}
-% \end{align*}
-
-% Let $\mathcal{E}(B)$ be the event that $B$ does not have
-% the stepwise bounded property.
-
-% \begin{align*}
-% \PROBABILITY\left[\mathcal{E}(B)\right]
-% &=
-% \PROBABILITY\left[\mathcal{E}(B)\right] \\
-% &\leq
-% \sum_{i=1}^{n} \PROBABILITY\left[Z_i \geq i+1\right] \\
-% &=
-% \sum_{i=1}^{n} \left(1 - \PROBABILITY\left[Z_i < i+1 \right]\right) \\
-% \end{align*}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-06-26 article.tex
figure

@@
     ]  
-    % nodes
-    % 1
-    \draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
-    \node [main node] (1a) at (0.5,10.5) {$\nu_1$};
-    \draw [draw,ultra thick,fill=black!5] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
-    \node [main node] (1b) at (2.5,11.5) {$\nu'_1$};
-    % 2
-    \draw [draw,ultra thick,fill=black!5] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
-    \node [main node] (2a) at (1.5,1.5) {$\nu_2$};
-    \draw [draw,ultra thick,fill=black!5] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
-    \node [main node] (2b) at (6.5,0.5) {$\nu'_2$};
-    % 3
-    \draw [draw,ultra thick,fill=black!5] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
-    \node [main node] (3a) at (3.5,8.5) {$\nu_3$};
-    \draw [draw,ultra thick,fill=black!5] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
-    \node [main node] (3b) at (7.5,9.5) {$\nu'_3$};
-    % 4
-    \draw [draw,ultra thick,fill=black!5] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
-    \node [main node] (3a) at (5.5,7.5) {$\nu_4$};
-    \draw [draw,ultra thick,fill=black!5] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
-    \node [main node] (3b) at (9.5,2.5) {$\nu'_4$};
-    % sigma 1
-    \draw [draw,ultra thick,fill=black!5] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
-    \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
-    % pi 1
-    \draw [draw,ultra thick,fill=black!5] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
-    \node [main node] (pi1) at (8.5,5.5) {$\pi'$};    
-    % pi 2
-    \draw [draw,ultra thick,fill=black!5] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
-    \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
-    % sigma 2
-    \draw [draw,ultra thick,fill=black!5] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
-    \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};    
     % edges
+    % pi 1 - pi 2
+    \draw[line width=1.5pt,black!20] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
+    % pi 2 - sigma 2
+    \draw[line width=1.5pt,black!20] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
     % 1
@@
     \node (claim edge 1) at (6.5,4) {Claim\ref{}};
-    % pi 1 - pi 2
-    \draw[line width=1.5pt,black!20] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
-    % pi 2 - sigma 2
-    \draw[line width=1.5pt,black!20] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
+    % nodes
+    % 1
+    \draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
+    \draw [->,ultra thick] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_1$};
+    %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
+    \draw [draw,ultra thick,fill=black!5] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
+    \draw [->,ultra thick] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_1$};
+    % 2
+    \draw [draw,ultra thick,fill=black!5] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
+    \draw [->,ultra thick] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_2$};
+    \draw [draw,ultra thick,fill=black!5] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
+    \draw [->,ultra thick] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_2$};
+    % 3
+    \draw [draw,ultra thick,fill=black!5] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
+    \draw [->,ultra thick] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_3$};
+    \draw [draw,ultra thick,fill=black!5] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
+    \draw [->,ultra thick] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_3$};
+    % 4
+    \draw [draw,ultra thick,fill=black!5] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
+    \draw [->,ultra thick] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_4$};
+    \draw [draw,ultra thick,fill=black!5] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
+    \draw [->,ultra thick] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_4$};
+    % sigma 1
+    \draw [draw,ultra thick,fill=black!5] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
+    \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
+    % pi 1
+    \draw [draw,ultra thick,fill=black!5] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
+    \node [main node] (pi1) at (8.5,5.5) {$\pi'$};    
+    % pi 2
+    \draw [draw,ultra thick,fill=black!5] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
+    \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
+    % sigma 2
+    \draw [draw,ultra thick,fill=black!5] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
+    \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};  
     % grid
-    \draw[step=1cm,gray,ultra thin] (0,0) grid (12,12);
+    \draw[step=1cm,black!30,ultra thin] (0,0) grid (12,12);
   \end{tikzpicture}

### 2015-06-26 article.tex
test commit
figure

--- 2015-06-26 article.tex
@@
 
-\newcommand\INCREASING[1]{%
-  \begin{tikzpicture}[>=latex]
-  \draw [->] (0,0) -- node[midway,below] {$#1$} -- ++ (1,1);
-  \end{tikzpicture}
-}
+\begin{proposition}
+Let $\pi \in S_{2n}$.
+Then,
+$\pi$ is the union of two order-isomorphic pattern 
+if and only if
+there exists a perfect containment-free mathing
+$\mathcal{M}$ of $G_\pi$ such that
+\end{proposition}
 
-\newpage
 \begin{figure}[t!]
@@
   \begin{tikzpicture}[
+    scale=.95,
     >=stealth',
     shorten >=1pt,
-    main node/.style={align=center}
-    %cell node/.style={fill=black!5}
+    main node/.style={align=center},
+    cell/.style={draw,ultra thick,fill=black!5},
+    structure link/.style={line width=1.5pt},
+    pattern link/.style={line width=1.5pt,black!20},
+    monotone/.style={->,thick}
     ]  
+    % edges
+    % pi 1 - pi 2
+    \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
+    % pi 2 - sigma 2
+    \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
+    % 1
+    \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
+    \node (claim edge 1) at (1.5,13) {Claim\ref{}};
+    % 2
+    \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
+    \node (claim edge 1) at (4,3.5) {Claim\ref{}};
+    % 3
+    \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
+    \node (claim edge 1) at (5.5,11.5) {Claim\ref{}};
+    % 4
+    \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
+    \node (claim edge 1) at (8.5,8.5) {Claim\ref{}};
+    % sigma 1 - pi 1
+    \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
+    \node (claim edge 1) at (6.5,4) {Claim\ref{}};
     % nodes
     % 1
-    \draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
-    \node [main node] (1a) at (0.5,10.5) {$\nu_1$};
-    \draw [draw,ultra thick,fill=black!5] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
-    \node [main node] (1b) at (2.5,11.5) {$\nu'_1$};
+    \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
+    \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_1$};
+    %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
+    \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
+    \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_1$};
     % 2
-    \draw [draw,ultra thick,fill=black!5] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
-    \node [main node] (2a) at (1.5,1.5) {$\nu_2$};
-    \draw [draw,ultra thick,fill=black!5] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
-    \node [main node] (2b) at (6.5,0.5) {$\nu'_2$};
+    \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
+    \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_2$};
+    \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
+    \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_2$};
     % 3
-    \draw [draw,ultra thick,fill=black!5] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
-    \node [main node] (3a) at (3.5,8.5) {$\nu_3$};
-    \draw [draw,ultra thick,fill=black!5] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
-    \node [main node] (3b) at (7.5,9.5) {$\nu'_3$};
+    \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
+    \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_3$};
+    \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
+    \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_3$};
     % 4
-    \draw [draw,ultra thick,fill=black!5] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
-    \node [main node] (3a) at (5.5,7.5) {$\nu_4$};
-    \draw [draw,ultra thick,fill=black!5] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
-    \node [main node] (3b) at (9.5,2.5) {$\nu'_4$};
+    \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
+    \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_4$};
+    \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
+    \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_4$};
     % sigma 1
-    \draw [draw,ultra thick,fill=black!5] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
+    \draw [cell] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
     \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
     % pi 1
-    \draw [draw,ultra thick,fill=black!5] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
+    \draw [cell] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
     \node [main node] (pi1) at (8.5,5.5) {$\pi'$};    
     % pi 2
-    \draw [draw,ultra thick,fill=black!5] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
+    \draw [cell] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
     \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
     % sigma 2
-    \draw [draw,ultra thick,fill=black!5] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
+    \draw [cell] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
     \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};  
-    % edges
-    % 1
-    \draw[line width=1.5pt] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
-    \node (claim edge 1) at (1.5,13) {Claim\ref{}};
-    % 2
-    \draw[line width=1.5pt] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
-    \node (claim edge 1) at (4,3.5) {Claim\ref{}};
-    % 3
-    \draw[line width=1.5pt] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
-    \node (claim edge 1) at (5.5,11.5) {Claim\ref{}};
-    % 4
-    \draw[line width=1.5pt] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
-    \node (claim edge 1) at (8.5,8.5) {Claim\ref{}};
-    % sigma 1 - pi 1
-    \draw[line width=1.5pt] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
-    \node (claim edge 1) at (6.5,4) {Claim\ref{}};
-    % pi 1 - pi 2
-    \draw[line width=1.5pt,black!20] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
-    % pi 2 - sigma 2
-    \draw[line width=1.5pt,black!20] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
     % grid
-    \draw[step=1cm,gray,ultra thin] (0,0) grid (12,12);
+    \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
+    % row size
+    \foreach \y/\N in {0.5/N_2,1.5/N_2}
+        \node at (-0.5,\y) (R\y) {$\N$};
+    % }
+    % column size
+    \foreach \x/\N in 
+    {0.5/N_1,1.5/N_2,2.5/N_1,3.5/N_3,4.5/k+2,5.5/N_4,6.5/N_2,7.5/N_3,8.5/n+2,9.5/N_4,10.5/n,11.5/k} {
+        \node at (\x,-0.5) (C\x) {$\N$};
+    }
   \end{tikzpicture}

### 2015-06-27 article.tex
reduction

@@
 
+\DeclareMathOperator{\LEFT}{\ell}
+\DeclareMathOperator{\RIGHT}{r}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   F-77454, Marne-la-Vallée, France\\
+  \email{samuele.giraudo@univ-mlv.fr} \\
   \email{vialette@univ-mlv.fr}
@@
 $\mathcal{M}$ of $G_\pi$ such that
+$\LEFT(\mathcal{M}) \simeq \RIGHT(\mathcal{M})$.
 \end{proposition}
+
+\begin{proposition}
+Deciding whether a permutation is the union of two
+order-isomorphic pattern is \NPC.
+\end{proposition}
+
+\begin{align*}
+N_4 &= 2(2n + k + 2) + 1  = 4n + 2k + 5 \\
+N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 12k + 9 \\
+N_2 &= 2(2N_3 + 2n + 2k + 4) + 1 \\
+N_1 &= 2(2N_2 + 2n + 2k + 4) + 1\text{.}
+\end{align*}
+
+\begin{align*}
+\sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
+\pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2] \\
+\sigma'' &= \sigma \; [2N_2 + N_4] \\
+\pi''    &= \pi \; [2N_2 + N_4 + k] \\
+\nu_1    &= \nearrow_{N_1} \; [2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
+\nu'_1   &= \nearrow_{N_1} \; [N_1 + 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
+\nu_2    &= \nearrow_{N_2} \; [N_2] \\
+\nu'_2   &= \searrow_{N_2} \\
+\nu_3    &= \nearrow_{N_3} \; [2N_2 + 2N_4 + 2n + 2k + 4] \\
+\nu'_3   &= \nearrow_{N_3} \; [2N_2 + N_3 + 2N_4 + 2n + 2k + 4] \\
+\nu_4    &= \searrow_{N_4} \; [2N_2 + N_4 + 2n + 2k + 4] \\
+\nu'_4   &= \searrow_{N_4} \; [2N_2]
+\end{align*}
+
+$$
+\mu
+=
+\nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
+$$
+
+\begin{claim}
+  \label{claim:no nu_1 - nu_1 edge}
+  There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
+\end{claim}
+
+\begin{proof}[of Claim~\ref{claim:no nu_1 - nu_1 edge}]
+Suppose, aiming at a contradiction that there exists
+a $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
+\begin{tikzpicture}
+
+\end{tikzpicture}
+\qed
+\end{proof}
+
+\begin{claim}
+There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+\end{claim}
 

### 2015-06-29 article.tex
reduction
Partie algebre (mise en place)

@@
 \DeclareMathOperator{\RIGHT}{r}
+\DeclareMathOperator{\QQ}{\mathbb{Q}}
+\DeclareMathOperator{\STD}{\mathrm{std}}
+\DeclareMathOperator{\SHUFFLE}{\bullet}
 
@@
 \label{section:Notations}
+
+% Définir :
+% - ensemble $S$
+% - standardisation $\STD$
+% - notation $\pi_{|P}$ et $\pi_{|\bar P}$
+% - notation $\pi^\sim$ (miroir)
+% - notation $\bar \pi$ (complémentaire)
+% - notation $\bar ^{-1}$ (inverse)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% ---- Algebraic issues ----
+%%
+\section{Algebraic issues}
+\label{section:Algebraic issues}
+Let use denote by $\QQ[S]$ the linear span of all permutations. We shall
+define on this space a kind of shuffle product. For this purpose, we
+first define a linear coproduct $\Delta$ on $\QQ[S]$ satisfying for
+any permutation $\pi$,
+\begin{equation} \label{equ:unshuffling_coproduct}
+    \Delta(\pi) =
+    \sum_{P \subseteq [|\pi|]}
+    \STD\left(\pi_{|P}\right) \otimes \STD\left(\pi_{|\bar P}\right).
+\end{equation}
+We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
+instance,
+\begin{equation}
+    \Delta(213) =
+    \epsilon \otimes 213 + 2. 1 \otimes 12 +
+    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
+\end{equation}
+\begin{equation}
+    \Delta(1234) =
+    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
+    4. 123 \otimes 1 + 1234 \otimes \epsilon,
+\end{equation}
+\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
+    \Delta(1432) & =
+    \epsilon \otimes 1432 + 3. 1 \otimes 132 + 1 \otimes 321 +
+    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
+    321 \otimes 1 + 1432 \otimes \epsilon.
+\end{split}\end{equation}
+\medskip
+
+As any coproduct, $\Delta$ leads to the definition of a product obtained
+by duality in the following way. From~\eqref{equ:unshuffling_coproduct},
+for any permutation $\pi$, we have
+\begin{equation}
+    \Delta(\pi) =
+    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^{(\pi)} \;
+    \sigma \otimes \nu,
+\end{equation}
+where the $\lambda_{\sigma, \nu}^{(\pi)}$ are nonnegative integers. Now,
+by definition of duality, the dual product of $\Delta$ denoted by
+$\SHUFFLE$ is a linear binary product on $\QQ[S]$. It satisfies, for any
+permutations
+$\sigma$ and $\nu$,
+\begin{equation}
+    \sigma \SHUFFLE \nu =
+    \sum_{\pi \in S}
+    \lambda_{\sigma, \nu}^{(\pi)} \; \pi.
+\end{equation}
+We call $\SHUFFLE$ the {\em shuffle product of permutations}. For
+instance,
+\begin{equation}\begin{split} \label{equ:example_shuffle_product}
+    12 \SHUFFLE 21 & =
+    1243 + 1324 + 2.1342 + 2.1423 + 3.1432 + 2134 + 2.2314 \\
+    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
+    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
+\end{split}\end{equation}
+Observe that the coefficient $3$ of the permutation $1432$
+in~\eqref{equ:example_shuffle_product} comes from the fact that the
+coefficient of the tensor $12 \otimes 21$ is $3$
+in~\eqref{equ:example_unshuffling_coproduct}.
+\medskip
+
+We say that a permutation $\pi$ {\em appears} in the shuffle
+$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
+coefficient $\lambda_{\sigma, \nu}^{(\pi)}$ is different from zero.
+Besides, a permutation $\pi$ is a {\em square} if there is a permutation
+$\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
+Equivalently, $\pi$ is a square if in the expansion of $\Delta(\pi)$,
+there is a tensor $\sigma \otimes \sigma$ with a nonzero coefficient.
+\medskip
+
+\begin{proposition}
+    The shuffle product $\SHUFFLE$ of permutations is associative and
+    commutative.
+\end{proposition}
+\begin{proof}
+
+\end{proof}
+\medskip
+
+\begin{proposition}
+    If $\pi$ is a square permutation, then $\pi^\sim$, $\bar \pi$,
+    and $\pi^{-1}$ are square.
+\end{proposition}
+\begin{proof}
+
+\end{proof}
+\medskip
 
@@
 a $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
+
 \begin{tikzpicture}
-
+  [scale=0.25]
+  % nu_1
+  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
+  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
+      \draw [fill=black] (\x,\y) circle (0.1);
+  }
+  % nu_4
+  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
+  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
+      \draw [fill=black] (\x,\y) circle (0.1);
+  }
+  % nu'_1
+  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
+  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
+      \draw [fill=black] (\x,\y) circle (0.1);
+  }
+  % edges
+  \draw [line width=1pt]
+  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
+  \draw [line width=1pt]
+  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
+\end{tikzpicture}
+\begin{tikzpicture}
+  [scale=0.25]
+  % nu_1
+  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
+  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
+      \draw [fill=black] (\x,\y) circle (0.1);
+  }
+  % nu_4
+  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
+  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
+      \draw [fill=black] (\x,\y) circle (0.1);
+  }
+  % nu'_1
+  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
+  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
+      \draw [fill=black] (\x,\y) circle (0.1);
+  }
+  % edges
+  \draw [line width=1pt]
+  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
+  \draw [line width=1pt]
+  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
 \end{tikzpicture}
@@
     % row size
-    \foreach \y/\N in {0.5/N_2,1.5/N_2}
+    \foreach \y/\N in {0.5/N_2,1.5/N_2} {
         \node at (-0.5,\y) (R\y) {$\N$};
-    % }
+    }
     % column size

### 2015-06-29 article.tex
Partie algebre (suite)

@@
 % Définir :
+% - notation $[n]$
 % - ensemble $S$
@@
     \Delta(\pi) =
-    \sum_{P \subseteq [|\pi|]}
-    \STD\left(\pi_{|P}\right) \otimes \STD\left(\pi_{|\bar P}\right).
+    \sum_{P_1 \uplus P_2 = [|\pi|]}
+    \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
 \end{equation}
@@
 
-\begin{proposition}
+\begin{proposition} \label{prop:shuffle_associative_commutative}
     The shuffle product $\SHUFFLE$ of permutations is associative and
@@
 \begin{proof}
+    To prove the associativity of $\SHUFFLE$, it is convenient to show
+    that its dual coproduct $\Delta$ is coassociative, that is
+    \begin{equation}
+        (\Delta \otimes I) \Delta = (I \otimes \Delta) \Delta,
+    \end{equation}
+    where $I$ denote the identity map. This strategy relies on the fact
+    that a product is associative if and only if its dual coproduct is
+    coassociative. For any permutation $\pi$, we have
+    \begin{equation} \begin{split}
+    \label{equ:shuffle_associative_commutative}
+        (\Delta \otimes I) \Delta(\pi) & =
+        (\Delta \otimes I)
+        \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
+        & =
+        \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes I\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & =
+        \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \sum_{Q_1 \uplus Q_2 = [|P_1|]}
+        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
+        \otimes
+        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_2}\right)
+        \otimes \STD\left(\pi_{|P_2}\right) \\
+        & =
+        \sum_{P_1 \uplus P_2 \uplus P_3 = [|\pi|]}
+        \STD\left(\pi_{|P_1}\right) \otimes
+        \STD\left(\pi_{|P_2}\right) \otimes
+        \STD\left(\pi_{|P_3}\right).
+    \end{split} \end{equation}
+    An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
+    is equal to the last member of~\eqref{equ:shuffle_associative_commutative},
+    whence the associativity of $\SHUFFLE$.
+    \smallskip
 
+    Finally, to prove the commutativity of $\SHUFFLE$, we shall show
+    that $\Delta$ is cocommutative, that is for any permutation $\pi$,
+    if in the expansion of $\Delta(\pi)$ there is a tensor
+    $\sigma \otimes \nu$ with a coefficient $\lambda$, there is in the
+    same expansion the tensor $\nu \otimes \sigma$ with the same
+    coefficient $\lambda$. Clearly, a product is associative if and only
+    if its dual coproduct is coassociative. Now, from the
+    definition~\eqref{equ:unshuffling_coproduct} of $\Delta$, one
+    observes that if the pair $(P_1, P_2)$ of subsets of $[|\pi|]$
+    contributes to the coefficient of
+    $\STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right)$,
+    the pair $(P_2, P_1)$ contributes to the coefficient of
+    $\STD\left(\pi_{|P_2}\right) \otimes \STD\left(\pi_{|P_1}\right)$.
+    This shows that $\Delta$ is cocommutative and hence, that $\SHUFFLE$
+    is commutative.
 \end{proof}
@@
 \begin{proposition}
-    If $\pi$ is a square permutation, then $\pi^\sim$, $\bar \pi$,
-    and $\pi^{-1}$ are square.
+    Let $\pi$ be a square permutation. Then, $\pi^\sim$, $\bar \pi$,
+    and $\pi^{-1}$ are square permutations.
 \end{proposition}

### 2015-06-29 article.tex
Partie algebre (suite)

@@
     \Delta(\pi) =
-    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^{(\pi)} \;
+    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
     \sigma \otimes \nu,
 \end{equation}
-where the $\lambda_{\sigma, \nu}^{(\pi)}$ are nonnegative integers. Now,
+where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
 by definition of duality, the dual product of $\Delta$ denoted by
@@
     \sum_{\pi \in S}
-    \lambda_{\sigma, \nu}^{(\pi)} \; \pi.
+    \lambda_{\sigma, \nu}^\pi \; \pi.
 \end{equation}
@@
 $\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
-coefficient $\lambda_{\sigma, \nu}^{(\pi)}$ is different from zero.
+coefficient $\lambda_{\sigma, \nu}^\pi$ is different from zero.
 Besides, a permutation $\pi$ is a {\em square} if there is a permutation
-$\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
-Equivalently, $\pi$ is a square if in the expansion of $\Delta(\pi)$,
-there is a tensor $\sigma \otimes \sigma$ with a nonzero coefficient.
+$\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$. In this
+case, we say that $\sigma$ is a {\em square root} of $\pi$. Equivalently,
+$\pi$ is a square if in the expansion of $\Delta(\pi)$, there is a
+tensor $\sigma \otimes \sigma$ with a nonzero coefficient.
 \medskip
@@
 \begin{proposition}
-    Let $\pi$ be a square permutation. Then, $\pi^\sim$, $\bar \pi$,
-    and $\pi^{-1}$ are square permutations.
+    Let $\pi$ be a square permutation and $\rho$ a square root of $\pi$.
+    Then,
+    \begin{enumerate}
+        \item the permutation $\pi^\sim$ is a square and $\rho^\sim$
+        is one of its square roots;
+        \item the permutation $\bar \pi$ is a square and $\bar \rho$
+        is one of its square roots;
+        \item the permutation $\pi^{-1}$ is a square and $\rho^{-1}$
+        is one of its square roots.
+    \end{enumerate}
 \end{proposition}

### 2015-06-29 article.tex
Partie algebre (suite)

@@
 \usepackage{subfigure}
+\usepackage{enumitem}
 
@@
 
-\begin{proposition}
+\begin{proposition} \label{prop:square_stability}
     Let $\pi$ be a square permutation and $\rho$ a square root of $\pi$.
     Then,
-    \begin{enumerate}
-        \item the permutation $\pi^\sim$ is a square and $\rho^\sim$
+    \begin{enumerate}[label={\it (\roman*)},fullwidth]
+        \item \label{item:square_stability_1}
+        the permutation $\pi^\sim$ is a square and $\rho^\sim$
         is one of its square roots;
-        \item the permutation $\bar \pi$ is a square and $\bar \rho$
+        \item \label{item:square_stability_2}
+        the permutation $\bar \pi$ is a square and $\bar \rho$
         is one of its square roots;
-        \item the permutation $\pi^{-1}$ is a square and $\rho^{-1}$
+        \item \label{item:square_stability_3}
+        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$
         is one of its square roots.
@@
 \begin{proof}
+    First of all, since $\pi$ is a square permutation and $\rho$ is a
+    square root of $\pi$, $\pi$ appears in $\rho \bullet \rho$. In
+    an equivalent way, this amounts to say that $\rho \otimes \rho$ has
+    a nonzero coefficient in the expansion of $\Delta(\pi)$. This
+    implies that there are two disjoint sets of indexes $P_1$ and $P_2$
+    such that $P_1 \uplus P_2 = [|\pi|]$ and
+    \begin{equation} \label{equ:square_stability_same_std}
+        \STD\left(\pi_{|P_1}\right) = \rho = \STD\left(\pi_{|P_2}\right).
+    \end{equation}
+    \smallskip
+
+    To prove~\ref{item:square_stability_1}, set
+    $P_1^\sim := \{|\pi| - i  + 1 : i \in P_1\}$ and
+    $P_2^\sim := \{|\pi| - i  + 1 : i \in P_2\}$.
+    Since~\eqref{equ:square_stability_same_std} holds, one has
+    \begin{equation}
+        \STD\left(\pi^\sim_{|P_1^\sim}\right) = \rho^\sim =
+        \STD\left(\pi^\sim_{|P_2^\sim}\right).
+    \end{equation}
+
+    TO CONTINUE
 

### 2015-06-29 article.tex
Partie algebre (suite)

@@
 case, we say that $\sigma$ is a {\em square root} of $\pi$. Equivalently,
-$\pi$ is a square if in the expansion of $\Delta(\pi)$, there is a
-tensor $\sigma \otimes \sigma$ with a nonzero coefficient.
+$\pi$ is a square with $\sigma$ as square root if and only if in the
+expansion of $\Delta(\pi)$, there is a tensor $\sigma \otimes \sigma$
+with a nonzero coefficient.
 \medskip
@@
     First of all, since $\pi$ is a square permutation and $\rho$ is a
-    square root of $\pi$, $\pi$ appears in $\rho \bullet \rho$. In
-    an equivalent way, this amounts to say that $\rho \otimes \rho$ has
-    a nonzero coefficient in the expansion of $\Delta(\pi)$. This
-    implies that there are two disjoint sets of indexes $P_1$ and $P_2$
-    such that $P_1 \uplus P_2 = [|\pi|]$ and
+    square root of $\pi$, $\rho \otimes \rho$ has a nonzero coefficient
+    in the expansion of $\Delta(\pi)$. This implies that there are two
+    disjoint sets of indexes $P_1$ and $P_2$ such that
+    $P_1 \uplus P_2 = [|\pi|]$ and, by denoting $\pi_{|P_1}$ by $u$ and
+    $\pi_{|P_2}$ by $v$,
     \begin{equation} \label{equ:square_stability_same_std}
-        \STD\left(\pi_{|P_1}\right) = \rho = \STD\left(\pi_{|P_2}\right).
+        \STD(u) = \rho = \STD(v).
     \end{equation}
@@
     To prove~\ref{item:square_stability_1}, set
-    $P_1^\sim := \{|\pi| - i  + 1 : i \in P_1\}$ and
-    $P_2^\sim := \{|\pi| - i  + 1 : i \in P_2\}$.
-    Since~\eqref{equ:square_stability_same_std} holds, one has
-    \begin{equation}
-        \STD\left(\pi^\sim_{|P_1^\sim}\right) = \rho^\sim =
-        \STD\left(\pi^\sim_{|P_2^\sim}\right).
-    \end{equation}
+    $P_1^\sim := \{|\pi| - i  + 1 : i \in P_1\}$,
+    $P_2^\sim := \{|\pi| - i  + 1 : i \in P_2\}$,
+    $u' := \pi^\sim_{|P_1^\sim}$, and $v' := \pi^\sim_{|P_2^\sim}$.
+    Now, by construction, ${u'}^\sim = u$ and ${v'}^\sim = v$. Then,
+    by~\eqref{equ:square_stability_same_std},
+    $\STD(u') = \rho^\sim = \STD(v')$, so that $\rho^\sim$ is a square
+    root of $\pi$.
+    \smallskip
+
+    Let us establish~\ref{item:square_stability_2}. Set
+    $u' := \bar \pi_{|P_1}$ and $v' := \bar \pi_{|P_2}$. By definition
+    of the operation $\bar\,$, for all indexes $i$ and $j$, we have
+    $u_i < u_j$ if and only if $\bar u'_i > \bar u'_j$, the same
+    property holding for $v$ and $v'$. Then,
+    by~\eqref{equ:square_stability_same_std},
+    $\STD(u') = \bar \rho = \STD(v')$, showing that $\bar \rho$ is a
+    square root of $\bar \pi$.
+    \smallskip
+
+    Finally, let us show~\ref{item:square_stability_3}. Set
+    $P^{-1}_1 := \{u_i : i \in [|u|]\}$,
+    $P^{-1}_2 := \{v_i : i \in [|v|]\}$, $u' := \pi^{-1}_{|P^{-1}_1}$,
+    and $v' := \pi^{-1}_{|P^{-1}_2}$.
 

### 2015-06-30 article.tex
Partie algebre (suite)

@@
 \DeclareMathOperator{\QQ}{\mathbb{Q}}
-\DeclareMathOperator{\STD}{\mathrm{std}}
+\DeclareMathOperator{\STD}{\mathrm{s}}
 \DeclareMathOperator{\SHUFFLE}{\bullet}
@@
 % - notation $\pi_{|P}$ et $\pi_{|\bar P}$
-% - notation $\pi^\sim$ (miroir)
+% - notation $\widetitle{\pi} (miroir)
 % - notation $\bar \pi$ (complémentaire)
@@
 
-\begin{proposition} \label{prop:square_stability}
-    Let $\pi$ be a square permutation and $\rho$ a square root of $\pi$.
-    Then,
-    \begin{enumerate}[label={\it (\roman*)},fullwidth]
-        \item \label{item:square_stability_1}
-        the permutation $\pi^\sim$ is a square and $\rho^\sim$
-        is one of its square roots;
-        \item \label{item:square_stability_2}
-        the permutation $\bar \pi$ is a square and $\bar \rho$
-        is one of its square roots;
-        \item \label{item:square_stability_3}
-        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$
-        is one of its square roots.
-    \end{enumerate}
+Proposition~\ref{prop:shuffle_associative_commutative} implies that
+$\QQ[S]$ endowed with the unshuffling coproduct $\Delta$ is a
+coassociative cocommutative coalgebra, or in an equivalent way, that
+$\QQ[S]$ endowed with the shuffle product $\SHUFFLE$ is an associative
+commutative algebra.
+\medskip
+
+\begin{proposition} \label{prop:endomorphisms}
+    The three linear maps
+    \begin{equation}
+        \phi_{\widetilde{\,}}, \: \phi_{\bar \,}, \: \phi_{-1}
+        : \QQ[S] \to \QQ[S]
+    \end{equation}
+    linearly sending a permutation $\pi$ to, respectively,
+    $\widetilde{\pi}$, $\bar \pi$, or $\pi^{-1}$ are endomorphims of associative
+    algebras.
 \end{proposition}
 \begin{proof}
+    ADAPT NEXT ARGUMENTS AND FINISH.
+
     First of all, since $\pi$ is a square permutation and $\rho$ is a
@@
     To prove~\ref{item:square_stability_1}, set
-    $P_1^\sim := \{|\pi| - i  + 1 : i \in P_1\}$,
-    $P_2^\sim := \{|\pi| - i  + 1 : i \in P_2\}$,
-    $u' := \pi^\sim_{|P_1^\sim}$, and $v' := \pi^\sim_{|P_2^\sim}$.
-    Now, by construction, ${u'}^\sim = u$ and ${v'}^\sim = v$. Then,
-    by~\eqref{equ:square_stability_same_std},
-    $\STD(u') = \rho^\sim = \STD(v')$, so that $\rho^\sim$ is a square
-    root of $\pi$.
+    $\widetilde{P_1} := \{|\pi| - i  + 1 : i \in P_1\}$,
+    $\widetilde{P_2} := \{|\pi| - i  + 1 : i \in P_2\}$,
+    $u' := \widetilde{\pi}_{|\widetilde{P_1}}$, and
+    $v' := \widetilde{\pi}_{|\widetilde{P_2}}$.
+    Now, by construction, $\widetilde{u'} = u$ and $\widetilde{v'} = v$.
+    Then, by~\eqref{equ:square_stability_same_std},
+    $\STD(u') = \widetilde{\rho} = \STD(v')$, so that $\widetilde{\rho}$
+    is a square root of $\pi$.
     \smallskip
@@
     TO CONTINUE
+\end{proof}
+\medskip
 
+\begin{proposition} \label{prop:square_stability}
+    Let $\pi$ be a square permutation and $\rho$ be a square root of
+    $\pi$. Then,
+    \begin{enumerate}[label={\it (\roman*)},fullwidth]
+        \item \label{item:square_stability_1}
+        the permutation $\widetilde{\pi}$ is a square and
+        $\widetilde{\rho}$ is one of its square roots;
+        \item \label{item:square_stability_2}
+        the permutation $\bar \pi$ is a square and $\bar \rho$
+        is one of its square roots;
+        \item \label{item:square_stability_3}
+        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$
+        is one of its square roots.
+    \end{enumerate}
+\end{proposition}
+\begin{proof}
+    Since $\pi$ is a square permutation and $\rho$ is a square root
+    of $\pi$, by definition $\pi$ appears in the product
+    $\rho \SHUFFLE \rho$. Now, by Proposition~\ref{prop:endomorphisms},
+    since $\phi_{\widetilde{\,}}$ is a morphism of associative algebras
+    from $\QQ[S]$ to $\QQ[S]$, $\phi_{\widetilde{\,}}$ commutes with
+    the shuffle product of permutations $\SHUFFLE$. Hence, in particular,
+    one has
+    \begin{equation}
+        \phi_{\widetilde{\,}}(\rho \SHUFFLE \rho) =
+        \phi_{\widetilde{\,}}(\rho) \SHUFFLE \phi_{\widetilde{\,}}(\rho).
+    \end{equation}
+    Then, since $\pi$ appears in $\rho \SHUFFLE \rho$,
+    $\widetilde{\pi}$ appears in $\widetilde{\rho} \SHUFFLE \widetilde{\rho}$.
+    This implies~\ref{item:square_stability_1}.
+    \smallskip
+
+    The proofs of~\ref{item:square_stability_2}
+    and~\ref{item:square_stability_3} are analogous.
 \end{proof}

### 2015-07-01 article.tex
Partie algebre (suite ; en plein dans les modifs)

@@
     \begin{equation}
-        \phi_{\widetilde{\,}}, \: \phi_{\bar \,}, \: \phi_{-1}
-        : \QQ[S] \to \QQ[S]
+        \phi_1, \phi_2, \phi_3 : \QQ[S] \to \QQ[S]
     \end{equation}
     linearly sending a permutation $\pi$ to, respectively,
-    $\widetilde{\pi}$, $\bar \pi$, or $\pi^{-1}$ are endomorphims of associative
-    algebras.
+    $\widetilde{\pi}$, $\bar \pi$, or $\pi^{-1}$ are endomorphims of
+    associative algebras.
 \end{proposition}
 \begin{proof}
+    To prove that, for $j = 1, 2, 3$, $\phi_j$ is a morphism of
+    associative algebras, we have to prove that for all permutations
+    $\sigma$ and $\nu$,
+    \begin{equation}
+        \phi_j(\sigma \SHUFFLE \nu) =
+        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
+    \end{equation}
+    We shall proceed here as we have done during the proof of
+    Proposition~\ref{prop:shuffle_associative_commutative} by showing
+    that $\phi_j$ is a morphism of coalgebras, that is
+    \begin{equation}
+        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
+    \end{equation}
+    \smallskip
+
+    Let $\pi$ be a permutation. If $P$ is a set of indexes of letters of
+    $\pi$, we denote by $\widetilde{\pi}$ the set
+    $\{|\pi| - i  + 1 : i \in P_1\}$. Now since this operation
+    $\widetilde{\,}$ define a bijection on the set of the subsets of
+    $[|\pi|]$ and the standardisation operation commutes with the
+    mirror operation on words, we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_1(\pi))
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]} \STD\left(\phi_1(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \uplus P_2 =
+        [|\pi|]} \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
+        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
+        & = \sum_{P_1 \uplus P_2 =
+        [|\pi|]} \STD\left(\widetilde{\pi}_{|P_1}\right)
+        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
+        & = \sum_{P_1 \uplus P_2 =
+        [|\pi|]} \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
+    $\phi_1$ is a morphism of associative algebras.
+    \smallskip
+
     ADAPT NEXT ARGUMENTS AND FINISH.
@@
         \item \label{item:square_stability_2}
-        the permutation $\bar \pi$ is a square and $\bar \rho$
-        is one of its square roots;
+        the permutation $\bar \pi$ is a square and $\bar \rho$ is one of
+        its square roots;
         \item \label{item:square_stability_3}
-        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$
-        is one of its square roots.
+        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$ is one of
+        its square roots.
     \end{enumerate}
@@
 \begin{proof}
-    Since $\pi$ is a square permutation and $\rho$ is a square root
-    of $\pi$, by definition $\pi$ appears in the product
+    All statements~\ref{item:square_stability_1},
+    \ref{item:square_stability_2}, and~\ref{item:square_stability_3} are
+    direct consequences of Proposition~\ref{prop:endomorphisms}. Indeed,
+    since $\pi$ is a square permutation and $\rho$ is a square root of
+    $\pi$, by definition $\pi$ appears in the product
     $\rho \SHUFFLE \rho$. Now, by Proposition~\ref{prop:endomorphisms},
-    since $\phi_{\widetilde{\,}}$ is a morphism of associative algebras
-    from $\QQ[S]$ to $\QQ[S]$, $\phi_{\widetilde{\,}}$ commutes with
-    the shuffle product of permutations $\SHUFFLE$. Hence, in particular,
+    for all $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
+    algebras from $\QQ[S]$ to $\QQ[S]$, $\phi_j$ commutes with the
+    shuffle product of permutations $\SHUFFLE$. Hence, in particular,
     one has
     \begin{equation}
-        \phi_{\widetilde{\,}}(\rho \SHUFFLE \rho) =
-        \phi_{\widetilde{\,}}(\rho) \SHUFFLE \phi_{\widetilde{\,}}(\rho).
+        \phi_j(\rho \SHUFFLE \rho) =
+        \phi_j(\rho) \SHUFFLE \phi_j(\rho).
     \end{equation}
-    Then, since $\pi$ appears in $\rho \SHUFFLE \rho$,
-    $\widetilde{\pi}$ appears in $\widetilde{\rho} \SHUFFLE \widetilde{\rho}$.
-    This implies~\ref{item:square_stability_1}.
-    \smallskip
-
-    The proofs of~\ref{item:square_stability_2}
-    and~\ref{item:square_stability_3} are analogous.
+    Then, since $\pi$ appears in $\rho \SHUFFLE \rho$, $\phi_j(\pi)$
+    appears in $\phi_j(\rho) \SHUFFLE \phi_j(\rho)$. This
+    implies~\ref{item:square_stability_1}, \ref{item:square_stability_2},
+    and~\ref{item:square_stability_3}.
 \end{proof}

### 2015-07-01 article.tex
Partie algebre (suite)

@@
     linearly sending a permutation $\pi$ to, respectively,
-    $\widetilde{\pi}$, $\bar \pi$, or $\pi^{-1}$ are endomorphims of
+    $\widetilde{\pi}$, $\bar \pi$, and $\pi^{-1}$ are endomorphims of
     associative algebras.
@@
     \end{equation}
+    In the sequel, $\pi$ is a permutation.
     \smallskip
 
-    Let $\pi$ be a permutation. If $P$ is a set of indexes of letters of
-    $\pi$, we denote by $\widetilde{\pi}$ the set
-    $\{|\pi| - i  + 1 : i \in P_1\}$. Now since this operation
-    $\widetilde{\,}$ define a bijection on the set of the subsets of
-    $[|\pi|]$ and the standardisation operation commutes with the
-    mirror operation on words, we have
+    If $P$ is a set of indexes of letters of $\pi$, we denote by
+    $\widetilde{\pi}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now since
+    this operation $\widetilde{\,}$ define a bijection on the set of the
+    subsets of $[|\pi|]$ and the standardisation operation commutes with
+    the mirror operation on words, we have
     \begin{equation} \begin{split}
         \Delta(\phi_1(\pi))
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]} \STD\left(\phi_1(\pi)_{|P_1}\right)
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\phi_1(\pi)_{|P_1}\right)
         \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 =
-        [|\pi|]} \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\widetilde{\pi}_{|P_1}\right)
+        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
         \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
-        & = \sum_{P_1 \uplus P_2 =
-        [|\pi|]} \STD\left(\widetilde{\pi}_{|P_1}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 =
-        [|\pi|]} \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \widetilde{\STD\left(\pi_{|P_1}\right)}
+        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
         \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
@@
 
-    ADAPT NEXT ARGUMENTS AND FINISH.
-
-    First of all, since $\pi$ is a square permutation and $\rho$ is a
-    square root of $\pi$, $\rho \otimes \rho$ has a nonzero coefficient
-    in the expansion of $\Delta(\pi)$. This implies that there are two
-    disjoint sets of indexes $P_1$ and $P_2$ such that
-    $P_1 \uplus P_2 = [|\pi|]$ and, by denoting $\pi_{|P_1}$ by $u$ and
-    $\pi_{|P_2}$ by $v$,
-    \begin{equation} \label{equ:square_stability_same_std}
-        \STD(u) = \rho = \STD(v).
-    \end{equation}
+    Next, since by definition of the complementation operation on
+    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
+    we have $\tau_i < \tau_j$ if and only if $\bar \tau_i > \bar \tau_j$,
+    we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_2(\pi))
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\phi_2(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\bar \pi_{|P_1}\right)
+        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
+    $\phi_2$ is a morphism of associative algebras.
     \smallskip
 
-    To prove~\ref{item:square_stability_1}, set
-    $\widetilde{P_1} := \{|\pi| - i  + 1 : i \in P_1\}$,
-    $\widetilde{P_2} := \{|\pi| - i  + 1 : i \in P_2\}$,
-    $u' := \widetilde{\pi}_{|\widetilde{P_1}}$, and
-    $v' := \widetilde{\pi}_{|\widetilde{P_2}}$.
-    Now, by construction, $\widetilde{u'} = u$ and $\widetilde{v'} = v$.
-    Then, by~\eqref{equ:square_stability_same_std},
-    $\STD(u') = \widetilde{\rho} = \STD(v')$, so that $\widetilde{\rho}$
-    is a square root of $\pi$.
-    \smallskip
-
-    Let us establish~\ref{item:square_stability_2}. Set
-    $u' := \bar \pi_{|P_1}$ and $v' := \bar \pi_{|P_2}$. By definition
-    of the operation $\bar\,$, for all indexes $i$ and $j$, we have
-    $u_i < u_j$ if and only if $\bar u'_i > \bar u'_j$, the same
-    property holding for $v$ and $v'$. Then,
-    by~\eqref{equ:square_stability_same_std},
-    $\STD(u') = \bar \rho = \STD(v')$, showing that $\bar \rho$ is a
-    square root of $\bar \pi$.
-    \smallskip
-
-    Finally, let us show~\ref{item:square_stability_3}. Set
-    $P^{-1}_1 := \{u_i : i \in [|u|]\}$,
-    $P^{-1}_2 := \{v_i : i \in [|v|]\}$, $u' := \pi^{-1}_{|P^{-1}_1}$,
-    and $v' := \pi^{-1}_{|P^{-1}_2}$.
-
-    TO CONTINUE
+    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
+    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
+    $\{\tau_i : i \in P\}$. Since the map sending a subset $P$ of
+    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection and
+    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
+    we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_3(\pi))
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\phi_3(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\pi^{-1}_{|P_1}\right)
+        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
+        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \STD\left(\pi_{|P_1}\right)^{-1}
+        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
+        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
+    $\phi_3$ is a morphism of associative algebras.
 \end{proof}

### 2015-07-01 article.tex
Quelques notations

@@
 
-% Définir :
-% - notation $[n]$
-% - ensemble $S$
-% - standardisation $\STD$
-% - notation $\pi_{|P}$ et $\pi_{|\bar P}$
-% - notation $\widetitle{\pi} (miroir)
-% - notation $\bar \pi$ (complémentaire)
-% - notation $\bar ^{-1}$ (inverse)
+We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
+set of all permutations. If $\pi$ is a permutation, we denote by $|\pi|$
+its size. If $P$ is a subset of $[|\pi|]$, $\pi_{|P}$ is the subword of
+$\pi$ consisting in the letters of $\pi$ at the positions specified by
+the elements of $P$. Besides, we denote by $\widetilde{\pi}$ (resp.
+$\bar \pi$, $\pi^{-1}$) the mirror image (resp. complementary, inverse)
+of $\pi$. If $u$ is a word of integers without multiple occurrences of a
+same letter, $\STD(u)$ denotes the unique permutation of same size as
+$u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only if
+$\STD(u)_i < \STD(u)_j$.
+\medskip
 

### 2015-07-02 article.tex
Partie algebre (modifications)

@@
 \end{split}\end{equation}
+Observe that the coefficient of the tensor $1 \otimes 132$ is $3$
+in~\eqref{equ:example_unshuffling_coproduct} because there are exactly
+three ways to extract two disjoint subwords of the permutation $1432$
+so that the letters of these two subwords have the same relative order
+as the letters of the permutations $1$ and $132$.
 \medskip
@@
 \end{equation}
-We call $\SHUFFLE$ the {\em shuffle product of permutations}. For
-instance,
+We call $\SHUFFLE$ the {\em shuffle product of permutations}. For instance,
 \begin{equation}\begin{split} \label{equ:example_shuffle_product}
@@
 in~\eqref{equ:example_unshuffling_coproduct}.
+\medskip
+
+Intuitively, this product shuffles the values and the positions of the
+letters of the permutations. There are other shuffle products on
+permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
+{\em convolution product}. Our shuffle product is different from these
+last two.
 \medskip

### 2015-07-02 article.tex
Ajouts divers

@@
 set of all permutations. If $\pi$ is a permutation, we denote by $|\pi|$
-its size. If $P$ is a subset of $[|\pi|]$, $\pi_{|P}$ is the subword of
-$\pi$ consisting in the letters of $\pi$ at the positions specified by
-the elements of $P$. Besides, we denote by $\widetilde{\pi}$ (resp.
+its size. The unique permutation of size $0$ is denoted by $\epsilon$.
+If $P$ is a subset of $[|\pi|]$, $\pi_{|P}$ is the subword of $\pi$
+consisting in the letters of $\pi$ at the positions specified by the
+elements of $P$. Besides, we denote by $\widetilde{\pi}$ (resp.
 $\bar \pi$, $\pi^{-1}$) the mirror image (resp. complementary, inverse)
@@
 \label{section:Algebraic issues}
-Let use denote by $\QQ[S]$ the linear span of all permutations. We shall
-define on this space a kind of shuffle product. For this purpose, we
-first define a linear coproduct $\Delta$ on $\QQ[S]$ satisfying for
-any permutation $\pi$,
+The purpose of this section is to define a shuffle product $\SHUFFLE$
+on permutations and establish some of its properties. To define this
+shuffle product, we shall first define a coproduct~\cite{Joni:Rota:1979}
+denoted by $\Delta$. By duality, $\Delta$ implies the definition of
+$\SHUFFLE$. It is worth to note that, as we will see, algebraic
+properties of $\Delta$ imply combinatorial properties of $\SHUFFLE$.
+\medskip
+
+Let use denote by $\QQ[S]$ the linear span of all permutations. As
+announced, we shall define on this space a kind of shuffle product and
+this purpose, we first define a linear coproduct $\Delta$ on $\QQ[S]$.
+For any permutation $\pi$, we set
 \begin{equation} \label{equ:unshuffling_coproduct}
@@
 $\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
-coefficient $\lambda_{\sigma, \nu}^\pi$ is different from zero.
-Besides, a permutation $\pi$ is a {\em square} if there is a permutation
-$\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$. In this
-case, we say that $\sigma$ is a {\em square root} of $\pi$. Equivalently,
-$\pi$ is a square with $\sigma$ as square root if and only if in the
-expansion of $\Delta(\pi)$, there is a tensor $\sigma \otimes \sigma$
-with a nonzero coefficient.
+coefficient $\lambda_{\sigma, \nu}^\pi$ is different from zero. Besides,
+a permutation $\pi$ is a {\em square} if there is a permutation $\sigma$
+such that $\pi$ appears in $\sigma \SHUFFLE \sigma$. In this case, we
+say that $\sigma$ is a {\em square root} of $\pi$. Equivalently, $\pi$
+is a square with $\sigma$ as square root if and only if in the expansion
+of $\Delta(\pi)$, there is a tensor $\sigma \otimes \sigma$ with a
+nonzero coefficient.
 \medskip
@@
     An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
-    is equal to the last member of~\eqref{equ:shuffle_associative_commutative},
-    whence the associativity of $\SHUFFLE$.
+    is equal to the last member
+    of~\eqref{equ:shuffle_associative_commutative}, whence the
+    associativity of $\SHUFFLE$.
     \smallskip
@@
     If $P$ is a set of indexes of letters of $\pi$, we denote by
-    $\widetilde{\pi}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now since
+    $\widetilde{\pi}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
     this operation $\widetilde{\,}$ define a bijection on the set of the
-    subsets of $[|\pi|]$ and the standardisation operation commutes with
-    the mirror operation on words, we have
+    subsets of $[|\pi|]$, and since the standardization operation
+    commutes with the mirror operation on words, we have
     \begin{equation} \begin{split}
@@
     $\{\tau_i : i \in P\}$. Since the map sending a subset $P$ of
-    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection and
+    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
     $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
@@
 \end{proof}
+\medskip
+
+We now use the algebraic properties of $\SHUFFLE$ exhibited by
+Proposition~\ref{prop:endomorphisms} to obtain combinatorial properties
+of square permutations.
 \medskip

### 2015-07-04 article.tex
reduction

--- 2015-07-01 article.tex
@@
 \usepackage{subfigure}
-\usepackage{enumitem}
 
@@
 \DeclareMathOperator{\RIGHT}{r}
-\DeclareMathOperator{\QQ}{\mathbb{Q}}
-\DeclareMathOperator{\STD}{\mathrm{s}}
-\DeclareMathOperator{\SHUFFLE}{\bullet}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newtheorem{fact}{Fact}
 
@@
 \author{%
-  Samuele Giraudo\inst{1} \and
-  St\'ephane Vialette\inst{1}
+  Samuele Giraudo \and
+  St\'ephane Vialette
 }% end author
@@
 
-We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
-set of all permutations. If $\pi$ is a permutation, we denote by $|\pi|$
-its size. If $P$ is a subset of $[|\pi|]$, $\pi_{|P}$ is the subword of
-$\pi$ consisting in the letters of $\pi$ at the positions specified by
-the elements of $P$. Besides, we denote by $\widetilde{\pi}$ (resp.
-$\bar \pi$, $\pi^{-1}$) the mirror image (resp. complementary, inverse)
-of $\pi$. If $u$ is a word of integers without multiple occurrences of a
-same letter, $\STD(u)$ denotes the unique permutation of same size as
-$u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only if
-$\STD(u)_i < \STD(u)_j$.
-\medskip
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Algebraic issues ----
-%%
-\section{Algebraic issues}
-\label{section:Algebraic issues}
-Let use denote by $\QQ[S]$ the linear span of all permutations. We shall
-define on this space a kind of shuffle product. For this purpose, we
-first define a linear coproduct $\Delta$ on $\QQ[S]$ satisfying for
-any permutation $\pi$,
-\begin{equation} \label{equ:unshuffling_coproduct}
-    \Delta(\pi) =
-    \sum_{P_1 \uplus P_2 = [|\pi|]}
-    \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
-\end{equation}
-We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
-instance,
-\begin{equation}
-    \Delta(213) =
-    \epsilon \otimes 213 + 2. 1 \otimes 12 +
-    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
-\end{equation}
-\begin{equation}
-    \Delta(1234) =
-    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
-    4. 123 \otimes 1 + 1234 \otimes \epsilon,
-\end{equation}
-\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
-    \Delta(1432) & =
-    \epsilon \otimes 1432 + 3. 1 \otimes 132 + 1 \otimes 321 +
-    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
-    321 \otimes 1 + 1432 \otimes \epsilon.
-\end{split}\end{equation}
-\medskip
-
-As any coproduct, $\Delta$ leads to the definition of a product obtained
-by duality in the following way. From~\eqref{equ:unshuffling_coproduct},
-for any permutation $\pi$, we have
-\begin{equation}
-    \Delta(\pi) =
-    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
-    \sigma \otimes \nu,
-\end{equation}
-where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
-by definition of duality, the dual product of $\Delta$ denoted by
-$\SHUFFLE$ is a linear binary product on $\QQ[S]$. It satisfies, for any
-permutations
-$\sigma$ and $\nu$,
-\begin{equation}
-    \sigma \SHUFFLE \nu =
-    \sum_{\pi \in S}
-    \lambda_{\sigma, \nu}^\pi \; \pi.
-\end{equation}
-We call $\SHUFFLE$ the {\em shuffle product of permutations}. For
-instance,
-\begin{equation}\begin{split} \label{equ:example_shuffle_product}
-    12 \SHUFFLE 21 & =
-    1243 + 1324 + 2.1342 + 2.1423 + 3.1432 + 2134 + 2.2314 \\
-    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
-    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
-\end{split}\end{equation}
-Observe that the coefficient $3$ of the permutation $1432$
-in~\eqref{equ:example_shuffle_product} comes from the fact that the
-coefficient of the tensor $12 \otimes 21$ is $3$
-in~\eqref{equ:example_unshuffling_coproduct}.
-\medskip
-
-We say that a permutation $\pi$ {\em appears} in the shuffle
-$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
-coefficient $\lambda_{\sigma, \nu}^\pi$ is different from zero.
-Besides, a permutation $\pi$ is a {\em square} if there is a permutation
-$\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$. In this
-case, we say that $\sigma$ is a {\em square root} of $\pi$. Equivalently,
-$\pi$ is a square with $\sigma$ as square root if and only if in the
-expansion of $\Delta(\pi)$, there is a tensor $\sigma \otimes \sigma$
-with a nonzero coefficient.
-\medskip
-
-\begin{proposition} \label{prop:shuffle_associative_commutative}
-    The shuffle product $\SHUFFLE$ of permutations is associative and
-    commutative.
-\end{proposition}
-\begin{proof}
-    To prove the associativity of $\SHUFFLE$, it is convenient to show
-    that its dual coproduct $\Delta$ is coassociative, that is
-    \begin{equation}
-        (\Delta \otimes I) \Delta = (I \otimes \Delta) \Delta,
-    \end{equation}
-    where $I$ denote the identity map. This strategy relies on the fact
-    that a product is associative if and only if its dual coproduct is
-    coassociative. For any permutation $\pi$, we have
-    \begin{equation} \begin{split}
-    \label{equ:shuffle_associative_commutative}
-        (\Delta \otimes I) \Delta(\pi) & =
-        (\Delta \otimes I)
-        \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
-        & =
-        \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes I\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & =
-        \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \sum_{Q_1 \uplus Q_2 = [|P_1|]}
-        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
-        \otimes
-        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_2}\right)
-        \otimes \STD\left(\pi_{|P_2}\right) \\
-        & =
-        \sum_{P_1 \uplus P_2 \uplus P_3 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right) \otimes
-        \STD\left(\pi_{|P_2}\right) \otimes
-        \STD\left(\pi_{|P_3}\right).
-    \end{split} \end{equation}
-    An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
-    is equal to the last member of~\eqref{equ:shuffle_associative_commutative},
-    whence the associativity of $\SHUFFLE$.
-    \smallskip
-
-    Finally, to prove the commutativity of $\SHUFFLE$, we shall show
-    that $\Delta$ is cocommutative, that is for any permutation $\pi$,
-    if in the expansion of $\Delta(\pi)$ there is a tensor
-    $\sigma \otimes \nu$ with a coefficient $\lambda$, there is in the
-    same expansion the tensor $\nu \otimes \sigma$ with the same
-    coefficient $\lambda$. Clearly, a product is associative if and only
-    if its dual coproduct is coassociative. Now, from the
-    definition~\eqref{equ:unshuffling_coproduct} of $\Delta$, one
-    observes that if the pair $(P_1, P_2)$ of subsets of $[|\pi|]$
-    contributes to the coefficient of
-    $\STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right)$,
-    the pair $(P_2, P_1)$ contributes to the coefficient of
-    $\STD\left(\pi_{|P_2}\right) \otimes \STD\left(\pi_{|P_1}\right)$.
-    This shows that $\Delta$ is cocommutative and hence, that $\SHUFFLE$
-    is commutative.
-\end{proof}
-\medskip
-
-Proposition~\ref{prop:shuffle_associative_commutative} implies that
-$\QQ[S]$ endowed with the unshuffling coproduct $\Delta$ is a
-coassociative cocommutative coalgebra, or in an equivalent way, that
-$\QQ[S]$ endowed with the shuffle product $\SHUFFLE$ is an associative
-commutative algebra.
-\medskip
-
-\begin{proposition} \label{prop:endomorphisms}
-    The three linear maps
-    \begin{equation}
-        \phi_1, \phi_2, \phi_3 : \QQ[S] \to \QQ[S]
-    \end{equation}
-    linearly sending a permutation $\pi$ to, respectively,
-    $\widetilde{\pi}$, $\bar \pi$, and $\pi^{-1}$ are endomorphims of
-    associative algebras.
-\end{proposition}
-\begin{proof}
-    To prove that, for $j = 1, 2, 3$, $\phi_j$ is a morphism of
-    associative algebras, we have to prove that for all permutations
-    $\sigma$ and $\nu$,
-    \begin{equation}
-        \phi_j(\sigma \SHUFFLE \nu) =
-        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
-    \end{equation}
-    We shall proceed here as we have done during the proof of
-    Proposition~\ref{prop:shuffle_associative_commutative} by showing
-    that $\phi_j$ is a morphism of coalgebras, that is
-    \begin{equation}
-        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
-    \end{equation}
-    In the sequel, $\pi$ is a permutation.
-    \smallskip
-
-    If $P$ is a set of indexes of letters of $\pi$, we denote by
-    $\widetilde{\pi}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now since
-    this operation $\widetilde{\,}$ define a bijection on the set of the
-    subsets of $[|\pi|]$ and the standardisation operation commutes with
-    the mirror operation on words, we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_1(\pi))
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\phi_1(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|P_1}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \widetilde{\STD\left(\pi_{|P_1}\right)}
-        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
-    $\phi_1$ is a morphism of associative algebras.
-    \smallskip
-
-    Next, since by definition of the complementation operation on
-    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
-    we have $\tau_i < \tau_j$ if and only if $\bar \tau_i > \bar \tau_j$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_2(\pi))
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\phi_2(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\bar \pi_{|P_1}\right)
-        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
-    $\phi_2$ is a morphism of associative algebras.
-    \smallskip
-
-    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
-    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
-    $\{\tau_i : i \in P\}$. Since the map sending a subset $P$ of
-    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection and
-    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_3(\pi))
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\phi_3(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|P_1}\right)
-        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
-        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right)^{-1}
-        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
-        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
-    $\phi_3$ is a morphism of associative algebras.
-\end{proof}
-\medskip
-
-\begin{proposition} \label{prop:square_stability}
-    Let $\pi$ be a square permutation and $\rho$ be a square root of
-    $\pi$. Then,
-    \begin{enumerate}[label={\it (\roman*)},fullwidth]
-        \item \label{item:square_stability_1}
-        the permutation $\widetilde{\pi}$ is a square and
-        $\widetilde{\rho}$ is one of its square roots;
-        \item \label{item:square_stability_2}
-        the permutation $\bar \pi$ is a square and $\bar \rho$ is one of
-        its square roots;
-        \item \label{item:square_stability_3}
-        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$ is one of
-        its square roots.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}
-    All statements~\ref{item:square_stability_1},
-    \ref{item:square_stability_2}, and~\ref{item:square_stability_3} are
-    direct consequences of Proposition~\ref{prop:endomorphisms}. Indeed,
-    since $\pi$ is a square permutation and $\rho$ is a square root of
-    $\pi$, by definition $\pi$ appears in the product
-    $\rho \SHUFFLE \rho$. Now, by Proposition~\ref{prop:endomorphisms},
-    for all $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
-    algebras from $\QQ[S]$ to $\QQ[S]$, $\phi_j$ commutes with the
-    shuffle product of permutations $\SHUFFLE$. Hence, in particular,
-    one has
-    \begin{equation}
-        \phi_j(\rho \SHUFFLE \rho) =
-        \phi_j(\rho) \SHUFFLE \phi_j(\rho).
-    \end{equation}
-    Then, since $\pi$ appears in $\rho \SHUFFLE \rho$, $\phi_j(\pi)$
-    appears in $\phi_j(\rho) \SHUFFLE \phi_j(\rho)$. This
-    implies~\ref{item:square_stability_1}, \ref{item:square_stability_2},
-    and~\ref{item:square_stability_3}.
-\end{proof}
-\medskip
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{proposition}
+  \label{proposition:matching}
   Let $\pi \in S_{2n}$.
   Then,
-$\pi$ is the union of two order-isomorphic pattern
+  $\pi$ is the union of two order-isomorphic patterns
   if and only if
@@
 \begin{proposition}
+  \label{proposition:hardness}
   Deciding whether a permutation is the union of two
-order-isomorphic pattern is \NPC.
+  order-isomorphic patterns is \NPC.
 \end{proposition}
-
-\begin{align*}
-N_4 &= 2(2n + k + 2) + 1  = 4n + 2k + 5 \\
-N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 12k + 9 \\
-N_2 &= 2(2N_3 + 2n + 2k + 4) + 1 \\
-N_1 &= 2(2N_2 + 2n + 2k + 4) + 1\text{.}
-\end{align*}
-
-\begin{align*}
-\sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
-\pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2] \\
-\sigma'' &= \sigma \; [2N_2 + N_4] \\
-\pi''    &= \pi \; [2N_2 + N_4 + k] \\
-\nu_1    &= \nearrow_{N_1} \; [2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
-\nu'_1   &= \nearrow_{N_1} \; [N_1 + 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
-\nu_2    &= \nearrow_{N_2} \; [N_2] \\
-\nu'_2   &= \searrow_{N_2} \\
-\nu_3    &= \nearrow_{N_3} \; [2N_2 + 2N_4 + 2n + 2k + 4] \\
-\nu'_3   &= \nearrow_{N_3} \; [2N_2 + N_3 + 2N_4 + 2n + 2k + 4] \\
-\nu_4    &= \searrow_{N_4} \; [2N_2 + N_4 + 2n + 2k + 4] \\
-\nu'_4   &= \searrow_{N_4} \; [2N_2]
-\end{align*}
-
-$$
-\mu
-=
-\nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
-$$
-
-\begin{claim}
-  \label{claim:no nu_1 - nu_1 edge}
-  There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
-\end{claim}
-
-\begin{proof}[of Claim~\ref{claim:no nu_1 - nu_1 edge}]
-Suppose, aiming at a contradiction that there exists
-a $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
-
-\begin{tikzpicture}
-  [scale=0.25]
-  % nu_1
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
-  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % nu_4
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
-  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % nu'_1
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
-  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % edges
-  \draw [line width=1pt]
-  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
-  \draw [line width=1pt]
-  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
-\end{tikzpicture}
-\begin{tikzpicture}
-  [scale=0.25]
-  % nu_1
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
-  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % nu_4
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
-  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % nu'_1
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
-  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % edges
-  \draw [line width=1pt]
-  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
-  \draw [line width=1pt]
-  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
-\end{tikzpicture}
-\qed
-\end{proof}
-
-\begin{claim}
-There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-\end{claim}
 
@@
   \end{tikzpicture}
+  \caption{\label{fig:reduction}}
 \end{figure}
+
+\begin{align*}
+N_4 &= 2(2n + k + 2) + 1  = 4n + 2k + 5 \\
+N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 12k + 9 \\
+N_2 &= 2(2N_3 + 2n + 2k + 4) + 1 \\
+N_1 &= 2(2N_2 + 2n + 2k + 4) + 1\text{.}
+\end{align*}
+
+\begin{align*}
+\sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
+\pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2] \\
+\sigma'' &= \sigma \; [2N_2 + N_4] \\
+\pi''    &= \pi \; [2N_2 + N_4 + k] \\
+\nu_1    &= \nearrow_{N_1} \; [2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
+\nu'_1   &= \nearrow_{N_1} \; [N_1 + 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
+\nu_2    &= \nearrow_{N_2} \; [N_2] \\
+\nu'_2   &= \searrow_{N_2} \\
+\nu_3    &= \nearrow_{N_3} \; [2N_2 + 2N_4 + 2n + 2k + 4] \\
+\nu'_3   &= \nearrow_{N_3} \; [2N_2 + N_3 + 2N_4 + 2n + 2k + 4] \\
+\nu_4    &= \searrow_{N_4} \; [2N_2 + N_4 + 2n + 2k + 4] \\
+\nu'_4   &= \searrow_{N_4} \; [2N_2]
+\end{align*}
+
+$$
+\mu
+=
+\nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
+$$
+
+\begin{figure}[t!]
+  \centering
+  \subfigure[A nested $(\nu_1,\nu'_1)$-edge.]{%
+    \begin{tikzpicture}
+      [
+        scale=0.3,
+        label/.style={anchor=base}
+      ]
+      % nu_1
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
+      \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % nu_4
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
+      \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % nu'_1
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
+      \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % labels
+      \node [label] (x) at (17,14.5) {$x$};
+      % \node [label] (y) at (14,14.5) {$y$};
+      \node [label] (a) at (0,8.5) {$a$};
+      \node [label] (b) at (8,-1.5) {$b$};
+      \node (nu1) at (2.5,7.7) {$\nu_1$};
+      \node (nu2) at (8.5,-2.5) {$\nu_2$};
+      \node (nu1') at (14.5,13.5) {$\nu'_1$};
+      % edges
+      \draw [line width=1pt]
+      (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
+      \draw [line width=1pt]
+      (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
+      \draw [line width=1pt]
+      (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
+      \draw [line width=1pt]
+      (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
+      \draw [line width=1pt]
+      (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
+      \draw [line width=1pt]
+      (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
+    \end{tikzpicture}
+    \label{subfig:no (nu_1, nu_2)-edge - 2}}
+  \qquad
+  \subfigure[A nested $(\nu_1,\nu'_1)$-edge.]{%
+    \begin{tikzpicture}
+      [
+        scale=0.3,
+        label/.style={anchor=base}
+      ]
+      % nu_1
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
+      \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % nu_4
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
+      \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % nu'_1
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
+      \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % labels
+      \node [label] (x) at (3,8.5) {$x$};
+      \node [label] (y) at (14,14.5) {$y$};
+      \node [label] (a) at (1,8.5) {$a$};
+      \node [label] (b) at (8,-1.5) {$b$};
+      \node (nu1) at (2.5,7.7) {$\nu_1$};
+      \node (nu2) at (8.5,-2.5) {$\nu_2$};
+      \node (nu1') at (14.5,13.5) {$\nu'_1$};
+      % edges
+      \draw [line width=1pt]
+      (3,13) .. controls +(0,9) and +(0,4) .. (14,18);
+      \draw [line width=1pt]
+      (1,11) .. controls +(0,6) and +(0,16) .. (8,3);
+    \end{tikzpicture}
+    \label{subfig:no (nu_1, nu_2)-edge - 2}}
+    \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
+            Claim~\ref{claim:no (nu_1, nu_2)-edge}.%
+            }
+  \end{figure}
+
+\begin{claim}
+  \label{claim:no (nu_1, nu_2)-edge}
+  There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
+\end{claim}
+
+\begin{proof}[of Claim~\ref{claim:no (nu_1, nu_2)-edge}]
+  Suppose, aiming at a contradiction that there exists
+  a $(\nu_1, \nu_2)$-edge $e = (a, b)$ in $\mathcal{M}$.
+  Without loss of generality, assume $(a, b)$ is the leftmost
+  $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
+  \begin{itemize}
+  \item $a=1$. 
+  Since $\mathcal{M}$ is containment-free, it does
+  not contain any $(\nu_1, \nu_1)$-edge.
+  Frthermore,
+  since $N_1 > 2(2N_2 + 2n + 2k + 4) > 
+  |\nu_2| + |\nu'_2| + |\nu_3| + |\nu'_3| + |\nu_4| + |\nu'_4|
+  + |\sigma'| + |\pi'| + |\pi''| + |\sigma''|$, there exists 
+  a $(\nu_1, \nu'_1)$-edge $e' = (x, y)$ in $\mathcal{M}$.
+  We have 
+  $\LEFT(e) = \mu(a) < \LEFT(e') = \mu(x)$
+  and
+  $\RIGHT(e') = \mu(b) < \RIGHT(e') = \mu(y)$
+  (see Fig.~\ref{fig:reduction}).
+  Therefore, since $\mathcal{M}$ is containment-free, it does
+  not contain any $(\nu_2, \nu_2)$-edge.
+  According to Proposition~\ref{proposition:matching}, this is 
+  a contradiciton.
+  \end{itemize}
+  \qed
+\end{proof}
+
+\begin{fact}
+  \label{fact:no (nu_2, nu_2)-edge}
+  There is no $(\nu_2, \nu_2)$-edge in $\mathcal{M}$.
+\end{fact}
+
+\begin{claim}
+  \label{fact:one (nu_1, nu'_1)-edge}
+  There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+\end{claim}
+
+\begin{claim}
+  \label{fact:no (nu_2, nu'_1)-edge}
+  There is no $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
+\end{claim}
+
+\begin{claim}
+  \label{fact:one (nu_2, nu'_2)-edge}
+  There is at least one $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
+\end{claim}
+
+\begin{fact}
+  \label{fact:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge}
+  There is 
+  no $(\nu'_1, \nu'_1)$-edge, 
+  no $(\nu_3, \nu_3)$-edge,
+  no $(\sigma', \sigma')$-edge and
+  no $(\nu_4, \nu_4)$-edge
+  in $\mathcal{M}$.
+\end{fact}
+
+\begin{claim}
+  \label{fact:no (nu_3, nu'_2)-edge}
+  There is no $(\nu_3, \nu'_2)$-edge in $\mathcal{M}$.
+\end{claim}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% ---- Patterns ----
+%%
+\section{Patterns}
+\label{section:Patterns}
+
+\begin{proposition}
+  \label{proposition:patterns:lower bound}
+  Let $\pi \in S_{n}$.
+  Then $\pi$ contains a pattern of size at least 
+  $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
+  is the union of two order-isomorphic patterns.
+\end{proposition}
+
+\begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
+  The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935} 
+  states that every permutation of 
+  length at least $pq+1$ must contain either the pattern
+  $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
+  The result now follows from the fact that
+  the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
+\qed
+\end{proof}
+
+\begin{proposition}
+  \label{proposition:patterns:upper bound}
+  If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
+  a permutation of $S_{n}$ 
+  that does not contains a pattern
+  of length $2k$ that is the union of two
+  order-isomorphic patterns.
+\end{proposition}
+
+\begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
+The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
+Let $\pi \in S_n$ be chosen randomly and uniformly.
+For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
+of $\pi$,
+let $X_\sigma$ be the indicator variable that 
+$\sigma$ is the union of two order-ismorphic patterns.
+According to Boole's inequality, we have
+$P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
+Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
+the probability that $\pi$ contains a pattern of length $2k$ 
+is the union of two order-ismorphic patterns is at most
+$\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
+Thus, with positive probability, no event $X_\sigma$ occurs and 
+there is a permutation of $S_n$ that does not contain a pattern
+of length $2k$ that is the union of two order-isomorphic patterns.
+\qed
+\end{proof}
 
@@
 bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla
-bla bla blabla bla bla blabla bla bla blabla bla bla blabla .
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla.
 
@@
 
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla
+bla bla blabla bla bla blabla bla bla blabla bla bla blabla.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-09-11 article.tex
reduction

--- 2015-07-02 article.tex
@@
 \usepackage{enumitem}
+\usepackage{shuffle}
 
 \usepackage{tikz}
-\usetikzlibrary{matrix, calc, arrows}
+\usetikzlibrary{matrix, calc, arrows, shapes.callouts}
 
@@
 \DeclareMathOperator{\SHUFFLE}{\bullet}
+\DeclareMathOperator{\BINTOPERM}{\mathrm{btp}}
+\DeclareMathOperator{\PERMTOBIN}{\mathrm{ptb}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newtheorem{fact}{Fact}
 
@@
 
+We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
+Let us recall here the most important ones. The {\em empty word} is
+denoted by $\epsilon$. If $u$ is a word, we denote by $\widetilde{u}$
+the {\em mirror image} of $u$. If $P$ is a subset of $[|u|]$, $u_{|P}$
+is the subword of $u$ consisting in the letters of $u$ at the positions
+specified by the elements of $P$. The {\em shuffle} of two words $u$ and
+$v$ is the set recursively defined by
+$u \shuffle \epsilon = u = \epsilon \shuffle u$ and
+$ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
+$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
+such that $u$ belongs to $v \shuffle v$.
+\medskip
+
 We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
-set of all permutations. If $\pi$ is a permutation, we denote by $|\pi|$
-its size. The unique permutation of size $0$ is denoted by $\epsilon$.
-If $P$ is a subset of $[|\pi|]$, $\pi_{|P}$ is the subword of $\pi$
-consisting in the letters of $\pi$ at the positions specified by the
-elements of $P$. Besides, we denote by $\widetilde{\pi}$ (resp.
-$\bar \pi$, $\pi^{-1}$) the mirror image (resp. complementary, inverse)
-of $\pi$. If $u$ is a word of integers without multiple occurrences of a
-same letter, $\STD(u)$ denotes the unique permutation of same size as
-$u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only if
-$\STD(u)_i < \STD(u)_j$.
-\medskip
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Algebraic issues ----
-%%
-\section{Algebraic issues}
-\label{section:Algebraic issues}
+set of all permutations. In this paper, permutations are specified by
+words without repetition of length $n$ on the alphabet $[n]$, so that
+all above definitions about words remain valid on permutations. If
+$\pi$ is a permutation, we denote by $\bar \pi$ (resp.~$\pi^{-1}$) the
+{\em complementary} (resp. {\em inverse}) of $\pi$. If $u$ is a word of
+integers without multiple occurrences of a same letter, $\STD(u)$ is the
+{\em standarized} of $u$, that is the unique permutation of the same
+size as $u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only
+if $\STD(u)_i < \STD(u)_j$. In particular, the image of the map $\STD$
+is the set $S$ of all permutations. Two words $u$ and $v$ having the
+same standarized are {\em order isomorphic}. If $\sigma$ is a permutation,
+there is an {\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if
+there is a set $P$ of indexes of letters of $\pi$ such that $\sigma$ and
+$\pi_{|P}$ are order isomorphic. When $\pi$ does not admit any occurrence
+of $\sigma$, $\pi$ {\em avoids} $\sigma$.
+\medskip
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% ---- Shuffle product on permutations ----
+%%
+\section{Shuffle product on permutations}
+\label{section:Shuffle product on permutations}
+
 The purpose of this section is to define a shuffle product $\SHUFFLE$
-on permutations and establish some of its properties. To define this
-shuffle product, we shall first define a coproduct~\cite{Joni:Rota:1979}
-denoted by $\Delta$. By duality, $\Delta$ implies the definition of
-$\SHUFFLE$. It is worth to note that, as we will see, algebraic
-properties of $\Delta$ imply combinatorial properties of $\SHUFFLE$.
+on permutations. To define this shuffle product, we shall first define
+a coproduct denoted by $\Delta$, enabling to unshuffle permutations.
+By duality, $\Delta$ implies the definition of $\SHUFFLE$. The reason
+why we need to pass by the definition of $\Delta$ to define $\SHUFFLE$
+is justified by the fact that a lot of properties of $\SHUFFLE$ depend
+of properties of $\Delta$ and that this strategy allows to write concise
+and clear proofs of them. We invite the reader unfamiliar with the
+concepts of coproduct and duality to consult~\cite{Joni:Rota:1979}
+or~\cite{Grinberg:Reiner:2014}.
 \medskip
@@
 announced, we shall define on this space a kind of shuffle product and
-this purpose, we first define a linear coproduct $\Delta$ on $\QQ[S]$.
+for this purpose, we first define a linear coproduct $\Delta$ on $\QQ[S]$.
 For any permutation $\pi$, we set
@@
     \Delta(1432) & =
-    \epsilon \otimes 1432 + 3. 1 \otimes 132 + 1 \otimes 321 +
+    \epsilon \otimes 1432 + 3. {\bf 1 \otimes 132} + 1 \otimes 321 +
     3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
@@
 three ways to extract two disjoint subwords of the permutation $1432$
-so that the letters of these two subwords have the same relative order
-as the letters of the permutations $1$ and $132$.
+so that the letters of these two subwords have respectively the same
+relative order as the letters of the permutations $1$ and $132$.
 \medskip
@@
 for any permutation $\pi$, we have
-\begin{equation}
+\begin{equation} \label{equ:schematic_coproduct}
     \Delta(\pi) =
@@
     \sum_{\pi \in S}
-    \lambda_{\sigma, \nu}^\pi \; \pi.
+    \lambda_{\sigma, \nu}^\pi \; \pi,
 \end{equation}
-We call $\SHUFFLE$ the {\em shuffle product of permutations}. For instance,
+where the coefficients $\lambda_{\sigma, \nu}^\pi$ are the ones
+of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
+{\em shuffle product of permutations}. For instance,
 \begin{equation}\begin{split} \label{equ:example_shuffle_product}
     12 \SHUFFLE 21 & =
-    1243 + 1324 + 2.1342 + 2.1423 + 3.1432 + 2134 + 2.2314 \\
+    1243 + 1324 + 2.1342 + 2.1423 + 3. {\bf 1432} + 2134 + 2.2314 \\
     & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
@@
 $\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
-coefficient $\lambda_{\sigma, \nu}^\pi$ is different from zero. Besides,
-a permutation $\pi$ is a {\em square} if there is a permutation $\sigma$
-such that $\pi$ appears in $\sigma \SHUFFLE \sigma$. In this case, we
-say that $\sigma$ is a {\em square root} of $\pi$. Equivalently, $\pi$
-is a square with $\sigma$ as square root if and only if in the expansion
-of $\Delta(\pi)$, there is a tensor $\sigma \otimes \sigma$ with a
-nonzero coefficient.
+coefficient $\lambda_{\sigma, \nu}^\pi$ defined above is different from
+zero. In a more combinatorial way, this is equivalent to say that there
+are two sets $P_1$ and $P_2$ of disjoints indexes of letters of $\pi$
+satisfying $P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$
+is order isomorphic to $\sigma$ and the subword $\pi_{|P_2}$ is order
+isomorphic to $\nu$.
+\medskip
+
+Besides, a permutation $\pi$ is a {\em square} if there is a
+permutation $\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
+In this case, we say that $\sigma$ is a {\em square root} of $\pi$.
+Equivalently, $\pi$ is a square with $\sigma$ as square root if and only
+if in the expansion of $\Delta(\pi)$, there is a tensor
+$\sigma \otimes \sigma$ with a nonzero coefficient. In a more
+combinatorial way, this is equivalent to say that there are two sets
+$P_1$ and $P_2$ of disjoints indexes of letters of $\pi$ satisfying
+$P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$ and
+$\pi_{|P_2}$ are order isomorphic. Computer experiments give us the
+first numbers of square permutations with respects to their size, which
+are, from size $0$ to $10$,
+\begin{equation}
+    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
+\end{equation}
+For instance, the square permutations of sizes $0$ to $4$ are
+\begin{multline}
+    \epsilon, 12, 21,
+    1234, 1243, 1423, 1324, 1342, 4132, 3124, 3142, \\ 3412, 4312,
+    2134, 2143, 2413, 4213, 2314, 2431, 4231, 3241, 3421, 4321.
+\end{multline}
+\medskip
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% ---- Binary square words and permutations ----
+%%
+\section{Binary square words and permutations}
+\label{section:Binary square words and permutations}
+In this section, we shall show that the square binary words are in
+one-to-one correspondence with square permutations avoiding some
+patterns. This property establishes a link between the shuffle of binary
+words and our shuffle of permutations and allows to obtain a new
+description of square binary words.
+\medskip
+
+Let $u$ be a binary word of length $n$ with $k$ occurrences of $0$.
+We denote by $\BINTOPERM$ the map sending any such word $u$ to the
+permutation obtained by replacing from left to right each occurrence of
+$0$ in $u$ by $1$, $2$, \dots, $k$, and from right to left each
+occurrence of $1$ in $u$ by $k + 1$, $k + 2$, \dots, $n$. For instance,
+\begin{equation}
+    \BINTOPERM({\bf 1}00{\bf 1}0{\bf 1}{\bf 1}0{\bf 1}000) =
+    {\bf C} 12 {\bf B} 3 {\bf A} {\bf 9} 4 {\bf 8} 567.
+\end{equation}
+Observe that for any nonempty permutation $\pi$ in the image of
+$\BINTOPERM$, there is exactly one binary word $u$ such that
+$\BINTOPERM(u0) = \BINTOPERM(u1) = \pi$. In support of this observation,
+when $\pi$ has an even size, we denote by $\PERMTOBIN(\pi)$ the word $ua$
+such that $|ua|_0$ and $|ua|_1$ are both even, where $a \in \{0, 1\}$.
+\medskip
+
+\begin{proposition} \label{prop:bijection_binary_to_permutations_squares}
+    For any $n \geq 0$, the map $\BINTOPERM$ restricted to the set of
+    square binary words of length $2n$ is a bijection between this last
+    set and the set of square permutations of size $2n$ avoiding the
+    patterns $213$ and $231$.
+\end{proposition}
+\begin{proof}
+    The statement of the proposition is a consequence of the
+    following claims implying that $\PERMTOBIN$ is the inverse
+    map of $\BINTOPERM$ over the set of square binary words.
+    \begin{claim} \label{claim:binary_to_permutation_avoiding}
+        The image of $\BINTOPERM$ is the set of all permutations
+        avoiding $213$ and $231$.
+    \end{claim}
+    \begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
+        Let us first show that the image of $\BINTOPERM$ contains only
+        permutations avoiding $213$ and $231$. Let $u$ be a binary word,
+        $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
+        positions of the occurrences of $0$ (resp. $1$) in $u$. By
+        definition of $\BINTOPERM$, from left to right, the subword
+        $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
+        is decreasing, and all letters of $w$ are greater than those
+        of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
+        Then, since $v$ is increasing and $w$ is decreasing, there is an
+        occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
+        occurrence of $21$ (resp. $2$, $1$). All these three cases
+        contradict the fact that all letters of $w$ are greater than
+        those of $v$. A similar argument shows that $\pi$ avoids~$231$
+        as well.
+        \smallskip
+
+        Finally, observe that any permutation $\pi$ avoiding $213$ and
+        $231$ necessarily starts by the smallest possible letter or the
+        greatest possible letter. This property is then true for the
+        suffix of $\pi$ obtained by deleting its first letter,
+        and so on for all of its suffixes. Thus, by
+        replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
+        role of a smallest (resp. greatest) letter, one obtains a binary
+        word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
+        avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
+        \qed
+    \end{proof}
+    \begin{claim} \label{claim:square_binary_to_square_permutation}
+        If $u$ is a square binary word, $\BINTOPERM(u)$ is a square
+        permutation.
+    \end{claim}
+    \begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
+        Since $u$ is a square binary word, there is a binary word $v$
+        such that $u \in v \shuffle v$. Then, there are two disjoint
+        sets $P$ and $Q$ of positions of letters of $u$ such that
+        $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
+        words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
+        same standarized $\rho$. Hence, and by definition of the shuffle
+        product of permutations, $\BINTOPERM(u)$ appears in
+        $\rho \SHUFFLE \rho$, showing that $\BINTOPERM(u)$ is a square
+        permutation.
+        \qed
+    \end{proof}
+    \begin{claim} \label{claim:square_permutation_to_square_binary}
+        If $\pi$ is a square permutation avoiding $213$ and $231$,
+        $\PERMTOBIN(\pi)$ is a square binary word.
+    \end{claim}
+    \begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
+        Let $\pi$ be a square permutation avoiding $213$ and $231$. By
+        Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
+        the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
+        well-defined binary word. Since $\pi$ is a square permutation,
+        there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
+        of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are order
+        isomorphic. This implies, by the definitions of $\BINTOPERM$ and
+        $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$ is a
+        square binary word.
+        \qed
+    \end{proof}
+\end{proof}
+\medskip
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% ---- Algebraic issues ----
+%%
+\section{Algebraic issues}
+\label{section:Algebraic issues}
+The aim of this section is to establish some of properties of the
+shuffle product of permutations $\SHUFFLE$. It is worth to note that, as
+we will see, algebraic properties of the unshuffling coproduct $\Delta$
+of permutations defined in
+Section~\ref{section:Shuffle product on permutations} lead to
+combinatorial properties of $\SHUFFLE$.
 \medskip
@@
         & =
-        \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \sum_{P_1 \uplus P_2 = [|\pi|]} \;
         \sum_{Q_1 \uplus Q_2 = [|P_1|]}
@@
     same expansion the tensor $\nu \otimes \sigma$ with the same
-    coefficient $\lambda$. Clearly, a product is associative if and only
-    if its dual coproduct is coassociative. Now, from the
+    coefficient $\lambda$. Clearly, a product is commutative if and only
+    if its dual coproduct is cocommutative. Now, from the
     definition~\eqref{equ:unshuffling_coproduct} of $\Delta$, one
@@
     is commutative.
+    \qed
 \end{proof}
@@
 \begin{proof}
-    To prove that, for $j = 1, 2, 3$, $\phi_j$ is a morphism of
+    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
     associative algebras, we have to prove that for all permutations
@@
     If $P$ is a set of indexes of letters of $\pi$, we denote by
-    $\widetilde{\pi}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
-    this operation $\widetilde{\,}$ define a bijection on the set of the
+    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
+    the operation $\widetilde{\,}$ define a bijection on the set of the
     subsets of $[|\pi|]$, and since the standardization operation
@@
     permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
-    we have $\tau_i < \tau_j$ if and only if $\bar \tau_i > \bar \tau_j$,
+    we have $\tau_i < \tau_k$ if and only if $\bar \tau_i > \bar \tau_k$,
     we have
@@
     $\phi_3$ is a morphism of associative algebras.
+    \qed
 \end{proof}
@@
     since $\pi$ is a square permutation and $\rho$ is a square root of
-    $\pi$, by definition $\pi$ appears in the product
+    $\pi$, by definition, $\pi$ appears in the product
     $\rho \SHUFFLE \rho$. Now, by Proposition~\ref{prop:endomorphisms},
-    for all $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
+    for any $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
     algebras from $\QQ[S]$ to $\QQ[S]$, $\phi_j$ commutes with the
@@
     Then, since $\pi$ appears in $\rho \SHUFFLE \rho$, $\phi_j(\pi)$
-    appears in $\phi_j(\rho) \SHUFFLE \phi_j(\rho)$. This
+    appears in $\phi_j(\rho \SHUFFLE \rho)$ and also in
+    $\phi_j(\rho) \SHUFFLE \phi_j(\rho)$. This shows that
+    $\phi_j(\rho)$ is a square root of $\phi_j(\pi)$ and
     implies~\ref{item:square_stability_1}, \ref{item:square_stability_2},
     and~\ref{item:square_stability_3}.
-\end{proof}
+    \qed
+\end{proof}
+\medskip
+
+Let us make an observation about Wilf-equivalence classes of
+permutations restrained on square permutations. The fact that there is a
+single Wilf-equivalence class of permutations of size $3$ together with
+Proposition~\ref{prop:square_stability} imply that $123$ and $321$ are
+in the same Wilf-equivalence class of square permutations, and that
+$132$, $213$, $231$, and $312$ are in the same Wilf-equivalence class of
+square permutations. Computer experiments show us that there are two
+Wilf-equivalence classes of square permutations of size $3$. Indeed, the
+number of square permutations avoiding $123$ begins by
+\begin{equation}
+    1, 0, 2, 0, 12, 0, 118, 0, 1218, 0, 14272,
+\end{equation}
+while the number of square permutations avoiding $132$ begins by
+\begin{equation}
+    1, 0, 2, 0, 11, 0, 84, 0, 743, 0, 7108.
+\end{equation}
+\medskip
+
+Besides, an other consequence of Proposition~\ref{prop:square_stability}
+is that its makes sense to enumerate the sets of square permutations
+quotiented by the operations of mirror image, complementary, and
+inverse. The sequence enumerating these sets begins by
+\begin{equation}
+    1, 0, 1, 0, 6, 0, 81, 0, 2774.
+\end{equation}
 \medskip
@@
 
+The following easy lemma will prove extremely useful for simplifying the proof of
+Proposition~\ref{proposition:hardness}.
+
+\begin{lemma}
+  \label{lemma:at most one edge monotone}
+  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
+  be a permutation with $n_1 = |\sigma_1| \geq 2$ and $n_2 = |\sigma_2| \geq 2$,
+  and $\mathcal{M}$ be a containment-free perfect matching of $G_\pi$.
+  The following assertions hold.
+  \begin{enumerate}
+    \item
+    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing and
+    $\sigma_1$ is above $\sigma_2$
+    (see Fig.~\ref{subfig:increasing before and aboce decreasing}),
+    then $\mathcal{M}$ sontains at most one $(\sigma_1, \sigma_2)$-edge.
+    \item
+    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing and
+    $\sigma_1$ is below $\sigma_2$
+    (see Fig.~\ref{subfig:decresing before and below increasing}),
+    then $\mathcal{M}$ sontains at most one $(\sigma_1, \sigma_2)$-edge.
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
+  \begin{figure}[t!]
+    \centering
+    \subfigure[%
+      An increasing pattern before and above a decreasing pattern.
+    ]{%
+      \begin{tikzpicture}
+        [
+          scale=0.45,
+          label/.style={anchor=base}
+        ]
+        % increasing pattern
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
+        \foreach \x/\y in {0/6,1/7,2/8,3/9,4/10,5/11} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % decreasing pattern
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
+        \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (a) at (1,5) {$a$};
+        \node [label] (b) at (4,5) {$b$};
+        \node [label] (ap) at (7,-1) {$a'$};
+        \node [label] (bp) at (10,-1) {$b'$};
+        \node (nu1) at (2.5,4) {$\sigma_1$};
+        \node (nu2) at (8.5,-2) {$\sigma_2$};
+        % crossings edges
+        \draw [line width=1pt]
+        (1,7) .. controls +(0,12) and +(0,10) .. (7,4);
+        \draw [line width=1pt]
+        (4,10) .. controls +(0,12) and +(0,10) .. (10,1);
+      \end{tikzpicture}
+      \label{subfig:increasing before and aboce decreasing}
+    }% end subfigure
+    \qquad
+    \subfigure[%
+      A decreasing pattern before and below an increasing pattern.
+    ]{%
+    \begin{tikzpicture}
+      [
+        scale=0.45,
+        label/.style={anchor=base}
+      ]
+      % decreasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (5.2,5.2);
+      \foreach \x/\y in {0/5,1/4,2/3,3/2,4/1,5/0} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % decreasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,5.8) grid (11.2,11.2);
+      \foreach \x/\y in {6/6,7/7,8/8,9/9,10/10,11/11} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % labels
+      \node [label] (a) at (1,-1) {$a$};
+      \node [label] (b) at (4,-1) {$b$};
+      \node [label] (ap) at (7,5) {$a'$};
+      \node [label] (bp) at (10,5) {$b'$};
+      \node (nu1) at (2.5,-2) {$\sigma_1$};
+      \node (nu2) at (8.5,4) {$\sigma_2$};
+      % crossings edges
+      \draw [line width=1pt]
+      (1,4) .. controls +(0,10) and +(0,12) .. (7,7);
+      \draw [line width=1pt]
+      (4,1) .. controls +(0,10) and +(0,12) .. (10,10);
+      \end{tikzpicture}
+      \label{subfig:decresing before and below increasing}
+    }% end subfigure
+    \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
+      Lemma~\ref{lemma:at most one edge monotone}:
+      Assuming a $(\nu_1,\nu_2)$-edge $(a, b)$ in $\mathcal{M}$.
+    }
+  \end{figure}
+  We first prove \emph{1}.
+  Suppose, aiming at a contradiction, that there exist two
+  $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in $\mathcal{M}$.
+  Since $\mathcal{M}$ is containment-free, we may assume $a < b$ and $a' < b'$.
+  Therefore, we have
+  $\pi(a) < \pi(b)$ (since $\sigma_1$ is increasing) and
+  $\pi(a') > \pi(b')$ (since $\sigma_2$ is decreasing),
+  which contradicts our hypothesis.
+
+  As for \emph{2}, suppose, still aiming at a contradiction,
+  that there exist two
+  $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in $\mathcal{M}$.
+  Since $\mathcal{M}$ is containment-free, we may assume $a < b$ and $a' < b'$.
+  Therefore, we have
+  $\pi(a) > \pi(b)$ (since $\sigma_1$ is decreasing) and
+  $\pi(a') < \pi(b')$ (since $\sigma_2$ is increasing),
+  which contradicts our hypothesis.
+  \qed
+\end{proof}
+
 \begin{proposition}
+  \label{proposition:hardness}
   Deciding whether a permutation is the union of two
-order-isomorphic pattern is \NPC.
+  order-isomorphic patterns is \NPC.
 \end{proposition}
 
-\begin{align*}
-N_4 &= 2(2n + k + 2) + 1  = 4n + 2k + 5 \\
-N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 12k + 9 \\
-N_2 &= 2(2N_3 + 2n + 2k + 4) + 1 \\
-N_1 &= 2(2N_2 + 2n + 2k + 4) + 1\text{.}
-\end{align*}
-
-\begin{align*}
-\sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
-\pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2] \\
-\sigma'' &= \sigma \; [2N_2 + N_4] \\
-\pi''    &= \pi \; [2N_2 + N_4 + k] \\
-\nu_1    &= \nearrow_{N_1} \; [2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
-\nu'_1   &= \nearrow_{N_1} \; [N_1 + 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
-\nu_2    &= \nearrow_{N_2} \; [N_2] \\
-\nu'_2   &= \searrow_{N_2} \\
-\nu_3    &= \nearrow_{N_3} \; [2N_2 + 2N_4 + 2n + 2k + 4] \\
-\nu'_3   &= \nearrow_{N_3} \; [2N_2 + N_3 + 2N_4 + 2n + 2k + 4] \\
-\nu_4    &= \searrow_{N_4} \; [2N_2 + N_4 + 2n + 2k + 4] \\
-\nu'_4   &= \searrow_{N_4} \; [2N_2]
-\end{align*}
-
-$$
-\mu
-=
-\nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
-$$
-
-\begin{claim}
-  \label{claim:no nu_1 - nu_1 edge}
-  There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
-\end{claim}
-
-\begin{proof}[of Claim~\ref{claim:no nu_1 - nu_1 edge}]
-Suppose, aiming at a contradiction that there exists
-a $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
-
-\begin{tikzpicture}
-  [scale=0.25]
-  % nu_1
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
-  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % nu_4
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
-  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % nu'_1
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
-  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % edges
-  \draw [line width=1pt]
-  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
-  \draw [line width=1pt]
-  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
-\end{tikzpicture}
-\begin{tikzpicture}
-  [scale=0.25]
-  % nu_1
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
-  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % nu_4
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
-  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % nu'_1
-  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
-  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
-      \draw [fill=black] (\x,\y) circle (0.1);
-  }
-  % edges
-  \draw [line width=1pt]
-  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
-  \draw [line width=1pt]
-  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
-\end{tikzpicture}
-\qed
-\end{proof}
-
-\begin{claim}
-There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-\end{claim}
+\begin{proof}
+  For the sake of clarity, we will that a bunch of positions $P$
+  is \emph{above} (resp. \emph{below}) above another bunch of positions $P'$
+  in permutation $\mu$
+  if $\mu(x) > \mu(y)$ (resp. $\mu(x) < \mu(y)$) for every $x \in P$
+  and every $y \in P'$.
+  For exemple, $\nu_1$ is above $\nu_2$ but below $\nu'_1$ in permutation $\mu$.
 
@@
       % row size
-    \foreach \y/\N in {0.5/N_2,1.5/N_2} {
+      \foreach \y/\N in {0.5/N_2,1.5/N_2,2.5/N_4,3.5/k,4.5/n,5.5/n+2,6.5/k+2,7.5/N_4,8.5/N_3,9.5/N_3,10.5/N_1,11.4/N_1} {
           \node at (-0.5,\y) (R\y) {$\N$};
@@
     \end{tikzpicture}
+    \caption{\label{fig:reduction}}
   \end{figure}
+
+  \begin{align*}
+  N_4 &= 2(2n + k + 2) + 1  = 4n + 2k + 5 \\
+  N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 12k + 9 \\
+  N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 \\
+  N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1\text{.}
+  \end{align*}
+  Notice that $N_1, N_2, N_3$ and $N_4$ are polynomial in $n$.
+  The crucial properties are that
+  (i) $N_1, N_2, N_3$ and $N_4$ are odd integers
+  and
+  (ii) $N_i > \left(\sum_{i < j \leq 4} 2N_j\right) + 2n + 2k + k$
+  for every $1 \leq i \leq 4$.
+
+  We now turn to defining various gadgets (sequences of integers)
+  that act as building blocks in our construction:
+  \begin{align*}
+  \sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
+  \pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2] \\
+  \sigma'' &= \sigma \; [2N_2 + N_4] \\
+  \pi''    &= \pi \; [2N_2 + N_4 + k] \\
+  \nu_1    &= \nearrow_{N_1} \; [2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
+  \nu'_1   &= \nearrow_{N_1} \; [N_1 + 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
+  \nu_2    &= \nearrow_{N_2} \; [N_2] \\
+  \nu'_2   &= \searrow_{N_2} \\
+  \nu_3    &= \nearrow_{N_3} \; [2N_2 + 2N_4 + 2n + 2k + 4] \\
+  \nu'_3   &= \nearrow_{N_3} \; [2N_2 + N_3 + 2N_4 + 2n + 2k + 4] \\
+  \nu_4    &= \searrow_{N_4} \; [2N_2 + N_4 + 2n + 2k + 4] \\
+  \nu'_4   &= \searrow_{N_4} \; [2N_2]\text{.}
+  \end{align*}
+  We are now in position to define our target permutation $\mu$:
+  $$
+  \mu
+  =
+  \nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
+  \text{.}
+  $$
+  We claim that $\sigma$ occurs as a pattern in $\pi$ if and only if
+  there exists a containment-free perfect matching
+  $\mathcal{M}$ in $G_\mu$ such that
+  for any two disctinct edges
+  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
+  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+
+  Suppose first that $\sigma$ occurs in $\pi$ as a pattern.
+  We show how to construct a containment-free perfect matching
+  $\mathcal{M}$ in $G_\mu$ such that
+  for any two disctinct edges
+  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
+  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+  \begin{itemize}
+    \item $\mathcal{M}$ contains $N_1$ pairwise crossing
+    $(\nu_1, \nu'_1)$-edges.
+    \item $\mathcal{M}$ contains $N_2$ pairwise crossing
+    $(\nu_2, \nu'_2)$-edges.
+    \item $\mathcal{M}$ contains $N_3$ pairwise crossing
+    $(\nu_3, \nu'_3)$-edges.
+    \item $\mathcal{M}$ contains $N_4$ pairwise crossing
+    $(\nu_4, \nu'_4)$-edges.
+    \item $\mathcal{M}$ contains $k+2$ pairwise crossing
+    $(\sigma', \pi')$-edges as depicted in Figure~\ref{}.
+    (Notice that we use here the fact that $\sigma$ occurs as a pattern in $\pi$).
+    \item $\mathcal{M}$ contains $n-k$ pairwise crossing
+    $(\pi', \pi'')$-edges as depicted in Figure~\ref{}.
+    \item $\mathcal{M}$ contains $k$ pairwise crossing
+    $(\pi'', \sigma'')$-edges as depicted in Figure~\ref{}.
+    (Notice that we use here again the fact that $\sigma$ occurs as a pattern in $\pi$).
+  \end{itemize}
+  It can be easily checked (probably referring to Figure~\ref{}) that
+  $\mathcal{M}$ is perfect and that
+  for any two disctinct edges
+  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
+  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+
+    \begin{figure}[t!]
+      \centering
+        \begin{tikzpicture}
+          [
+            scale=0.45,
+            label/.style={anchor=base},
+            cell/.style={draw,thick,fill=black!5}
+          ]
+          % sigma'
+          \draw [cell,rounded corners] (1.5,16.5) -- (4.5,16.5) -- (4.5,19.5) -- (1.5,19.5) -- cycle;
+          \draw[step=1cm,black!60,ultra thin,fill=black!10] (0.8,16.8) grid (5.2,21.2);
+          \foreach \x/\y in {1/20,2/19,3/17,4/18,5/21} {
+              \draw [fill=black] (\x,\y) circle (0.1);
+          }
+          \node (pip) at (3,22) {$\sigma'$};
+          \node[%
+            draw,
+            text width=5.5cm,text justified,color=black!50,rounded corners,
+            rectangle callout,callout relative pointer={(0.0cm,-1.5cm)}
+          ] at (3,26)
+          {%
+            $((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$
+          };%
+          %\node at (3,-0.5) {$((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$};
+          \node at (-0.5,20) {$k+1$};
+          \node at (-0.5,21) {$k+2$};
+          % pi'
+          \draw [cell,rounded corners] (6.5,8.5) -- (12.5,8.5) -- (12.5,14.5) -- (6.5,14.5) -- cycle;
+          \draw[step=1cm,black!60,ultra thin,fill=black!10] (5.8,8.8) grid (13.2,16.2);
+          \foreach \x/\y in {6/15,7/12,8/13,9/10,10/9,11/11,12/14,13/16} {
+              \draw [fill=black] (\x,\y) circle (0.1);
+          }
+          \node (pip) at (9.5,17) {$\pi'$};
+          \node[%
+            draw,
+            text width=5.5cm,text justified,color=black!50,rounded corners,
+            rectangle callout,callout relative pointer={(0.0cm,3cm)}
+          ] at (9.5,0)
+          {%
+            $((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$
+          };%
+          %\node at (9.5,-2) {$((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$};
+          \node at (-0.5,15) {$n+1$};
+          \node at (-0.5,16) {$n+2$};
+          % pi''
+          \draw [cell,rounded corners] (13.5,2.5) -- (19.5,2.5) -- (19.5,8.5) -- (13.5,8.5) -- cycle;
+          \draw[step=1cm,black!60,ultra thin,fill=black!10] (13.8,2.8) grid (19.2,8.2);
+          \foreach \x/\y in {14/6,15/7,16/4,17/3,18/5,19/8} {
+              \draw [fill=black] (\x,\y) circle (0.1);
+          }
+          \node (pipp) at (16.75,9) {$\pi''$};
+          \node[%
+            draw,
+            text width=2.25cm,text justified,color=black!50,rounded corners,
+            rectangle callout,callout relative pointer={(0.0cm,-4cm)}
+          ] at (16.5,19)
+          {%
+            $\pi [2N_2 + N_4 + k]$
+          };%
+          %\node at (16.5,-3.5) {$\pi [2N_2 + N_4 + k]$};
+          % sigma''
+          \draw [cell,rounded corners] (19.5,-0.5) -- (22.5,-.5) -- (22.5,2.5) -- (19.5,2.5) -- cycle;
+          \draw[step=1cm,black!60,ultra thin,fill=black!10] (19.8,-0.2) grid (22.2,2.2);
+          \foreach \x/\y in {20/2,21/0,22/1} {
+              \draw [fill=black] (\x,\y) circle (0.1);
+          }
+          \node (sigmapp) at (21,3) {$\sigma''$};
+          \node[%
+            draw,
+            text width=1.75cm,text justified,color=black!50,rounded corners,
+            rectangle callout,callout relative pointer={(0.0cm,1.25cm)}
+          ] at (21,-4.5)
+          {%
+            $\sigma [2N_2 + N_4]$
+          };%
+          %\node at (21,-5) {$\sigma [2N_2 + N_4]$};
+          % edges sigma' - pi'
+          \draw [line width=1pt]
+          (1,20) .. controls +(0,9) and +(0,4) .. (6,15);
+          \draw [line width=1pt]
+          (5,21) .. controls +(0,9) and +(0,4) .. (13,16);
+          \draw [line width=1pt]
+          (2,19) .. controls +(0,-17) and +(0,-7) .. (8,13);
+          \draw [line width=1pt]
+          (3,17) .. controls +(0,-17) and +(0,-7) .. (9,10);
+          \draw [line width=1pt]
+          (4,18) .. controls +(0,-17) and +(0,-7) .. (11,11);
+          % edges pi' - pi''
+          \draw [line width=1pt]
+          (7,12) .. controls +(0,9) and +(0,4) .. (14,6);
+          \draw [line width=1pt]
+          (10,9) .. controls +(0,9) and +(0,4) .. (17,3);
+          \draw [line width=1pt]
+          (12,14) .. controls +(0,9) and +(0,4) .. (19,8);
+          % edges pi'' - sigma''
+          \draw [line width=1pt]
+          (15,7) .. controls +(0,-9) and +(0,-4) .. (20,2);
+          \draw [line width=1pt]
+          (16,4) .. controls +(0,-9) and +(0,-4) .. (21,0);
+          \draw [line width=1pt]
+          (18,5) .. controls +(0,-9) and +(0,-4) .. (22,1);
+          % text
+          % \node[%
+          %   draw,
+          %   text width=5.5cm,text justified,color=black!50,
+          %   rectangle callout,callout relative pointer={(0.5cm,-.75cm)}
+          % ] at (1,27)
+          % {%
+          % $2$ pairwise crossing $(\sigma',\pi')$-edges that
+          % link the guards
+          % };%
+          % \node[%
+          %   draw,
+          %   text width=5.5cm,text justified,color=black!50,
+          %   rectangle callout,callout relative pointer={(-0.1cm,-2.5cm)}
+          % ] at (17,24)
+          % {%
+          % $n-k$ pairwise crossing $(\sigma',\pi')$-edges that
+          % denote positions that do not correspond to
+          % a specific occurence of $\sigma$ in $\pi$
+          % (at positions $2$, $3$ and $5$).
+          % };%
+          % \node[%
+          %   draw,
+          %   text width=5.5cm,text justified,color=black!50,
+          %   rectangle callout,callout relative pointer={(0.75cm,1.5cm)}
+          % ] at (2.5,4,5)
+          % {%
+          % $k$ pairwise crossing $(\sigma',\pi')$-edges that
+          % denote a specific occurence of $\sigma$ in $\pi$
+          % (at positions $2$, $3$ and $5$).
+          % };%
+          % \node[%
+          %   draw,
+          %   text width=5.5cm,text justified,color=black!50,
+          %   rectangle callout,callout relative pointer={(2cm,.75cm)}
+          % ] at (7,-3)
+          % {%
+          % $k$ pairwise crossing $(\pi''; \sigma'')$-edges that
+          % denote a specific occurence of $\sigma$ in $\pi$
+          % (at positions $2$, $3$ and $5$).
+          % };%
+        \end{tikzpicture}
+        \caption{\label{fig:subfig:sigma' - pi' - pi'' - sigma''}%
+        $\sigma = 312$ and $\pi = 4\mathbf{5}\mathbf{2}1\mathbf{3}6$
+        (where a specific occurrence of $\sigma$ in $\pi$ is depicted in bold).
+        }
+      \end{figure}
+
+  More precisely,
+  (i) the first integer of $\sigma'$
+  (\emph{i.e.}, $(k+1) + (2N_2 + N_4 + 2n + k + 2)$) is linked
+  to the first integer of $\pi'$
+  ((\emph{i.e.}, $(n+1) + (2N_2 + N_4 + n + k + 2)$)),
+  (ii)
+
+  Conversely, suppose that there exists a containment-free perfect matching
+  $\mathcal{M}$ in $G_\mu$ such that for any two disctinct edges
+  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
+  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+  We show that $\sigma$ occurs as a pattern in $\pi$.
+  Whereas the matching $\mathcal{M}$ may not be as regular as in the forward
+  direction, the main idea is to prove that $\mathcal{M}$ contains enough structure
+  (more precisely, enough $(\sigma', \pi')$-edges) so that we can conclude.
+  We have divided the reverse direction into a serie of basic claims that
+  progressively defines the overall structure of $\mathcal{M}$.
+
+    \begin{figure}[t!]
+      \centering
+      \subfigure[%
+        A $(\nu_1, \nu_2)$-edge $(a, b)$ together with a
+        $(\nu_2, \nu'_1)$-edge $(x, y)$.
+        Notice that we may have $x<b$
+        (as long as $N_1 < x \leq N_1+N_2$ and $N_1 < b \leq N_1+N_2$
+        do hold).
+      ]{%
+        \begin{tikzpicture}
+          [
+            scale=0.25,
+            label/.style={anchor=base}
+          ]
+          % nu_1
+          \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
+          \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
+              \draw [fill=black] (\x,\y) circle (0.1);
+          }
+          % nu_4
+          \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
+          \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
+              \draw [fill=black] (\x,\y) circle (0.1);
+          }
+          % nu'_1
+          \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
+          \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
+              \draw [fill=black] (\x,\y) circle (0.1);
+          }
+          % labels
+          \node [label] (y) at (17,14.5) {$y$};
+          \node [label] (yp) at (10,-1.5) {$y'$};
+          \node [label] (a) at (0,8.5) {$a$};
+          \node [label] (ap) at (8,-1.5) {$a'$};
+          \node (nu1) at (2.5,7.7) {$\nu_1$};
+          \node (nu2) at (8.5,-2.5) {$\nu_2$};
+          \node (nu1') at (14.5,13.5) {$\nu'_1$};
+          % crossings edges
+          \draw [line width=1pt]
+          (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
+          \draw [line width=1pt]
+          (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
+          \draw [line width=1pt]
+          (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
+          \draw [line width=1pt]
+          (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
+          \draw [line width=1pt]
+          (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
+          % data edges
+          \draw [line width=1pt]
+          (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
+          \draw [line width=1pt]
+          (10,1) .. controls +(0,32) and +(0,10) .. (17,21);
+        \end{tikzpicture}
+        \label{subfig:no (nu_1, nu_2)-edge - 1}
+      }% end subfigure
+      \qquad
+      \subfigure[%
+        A $(\nu_1, \nu'_1)$-edge $(x', y')$ together with an
+        edge $(y, z)$ which is neither
+        a $(\nu_1, \nu'_1)$-edge,
+        nor a $(\nu_2, \nu'_1)$-edge,
+        nor a $(\nu'_1, \nu'_1)$-edge.
+        Notice that $y$ may be any position satisfying
+        $N_1+N_2 < y \leq 2N_1+N_2$.
+      ]{%
+      \begin{tikzpicture}
+        [
+          scale=0.25,
+          label/.style={anchor=base}
+        ]
+        % nu_1
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
+        \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % nu_4
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
+        \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % nu'_1
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
+        \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (y) at (17,14.5) {$y$};
+        \node [label] (ypp) at (23,-4.5) {$y''$};
+        \node [label] (a) at (0,8.5) {$a$};
+        \node [label] (a) at (2,8.5) {$x'$};
+        \node [label] (a) at (14,14.5) {$y'$};
+        \node [label] (ap) at (8,-1.5) {$a'$};
+        \node (nu1) at (2.5,7.7) {$\nu_1$};
+        \node (nu2) at (8.5,-2.5) {$\nu_2$};
+        \node (nu1') at (14.5,13.5) {$\nu'_1$};
+        % crossings edges
+        \draw [line width=1pt]
+        (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
+        \draw [line width=1pt]
+        (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
+        \draw [line width=1pt]
+        (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
+        \draw [line width=1pt]
+        (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
+        \draw [line width=1pt]
+        (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
+        % data edges
+        \draw [line width=1pt]
+        (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
+        \draw [line width=1pt]
+        (17,21) .. controls +(0,10) and +(0,35) .. (23,-3);
+        \end{tikzpicture}
+        \label{subfig:no (nu_1, nu_2)-edge - 2}
+      }% end subfigure
+      \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
+        Claim~\ref{claim:no (nu_1, nu_2)-edge}:
+        Assuming a $(\nu_1,\nu_2)$-edge $(a, b)$ in $\mathcal{M}$.
+      }
+    \end{figure}
+
+  \begin{claim}
+    \label{claim:no (nu_1, nu_2)-edge}
+    There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_2)-edge}]
+    First, according to Lemma~\ref{lemma:at most one edge monotone},
+    there exists at most one $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
+    Suppose now, aiming at a contradiction, that there exists
+    exactly one $(\nu_1, \nu_2)$-edge $e = (a, a')$ in $\mathcal{M}$.
+    Under this assumption, we show that $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
+    Indeed, suppose, aiming at a contradiction,
+    that there exists a $(\nu_1, \nu_1)$-edge $(x, x')$ in $\mathcal{M}$.
+    Since $\mathcal{M}$ is containment-free,
+    we must have $x < a$.
+    Then it follows that $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing) and
+    $\mu(x') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
+    which contradicts our hypothesis.
+    Hence $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
+    Furthermore, since $\nu_1$ is increasing and
+    there exists only one $(\nu_1, \nu_2)$-edge $(a, a')$ in $\mathcal{M}$,
+    we conclude that
+    $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
+    But $|\nu_1| = |\nu'_1| = N_1$ and $\mathcal{M}$ is perfect,
+    and hence there exists a position, say $y$, in $\nu'_1$ that is
+    not involved in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+    We rule out this configuration by considering two cases:
+    \begin{itemize}
+      \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-edge,
+      say $(y', y)$. Then, we have $\mu(a) > \mu(y')$ (since $\nu_1$ is above $\nu_2$)
+      and $\mu(a') < \mu(y)$ (since $\nu_2$ is below $\nu'_1$),
+      which contradicts our hypothesis.
+      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
+      \item Position $y$ is not involved in a $(\nu_2, \nu'_1)$-edge.
+      Since $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges
+      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge $(y, y'')$ in $\mathcal{M}$
+      that is neither a $(\nu_1, \nu'_1)$-edge nor a $(\nu_2, \nu'_1)$-edge
+      nor a $(\nu'_1, \nu'_1)$-edge.
+      In other words, $y'' > 2N_1 + N_2$.
+      Now, for any $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$, we
+      have $\mu(x) < \mu(y)$ (since $\nu_1$ is below $\nu'_1$)
+      and $\mu(x') > \mu(y'')$ (since $\nu'_1$ is above any other gadget),
+      which contradicts our hypothesis.
+      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 2}.
+    \end{itemize}
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:one (nu_1, nu'_1)-edge}
+    There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:one (nu_1, nu'_1)-edge}]
+    Suppose, aiming at a contradiction, that there is no
+    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+    Then it follows that there exists an edge $(x, y)$ in $\mathcal{M}$,
+    $1 \leq x \leq N_1$, that is neither a
+    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
+    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
+    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
+    (In other words, $x \leq N_1$ and $y > 2N_1 + N_2$.)
+    Therefore, since $\mathcal{M}$ is containment-free,
+    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
+    in $\mathcal{M}$.
+    Then it follows that every position $x$ of $\nu'_1$
+    (\emph{i.e.}, $N_1+N_2 < x \leq 2N_1+N_2$)
+    is involved in an edge $(x, y)$ of $\mathcal{M}$ with $y \geq 2N_1+N_2$.
+    But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
+    > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
+    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
+    Then it follows that $\mathcal{M}$ is not a perfect matching,
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_2, nu_2)-edge}
+    There is no $(\nu_2, \nu_2)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_2)-edge}]
+  Combine Claim~\ref{claim:one (nu_1, nu'_1)-edge} together with
+  the fact that $\mathcal{M}$ is containment-free.
+  \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_2, nu'_1)-edge}
+    There is no $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_1)-edge}]
+    We first prove that there exists at most one
+    $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
+    Indeed, suppose, aiming at a contradiction, that there exist
+    two $(\nu_2, \nu'_1)$-edges $e = (a, b)$ and $e' = (a', b')$ in $\mathcal{M}$.
+    Since $\mathcal{M}$ is containment-free,
+    we may assume that $a < a'$ and $b < b'$.
+    Then it follows that $\mu(a) > \mu(a')$
+    (since $\nu_2$ is decreasing) and $\mu(b) < \mu(b')$
+    (since $\nu'_1$ is increasing),
+    which contradicts our hypothesis.
+
+    Suppose now, still aiming at a contradiction, that there exists
+    exactly one $(\nu_2, \nu'_1)$-edge $e = (a, b)$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge}, there exists at least
+    one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+    Furtheremore, since
+    there is no $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge}),
+    every position in $\nu_1$ is involved either in a
+    $(\nu_1, \nu_1)$-edge or in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$
+    (otherwise $\mathcal{M}$ would not be containment-free).
+    Then it follows that that there is an odd number of $(\nu_1, \nu'_1)$-edges
+    in $\mathcal{M}$ since $\mathcal{M}$ is perfect and $|\nu_1| = N_1$ is odd.
+    Combining this with the fact that there exists exactly one $(\nu_2, \nu'_1)$-edge
+    in $\mathcal{M}$ and $|\nu'_1| = N_1$ is odd,
+    we conclude that there exists a position, say $x$, in $\nu'_1$
+    that is neither involved in a $(\nu_1, \nu'_1)$-edge nor a $(\nu_1, \nu'_1)$-edge
+    nor a $(\nu'_1, \nu'_1)$-edge.
+    Write $(x, y)$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $y > 2N_1 + N_2$, this edge.
+    We have $\mu(a) < \mu(x)$ (since $\nu_2$ is below $\nu'_1$)
+    and $\mu(b) > \mu(y)$ (since $\nu'_1$ is above any other gadget),
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:one (nu_2, nu'_2)-edge}
+    There is at least one $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:one (nu_2, nu'_2)-edge}]
+    Suppose, aiming at a contradiction there there is no
+    $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
+    Notice that there is neither
+    a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
+    nor a $(\nu_2, \nu_2)$-edge (Claim~\ref{claim:no (nu_2, nu_2)-edge})
+    nor a $(\nu_2, \nu'_1)$-edge (Claim~\ref{claim:no (nu_2, nu'_1)-edge})
+    in $\mathcal{M}$.
+    But $|\nu_2| = N_2 > 2N_3 + 2N_4 + 2n + 2k$.
+    Hence $\mathcal{M}$ is not a perfect matching,
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge}
+    There is
+    neither a $(\nu'_1, \nu'_1)$-edge,
+    nor a $(\nu_3, \nu_3)$-edge,
+    nor a $(\sigma', \sigma')$-edge,
+    nor a $(\nu_4, \nu_4)$-edge
+    in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge}]
+    Combine Claim~\ref{claim:one (nu_2, nu'_2)-edge} with the fact that
+    $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}).
+    \qed
+  \end{proof}
+
+  \begin{figure}[t!]
+    \centering
+      \begin{tikzpicture}
+        [
+          scale=0.25,
+          label/.style={anchor=base}
+        ]
+        % nu_2
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
+        \foreach \x/\y in {0/11,1/10,2/9,3/8,4/7,5/6} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % nu_3
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,15.8) grid (11.2,21.2);
+        \foreach \x/\y in {6/16,7/17,8/18,9/19,10/20,11/21} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % nu'_2
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,-0.2) grid (17.2,5.2);
+
+        \foreach \x/\y in {12/5,13/4,14/3,15/2,16/1,17/0} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (x) at (2,4.5) {$x$};
+        \node [label] (a) at (8,14.5) {$a$};
+        \node [label] (y) at (14,-1.5) {$y$};
+        \node [label] (b) at (16,-1.5) {$b$};
+        \node (nu1) at (2.5,2.5) {$\nu_2$};
+        \node (nu2) at (8.5,12.5) {$\nu_3$};
+        \node (nu1') at (14.5,-3.5) {$\nu'_2$};
+        % crossings edges
+        \draw [line width=1pt]
+        (2,9) .. controls +(0,25) and +(0,30) .. (14,3);
+        \draw [line width=1pt]
+        (8,18) .. controls +(0,15) and +(0,20) .. (16,1);
+      \end{tikzpicture}
+    \caption{\label{fig:subfig:no (nu_3, nu'_2)-edge}%
+      Claim~\ref{claim:no (nu_3, nu'_2)-edge}:
+      Assuming a $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
+    }
+  \end{figure}
+
+  \begin{claim}
+    \label{claim:no (nu_3, nu'_2)-edge}
+    There is no $(\nu_3, \nu'_2)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
+    Suppose, aiming at a contradiction, that there exists
+    $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge}, there
+    exists at least one $(\nu_2, \nu'_2)$-edge $(x, y)$ in $\mathcal{M}$.
+    Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
+    we have $y < b$.
+    Therefore,
+    we have $\mu(x) < \mu(a)$ (by construction) and
+    $\mu(y) > \mu(b)$ (since $\nu'_2$ is decreasing),
+    which contradicts our hypothesis.
+    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge}
+    There is neither a $(\nu_1, \nu'_3)$-edge nor a $(\nu'_1, \nu_3)$-edge
+    in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge}]
+    Suppose, aiming at a contradiction, that there exists a
+    $(\nu_1, \nu'_3)$-edge $(a, b)$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge} there exists
+    $(\nu_1, \nu'_1)$-edge $(x, y)$ in $\mathcal{M}$,
+    and ince $\mathcal{M}$ is containment-free we have $x < a$.
+    Therefore, we have
+    $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing)
+    and
+    $\mu(y) > \mu(b)$ (by construction),
+    which contradicts our hypothesis.
+    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
+
+    A similar argument shows that there is no
+    $(\nu'_1, \nu_3)$-edge in $\mathcal{M}$.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:at most one (nu_2, nu_3)-edge}
+    There is at most one $(\nu_2, \nu_3)$-edge
+    in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
+    Suppose, aiming at a contradiction, that there exist two
+    $(\nu_2, \nu_3)$-edges $(a, b)$ and $(a', b')$ in $\mathcal{M}$.
+    Since these two edges have to be crossing
+    (Proposition~\ref{proposition:matching}), we may assume that
+    $a < a'$ and $b < b'$.
+    Therefore, we have
+    $\mu(a) > \mu(a')$ (since $\nu_2$ is decreasing)
+    and
+    $\mu(b) < \mu(b')$ (since $\nu_3$ is increasing).
+    According to Proposition~\ref{proposition:matching}, this is
+    a contradiction.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:one (nu_3, nu'_3)-edge}
+    There is at least one $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:one (nu_3, nu'_3)-edge}]
+    Suppose, aiming at a contradiction, that there is no
+    $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:one (nu_4, nu'_4)-edge}
+    There is at least one $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:one (nu_4, nu'_4)-edge}]
+    Suppose, aiming at a contradiction, that there is no
+    $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_4, nu'_3)-edge}
+    There is no $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_4, nu'_3)-edge}]
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_4, pi')-edge}
+    There is no $(\nu_4, \pi')$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_4, pi')-edge}]
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:one (nu_4, nu'_4)-edge}
+    There is at least $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:one (nu_4, nu'_4)-edge}]
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (pi', pi'')-edge}
+    There is no $(\pi', \pi'')$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (pi', pi'')-edge}]
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu'_2, pi')-edge}
+    There is no $(\nu'_2, \pi')$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu'_2, pi')-edge}]
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu'_3, pi')-edge}
+    There is no $(\nu'_3, \pi')$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu'_3, pi')-edge}]
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_2, nu'_2)-edge}
+    There is no $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_2)-edge}]
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_3, nu'_3)-edge}
+    There is no $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_3)-edge}]
+    \qed
+  \end{proof}
+\qed
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% ---- Patterns ----
+%%
+\section{From square permutations to binary squares}
+\label{section:From square permutations to binary squares}
+
+It is conjectured in \cite{Henshall:Rampersad:Shallit:2011} that the number of binary strings
+of length $2n$ is given asymptoticaly by
+$\frac{\binom{2n}{n}}{n+1} 2^n - \binom{2n-1}{n+1} 2^{n-1} + O(2^{n-2})$.
+
+\begin{proposition}
+  \label{proposition:bijection binary - (213,231)-avoiding}
+  There exists a bijection between square binary strings of length $2n$ and
+  square $(213,231)$-avoiding permutations of length $2n$.
+\end{proposition}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% %%
+% %% ---- Patterns ----
+% %%
+% \section{Patterns}
+% \label{section:Patterns}
+%
+% \begin{proposition}
+%   \label{proposition:patterns:lower bound}
+%   Let $\pi \in S_{n}$.
+%   Then $\pi$ contains a pattern of size at least
+%   $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
+%   is the union of two order-isomorphic patterns.
+% \end{proposition}
+%
+% \begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
+%   The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935}
+%   states that every permutation of
+%   length at least $pq+1$ must contain either the pattern
+%   $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
+%   The result now follows from the fact that
+%   the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
+% \qed
+% \end{proof}
+%
+% \begin{proposition}
+%   \label{proposition:patterns:upper bound}
+%   If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
+%   a permutation of $S_{n}$
+%   that does not contains a pattern
+%   of length $2k$ that is the union of two
+%   order-isomorphic patterns.
+% \end{proposition}
+%
+% \begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
+% The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
+% Let $\pi \in S_n$ be chosen randomly and uniformly.
+% For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
+% of $\pi$,
+% let $X_\sigma$ be the indicator variable that
+% $\sigma$ is the union of two order-ismorphic patterns.
+% According to Boole's inequality, we have
+% $P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
+% Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
+% the probability that $\pi$ contains a pattern of length $2k$
+% is the union of two order-ismorphic patterns is at most
+% $\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
+% Thus, with positive probability, no event $X_\sigma$ occurs and
+% there is a permutation of $S_n$ that does not contain a pattern
+% of length $2k$ that is the union of two order-isomorphic patterns.
+% \qed
+% \end{proof}
+
+\bigskip
+\bigskip
 

### 2015-09-11 article.tex
reduction

@@
     $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-    Then it follows that there exists an edge $(x, y)$ in $\mathcal{M}$,
+    Then it follows that there exists an edge $(x, x')$ in $\mathcal{M}$,
     $1 \leq x \leq N_1$, that is neither a
@@
     nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
-    (In other words, $x \leq N_1$ and $y > 2N_1 + N_2$.)
+    (In other words, $x \leq N_1$ and $x' > 2N_1 + N_2$.)
     Therefore, since $\mathcal{M}$ is containment-free,
@@
     in $\mathcal{M}$.
-    Then it follows that every position $x$ of $\nu'_1$
-    (\emph{i.e.}, $N_1+N_2 < x \leq 2N_1+N_2$)
-    is involved in an edge $(x, y)$ of $\mathcal{M}$ with $y \geq 2N_1+N_2$.
+    Then it follows that every position $y$ of $\nu'_1$
+    (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
+    is involved in an edge $(y, y'$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
+    (otherwise $\mathcal{M}$ would not be containment-free).
     But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
@@
   \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_1)-edge}]
-    We first prove that there exists at most one
-    $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
-    Indeed, suppose, aiming at a contradiction, that there exist
-    two $(\nu_2, \nu'_1)$-edges $e = (a, b)$ and $e' = (a', b')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free,
-    we may assume that $a < a'$ and $b < b'$.
-    Then it follows that $\mu(a) > \mu(a')$
-    (since $\nu_2$ is decreasing) and $\mu(b) < \mu(b')$
-    (since $\nu'_1$ is increasing),
-    which contradicts our hypothesis.
-
-    Suppose now, still aiming at a contradiction, that there exists
-    exactly one $(\nu_2, \nu'_1)$-edge $e = (a, b)$ in $\mathcal{M}$.
+    First, according to Lemma~\ref{lemma:at most one edge monotone},
+    there exists at most one $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
+    Suppose now, aiming at a contradiction, that there exists
+    exactly one $(\nu_2, \nu'_1)$-edge $e = (a, a')$ in $\mathcal{M}$.
     According to Claim~\ref{claim:one (nu_1, nu'_1)-edge}, there exists at least
@@
     nor a $(\nu'_1, \nu'_1)$-edge.
-    Write $(x, y)$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $y > 2N_1 + N_2$, this edge.
+    Write $(x, x')$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $x' > 2N_1 + N_2$, this edge.
     We have $\mu(a) < \mu(x)$ (since $\nu_2$ is below $\nu'_1$)
-    and $\mu(b) > \mu(y)$ (since $\nu'_1$ is above any other gadget),
+    and $\mu(a') > \mu(x')$ (since $\nu'_1$ is above any other gadget),
     which contradicts our hypothesis.
@@
     in $\mathcal{M}$.
-    But $|\nu_2| = N_2 > 2N_3 + 2N_4 + 2n + 2k$.
+    But $|\nu_2| = N_2 > 2N_3 + 2N_4 + 2n + 2k +4$.
     Hence $\mathcal{M}$ is not a perfect matching,
@@
     Combine Claim~\ref{claim:one (nu_2, nu'_2)-edge} with the fact that
-    $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}).
+    $\mathcal{M}$ is containment-free.
     \qed
@@
         [
-          scale=0.25,
+          scale=0.35,
           label/.style={anchor=base}
@@
         \node [label] (a) at (8,14.5) {$a$};
-        \node [label] (y) at (14,-1.5) {$y$};
-        \node [label] (b) at (16,-1.5) {$b$};
+        \node [label] (xp) at (14,-1.5) {$x'$};
+        \node [label] (ap) at (16,-1.5) {$a'$};
         \node (nu1) at (2.5,2.5) {$\nu_2$};
@@
     Suppose, aiming at a contradiction, that there exists
-    $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
+    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
     According to Claim~\ref{claim:one (nu_2, nu'_2)-edge}, there
-    exists at least one $(\nu_2, \nu'_2)$-edge $(x, y)$ in $\mathcal{M}$.
+    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
     Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
-    we have $y < b$.
+    we have $x' < a'$.
     Therefore,
-    we have $\mu(x) < \mu(a)$ (by construction) and
-    $\mu(y) > \mu(b)$ (since $\nu'_2$ is decreasing),
+    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
+    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
     which contradicts our hypothesis.
@@
     \label{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge}
-    There is neither a $(\nu_1, \nu'_3)$-edge nor a $(\nu'_1, \nu_3)$-edge
+    There is neither a $(\nu_1, \nu_3)$-edge nor a $(\nu'_1, \nu_3)$-edge
     in $\mathcal{M}$.
@@
     Suppose, aiming at a contradiction, that there exists a
-    $(\nu_1, \nu'_3)$-edge $(a, b)$ in $\mathcal{M}$.
+    $(\nu_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
     According to Claim~\ref{claim:one (nu_1, nu'_1)-edge} there exists
-    $(\nu_1, \nu'_1)$-edge $(x, y)$ in $\mathcal{M}$,
-    and ince $\mathcal{M}$ is containment-free we have $x < a$.
+    $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$,
+    and since $\mathcal{M}$ is containment-free we have $x < a$.
     Therefore, we have
@@
     and
-    $\mu(y) > \mu(b)$ (by construction),
+    $\mu(x') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
     which contradicts our hypothesis.
@@
 
-    A similar argument shows that there is no
+    The same argument shows that there is no
     $(\nu'_1, \nu_3)$-edge in $\mathcal{M}$.
@@
   \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
-    Suppose, aiming at a contradiction, that there exist two
-    $(\nu_2, \nu_3)$-edges $(a, b)$ and $(a', b')$ in $\mathcal{M}$.
-    Since these two edges have to be crossing
-    (Proposition~\ref{proposition:matching}), we may assume that
-    $a < a'$ and $b < b'$.
-    Therefore, we have
-    $\mu(a) > \mu(a')$ (since $\nu_2$ is decreasing)
-    and
-    $\mu(b) < \mu(b')$ (since $\nu_3$ is increasing).
-    According to Proposition~\ref{proposition:matching}, this is
-    a contradiction.
+    Apply Lemma~\ref{lemma:at most one edge monotone}.
     \qed
@@
     $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
+    Combining Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge},
+    Claim~\ref{claim:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge},
+    Claim~\ref{claim:at most one (nu_2, nu_3)-edge},
+    and Claim~\ref{claim:no (nu_3, nu'_2)-edge},
+    we conclude that $N_3-1$ positions in $\nu_3$ are involved
+    in edges of $\mathcal{M}$ that are
+    neither ($\nu_1, \nu_3)$-edges,
+    nor $(\nu_2, \nu_3)$-edges
+    nor $(\nu'_1, \nu_3)$-edges
+    nor $(\nu_3, \nu_3)$-edges.
+    But $N_3 - 1 > 2N_4 + 2n + 2k + 4$ (the positions left) and hence
+    $\mathcal{M}$ is not a perfect matching,
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+  \begin{claim}
+    \label{claim:no (sigma', nu_4)-edge}
+    There is no $(\sigma', \nu_4)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (sigma', nu_4)-edge}]
+    Combine Claim~\ref{claim:one (nu_3, nu'_3)-edge} with the fact that
+    $\mathcal{M}$ is containment-free.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_4, nu'_3)-edge}
+    There is no $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_4, nu'_3)-edge}]
+    Suppose, aiming at a contradiction, that there exists
+    a $(\nu_4, \nu'_3)$-edge $(a, a')$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:one (nu_3, nu'_3)-edge},
+    there exists a $(\nu_3, \nu'_3)$-edge $(x, x')$
+    in $\mathcal{M}$.
+    Since $\mathcal{M}$ is containment-free, we must have
+    $x' < a'$.
+    Therefore,
+    we have $\mu(x) > \mu(a)$ (since $\nu_3$ is above $\nu_4$) and
+    $\mu(x') < \nu(a')$ (since $\nu'_3$ is increasing),
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_4, pi')-edge}
+    There is no $(\nu_4, \pi')$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_4, pi')-edge}]
     \qed
@@
     $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_4, nu'_3)-edge}
-    There is no $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_4, nu'_3)-edge}]
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_4, pi')-edge}
-    There is no $(\nu_4, \pi')$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_4, pi')-edge}]
     \qed

### 2015-09-11 article.tex
reduction

@@
   \begin{claim}
-    \label{claim:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge}
+    \label{claim:no edge below one (nu_2, nu_2)-edge}
     There is
     neither a $(\nu'_1, \nu'_1)$-edge,
+    nor a $(\nu'_1, \nu_3)$-edge,
+    nor a $(\nu'_1, \sigma')$-edge,
+    nor a $(\nu'_1, \nu_4)$-edge
     nor a $(\nu_3, \nu_3)$-edge,
+    nor a $(\nu_3, \sigma')$-edge,
+    nor a $(\nu_3, \nu_4)$-edge,
     nor a $(\sigma', \sigma')$-edge,
+    nor a $(\sigma', \nu_4)$-edge,
     nor a $(\nu_4, \nu_4)$-edge
@@
 
-  \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge}]
-    Combine Claim~\ref{claim:one (nu_2, nu'_2)-edge} with the fact that
+  \begin{proof}[of Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge}]
+    Combine Claim~\ref{claim:one (nu_2, nu'_2)-edge} together with the fact that
     $\mathcal{M}$ is containment-free.
+    \qed
+  \end{proof}
+
+  The next two claims state that $\mathcal{M}$ actually contains
+  all $(\nu_1, \nu'_1)$-edges and all $(\nu_2, \nu'_2)$-edges.
+
+  \begin{claim}
+    \label{claim: all (nu_1, nu'_1)-edges}
+    $\mathcal{M}$ contains $N_1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim: all (nu_1, nu'_1)-edges}]
+    Suppose, aiming at a contradiction, that
+    $\mathcal{M}$ does not contain $N_1$ $(\nu_1, \nu'_1)$-edge.
+    Combing Claim~\ref{claim:no (nu_2, nu'_1)-edge} and
+    Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge},
+    we conclude that there exists a position $a$ in $\nu'_1$ that is
+    involded in an edge $(a, a')$ in $\mathcal{M}$ that is
+    neither a $\nu_1, \nu'_1)$-edge,
+    nor a $(\nu_2, \nu'_1)$-edge,
+    nor a $(\nu'_1, \nu'_1)$-edge.
+    (In other words, $a' > 2N_1 + N_2$.)
+    Furthermore, according to Claim~\ref{claim:one (nu_1, nu'_1)-edge},
+    there exists a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+    Then, we have
+    $\mu(x) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
+    $\mu(x') < \mu(a')$ (since $\nu'_1$ is above any other gadget),
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim: all (nu_2, nu'_2)-edges}
+    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim: all (nu_2, nu'_2)-edges}]
+    Suppose, aiming at a contradiction, that
+    $\mathcal{M}$ does not contain $N_2$ $(\nu_2, \nu'_2)$-edge.
+    Combing Claim~\ref{claim:no (nu_1, nu_2)-edge} and
+    Claim~\ref{claim:one (nu_2, nu'_2)-edge},
+    we conclude that there exists a position $a$ in $\nu_2$ that is
+    involded in an edge $(a, a')$ in $\mathcal{M}$ that is
+    neither a $\nu_1, \nu_2)$-edge,
+    nor a $(\nu_2, \nu_2)$-edge.
+    Furthermore, according to Claim~\ref{claim:one (nu_2, nu'_2)-edge},
+    there exists a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+    Then, we have
+    $\mu(x) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
+    $\mu(x') < \mu(a')$ (since $\nu'_1$ is above any other gadget),
+    which contradicts our hypothesis.
     \qed
@@
     Combining Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge},
-    Claim~\ref{claim:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge},
+    Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge},
     Claim~\ref{claim:at most one (nu_2, nu_3)-edge},
@@
   \end{proof}
-  \begin{claim}
-    \label{claim:no (sigma', nu_4)-edge}
-    There is no $(\sigma', \nu_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (sigma', nu_4)-edge}]
-    Combine Claim~\ref{claim:one (nu_3, nu'_3)-edge} with the fact that
-    $\mathcal{M}$ is containment-free.
+
+  \begin{claim}
+    \label{claim:no (nu_1, nu_4)-edge}
+    There is no $(\nu_1, \nu_4)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_4)-edge}]
+    Suppose, aiming at a contradiction, that there exists
+    a $(\nu_1, \nu_4)$-edge $(a, b)$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge},
+    there exists a $(\nu_1, \nu'_1)$-edge $(x, x')$
+    in $\mathcal{M}$.
+    Since $\mathcal{M}$ is containment-free, we must have $x < a$.
+    Therefore, we have
+    $\mu(x) < \mu(a)$ (since $\nu_1$ is incresing)
+    and $\mu(x') > \mu(a')$ since $\nu'_1$ is above $\nu_4$),
+    which contradicts our hypothesis.
     \qed
@@
   \begin{claim}
-    \label{claim:no (nu_4, pi')-edge}
-    There is no $(\nu_4, \pi')$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_4, pi')-edge}]
+    \label{claim:no (nu_2, nu_4)-edge}
+    There is no $(\nu_2, \nu_4)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_4)-edge}]
+    Suppose, aiming at a contradiction, that there exists
+    a $(\nu_2, \nu_4)$-edge $(a, a')$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:one (nu_3, nu'_3)-edge},
+    there exists a $(\nu_3, \nu'_3)$-edge $(x, x')$
+    in $\mathcal{M}$.
+    Since $\mathcal{M}$ is containment-free, we must have
+    $x' < a'$.
+    Therefore,
+    we have $\mu(x) > \mu(a)$ (since $\nu_3$ is above $\nu_4$) and
+    $\mu(x') < \nu(a')$ (since $\nu'_3$ is increasing),
+    which contradicts our hypothesis.
     \qed
@@
     $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
+    Combining
+    Claim~\ref{claim:no (nu_1, nu_4), (nu'_1, nu_4)-edge}
+    Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge}
+    Claim~\ref{}
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no (nu_4, pi')-edge}
+    There is no $(\nu_4, \pi')$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_4, pi')-edge}]
     \qed

### 2015-09-12 article.tex
reduction

@@
     $\mathcal{M}$ does not contain $N_2$ $(\nu_2, \nu'_2)$-edge.
-    Combing Claim~\ref{claim:no (nu_1, nu_2)-edge} and
+    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge},
+    there exists a $(\nu_2; \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
+    Combing Claim~\ref{claim:no (nu_1, nu_2)-edge},
+    Claim~\ref{claim:no (nu_2, nu'_1)-edge}  and
     Claim~\ref{claim:one (nu_2, nu'_2)-edge},
-    we conclude that there exists a position $a$ in $\nu_2$ that is
-    involded in an edge $(a, a')$ in $\mathcal{M}$ that is
+    we conclude that there exists a position $x$ in $\nu_2$ that is
+    involded in an edge $(x, x')$ in $\mathcal{M}$ that is
     neither a $\nu_1, \nu_2)$-edge,
-    nor a $(\nu_2, \nu_2)$-edge.
+    nor a $(\nu_2, \nu_2)$-edge,
+    nor a $(\nu_2, \nu'_1)$-edge
+    nor a $(\nu_2, \nu'_2)$-edge (our hypothesis).
+    We need to consider two cases.
+    \begin{itemize}
+      \item
+      $(x, x')$ is
+      either a $(\nu_2, \nu'_3)$-edge
+      or a $(\nu_2, \pi')$-edge
+      or a $(\nu_2, \nu'_4)$-edge
+      or a $(\nu_2, \pi'')$-edge
+      or a $(\nu_2, \sigma'')$-edge.
+      Since $\mathcal{M}$ is containment-free,
+      we must have $a < x$.
+      Then it follows that
+      $\mu(a) > \mu(x)$ (since $\nu_2$ is decreasing) and
+      $\mu(a') < \mu(x')$
+      (since $\nu'_3$, $\pi'$, $\nu'_4$, $\pi''$ and $\sigma''$ are all above
+      $\nu'_2$),
+      which contradicts our hypothesis.
+      \item
+      $(x, x')$ is
+      either a $(\nu_2, \nu_3)$-edge
+      or a $(\nu_2, \sigma')$-edge
+      or a $(\nu_2, \nu_4)$-edge.
+    \end{itemize}
     Furthermore, according to Claim~\ref{claim:one (nu_2, nu'_2)-edge},

### 2015-09-12 article.tex
reduction

--- 2015-09-11 article.tex
@@
 \usepackage{graphicx}
-%\usepackage{llncsdoc}
+\usepackage{llncsdoc}
 \usepackage[english]{babel}
@@
   \begin{proof}[of Claim~\ref{claim:one (nu_2, nu'_2)-edge}]
-    Suppose, aiming at a contradiction there there is no
+    Suppose, aiming at a contradiction that there is no
     $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
@@
   \begin{claim}
-    \label{claim:no edge below one (nu_2, nu_2)-edge}
+    \label{claim:no edge below one (nu_2, nu'_2)-edge}
     There is
@@
 
-  \begin{proof}[of Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge}]
-    Combine Claim~\ref{claim:one (nu_2, nu'_2)-edge} together with the fact that
+  \begin{proof}[of Claim~\ref{claim:no edge below one (nu_2, nu'_2)-edge}]
+    Combine Claim~\ref{claim:one (nu_2, nu'_2)-edge} with the fact that
+    $\mathcal{M}$ is containment-free.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim: no edge right of nu'_2}
+    There is
+    neither a $(\nu'_2, \nu'_3)$-edge,
+    nor a $(\nu'_2, \pi')$-edge,
+    nor a $(\nu'_2, \nu'_4)$-edge,
+    nor a $(\nu'_2, \pi'')$-edge
+    nor a $(\nu'_2, \sigma'')$-edge
+    in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim: no edge right of nu'_2}]
+    Suppose, aiming at a contradiction, that
+    $\mathcal{M}$ contains an edge $(a, a')$ which is
+    either a $(\nu'_2, \nu'_3)$-edge
+    or a $(\nu'_2, \pi')$-edge
+    or a $(\nu'_2, \nu'_4)$-edge
+    or a $(\nu'_2, \pi'')$-edge
+    or a $(\nu'_2, \sigma'')$-edge.
+    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge},
+    there exists a $(\nu_2, \nu'_2)$-edge $(b, b')$ in $\mathcal{M}$.
+    Then it follows that we have
+    $\mu(b) > \mu(a)$ (since $\nu_2$ is above $\nu'_2$) and
+    $\mu(b') < \mu(a')$ (since $\nu'_3$, $\pi'$, $\nu'_4$, $\pi''$ and
+    $\sigma''$ are all above $\nu'_2$),
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:one (nu_1, nu'_2)-edge}
+    There is no $(\nu_1, \nu'_2)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:one (nu_1, nu'_2)-edge}]
+    Suppose, aiming at a contradiction that there is a
+    $(\nu_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge},
+    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
+    Since $\mathcal{M}$ is containment-free, we must have $b < a$.
+    Then it follows that we have
+    $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
+    $\mu(b') > \mu(a')$ (since $\nu'_2$ is above $\nu'_2$),
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:one (nu'_1, nu'_2)-edge}
+    There is no $(\nu'_1, \nu'_2)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:one (nu'_1, nu'_2)-edge}]
+    Suppose, aiming at a contradiction that there is a
+    $(\nu'_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge},
+    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
+    Then it follows that we have
+    $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
+    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu'_2$),
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  \begin{figure}[t!]
+    \centering
+      \begin{tikzpicture}
+        [
+          scale=0.35,
+          label/.style={anchor=base}
+        ]
+        % nu_2
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
+        \foreach \x/\y in {0/11,1/10,2/9,3/8,4/7,5/6} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % nu_3
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,15.8) grid (11.2,21.2);
+        \foreach \x/\y in {6/16,7/17,8/18,9/19,10/20,11/21} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % nu'_2
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,-0.2) grid (17.2,5.2);
+
+        \foreach \x/\y in {12/5,13/4,14/3,15/2,16/1,17/0} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (x) at (2,4.5) {$x$};
+        \node [label] (a) at (8,14.5) {$a$};
+        \node [label] (xp) at (14,-1.5) {$x'$};
+        \node [label] (ap) at (16,-1.5) {$a'$};
+        \node (nu1) at (2.5,2.5) {$\nu_2$};
+        \node (nu2) at (8.5,12.5) {$\nu_3$};
+        \node (nu1') at (14.5,-3.5) {$\nu'_2$};
+        % crossings edges
+        \draw [line width=1pt]
+        (2,9) .. controls +(0,25) and +(0,30) .. (14,3);
+        \draw [line width=1pt]
+        (8,18) .. controls +(0,15) and +(0,20) .. (16,1);
+      \end{tikzpicture}
+    \caption{\label{fig:subfig:no (nu_3, nu'_2)-edge}%
+      Claim~\ref{claim:no (nu_3, nu'_2)-edge}:
+      Assuming a $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
+    }
+  \end{figure}
+
+  \begin{claim}
+    \label{claim:no (nu_3, nu'_2)-edge}
+    There is no $(\nu_3, \nu'_2)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
+    Suppose, aiming at a contradiction, that there exists
+    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge}, there
+    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
+    Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
+    we have $x' < a'$.
+    Therefore,
+    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
+    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
+    which contradicts our hypothesis.
+    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:at most one (nu_2, nu_3)-edge}
+    There is at most one $(\nu_2, \nu_3)$-edge
+    in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
+    Apply Lemma~\ref{lemma:at most one edge monotone}.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:one (nu_3, nu'_3)-edge}
+    There is at least one $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:one (nu_3, nu'_3)-edge}]
+    Suppose, aiming at a contradiction, that there is no
+    $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
+    Combining Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge},
+    Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge},
+    Claim~\ref{claim:at most one (nu_2, nu_3)-edge},
+    and Claim~\ref{claim:no (nu_3, nu'_2)-edge},
+    we conclude that $N_3-1$ positions in $\nu_3$ are involved
+    in edges of $\mathcal{M}$ that are
+    neither ($\nu_1, \nu_3)$-edges,
+    nor $(\nu_2, \nu_3)$-edges
+    nor $(\nu'_1, \nu_3)$-edges
+    nor $(\nu_3, \nu_3)$-edges.
+    But $N_3 - 1 > 2N_4 + 2n + 2k + 4$ (the positions left) and hence
+    $\mathcal{M}$ is not a perfect matching,
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:no edge below one (nu_3, nu'_3)-edge}
+    There is
+    neither a $(\sigma', \sigma')$-edge,
+    nor a $(\sigma', \nu_4)$-edge,
+    nor a $(\sigma', \nu'_2)$-edge,
+    nor a $(\nu_4, \nu_4)$-edge
+    nor a $(\nu_4, \nu'_2)$-edge,
+    nor a $(\nu'_2, \nu'_2)$-edge
+    in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge}]
+    Combine Claim~\ref{claim:one (nu_3, nu'_3)-edge} with the fact that
     $\mathcal{M}$ is containment-free.
@@
   \begin{proof}[of Claim~\ref{claim: all (nu_2, nu'_2)-edges}]
-    Suppose, aiming at a contradiction, that
-    $\mathcal{M}$ does not contain $N_2$ $(\nu_2, \nu'_2)$-edge.
-    Combing Claim~\ref{claim:no (nu_1, nu_2)-edge} and
-    Claim~\ref{claim:one (nu_2, nu'_2)-edge},
-    we conclude that there exists a position $a$ in $\nu_2$ that is
-    involded in an edge $(a, a')$ in $\mathcal{M}$ that is
-    neither a $\nu_1, \nu_2)$-edge,
-    nor a $(\nu_2, \nu_2)$-edge.
-    Furthermore, according to Claim~\ref{claim:one (nu_2, nu'_2)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-    Then, we have
-    $\mu(x) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
-    $\mu(x') < \mu(a')$ (since $\nu'_1$ is above any other gadget),
-    which contradicts our hypothesis.
-    \qed
-  \end{proof}
-
-  \begin{figure}[t!]
-    \centering
-      \begin{tikzpicture}
-        [
-          scale=0.35,
-          label/.style={anchor=base}
-        ]
-        % nu_2
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
-        \foreach \x/\y in {0/11,1/10,2/9,3/8,4/7,5/6} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu_3
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,15.8) grid (11.2,21.2);
-        \foreach \x/\y in {6/16,7/17,8/18,9/19,10/20,11/21} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu'_2
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,-0.2) grid (17.2,5.2);
-
-        \foreach \x/\y in {12/5,13/4,14/3,15/2,16/1,17/0} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % labels
-        \node [label] (x) at (2,4.5) {$x$};
-        \node [label] (a) at (8,14.5) {$a$};
-        \node [label] (xp) at (14,-1.5) {$x'$};
-        \node [label] (ap) at (16,-1.5) {$a'$};
-        \node (nu1) at (2.5,2.5) {$\nu_2$};
-        \node (nu2) at (8.5,12.5) {$\nu_3$};
-        \node (nu1') at (14.5,-3.5) {$\nu'_2$};
-        % crossings edges
-        \draw [line width=1pt]
-        (2,9) .. controls +(0,25) and +(0,30) .. (14,3);
-        \draw [line width=1pt]
-        (8,18) .. controls +(0,15) and +(0,20) .. (16,1);
-      \end{tikzpicture}
-    \caption{\label{fig:subfig:no (nu_3, nu'_2)-edge}%
-      Claim~\ref{claim:no (nu_3, nu'_2)-edge}:
-      Assuming a $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
-    }
-  \end{figure}
-
-  \begin{claim}
-    \label{claim:no (nu_3, nu'_2)-edge}
-    There is no $(\nu_3, \nu'_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
-    Suppose, aiming at a contradiction, that there exists
-    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge}, there
-    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
-    we have $x' < a'$.
-    Therefore,
-    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
-    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
-    which contradicts our hypothesis.
-    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
+    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge},
+    Claim~\ref{}
+    all positions in $\nu'_2$ are involved in a $(\nu_2, \nu'_2)$.
     \qed
@@
   \begin{claim}
-    \label{claim:at most one (nu_2, nu_3)-edge}
-    There is at most one $(\nu_2, \nu_3)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
-    Apply Lemma~\ref{lemma:at most one edge monotone}.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:one (nu_3, nu'_3)-edge}
-    There is at least one $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:one (nu_3, nu'_3)-edge}]
-    Suppose, aiming at a contradiction, that there is no
-    $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
-    Combining Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge},
-    Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge},
-    Claim~\ref{claim:at most one (nu_2, nu_3)-edge},
-    and Claim~\ref{claim:no (nu_3, nu'_2)-edge},
-    we conclude that $N_3-1$ positions in $\nu_3$ are involved
-    in edges of $\mathcal{M}$ that are
-    neither ($\nu_1, \nu_3)$-edges,
-    nor $(\nu_2, \nu_3)$-edges
-    nor $(\nu'_1, \nu_3)$-edges
-    nor $(\nu_3, \nu_3)$-edges.
-    But $N_3 - 1 > 2N_4 + 2n + 2k + 4$ (the positions left) and hence
-    $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis.
-    \qed
-  \end{proof}
-
-  \begin{claim}
     \label{claim:no (nu_1, nu_4)-edge}
@@
   \begin{claim}
-    \label{claim:one (nu_4, nu'_4)-edge}
-    There is at least $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:one (nu_4, nu'_4)-edge}]
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (pi', pi'')-edge}
-    There is no $(\pi', \pi'')$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (pi', pi'')-edge}]
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu'_2, pi')-edge}
-    There is no $(\nu'_2, \pi')$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu'_2, pi')-edge}]
+    \label{claim:no (pi', pi)-edge}
+    There is no $(\pi', \pi')$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (pi', pi')-edge}]
+    Combine Claim~\ref{claim:one (nu_4, nu'_4)-edge} together with the
+    fact that $\mathcal{M}$ is containment-free.
     \qed
@@
   \begin{proof}[of Claim~\ref{claim:no (nu'_3, pi')-edge}]
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_2, nu'_2)-edge}
-    There is no $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_2)-edge}]
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_3, nu'_3)-edge}
-    There is no $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_3)-edge}]
-    \qed
-  \end{proof}
+    Suppose aiming at a contradiction that there exists
+    a $(\nu'_3, \pi')$-edge $(a, a')$ in $\mathcal{M}$.
+    Accordint to Claim~\ref{claim:one (nu_3, nu'_3)-edge},
+    there exists a $(\nu_3, \nu'_3)$-edge $(b, b')$ in $\mathcal{M}$.
+    Then it follows that we have
+    $\mu(b) < \mu(a)$ (since $\nu_3$ is below $\nu'_3$) and
+    $\mu(b') > \mu(a')$ (since $\nu'_3$ is above $\pi'$),
+    which contradicts our hypothesis.
+    \qed
+  \end{proof}
+
+  Combinf the above claims
 \qed

### 2015-09-12 article.tex
New term in enumeration of square classes + some very slight modifications

@@
 \begin{proof}
+    [of Proposition~\ref{prop:bijection_binary_to_permutations_squares}]
     The statement of the proposition is a consequence of the
@@
 \end{proposition}
-\begin{proof}
+\begin{proof}[of Proposition~\ref{prop:shuffle_associative_commutative}]
     To prove the associativity of $\SHUFFLE$, it is convenient to show
@@
 \end{proposition}
-\begin{proof}
+\begin{proof}[of Proposition~\ref{prop:endomorphisms}]
     To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
@@
 \end{proposition}
-\begin{proof}
+\begin{proof}[of Proposition~\ref{prop:square_stability}]
     All statements~\ref{item:square_stability_1},
@@
 \begin{equation}
-    1, 0, 1, 0, 6, 0, 81, 0, 2774.
+    1, 0, 1, 0, 6, 0, 81, 0, 2774, 0, 162945.
 \end{equation}

### 2015-09-12 article.tex
reduction

--- 2015-09-12 article.tex
@@
   \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_2)-edge}]
-  Combine Claim~\ref{claim:one (nu_1, nu'_1)-edge} together with
+  Combine Claim~\ref{claim:one (nu_1, nu'_1)-edge} with
   the fact that $\mathcal{M}$ is containment-free.
@@
   \begin{claim}
-    \label{claim:one (nu_1, nu'_2)-edge}
+    \label{claim:no (nu_1, nu'_2)-edge}
     There is no $(\nu_1, \nu'_2)$-edge in $\mathcal{M}$.
@@
 
-  \begin{proof}[of Claim~\ref{claim:one (nu_1, nu'_2)-edge}]
+  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu'_2)-edge}]
     Suppose, aiming at a contradiction that there is a
@@
   \begin{claim}
-    \label{claim:one (nu'_1, nu'_2)-edge}
+    \label{claim:no (nu'_1, nu'_2)-edge}
     There is no $(\nu'_1, \nu'_2)$-edge in $\mathcal{M}$.
@@
 
-  \begin{proof}[of Claim~\ref{claim:one (nu'_1, nu'_2)-edge}]
+  \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_2)-edge}]
     Suppose, aiming at a contradiction that there is a
@@
     Combing Claim~\ref{claim:no (nu_2, nu'_1)-edge} and
-    Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge},
+    Claim~\ref{claim:no edge below one (nu_2, nu'_2)-edge},
     we conclude that there exists a position $a$ in $\nu'_1$ that is
     involded in an edge $(a, a')$ in $\mathcal{M}$ that is
-    neither a $\nu_1, \nu'_1)$-edge,
+    neither a $(\nu_1, \nu'_1)$-edge,
     nor a $(\nu_2, \nu'_1)$-edge,
@@
   \begin{claim}
-    \label{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge}
+    \label{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge no (nu_1, nu_4)-edge}
     There is neither a $(\nu_1, \nu_3)$-edge nor a $(\nu'_1, \nu_3)$-edge
+    nor a $(\nu_1, \nu_4)$-edge
     in $\mathcal{M}$.
@@
 
-  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge}]
-    Suppose, aiming at a contradiction, that there exists a
-    $(\nu_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge} there exists
-    $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$,
-    and since $\mathcal{M}$ is containment-free we have $x < a$.
-    Therefore, we have
-    $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing)
-    and
-    $\mu(x') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
-    which contradicts our hypothesis.
-    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
-
-    The same argument shows that there is no
-    $(\nu'_1, \nu_3)$-edge in $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_1, nu_4)-edge}
-    There is no $(\nu_1, \nu_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_4)-edge}]
-    Suppose, aiming at a contradiction, that there exists
-    a $(\nu_1, \nu_4)$-edge $(a, b)$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge $(x, x')$
-    in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free, we must have $x < a$.
-    Therefore, we have
-    $\mu(x) < \mu(a)$ (since $\nu_1$ is incresing)
-    and $\mu(x') > \mu(a')$ since $\nu'_1$ is above $\nu_4$),
-    which contradicts our hypothesis.
+  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge no (nu_1, nu_4)-edge}]
+    According to Claim~\ref{claim: all (nu_1, nu'_1)-edges},
+    all positions in $\nu_1$ and $\nu'_1$ are involved in
+    $(\nu_1, \nu'_1)$-edges in $\mathcal{M}$.
     \qed
@@
   \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_4)-edge}]
-    Suppose, aiming at a contradiction, that there exists
-    a $(\nu_2, \nu_4)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:one (nu_3, nu'_3)-edge},
-    there exists a $(\nu_3, \nu'_3)$-edge $(x, x')$
-    in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free, we must have
-    $x' < a'$.
-    Therefore,
-    we have $\mu(x) > \mu(a)$ (since $\nu_3$ is above $\nu_4$) and
-    $\mu(x') < \nu(a')$ (since $\nu'_3$ is increasing),
-    which contradicts our hypothesis.
+    According to Claim~\ref{claim: all (nu_2, nu'_2)-edges},
+    all positions in $\nu_2$ and $\nu'_2$ are involved in
+    $(\nu_2, \nu'_2)$-edges in $\mathcal{M}$.
     \qed
@@
   \begin{claim}
-    \label{claim:no (nu_4, pi')-edge}
-    There is no $(\nu_4, \pi')$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_4, pi')-edge}]
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (pi', pi)-edge}
-    There is no $(\pi', \pi')$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (pi', pi')-edge}]
-    Combine Claim~\ref{claim:one (nu_4, nu'_4)-edge} together with the
+    \label{claim:no (pi', pi)-edge no (sigma', pi'')-edge no (sigma',sigma'')-edge}
+    There is
+    neither a $(\pi', \pi')$-edge
+    nor a $(\sigma', \pi'')$-edge
+    nor a $(\sigma', \sigma'')$-edge
+    in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (pi', pi)-edge no (sigma', pi'')-edge no (sigma',sigma'')-edge}]
+    Combine Claim~\ref{claim:one (nu_4, nu'_4)-edge} with the
     fact that $\mathcal{M}$ is containment-free.
@@
 
-  \begin{claim}
-    \label{claim:no (nu'_3, pi')-edge}
-    There is no $(\nu'_3, \pi')$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu'_3, pi')-edge}]
-    Suppose aiming at a contradiction that there exists
-    a $(\nu'_3, \pi')$-edge $(a, a')$ in $\mathcal{M}$.
-    Accordint to Claim~\ref{claim:one (nu_3, nu'_3)-edge},
-    there exists a $(\nu_3, \nu'_3)$-edge $(b, b')$ in $\mathcal{M}$.
-    Then it follows that we have
-    $\mu(b) < \mu(a)$ (since $\nu_3$ is below $\nu'_3$) and
-    $\mu(b') > \mu(a')$ (since $\nu'_3$ is above $\pi'$),
-    which contradicts our hypothesis.
-    \qed
-  \end{proof}
-
-  Combinf the above claims
+  Then it follows that that every position in $\sigma$ is involved in
+  either a $(\sigma', \nu'_3)$-edge
+  or a $(\sigma', \pi')$-edge
+  or a $(\sigma', \nu'_4)$-edge
+  in $\mathcal{M}$.
   \qed

### 2015-09-12 article.tex
notations

@@
 
-\documentclass[a4paper,10pt]{llncs}
+\documentclass[a4paper]{llncs}
 
@@
 \input{complexity}
+\input{graph}
 
@@
 such that $u$ belongs to $v \shuffle v$.
-\medskip
 
@@
 of $\sigma$, $\pi$ {\em avoids} $\sigma$.
-\medskip
+
+For a graph $G$, we denote $\VS{G}$ as the set of vertices and
+$\ES{G}$ as the set of edges.
+Let $u = u_1 \, u_2 \,\ldots\, u_n$ be a word.
+The \emph{clique graph associated to $u$}, denoted $K_u$,
+is the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$.
+The structure of this underlying graph is linear,
+\emph{i.e.}, the set of vertices is equipped with a natural total order
+$<$ defined by $u_i < u_j$ if and only if $i < j$.
+(For a linear graph, we write $(u_i, u_j)$ instead of $\{u_i, u_j\}$
+for an edge if it is clear from the context that $i < j$.)
+Recall that two edges of a graph are \emph{independent} if they do not share
+a common vertex, and that a \emph{matching} $\mathcal{M}$ in $G$ is a set of
+pairwise independent edges.
+A matching is \emph{perfect} if it covers all the vertices of the graph.
+In case the set of vertices is equipped with a total order,
+a matching $\mathcal{M}$ is said to be \emph{containment-free}
+if there do not exist (independent) edges $(u_i, u_j)$ and
+$(u_k, u_\ell)$ in $\mathcal{M}$ such that
+$u_i < u_k < u_\ell < u_j$ or  $u_k < u_i < u_j < u_\ell$.
 

### 2015-09-14 article.tex
notations

@@
   $\pi(a') > \pi(b')$ (since $\sigma_2$ is decreasing),
-  which contradicts our hypothesis.
+  which contradicts our hypothesis about $\mathcal{M}$.
 
@@
   $\pi(a') < \pi(b')$ (since $\sigma_2$ is increasing),
-  which contradicts our hypothesis.
+  which contradicts our hypothesis about $\mathcal{M}$.
   \qed
@@
       \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
-      \node (claim edge 1) at (1.5,13) {Claim\ref{}};
+      \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
+      \node (claim edge 1) at (1.5,13) {Claim~\ref{claim:all (nu_1, nu'_1)-edges}};
       % 2
       \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
-      \node (claim edge 1) at (4,3.5) {Claim\ref{}};
+      \node (claim edge 2) at (4,4) {Claim\ref{claim:(nu_2, nu'_2)-edge}};
+      \node (claim edge 2) at (4,3.5) {Claim\ref{claim:all (nu_2, nu'_2)-edges}};
       % 3
       \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
-      \node (claim edge 1) at (5.5,11.5) {Claim\ref{}};
+      \node (claim edge 1) at (5.5,11.5) {Claim\ref{claim:(nu_3, nu'_3)-edge}};
       % 4
@@
     $\mu(x') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
-    which contradicts our hypothesis.
+    which contradicts our hypothesis about $\mathcal{M}$.
     Hence $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
@@
       and $\mu(a') < \mu(y)$ (since $\nu_2$ is below $\nu'_1$),
-      which contradicts our hypothesis.
+      which contradicts our hypothesis about $\mathcal{M}$.
       See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
@@
       and $\mu(x') > \mu(y'')$ (since $\nu'_1$ is above any other gadget),
-      which contradicts our hypothesis.
+      which contradicts our hypothesis about $\mathcal{M}$.
       See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 2}.
@@
   \begin{claim}
-    \label{claim:one (nu_1, nu'_1)-edge}
+    \label{claim:(nu_1, nu'_1)-edge}
     There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
@@
 
-  \begin{proof}[of Claim~\ref{claim:one (nu_1, nu'_1)-edge}]
+  \begin{proof}[of Claim~\ref{claim:(nu_1, nu'_1)-edge}]
     Suppose, aiming at a contradiction, that there is no
@@
     Then it follows that $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis.
-    \qed
-  \end{proof}
+    which contradicts our hypothesis about $\mathcal{M}$.
+    \qed
+  \end{proof}
+
+  The above claim will be complement in upcoming Claim~\ref{claim: all (nu_1, nu'_1)-edges}.
 
@@
   \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_2)-edge}]
-  Combine Claim~\ref{claim:one (nu_1, nu'_1)-edge} with
+  Combine Claim~\ref{claim:(nu_1, nu'_1)-edge} with
   the fact that $\mathcal{M}$ is containment-free.
@@
     exactly one $(\nu_2, \nu'_1)$-edge $e = (a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge}, there exists at least
+    According to Claim~\ref{claim:(nu_1, nu'_1)-edge}, there exists at least
     one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
@@
     and $\mu(a') > \mu(x')$ (since $\nu'_1$ is above any other gadget),
-    which contradicts our hypothesis.
+    which contradicts our hypothesis about $\mathcal{M}$.
     \qed
@@
   \begin{claim}
-    \label{claim:one (nu_2, nu'_2)-edge}
+    \label{claim:(nu_2, nu'_2)-edge}
     There is at least one $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
@@
 
-  \begin{proof}[of Claim~\ref{claim:one (nu_2, nu'_2)-edge}]
+  \begin{proof}[of Claim~\ref{claim:(nu_2, nu'_2)-edge}]
     Suppose, aiming at a contradiction that there is no
@@
     Hence $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis.
+    which contradicts our hypothesis about $\mathcal{M}$.
     \qed
@@
   \begin{claim}
-    \label{claim:no edge below one (nu_2, nu'_2)-edge}
+    \label{claim:no edge below (nu_2, nu'_2)-edge}
     There is
@@
 
-  \begin{proof}[of Claim~\ref{claim:no edge below one (nu_2, nu'_2)-edge}]
-    Combine Claim~\ref{claim:one (nu_2, nu'_2)-edge} with the fact that
+  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
+    Combine Claim~\ref{claim:(nu_2, nu'_2)-edge} with the fact that
     $\mathcal{M}$ is containment-free.
@@
     or a $(\nu'_2, \sigma'')$-edge.
-    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge},
+    According to Claim~\ref{claim:(nu_2, nu'_2)-edge},
     there exists a $(\nu_2, \nu'_2)$-edge $(b, b')$ in $\mathcal{M}$.
@@
     $\sigma''$ are all above $\nu'_2$),
-    which contradicts our hypothesis.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_1, nu'_2)-edge}
-    There is no $(\nu_1, \nu'_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu'_2)-edge}]
-    Suppose, aiming at a contradiction that there is a
-    $(\nu_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free, we must have $b < a$.
-    Then it follows that we have
-    $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
-    $\mu(b') > \mu(a')$ (since $\nu'_2$ is above $\nu'_2$),
-    which contradicts our hypothesis.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu'_1, nu'_2)-edge}
-    There is no $(\nu'_1, \nu'_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_2)-edge}]
-    Suppose, aiming at a contradiction that there is a
-    $(\nu'_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:one (nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-    Then it follows that we have
-    $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
-    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu'_2$),
-    which contradicts our hypothesis.
-    \qed
-  \end{proof}
+    which contradicts our hypothesis about $\mathcal{M}$.
+    \qed
+  \end{proof}
+
+  % \begin{claim}
+  %   \label{claim:no (nu_1, nu'_2)-edge}
+  %   There is no $(\nu_1, \nu'_2)$-edge in $\mathcal{M}$.
+  % \end{claim}
+  %
+  % \begin{proof}[of Claim~\ref{claim:no (nu_1, nu'_2)-edge}]
+  %   Suppose, aiming at a contradiction that there is a
+  %   $(\nu_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
+  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
+  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
+  %   Since $\mathcal{M}$ is containment-free, we must have $b < a$.
+  %   Then it follows that we have
+  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
+  %   $\mu(b') > \mu(a')$ (since $\nu'_2$ is above $\nu'_2$),
+  %   which contradicts our hypothesis about $\mathcal{M}$.
+  %   \qed
+  % \end{proof}
+  %
+  % \begin{claim}
+  %   \label{claim:no (nu'_1, nu'_2)-edge}
+  %   There is no $(\nu'_1, \nu'_2)$-edge in $\mathcal{M}$.
+  % \end{claim}
+  %
+  % \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_2)-edge}]
+  %   Suppose, aiming at a contradiction that there is a
+  %   $(\nu'_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
+  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
+  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
+  %   Then it follows that we have
+  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
+  %   $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu'_2$),
+  %   which contradicts our hypothesis about $\mathcal{M}$.
+  %   \qed
+  % \end{proof}
 
@@
     $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge}, there
+    According to Claim~\ref{claim:(nu_2, nu'_2)-edge}, there
     exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
@@
     $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
-    which contradicts our hypothesis.
+    which contradicts our hypothesis about $\mathcal{M}$.
     See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
@@
 
+    We will see soon (upcoming Claim~\ref{claim:all (nu_2, nu'_2)-edges})
+    that there exists actually no $(\nu_2, \nu_3)$-edge
+    in $\mathcal{M}$.
+
   \begin{claim}
-    \label{claim:one (nu_3, nu'_3)-edge}
-    There is at least one $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:one (nu_3, nu'_3)-edge}]
+    \label{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}
+    There is neither a $(\nu_1, \nu_3)$-edge not a $(\nu'_1, \nu_3)$-edge
+    in $\mathcal{M}$.
+  \end{claim}
+
+ \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}]
+    We first prove that there is no $(\nu_1, \nu_3)$-edge in $\mathcal{M}$.
+    Indeed, suppose, aiming at a contradiction, that these exists
+    a $(\nu_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
+    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
+    Since $\mathcal{M}$ is containment-free, we must have
+    $b < a'$.
+    Then it follows that
+    $\mu(b) < \mu(a)$ (since $\nu_1$ is incresing) and
+    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
+    which contradicts our hypothesis about $\mathcal{M}$.
+
+    As for no $(\nu_1, \nu_3)$-edges in $\mathcal{M}$, suppose,
+    aiming at a contradiction, that these exists
+    a $(\nu'_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
+    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
+    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
+    Then it follows that
+    $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
+    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
+    which contradicts our hypothesis about $\mathcal{M}$.
+    \qed
+  \end{proof}
+
+  \begin{claim}
+    \label{claim:(nu_3, nu'_3)-edge}
+    There exists a $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:(nu_3, nu'_3)-edge}]
     Suppose, aiming at a contradiction, that there is no
     $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
-    Combining Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge},
-    Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge},
-    Claim~\ref{claim:at most one (nu_2, nu_3)-edge},
-    and Claim~\ref{claim:no (nu_3, nu'_2)-edge},
+    Combining
+    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
+    Claim~\ref{claim:no (nu_3, nu'_2)-edge},
+    Claim~\ref{claim:at most one (nu_2, nu_3)-edge}
+    and Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge},
     we conclude that $N_3-1$ positions in $\nu_3$ are involved
@@
     $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis.
+    which contradicts our hypothesis about $\mathcal{M}$.
     \qed
@@
   \begin{proof}[of Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge}]
-    Combine Claim~\ref{claim:one (nu_3, nu'_3)-edge} with the fact that
+    Combine Claim~\ref{claim:(nu_3, nu'_3)-edge} with the fact that
     $\mathcal{M}$ is containment-free.
@@
   \begin{claim}
-    \label{claim: all (nu_1, nu'_1)-edges}
+    \label{claim:all (nu_1, nu'_1)-edges}
     $\mathcal{M}$ contains $N_1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
@@
 
-  \begin{proof}[of Claim~\ref{claim: all (nu_1, nu'_1)-edges}]
+  \begin{proof}[of Claim~\ref{claim:all (nu_1, nu'_1)-edges}]
     Suppose, aiming at a contradiction, that
@@
     Combing Claim~\ref{claim:no (nu_2, nu'_1)-edge} and
-    Claim~\ref{claim:no edge below one (nu_2, nu'_2)-edge},
+    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
     we conclude that there exists a position $a$ in $\nu'_1$ that is
@@
     (In other words, $a' > 2N_1 + N_2$.)
-    Furthermore, according to Claim~\ref{claim:one (nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+    Furthermore, according to Claim~\ref{claim:(nu_1, nu'_1)-edge},
+    there exists a $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$.
     Then, we have
@@
     $\mu(x') < \mu(a')$ (since $\nu'_1$ is above any other gadget),
-    which contradicts our hypothesis.
+    which contradicts our hypothesis about $\mathcal{M}$.
     \qed
@@
   \begin{claim}
-    \label{claim: all (nu_2, nu'_2)-edges}
+    \label{claim:all (nu_2, nu'_2)-edges}
     $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
@@
 
-  \begin{proof}[of Claim~\ref{claim: all (nu_2, nu'_2)-edges}]
-    According to Claim~\ref{claim:one (nu_2, nu'_2)-edge},
-    Claim~\ref{}
-    all positions in $\nu'_2$ are involved in a $(\nu_2, \nu'_2)$.
+  \begin{proof}[of Claim~\ref{claim:all (nu_2, nu'_2)-edges}]
+    The key idea is to focus on $\nu'_2$.
+    Combining Claim~\ref{claim:no (nu_3, nu'_2)-edge},
+    Claim~\ref{claim: no edge right of nu'_2} and
+    Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge},
+    we obtain that every position in $\nu'_2$ is involved in
+    either a $(\nu_1, \nu'_2$)-edge
+    or a $(\nu'_1, \nu'_2$)-edge
+    or $(\nu_2, \nu'_2$)-edge.
+    But, according to Claim~\ref{claim:all (nu_1, nu'_1)-edges},
+    all positions in $\nu_1$ and $\nu'_1$ are involved in
+    $(\nu_1, \nu'_1)$-edges.
+    Then it follows that
+    all positions in $\nu'_2$ are involved in a $(\nu_2, \nu'_2)$,
+    and hence
+    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
     \qed
@@
   \begin{claim}
-    \label{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge no (nu_1, nu_4)-edge}
-    There is neither a $(\nu_1, \nu_3)$-edge nor a $(\nu'_1, \nu_3)$-edge
-    nor a $(\nu_1, \nu_4)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge no (nu'_1, nu_3)-edge no (nu_1, nu_4)-edge}]
-    According to Claim~\ref{claim: all (nu_1, nu'_1)-edges},
-    all positions in $\nu_1$ and $\nu'_1$ are involved in
-    $(\nu_1, \nu'_1)$-edges in $\mathcal{M}$.
+    \label{claim:no (nu_2, nu_4)-edge}
+    There is no $(\nu_2, \nu_4)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_4)-edge}]
+    According to Claim~\ref{claim: all (nu_2, nu'_2)-edges},
+    all positions in $\nu_2$ and $\nu'_2$ are involved in
+    $(\nu_2, \nu'_2)$-edges in $\mathcal{M}$.
     \qed
@@
     a $(\nu_4, \nu'_3)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:one (nu_3, nu'_3)-edge},
+    According to Claim~\ref{claim:(nu_3, nu'_3)-edge},
     there exists a $(\nu_3, \nu'_3)$-edge $(x, x')$
@@
     $\mu(x') < \nu(a')$ (since $\nu'_3$ is increasing),
-    which contradicts our hypothesis.
+    which contradicts our hypothesis about $\mathcal{M}$.
     \qed
@@
   \begin{claim}
-    \label{claim:no (nu_2, nu_4)-edge}
-    There is no $(\nu_2, \nu_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_4)-edge}]
-    According to Claim~\ref{claim: all (nu_2, nu'_2)-edges},
-    all positions in $\nu_2$ and $\nu'_2$ are involved in
-    $(\nu_2, \nu'_2)$-edges in $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:one (nu_4, nu'_4)-edge}
+    \label{claim:(nu_4, nu'_4)-edge}
     There is at least one $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
@@
 
-  \begin{proof}[of Claim~\ref{claim:one (nu_4, nu'_4)-edge}]
+  \begin{proof}[of Claim~\ref{claim:(nu_4, nu'_4)-edge}]
     Suppose, aiming at a contradiction, that there is no
     $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
-    Combining
-    Claim~\ref{claim:no (nu_1, nu_4), (nu'_1, nu_4)-edge}
-    Claim~\ref{claim:no edge below one (nu_2, nu_2)-edge}
-    Claim~\ref{}
+    First, according to
+    Claim~\ref{claim:all (nu_1, nu'_1)-edges} and
+    Claim~\ref{claim:all (nu_2, nu'_2)-edges},
+    there is neither a $(\nu_1, \nu_4)$-edge
+    nor a $(\nu'_1, \nu_4)$-edge
+    nor a $(\nu_2, \nu_4)$-edge
+    nor a $(\nu_4, \nu'_2)$-edge in $\mathcal{M}$.
+    Furthermore, according to
+    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge} and
+    Claim~\ref{claim:no (nu_4, nu'_3)-edge},
+    there is neither a $(\nu_3, \nu_4)$-edge
+    nor a $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
+    But $N_4 > 2n + 2k + 2$, and
+    hence $\mathcal{M}$ is not perfect,
+    which contradicts our hypothesis about $\mathcal{M}$.
     \qed
@@
   \begin{claim}
-    \label{claim:no (pi', pi)-edge no (sigma', pi'')-edge no (sigma',sigma'')-edge}
+    \label{claim:no edge below (nu_4, nu'_4)-edge}
     There is
@@
 
-  \begin{proof}[of Claim~\ref{claim:no (pi', pi)-edge no (sigma', pi'')-edge no (sigma',sigma'')-edge}]
-    Combine Claim~\ref{claim:one (nu_4, nu'_4)-edge} with the
+  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
+    Combine Claim~\ref{claim:(nu_4, nu'_4)-edge} with the
     fact that $\mathcal{M}$ is containment-free.
@@
 
-  Then it follows that that every position in $\sigma$ is involved in
+  \begin{claim}
+    \label{claim:no (sigma', nu'_3)-edge, no (sigma', nu'_4)-edge}
+    There is neither a $(\sigma', \nu'_3)$-edge
+    nor a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
+  \end{claim}
+
+  \begin{proof}[of Claim~\label{claim:no (sigma', nu'_3)-edge}]
+    According to
+    Claim~\ref{claim: no edge right of nu'_2},
+    Claim~\ref{claim:all (nu_1, nu'_1)-edges}
+    Claim~\ref{claim:all (nu_2, nu'_2)-edges} and
+    Claim~\ref{claim:no (nu_4, nu'_3)-edge}
+    every position in $\sigma'$ is involved in
     either a $(\sigma', \nu'_3)$-edge
@@
     in $\mathcal{M}$.
-  \qed
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Patterns ----
-%%
-\section{From square permutations to binary squares}
-\label{section:From square permutations to binary squares}
-
-It is conjectured in \cite{Henshall:Rampersad:Shallit:2011} that the number of binary strings
-of length $2n$ is given asymptoticaly by
-$\frac{\binom{2n}{n}}{n+1} 2^n - \binom{2n-1}{n+1} 2^{n-1} + O(2^{n-2})$.
-
-\begin{proposition}
-  \label{proposition:bijection binary - (213,231)-avoiding}
-  There exists a bijection between square binary strings of length $2n$ and
-  square $(213,231)$-avoiding permutations of length $2n$.
-\end{proposition}
+    Let $a$ be the first position of $\sigma'$
+    (\emph{i.e.}, $a = 2N_1 + N_2 + N_3 + 1$)
+    and
+    $b$ be the last position of $\sigma'$
+    (\emph{i.e.}, $b = 2N_1 + N_2 + N_3 + k + 2$),
+    and write $(a, a')$ and $(b, b')$ for the two associated edges of $\mathcal{M}$.
+    We consider the presence of a $(\sigma', \nu'_3)$-edge
+    or a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$ separately.
+    \begin{itemize}
+      \item
+      Suppose, aiming at a contradiction, that there exists
+      a $(\sigma', \nu'_3)$-edge in $\mathcal{M}$.
+      Since $\mathcal{M}$ is containment-free, we conclude that
+      $(a, a')$ must be a $(\sigma', \nu'_3)$-edge.
+      But $\mu(a) = 2N_2 + N_4 + 2n + k + 2 + (k+1) <
+      \mu(b) = 2N_2 + N_4 + 2n + k + 2 + (k+2)$, and hence
+      $(b, b')$ is also a $(\sigma', \nu'_3)$-edge
+      (otherwise we would have $\mu(a') > \mu(b')$ since
+      $\nu'_3$ is above $\pi'$ and $\nu'_4$,
+      which contradicts our hypothesis).
+      Consider any position $c$ but $a$ and $b$ of $\sigma'$
+      (\emph{i.e.},
+      $a = 2N_1 + N_2 + N_3 + 1 < c < 2N_1 + N_2 + N_3 + k + 2 = b$),
+      and let $(c, c')$ be the associate edge in $\mathcal{M}$.
+      We first observe that $(c, c')$ is not a $(\sigma', \nu'_3)$-edge,
+      since otherwise we would have
+      $\mu(a) > \mu(c)$ and $\mu(a') < \mu(c')$
+      (since $\mathcal{M}$ is containment-free and $\nu'_3$ is increasing),
+      thereby contradicting our hypothesis about $\mathcal{M}$.
+      Therefore, $(c, c')$ is either a $(\sigma', \pi)$-edge or a $(\sigma', \nu'_4)$-edge.
+      In both cases, $(b, b')$ is contained in $(c, c')$,
+      which contradicts our hypothesis about $\mathcal{M}$ being containment-free.
+      \item
+      Suppose, aiming at a contradiction, that there exists
+      a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
+      Since $\mathcal{M}$ is containment-free, we conclude that
+      $(b, b')$ must be a $(\sigma', \nu'_4)$-edge.
+      Since there is no $(\sigma', \nu'_3)$-edge in $\mathcal{M}$
+      (preceding point),
+      $(a, a')$ is either a
+      $(\sigma', \pi')$-edge or a $(\sigma', \nu'_4)$-edge.
+      In both cases, we have $a' < a$.
+      Then it follows that
+      $\mu(a) < \mu(b)$
+      (since $2N_2 + N_4 + 2n + k + 2 + (k+1) < 2N_2 + N_4 + 2n + k + 2 + (k+2)$)
+      and
+      $\mu(a') > \mu(b')$ (since $\pi'$ is above $\nu'_4$ if $(a, a')$ is
+      a $(\sigma', \pi')$-edge and
+      $\nu'_4$ is decreasing if $(a, a')$ is
+      a $(\sigma', \nu'_4)$-edge),
+      which contradicts our hypothesis about $\mathcal{M}$.
+    \end{itemize}
+    \qed
+  \end{proof}
+
+  Combining the above claims, we conclude that there are
+  $k+2$ $(\sigma', \pi')$-edges in $\mathcal{M}$.
+  Recall that
+  $\sigma' = ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$ and that
+  $\pi' = ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$.
+  Then it follows we have
+  at least $k$ (possibly $k+1$ or $k+2$) independent $(\sigma', \pi')$-edges
+  $(a, a')$ in $\mathcal{M}$ with
+  $2N_1 + N_2 + N_3 + 1 < a < 2N_1 + N_2 + N_3 + k + 2$
+  and
+  $2N_1 + N_2 + N_3 + (k + 2) + 1 < a' < 2N_1 + N_2 + N_3 + (k + 2) + n + 2$.
+  Therefore, by our hypothesis about $\mathcal{M}$,
+  $\sigma$ occurs as a pattern in $\pi$.
+  \qed
+\end{proof}
 
@@
 
-bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla
-bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla
-bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla
-bla bla blabla bla bla blabla bla bla blabla bla bla blabla bla bla blabla
-bla bla blabla bla bla blabla bla bla blabla bla bla blabla.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-09-14 article.tex
conclusion

@@
 
-\bigskip
-\bigskip
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+There are a number of further directions of investigation in this general subject.
+We mention several -~ñot necessarily new~- open problems that are, in our opinion,
+the most interesting.
+How many permutations of $S_{2n}$ are squares?
+How many $(213,231)$-avoiding permutations of $S_{2n}$ are squares?
+(Equivalentely, how many strings in $\{0,1\}^{2n}$ are squares;
+see Problem~4 in \cite{Henshall:Rampersad:Shallit:2011})?
+How are is the problem of deciding whether whether a $(213,231)$9avoiding permutation
+is a square (see Problem~4 in \cite{Henshall:Rampersad:Shallit:2011})?)
+Given two permutations $\pi$ and $\sigma$, how hard is the problem of deciding
+whether $\pi$ is the union of two patterns order-isomorphic to $\sigma$?
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-09-14 article.tex
conclusion

--- 2015-09-14 article.tex
@@
 
+There are a number of further directions of investigation in this general subject.
+We mention several -~not necessarily new~- open problems that are, in our opinion,
+the most interesting.
+How many permutations of $S_{2n}$ are squares?
+How many $(213,231)$-avoiding permutations of $S_{2n}$ are squares?
+(Equivalentely, how many binary strings of length $2n$ are squares;
+see Problem~4 in \cite{Henshall:Rampersad:Shallit:2011})?
+How hard is the problem of deciding whether a $(213,231)$-avoiding permutation
+is a square (see Problem~4 in \cite{Henshall:Rampersad:Shallit:2011})?)
+Given two permutations $\pi$ and $\sigma$, how hard is the problem of deciding
+whether $\pi$ is the union of two patterns order-isomorphic to $\sigma$?
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-09-14 article.tex
abstract

@@
 \begin{abstract}
-bla bla bla \ldots
+  A permutation is said to be a square if it is the union of two order-isomorphic
+  patterns.
+  The definition is intended to be the natural counterpart to the ordinary shuffle
+  of words and languages.
+  In this paper, we tackle the problem of recognising square permutations from both
+  the point of view of algebra and algorithmic.
+  On the one hand, we present some algebraic properties of the shuffle product of
+  permutations and exhibit a bijection between square $(213,231)$-avoiding permutations
+  and square binary words.
+  On the other hand, we prove that recognising square permutations is \NPC.
 \end{abstract}
@@
 \label{section:Introduction}
+
+$123$-avoiding (resp. $321$-avoiding) permutations are the union of two decreasing
+(resp. incresing) patterns.
 

### 2015-09-14 article.tex
Some typos and improvements

@@
 \usepackage{shuffle}
+\usepackage{multirow}
 
@@
 % \renewcommand{\textfraction}{.1}
+\renewcommand{\arraystretch}{1.2}
 
@@
 
+For any nonnegative integer $n$, $[n]$ is the set $\{1, \dots, n\}$.
 We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
@@
 We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
-set of all permutations. In this paper, permutations are specified by
-words without repetition of length $n$ on the alphabet $[n]$, so that
-all above definitions about words remain valid on permutations. If
-$\pi$ is a permutation, we denote by $\bar \pi$ (resp.~$\pi^{-1}$) the
-{\em complementary} (resp. {\em inverse}) of $\pi$. If $u$ is a word of
-integers without multiple occurrences of a same letter, $\STD(u)$ is the
-{\em standarized} of $u$, that is the unique permutation of the same
-size as $u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only
-if $\STD(u)_i < \STD(u)_j$. In particular, the image of the map $\STD$
-is the set $S$ of all permutations. Two words $u$ and $v$ having the
-same standarized are {\em order isomorphic}. If $\sigma$ is a permutation,
-there is an {\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if
-there is a set $P$ of indexes of letters of $\pi$ such that $\sigma$ and
-$\pi_{|P}$ are order isomorphic. When $\pi$ does not admit any occurrence
-of $\sigma$, $\pi$ {\em avoids} $\sigma$.
+set of all permutations. In this paper, permutations of a size $n$
+are specified by words of length $n$ on the alphabet $[n]$ and without
+multiple occurrence of a letter, so that all above definitions about
+words remain valid on permutations. If $\pi$ is a permutation, we denote
+by $\bar \pi$ (resp.~$\pi^{-1}$) the {\em complementary} (resp.
+{\em inverse}) of $\pi$. If $u$ is a word of integers without multiple
+occurrences of a same letter, $\STD(u)$ is the {\em standarized} of $u$,
+that is the unique permutation of the same size as $u$ such that for all
+$i, j \in [|u|]$, $u_i < u_j$ if and only if $\STD(u)_i < \STD(u)_j$. In
+particular, the image of the map $\STD$ is the set $S$ of all
+permutations. Two words $u$ and $v$ having the same standarized are
+{\em order-isomorphic}. If $\sigma$ is a permutation, there is an
+{\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if there is a
+set $P$ of indexes of letters of $\pi$ such that $\sigma$ and $\pi_{|P}$
+are order-isomorphic. When $\pi$ does not admit any occurrence of
+$\sigma$, $\pi$ {\em avoids} $\sigma$.
 
@@
 
-Let use denote by $\QQ[S]$ the linear span of all permutations. As
-announced, we shall define on this space a kind of shuffle product and
-for this purpose, we first define a linear coproduct $\Delta$ on $\QQ[S]$.
-For any permutation $\pi$, we set
+Let use denote by $\QQ[S]$ the linear span of all permutations. We
+define a linear coproduct $\Delta$ on $\QQ[S]$ in the following way. For
+any permutation $\pi$, we set
 \begin{equation} \label{equ:unshuffling_coproduct}
     \Delta(\pi) =
-    \sum_{P_1 \uplus P_2 = [|\pi|]}
+    \sum_{P_1 \sqcup P_2 = [|\pi|]}
     \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
@@
 in~\eqref{equ:example_unshuffling_coproduct} because there are exactly
-three ways to extract two disjoint subwords of the permutation $1432$
-so that the letters of these two subwords have respectively the same
-relative order as the letters of the permutations $1$ and $132$.
+three ways to extract from the permutation $1432$ two disjoint subwords
+respectively order-isomorphic to the permutations $1$ and $132$.
 \medskip
 
-As any coproduct, $\Delta$ leads to the definition of a product obtained
-by duality in the following way. From~\eqref{equ:unshuffling_coproduct},
-for any permutation $\pi$, we have
+As announced, let us now use $\Delta$ to define a shuffle product on
+permutations. As any coproduct, $\Delta$ leads to the definition of a
+product obtained by duality in the following way.
+From~\eqref{equ:unshuffling_coproduct}, for any permutation $\pi$, we
+have
 \begin{equation} \label{equ:schematic_coproduct}
@@
 where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
-by definition of duality, the dual product of $\Delta$ denoted by
-$\SHUFFLE$ is a linear binary product on $\QQ[S]$. It satisfies, for any
-permutations
-$\sigma$ and $\nu$,
+by definition of duality, the dual product of $\Delta$, denoted by
+$\SHUFFLE$, is a linear binary product on $\QQ[S]$. It satisfies, for
+any permutations $\sigma$ and $\nu$,
 \begin{equation}
@@
 Intuitively, this product shuffles the values and the positions of the
-letters of the permutations. There are other shuffle products on
+letters of the permutations. As first and easy properties of $\SHUFFLE$,
+one can observe that the empty permutation $\epsilon$ is a unit for
+$\SHUFFLE$ and that this product is graded by the sizes of the
+permutations ({\em i.e.}, the product of a permutation of size $n$ with
+a permutation of size $m$ produces a sum of permutations of size $n + m$.
+Notice that there are other shuffle products on
 permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
@@
 satisfying $P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$
-is order isomorphic to $\sigma$ and the subword $\pi_{|P_2}$ is order
-isomorphic to $\nu$.
+is order-isomorphic to $\sigma$ and the subword $\pi_{|P_2}$ is
+order-isomorphic to $\nu$.
 \medskip
@@
 $P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$ and
-$\pi_{|P_2}$ are order isomorphic. Computer experiments give us the
+$\pi_{|P_2}$ are order-isomorphic. Computer experiments give us the
 first numbers of square permutations with respects to their size, which
@@
 For instance, the square permutations of sizes $0$ to $4$ are
-\begin{multline}
-    \epsilon, 12, 21,
-    1234, 1243, 1423, 1324, 1342, 4132, 3124, 3142, \\ 3412, 4312,
-    2134, 2143, 2413, 4213, 2314, 2431, 4231, 3241, 3421, 4321.
-\end{multline}
+\smallskip
+
+\begin{tabular}{c|c|c}
+    Size $0$ \, & Size $2$ & Size $4$ \\ \hline
+    \multirow{2}{*}{\, $\epsilon$} &
+    \multirow{2}{*}{\, $12$, $21$ \,} &
+    $1234$, $1243$, $1423$, $1324$, $1342$, $4132$, $3124$, $3142$,
+    $3412$, $4312$, \\
+    & & $2134$, $2143$, $2413$, $4213$, $2314$, $2431$, $4231$, $3241$,
+    $3421$, $4321$
+\end{tabular}
 \medskip
@@
         there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
-        of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are order
-        isomorphic. This implies, by the definitions of $\BINTOPERM$ and
-        $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$ is a
-        square binary word.
+        of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
+        order-isomorphic. This implies, by the definitions of $\BINTOPERM$
+        and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
+        is a square binary word.
         \qed
@@
         (\Delta \otimes I)
-        \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
         & =
-        \sum_{P_1 \uplus P_2 = [|\pi|]}
+        \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
@@
         & =
-        \sum_{P_1 \uplus P_2 = [|\pi|]} \;
-        \sum_{Q_1 \uplus Q_2 = [|P_1|]}
+        \sum_{P_1 \sqcup P_2 = [|\pi|]} \;
+        \sum_{Q_1 \sqcup Q_2 = [|P_1|]}
         \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
@@
         & =
-        \sum_{P_1 \uplus P_2 \uplus P_3 = [|\pi|]}
+        \sum_{P_1 \sqcup P_2 \sqcup P_3 = [|\pi|]}
         \STD\left(\pi_{|P_1}\right) \otimes
@@
     $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
-    the operation $\widetilde{\,}$ define a bijection on the set of the
+    the operation $\widetilde{\,}$ defines a bijection on the set of the
     subsets of $[|\pi|]$, and since the standardization operation
-    commutes with the mirror operation on words, we have
+    commutes with the mirror operation on words without multiple
+    occurrence of a letter, we have
     \begin{equation} \begin{split}
         \Delta(\phi_1(\pi))
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\phi_1(\pi)_{|P_1}\right)
         \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\widetilde{\pi}_{|P_1}\right)
         \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
         \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \widetilde{\STD\left(\pi_{|P_1}\right)}
         \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
@@
         \Delta(\phi_2(\pi))
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\phi_2(\pi)_{|P_1}\right)
         \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\bar \pi_{|P_1}\right)
         \otimes \STD\left(\bar \pi_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
@@
         \Delta(\phi_3(\pi))
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\phi_3(\pi)_{|P_1}\right)
         \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\pi^{-1}_{|P_1}\right)
         \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
         \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \STD\left(\pi_{|P_1}\right)^{-1}
         \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
-        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
         \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
@@
     \ref{item:square_stability_2}, and~\ref{item:square_stability_3} are
-    direct consequences of Proposition~\ref{prop:endomorphisms}. Indeed,
+    consequences of Proposition~\ref{prop:endomorphisms}. Indeed,
     since $\pi$ is a square permutation and $\rho$ is a square root of
@@
     Then, since $\pi$ appears in $\rho \SHUFFLE \rho$, $\phi_j(\pi)$
-    appears in $\phi_j(\rho \SHUFFLE \rho)$ and also in
+    appears in $\phi_j(\rho \SHUFFLE \rho)$ and appears also in
     $\phi_j(\rho) \SHUFFLE \phi_j(\rho)$. This shows that
@@
 Let us make an observation about Wilf-equivalence classes of
-permutations restrained on square permutations. The fact that there is a
-single Wilf-equivalence class of permutations of size $3$ together with
+permutations restrained on square permutations. The
+well-known~\cite{Simion:Schmidt:EJC:1985} fact that there is a single
+Wilf-equivalence class of permutations of size $3$ together with
 Proposition~\ref{prop:square_stability} imply that $123$ and $321$ are

### 2015-09-15 article.tex
article

@@
 \label{section:Introduction}
+
+Given three words $u$, $v$ and $w$, $u$ is said to a
+\emph{shuffle} of $v$ and $w$ if it
+can be formed by interleaving the letters from $v$ and $w$ in a way
+that maintains the left-to-right ordering of the letters from each word
+(it is a classical textbook exercice to
+design an efficient dynamic programming algorithm that determines
+whether $u$ is a shuffle of $v$ and $w$).
+A word $u$ is said to be a \emph{square} (w.r.t. the shuffle) if there exists
+a word $v$ such that $u$ is a shuffle of $v$ with itself.
+
+This paper is intended to study a natural generalisation of the shuffle
+to permutations.
+Given permutations $\pi$, $\mu$ and $\nu$, $\pi$ is said to be a shuffle of
+$\mu$ and $\nu$ if $\pi$ (viewed as a word) is a shuffle of two words that
+are order-isomorphic to $\mu$ and $\nu$.
+For example, $\pi$ is the shuffle
+of $\mu=$ and $\nu=$ as
+$\pi$.
+A permutation $\pi$ is said to be a square (w.r.t. the shuffle) if there exists
+a permutation $\mu$ such that $\pi$ is a shuffle of $\mu$ with itself.
+For example, $\pi = 241356$ is a square as it is the shuffle of $213$ with itself
+(\emph{e.g.},
+$\pi = \dot{2}\ddot{4}\dot{1}\ddot{3}\ddot{5}\dot{6}$, where
+$\dot{i}$ and $\ddot{i}$),
+whereas $\pi = 4123$ is not.
+
+Formally, the shuffle of two words $u$ and $v$ of $A^*$ is the language
+$u \shuffle v$ consisting of all words
+$u_1 v_1 u_2 v_2 \ldots u_k v_k$, where $k \geq 0$ and
+the $u_i$ and the $v_i$ are the words  of $A^*$
+such that $u = u_1 u_2 \ldots u_k$ and
+$v = v_1 v_2 \ldots v_k$.
+We first briefly review the key results for words
+that arise in our context.
+Given words $u$, $v_1$ and $v_2$, it can be tested in
+$O\left(|u|^2 / \log(|u|)\right)$ time whether or not
+$u \in v_1 \shuffle v_2$~\cite{Leeuwen:Nivat:IPL:1982}.
+(To the best of our knowledge, the first
+$O(|u|^2)$ time algorithm appeared in~\cite{Mansfield:DAM:1983}.
+This algorithm can easily be extended to check in polynomial-time
+whether or not a word is the shuffle of any fixed number of given words.)
+The shuffle $u \shuffle v$ of words $u$ and $v$ can be computed in
+$O\left((|u|+|v|) \; \binom{|u|+|v|}{|u|}\right)$
+time~\cite{Spehner:TCS:1986}.
+An improvement and generalization has been proposed in
+\cite{Allauzen:IGM:2000}, where it is proved that,
+given words $u_1, u_2, \ldots, u_k$, the shuffle
+$\shuffle_{i=1}^{k} u_i$ can be computed in
+$O\left(\binom{|u_1|+|u_2|+\ldots+|u_k|}{|u_1|,|u_2|,\ldots,|u_k|}\right)$ time.
+Given words $u, v_1, v_2, \ldots, v_n \in A^*$, it is however \NPC to
+decide whether or not
+$u \in \shuffle_{i=1}^{k} v_i$.
+\cite{Mansfield:DAM:1983,Warmuth:Haussler:JCSS:1984}.
+This remains true even if the alphabet has size $3$~\cite{Warmuth:Haussler:JCSS:1984}.
+Of particular interest,
+it is shown in~\cite{Warmuth:Haussler:JCSS:1984} that this problem remains
+\NPC even if all
+words $v_1, v_2, \ldots, v_n$ are identical,
+thereby proving that,
+for two words $u$ and $v$, it is \NPC to decide whether or not $u$
+is in the iterated shuffle of $v$.
+Again, this remains true even if the alphabet has size $3$.
+As for squares, given a word $z$ it is \NPC to determine whether
+there exists a word $x$ such that $z$ is the shuffle of $x$ with itself
+\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}.
 

### 2015-09-15 article.tex
Ajouts dans conclusion et dans introduction

@@
 \begin{abstract}
-  A permutation is said to be a square if it is the union of two order-isomorphic
-  patterns.
-  The definition is intended to be the natural counterpart to the ordinary shuffle
-  of words and languages.
-  In this paper, we tackle the problem of recognising square permutations from both
-  the point of view of algebra and algorithmic.
-  On the one hand, we present some algebraic properties of the shuffle product of
-  permutations and exhibit a bijection between square $(213,231)$-avoiding permutations
-  and square binary words.
-  On the other hand, we prove that recognising square permutations is \NPC.
+    A permutation is said to be a square if it is the union of two
+    order-isomorphic patterns. The definition is intended to be the
+    natural counterpart to the ordinary shuffle of words and languages.
+    In this paper, we tackle the problem of recognizing square
+    permutations from both the point of view of algebra and algorithmic.
+    On the one hand, we present some algebraic properties of the shuffle
+    product of permutations and exhibit a bijection between square
+    $(213,231)$-avoiding permutations and square binary words. On the
+    other hand, we prove that recognizing square permutations is \NPC.
 \end{abstract}
@@
 
-Given three words $u$, $v$ and $w$, $u$ is said to a
-\emph{shuffle} of $v$ and $w$ if it
-can be formed by interleaving the letters from $v$ and $w$ in a way
-that maintains the left-to-right ordering of the letters from each word
-(it is a classical textbook exercice to
-design an efficient dynamic programming algorithm that determines
-whether $u$ is a shuffle of $v$ and $w$).
-A word $u$ is said to be a \emph{square} (w.r.t. the shuffle) if there exists
-a word $v$ such that $u$ is a shuffle of $v$ with itself.
-
-This paper is intended to study a natural generalisation of the shuffle
-to permutations.
-Given permutations $\pi$, $\mu$ and $\nu$, $\pi$ is said to be a shuffle of
-$\mu$ and $\nu$ if $\pi$ (viewed as a word) is a shuffle of two words that
-are order-isomorphic to $\mu$ and $\nu$.
-For example, $\pi$ is the shuffle
-of $\mu=$ and $\nu=$ as
-$\pi$.
-A permutation $\pi$ is said to be a square (w.r.t. the shuffle) if there exists
-a permutation $\mu$ such that $\pi$ is a shuffle of $\mu$ with itself.
-For example, $\pi = 241356$ is a square as it is the shuffle of $213$ with itself
-(\emph{e.g.},
-$\pi = \dot{2}\ddot{4}\dot{1}\ddot{3}\ddot{5}\dot{6}$, where
-$\dot{i}$ and $\ddot{i}$),
-whereas $\pi = 4123$ is not.
+Given three words $u$, $v$ and $w$, $u$ is said to a \emph{shuffle} of
+$v$ and $w$ if it can be formed by interleaving the letters from $v$ and
+$w$ in a way that maintains the left-to-right ordering of the letters
+from each word (it is a classical textbook exercise to design an
+efficient dynamic programming algorithm that determines whether $u$ is a
+shuffle of $v$ and $w$). A word $u$ is said to be a \emph{square} (w.r.t.
+the shuffle) if there exists a word $v$ such that $u$ is a shuffle of $v$
+with itself.
+\smallskip
+
+This paper is intended to study a natural generalization of the shuffle
+to permutations. Given permutations $\pi$, $\mu$ and $\nu$, $\pi$ is
+said to be a shuffle of $\mu$ and $\nu$ if $\pi$ (viewed as a word) is
+a shuffle of two words that are order-isomorphic to $\mu$ and $\nu$. For
+example, $\pi$ is the shuffle of $\mu=$ and $\nu=$ as $\pi$. A
+permutation $\pi$ is said to be a square (w.r.t. the shuffle) if there
+exists a permutation $\mu$ such that $\pi$ is a shuffle of $\mu$ with
+itself. For example, $\pi = 241356$ is a square as it is the shuffle of
+$213$ with itself (\emph{e.g.},
+$\pi = \dot{2}\ddot{4}\dot{1}\ddot{3}\ddot{5}\dot{6}$, where $\dot{i}$
+and $\ddot{i}$), whereas $\pi = 4123$ is not.
+The shuffle product defined and considered in this paper is different
+from other existing and well-known shuffle products on permutations.
+For instance, in~\cite{DHT:IJAC:2002}, the authors defined the
+{\em shifted shuffle product}. In this product $\pi$ is a shuffle of
+$\mu$ and $\nu$ if $\pi$ is the shuffle, as words, of $\mu$ and the
+word obtained by incrementing all letters of $\nu$ by the size of $\mu$.
+It is a simple exercise to prove that, give a permutation $\pi$, the
+decision problem to test if $\pi$ is a square for the shifted shuffle
+product is linear.
+\smallskip
 
 Formally, the shuffle of two words $u$ and $v$ of $A^*$ is the language
-$u \shuffle v$ consisting of all words
-$u_1 v_1 u_2 v_2 \ldots u_k v_k$, where $k \geq 0$ and
-the $u_i$ and the $v_i$ are the words  of $A^*$
-such that $u = u_1 u_2 \ldots u_k$ and
-$v = v_1 v_2 \ldots v_k$.
-We first briefly review the key results for words
-that arise in our context.
+$u \shuffle v$ consisting of all words $u_1 v_1 u_2 v_2 \ldots u_k v_k$,
+where $k \geq 0$ and the $u_i$ and the $v_i$ are the words  of $A^*$
+such that $u = u_1 u_2 \ldots u_k$ and $v = v_1 v_2 \ldots v_k$. We
+first briefly review the key results for words that arise in our context.
 Given words $u$, $v_1$ and $v_2$, it can be tested in
@@
 $u \in v_1 \shuffle v_2$~\cite{Leeuwen:Nivat:IPL:1982}.
-(To the best of our knowledge, the first
-$O(|u|^2)$ time algorithm appeared in~\cite{Mansfield:DAM:1983}.
-This algorithm can easily be extended to check in polynomial-time
-whether or not a word is the shuffle of any fixed number of given words.)
-The shuffle $u \shuffle v$ of words $u$ and $v$ can be computed in
-$O\left((|u|+|v|) \; \binom{|u|+|v|}{|u|}\right)$
-time~\cite{Spehner:TCS:1986}.
+(To the best of our knowledge, the first $O(|u|^2)$ time algorithm a
+ppeared in~\cite{Mansfield:DAM:1983}. This algorithm can easily be
+extended to check in polynomial-time whether or not a word is the
+shuffle of any fixed number of given words.) The shuffle $u \shuffle v$
+of words $u$ and $v$ can be computed in
+$O\left((|u|+|v|) \; \binom{|u|+|v|}{|u|}\right)$ time~\cite{Spehner:TCS:1986}.
 An improvement and generalization has been proposed in
-\cite{Allauzen:IGM:2000}, where it is proved that,
-given words $u_1, u_2, \ldots, u_k$, the shuffle
-$\shuffle_{i=1}^{k} u_i$ can be computed in
-$O\left(\binom{|u_1|+|u_2|+\ldots+|u_k|}{|u_1|,|u_2|,\ldots,|u_k|}\right)$ time.
-Given words $u, v_1, v_2, \ldots, v_n \in A^*$, it is however \NPC to
-decide whether or not
-$u \in \shuffle_{i=1}^{k} v_i$.
+\cite{Allauzen:IGM:2000}, where it is proved that, given words
+$u_1, u_2, \ldots, u_k$, the shuffle $\shuffle_{i=1}^{k} u_i$ can be
+computed in
+$O\left(\binom{|u_1|+|u_2|+\ldots+|u_k|}{|u_1|,|u_2|,\ldots,|u_k|}\right)$
+time. Given words $u, v_1, v_2, \ldots, v_n \in A^*$, it is however \NPC
+to decide whether or not $u \in \shuffle_{i=1}^{k} v_i$.
 \cite{Mansfield:DAM:1983,Warmuth:Haussler:JCSS:1984}.
@@
 \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}.
+\smallskip
 
 $123$-avoiding (resp. $321$-avoiding) permutations are the union of two decreasing
-(resp. incresing) patterns.
+(resp. increasing) patterns.
 
@@
 
-There are a number of further directions of investigation in this general subject.
-We mention several -~not necessarily new~- open problems that are, in our opinion,
-the most interesting.
-How many permutations of $S_{2n}$ are squares?
-How many $(213,231)$-avoiding permutations of $S_{2n}$ are squares?
-(Equivalentely, how many binary strings of length $2n$ are squares;
-see Problem~4 in \cite{Henshall:Rampersad:Shallit:2011})?
-How hard is the problem of deciding whether a $(213,231)$-avoiding permutation
-is a square (see Problem~4 in \cite{Henshall:Rampersad:Shallit:2011})?)
-Given two permutations $\pi$ and $\sigma$, how hard is the problem of deciding
+There are a number of further directions of investigation in this
+general subject. We mention several -~not necessarily new~- open
+problems that are, in our opinion, the most interesting. How many
+permutations of $S_{2n}$ are squares? How many $(213,231)$-avoiding
+permutations of $S_{2n}$ are squares? (Equivalently, how many binary
+strings of length $2n$ are squares; see Problem~4
+in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of
+deciding whether a $(213,231)$-avoiding permutation is a square
+(see Problem~4 in \cite{Henshall:Rampersad:Shallit:2011})?) Given two
+permutations $\pi$ and $\sigma$, how hard is the problem of deciding
 whether $\pi$ is the union of two patterns order-isomorphic to $\sigma$?
+\smallskip
+
+Besides, in this work, we investigate some algebraic properties of the
+shuffle of permutations which have direct combinatorial consequences.
+However, this work raises several additional classical algebraic
+questions. Among these, one can ask for a complete algebraic study of
+$\QQ[S]$ as a graded associative algebra for the shuffle  product
+$\SHUFFLE$. Describing a generating family for $\QQ[S]$, defining
+multiplicative bases of $\QQ[S]$, and determining whether $\QQ[S]$ is
+free as an associative algebra are worthwhile questions.
 

### 2015-09-16 article.tex
Introduction hardness

@@
 
-\begin{proposition}
-  \label{proposition:matching}
-  Let $\pi \in S_{2n}$.
-  Then,
-  $\pi$ is the union of two order-isomorphic patterns
-  if and only if
-  there exists a perfect containment-free mathing
+This section is devoted to proving hardness of recognizing square permutations.
+In the same way as happens with words, we shall use a linear graph framework
+where deciding whether a permutation is a square reduces to computing some
+specific matching in the associated linear graph \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}.
+It reads as follows.
+
+\begin{lemma}
+  \label{lemma:matching}
+  Let $\pi$ be a permutation. The following statements are equivalent:
+  \begin{enumerate}
+    \item
+    $\pi$ is a square.
+    \item
+    There exists a perfect containment-free mathing
     $\mathcal{M}$ of $G_\pi$ such that
-  $\LEFT(\mathcal{M}) \simeq \RIGHT(\mathcal{M})$.
-\end{proposition}
-
-The following easy lemma will prove extremely useful for simplifying the proof of
+    for any two disctinct edges
+    $(a, b)$ and $(a', b')$ in $\mathcal{M}$
+    we have $\pi(a) < \pi(a')$ if and only if $\pi(b) < \pi(b')$.
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}[of Lemma~\ref{lemma:matching}]
+    \qed
+\end{proof}
+
+Before proving hardness, we give an easy lemma that will prove extremely
+useful for simplifying the proof of upcoming
 Proposition~\ref{proposition:hardness}.
@@
   Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
-  be a permutation with $n_1 = |\sigma_1| \geq 2$ and $n_2 = |\sigma_2| \geq 2$,
+  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
   and $\mathcal{M}$ be a containment-free perfect matching of $G_\pi$.
-  The following assertions hold.
+  The following assertions hold:
   \begin{enumerate}
@@
 \begin{proof}
+  The problem is certainly in $\NPclass$.
+  We propose a reduction from the pattern involvement problem which is
+  known to be \NPC~\cite{Bose:Buss:Lubiw:1998}:
+  Given permutation $\pi$ and $\sigma$, decide whether $\sigma$
+  occurs in $\pi$ (as an order-isomorphic pattern).
+
   For the sake of clarity, we will that a bunch of positions $P$
@@
 
+  Let $\pi \in S_n$ and $\sigma \in S_k$ be two arbitrary permutations.
+  Define
   \begin{align*}
@@
   We now turn to defining various gadgets (sequences of integers)
-  that act as building blocks in our construction:
+  that act as building blocks in our construction of a new permutation $\mu$:
   \begin{align*}
@@
   $$
+
+  It is immediate that H can be constructed in polynomial-time in $n$ and $k$.
   We claim that $\sigma$ occurs as a pattern in $\pi$ if and only if

### 2015-09-16 article.tex
Introduction hardness

@@
 $\sigma$, $\pi$ {\em avoids} $\sigma$.
+We write $\nearrow_{n}$ (resp. $\searrow_{n}$)
+for the permutation $1 2 \ldots n$ (resp. $n (n-1) \ldots 1$).
+we write seqeunce of integers $s = s_1 s_2 \ldots s_n$, we
+we let $s\;[k]$ stand for the lifted sequence
+$(s_1+k) s_2+k) \ldots (s_n+k)$.
 
@@
 
+Let $\pi$ be a permutation.
+For the sake of clarity, we will say that a bunch of consecutive positions $P$
+of $\pi$ is \emph{above} (resp. \emph{below}) above another bunch of consecutive
+positions $P'$ in permutation $\pi$
+if $\pi(i) > \pi(j))$ (resp. $\pi(i) < \pi(j)$) for every
+$i \in P$ and every $j \in P'$.
+For exemple, $\sigma_1$ is above $\sigma_2$
+(in an equivalent manner, $\sigma_2$ is below $\sigma_1$)
+in Fig.~\ref{subfig:increasing before and aboce decreasing},
+whereas $\sigma_1$ is below $\sigma_2$
+(in an equivalent manner, $\sigma_2$ is above $\sigma_1$)
+in Fig.~\ref{subfig:decreasing before and below increasing}.
+
 Before proving hardness, we give an easy lemma that will prove extremely
@@
     $\sigma_1$ is below $\sigma_2$
-    (see Fig.~\ref{subfig:decresing before and below increasing}),
+    (see Fig.~\ref{subfig:decreasing before and below increasing}),
     then $\mathcal{M}$ sontains at most one $(\sigma_1, \sigma_2)$-edge.
@@
       \end{tikzpicture}
-      \label{subfig:decresing before and below increasing}
+      \label{subfig:decreasing before and below increasing}
     }% end subfigure
@@
   occurs in $\pi$ (as an order-isomorphic pattern).
-
-  For the sake of clarity, we will that a bunch of positions $P$
-  is \emph{above} (resp. \emph{below}) above another bunch of positions $P'$
-  in permutation $\mu$
-  if $\mu(x) > \mu(y)$ (resp. $\mu(x) < \mu(y)$) for every $x \in P$
-  and every $y \in P'$.
-  For exemple, $\nu_1$ is above $\nu_2$ but below $\nu'_1$ in permutation $\mu$.
 
@@
       \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
-      \node (claim edge 2) at (4,4) {Claim\ref{claim:(nu_2, nu'_2)-edge}};
-      \node (claim edge 2) at (4,3.5) {Claim\ref{claim:all (nu_2, nu'_2)-edges}};
+      \node (claim edge 2) at (4,4) {Claim~\ref{claim:(nu_2, nu'_2)-edge}};
+      \node (claim edge 2) at (4,3.5) {Claim~\ref{claim:all (nu_2, nu'_2)-edges}};
       % 3
       \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
-      \node (claim edge 1) at (5.5,11.5) {Claim\ref{claim:(nu_3, nu'_3)-edge}};
+      \node (claim edge 1) at (5.5,11.5) {Claim~\ref{claim:(nu_3, nu'_3)-edge}};
       % 4
       \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
-      \node (claim edge 1) at (8.5,8.5) {Claim\ref{}};
+      \node (claim edge 1) at (8.5,8.5) {Claim~\ref{}};
       % sigma 1 - pi 1
       \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
-      \node (claim edge 1) at (6.5,4) {Claim\ref{}};
+      \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
       % nodes
@@
   \end{align*}
-  We are now in position to define our target permutation $\mu$:
+  We are now in position to define our target permutation $\mu$
+  (see Fig.~\ref{fig:reduction} for an illustration):
   $$
@@
   It is immediate that H can be constructed in polynomial-time in $n$ and $k$.
-  We claim that $\sigma$ occurs as a pattern in $\pi$ if and only if
+  We claim that $\sigma$ occurs in $\pi$ if and only if
   there exists a containment-free perfect matching
@@
 
-  Suppose first that $\sigma$ occurs in $\pi$ as a pattern.
+  Suppose first that $\sigma$ occurs in $\pi$.
   We show how to construct a containment-free perfect matching

### 2015-09-16 article.tex
reduction

@@
 $213$ with itself (\emph{e.g.},
-$\pi = \dot{2}\ddot{4}\dot{1}\ddot{3}\ddot{5}\dot{6}$, where $\dot{i}$
-and $\ddot{i}$), whereas $\pi = 4123$ is not.
+$\pi = \dot{2}{4}\dot{1}{3}{5}\dot{6}$, where the $\dot{x}$'s
+denote the second occurrence), whereas $\pi = 4123$ is not.
 The shuffle product defined and considered in this paper is different
@@
 positions $P'$ in permutation $\pi$
-if $\pi(i) > \pi(j))$ (resp. $\pi(i) < \pi(j)$) for every
+if $\pi(i) > \pi(j)$ (resp. $\pi(i) < \pi(j)$) for every
 $i \in P$ and every $j \in P'$.
@@
         [
-          scale=0.45,
+          scale=0.325,
           label/.style={anchor=base}
@@
     }% end subfigure
-    \qquad
+    \qquad\qquad\qquad
     \subfigure[%
@@
       [
-        scale=0.45,
+        scale=0.325,
         label/.style={anchor=base}
@@
   \label{proposition:hardness}
-  Deciding whether a permutation is the union of two
-  order-isomorphic patterns is \NPC.
+  Deciding whether a permutation is a square is \NPC.
 \end{proposition}
@@
     \begin{tikzpicture}[
-      scale=.95,
+      scale=.9,
       >=stealth',
@@
     \end{tikzpicture}
-    \caption{\label{fig:reduction}}
+    \caption{\label{fig:reduction}%
+      Schematic representation of the permutation $\mu$ used
+      in Proposition~\ref{proposition:hardness}.
+      Black edges denote the presence of at least one edge between two bunches of
+      positions in $\mu$.
+      Grey edges denote edges that are  only considered in the forward direction of the
+      proof.
+    }%
   \end{figure}
@@
 
-  Suppose first that $\sigma$ occurs in $\pi$.
-  We show how to construct a containment-free perfect matching
-  $\mathcal{M}$ in $G_\mu$ such that
-  for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+  Suppose first that $\sigma$ occurs in $\pi$ and fix an occurrence.
+  Construct a matching $\mathcal{M}$ as follows:
   \begin{itemize}
@@
     \item $\mathcal{M}$ contains $k+2$ pairwise crossing
-    $(\sigma', \pi')$-edges as depicted in Figure~\ref{}.
-    (Notice that we use here the fact that $\sigma$ occurs as a pattern in $\pi$).
+    $(\sigma', \pi')$-edges as depicted in
+    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
+    More precisely,
+    (i) the first position of $\sigma'$
+    (\emph{i.e.}, $(2N_1+N_2+N_3) + 1$) is linked
+    to the first position of $\pi'$
+    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + 1$),
+    (ii) the last position of $\sigma'$
+    (\emph{i.e.}, $(2N_1+N_2+N_3) + k+2$) is linked
+    to the last position of $\pi'$
+    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + n+2$),
+    and all other positions in $\sigma'$ are linked by means of $k$ pairwise
+    crossing edges to the positions in
+    $\pi'$ that correspond to the fixed occurence of $\sigma$ in $\pi$.
+    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
     \item $\mathcal{M}$ contains $n-k$ pairwise crossing
-    $(\pi', \pi'')$-edges as depicted in Figure~\ref{}.
+    $(\pi', \pi'')$-edges as depicted in
+    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
+    More precisely,
+    all positions in $\pi'$ that do not correspond to the fixed occurence of
+    $\sigma$ in $\pi$ are linked by means of $n-k$ pairwise crossing edges
+    to the positions in $\pi''$ that do not correspond to the fixed
+    occurrence of $\sigma$ in $\pi$.
     \item $\mathcal{M}$ contains $k$ pairwise crossing
-    $(\pi'', \sigma'')$-edges as depicted in Figure~\ref{}.
-    (Notice that we use here again the fact that $\sigma$ occurs as a pattern in $\pi$).
+    $(\pi'', \sigma'')$-edges as depicted in
+    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
+    More precisely, the positions in $\pi''$ that correspond to
+    the fixed occurrence of $\sigma$ in $\pi$ are llinked
+    linked by means of $k$ pairwise crossing edges to all positions in
+    $\sigma''$.
+    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
   \end{itemize}
-  It can be easily checked (probably referring to Figure~\ref{}) that
-  $\mathcal{M}$ is perfect and that
+  It can be easily checked (probably referring to Figure~\ref{fig:reduction}) that
+  $\mathcal{M}$ is perfect, containment-free and that
   for any two disctinct edges
@@
           [
-            scale=0.45,
+            scale=0.35,
             label/.style={anchor=base},
@@
             text width=5.5cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,-1.5cm)}
+            rectangle callout,callout relative pointer={(0.0cm,-1cm)}
           ] at (3,26)
@@
             text width=5.5cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,3cm)}
-          ] at (9.5,0)
+            rectangle callout,callout relative pointer={(0.0cm,3.5cm)}
+          ] at (9.5,-3.5)
           {%
@@
             text width=2.25cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,-4cm)}
+            rectangle callout,callout relative pointer={(0.0cm,-3cm)}
           ] at (16.5,19)
@@
             text width=1.75cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,1.25cm)}
+            rectangle callout,callout relative pointer={(0.0cm,0.75cm)}
           ] at (21,-4.5)
@@
         \caption{\label{fig:subfig:sigma' - pi' - pi'' - sigma''}%
-        $\sigma = 312$ and $\pi = 4\mathbf{5}\mathbf{2}1\mathbf{3}6$
+        Illustration of the containment-free perfect matching $\mathcal{M}$
+        between gadgets $\sigma'$, $\pi'$, $\pi''$ and $\sigma''$ assuming two
+        input permutation $\sigma = 312$ and $\pi = 4\mathbf{5}\mathbf{2}1\mathbf{3}6$
         (where a specific occurrence of $\sigma$ in $\pi$ is depicted in bold).
@@
       \end{figure}
-
-  More precisely,
-  (i) the first integer of $\sigma'$
-  (\emph{i.e.}, $(k+1) + (2N_2 + N_4 + 2n + k + 2)$) is linked
-  to the first integer of $\pi'$
-  ((\emph{i.e.}, $(n+1) + (2N_2 + N_4 + n + k + 2)$)),
-  (ii)
 
@@
   direction, the main idea is to prove that $\mathcal{M}$ contains enough structure
-  (more precisely, enough $(\sigma', \pi')$-edges) so that we can conclude.
+  (more precisely, $k+2$ $(\sigma', \pi')$-edges) so that we can conclude
+  that $\sigma$ occurs in $\pi$.
   We have divided the reverse direction into a serie of basic claims that

### 2015-09-16 article.tex
matching

@@
 \begin{proof}[of Lemma~\ref{lemma:matching}]
+  The proof is, naturally enough, a slight extension of
+  Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and
+  is added here in full for the sake of completeness.
+
+  (\emph{1} $\Rightarrow$ \emph{2}).
+  Let $\sigma$ be a permutation such that $\pi$ is the union of two
+  patterns that are order-isomorphic to $\sigma$.
+  Let $2n = |\pi|$ (and hence $|\sigma| = n$).
+  Fix the occurrences in $\pi$ of the two disjoint patterns that
+  are order-isomorphic to $\sigma$ and
+  write
+  $I^1 = \{i^1_1, i^1_2, \ldots, i^1_n\}$,
+  $i^1_1 < i^1_2 < \ldots < i^1_n$, for the positions in $\pi$
+  of the first pattern order-ismorphic to $\sigma$ and
+  $I^2 = \{i^2_1, i^2_2, \ldots, i^2_n\}$,
+  $i^2_1 < i^2_2 < \ldots < i^2_n$, for the positions in $\pi$
+  of the second pattern order-ismorphic to $\sigma$.
+  Let $G_\pi$ the linear clique associated to $\pi$ with
+  vertices $\VS{G_\pi} =
+  \{i^1_j : 1 \leq j \leq n\} \cup \{i^2_j : 1 \leq j \leq n\}$.
+  It is easily seen that
+  $\mathcal{M} = \{\{i^1_j, i^2_j\} : 1 \leq j \leq n\}$
+  is a subset of $\ES{G_\pi}$.
+  Furthermore, $\mathcal{M}$ is a perfect matching since
+  $I^1 \cap I^2 = \emptyset$,
+  and $I^1 \cup I^2 = \{1, 2,  \ldots 2n\}$.
+  It is also containment-free.
+  Indeed, if it were not the case then there would exist two edges
+  $e = \{i^1_j, i^2_j\}$ and
+  $e' = \{i^1_k, i^2_k\}$, $j < k$,
+  in $\mathcal{M}$ such that
+  $i^1_j < i^1_k < i^2_k < i^2_j$ or
+  $i^1_j < i^2_k < i^1_k < i^2_j$.
+  This is a contradiction since
+  $i^2_k > i^2_j$ if $k > j$.
+  Finally, for any two distinct edges
+  $e = \{\i^1_j, i^2_j\}$ and
+  $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
+  we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
+  $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing
+  in both cases two elements (at position $j$ and $k$)
+  in two patterns that are order-isomorphic to $\sigma$.
+
+  (\emph{2} $\Rightarrow$ \emph{1}).
+  Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
+  \;\text{with}\; j < k\}$, and
+  $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
+  \;\text{with}\; j < k\}$.
+  Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right) endpoints
+  of the edges in $\mathcal{M}$ assuming the positions $1, 2, \ldots, 2$
+  are drawn along an horizontal line.
+  Without loss of generality,
+  assume $i^1_1 < i^1_2 < \ldots < i^1_n$.
+  % and
+  % $I^2 = \{i^2_1, i^2_2, \ldots, i^2_n\}$.
+  Let
+  $p^1 = \pi(i^1_1) \pi(i^1_2) \ldots \pi(i^1_n)$ and
+  $p^2 = \pi(i^2_1) \pi(i^2_2) \ldots \pi(i^2_n)$ be the two patterns
+  from $\pi$ induced by $I^1$ and $I^2$, respectivelly.
+  Clearly $p^1$ and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is
+  a matching) and cover $\pi$ (since $\mathcal{M}$ is perfect).
+  We also claim that $p^1$ and $p^2$ are order-isomorphic.
+  Suppose, aiming at a contradiction, that this is not the case.
+  Therefore, there exist two positions $j$ and $k$, say $j < k$, such that
+  $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
+  $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
+  But since $\mathcal{M}$ is containment-free, it follows that
+  both $\{i^1_j, i^2_j\}$ and $\{i^1_k, i^2_k\}$.
+  This is a contradiction since
+  for any two disctinct edges
+  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
+  we have $\pi(a) < \pi(a')$ if and only if $\pi(b) < \pi(b')$.
   \qed
@@
     \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
-      Lemma~\ref{lemma:at most one edge monotone}:
-      Assuming a $(\nu_1,\nu_2)$-edge $(a, b)$ in $\mathcal{M}$.
+      Illustration of Lemma~\ref{lemma:at most one edge monotone}.
     }
@@
 permutations of $S_{2n}$ are squares? (Equivalently, how many binary
-strings of length $2n$ are squares; see Problem~4
+strings of length $2n$ are squares; Problem~4
 in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of
 deciding whether a $(213,231)$-avoiding permutation is a square
-(see Problem~4 in \cite{Henshall:Rampersad:Shallit:2011})?) Given two
-permutations $\pi$ and $\sigma$, how hard is the problem of deciding
-whether $\pi$ is the union of two patterns order-isomorphic to $\sigma$?
+(Problem~4 in \cite{Henshall:Rampersad:Shallit:2011},
+see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?)
+% Given two
+% permutations $\pi$ and $\sigma$, how hard is the problem of deciding
+% whether $\pi$ is the union of two patterns order-isomorphic to $\sigma$?
 \smallskip

### 2015-09-16 article.tex
nouvelle version avec les preuves qui manquaient

@@
 $<$ defined by $u_i < u_j$ if and only if $i < j$.
-(For a linear graph, we write $(u_i, u_j)$ instead of $\{u_i, u_j\}$
-for an edge if it is clear from the context that $i < j$.)
+For a linear graph, we write $(u_i, u_j)$ instead of $\{u_i, u_j\}$
+for an edge if it is clear from the context that $i < j$.
 Recall that two edges of a graph are \emph{independent} if they do not share
@@
   But since $\mathcal{M}$ is containment-free, it follows that
-  both $\{i^1_j, i^2_j\}$ and $\{i^1_k, i^2_k\}$.
+  both $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and
+  $\{i^1_k, i^2_k\}$, $i^1_k, i^2_k$, belong to $\mathcal{M}$.
   This is a contradiction since

### 2015-09-16 article.tex
nouvelle version avec les preuves qui manquaient
Debut homogeneisation

@@
 
-For any nonnegative integer $n$, $[n]$ is the set $\{1, \dots, n\}$.
+If $S$ is a finite set, the cardinality of $S$ is denoted by $|S|$,
+and if $P$ and $Q$ are two disjoint sets, $P \sqcup Q$ denotes the
+disjoint union of $P$ and $Q$. For any nonnegative integer $n$, $[n]$
+is the set $\{1, \dots, n\}$.
+
 We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
-Let us recall here the most important ones. The {\em empty word} is
-denoted by $\epsilon$. If $u$ is a word, we denote by $\widetilde{u}$
-the {\em mirror image} of $u$. If $P$ is a subset of $[|u|]$, $u_{|P}$
+Let us recall here the most important ones. Let $u$ be a word. The
+length of $u$ is denoted by $|u|$. The {\em empty word}, the only word
+of null length, is denoted by $\epsilon$. We denote by $\widetilde{u}$
+the {\em mirror image} of $u$ that is the word
+$u_{|u|} u_{|u| - 1} \dots u_1$. If $P$ is a subset of $[|u|]$, $u_{|P}$
 is the subword of $u$ consisting in the letters of $u$ at the positions
-specified by the elements of $P$. The {\em shuffle} of two words $u$ and
+specified by the elements of $P$.
+If $u$ is a word of integers and $k$ is an integer, we denote by
+$u[k]$ the word obtained by incrementing by $k$ all letters of $u$.
+The {\em shuffle} of two words $u$ and
 $v$ is the set recursively defined by
@@
 multiple occurrence of a letter, so that all above definitions about
-words remain valid on permutations. If $\pi$ is a permutation, we denote
-by $\bar \pi$ (resp.~$\pi^{-1}$) the {\em complementary} (resp.
-{\em inverse}) of $\pi$. If $u$ is a word of integers without multiple
+words remain valid on permutations.
+For any nonnegative integer $n$, we write $\nearrow_{n}$
+(resp. $\searrow_{n}$)
+for the permutation $1 2 \ldots n$ (resp. $n (n-1) \ldots 1$).
+If $\pi$ is a permutation of $S_n$,
+we denote by $\bar \pi$ the {\em complementary} of $\pi$ that is the
+permutation satisfying $\bar \pi(i) = n - \pi(i) + 1$ for all
+$i \in [n]$. The {\em inverse} of $\pi$ is denoted by $\pi^{-1}$.
+
+If $u$ is a word of integers without multiple
 occurrences of a same letter, $\STD(u)$ is the {\em standarized} of $u$,
@@
 $\sigma$, $\pi$ {\em avoids} $\sigma$.
-We write $\nearrow_{n}$ (resp. $\searrow_{n}$)
-for the permutation $1 2 \ldots n$ (resp. $n (n-1) \ldots 1$).
-we write seqeunce of integers $s = s_1 s_2 \ldots s_n$, we
-we let $s\;[k]$ stand for the lifted sequence
-$(s_1+k) s_2+k) \ldots (s_n+k)$.
 
@@
     permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
-    we have $\tau_i < \tau_k$ if and only if $\bar \tau_i > \bar \tau_k$,
+    we have $\tau(i) < \tau(k)$ if and only if $\bar \tau(i) > \bar \tau(k)$,
     we have
@@
     letters of $\tau$, we denote by $P_\tau^{-1}$ the set
-    $\{\tau_i : i \in P\}$. Since the map sending a subset $P$ of
+    $\{\tau(i) : i \in P\}$. Since the map sending a subset $P$ of
     $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
@@
     \item
-    $\pi$ is a square.
+    The permutation $\pi$ is a square.
     \item
@@
   (\emph{1} $\Rightarrow$ \emph{2}).
-  Let $\sigma$ be a permutation such that $\pi$ is the union of two
-  patterns that are order-isomorphic to $\sigma$.
+  Since $\pi$ is a quare, it admits a square root $\sigma$. Then, by
+  definition, there
+  are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$ and
+  $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
+  letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
+  order-isomorphic to $\sigma$.
+  %Let $\sigma$ be a permutation such that $\pi$ is the union of two
+  %patterns that are order-isomorphic to $\sigma$.
   Let $2n = |\pi|$ (and hence $|\sigma| = n$).
-  Fix the occurrences in $\pi$ of the two disjoint patterns that
-  are order-isomorphic to $\sigma$ and
-  write
-  $I^1 = \{i^1_1, i^1_2, \ldots, i^1_n\}$,
-  $i^1_1 < i^1_2 < \ldots < i^1_n$, for the positions in $\pi$
-  of the first pattern order-ismorphic to $\sigma$ and
-  $I^2 = \{i^2_1, i^2_2, \ldots, i^2_n\}$,
-  $i^2_1 < i^2_2 < \ldots < i^2_n$, for the positions in $\pi$
-  of the second pattern order-ismorphic to $\sigma$.
-  Let $G_\pi$ the linear clique associated to $\pi$ with
+  %Fix the occurrences in $\pi$ of the two disjoint patterns that
+  %are order-isomorphic to $\sigma$ and
+  %write
+  %$I^1 = \{i^1_1, i^1_2, \ldots, i^1_n\}$,
+  %$i^1_1 < i^1_2 < \ldots < i^1_n$, for the positions in $\pi$
+  %of the first pattern order-ismorphic to $\sigma$ and
+  %$I^2 = \{i^2_1, i^2_2, \ldots, i^2_n\}$,
+  %$i^2_1 < i^2_2 < \ldots < i^2_n$, for the positions in $\pi$
+  %of the second pattern order-ismorphic to $\sigma$.
+  Let $G_\pi$ be the linear clique associated to $\pi$ with
   vertices $\VS{G_\pi} =
@@
   It is also containment-free.
-  Indeed, if it were not the case then there would exist two edges
+  Indeed, if it is not the case then there would exist two edges
   $e = \{i^1_j, i^2_j\}$ and
@@
   known to be \NPC~\cite{Bose:Buss:Lubiw:1998}:
-  Given permutation $\pi$ and $\sigma$, decide whether $\sigma$
+  Given two permutations $\pi$ and $\sigma$, decide whether $\sigma$
   occurs in $\pi$ (as an order-isomorphic pattern).
@@
       \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
-      \node (claim edge 1) at (8.5,8.5) {Claim~\ref{}};
+      \node (claim edge 1) at (8.5,8.5) {Claim~\ref{claim:(nu_4, nu'_4)-edge}};
       % sigma 1 - pi 1
       \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
-      \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
+      % \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
       % nodes
@@
 permutations of $S_{2n}$ are squares? How many $(213,231)$-avoiding
-permutations of $S_{2n}$ are squares? (Equivalently, how many binary
-strings of length $2n$ are squares; Problem~4
+permutations of $S_{2n}$ are squares? (Equivalently, by
+Proposition~\ref{prop:bijection_binary_to_permutations_squares},
+how many binary strings of length $2n$ are squares; Problem~4
 in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of

### 2015-09-17 article.tex
Modifs intro (je continue ce soir)

@@
 
-Given three words $u$, $v$ and $w$, $u$ is said to a \emph{shuffle} of
+Given three words $u$, $v$, and $w$, $u$ is said to a \emph{shuffle} of
 $v$ and $w$ if it can be formed by interleaving the letters from $v$ and
 $w$ in a way that maintains the left-to-right ordering of the letters
-from each word (it is a classical textbook exercise to design an
-efficient dynamic programming algorithm that determines whether $u$ is a
-shuffle of $v$ and $w$). A word $u$ is said to be a \emph{square} (w.r.t.
-the shuffle) if there exists a word $v$ such that $u$ is a shuffle of $v$
-with itself.
+from each word. In addition to purely combinatorial questions, the
+shuffle product of words leads to the following purely computational
+questions:
+\begin{enumerate}
+    \item \label{item:problem_1}
+    Given three words $u$, $v$, and $w$, decide if $u$ is a shuffle of
+    $v$ and $w$.
+    \item \label{item:problem_2}
+    Given a word $u$, decide if there exist two words $v$ and $w$ such
+    that $u$ is a shuffle of $v$ and $w$.
+\end{enumerate}
+Problem~\ref{item:problem_1} is in $\Pclass$ (it is a classical textbook
+exercise to design an efficient dynamic programming algorithm that
+determines whether $u$ is a shuffle of $v$ and $w$), while
+Problem~\ref{item:problem_2} is in $\NPclass$ as shown
+in~\cite{Rizzi:Vialette:CSR:2013}. Analogous questions can be stated to
+decide if a word $u$ is a \emph{square} (w.r.t. the shuffle). A square
+is a word $u$ with the property that there exists a word $v$ such that
+$u$ is a shuffle of $v$ with itself.
 \smallskip
@@
 This paper is intended to study a natural generalization of the shuffle
-to permutations. Given permutations $\pi$, $\mu$ and $\nu$, $\pi$ is
-said to be a shuffle of $\mu$ and $\nu$ if $\pi$ (viewed as a word) is
-a shuffle of two words that are order-isomorphic to $\mu$ and $\nu$. For
-example, $\pi$ is the shuffle of $\mu=$ and $\nu=$ as $\pi$. A
-permutation $\pi$ is said to be a square (w.r.t. the shuffle) if there
-exists a permutation $\mu$ such that $\pi$ is a shuffle of $\mu$ with
-itself. For example, $\pi = 241356$ is a square as it is the shuffle of
-$213$ with itself (\emph{e.g.},
-$\pi = \dot{2}{4}\dot{1}{3}{5}\dot{6}$, where the $\dot{x}$'s
-denote the second occurrence), whereas $\pi = 4123$ is not.
-The shuffle product defined and considered in this paper is different
-from other existing and well-known shuffle products on permutations.
-For instance, in~\cite{DHT:IJAC:2002}, the authors defined the
+to permutations. Roughly speaking, given three permutations $\pi$, $\mu$,
+and $\nu$, $\pi$ is said to be a shuffle of $\mu$ and $\nu$ if $\pi$
+(viewed as a word) is a shuffle of two words that are order-isomorphic
+to $\mu$ and $\nu$.
+%For
+%example, $\pi$ is the shuffle of $\mu=$ and $\nu=$ as $\pi$. A
+%permutation $\pi$ is said to be a square (w.r.t. the shuffle) if there
+%exists a permutation $\mu$ such that $\pi$ is a shuffle of $\mu$ with
+%itself. For example, $\pi = 241356$ is a square as it is the shuffle of
+%$213$ with itself (\emph{e.g.},
+%$\pi = \dot{2}{4}\dot{1}{3}{5}\dot{6}$, where the $\dot{x}$'s
+%denote the second occurrence), whereas $\pi = 4123$ is not.
+Our intention in this paper is to to study this shuffle product of
+permutations both from a combinatorial and from a computational point of
+view by focusing on {\em square} permutations, that are permutations
+$\pi$ being in the shuffle of a permutation $\mu$ with itself. The
+shuffle product defined and considered in this paper is different from
+other existing and well-known shuffle products on permutations. For
+instance, in~\cite{DHT:IJAC:2002}, the authors define the
 {\em shifted shuffle product}. In this product $\pi$ is a shuffle of
@@
 word obtained by incrementing all letters of $\nu$ by the size of $\mu$.
-It is a simple exercise to prove that, give a permutation $\pi$, the
-decision problem to test if $\pi$ is a square for the shifted shuffle
-product is linear.
+It is a simple exercise to prove that, given three permutations $\pi$,
+$\mu$, and $\nu$, deciding if $\pi$ is a in the shifted shuffle of
+$\mu$ and $\nu$ is in $\Pclass$.
 \smallskip
 
-Formally, the shuffle of two words $u$ and $v$ of $A^*$ is the language
-$u \shuffle v$ consisting of all words $u_1 v_1 u_2 v_2 \ldots u_k v_k$,
-where $k \geq 0$ and the $u_i$ and the $v_i$ are the words  of $A^*$
-such that $u = u_1 u_2 \ldots u_k$ and $v = v_1 v_2 \ldots v_k$. We
-first briefly review the key results for words that arise in our context.
-Given words $u$, $v_1$ and $v_2$, it can be tested in
+Let us now  briefly review the key results for words that arise in our context.
+%Formally, the shuffle of two words $u$ and $v$ of $A^*$ is the language
+%$u \shuffle v$ consisting of all words $u_1 v_1 u_2 v_2 \ldots u_k v_k$,
+%where $k \geq 0$ and the $u_i$ and the $v_i$ are the words  of $A^*$
+%such that $u = u_1 u_2 \ldots u_k$ and $v = v_1 v_2 \ldots v_k$. We
+%first briefly review the key results for words that arise in our context.
+Given three words $u$, $v_1$ and $v_2$, it can be tested in
 $O\left(|u|^2 / \log(|u|)\right)$ time whether or not
-$u \in v_1 \shuffle v_2$~\cite{Leeuwen:Nivat:IPL:1982}.
-(To the best of our knowledge, the first $O(|u|^2)$ time algorithm a
-ppeared in~\cite{Mansfield:DAM:1983}. This algorithm can easily be
+$u$ is in $v_1 \shuffle v_2$~\cite{Leeuwen:Nivat:IPL:1982}.
+To the best of our knowledge, the first $O(|u|^2)$ time algorithm
+appeared in~\cite{Mansfield:DAM:1983}. This algorithm can easily be
 extended to check in polynomial-time whether or not a word is the
-shuffle of any fixed number of given words.) The shuffle $u \shuffle v$
+shuffle of any fixed number of given words. The shuffle $u \shuffle v$
 of words $u$ and $v$ can be computed in
-$O\left((|u|+|v|) \; \binom{|u|+|v|}{|u|}\right)$ time~\cite{Spehner:TCS:1986}.
+\begin{equation}
+    O\left((|u|+|v|) \; \binom{|u|+|v|}{|u|}\right)
+\end{equation}
+time~\cite{Spehner:TCS:1986}.
 An improvement and generalization has been proposed in
@@
 computed in
-$O\left(\binom{|u_1|+|u_2|+\ldots+|u_k|}{|u_1|,|u_2|,\ldots,|u_k|}\right)$
+\begin{equation}
+    O\left(\binom{|u_1|+|u_2|+\ldots+|u_k|}{|u_1|,|u_2|,\ldots,|u_k|}\right)
+\end{equation}
 time. Given words $u, v_1, v_2, \ldots, v_n \in A^*$, it is however \NPC
@@
 
+This paper is organized as follows...
+
 $123$-avoiding (resp. $321$-avoiding) permutations are the union of two decreasing
@@
 multiple occurrence of a letter, so that all above definitions about
-words remain valid on permutations.
+words remain valid on permutations. The only difference lies on the
+fact that we shall denote by $\pi(i)$ (instead of $\pi_i$) the
+$i$-th letter of any permutation $\pi$.
 For any nonnegative integer $n$, we write $\nearrow_{n}$

### 2015-09-17 article.tex
Suite modifs intro

@@
 from each word. In addition to purely combinatorial questions, the
-shuffle product of words leads to the following purely computational
-questions:
+shuffle product of words leads to the following computational questions:
 \begin{enumerate}
@@
 \end{enumerate}
-Problem~\ref{item:problem_1} is in $\Pclass$ (it is a classical textbook
-exercise to design an efficient dynamic programming algorithm that
-determines whether $u$ is a shuffle of $v$ and $w$), while
-Problem~\ref{item:problem_2} is in $\NPclass$ as shown
+Problem~\ref{item:problem_1} is in $\Pclass$ (it is a classical
+textbook exercise to design an efficient dynamic programming algorithm
+that determines whether $u$ is a shuffle of $v$ and $w$), while
+Problem~\ref{item:problem_2} is $\NPclass$-complete as shown
 in~\cite{Rizzi:Vialette:CSR:2013}. Analogous questions can be stated to
-decide if a word $u$ is a \emph{square} (w.r.t. the shuffle). A square
-is a word $u$ with the property that there exists a word $v$ such that
+decide if a word $u$ is a \emph{square} (w.r.t. the shuffle), that is
+a word $u$ with the property that there exists a word $v$ such that
 $u$ is a shuffle of $v$ with itself.
@@
 
-This paper is intended to study a natural generalization of the shuffle
-to permutations. Roughly speaking, given three permutations $\pi$, $\mu$,
-and $\nu$, $\pi$ is said to be a shuffle of $\mu$ and $\nu$ if $\pi$
+This paper is intended to study a natural generalization $\SHUFFLE$
+of the shuffle
+of permutations. Roughly speaking, given three permutations $\pi$, $\mu$,
+and $\nu$, $\pi$ is said to be a {\em shuffle} of $\mu$ and $\nu$ if $\pi$
 (viewed as a word) is a shuffle of two words that are order-isomorphic
@@
 %denote the second occurrence), whereas $\pi = 4123$ is not.
-Our intention in this paper is to to study this shuffle product of
+Our intention in this paper is to study this shuffle product of
 permutations both from a combinatorial and from a computational point of
@@
 instance, in~\cite{DHT:IJAC:2002}, the authors define the
-{\em shifted shuffle product}. In this product $\pi$ is a shuffle of
-$\mu$ and $\nu$ if $\pi$ is the shuffle, as words, of $\mu$ and the
-word obtained by incrementing all letters of $\nu$ by the size of $\mu$.
+{\em shifted shuffle product}. For this product, $\pi$ is a shuffle of
+$\mu$ and $\nu$ if $\pi$ is in the shuffle, as words, of $\mu$ and the
+word obtained by incrementing all the letters of $\nu$ by the size of $\mu$.
 It is a simple exercise to prove that, given three permutations $\pi$,
@@
 %first briefly review the key results for words that arise in our context.
-Given three words $u$, $v_1$ and $v_2$, it can be tested in
+Given three words $u$, $v_1$, and $v_2$, it can be tested in
 $O\left(|u|^2 / \log(|u|)\right)$ time whether or not
-$u$ is in $v_1 \shuffle v_2$~\cite{Leeuwen:Nivat:IPL:1982}.
+$u$ is in the shuffle $v_1 \shuffle v_2$ of $v_1$ and
+$v_2$~\cite{Leeuwen:Nivat:IPL:1982}.
 To the best of our knowledge, the first $O(|u|^2)$ time algorithm
@@
 \cite{Allauzen:IGM:2000}, where it is proved that, given words
-$u_1, u_2, \ldots, u_k$, the shuffle $\shuffle_{i=1}^{k} u_i$ can be
+$u_1$, $u_2$, \dots, $u_k$, the shuffle $\shuffle_{i=1}^{k} u_i$ can be
 computed in
@@
 \end{equation}
-time. Given words $u, v_1, v_2, \ldots, v_n \in A^*$, it is however \NPC
-to decide whether or not $u \in \shuffle_{i=1}^{k} v_i$.
+time. Given words $u$, $v_1$, $v_2$, \dots, $v_n$, it is however \NPC
+to decide whether or not $u$ is in $\shuffle_{i=1}^{k} v_i$.
 \cite{Mansfield:DAM:1983,Warmuth:Haussler:JCSS:1984}.
-This remains true even if the alphabet has size $3$~\cite{Warmuth:Haussler:JCSS:1984}.
+This remains true even if the ground alphabet has size $3$~\cite{Warmuth:Haussler:JCSS:1984}.
 Of particular interest,
@@
 is in the iterated shuffle of $v$.
-Again, this remains true even if the alphabet has size $3$.
+Again, this remains true even if the ground alphabet has size $3$.
 As for squares, given a word $z$ it is \NPC to determine whether
@@
 
-This paper is organized as follows...
+This paper is organized as follows. In
+Section~\ref{section:Shuffle product on permutations} we provide a precise
+definition of $\SHUFFLE$. This definition passes
+through the preliminary definition of an operator $\Delta$ to unshuffle
+permutations. This operator is in fact a coproduct, endowing the space
+of all permutations with a coalgebra structure. By duality, the
+unshuffling operator $\Delta$ leads to the definition of our shuffle operation on
+permutations. This approach has many advantages. First, some combinatorial
+properties of $\SHUFFLE$ depend on properties of $\Delta$
+and are more easy to prove on the coproduct
+side. Second, this way of doing allows to obtain a clear description of
+the multiplicities of the elements appearing in the shuffle of two
+permutations, which are worthy of interest.
+Section~\ref{section:Binary square words and permutations}
+is devoted to show that the problems related to the shuffle of words has
+links with the shuffle of permutations. In particular, we show that
+binary words that are square are in one-to-one bijection with square
+permutations (Proposition~\ref{prop:bijection_binary_to_permutations_squares}).
+Next, Section~\ref{section:Algebraic issues} presents some algebraic
+and combinatorial properties of~$\SHUFFLE$. In particular, we show that
+$\SHUFFLE$ is associative and commutative
+(Proposition~\ref{prop:shuffle_associative_commutative}), and that
+if a permutation if a square, its mirror, complement, and inverse also
+are square (Proposition~\ref{prop:square_stability}). Finally,
+Section~\ref{section:Algorithmic issues} presents
+the most import result of this paper: the fact that deciding if
+a permutation is a square is $\NPclass$-complete
+(Proposition~\ref{proposition:hardness}).
 
@@
 (Problem~4 in \cite{Henshall:Rampersad:Shallit:2011},
-see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?)
+see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?
 % Given two

### 2015-09-17 article.tex
Reorganisation intro

@@
 \label{section:Introduction}
-
-Given three words $u$, $v$, and $w$, $u$ is said to a \emph{shuffle} of
-$v$ and $w$ if it can be formed by interleaving the letters from $v$ and
-$w$ in a way that maintains the left-to-right ordering of the letters
-from each word. In addition to purely combinatorial questions, the
-shuffle product of words leads to the following computational questions:
+The {\em shuffle product} $\shuffle$ is a very famous operation on words
+first defined by Eilenberg and MacLane~\cite{Eilenberg:MacLane:1953}.
+Given three words $u$, $v$, and $w$, $u$ is said to be a \emph{shuffle}
+of $v$ and $w$ if it can be formed by interleaving the letters from $v$
+and $w$ in a way that maintains the left-to-right ordering of the
+letters from each word. In addition to purely combinatorial questions,
+the shuffle product of words leads to the following computational ones:
 \begin{enumerate}
     \item \label{item:problem_1}
-    Given three words $u$, $v$, and $w$, decide if $u$ is a shuffle of
-    $v$ and $w$.
+    Given two words $v_1$ and $v_2$, compute the set $v_1 \shuffle v_2$.
     \item \label{item:problem_2}
-    Given a word $u$, decide if there exist two words $v$ and $w$ such
-    that $u$ is a shuffle of $v$ and $w$.
+    Given three words $u$, $v_1$, and $v_2$, decide if $u$ is a shuffle
+    of $v_1$ and $v_2$.
+    \item \label{item:problem_3}
+    Given words $u$, $v_1$, \dots, $v_k$, decide if $u$ is in
+    $v_1 \shuffle \dots \shuffle v_k$.
+    \item \label{item:problem_4}
+    Given a word $u$, decide if there exist two words $v_1$ and $v_2$
+    such that $u$ is a shuffle of $v_1$ and $v_2$.
 \end{enumerate}
-Problem~\ref{item:problem_1} is in $\Pclass$ (it is a classical
-textbook exercise to design an efficient dynamic programming algorithm
-that determines whether $u$ is a shuffle of $v$ and $w$), while
-Problem~\ref{item:problem_2} is $\NPclass$-complete as shown
-in~\cite{Rizzi:Vialette:CSR:2013}. Analogous questions can be stated to
+Even if these problems seem similar, they are very different in terms
+of complexity. Let us now review some facts about these. In what follows,
+$n$ denotes the size of $u$ and $m_i$ denotes the size of each $v_i$.
+A solution to Problem~\ref{item:problem_1} can be computed in
+\begin{equation}
+    O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)
+\end{equation}
+time~\cite{Spehner:TCS:1986}. An improvement and a generalization
+of~Problem~\ref{item:problem_1} has been proposed
+in~\cite{Allauzen:IGM:2000}, where it is proved that, given words
+$v_1$, \dots, $v_k$, the iterated shuffle
+$v_1 \shuffle \dots \shuffle v_k$ can be computed in
+\begin{equation}
+    O\left(\binom{m_1 + \dots + m_k}{m_1, \dots, m_k}\right)
+\end{equation}
+time. Problem~\ref{item:problem_2} is in \Pclass; it is indeed a
+classical textbook exercise to design an efficient dynamic programming
+algorithm solving it. It can be tested in $O\left(n^2 / \log(n)\right)$
+time~\cite{Leeuwen:Nivat:IPL:1982}. To the best of our knowledge, the
+first $O(n^2)$ time algorithm for this problem appeared
+in~\cite{Mansfield:DAM:1983}. This algorithm can easily be extended to
+check in polynomial-time whether or not a word is in the shuffle of any
+fixed number of given words. Nevertheless, Problem~\ref{item:problem_3}
+is \NPC~\cite{Mansfield:DAM:1983,Warmuth:Haussler:JCSS:1984}. This
+remains true even if the ground alphabet has size
+$3$~\cite{Warmuth:Haussler:JCSS:1984}. Of particular interest,
+it is shown in~\cite{Warmuth:Haussler:JCSS:1984} that
+Problem~\ref{item:problem_3} remains \NPC even if all the words $v_i$,
+$i \in [k]$, are identical, thereby proving that, for two words $u$ and
+$v$, it is \NPC to decide whether or not $u$ is in the iterated shuffle
+of $v$. Again, this remains true even if the ground alphabet has size $3$.
+Let us now focus on Problem~\ref{item:problem_4}. It is shown
+in~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013} that it is \NPC to
 decide if a word $u$ is a \emph{square} (w.r.t. the shuffle), that is
-a word $u$ with the property that there exists a word $v$ such that
-$u$ is a shuffle of $v$ with itself.
+a word $u$ with the property that there exists a word $v$ such that $u$
+is a shuffle of $v$ with itself. The hardness of this last particular
+problem implies that Problem~\ref{item:problem_4} is \NPC.
 \smallskip
 
-This paper is intended to study a natural generalization $\SHUFFLE$
-of the shuffle
-of permutations. Roughly speaking, given three permutations $\pi$, $\mu$,
-and $\nu$, $\pi$ is said to be a {\em shuffle} of $\mu$ and $\nu$ if $\pi$
-(viewed as a word) is a shuffle of two words that are order-isomorphic
-to $\mu$ and $\nu$.
-%For
-%example, $\pi$ is the shuffle of $\mu=$ and $\nu=$ as $\pi$. A
-%permutation $\pi$ is said to be a square (w.r.t. the shuffle) if there
-%exists a permutation $\mu$ such that $\pi$ is a shuffle of $\mu$ with
-%itself. For example, $\pi = 241356$ is a square as it is the shuffle of
-%$213$ with itself (\emph{e.g.},
-%$\pi = \dot{2}{4}\dot{1}{3}{5}\dot{6}$, where the $\dot{x}$'s
-%denote the second occurrence), whereas $\pi = 4123$ is not.
+This paper is intended to study a natural generalization of $\shuffle$,
+denoted by $\SHUFFLE$, as a shuffle of permutations. Roughly speaking,
+given three permutations $\pi$, $\mu$, and $\nu$, $\pi$ is said to be a
+{\em shuffle} of $\mu$ and $\nu$ if $\pi$ (viewed as a word) is a
+shuffle of two words that are order-isomorphic to $\mu$ and $\nu$.
 Our intention in this paper is to study this shuffle product of
-permutations both from a combinatorial and from a computational point of
-view by focusing on {\em square} permutations, that are permutations
-$\pi$ being in the shuffle of a permutation $\mu$ with itself. The
-shuffle product defined and considered in this paper is different from
-other existing and well-known shuffle products on permutations. For
-instance, in~\cite{DHT:IJAC:2002}, the authors define the
-{\em shifted shuffle product}. For this product, $\pi$ is a shuffle of
-$\mu$ and $\nu$ if $\pi$ is in the shuffle, as words, of $\mu$ and the
-word obtained by incrementing all the letters of $\nu$ by the size of $\mu$.
-It is a simple exercise to prove that, given three permutations $\pi$,
-$\mu$, and $\nu$, deciding if $\pi$ is a in the shifted shuffle of
-$\mu$ and $\nu$ is in $\Pclass$.
+permutations $\SHUFFLE$ both from a combinatorial and from a
+computational point of view by focusing on {\em square} permutations,
+that are permutations $\pi$ being in the shuffle of a permutation $\mu$
+with itself. The shuffle product defined and considered in this paper
+is different from other existing and well-known shuffle products on
+permutations. For instance, in~\cite{DHT:IJAC:2002}, the authors define
+the {\em shifted shuffle product}. For this product, $\pi$ is a shuffle
+of $\mu$ and $\nu$ if $\pi$ is in the shuffle, as words, of $\mu$ and
+the word obtained by incrementing all the letters of $\nu$ by the size
+of $\mu$. It is a simple exercise to prove that, given three
+permutations $\pi$, $\mu$, and $\nu$, deciding if $\pi$ is a in the
+shifted shuffle of $\mu$ and $\nu$ is in~\Pclass.
 \smallskip
 
-Let us now  briefly review the key results for words that arise in our context.
-%Formally, the shuffle of two words $u$ and $v$ of $A^*$ is the language
-%$u \shuffle v$ consisting of all words $u_1 v_1 u_2 v_2 \ldots u_k v_k$,
-%where $k \geq 0$ and the $u_i$ and the $v_i$ are the words  of $A^*$
-%such that $u = u_1 u_2 \ldots u_k$ and $v = v_1 v_2 \ldots v_k$. We
-%first briefly review the key results for words that arise in our context.
-Given three words $u$, $v_1$, and $v_2$, it can be tested in
-$O\left(|u|^2 / \log(|u|)\right)$ time whether or not
-$u$ is in the shuffle $v_1 \shuffle v_2$ of $v_1$ and
-$v_2$~\cite{Leeuwen:Nivat:IPL:1982}.
-To the best of our knowledge, the first $O(|u|^2)$ time algorithm
-appeared in~\cite{Mansfield:DAM:1983}. This algorithm can easily be
-extended to check in polynomial-time whether or not a word is the
-shuffle of any fixed number of given words. The shuffle $u \shuffle v$
-of words $u$ and $v$ can be computed in
-\begin{equation}
-    O\left((|u|+|v|) \; \binom{|u|+|v|}{|u|}\right)
-\end{equation}
-time~\cite{Spehner:TCS:1986}.
-An improvement and generalization has been proposed in
-\cite{Allauzen:IGM:2000}, where it is proved that, given words
-$u_1$, $u_2$, \dots, $u_k$, the shuffle $\shuffle_{i=1}^{k} u_i$ can be
-computed in
-\begin{equation}
-    O\left(\binom{|u_1|+|u_2|+\ldots+|u_k|}{|u_1|,|u_2|,\ldots,|u_k|}\right)
-\end{equation}
-time. Given words $u$, $v_1$, $v_2$, \dots, $v_n$, it is however \NPC
-to decide whether or not $u$ is in $\shuffle_{i=1}^{k} v_i$.
-\cite{Mansfield:DAM:1983,Warmuth:Haussler:JCSS:1984}.
-This remains true even if the ground alphabet has size $3$~\cite{Warmuth:Haussler:JCSS:1984}.
-Of particular interest,
-it is shown in~\cite{Warmuth:Haussler:JCSS:1984} that this problem remains
-\NPC even if all
-words $v_1, v_2, \ldots, v_n$ are identical,
-thereby proving that,
-for two words $u$ and $v$, it is \NPC to decide whether or not $u$
-is in the iterated shuffle of $v$.
-Again, this remains true even if the ground alphabet has size $3$.
-As for squares, given a word $z$ it is \NPC to determine whether
-there exists a word $x$ such that $z$ is the shuffle of $x$ with itself
-\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}.
-\smallskip
-
 This paper is organized as follows. In
-Section~\ref{section:Shuffle product on permutations} we provide a precise
-definition of $\SHUFFLE$. This definition passes
-through the preliminary definition of an operator $\Delta$ to unshuffle
-permutations. This operator is in fact a coproduct, endowing the space
-of all permutations with a coalgebra structure. By duality, the
-unshuffling operator $\Delta$ leads to the definition of our shuffle operation on
-permutations. This approach has many advantages. First, some combinatorial
-properties of $\SHUFFLE$ depend on properties of $\Delta$
-and are more easy to prove on the coproduct
-side. Second, this way of doing allows to obtain a clear description of
-the multiplicities of the elements appearing in the shuffle of two
-permutations, which are worthy of interest.
-Section~\ref{section:Binary square words and permutations}
+Section~\ref{section:Shuffle product on permutations} we provide a
+precise definition of $\SHUFFLE$. This definition passes through the
+preliminary definition of an operator $\Delta$ to unshuffle permutations.
+This operator is in fact a coproduct, endowing the space of all
+permutations with a coalgebra structure. By duality, the unshuffling
+operator $\Delta$ leads to the definition of our shuffle operation on
+permutations. This approach has many advantages. First, some
+combinatorial properties of $\SHUFFLE$ depend on properties of $\Delta$
+and are more easy to prove on the coproduct side. Second, this way of
+doing allows to obtain a clear description of the multiplicities of the
+elements appearing in the shuffle of two permutations, which are worthy
+of interest. Section~\ref{section:Binary square words and permutations}
 is devoted to show that the problems related to the shuffle of words has
@@
 binary words that are square are in one-to-one bijection with square
-permutations (Proposition~\ref{prop:bijection_binary_to_permutations_squares}).
-Next, Section~\ref{section:Algebraic issues} presents some algebraic
-and combinatorial properties of~$\SHUFFLE$. In particular, we show that
+permutations
+(Proposition~\ref{prop:bijection_binary_to_permutations_squares}). Next,
+Section~\ref{section:Algebraic issues} presents some algebraic and
+combinatorial properties of~$\SHUFFLE$. In particular, we show that
 $\SHUFFLE$ is associative and commutative
@@
 Section~\ref{section:Algorithmic issues} presents
-the most import result of this paper: the fact that deciding if
-a permutation is a square is $\NPclass$-complete
-(Proposition~\ref{proposition:hardness}).
-
-$123$-avoiding (resp. $321$-avoiding) permutations are the union of two decreasing
-(resp. increasing) patterns.
+the most import result of this paper: the fact that deciding if a
+permutation is a square is \NPC (Proposition~\ref{proposition:hardness}).
 

### 2015-09-17 article.tex
Diverses petites modif

@@
 The {\em shuffle product} $\shuffle$ is a very famous operation on words
-first defined by Eilenberg and MacLane~\cite{Eilenberg:MacLane:1953}.
-Given three words $u$, $v$, and $w$, $u$ is said to be a \emph{shuffle}
-of $v$ and $w$ if it can be formed by interleaving the letters from $v$
-and $w$ in a way that maintains the left-to-right ordering of the
+first defined by Eilenberg and Mac Lane~\cite{Eilenberg:MacLane:1953}.
+Given three words $u$, $v_1$, and $v_2$, $u$ is said to be a \emph{shuffle}
+of $v_1$ and $v_2$ if it can be formed by interleaving the letters from
+$v_1$ and $v_2$ in a way that maintains the left-to-right ordering of the
 letters from each word. In addition to purely combinatorial questions,
@@
 denoted by $\SHUFFLE$, as a shuffle of permutations. Roughly speaking,
-given three permutations $\pi$, $\mu$, and $\nu$, $\pi$ is said to be a
-{\em shuffle} of $\mu$ and $\nu$ if $\pi$ (viewed as a word) is a
-shuffle of two words that are order-isomorphic to $\mu$ and $\nu$.
-Our intention in this paper is to study this shuffle product of
-permutations $\SHUFFLE$ both from a combinatorial and from a
-computational point of view by focusing on {\em square} permutations,
-that are permutations $\pi$ being in the shuffle of a permutation $\mu$
-with itself. The shuffle product defined and considered in this paper
-is different from other existing and well-known shuffle products on
-permutations. For instance, in~\cite{DHT:IJAC:2002}, the authors define
-the {\em shifted shuffle product}. For this product, $\pi$ is a shuffle
-of $\mu$ and $\nu$ if $\pi$ is in the shuffle, as words, of $\mu$ and
-the word obtained by incrementing all the letters of $\nu$ by the size
-of $\mu$. It is a simple exercise to prove that, given three
-permutations $\pi$, $\mu$, and $\nu$, deciding if $\pi$ is a in the
-shifted shuffle of $\mu$ and $\nu$ is in~\Pclass.
+given three permutations $\pi$, $\sigma_1$, and $\sigma_2$, $\pi$ is
+said to be a {\em shuffle} of $\sigma_1$ and $\sigma_2$ if $\pi$ (viewed
+as a word) is a shuffle of two words that are order-isomorphic to
+$\sigma_1$ and $\sigma_2$. Our intention in this paper is to study this
+shuffle product of permutations $\SHUFFLE$ both from a combinatorial and
+from a computational point of view by focusing on {\em square}
+permutations, that are permutations $\pi$ being in the shuffle of a
+permutation $\sigma$ with itself. The shuffle product defined and
+considered in this paper is different from other existing and well-known
+shuffle products on permutations. For instance, in~\cite{DHT:IJAC:2002},
+the authors define the {\em shifted shuffle product}. For this product,
+$\pi$ is a shuffle of $\sigma_1$ and $\sigma_2$ if $\pi$ is in the
+shuffle, as words, of $\sigma_1$ and the word obtained by incrementing
+all the letters of $\sigma_2$ by the size of $\sigma_1$. It is a simple
+exercise to prove that, given three permutations $\pi$, $\sigma_1$, and
+$\sigma_2$, deciding if $\pi$ is a in the shifted shuffle of $\sigma_1$
+and $\sigma_2$ is in~\Pclass.
 \smallskip
@@
 precise definition of $\SHUFFLE$. This definition passes through the
-preliminary definition of an operator $\Delta$ to unshuffle permutations.
-This operator is in fact a coproduct, endowing the space of all
-permutations with a coalgebra structure. By duality, the unshuffling
+preliminary definition of an operator~$\Delta$, allowing to
+{\em unshuffle} permutations. This operator is in fact a coproduct,
+endowing the space of all permutations with a coalgebra structure
+(see~\cite{Joni:Rota:1979} or~\cite{Grinberg:Reiner:2014} for the
+definition of these algebraic structures). By duality, the unshuffling
 operator $\Delta$ leads to the definition of our shuffle operation on
@@
 elements appearing in the shuffle of two permutations, which are worthy
-of interest. Section~\ref{section:Binary square words and permutations}
-is devoted to show that the problems related to the shuffle of words has
-links with the shuffle of permutations. In particular, we show that
-binary words that are square are in one-to-one bijection with square
-permutations
+of interest from a combinatorial point of view.
+Section~\ref{section:Binary square words and permutations} is devoted to
+show that the problems related to the shuffle of words has links with
+the shuffle of permutations. In particular, we show that binary words
+that are square are in one-to-one bijection with square permutations
+avoiding some patterns
 (Proposition~\ref{prop:bijection_binary_to_permutations_squares}). Next,
 Section~\ref{section:Algebraic issues} presents some algebraic and
-combinatorial properties of~$\SHUFFLE$. In particular, we show that
-$\SHUFFLE$ is associative and commutative
-(Proposition~\ref{prop:shuffle_associative_commutative}), and that
-if a permutation if a square, its mirror, complement, and inverse also
-are square (Proposition~\ref{prop:square_stability}). Finally,
-Section~\ref{section:Algorithmic issues} presents
-the most import result of this paper: the fact that deciding if a
-permutation is a square is \NPC (Proposition~\ref{proposition:hardness}).
+combinatorial properties of~$\SHUFFLE$. We show that $\SHUFFLE$ is
+associative and commutative
+(Proposition~\ref{prop:shuffle_associative_commutative}), and that if a
+permutation is a square, its mirror, complement, and inverse are also
+squares (Proposition~\ref{prop:square_stability}). Finally,
+Section~\ref{section:Algorithmic issues} presents the most import result
+of this paper: the fact that deciding if a permutation is a square is
+\NPC (Proposition~\ref{proposition:hardness}).
 
@@
 
-\begin{proposition} \label{prop:endomorphisms}
+\begin{lemma} \label{lem:endomorphisms}
     The three linear maps
@@
     associative algebras.
-\end{proposition}
-\begin{proof}[of Proposition~\ref{prop:endomorphisms}]
+\end{lemma}
+\begin{proof}[of Lemma~\ref{lem:endomorphisms}]
     To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
@@
 We now use the algebraic properties of $\SHUFFLE$ exhibited by
-Proposition~\ref{prop:endomorphisms} to obtain combinatorial properties
+Lemma~\ref{lem:endomorphisms} to obtain combinatorial properties
 of square permutations.
@@
     \ref{item:square_stability_2}, and~\ref{item:square_stability_3} are
-    consequences of Proposition~\ref{prop:endomorphisms}. Indeed,
+    consequences of Lemma~\ref{lem:endomorphisms}. Indeed,
     since $\pi$ is a square permutation and $\rho$ is a square root of
     $\pi$, by definition, $\pi$ appears in the product
-    $\rho \SHUFFLE \rho$. Now, by Proposition~\ref{prop:endomorphisms},
+    $\rho \SHUFFLE \rho$. Now, by Lemma~\ref{lem:endomorphisms},
     for any $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative

### 2015-09-18 article.tex
Improvements

@@
 is the subword of $u$ consisting in the letters of $u$ at the positions
-specified by the elements of $P$.
-If $u$ is a word of integers and $k$ is an integer, we denote by
-$u[k]$ the word obtained by incrementing by $k$ all letters of $u$.
-The {\em shuffle} of two words $u$ and
-$v$ is the set recursively defined by
+specified by the elements of $P$. If $u$ is a word of integers and $k$
+is an integer, we denote by $u[k]$ the word obtained by incrementing by
+$k$ all letters of $u$. The {\em shuffle} of two words $u$ and $v$ is
+the set recursively defined by
 $u \shuffle \epsilon = u = \epsilon \shuffle u$ and
@@
 We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
-set of all permutations. In this paper, permutations of a size $n$
-are specified by words of length $n$ on the alphabet $[n]$ and without
+set of all permutations. In this paper, permutations of a size $n$ are
+specified by words of length $n$ on the alphabet $[n]$ and without
 multiple occurrence of a letter, so that all above definitions about
 words remain valid on permutations. The only difference lies on the
-fact that we shall denote by $\pi(i)$ (instead of $\pi_i$) the
-$i$-th letter of any permutation $\pi$.
-For any nonnegative integer $n$, we write $\nearrow_{n}$
-(resp. $\searrow_{n}$)
-for the permutation $1 2 \ldots n$ (resp. $n (n-1) \ldots 1$).
-If $\pi$ is a permutation of $S_n$,
-we denote by $\bar \pi$ the {\em complementary} of $\pi$ that is the
-permutation satisfying $\bar \pi(i) = n - \pi(i) + 1$ for all
+fact that we shall denote by $\pi(i)$ (instead of $\pi_i$) the $i$-th
+letter of any permutation $\pi$. For any nonnegative integer $n$, we
+write $\nearrow_{n}$ (resp. $\searrow_{n}$) for the permutation
+$1 2 \dots n$ (resp. $n (n-1) \dots 1$). If $\pi$ is a permutation of
+$S_n$, we denote by $\bar \pi$ the {\em complementary} of $\pi$ that is
+the permutation satisfying $\bar \pi(i) = n - \pi(i) + 1$ for all
 $i \in [n]$. The {\em inverse} of $\pi$ is denoted by $\pi^{-1}$.
 
-If $u$ is a word of integers without multiple
-occurrences of a same letter, $\STD(u)$ is the {\em standarized} of $u$,
-that is the unique permutation of the same size as $u$ such that for all
-$i, j \in [|u|]$, $u_i < u_j$ if and only if $\STD(u)_i < \STD(u)_j$. In
-particular, the image of the map $\STD$ is the set $S$ of all
-permutations. Two words $u$ and $v$ having the same standarized are
-{\em order-isomorphic}. If $\sigma$ is a permutation, there is an
-{\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if there is a
-set $P$ of indexes of letters of $\pi$ such that $\sigma$ and $\pi_{|P}$
-are order-isomorphic. When $\pi$ does not admit any occurrence of
-$\sigma$, $\pi$ {\em avoids} $\sigma$.
-
-For a graph $G$, we denote $\VS{G}$ as the set of vertices and
-$\ES{G}$ as the set of edges.
-Let $u = u_1 \, u_2 \,\ldots\, u_n$ be a word.
-The \emph{clique graph associated to $u$}, denoted $K_u$,
-is the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$.
-The structure of this underlying graph is linear,
-\emph{i.e.}, the set of vertices is equipped with a natural total order
-$<$ defined by $u_i < u_j$ if and only if $i < j$.
-For a linear graph, we write $(u_i, u_j)$ instead of $\{u_i, u_j\}$
-for an edge if it is clear from the context that $i < j$.
-Recall that two edges of a graph are \emph{independent} if they do not share
-a common vertex, and that a \emph{matching} $\mathcal{M}$ in $G$ is a set of
-pairwise independent edges.
-A matching is \emph{perfect} if it covers all the vertices of the graph.
-In case the set of vertices is equipped with a total order,
-a matching $\mathcal{M}$ is said to be \emph{containment-free}
-if there do not exist (independent) edges $(u_i, u_j)$ and
-$(u_k, u_\ell)$ in $\mathcal{M}$ such that
-$u_i < u_k < u_\ell < u_j$ or  $u_k < u_i < u_j < u_\ell$.
+If $u$ is a word of integers without multiple occurrences of a same
+letter, $\STD(u)$ is the {\em standarized} of $u$, that is the unique
+permutation of the same size as $u$ such that for all $i, j \in [|u|]$,
+$u_i < u_j$ if and only if $\STD(u)(i) < \STD(u)(j)$. In particular, the
+image of the map $\STD$ is the set $S$ of all permutations. Two words $u$
+and $v$ having the same standarized are {\em order-isomorphic}. If
+$\sigma$ is a permutation, there is an {\em occurrence} of (the
+{\em pattern}) $\sigma$ in $\pi$ if there is a set $P$ of indexes of
+letters of $\pi$ such that $\sigma$ and $\pi_{|P}$ are order-isomorphic.
+When $\pi$ does not admit any occurrence of $\sigma$, $\pi$ {\em avoids}
+$\sigma$.
+
+Let us now provide some definitions about graphs and matching, used in
+the sequel. For a graph $G$, we denote $\VS{G}$ as the set of vertices
+and $\ES{G}$ as the set of edges of $G$. Let $u = u_1 u_2 \dots u_n$ be
+a word. The \emph{clique graph associated to $u$}, denoted by $K_u$, is
+the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$. The
+structure of this underlying graph is linear, \emph{i.e.}, the set of
+its vertices is equipped with a natural total order $<$ defined by
+$u_i < u_j$ if and only if $i < j$. For a linear graph, we write
+$(u_i, u_j)$ instead of $\{u_i, u_j\}$ for an edge if it is clear from
+the context that $i < j$. Recall that two edges of a graph are
+\emph{independent} if they do not share a common vertex, and that a
+\emph{matching} $\mathcal{M}$ in $G$ is a set of pairwise independent
+edges. A matching is \emph{perfect} if it covers all the vertices of the
+graph. In case the set of vertices is equipped with a total order, a
+matching $\mathcal{M}$ is said to be \emph{containment-free} if there do
+not exist (independent) edges $(u_i, u_j)$ and $(u_k, u_\ell)$ in
+$\mathcal{M}$ such that $u_i < u_k < u_\ell < u_j$ or
+$u_k < u_i < u_j < u_\ell$.
 
@@
         words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
-        same standarized $\rho$. Hence, and by definition of the shuffle
-        product of permutations, $\BINTOPERM(u)$ appears in
-        $\rho \SHUFFLE \rho$, showing that $\BINTOPERM(u)$ is a square
-        permutation.
+        same standarized $\sigma$. Hence, and by definition of
+        the shuffle product of permutations, $\BINTOPERM(u)$ appears in
+        $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
+        square permutation.
         \qed
@@
 \begin{proposition} \label{prop:square_stability}
-    Let $\pi$ be a square permutation and $\rho$ be a square root of
+    Let $\pi$ be a square permutation and $\sigma$ be a square root of
     $\pi$. Then,
@@
         the permutation $\widetilde{\pi}$ is a square and
-        $\widetilde{\rho}$ is one of its square roots;
+        $\widetilde{\sigma}$ is one of its square roots;
         \item \label{item:square_stability_2}
-        the permutation $\bar \pi$ is a square and $\bar \rho$ is one of
+        the permutation $\bar \pi$ is a square and $\bar \sigma$ is one of
         its square roots;
         \item \label{item:square_stability_3}
-        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$ is one of
+        the permutation $\pi^{-1}$ is a square and $\sigma^{-1}$ is one of
         its square roots.
@@
     consequences of Lemma~\ref{lem:endomorphisms}. Indeed,
-    since $\pi$ is a square permutation and $\rho$ is a square root of
+    since $\pi$ is a square permutation and $\sigma$ is a square root of
     $\pi$, by definition, $\pi$ appears in the product
-    $\rho \SHUFFLE \rho$. Now, by Lemma~\ref{lem:endomorphisms},
+    $\sigma \SHUFFLE \sigma$. Now, by Lemma~\ref{lem:endomorphisms},
     for any $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
@@
     \begin{equation}
-        \phi_j(\rho \SHUFFLE \rho) =
-        \phi_j(\rho) \SHUFFLE \phi_j(\rho).
+        \phi_j(\sigma \SHUFFLE \sigma) =
+        \phi_j(\sigma) \SHUFFLE \phi_j(\sigma).
     \end{equation}
-    Then, since $\pi$ appears in $\rho \SHUFFLE \rho$, $\phi_j(\pi)$
-    appears in $\phi_j(\rho \SHUFFLE \rho)$ and appears also in
-    $\phi_j(\rho) \SHUFFLE \phi_j(\rho)$. This shows that
-    $\phi_j(\rho)$ is a square root of $\phi_j(\pi)$ and
+    Then, since $\pi$ appears in $\sigma \SHUFFLE \sigma$, $\phi_j(\pi)$
+    appears in $\phi_j(\sigma \SHUFFLE \sigma)$ and appears also in
+    $\phi_j(\sigma) \SHUFFLE \phi_j(\sigma)$. This shows that
+    $\phi_j(\sigma)$ is a square root of $\phi_j(\pi)$ and
     implies~\ref{item:square_stability_1}, \ref{item:square_stability_2},
@@
 
-This section is devoted to proving hardness of recognizing square permutations.
-In the same way as happens with words, we shall use a linear graph framework
-where deciding whether a permutation is a square reduces to computing some
-specific matching in the associated linear graph \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}.
-It reads as follows.
+This section is devoted to proving hardness of recognizing square
+permutations. In the same way as happens with words, we shall use a
+linear graph framework where deciding whether a permutation is a square
+reduces to computing some specific matching in the associated linear
+graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows.
 
@@
   \begin{enumerate}
-    \item
-    The permutation $\pi$ is a square.
-    \item
-    There exists a perfect containment-free mathing
-    $\mathcal{M}$ of $G_\pi$ such that
-    for any two disctinct edges
-    $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-    we have $\pi(a) < \pi(a')$ if and only if $\pi(b) < \pi(b')$.
+    \item The permutation $\pi$ is a square.
+    \item There exists a perfect containment-free matching $\mathcal{M}$
+    of $G_\pi$ such that for any two distinct edges $(a, b)$ and
+    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
+    if $\pi(b) < \pi(b')$.
   \end{enumerate}
@@
     The proof is, naturally enough, a slight extension of
-  Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and
-  is added here in full for the sake of completeness.
+    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
+    for the sake of completeness.
 
     (\emph{1} $\Rightarrow$ \emph{2}).
-  Since $\pi$ is a quare, it admits a square root $\sigma$. Then, by
-  definition, there
-  are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$ and
-  $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
+    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
+    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
+    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
     letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
+    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
+    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
+    with vertices
+    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
+    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
+    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
+    matching since $I^1 \cap I^2 = \emptyset$, and
+    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
+    Indeed, if it is not the case then there would exist two edges
+    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
+    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
+    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
+    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
+    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
+    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
+    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
+    elements (at position $j$ and $k$) in two patterns that are
     order-isomorphic to $\sigma$.
-  %Let $\sigma$ be a permutation such that $\pi$ is the union of two
-  %patterns that are order-isomorphic to $\sigma$.
-  Let $2n = |\pi|$ (and hence $|\sigma| = n$).
-  %Fix the occurrences in $\pi$ of the two disjoint patterns that
-  %are order-isomorphic to $\sigma$ and
-  %write
-  %$I^1 = \{i^1_1, i^1_2, \ldots, i^1_n\}$,
-  %$i^1_1 < i^1_2 < \ldots < i^1_n$, for the positions in $\pi$
-  %of the first pattern order-ismorphic to $\sigma$ and
-  %$I^2 = \{i^2_1, i^2_2, \ldots, i^2_n\}$,
-  %$i^2_1 < i^2_2 < \ldots < i^2_n$, for the positions in $\pi$
-  %of the second pattern order-ismorphic to $\sigma$.
-  Let $G_\pi$ be the linear clique associated to $\pi$ with
-  vertices $\VS{G_\pi} =
-  \{i^1_j : 1 \leq j \leq n\} \cup \{i^2_j : 1 \leq j \leq n\}$.
-  It is easily seen that
-  $\mathcal{M} = \{\{i^1_j, i^2_j\} : 1 \leq j \leq n\}$
-  is a subset of $\ES{G_\pi}$.
-  Furthermore, $\mathcal{M}$ is a perfect matching since
-  $I^1 \cap I^2 = \emptyset$,
-  and $I^1 \cup I^2 = \{1, 2,  \ldots 2n\}$.
-  It is also containment-free.
-  Indeed, if it is not the case then there would exist two edges
-  $e = \{i^1_j, i^2_j\}$ and
-  $e' = \{i^1_k, i^2_k\}$, $j < k$,
-  in $\mathcal{M}$ such that
-  $i^1_j < i^1_k < i^2_k < i^2_j$ or
-  $i^1_j < i^2_k < i^1_k < i^2_j$.
-  This is a contradiction since
-  $i^2_k > i^2_j$ if $k > j$.
-  Finally, for any two distinct edges
-  $e = \{\i^1_j, i^2_j\}$ and
-  $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
-  we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
-  $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing
-  in both cases two elements (at position $j$ and $k$)
-  in two patterns that are order-isomorphic to $\sigma$.
 
@@
     \;\text{with}\; j < k\}$.
-  Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right) endpoints
-  of the edges in $\mathcal{M}$ assuming the positions $1, 2, \ldots, 2$
-  are drawn along an horizontal line.
-  Without loss of generality,
-  assume $i^1_1 < i^1_2 < \ldots < i^1_n$.
-  % and
-  % $I^2 = \{i^2_1, i^2_2, \ldots, i^2_n\}$.
-  Let
-  $p^1 = \pi(i^1_1) \pi(i^1_2) \ldots \pi(i^1_n)$ and
-  $p^2 = \pi(i^2_1) \pi(i^2_2) \ldots \pi(i^2_n)$ be the two patterns
-  from $\pi$ induced by $I^1$ and $I^2$, respectivelly.
-  Clearly $p^1$ and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is
-  a matching) and cover $\pi$ (since $\mathcal{M}$ is perfect).
-  We also claim that $p^1$ and $p^2$ are order-isomorphic.
-  Suppose, aiming at a contradiction, that this is not the case.
-  Therefore, there exist two positions $j$ and $k$, say $j < k$, such that
+    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
+    endpoints of the edges in $\mathcal{M}$ assuming the positions
+    $1, 2, \dots, 2$ are drawn along an horizontal line.
+    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
+    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
+    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
+    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
+    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
+    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
+    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
+    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
+    contradiction, that this is not the case. Therefore, there exist two
+    positions $j$ and $k$, say $j < k$, such that
     $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
     $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
-  But since $\mathcal{M}$ is containment-free, it follows that
-  both $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and
-  $\{i^1_k, i^2_k\}$, $i^1_k, i^2_k$, belong to $\mathcal{M}$.
-  This is a contradiction since
-  for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\pi(a) < \pi(a')$ if and only if $\pi(b) < \pi(b')$.
-  \qed
-\end{proof}
-
-Let $\pi$ be a permutation.
-For the sake of clarity, we will say that a bunch of consecutive positions $P$
-of $\pi$ is \emph{above} (resp. \emph{below}) above another bunch of consecutive
-positions $P'$ in permutation $\pi$
-if $\pi(i) > \pi(j)$ (resp. $\pi(i) < \pi(j)$) for every
-$i \in P$ and every $j \in P'$.
-For exemple, $\sigma_1$ is above $\sigma_2$
-(in an equivalent manner, $\sigma_2$ is below $\sigma_1$)
-in Fig.~\ref{subfig:increasing before and aboce decreasing},
-whereas $\sigma_1$ is below $\sigma_2$
-(in an equivalent manner, $\sigma_2$ is above $\sigma_1$)
-in Fig.~\ref{subfig:decreasing before and below increasing}.
+    But since $\mathcal{M}$ is containment-free, it follows that both
+    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
+    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
+    since for any two distinct edges $(a, b)$ and $(a', b')$ in
+    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
+    $\pi(b) < \pi(b')$.
+    \qed
+\end{proof}
+
+Let $\pi$ be a permutation. For the sake of clarity, we will say that a
+bunch of consecutive positions $P$ of $\pi$ is \emph{above} (resp.
+\emph{below}) above another bunch of consecutive positions $P'$ in $\pi$
+if $\pi(i) > \pi(j)$ (resp. $\pi(i) < \pi(j)$) for every $i \in P$ and
+every $j \in P'$. For example, $\sigma_1$ is above $\sigma_2$ (in an
+equivalent manner, $\sigma_2$ is below $\sigma_1$) in
+Fig.~\ref{subfig:increasing before and above decreasing}, whereas
+$\sigma_1$ is below $\sigma_2$ (in an equivalent manner, $\sigma_2$ is
+above $\sigma_1$) in Fig.~\ref{subfig:decreasing before and below
+increasing}.
 
@@
     \item
-    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing and
-    $\sigma_1$ is above $\sigma_2$
-    (see Fig.~\ref{subfig:increasing before and aboce decreasing}),
-    then $\mathcal{M}$ sontains at most one $(\sigma_1, \sigma_2)$-edge.
+    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
+    $\sigma_1$ is above $\sigma_2$ (see Fig.~\ref{subfig:increasing before
+    and above decreasing}), then $\mathcal{M}$ contains at most one
+    $(\sigma_1, \sigma_2)$-edge.
     \item
-    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing and
-    $\sigma_1$ is below $\sigma_2$
-    (see Fig.~\ref{subfig:decreasing before and below increasing}),
-    then $\mathcal{M}$ sontains at most one $(\sigma_1, \sigma_2)$-edge.
+    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
+    is below $\sigma_2$ (see Fig.~\ref{subfig:decreasing before and below
+    increasing}), then $\mathcal{M}$ contains at most one
+    $(\sigma_1, \sigma_2)$-edge.
   \end{enumerate}
@@
       \end{tikzpicture}
-      \label{subfig:increasing before and aboce decreasing}
+      \label{subfig:increasing before and above decreasing}
     }% end subfigure
@@
   \end{figure}
-  We first prove \emph{1}.
-  Suppose, aiming at a contradiction, that there exist two
-  $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in $\mathcal{M}$.
-  Since $\mathcal{M}$ is containment-free, we may assume $a < b$ and $a' < b'$.
-  Therefore, we have
-  $\pi(a) < \pi(b)$ (since $\sigma_1$ is increasing) and
-  $\pi(a') > \pi(b')$ (since $\sigma_2$ is decreasing),
-  which contradicts our hypothesis about $\mathcal{M}$.
-
-  As for \emph{2}, suppose, still aiming at a contradiction,
-  that there exist two
-  $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in $\mathcal{M}$.
-  Since $\mathcal{M}$ is containment-free, we may assume $a < b$ and $a' < b'$.
-  Therefore, we have
-  $\pi(a) > \pi(b)$ (since $\sigma_1$ is decreasing) and
-  $\pi(a') < \pi(b')$ (since $\sigma_2$ is increasing),
-  which contradicts our hypothesis about $\mathcal{M}$.
+  We first prove \emph{1}. Suppose, aiming at a contradiction, that
+  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
+  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
+  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
+  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
+  $\sigma_2$ is decreasing), which contradicts our hypothesis about
+  $\mathcal{M}$.
+
+  As for \emph{2}, suppose, still aiming at a contradiction, that there
+  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
+  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
+  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
+  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
+  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
   \qed
@@
 
-\begin{proof}
-  The problem is certainly in $\NPclass$.
-  We propose a reduction from the pattern involvement problem which is
-  known to be \NPC~\cite{Bose:Buss:Lubiw:1998}:
-  Given two permutations $\pi$ and $\sigma$, decide whether $\sigma$
-  occurs in $\pi$ (as an order-isomorphic pattern).
+\begin{proof}[of Proposition~\ref{proposition:hardness}]
+  The problem is certainly in $\NPclass$. We propose a reduction from
+  the pattern involvement problem which is known to be
+  \NPC~\cite{Bose:Buss:Lubiw:1998}: Given two permutations $\pi$ and
+  $\sigma$, decide whether $\sigma$ occurs in $\pi$ (as an
+  order-isomorphic pattern).
 
@@
   \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_4)-edge}]
-    According to Claim~\ref{claim: all (nu_2, nu'_2)-edges},
+    According to Claim~\ref{claim:all (nu_2, nu'_2)-edges},
     all positions in $\nu_2$ and $\nu'_2$ are involved in
@@
 see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?
-% Given two
-% permutations $\pi$ and $\sigma$, how hard is the problem of deciding
-% whether $\pi$ is the union of two patterns order-isomorphic to $\sigma$?
+Given two permutations $\pi$ and $\sigma$, how hard is the problem of
+deciding whether $\sigma$ is a square root of $\pi$?
 \smallskip

### 2015-09-18 article.tex
Abstract

@@
 \begin{abstract}
-    A permutation is said to be a square if it is the union of two
-    order-isomorphic patterns. The definition is intended to be the
-    natural counterpart to the ordinary shuffle of words and languages.
-    In this paper, we tackle the problem of recognizing square
+    A permutation is said to be a square if it can be obtained by
+    shuffling two order-isomorphic patterns. The definition is intended
+    to be the natural counterpart to the ordinary shuffle of words and
+    languages. In this paper, we tackle the problem of recognizing square
     permutations from both the point of view of algebra and algorithmic.
-    On the one hand, we present some algebraic properties of the shuffle
-    product of permutations and exhibit a bijection between square
-    $(213,231)$-avoiding permutations and square binary words. On the
-    other hand, we prove that recognizing square permutations is \NPC.
+    On the one hand, we present some algebraic and combinatorial
+    properties of the shuffle product of permutations. We follow an
+    unusual line consisting in defining the shuffle of permutations by
+    means of an unshuffling operator, known as a coproduct. This
+    strategy allows to obtain easy proofs for algebraic and combinatorial
+    properties of our shuffle product. On the other hand, we prove that
+    recognizing square permutations is \NPC. This property is proved by
+    establishing a reduction from the pattern involvement problem and by
+    using a containment-free perfect matching criterion to decide if a
+    permutation is a square.
+    %We exhibit a bijection between square
+    %$(213,231)$-avoiding permutations and square binary words.
 \end{abstract}

### 2015-09-20 article.tex
Abstract
Resolution of the ??

@@
     strategy allows to obtain easy proofs for algebraic and combinatorial
-    properties of our shuffle product. On the other hand, we prove that
-    recognizing square permutations is \NPC. This property is proved by
-    establishing a reduction from the pattern involvement problem and by
-    using a containment-free perfect matching criterion to decide if a
+    properties of our shuffle product. We besides exhibit a bijection
+    between square $(213,231)$-avoiding permutations and square binary
+    words. On the other hand, we prove that recognizing square
+    permutations is \NPC. This property is proved by establishing a
+    reduction from the pattern involvement problem and by using a
+    containment-free perfect matching criterion to decide if a
     permutation is a square.
-    %We exhibit a bijection between square
-    %$(213,231)$-avoiding permutations and square binary words.
 \end{abstract}
@@
 
-  The above claim will be complement in upcoming Claim~\ref{claim: all (nu_1, nu'_1)-edges}.
+  The above claim will be complement in upcoming Claim~\ref{claim:all (nu_1, nu'_1)-edges}.
 
@@
     exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
+    Since $\mathcal{M}$ is containment-free (Lemma~\ref{lemma:matching}),
     we have $x' < a'$.

### 2015-09-20 article.tex
Some improvements

@@
     \item \label{item:problem_4}
-    Given a word $u$, decide if there exist two words $v_1$ and $v_2$
-    such that $u$ is a shuffle of $v_1$ and $v_2$.
+    Given a word $u$, decide if there is a word $v$ such that $u$ is
+    in $v \shuffle v$.
 \end{enumerate}
@@
 of~Problem~\ref{item:problem_1} has been proposed
-in~\cite{Allauzen:IGM:2000}, where it is proved that, given words
+in~\cite{Allauzen:IGM:2000}, where it is proved that given words
 $v_1$, \dots, $v_k$, the iterated shuffle
@@
 of $v$. Again, this remains true even if the ground alphabet has size $3$.
-Let us now focus on Problem~\ref{item:problem_4}. It is shown
+Let us now finally focus on Problem~\ref{item:problem_4}. It is shown
 in~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013} that it is \NPC to
@@
 a word $u$ with the property that there exists a word $v$ such that $u$
-is a shuffle of $v$ with itself. The hardness of this last particular
-problem implies that Problem~\ref{item:problem_4} is \NPC.
+is a shuffle of $v$ with itself. Hence, Problem~\ref{item:problem_4}
+is \NPC.
 \smallskip
@@
 {\em unshuffle} permutations. This operator is in fact a coproduct,
-endowing the space of all permutations with a coalgebra structure
+endowing the linear span of all permutations with a coalgebra structure
 (see~\cite{Joni:Rota:1979} or~\cite{Grinberg:Reiner:2014} for the
@@
 the shuffle of permutations. In particular, we show that binary words
-that are square are in one-to-one bijection with square permutations
-avoiding some patterns
+that are square are in one-to-one correspondence with square
+permutations avoiding some patterns
 (Proposition~\ref{prop:bijection_binary_to_permutations_squares}). Next,
@@
 of this paper: the fact that deciding if a permutation is a square is
-\NPC (Proposition~\ref{proposition:hardness}).
+\NPC (Proposition~\ref{proposition:hardness}). This result is obtained
+by exhibiting a reduction from the pattern involvement
+problem~\cite{Bose:Buss:Lubiw:1998} which is \NPC.
 
@@
 permutations ({\em i.e.}, the product of a permutation of size $n$ with
-a permutation of size $m$ produces a sum of permutations of size $n + m$.
+a permutation of size $m$ produces a sum of permutations of size $n + m$).
 Notice that there are other shuffle products on
@@
 \end{equation}
-For instance, the square permutations of sizes $0$ to $4$ are
+This sequence (and its subsequence obtained by removing the $0$s) is for
+the time being not listed in~\cite{Slo}. The square permutations of
+sizes $0$ to $4$ are
 \smallskip
@@
 \end{proof}
+\medskip
+
+The number of square binary words is Sequence A191755 of~\cite{Slo}
+beginning by
+\begin{equation}
+1, 0, 2, 0, 6, 0, 22, 0, 82, 0, 320, 0, 1268, 0, 5102, 0, 020632.
+\end{equation}
+By Proposition~\ref{prop:bijection_binary_to_permutations_squares}, this
+is also the sequence enumerating square permutations avoiding $213$
+and $231$.
 \medskip
@@
 number of square permutations avoiding $123$ begins by
-\begin{equation}
+\begin{equation} \label{equ:sequence_square_123}
     1, 0, 2, 0, 12, 0, 118, 0, 1218, 0, 14272,
@@
 while the number of square permutations avoiding $132$ begins by
-\begin{equation}
+\begin{equation} \label{equ:sequence_square_132}
     1, 0, 2, 0, 11, 0, 84, 0, 743, 0, 7108.
@@
 inverse. The sequence enumerating these sets begins by
-\begin{equation}
+\begin{equation} \label{equ:sequence_square_classes}
     1, 0, 1, 0, 6, 0, 81, 0, 2774, 0, 162945.
 \end{equation}
+\medskip
+
+All Sequences~\eqref{equ:sequence_square_123}, \eqref{equ:sequence_square_132},
+and~\eqref{equ:sequence_square_classes} (and their subsequences obtained
+by removing the $0$s) are for the time being not listed in~\cite{Slo}.
 \medskip

### 2015-09-22 submission.tex
Deplacement des demos dans submission.tex
abstract

--- 2015-09-20 article.tex
@@
     between square $(213,231)$-avoiding permutations and square binary
-    words. On the other hand, we prove that recognizing square
-    permutations is \NPC. This property is proved by establishing a
-    reduction from the pattern involvement problem and by using a
-    containment-free perfect matching criterion to decide if a
-    permutation is a square.
+    words.
+    On the other hand,
+    using a containment-free perfect matching criterion,
+    we prove that recognizing square permutations is \NPC. 
 \end{abstract}
@@
 or~\cite{Grinberg:Reiner:2014}.
-\medskip
 
@@
 respectively order-isomorphic to the permutations $1$ and $132$.
-\medskip
 
@@
 in~\eqref{equ:example_unshuffling_coproduct}.
-\medskip
 
@@
 last two.
-\medskip
 
@@
 order-isomorphic to $\nu$.
-\medskip
 
@@
 \end{tabular}
-\medskip
 
@@
 description of square binary words.
-\medskip
 
@@
 such that $|ua|_0$ and $|ua|_1$ are both even, where $a \in \{0, 1\}$.
-\medskip
 
@@
     \end{claim}
-    \begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
-        Let us first show that the image of $\BINTOPERM$ contains only
-        permutations avoiding $213$ and $231$. Let $u$ be a binary word,
-        $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
-        positions of the occurrences of $0$ (resp. $1$) in $u$. By
-        definition of $\BINTOPERM$, from left to right, the subword
-        $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
-        is decreasing, and all letters of $w$ are greater than those
-        of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
-        Then, since $v$ is increasing and $w$ is decreasing, there is an
-        occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
-        occurrence of $21$ (resp. $2$, $1$). All these three cases
-        contradict the fact that all letters of $w$ are greater than
-        those of $v$. A similar argument shows that $\pi$ avoids~$231$
-        as well.
-        \smallskip
-
-        Finally, observe that any permutation $\pi$ avoiding $213$ and
-        $231$ necessarily starts by the smallest possible letter or the
-        greatest possible letter. This property is then true for the
-        suffix of $\pi$ obtained by deleting its first letter,
-        and so on for all of its suffixes. Thus, by
-        replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
-        role of a smallest (resp. greatest) letter, one obtains a binary
-        word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
-        avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
-        \qed
-    \end{proof}
+
     \begin{claim} \label{claim:square_binary_to_square_permutation}
@@
     \end{claim}
-    \begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
-        Since $u$ is a square binary word, there is a binary word $v$
-        such that $u \in v \shuffle v$. Then, there are two disjoint
-        sets $P$ and $Q$ of positions of letters of $u$ such that
-        $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
-        words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
-        same standarized $\sigma$. Hence, and by definition of
-        the shuffle product of permutations, $\BINTOPERM(u)$ appears in
-        $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
-        square permutation.
-        \qed
-    \end{proof}
+
     \begin{claim} \label{claim:square_permutation_to_square_binary}
@@
     \end{claim}
-    \begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
-        Let $\pi$ be a square permutation avoiding $213$ and $231$. By
-        Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
-        the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
-        well-defined binary word. Since $\pi$ is a square permutation,
-        there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
-        of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
-        order-isomorphic. This implies, by the definitions of $\BINTOPERM$
-        and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
-        is a square binary word.
-        \qed
-    \end{proof}
-\end{proof}
-\medskip
+\end{proof}
 
@@
 and $231$.
-\medskip
 
@@
 combinatorial properties of $\SHUFFLE$.
-\medskip
 
@@
 \end{proof}
-\medskip
 
@@
 commutative algebra.
-\medskip
 
@@
 \end{lemma}
-\begin{proof}[of Lemma~\ref{lem:endomorphisms}]
-    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
-    associative algebras, we have to prove that for all permutations
-    $\sigma$ and $\nu$,
-    \begin{equation}
-        \phi_j(\sigma \SHUFFLE \nu) =
-        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
-    \end{equation}
-    We shall proceed here as we have done during the proof of
-    Proposition~\ref{prop:shuffle_associative_commutative} by showing
-    that $\phi_j$ is a morphism of coalgebras, that is
-    \begin{equation}
-        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
-    \end{equation}
-    In the sequel, $\pi$ is a permutation.
-    \smallskip
-
-    If $P$ is a set of indexes of letters of $\pi$, we denote by
-    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
-    the operation $\widetilde{\,}$ defines a bijection on the set of the
-    subsets of $[|\pi|]$, and since the standardization operation
-    commutes with the mirror operation on words without multiple
-    occurrence of a letter, we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_1(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_1(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|P_1}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \widetilde{\STD\left(\pi_{|P_1}\right)}
-        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
-    $\phi_1$ is a morphism of associative algebras.
-    \smallskip
-
-    Next, since by definition of the complementation operation on
-    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
-    we have $\tau(i) < \tau(k)$ if and only if $\bar \tau(i) > \bar \tau(k)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_2(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_2(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\bar \pi_{|P_1}\right)
-        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
-    $\phi_2$ is a morphism of associative algebras.
-    \smallskip
-
-    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
-    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
-    $\{\tau(i) : i \in P\}$. Since the map sending a subset $P$ of
-    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
-    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_3(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_3(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|P_1}\right)
-        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
-        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right)^{-1}
-        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
-    $\phi_3$ is a morphism of associative algebras.
-    \qed
-\end{proof}
-\medskip
 
@@
 of square permutations.
-\medskip
 
@@
 \end{equation}
-\medskip
 
@@
 by removing the $0$s) are for the time being not listed in~\cite{Slo}.
-\medskip
 
@@
 
-\begin{proof}[of Lemma~\ref{lemma:matching}]
-    The proof is, naturally enough, a slight extension of
-    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
-    for the sake of completeness.
-
-    (\emph{1} $\Rightarrow$ \emph{2}).
-    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
-    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
-    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
-    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
-    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
-    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
-    with vertices
-    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
-    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
-    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
-    matching since $I^1 \cap I^2 = \emptyset$, and
-    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
-    Indeed, if it is not the case then there would exist two edges
-    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
-    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
-    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
-    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
-    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
-    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
-    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
-    elements (at position $j$ and $k$) in two patterns that are
-    order-isomorphic to $\sigma$.
-
-    (\emph{2} $\Rightarrow$ \emph{1}).
-    Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$, and
-    $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$.
-    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
-    endpoints of the edges in $\mathcal{M}$ assuming the positions
-    $1, 2, \dots, 2$ are drawn along an horizontal line.
-    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
-    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
-    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
-    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
-    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
-    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
-    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
-    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
-    contradiction, that this is not the case. Therefore, there exist two
-    positions $j$ and $k$, say $j < k$, such that
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
-    But since $\mathcal{M}$ is containment-free, it follows that both
-    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
-    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
-    since for any two distinct edges $(a, b)$ and $(a', b')$ in
-    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
-    $\pi(b) < \pi(b')$.
-    \qed
-\end{proof}
-
 Let $\pi$ be a permutation. For the sake of clarity, we will say that a
@@
 increasing}.
-
-Before proving hardness, we give an easy lemma that will prove extremely
-useful for simplifying the proof of upcoming
-Proposition~\ref{proposition:hardness}.
-
-\begin{lemma}
-  \label{lemma:at most one edge monotone}
-  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
-  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
-  and $\mathcal{M}$ be a containment-free perfect matching of $G_\pi$.
-  The following assertions hold:
-  \begin{enumerate}
-    \item
-    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
-    $\sigma_1$ is above $\sigma_2$ (see Fig.~\ref{subfig:increasing before
-    and above decreasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
-    \item
-    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
-    is below $\sigma_2$ (see Fig.~\ref{subfig:decreasing before and below
-    increasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
   \begin{figure}[t!]
@@
   \end{figure}
-  We first prove \emph{1}. Suppose, aiming at a contradiction, that
-  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
-  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
-  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
-  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
-  $\sigma_2$ is decreasing), which contradicts our hypothesis about
-  $\mathcal{M}$.
-
-  As for \emph{2}, suppose, still aiming at a contradiction, that there
-  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
-  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
-  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
-  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
-  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
-  \qed
-\end{proof}
+
+Before proving hardness, we give an easy lemma that will prove extremely
+useful for simplifying the proof of upcoming
+Proposition~\ref{proposition:hardness}.
+
+\begin{lemma}
+  \label{lemma:at most one edge monotone}
+  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
+  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
+  and $\mathcal{M}$ be a containment-free perfect matching of $G_\pi$.
+  The following assertions hold:
+  \begin{enumerate}
+    \item
+    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
+    $\sigma_1$ is above $\sigma_2$ (see Fig.~\ref{subfig:increasing before
+    and above decreasing}), then $\mathcal{M}$ contains at most one
+    $(\sigma_1, \sigma_2)$-edge.
+    \item
+    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
+    is below $\sigma_2$ (see Fig.~\ref{subfig:decreasing before and below
+    increasing}), then $\mathcal{M}$ contains at most one
+    $(\sigma_1, \sigma_2)$-edge.
+  \end{enumerate}
+\end{lemma}
 
@@
 \newpage
-\section*{Appendix (Reviewers' version only)}
+\section*{Appendix (Reviewers version only)}
+
+\begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
+    Let us first show that the image of $\BINTOPERM$ contains only
+    permutations avoiding $213$ and $231$. Let $u$ be a binary word,
+    $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
+    positions of the occurrences of $0$ (resp. $1$) in $u$. By
+    definition of $\BINTOPERM$, from left to right, the subword
+    $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
+    is decreasing, and all letters of $w$ are greater than those
+    of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
+    Then, since $v$ is increasing and $w$ is decreasing, there is an
+    occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
+    occurrence of $21$ (resp. $2$, $1$). All these three cases
+    contradict the fact that all letters of $w$ are greater than
+    those of $v$. A similar argument shows that $\pi$ avoids~$231$
+    as well.
+    \smallskip
+
+    Finally, observe that any permutation $\pi$ avoiding $213$ and
+    $231$ necessarily starts by the smallest possible letter or the
+    greatest possible letter. This property is then true for the
+    suffix of $\pi$ obtained by deleting its first letter,
+    and so on for all of its suffixes. Thus, by
+    replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
+    role of a smallest (resp. greatest) letter, one obtains a binary
+    word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
+    avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
+    Since $u$ is a square binary word, there is a binary word $v$
+    such that $u \in v \shuffle v$. Then, there are two disjoint
+    sets $P$ and $Q$ of positions of letters of $u$ such that
+    $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
+    words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
+    same standarized $\sigma$. Hence, and by definition of
+    the shuffle product of permutations, $\BINTOPERM(u)$ appears in
+    $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
+    square permutation.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
+    Let $\pi$ be a square permutation avoiding $213$ and $231$. By
+    Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
+    the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
+    well-defined binary word. Since $\pi$ is a square permutation,
+    there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
+    of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
+    order-isomorphic. This implies, by the definitions of $\BINTOPERM$
+    and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
+    is a square binary word.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Lemma~\ref{lem:endomorphisms}]
+    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
+    associative algebras, we have to prove that for all permutations
+    $\sigma$ and $\nu$,
+    \begin{equation}
+        \phi_j(\sigma \SHUFFLE \nu) =
+        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
+    \end{equation}
+    We shall proceed here as we have done during the proof of
+    Proposition~\ref{prop:shuffle_associative_commutative} by showing
+    that $\phi_j$ is a morphism of coalgebras, that is
+    \begin{equation}
+        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
+    \end{equation}
+    In the sequel, $\pi$ is a permutation.
+    \smallskip
+
+    If $P$ is a set of indexes of letters of $\pi$, we denote by
+    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
+    the operation $\widetilde{\,}$ defines a bijection on the set of the
+    subsets of $[|\pi|]$, and since the standardization operation
+    commutes with the mirror operation on words without multiple
+    occurrence of a letter, we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_1(\pi))
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\phi_1(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\widetilde{\pi}_{|P_1}\right)
+        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
+        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \widetilde{\STD\left(\pi_{|P_1}\right)}
+        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
+    $\phi_1$ is a morphism of associative algebras.
+    \smallskip
+
+    Next, since by definition of the complementation operation on
+    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
+    we have $\tau(i) < \tau(k)$ if and only if $\bar \tau(i) > \bar \tau(k)$,
+    we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_2(\pi))
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\phi_2(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\bar \pi_{|P_1}\right)
+        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
+    $\phi_2$ is a morphism of associative algebras.
+    \smallskip
+
+    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
+    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
+    $\{\tau(i) : i \in P\}$. Since the map sending a subset $P$ of
+    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
+    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
+    we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_3(\pi))
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\phi_3(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\pi^{-1}_{|P_1}\right)
+        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
+        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\pi_{|P_1}\right)^{-1}
+        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
+    $\phi_3$ is a morphism of associative algebras.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Lemma~\ref{lemma:matching}]
+    The proof is, naturally enough, a slight extension of
+    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
+    for the sake of completeness.
+
+    (\emph{1} $\Rightarrow$ \emph{2}).
+    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
+    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
+    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
+    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
+    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
+    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
+    with vertices
+    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
+    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
+    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
+    matching since $I^1 \cap I^2 = \emptyset$, and
+    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
+    Indeed, if it is not the case then there would exist two edges
+    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
+    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
+    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
+    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
+    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
+    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
+    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
+    elements (at position $j$ and $k$) in two patterns that are
+    order-isomorphic to $\sigma$.
+
+    (\emph{2} $\Rightarrow$ \emph{1}).
+    Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
+    \;\text{with}\; j < k\}$, and
+    $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
+    \;\text{with}\; j < k\}$.
+    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
+    endpoints of the edges in $\mathcal{M}$ assuming the positions
+    $1, 2, \dots, 2$ are drawn along an horizontal line.
+    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
+    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
+    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
+    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
+    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
+    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
+    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
+    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
+    contradiction, that this is not the case. Therefore, there exist two
+    positions $j$ and $k$, say $j < k$, such that
+    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
+    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
+    But since $\mathcal{M}$ is containment-free, it follows that both
+    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
+    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
+    since for any two distinct edges $(a, b)$ and $(a', b')$ in
+    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
+    $\pi(b) < \pi(b')$.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
+  We first prove \emph{1}. Suppose, aiming at a contradiction, that
+  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
+  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
+  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
+  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
+  $\sigma_2$ is decreasing), which contradicts our hypothesis about
+  $\mathcal{M}$.
+
+  As for \emph{2}, suppose, still aiming at a contradiction, that there
+  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
+  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
+  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
+  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
+  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
+  \qed
+\end{proof}
+\bigskip
 

### 2015-09-22 submission.tex
add figure

@@
 \renewcommand{\arraystretch}{1.2}
+\newcommand{\rt}{\textcolor{red}}
 
@@
 \DeclareMathOperator{\PERMTOBIN}{\mathrm{ptb}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% 1 2 2 1
+\newcommand{\MWithContainment}{%
+\begin{tikzpicture}[scale=0.1,inner sep=.8pt,node distance=.25cm]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw (U2.north) .. controls ($ (U2.north) + (0,1.75) $) and ($ (U3.north) + (0,1.75) $) .. (U3.north);
+\draw (U1.north) .. controls ($ (U1.north) + (0,2.5) $) and ($ (U4.north) + (0,2.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \MWithContainment
 
@@
 \label{section:Introduction}
-The {\em shuffle product} $\shuffle$ is a very famous operation on words
+The {\em shuffle product}, denoted $\shuffle$, is a very famous operation on words
 first defined by Eilenberg and Mac Lane~\cite{Eilenberg:MacLane:1953}.
@@
 $v_1$ and $v_2$ in a way that maintains the left-to-right ordering of the
-letters from each word. In addition to purely combinatorial questions,
+letters from each word.
+Besides purely combinatorial questions,
 the shuffle product of words leads to the following computational ones:
@@
 
-Let us now provide some definitions about graphs and matching, used in
+Let us now provide some definitions about graphs and matching that are used in
 the sequel. For a graph $G$, we denote $\VS{G}$ as the set of vertices
@@
 edges. A matching is \emph{perfect} if it covers all the vertices of the
-graph. In case the set of vertices is equipped with a total order, a
+graph.
+For an edge $e = \{u,v\}$, we shall occasionally say that $u$ and $v$
+are \emph{involved} in $e$.
+In case the set of vertices is equipped with a total order, a
 matching $\mathcal{M}$ is said to be \emph{containment-free} if there do
@@
 $\mathcal{M}$ such that $u_i < u_k < u_\ell < u_j$ or
-$u_k < u_i < u_j < u_\ell$.
+$u_k < u_i < u_j < u_\ell$
+(\emph{i.e.}, $\mathcal{M}$ avoids \MWithContainment).
 
@@
 is justified by the fact that a lot of properties of $\SHUFFLE$ depend
-of properties of $\Delta$ and that this strategy allows to write concise
+of properties of $\Delta$, and that this strategy allows to write concise
 and clear proofs of them. We invite the reader unfamiliar with the
@@
     \end{claim}
+    \qed
 \end{proof}
@@
 reduces to computing some specific matching in the associated linear
-graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows.
+graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows
+(see Fig.\ref{fig:example containment-free matching}).
 
@@
     \item There exists a perfect containment-free matching $\mathcal{M}$
-    of $G_\pi$ such that for any two distinct edges $(a, b)$ and
+    of $K_\pi$ such that for any two distinct edges $(a, b)$ and
     $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
@@
 \end{lemma}
+
+\begin{figure}[ht!]
+  \centering
+  \begin{tikzpicture}[scale=0.5,inner sep=2pt,node distance=1cm]
+    \draw node (K) {$K_\pi$};
+    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=K]   (U01) {$2$};
+    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U01] (U02) {$A$};
+    \draw node [draw, circle, minimum size=.55cm] [right of=U02] (U03) {$4$};
+    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U03] (U04) {$7$};
+    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U04] (U05) {$1$};
+    \draw node [draw, circle, minimum size=.55cm] [right of=U05] (U06) {$9$};
+    \draw node [draw, circle, minimum size=.55cm] [right of=U06] (U07) {$5$};
+    \draw node [draw, circle, minimum size=.55cm] [right of=U07] (U08) {$3$};
+    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U08] (U09) {$8$};
+    \draw node [draw, circle, minimum size=.55cm] [right of=U09] (U10) {$6$};
+    \draw [thick]
+      (U01.north) .. controls ($ (U01.north) + (0,2.25) $) and ($ (U03.north) + (0,2.25) $) .. (U03.north);
+    \draw [thick]
+      (U02.north) .. controls ($ (U02.north) + (0,2.25) $) and ($ (U06.north) + (0,2.25) $) .. (U06.north);
+    \draw [thick]
+      (U04.north) .. controls ($ (U04.north) + (0,2.5) $) and ($ (U07.north) + (0,2.5) $) .. (U07.north);
+    \draw [thick]
+      (U05.north) .. controls ($ (U05.north) + (0,2.5) $) and ($ (U08.north) + (0,2.5) $) .. (U08.north);
+    \draw [thick]
+      (U09.north) .. controls ($ (U09.north) + (0,2.5) $) and ($ (U10.north) + (0,2.5) $) .. (U10.north);
+  \end{tikzpicture}
+  \caption{\label{fig:example containment-free matching}%
+    The linear graph $K_\pi$ associated to the permutation
+    $\pi = 2A47195386$ together with
+    a containment-free perfect matching $\mathcal{M}$
+    such that for any two distinct edges $(a, b)$ and
+    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
+    if $\pi(b) < \pi(b')$.
+    From $\mathcal{M}$, it follows that $\pi$ is square as it is
+    the shuffle of $2A718$ and $49536$ that are both
+    order-isomorphic to $25314$.
+  }% end caption
+\end{figure}
 

### 2015-09-24 submission.tex
typos

@@
 of null length, is denoted by $\epsilon$. We denote by $\widetilde{u}$
-the {\em mirror image} of $u$ that is the word
+the {\em mirror image} of $u$, that is the word
 $u_{|u|} u_{|u| - 1} \dots u_1$. If $P$ is a subset of $[|u|]$, $u_{|P}$
@@
 $1 2 \dots n$ (resp. $n (n-1) \dots 1$). If $\pi$ is a permutation of
-$S_n$, we denote by $\bar \pi$ the {\em complementary} of $\pi$ that is
+$S_n$, we denote by $\bar \pi$ the {\em complementary} of $\pi$, that is
 the permutation satisfying $\bar \pi(i) = n - \pi(i) + 1$ for all

### 2015-09-24 submission.tex
typos

@@
 $P_1$ and $P_2$ of disjoints indexes of letters of $\pi$ satisfying
-$P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$ and
+$P_1 \sqcup P_2 = [|\pi|]$ such that the subwords $\pi_{|P_1}$ and
 $\pi_{|P_2}$ are order-isomorphic. Computer experiments give us the
@@
 \end{equation}
-This sequence (and its subsequence obtained by removing the $0$s) is for
+This sequence (and its subsequence obtained by removing the $0$'s) is for
 the time being not listed in~\cite{Slo}. The square permutations of

### 2015-09-25 submission.tex
typos
N1, N2, N3, N4
introduction
Typos

@@
 A solution to Problem~\ref{item:problem_1} can be computed in
-\begin{equation}
-    O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)
-\end{equation}
+% \begin{equation}
+%     O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)
+% \end{equation}
+$O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)$
 time~\cite{Spehner:TCS:1986}. An improvement and a generalization
@@
 $v_1 \shuffle \dots \shuffle v_k$ can be computed in
-\begin{equation}
-    O\left(\binom{m_1 + \dots + m_k}{m_1, \dots, m_k}\right)
-\end{equation}
+% \begin{equation}
+%     O\left(\binom{m_1 + \dots + m_k}{m_1, \dots, m_k}\right)
+% \end{equation}
+$O\left(\binom{m_1 + \dots + m_k}{m_1, \dots, m_k}\right)$
 time. Problem~\ref{item:problem_2} is in \Pclass; it is indeed a
@@
 the set recursively defined by
-$u \shuffle \epsilon = u = \epsilon \shuffle u$ and
+$u \shuffle \epsilon = \{u\} = \epsilon \shuffle u$ and
 $ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
@@
 $u_k < u_i < u_j < u_\ell$
-(\emph{i.e.}, $\mathcal{M}$ avoids \MWithContainment).
+(\emph{i.e.}, $\mathcal{M}$ avoids~\MWithContainment).
 
@@
 (see Fig.\ref{fig:example containment-free matching}).
-
-\begin{lemma}
-  \label{lemma:matching}
-  Let $\pi$ be a permutation. The following statements are equivalent:
-  \begin{enumerate}
-    \item The permutation $\pi$ is a square.
-    \item There exists a perfect containment-free matching $\mathcal{M}$
-    of $K_\pi$ such that for any two distinct edges $(a, b)$ and
-    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
-    if $\pi(b) < \pi(b')$.
-  \end{enumerate}
-\end{lemma}
-
 \begin{figure}[ht!]
@@
     if $\pi(b) < \pi(b')$.
-    From $\mathcal{M}$, it follows that $\pi$ is square as it is
-    the shuffle of $2A718$ and $49536$ that are both
+    From $\mathcal{M}$, it follows that $\pi$ is square as it is in
+    the shuffle of $2A718$ and $49536$, both being
     order-isomorphic to $25314$.
@@
 \end{figure}
+
+\begin{lemma}
+  \label{lemma:matching}
+  Let $\pi$ be a permutation. The following statements are equivalent:
+  \begin{enumerate}
+    \item The permutation $\pi$ is a square.
+    \item There exists a perfect containment-free matching $\mathcal{M}$
+    of $K_\pi$ such that for any two distinct edges $(a, b)$ and
+    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
+    if $\pi(b) < \pi(b')$.
+  \end{enumerate}
+\end{lemma}
 
@@
   \begin{align*}
-  N_4 &= 2(2n + k + 2) + 1  = 4n + 2k + 5 \\
-  N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 12k + 9 \\
-  N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 \\
-  N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1\text{.}
+  N_4 &= 2(2n + 2k + 4) + 1  = 4n + 4k + 9 \\
+  N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45 \\
+  N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225 \\
+  N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325\text{.}
   \end{align*}
@@
 see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?
-Given two permutations $\pi$ and $\sigma$, how hard is the problem of
-deciding whether $\sigma$ is a square root of $\pi$?
-\smallskip
+Given two permutations $\pi$ and~$\sigma$, how hard is the problem of
+deciding whether $\sigma$ is a square root of~$\pi$?
 

### 2015-09-25 submission.tex
move proof

@@
       % grid
-      \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
+      \draw[step=1cm,black!30,thin,fill=black!10] (0,0) grid (12,12);
       % row size
@@
 
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% %%
+% %% ---- Patterns ----
+% %%
+% \section{Patterns}
+% \label{section:Patterns}
+%
+% \begin{proposition}
+%   \label{proposition:patterns:lower bound}
+%   Let $\pi \in S_{n}$.
+%   Then $\pi$ contains a pattern of size at least
+%   $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
+%   is the union of two order-isomorphic patterns.
+% \end{proposition}
+%
+% \begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
+%   The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935}
+%   states that every permutation of
+%   length at least $pq+1$ must contain either the pattern
+%   $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
+%   The result now follows from the fact that
+%   the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
+% \qed
+% \end{proof}
+%
+% \begin{proposition}
+%   \label{proposition:patterns:upper bound}
+%   If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
+%   a permutation of $S_{n}$
+%   that does not contains a pattern
+%   of length $2k$ that is the union of two
+%   order-isomorphic patterns.
+% \end{proposition}
+%
+% \begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
+% The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
+% Let $\pi \in S_n$ be chosen randomly and uniformly.
+% For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
+% of $\pi$,
+% let $X_\sigma$ be the indicator variable that
+% $\sigma$ is the union of two order-ismorphic patterns.
+% According to Boole's inequality, we have
+% $P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
+% Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
+% the probability that $\pi$ contains a pattern of length $2k$
+% is the union of two order-ismorphic patterns is at most
+% $\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
+% Thus, with positive probability, no event $X_\sigma$ occurs and
+% there is a permutation of $S_n$ that does not contain a pattern
+% of length $2k$ that is the union of two order-isomorphic patterns.
+% \qed
+% \end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% ---- Conclusion ----
+%%
+\section{Conclusion}
+\label{section:Conclusion}
+
+There are a number of further directions of investigation in this
+general subject. We mention several -~not necessarily new~- open
+problems that are, in our opinion, the most interesting. How many
+permutations of $S_{2n}$ are squares? How many $(213,231)$-avoiding
+permutations of $S_{2n}$ are squares? (Equivalently, by
+Proposition~\ref{prop:bijection_binary_to_permutations_squares},
+how many binary strings of length $2n$ are squares; Problem~4
+in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of
+deciding whether a $(213,231)$-avoiding permutation is a square
+(Problem~4 in \cite{Henshall:Rampersad:Shallit:2011},
+see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?
+Given two permutations $\pi$ and~$\sigma$, how hard is the problem of
+deciding whether $\sigma$ is a square root of~$\pi$?
+
+Besides, in this work, we investigate some algebraic properties of the
+shuffle of permutations which have direct combinatorial consequences.
+However, this work raises several additional classical algebraic
+questions. Among these, one can ask for a complete algebraic study of
+$\QQ[S]$ as a graded associative algebra for the shuffle  product
+$\SHUFFLE$. Describing a generating family for $\QQ[S]$, defining
+multiplicative bases of $\QQ[S]$, and determining whether $\QQ[S]$ is
+free as an associative algebra are worthwhile questions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% Bibliography
+%%
+
+\bibliographystyle{plain}
+\bibliography{biblio}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+\section*{Appendix (Reviewers version only)}
+
+\begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
+    Let us first show that the image of $\BINTOPERM$ contains only
+    permutations avoiding $213$ and $231$. Let $u$ be a binary word,
+    $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
+    positions of the occurrences of $0$ (resp. $1$) in $u$. By
+    definition of $\BINTOPERM$, from left to right, the subword
+    $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
+    is decreasing, and all letters of $w$ are greater than those
+    of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
+    Then, since $v$ is increasing and $w$ is decreasing, there is an
+    occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
+    occurrence of $21$ (resp. $2$, $1$). All these three cases
+    contradict the fact that all letters of $w$ are greater than
+    those of $v$. A similar argument shows that $\pi$ avoids~$231$
+    as well.
+    \smallskip
+
+    Finally, observe that any permutation $\pi$ avoiding $213$ and
+    $231$ necessarily starts by the smallest possible letter or the
+    greatest possible letter. This property is then true for the
+    suffix of $\pi$ obtained by deleting its first letter,
+    and so on for all of its suffixes. Thus, by
+    replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
+    role of a smallest (resp. greatest) letter, one obtains a binary
+    word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
+    avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
+    Since $u$ is a square binary word, there is a binary word $v$
+    such that $u \in v \shuffle v$. Then, there are two disjoint
+    sets $P$ and $Q$ of positions of letters of $u$ such that
+    $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
+    words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
+    same standarized $\sigma$. Hence, and by definition of
+    the shuffle product of permutations, $\BINTOPERM(u)$ appears in
+    $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
+    square permutation.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
+    Let $\pi$ be a square permutation avoiding $213$ and $231$. By
+    Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
+    the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
+    well-defined binary word. Since $\pi$ is a square permutation,
+    there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
+    of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
+    order-isomorphic. This implies, by the definitions of $\BINTOPERM$
+    and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
+    is a square binary word.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Lemma~\ref{lem:endomorphisms}]
+    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
+    associative algebras, we have to prove that for all permutations
+    $\sigma$ and $\nu$,
+    \begin{equation}
+        \phi_j(\sigma \SHUFFLE \nu) =
+        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
+    \end{equation}
+    We shall proceed here as we have done during the proof of
+    Proposition~\ref{prop:shuffle_associative_commutative} by showing
+    that $\phi_j$ is a morphism of coalgebras, that is
+    \begin{equation}
+        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
+    \end{equation}
+    In the sequel, $\pi$ is a permutation.
+    \smallskip
+
+    If $P$ is a set of indexes of letters of $\pi$, we denote by
+    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
+    the operation $\widetilde{\,}$ defines a bijection on the set of the
+    subsets of $[|\pi|]$, and since the standardization operation
+    commutes with the mirror operation on words without multiple
+    occurrence of a letter, we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_1(\pi))
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\phi_1(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\widetilde{\pi}_{|P_1}\right)
+        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
+        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \widetilde{\STD\left(\pi_{|P_1}\right)}
+        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
+    $\phi_1$ is a morphism of associative algebras.
+    \smallskip
+
+    Next, since by definition of the complementation operation on
+    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
+    we have $\tau(i) < \tau(k)$ if and only if $\bar \tau(i) > \bar \tau(k)$,
+    we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_2(\pi))
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\phi_2(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\bar \pi_{|P_1}\right)
+        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
+    $\phi_2$ is a morphism of associative algebras.
+    \smallskip
+
+    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
+    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
+    $\{\tau(i) : i \in P\}$. Since the map sending a subset $P$ of
+    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
+    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
+    we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_3(\pi))
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\phi_3(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\pi^{-1}_{|P_1}\right)
+        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
+        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\pi_{|P_1}\right)^{-1}
+        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
+    $\phi_3$ is a morphism of associative algebras.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Lemma~\ref{lemma:matching}]
+    The proof is, naturally enough, a slight extension of
+    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
+    for the sake of completeness.
+
+    (\emph{1} $\Rightarrow$ \emph{2}).
+    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
+    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
+    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
+    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
+    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
+    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
+    with vertices
+    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
+    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
+    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
+    matching since $I^1 \cap I^2 = \emptyset$, and
+    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
+    Indeed, if it is not the case then there would exist two edges
+    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
+    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
+    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
+    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
+    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
+    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
+    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
+    elements (at position $j$ and $k$) in two patterns that are
+    order-isomorphic to $\sigma$.
+
+    (\emph{2} $\Rightarrow$ \emph{1}).
+    Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
+    \;\text{with}\; j < k\}$, and
+    $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
+    \;\text{with}\; j < k\}$.
+    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
+    endpoints of the edges in $\mathcal{M}$ assuming the positions
+    $1, 2, \dots, 2$ are drawn along an horizontal line.
+    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
+    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
+    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
+    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
+    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
+    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
+    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
+    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
+    contradiction, that this is not the case. Therefore, there exist two
+    positions $j$ and $k$, say $j < k$, such that
+    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
+    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
+    But since $\mathcal{M}$ is containment-free, it follows that both
+    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
+    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
+    since for any two distinct edges $(a, b)$ and $(a', b')$ in
+    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
+    $\pi(b) < \pi(b')$.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
+  We first prove \emph{1}. Suppose, aiming at a contradiction, that
+  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
+  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
+  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
+  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
+  $\sigma_2$ is decreasing), which contradicts our hypothesis about
+  $\mathcal{M}$.
+
+  As for \emph{2}, suppose, still aiming at a contradiction, that there
+  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
+  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
+  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
+  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
+  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
+  \qed
+\end{proof}
+
+\bigskip
+
+\begin{proof}[of Proposition~\ref{proposition:hardness}]
   Suppose first that $\sigma$ occurs in $\pi$ and fix an occurrence.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% %%
-% %% ---- Patterns ----
-% %%
-% \section{Patterns}
-% \label{section:Patterns}
-%
-% \begin{proposition}
-%   \label{proposition:patterns:lower bound}
-%   Let $\pi \in S_{n}$.
-%   Then $\pi$ contains a pattern of size at least
-%   $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
-%   is the union of two order-isomorphic patterns.
-% \end{proposition}
-%
-% \begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
-%   The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935}
-%   states that every permutation of
-%   length at least $pq+1$ must contain either the pattern
-%   $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
-%   The result now follows from the fact that
-%   the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
-% \qed
-% \end{proof}
-%
-% \begin{proposition}
-%   \label{proposition:patterns:upper bound}
-%   If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
-%   a permutation of $S_{n}$
-%   that does not contains a pattern
-%   of length $2k$ that is the union of two
-%   order-isomorphic patterns.
-% \end{proposition}
-%
-% \begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
-% The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
-% Let $\pi \in S_n$ be chosen randomly and uniformly.
-% For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
-% of $\pi$,
-% let $X_\sigma$ be the indicator variable that
-% $\sigma$ is the union of two order-ismorphic patterns.
-% According to Boole's inequality, we have
-% $P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
-% Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
-% the probability that $\pi$ contains a pattern of length $2k$
-% is the union of two order-ismorphic patterns is at most
-% $\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
-% Thus, with positive probability, no event $X_\sigma$ occurs and
-% there is a permutation of $S_n$ that does not contain a pattern
-% of length $2k$ that is the union of two order-isomorphic patterns.
-% \qed
-% \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Conclusion ----
-%%
-\section{Conclusion}
-\label{section:Conclusion}
-
-There are a number of further directions of investigation in this
-general subject. We mention several -~not necessarily new~- open
-problems that are, in our opinion, the most interesting. How many
-permutations of $S_{2n}$ are squares? How many $(213,231)$-avoiding
-permutations of $S_{2n}$ are squares? (Equivalently, by
-Proposition~\ref{prop:bijection_binary_to_permutations_squares},
-how many binary strings of length $2n$ are squares; Problem~4
-in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of
-deciding whether a $(213,231)$-avoiding permutation is a square
-(Problem~4 in \cite{Henshall:Rampersad:Shallit:2011},
-see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?
-Given two permutations $\pi$ and~$\sigma$, how hard is the problem of
-deciding whether $\sigma$ is a square root of~$\pi$?
-
-Besides, in this work, we investigate some algebraic properties of the
-shuffle of permutations which have direct combinatorial consequences.
-However, this work raises several additional classical algebraic
-questions. Among these, one can ask for a complete algebraic study of
-$\QQ[S]$ as a graded associative algebra for the shuffle  product
-$\SHUFFLE$. Describing a generating family for $\QQ[S]$, defining
-multiplicative bases of $\QQ[S]$, and determining whether $\QQ[S]$ is
-free as an associative algebra are worthwhile questions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Bibliography
-%%
-
-\bibliographystyle{plain}
-\bibliography{biblio}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\newpage
-\section*{Appendix (Reviewers version only)}
-
-\begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
-    Let us first show that the image of $\BINTOPERM$ contains only
-    permutations avoiding $213$ and $231$. Let $u$ be a binary word,
-    $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
-    positions of the occurrences of $0$ (resp. $1$) in $u$. By
-    definition of $\BINTOPERM$, from left to right, the subword
-    $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
-    is decreasing, and all letters of $w$ are greater than those
-    of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
-    Then, since $v$ is increasing and $w$ is decreasing, there is an
-    occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
-    occurrence of $21$ (resp. $2$, $1$). All these three cases
-    contradict the fact that all letters of $w$ are greater than
-    those of $v$. A similar argument shows that $\pi$ avoids~$231$
-    as well.
-    \smallskip
-
-    Finally, observe that any permutation $\pi$ avoiding $213$ and
-    $231$ necessarily starts by the smallest possible letter or the
-    greatest possible letter. This property is then true for the
-    suffix of $\pi$ obtained by deleting its first letter,
-    and so on for all of its suffixes. Thus, by
-    replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
-    role of a smallest (resp. greatest) letter, one obtains a binary
-    word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
-    avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
-    Since $u$ is a square binary word, there is a binary word $v$
-    such that $u \in v \shuffle v$. Then, there are two disjoint
-    sets $P$ and $Q$ of positions of letters of $u$ such that
-    $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
-    words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
-    same standarized $\sigma$. Hence, and by definition of
-    the shuffle product of permutations, $\BINTOPERM(u)$ appears in
-    $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
-    square permutation.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
-    Let $\pi$ be a square permutation avoiding $213$ and $231$. By
-    Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
-    the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
-    well-defined binary word. Since $\pi$ is a square permutation,
-    there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
-    of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
-    order-isomorphic. This implies, by the definitions of $\BINTOPERM$
-    and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
-    is a square binary word.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lem:endomorphisms}]
-    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
-    associative algebras, we have to prove that for all permutations
-    $\sigma$ and $\nu$,
-    \begin{equation}
-        \phi_j(\sigma \SHUFFLE \nu) =
-        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
-    \end{equation}
-    We shall proceed here as we have done during the proof of
-    Proposition~\ref{prop:shuffle_associative_commutative} by showing
-    that $\phi_j$ is a morphism of coalgebras, that is
-    \begin{equation}
-        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
-    \end{equation}
-    In the sequel, $\pi$ is a permutation.
-    \smallskip
-
-    If $P$ is a set of indexes of letters of $\pi$, we denote by
-    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
-    the operation $\widetilde{\,}$ defines a bijection on the set of the
-    subsets of $[|\pi|]$, and since the standardization operation
-    commutes with the mirror operation on words without multiple
-    occurrence of a letter, we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_1(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_1(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|P_1}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \widetilde{\STD\left(\pi_{|P_1}\right)}
-        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
-    $\phi_1$ is a morphism of associative algebras.
-    \smallskip
-
-    Next, since by definition of the complementation operation on
-    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
-    we have $\tau(i) < \tau(k)$ if and only if $\bar \tau(i) > \bar \tau(k)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_2(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_2(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\bar \pi_{|P_1}\right)
-        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
-    $\phi_2$ is a morphism of associative algebras.
-    \smallskip
-
-    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
-    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
-    $\{\tau(i) : i \in P\}$. Since the map sending a subset $P$ of
-    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
-    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_3(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_3(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|P_1}\right)
-        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
-        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right)^{-1}
-        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
-    $\phi_3$ is a morphism of associative algebras.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lemma:matching}]
-    The proof is, naturally enough, a slight extension of
-    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
-    for the sake of completeness.
-
-    (\emph{1} $\Rightarrow$ \emph{2}).
-    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
-    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
-    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
-    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
-    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
-    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
-    with vertices
-    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
-    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
-    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
-    matching since $I^1 \cap I^2 = \emptyset$, and
-    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
-    Indeed, if it is not the case then there would exist two edges
-    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
-    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
-    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
-    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
-    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
-    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
-    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
-    elements (at position $j$ and $k$) in two patterns that are
-    order-isomorphic to $\sigma$.
-
-    (\emph{2} $\Rightarrow$ \emph{1}).
-    Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$, and
-    $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$.
-    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
-    endpoints of the edges in $\mathcal{M}$ assuming the positions
-    $1, 2, \dots, 2$ are drawn along an horizontal line.
-    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
-    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
-    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
-    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
-    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
-    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
-    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
-    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
-    contradiction, that this is not the case. Therefore, there exist two
-    positions $j$ and $k$, say $j < k$, such that
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
-    But since $\mathcal{M}$ is containment-free, it follows that both
-    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
-    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
-    since for any two distinct edges $(a, b)$ and $(a', b')$ in
-    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
-    $\pi(b) < \pi(b')$.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
-  We first prove \emph{1}. Suppose, aiming at a contradiction, that
-  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
-  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
-  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
-  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
-  $\sigma_2$ is decreasing), which contradicts our hypothesis about
-  $\mathcal{M}$.
-
-  As for \emph{2}, suppose, still aiming at a contradiction, that there
-  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
-  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
-  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
-  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
-  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
-  \qed
-\end{proof}
-\bigskip
+\end{document}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-09-25 submission.tex
reduce figure

@@
 
-\begin{lemma}
-  \label{lemma:matching}
-  Let $\pi$ be a permutation. The following statements are equivalent:
-  \begin{enumerate}
-    \item The permutation $\pi$ is a square.
-    \item There exists a perfect containment-free matching $\mathcal{M}$
-    of $K_\pi$ such that for any two distinct edges $(a, b)$ and
-    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
-    if $\pi(b) < \pi(b')$.
-  \end{enumerate}
-\end{lemma}
-
-Let $\pi$ be a permutation. For the sake of clarity, we will say that a
-bunch of consecutive positions $P$ of $\pi$ is \emph{above} (resp.
-\emph{below}) above another bunch of consecutive positions $P'$ in $\pi$
-if $\pi(i) > \pi(j)$ (resp. $\pi(i) < \pi(j)$) for every $i \in P$ and
-every $j \in P'$. For example, $\sigma_1$ is above $\sigma_2$ (in an
-equivalent manner, $\sigma_2$ is below $\sigma_1$) in
-Fig.~\ref{subfig:increasing before and above decreasing}, whereas
-$\sigma_1$ is below $\sigma_2$ (in an equivalent manner, $\sigma_2$ is
-above $\sigma_1$) in Fig.~\ref{subfig:decreasing before and below
-increasing}.
 \begin{figure}[t!]
@@
 
+\begin{lemma}
+  \label{lemma:matching}
+  Let $\pi$ be a permutation. The following statements are equivalent:
+  \begin{enumerate}
+    \item The permutation $\pi$ is a square.
+    \item There exists a perfect containment-free matching $\mathcal{M}$
+    of $K_\pi$ such that for any two distinct edges $(a, b)$ and
+    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
+    if $\pi(b) < \pi(b')$.
+  \end{enumerate}
+\end{lemma}
+
+Let $\pi$ be a permutation. For the sake of clarity, we will say that a
+bunch of consecutive positions $P$ of $\pi$ is \emph{above} (resp.
+\emph{below}) above another bunch of consecutive positions $P'$ in $\pi$
+if $\pi(i) > \pi(j)$ (resp. $\pi(i) < \pi(j)$) for every $i \in P$ and
+every $j \in P'$. For example, $\sigma_1$ is above $\sigma_2$ (in an
+equivalent manner, $\sigma_2$ is below $\sigma_1$) in
+Fig.~\ref{subfig:increasing before and above decreasing}, whereas
+$\sigma_1$ is below $\sigma_2$ (in an equivalent manner, $\sigma_2$ is
+above $\sigma_1$) in Fig.~\ref{subfig:decreasing before and below
+increasing}.
+
 Before proving hardness, we give an easy lemma that will prove extremely
@@
     \begin{tikzpicture}[
-      scale=.9,
+      scale=.8,
       >=stealth',
@@
       \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
-      \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
-      \node (claim edge 1) at (1.5,13) {Claim~\ref{claim:all (nu_1, nu'_1)-edges}};
+      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
+      % \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
+      % \node (claim edge 1) at (1.5,13) {Claim~\ref{claim:all (nu_1, nu'_1)-edges}};
       % 2
       \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
-      \node (claim edge 2) at (4,4) {Claim~\ref{claim:(nu_2, nu'_2)-edge}};
-      \node (claim edge 2) at (4,3.5) {Claim~\ref{claim:all (nu_2, nu'_2)-edges}};
+      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
+      % \node (claim edge 2) at (4,4) {Claim~\ref{claim:(nu_2, nu'_2)-edge}};
+      % \node (claim edge 2) at (4,3.5) {Claim~\ref{claim:all (nu_2, nu'_2)-edges}};
       % 3
       \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
-      \node (claim edge 1) at (5.5,11.5) {Claim~\ref{claim:(nu_3, nu'_3)-edge}};
+      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
+      % \node (claim edge 1) at (5.5,11.5) {Claim~\ref{claim:(nu_3, nu'_3)-edge}};
       % 4
       \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
-      \node (claim edge 1) at (8.5,8.5) {Claim~\ref{claim:(nu_4, nu'_4)-edge}};
+      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
+      % \node (claim edge 1) at (8.5,8.5) {Claim~\ref{claim:(nu_4, nu'_4)-edge}};
       % sigma 1 - pi 1
@@
   Define
-  \begin{align*}
-  N_4 &= 2(2n + 2k + 4) + 1  = 4n + 4k + 9 \\
-  N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45 \\
-  N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225 \\
-  N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325\text{.}
-  \end{align*}
+  % \begin{align*}
+  % N_4 &= 2(2n + 2k + 4) + 1  = 4n + 4k + 9 \\
+  % N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45 \\
+  % N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225 \\
+  % N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325\text{.}
+  % \end{align*}
+  $N_4 = 2(2n + 2k + 4) + 1  = 4n + 4k + 9$,
+  $N_3 = 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45$,
+  $N_2 = 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225$ and
+  $N_1 = 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325$.
   Notice that $N_1, N_2, N_3$ and $N_4$ are polynomial in $n$.

### 2015-09-25 submission.tex
correction pi prime

@@
   It is immediate that H can be constructed in polynomial-time in $n$ and $k$.
-  We claim that $\sigma$ occurs in $\pi$ if and only if
+  % We claim that $\sigma$ occurs in $\pi$ if and only if
+  % there exists a containment-free perfect matching
+  % $\mathcal{M}$ in $G_\mu$ such that
+  % for any two disctinct edges
+  % $(a, b)$ and $(a', b')$ in $\mathcal{M}$
+  % we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+  It can be shown (see Appendix) that
+  that $\sigma$ occurs in $\pi$ if and only if
   there exists a containment-free perfect matching
@@
   we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
-
+  \qed
 \end{proof}

### 2015-09-25 submission.tex
Conclusion

--- 2015-09-25 submission.tex
@@
       % grid
-      \draw[step=1cm,black!30,thin,fill=black!10] (0,0) grid (12,12);
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
       % row size
@@
 
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% %%
-% %% ---- Patterns ----
-% %%
-% \section{Patterns}
-% \label{section:Patterns}
-%
-% \begin{proposition}
-%   \label{proposition:patterns:lower bound}
-%   Let $\pi \in S_{n}$.
-%   Then $\pi$ contains a pattern of size at least
-%   $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
-%   is the union of two order-isomorphic patterns.
-% \end{proposition}
-%
-% \begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
-%   The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935}
-%   states that every permutation of
-%   length at least $pq+1$ must contain either the pattern
-%   $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
-%   The result now follows from the fact that
-%   the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
-% \qed
-% \end{proof}
-%
-% \begin{proposition}
-%   \label{proposition:patterns:upper bound}
-%   If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
-%   a permutation of $S_{n}$
-%   that does not contains a pattern
-%   of length $2k$ that is the union of two
-%   order-isomorphic patterns.
-% \end{proposition}
-%
-% \begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
-% The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
-% Let $\pi \in S_n$ be chosen randomly and uniformly.
-% For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
-% of $\pi$,
-% let $X_\sigma$ be the indicator variable that
-% $\sigma$ is the union of two order-ismorphic patterns.
-% According to Boole's inequality, we have
-% $P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
-% Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
-% the probability that $\pi$ contains a pattern of length $2k$
-% is the union of two order-ismorphic patterns is at most
-% $\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
-% Thus, with positive probability, no event $X_\sigma$ occurs and
-% there is a permutation of $S_n$ that does not contain a pattern
-% of length $2k$ that is the union of two order-isomorphic patterns.
-% \qed
-% \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Conclusion ----
-%%
-\section{Conclusion}
-\label{section:Conclusion}
-
-There are a number of further directions of investigation in this
-general subject. We mention several -~not necessarily new~- open
-problems that are, in our opinion, the most interesting. How many
-permutations of $S_{2n}$ are squares? How many $(213,231)$-avoiding
-permutations of $S_{2n}$ are squares? (Equivalently, by
-Proposition~\ref{prop:bijection_binary_to_permutations_squares},
-how many binary strings of length $2n$ are squares; Problem~4
-in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of
-deciding whether a $(213,231)$-avoiding permutation is a square
-(Problem~4 in \cite{Henshall:Rampersad:Shallit:2011},
-see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?
-Given two permutations $\pi$ and~$\sigma$, how hard is the problem of
-deciding whether $\sigma$ is a square root of~$\pi$?
-
-Besides, in this work, we investigate some algebraic properties of the
-shuffle of permutations which have direct combinatorial consequences.
-However, this work raises several additional classical algebraic
-questions. Among these, one can ask for a complete algebraic study of
-$\QQ[S]$ as a graded associative algebra for the shuffle  product
-$\SHUFFLE$. Describing a generating family for $\QQ[S]$, defining
-multiplicative bases of $\QQ[S]$, and determining whether $\QQ[S]$ is
-free as an associative algebra are worthwhile questions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Bibliography
-%%
-
-\bibliographystyle{plain}
-\bibliography{biblio}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\section*{Appendix (Reviewers version only)}
-
-\begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
-    Let us first show that the image of $\BINTOPERM$ contains only
-    permutations avoiding $213$ and $231$. Let $u$ be a binary word,
-    $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
-    positions of the occurrences of $0$ (resp. $1$) in $u$. By
-    definition of $\BINTOPERM$, from left to right, the subword
-    $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
-    is decreasing, and all letters of $w$ are greater than those
-    of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
-    Then, since $v$ is increasing and $w$ is decreasing, there is an
-    occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
-    occurrence of $21$ (resp. $2$, $1$). All these three cases
-    contradict the fact that all letters of $w$ are greater than
-    those of $v$. A similar argument shows that $\pi$ avoids~$231$
-    as well.
-    \smallskip
-
-    Finally, observe that any permutation $\pi$ avoiding $213$ and
-    $231$ necessarily starts by the smallest possible letter or the
-    greatest possible letter. This property is then true for the
-    suffix of $\pi$ obtained by deleting its first letter,
-    and so on for all of its suffixes. Thus, by
-    replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
-    role of a smallest (resp. greatest) letter, one obtains a binary
-    word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
-    avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
-    Since $u$ is a square binary word, there is a binary word $v$
-    such that $u \in v \shuffle v$. Then, there are two disjoint
-    sets $P$ and $Q$ of positions of letters of $u$ such that
-    $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
-    words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
-    same standarized $\sigma$. Hence, and by definition of
-    the shuffle product of permutations, $\BINTOPERM(u)$ appears in
-    $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
-    square permutation.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
-    Let $\pi$ be a square permutation avoiding $213$ and $231$. By
-    Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
-    the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
-    well-defined binary word. Since $\pi$ is a square permutation,
-    there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
-    of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
-    order-isomorphic. This implies, by the definitions of $\BINTOPERM$
-    and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
-    is a square binary word.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lem:endomorphisms}]
-    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
-    associative algebras, we have to prove that for all permutations
-    $\sigma$ and $\nu$,
-    \begin{equation}
-        \phi_j(\sigma \SHUFFLE \nu) =
-        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
-    \end{equation}
-    We shall proceed here as we have done during the proof of
-    Proposition~\ref{prop:shuffle_associative_commutative} by showing
-    that $\phi_j$ is a morphism of coalgebras, that is
-    \begin{equation}
-        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
-    \end{equation}
-    In the sequel, $\pi$ is a permutation.
-    \smallskip
-
-    If $P$ is a set of indexes of letters of $\pi$, we denote by
-    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
-    the operation $\widetilde{\,}$ defines a bijection on the set of the
-    subsets of $[|\pi|]$, and since the standardization operation
-    commutes with the mirror operation on words without multiple
-    occurrence of a letter, we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_1(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_1(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|P_1}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \widetilde{\STD\left(\pi_{|P_1}\right)}
-        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
-    $\phi_1$ is a morphism of associative algebras.
-    \smallskip
-
-    Next, since by definition of the complementation operation on
-    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
-    we have $\tau(i) < \tau(k)$ if and only if $\bar \tau(i) > \bar \tau(k)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_2(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_2(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\bar \pi_{|P_1}\right)
-        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
-    $\phi_2$ is a morphism of associative algebras.
-    \smallskip
-
-    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
-    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
-    $\{\tau(i) : i \in P\}$. Since the map sending a subset $P$ of
-    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
-    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_3(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_3(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|P_1}\right)
-        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
-        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right)^{-1}
-        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
-    $\phi_3$ is a morphism of associative algebras.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lemma:matching}]
-    The proof is, naturally enough, a slight extension of
-    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
-    for the sake of completeness.
-
-    (\emph{1} $\Rightarrow$ \emph{2}).
-    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
-    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
-    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
-    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
-    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
-    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
-    with vertices
-    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
-    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
-    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
-    matching since $I^1 \cap I^2 = \emptyset$, and
-    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
-    Indeed, if it is not the case then there would exist two edges
-    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
-    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
-    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
-    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
-    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
-    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
-    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
-    elements (at position $j$ and $k$) in two patterns that are
-    order-isomorphic to $\sigma$.
-
-    (\emph{2} $\Rightarrow$ \emph{1}).
-    Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$, and
-    $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$.
-    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
-    endpoints of the edges in $\mathcal{M}$ assuming the positions
-    $1, 2, \dots, 2$ are drawn along an horizontal line.
-    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
-    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
-    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
-    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
-    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
-    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
-    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
-    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
-    contradiction, that this is not the case. Therefore, there exist two
-    positions $j$ and $k$, say $j < k$, such that
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
-    But since $\mathcal{M}$ is containment-free, it follows that both
-    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
-    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
-    since for any two distinct edges $(a, b)$ and $(a', b')$ in
-    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
-    $\pi(b) < \pi(b')$.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
-  We first prove \emph{1}. Suppose, aiming at a contradiction, that
-  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
-  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
-  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
-  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
-  $\sigma_2$ is decreasing), which contradicts our hypothesis about
-  $\mathcal{M}$.
-
-  As for \emph{2}, suppose, still aiming at a contradiction, that there
-  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
-  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
-  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
-  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
-  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
-  \qed
-\end{proof}
-
-\bigskip
-
-\begin{proof}[of Proposition~\ref{proposition:hardness}]
   Suppose first that $\sigma$ occurs in $\pi$ and fix an occurrence.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% %%
+% %% ---- Patterns ----
+% %%
+% \section{Patterns}
+% \label{section:Patterns}
+%
+% \begin{proposition}
+%   \label{proposition:patterns:lower bound}
+%   Let $\pi \in S_{n}$.
+%   Then $\pi$ contains a pattern of size at least
+%   $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
+%   is the union of two order-isomorphic patterns.
+% \end{proposition}
+%
+% \begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
+%   The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935}
+%   states that every permutation of
+%   length at least $pq+1$ must contain either the pattern
+%   $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
+%   The result now follows from the fact that
+%   the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
+% \qed
+% \end{proof}
+%
+% \begin{proposition}
+%   \label{proposition:patterns:upper bound}
+%   If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
+%   a permutation of $S_{n}$
+%   that does not contains a pattern
+%   of length $2k$ that is the union of two
+%   order-isomorphic patterns.
+% \end{proposition}
+%
+% \begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
+% The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
+% Let $\pi \in S_n$ be chosen randomly and uniformly.
+% For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
+% of $\pi$,
+% let $X_\sigma$ be the indicator variable that
+% $\sigma$ is the union of two order-ismorphic patterns.
+% According to Boole's inequality, we have
+% $P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
+% Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
+% the probability that $\pi$ contains a pattern of length $2k$
+% is the union of two order-ismorphic patterns is at most
+% $\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
+% Thus, with positive probability, no event $X_\sigma$ occurs and
+% there is a permutation of $S_n$ that does not contain a pattern
+% of length $2k$ that is the union of two order-isomorphic patterns.
+% \qed
+% \end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% ---- Conclusion ----
+%%
+\section{Conclusion}
+\label{section:Conclusion}
+
+% Unifier (une seule partie)
+
+There are a number of further directions of investigation in this
+general subject. They cover several areas: algorithmic, combinatorics,
+and algebra. Let us mention several -~not necessarily new~- open
+problems that are, in our opinion, the most interesting. How many
+permutations of $S_{2n}$ are squares? How many $(213,231)$-avoiding
+permutations of $S_{2n}$ are squares? (Equivalently, by
+Proposition~\ref{prop:bijection_binary_to_permutations_squares},
+how many binary strings of length $2n$ are squares; Problem~4
+in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of
+deciding whether a $(213,231)$-avoiding permutation is a square
+(Problem~4 in \cite{Henshall:Rampersad:Shallit:2011},
+see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?
+Given two permutations $\pi$ and~$\sigma$, how hard is the problem of
+deciding whether $\sigma$ is a square root of~$\pi$?
+%Besides, in this work, we investigate some algebraic properties of the
+%shuffle of permutations which have direct combinatorial consequences.
+%However, this work raises several additional classical algebraic
+%questions. Among these,
+Moreover, one can ask for a complete algebraic study of
+$\QQ[S]$ as a graded associative algebra for the shuffle  product
+$\SHUFFLE$. Describing a generating family for $\QQ[S]$, defining
+multiplicative bases of $\QQ[S]$, and determining whether $\QQ[S]$ is
+free as an associative algebra are worthwhile questions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+%% Bibliography
+%%
+
+\bibliographystyle{plain}
+\bibliography{biblio}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\newpage
+\section*{Appendix (Reviewers version only)}
+
+\begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
+    Let us first show that the image of $\BINTOPERM$ contains only
+    permutations avoiding $213$ and $231$. Let $u$ be a binary word,
+    $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
+    positions of the occurrences of $0$ (resp. $1$) in $u$. By
+    definition of $\BINTOPERM$, from left to right, the subword
+    $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
+    is decreasing, and all letters of $w$ are greater than those
+    of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
+    Then, since $v$ is increasing and $w$ is decreasing, there is an
+    occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
+    occurrence of $21$ (resp. $2$, $1$). All these three cases
+    contradict the fact that all letters of $w$ are greater than
+    those of $v$. A similar argument shows that $\pi$ avoids~$231$
+    as well.
+    \smallskip
+
+    Finally, observe that any permutation $\pi$ avoiding $213$ and
+    $231$ necessarily starts by the smallest possible letter or the
+    greatest possible letter. This property is then true for the
+    suffix of $\pi$ obtained by deleting its first letter,
+    and so on for all of its suffixes. Thus, by
+    replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
+    role of a smallest (resp. greatest) letter, one obtains a binary
+    word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
+    avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
+    Since $u$ is a square binary word, there is a binary word $v$
+    such that $u \in v \shuffle v$. Then, there are two disjoint
+    sets $P$ and $Q$ of positions of letters of $u$ such that
+    $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
+    words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
+    same standarized $\sigma$. Hence, and by definition of
+    the shuffle product of permutations, $\BINTOPERM(u)$ appears in
+    $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
+    square permutation.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
+    Let $\pi$ be a square permutation avoiding $213$ and $231$. By
+    Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
+    the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
+    well-defined binary word. Since $\pi$ is a square permutation,
+    there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
+    of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
+    order-isomorphic. This implies, by the definitions of $\BINTOPERM$
+    and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
+    is a square binary word.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Lemma~\ref{lem:endomorphisms}]
+    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
+    associative algebras, we have to prove that for all permutations
+    $\sigma$ and $\nu$,
+    \begin{equation}
+        \phi_j(\sigma \SHUFFLE \nu) =
+        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
+    \end{equation}
+    We shall proceed here as we have done during the proof of
+    Proposition~\ref{prop:shuffle_associative_commutative} by showing
+    that $\phi_j$ is a morphism of coalgebras, that is
+    \begin{equation}
+        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
+    \end{equation}
+    In the sequel, $\pi$ is a permutation.
+    \smallskip
+
+    If $P$ is a set of indexes of letters of $\pi$, we denote by
+    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
+    the operation $\widetilde{\,}$ defines a bijection on the set of the
+    subsets of $[|\pi|]$, and since the standardization operation
+    commutes with the mirror operation on words without multiple
+    occurrence of a letter, we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_1(\pi))
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\phi_1(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\widetilde{\pi}_{|P_1}\right)
+        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
+        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \widetilde{\STD\left(\pi_{|P_1}\right)}
+        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
+    $\phi_1$ is a morphism of associative algebras.
+    \smallskip
+
+    Next, since by definition of the complementation operation on
+    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
+    we have $\tau(i) < \tau(k)$ if and only if $\bar \tau(i) > \bar \tau(k)$,
+    we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_2(\pi))
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\phi_2(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\bar \pi_{|P_1}\right)
+        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
+    $\phi_2$ is a morphism of associative algebras.
+    \smallskip
+
+    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
+    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
+    $\{\tau(i) : i \in P\}$. Since the map sending a subset $P$ of
+    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
+    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
+    we have
+    \begin{equation} \begin{split}
+        \Delta(\phi_3(\pi))
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\phi_3(\pi)_{|P_1}\right)
+        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\pi^{-1}_{|P_1}\right)
+        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
+        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \STD\left(\pi_{|P_1}\right)^{-1}
+        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
+        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
+        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
+        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
+        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
+    \end{split} \end{equation}
+    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
+    $\phi_3$ is a morphism of associative algebras.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Lemma~\ref{lemma:matching}]
+    The proof is, naturally enough, a slight extension of
+    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
+    for the sake of completeness.
+
+    (\emph{1} $\Rightarrow$ \emph{2}).
+    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
+    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
+    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
+    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
+    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
+    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
+    with vertices
+    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
+    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
+    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
+    matching since $I^1 \cap I^2 = \emptyset$, and
+    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
+    Indeed, if it is not the case then there would exist two edges
+    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
+    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
+    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
+    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
+    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
+    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
+    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
+    elements (at position $j$ and $k$) in two patterns that are
+    order-isomorphic to $\sigma$.
+
+    (\emph{2} $\Rightarrow$ \emph{1}).
+    Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
+    \;\text{with}\; j < k\}$, and
+    $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
+    \;\text{with}\; j < k\}$.
+    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
+    endpoints of the edges in $\mathcal{M}$ assuming the positions
+    $1, 2, \dots, 2$ are drawn along an horizontal line.
+    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
+    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
+    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
+    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
+    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
+    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
+    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
+    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
+    contradiction, that this is not the case. Therefore, there exist two
+    positions $j$ and $k$, say $j < k$, such that
+    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
+    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
+    But since $\mathcal{M}$ is containment-free, it follows that both
+    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
+    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
+    since for any two distinct edges $(a, b)$ and $(a', b')$ in
+    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
+    $\pi(b) < \pi(b')$.
+    \qed
+\end{proof}
+\bigskip
+
+\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
+  We first prove \emph{1}. Suppose, aiming at a contradiction, that
+  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
+  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
+  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
+  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
+  $\sigma_2$ is decreasing), which contradicts our hypothesis about
+  $\mathcal{M}$.
+
+  As for \emph{2}, suppose, still aiming at a contradiction, that there
+  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
+  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
+  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
+  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
+  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
+  \qed
+\end{proof}
+\bigskip
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2015-09-25 conclusion.tex

--- 2015-09-25 submission.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[a4paper]{llncs}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\usepackage[utf8x]{inputenc}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{llncsdoc}
-\usepackage[english]{babel}
-\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
-\usepackage{algorithmic}
-\usepackage{float}
-\usepackage{todonotes}
-\usepackage{subfigure}
-\usepackage{enumitem}
-\usepackage{shuffle}
-\usepackage{multirow}
-
-\usepackage{tikz}
-\usetikzlibrary{matrix, calc, arrows, shapes.callouts}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{complexity}
-\input{graph}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \renewcommand{\floatpagefraction}{.9}
-% \renewcommand{\textfraction}{.1}
-\renewcommand{\arraystretch}{1.2}
-\newcommand{\rt}{\textcolor{red}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\DeclareMathOperator{\LEFT}{\ell}
-\DeclareMathOperator{\RIGHT}{r}
-\DeclareMathOperator{\QQ}{\mathbb{Q}}
-\DeclareMathOperator{\STD}{\mathrm{s}}
-\DeclareMathOperator{\SHUFFLE}{\bullet}
-\DeclareMathOperator{\BINTOPERM}{\mathrm{btp}}
-\DeclareMathOperator{\PERMTOBIN}{\mathrm{ptb}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% 1 2 2 1
-\newcommand{\MWithContainment}{%
-\begin{tikzpicture}[scale=0.1,inner sep=.8pt,node distance=.25cm]
-\draw node [draw,circle,fill=black] (U1)               {};
-\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
-\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
-\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
-\draw (U2.north) .. controls ($ (U2.north) + (0,1.75) $) and ($ (U3.north) + (0,1.75) $) .. (U3.north);
-\draw (U1.north) .. controls ($ (U1.north) + (0,2.5) $) and ($ (U4.north) + (0,2.5) $) .. (U4.north);
-\end{tikzpicture}
-}% end \MWithContainment
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newtheorem{fact}{Fact}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\makeatletter
-\newcommand{\pushright}[1]{\ifmeasuring@#1\else\omit\hfill$\displaystyle#1$\fi\ignorespaces}
-\newcommand{\pushleft}[1]{\ifmeasuring@#1\else\omit$\displaystyle#1$\hfill\fi\ignorespaces}
-\makeatother
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{%
-Unshuffling Permutations}%
-
-\author{%
-  Samuele Giraudo \and
-  St\'ephane Vialette
-}% end author
-\institute{%
-  Universit\'e Paris-Est, LIGM (UMR 8049), CNRS, UPEM, ESIEE Paris, ENPC,
-  F-77454, Marne-la-Vallée, France\\
-  \email{samuele.giraudo@univ-mlv.fr} \\
-  \email{vialette@univ-mlv.fr}
-}% end institute
-\date{\today}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{abstract}
-    A permutation is said to be a square if it can be obtained by
-    shuffling two order-isomorphic patterns. The definition is intended
-    to be the natural counterpart to the ordinary shuffle of words and
-    languages. In this paper, we tackle the problem of recognizing square
-    permutations from both the point of view of algebra and algorithmic.
-    On the one hand, we present some algebraic and combinatorial
-    properties of the shuffle product of permutations. We follow an
-    unusual line consisting in defining the shuffle of permutations by
-    means of an unshuffling operator, known as a coproduct. This
-    strategy allows to obtain easy proofs for algebraic and combinatorial
-    properties of our shuffle product. We besides exhibit a bijection
-    between square $(213,231)$-avoiding permutations and square binary
-    words.
-    On the other hand,
-    using a containment-free perfect matching criterion,
-    we prove that recognizing square permutations is \NPC.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%%
-%% ---- Introduction ----
-%%
-\section{Introduction}
-\label{section:Introduction}
-The {\em shuffle product}, denoted $\shuffle$, is a very famous operation on words
-first defined by Eilenberg and Mac Lane~\cite{Eilenberg:MacLane:1953}.
-Given three words $u$, $v_1$, and $v_2$, $u$ is said to be a \emph{shuffle}
-of $v_1$ and $v_2$ if it can be formed by interleaving the letters from
-$v_1$ and $v_2$ in a way that maintains the left-to-right ordering of the
-letters from each word.
-Besides purely combinatorial questions,
-the shuffle product of words leads to the following computational ones:
-\begin{enumerate}
-    \item \label{item:problem_1}
-    Given two words $v_1$ and $v_2$, compute the set $v_1 \shuffle v_2$.
-    \item \label{item:problem_2}
-    Given three words $u$, $v_1$, and $v_2$, decide if $u$ is a shuffle
-    of $v_1$ and $v_2$.
-    \item \label{item:problem_3}
-    Given words $u$, $v_1$, \dots, $v_k$, decide if $u$ is in
-    $v_1 \shuffle \dots \shuffle v_k$.
-    \item \label{item:problem_4}
-    Given a word $u$, decide if there is a word $v$ such that $u$ is
-    in $v \shuffle v$.
-\end{enumerate}
-Even if these problems seem similar, they are very different in terms
-of complexity. Let us now review some facts about these. In what follows,
-$n$ denotes the size of $u$ and $m_i$ denotes the size of each $v_i$.
-A solution to Problem~\ref{item:problem_1} can be computed in
-% \begin{equation}
-%     O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)
-% \end{equation}
-$O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)$
-time~\cite{Spehner:TCS:1986}. An improvement and a generalization
-of~Problem~\ref{item:problem_1} has been proposed
-in~\cite{Allauzen:IGM:2000}, where it is proved that given words
-$v_1$, \dots, $v_k$, the iterated shuffle
-$v_1 \shuffle \dots \shuffle v_k$ can be computed in
-% \begin{equation}
-%     O\left(\binom{m_1 + \dots + m_k}{m_1, \dots, m_k}\right)
-% \end{equation}
-$O\left(\binom{m_1 + \dots + m_k}{m_1, \dots, m_k}\right)$
-time. Problem~\ref{item:problem_2} is in \Pclass; it is indeed a
-classical textbook exercise to design an efficient dynamic programming
-algorithm solving it. It can be tested in $O\left(n^2 / \log(n)\right)$
-time~\cite{Leeuwen:Nivat:IPL:1982}. To the best of our knowledge, the
-first $O(n^2)$ time algorithm for this problem appeared
-in~\cite{Mansfield:DAM:1983}. This algorithm can easily be extended to
-check in polynomial-time whether or not a word is in the shuffle of any
-fixed number of given words. Nevertheless, Problem~\ref{item:problem_3}
-is \NPC~\cite{Mansfield:DAM:1983,Warmuth:Haussler:JCSS:1984}. This
-remains true even if the ground alphabet has size
-$3$~\cite{Warmuth:Haussler:JCSS:1984}. Of particular interest,
-it is shown in~\cite{Warmuth:Haussler:JCSS:1984} that
-Problem~\ref{item:problem_3} remains \NPC even if all the words $v_i$,
-$i \in [k]$, are identical, thereby proving that, for two words $u$ and
-$v$, it is \NPC to decide whether or not $u$ is in the iterated shuffle
-of $v$. Again, this remains true even if the ground alphabet has size $3$.
-Let us now finally focus on Problem~\ref{item:problem_4}. It is shown
-in~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013} that it is \NPC to
-decide if a word $u$ is a \emph{square} (w.r.t. the shuffle), that is
-a word $u$ with the property that there exists a word $v$ such that $u$
-is a shuffle of $v$ with itself. Hence, Problem~\ref{item:problem_4}
-is \NPC.
-\smallskip
-
-This paper is intended to study a natural generalization of $\shuffle$,
-denoted by $\SHUFFLE$, as a shuffle of permutations. Roughly speaking,
-given three permutations $\pi$, $\sigma_1$, and $\sigma_2$, $\pi$ is
-said to be a {\em shuffle} of $\sigma_1$ and $\sigma_2$ if $\pi$ (viewed
-as a word) is a shuffle of two words that are order-isomorphic to
-$\sigma_1$ and $\sigma_2$. Our intention in this paper is to study this
-shuffle product of permutations $\SHUFFLE$ both from a combinatorial and
-from a computational point of view by focusing on {\em square}
-permutations, that are permutations $\pi$ being in the shuffle of a
-permutation $\sigma$ with itself. The shuffle product defined and
-considered in this paper is different from other existing and well-known
-shuffle products on permutations. For instance, in~\cite{DHT:IJAC:2002},
-the authors define the {\em shifted shuffle product}. For this product,
-$\pi$ is a shuffle of $\sigma_1$ and $\sigma_2$ if $\pi$ is in the
-shuffle, as words, of $\sigma_1$ and the word obtained by incrementing
-all the letters of $\sigma_2$ by the size of $\sigma_1$. It is a simple
-exercise to prove that, given three permutations $\pi$, $\sigma_1$, and
-$\sigma_2$, deciding if $\pi$ is a in the shifted shuffle of $\sigma_1$
-and $\sigma_2$ is in~\Pclass.
-\smallskip
-
-This paper is organized as follows. In
-Section~\ref{section:Shuffle product on permutations} we provide a
-precise definition of $\SHUFFLE$. This definition passes through the
-preliminary definition of an operator~$\Delta$, allowing to
-{\em unshuffle} permutations. This operator is in fact a coproduct,
-endowing the linear span of all permutations with a coalgebra structure
-(see~\cite{Joni:Rota:1979} or~\cite{Grinberg:Reiner:2014} for the
-definition of these algebraic structures). By duality, the unshuffling
-operator $\Delta$ leads to the definition of our shuffle operation on
-permutations. This approach has many advantages. First, some
-combinatorial properties of $\SHUFFLE$ depend on properties of $\Delta$
-and are more easy to prove on the coproduct side. Second, this way of
-doing allows to obtain a clear description of the multiplicities of the
-elements appearing in the shuffle of two permutations, which are worthy
-of interest from a combinatorial point of view.
-Section~\ref{section:Binary square words and permutations} is devoted to
-show that the problems related to the shuffle of words has links with
-the shuffle of permutations. In particular, we show that binary words
-that are square are in one-to-one correspondence with square
-permutations avoiding some patterns
-(Proposition~\ref{prop:bijection_binary_to_permutations_squares}). Next,
-Section~\ref{section:Algebraic issues} presents some algebraic and
-combinatorial properties of~$\SHUFFLE$. We show that $\SHUFFLE$ is
-associative and commutative
-(Proposition~\ref{prop:shuffle_associative_commutative}), and that if a
-permutation is a square, its mirror, complement, and inverse are also
-squares (Proposition~\ref{prop:square_stability}). Finally,
-Section~\ref{section:Algorithmic issues} presents the most import result
-of this paper: the fact that deciding if a permutation is a square is
-\NPC (Proposition~\ref{proposition:hardness}). This result is obtained
-by exhibiting a reduction from the pattern involvement
-problem~\cite{Bose:Buss:Lubiw:1998} which is \NPC.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Notations ----
-%%
-\section{Notations}
-\label{section:Notations}
-
-If $S$ is a finite set, the cardinality of $S$ is denoted by $|S|$,
-and if $P$ and $Q$ are two disjoint sets, $P \sqcup Q$ denotes the
-disjoint union of $P$ and $Q$. For any nonnegative integer $n$, $[n]$
-is the set $\{1, \dots, n\}$.
-
-We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
-Let us recall here the most important ones. Let $u$ be a word. The
-length of $u$ is denoted by $|u|$. The {\em empty word}, the only word
-of null length, is denoted by $\epsilon$. We denote by $\widetilde{u}$
-the {\em mirror image} of $u$, that is the word
-$u_{|u|} u_{|u| - 1} \dots u_1$. If $P$ is a subset of $[|u|]$, $u_{|P}$
-is the subword of $u$ consisting in the letters of $u$ at the positions
-specified by the elements of $P$. If $u$ is a word of integers and $k$
-is an integer, we denote by $u[k]$ the word obtained by incrementing by
-$k$ all letters of $u$. The {\em shuffle} of two words $u$ and $v$ is
-the set recursively defined by
-$u \shuffle \epsilon = \{u\} = \epsilon \shuffle u$ and
-$ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
-$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
-such that $u$ belongs to $v \shuffle v$.
-
-We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
-set of all permutations. In this paper, permutations of a size $n$ are
-specified by words of length $n$ on the alphabet $[n]$ and without
-multiple occurrence of a letter, so that all above definitions about
-words remain valid on permutations. The only difference lies on the
-fact that we shall denote by $\pi(i)$ (instead of $\pi_i$) the $i$-th
-letter of any permutation $\pi$. For any nonnegative integer $n$, we
-write $\nearrow_{n}$ (resp. $\searrow_{n}$) for the permutation
-$1 2 \dots n$ (resp. $n (n-1) \dots 1$). If $\pi$ is a permutation of
-$S_n$, we denote by $\bar \pi$ the {\em complementary} of $\pi$, that is
-the permutation satisfying $\bar \pi(i) = n - \pi(i) + 1$ for all
-$i \in [n]$. The {\em inverse} of $\pi$ is denoted by $\pi^{-1}$.
-
-If $u$ is a word of integers without multiple occurrences of a same
-letter, $\STD(u)$ is the {\em standarized} of $u$, that is the unique
-permutation of the same size as $u$ such that for all $i, j \in [|u|]$,
-$u_i < u_j$ if and only if $\STD(u)(i) < \STD(u)(j)$. In particular, the
-image of the map $\STD$ is the set $S$ of all permutations. Two words $u$
-and $v$ having the same standarized are {\em order-isomorphic}. If
-$\sigma$ is a permutation, there is an {\em occurrence} of (the
-{\em pattern}) $\sigma$ in $\pi$ if there is a set $P$ of indexes of
-letters of $\pi$ such that $\sigma$ and $\pi_{|P}$ are order-isomorphic.
-When $\pi$ does not admit any occurrence of $\sigma$, $\pi$ {\em avoids}
-$\sigma$.
-
-Let us now provide some definitions about graphs and matching that are used in
-the sequel. For a graph $G$, we denote $\VS{G}$ as the set of vertices
-and $\ES{G}$ as the set of edges of $G$. Let $u = u_1 u_2 \dots u_n$ be
-a word. The \emph{clique graph associated to $u$}, denoted by $K_u$, is
-the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$. The
-structure of this underlying graph is linear, \emph{i.e.}, the set of
-its vertices is equipped with a natural total order $<$ defined by
-$u_i < u_j$ if and only if $i < j$. For a linear graph, we write
-$(u_i, u_j)$ instead of $\{u_i, u_j\}$ for an edge if it is clear from
-the context that $i < j$. Recall that two edges of a graph are
-\emph{independent} if they do not share a common vertex, and that a
-\emph{matching} $\mathcal{M}$ in $G$ is a set of pairwise independent
-edges. A matching is \emph{perfect} if it covers all the vertices of the
-graph.
-For an edge $e = \{u,v\}$, we shall occasionally say that $u$ and $v$
-are \emph{involved} in $e$.
-In case the set of vertices is equipped with a total order, a
-matching $\mathcal{M}$ is said to be \emph{containment-free} if there do
-not exist (independent) edges $(u_i, u_j)$ and $(u_k, u_\ell)$ in
-$\mathcal{M}$ such that $u_i < u_k < u_\ell < u_j$ or
-$u_k < u_i < u_j < u_\ell$
-(\emph{i.e.}, $\mathcal{M}$ avoids~\MWithContainment).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Shuffle product on permutations ----
-%%
-\section{Shuffle product on permutations}
-\label{section:Shuffle product on permutations}
-
-The purpose of this section is to define a shuffle product $\SHUFFLE$
-on permutations. To define this shuffle product, we shall first define
-a coproduct denoted by $\Delta$, enabling to unshuffle permutations.
-By duality, $\Delta$ implies the definition of $\SHUFFLE$. The reason
-why we need to pass by the definition of $\Delta$ to define $\SHUFFLE$
-is justified by the fact that a lot of properties of $\SHUFFLE$ depend
-of properties of $\Delta$, and that this strategy allows to write concise
-and clear proofs of them. We invite the reader unfamiliar with the
-concepts of coproduct and duality to consult~\cite{Joni:Rota:1979}
-or~\cite{Grinberg:Reiner:2014}.
-
-Let use denote by $\QQ[S]$ the linear span of all permutations. We
-define a linear coproduct $\Delta$ on $\QQ[S]$ in the following way. For
-any permutation $\pi$, we set
-\begin{equation} \label{equ:unshuffling_coproduct}
-    \Delta(\pi) =
-    \sum_{P_1 \sqcup P_2 = [|\pi|]}
-    \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
-\end{equation}
-We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
-instance,
-\begin{equation}
-    \Delta(213) =
-    \epsilon \otimes 213 + 2. 1 \otimes 12 +
-    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
-\end{equation}
-\begin{equation}
-    \Delta(1234) =
-    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
-    4. 123 \otimes 1 + 1234 \otimes \epsilon,
-\end{equation}
-\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
-    \Delta(1432) & =
-    \epsilon \otimes 1432 + 3. {\bf 1 \otimes 132} + 1 \otimes 321 +
-    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
-    321 \otimes 1 + 1432 \otimes \epsilon.
-\end{split}\end{equation}
-Observe that the coefficient of the tensor $1 \otimes 132$ is $3$
-in~\eqref{equ:example_unshuffling_coproduct} because there are exactly
-three ways to extract from the permutation $1432$ two disjoint subwords
-respectively order-isomorphic to the permutations $1$ and $132$.
-
-As announced, let us now use $\Delta$ to define a shuffle product on
-permutations. As any coproduct, $\Delta$ leads to the definition of a
-product obtained by duality in the following way.
-From~\eqref{equ:unshuffling_coproduct}, for any permutation $\pi$, we
-have
-\begin{equation} \label{equ:schematic_coproduct}
-    \Delta(\pi) =
-    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
-    \sigma \otimes \nu,
-\end{equation}
-where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
-by definition of duality, the dual product of $\Delta$, denoted by
-$\SHUFFLE$, is a linear binary product on $\QQ[S]$. It satisfies, for
-any permutations $\sigma$ and $\nu$,
-\begin{equation}
-    \sigma \SHUFFLE \nu =
-    \sum_{\pi \in S}
-    \lambda_{\sigma, \nu}^\pi \; \pi,
-\end{equation}
-where the coefficients $\lambda_{\sigma, \nu}^\pi$ are the ones
-of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
-{\em shuffle product of permutations}. For instance,
-\begin{equation}\begin{split} \label{equ:example_shuffle_product}
-    12 \SHUFFLE 21 & =
-    1243 + 1324 + 2.1342 + 2.1423 + 3. {\bf 1432} + 2134 + 2.2314 \\
-    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
-    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
-\end{split}\end{equation}
-Observe that the coefficient $3$ of the permutation $1432$
-in~\eqref{equ:example_shuffle_product} comes from the fact that the
-coefficient of the tensor $12 \otimes 21$ is $3$
-in~\eqref{equ:example_unshuffling_coproduct}.
-
-Intuitively, this product shuffles the values and the positions of the
-letters of the permutations. As first and easy properties of $\SHUFFLE$,
-one can observe that the empty permutation $\epsilon$ is a unit for
-$\SHUFFLE$ and that this product is graded by the sizes of the
-permutations ({\em i.e.}, the product of a permutation of size $n$ with
-a permutation of size $m$ produces a sum of permutations of size $n + m$).
-Notice that there are other shuffle products on
-permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
-{\em convolution product}. Our shuffle product is different from these
-last two.
-
-We say that a permutation $\pi$ {\em appears} in the shuffle
-$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
-coefficient $\lambda_{\sigma, \nu}^\pi$ defined above is different from
-zero. In a more combinatorial way, this is equivalent to say that there
-are two sets $P_1$ and $P_2$ of disjoints indexes of letters of $\pi$
-satisfying $P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$
-is order-isomorphic to $\sigma$ and the subword $\pi_{|P_2}$ is
-order-isomorphic to $\nu$.
-
-Besides, a permutation $\pi$ is a {\em square} if there is a
-permutation $\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
-In this case, we say that $\sigma$ is a {\em square root} of $\pi$.
-Equivalently, $\pi$ is a square with $\sigma$ as square root if and only
-if in the expansion of $\Delta(\pi)$, there is a tensor
-$\sigma \otimes \sigma$ with a nonzero coefficient. In a more
-combinatorial way, this is equivalent to say that there are two sets
-$P_1$ and $P_2$ of disjoints indexes of letters of $\pi$ satisfying
-$P_1 \sqcup P_2 = [|\pi|]$ such that the subwords $\pi_{|P_1}$ and
-$\pi_{|P_2}$ are order-isomorphic. Computer experiments give us the
-first numbers of square permutations with respects to their size, which
-are, from size $0$ to $10$,
-\begin{equation}
-    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
-\end{equation}
-This sequence (and its subsequence obtained by removing the $0$'s) is for
-the time being not listed in~\cite{Slo}. The square permutations of
-sizes $0$ to $4$ are
-\smallskip
-
-\begin{tabular}{c|c|c}
-    Size $0$ \, & Size $2$ & Size $4$ \\ \hline
-    \multirow{2}{*}{\, $\epsilon$} &
-    \multirow{2}{*}{\, $12$, $21$ \,} &
-    $1234$, $1243$, $1423$, $1324$, $1342$, $4132$, $3124$, $3142$,
-    $3412$, $4312$, \\
-    & & $2134$, $2143$, $2413$, $4213$, $2314$, $2431$, $4231$, $3241$,
-    $3421$, $4321$
-\end{tabular}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Binary square words and permutations ----
-%%
-\section{Binary square words and permutations}
-\label{section:Binary square words and permutations}
-In this section, we shall show that the square binary words are in
-one-to-one correspondence with square permutations avoiding some
-patterns. This property establishes a link between the shuffle of binary
-words and our shuffle of permutations and allows to obtain a new
-description of square binary words.
-
-Let $u$ be a binary word of length $n$ with $k$ occurrences of $0$.
-We denote by $\BINTOPERM$ the map sending any such word $u$ to the
-permutation obtained by replacing from left to right each occurrence of
-$0$ in $u$ by $1$, $2$, \dots, $k$, and from right to left each
-occurrence of $1$ in $u$ by $k + 1$, $k + 2$, \dots, $n$. For instance,
-\begin{equation}
-    \BINTOPERM({\bf 1}00{\bf 1}0{\bf 1}{\bf 1}0{\bf 1}000) =
-    {\bf C} 12 {\bf B} 3 {\bf A} {\bf 9} 4 {\bf 8} 567.
-\end{equation}
-Observe that for any nonempty permutation $\pi$ in the image of
-$\BINTOPERM$, there is exactly one binary word $u$ such that
-$\BINTOPERM(u0) = \BINTOPERM(u1) = \pi$. In support of this observation,
-when $\pi$ has an even size, we denote by $\PERMTOBIN(\pi)$ the word $ua$
-such that $|ua|_0$ and $|ua|_1$ are both even, where $a \in \{0, 1\}$.
-
-\begin{proposition} \label{prop:bijection_binary_to_permutations_squares}
-    For any $n \geq 0$, the map $\BINTOPERM$ restricted to the set of
-    square binary words of length $2n$ is a bijection between this last
-    set and the set of square permutations of size $2n$ avoiding the
-    patterns $213$ and $231$.
-\end{proposition}
-\begin{proof}
-    [of Proposition~\ref{prop:bijection_binary_to_permutations_squares}]
-    The statement of the proposition is a consequence of the
-    following claims implying that $\PERMTOBIN$ is the inverse
-    map of $\BINTOPERM$ over the set of square binary words.
-    \begin{claim} \label{claim:binary_to_permutation_avoiding}
-        The image of $\BINTOPERM$ is the set of all permutations
-        avoiding $213$ and $231$.
-    \end{claim}
-
-    \begin{claim} \label{claim:square_binary_to_square_permutation}
-        If $u$ is a square binary word, $\BINTOPERM(u)$ is a square
-        permutation.
-    \end{claim}
-
-    \begin{claim} \label{claim:square_permutation_to_square_binary}
-        If $\pi$ is a square permutation avoiding $213$ and $231$,
-        $\PERMTOBIN(\pi)$ is a square binary word.
-    \end{claim}
-    \qed
-\end{proof}
-
-The number of square binary words is Sequence A191755 of~\cite{Slo}
-beginning by
-\begin{equation}
-1, 0, 2, 0, 6, 0, 22, 0, 82, 0, 320, 0, 1268, 0, 5102, 0, 020632.
-\end{equation}
-By Proposition~\ref{prop:bijection_binary_to_permutations_squares}, this
-is also the sequence enumerating square permutations avoiding $213$
-and $231$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Algebraic issues ----
-%%
-\section{Algebraic issues}
-\label{section:Algebraic issues}
-The aim of this section is to establish some of properties of the
-shuffle product of permutations $\SHUFFLE$. It is worth to note that, as
-we will see, algebraic properties of the unshuffling coproduct $\Delta$
-of permutations defined in
-Section~\ref{section:Shuffle product on permutations} lead to
-combinatorial properties of $\SHUFFLE$.
-
-\begin{proposition} \label{prop:shuffle_associative_commutative}
-    The shuffle product $\SHUFFLE$ of permutations is associative and
-    commutative.
-\end{proposition}
-\begin{proof}[of Proposition~\ref{prop:shuffle_associative_commutative}]
-    To prove the associativity of $\SHUFFLE$, it is convenient to show
-    that its dual coproduct $\Delta$ is coassociative, that is
-    \begin{equation}
-        (\Delta \otimes I) \Delta = (I \otimes \Delta) \Delta,
-    \end{equation}
-    where $I$ denote the identity map. This strategy relies on the fact
-    that a product is associative if and only if its dual coproduct is
-    coassociative. For any permutation $\pi$, we have
-    \begin{equation} \begin{split}
-    \label{equ:shuffle_associative_commutative}
-        (\Delta \otimes I) \Delta(\pi) & =
-        (\Delta \otimes I)
-        \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
-        & =
-        \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes I\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & =
-        \sum_{P_1 \sqcup P_2 = [|\pi|]} \;
-        \sum_{Q_1 \sqcup Q_2 = [|P_1|]}
-        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
-        \otimes
-        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_2}\right)
-        \otimes \STD\left(\pi_{|P_2}\right) \\
-        & =
-        \sum_{P_1 \sqcup P_2 \sqcup P_3 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right) \otimes
-        \STD\left(\pi_{|P_2}\right) \otimes
-        \STD\left(\pi_{|P_3}\right).
-    \end{split} \end{equation}
-    An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
-    is equal to the last member
-    of~\eqref{equ:shuffle_associative_commutative}, whence the
-    associativity of $\SHUFFLE$.
-    \smallskip
-
-    Finally, to prove the commutativity of $\SHUFFLE$, we shall show
-    that $\Delta$ is cocommutative, that is for any permutation $\pi$,
-    if in the expansion of $\Delta(\pi)$ there is a tensor
-    $\sigma \otimes \nu$ with a coefficient $\lambda$, there is in the
-    same expansion the tensor $\nu \otimes \sigma$ with the same
-    coefficient $\lambda$. Clearly, a product is commutative if and only
-    if its dual coproduct is cocommutative. Now, from the
-    definition~\eqref{equ:unshuffling_coproduct} of $\Delta$, one
-    observes that if the pair $(P_1, P_2)$ of subsets of $[|\pi|]$
-    contributes to the coefficient of
-    $\STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right)$,
-    the pair $(P_2, P_1)$ contributes to the coefficient of
-    $\STD\left(\pi_{|P_2}\right) \otimes \STD\left(\pi_{|P_1}\right)$.
-    This shows that $\Delta$ is cocommutative and hence, that $\SHUFFLE$
-    is commutative.
-    \qed
-\end{proof}
-
-Proposition~\ref{prop:shuffle_associative_commutative} implies that
-$\QQ[S]$ endowed with the unshuffling coproduct $\Delta$ is a
-coassociative cocommutative coalgebra, or in an equivalent way, that
-$\QQ[S]$ endowed with the shuffle product $\SHUFFLE$ is an associative
-commutative algebra.
-
-\begin{lemma} \label{lem:endomorphisms}
-    The three linear maps
-    \begin{equation}
-        \phi_1, \phi_2, \phi_3 : \QQ[S] \to \QQ[S]
-    \end{equation}
-    linearly sending a permutation $\pi$ to, respectively,
-    $\widetilde{\pi}$, $\bar \pi$, and $\pi^{-1}$ are endomorphims of
-    associative algebras.
-\end{lemma}
-
-We now use the algebraic properties of $\SHUFFLE$ exhibited by
-Lemma~\ref{lem:endomorphisms} to obtain combinatorial properties
-of square permutations.
-
-\begin{proposition} \label{prop:square_stability}
-    Let $\pi$ be a square permutation and $\sigma$ be a square root of
-    $\pi$. Then,
-    \begin{enumerate}[label={\it (\roman*)},fullwidth]
-        \item \label{item:square_stability_1}
-        the permutation $\widetilde{\pi}$ is a square and
-        $\widetilde{\sigma}$ is one of its square roots;
-        \item \label{item:square_stability_2}
-        the permutation $\bar \pi$ is a square and $\bar \sigma$ is one of
-        its square roots;
-        \item \label{item:square_stability_3}
-        the permutation $\pi^{-1}$ is a square and $\sigma^{-1}$ is one of
-        its square roots.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}[of Proposition~\ref{prop:square_stability}]
-    All statements~\ref{item:square_stability_1},
-    \ref{item:square_stability_2}, and~\ref{item:square_stability_3} are
-    consequences of Lemma~\ref{lem:endomorphisms}. Indeed,
-    since $\pi$ is a square permutation and $\sigma$ is a square root of
-    $\pi$, by definition, $\pi$ appears in the product
-    $\sigma \SHUFFLE \sigma$. Now, by Lemma~\ref{lem:endomorphisms},
-    for any $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
-    algebras from $\QQ[S]$ to $\QQ[S]$, $\phi_j$ commutes with the
-    shuffle product of permutations $\SHUFFLE$. Hence, in particular,
-    one has
-    \begin{equation}
-        \phi_j(\sigma \SHUFFLE \sigma) =
-        \phi_j(\sigma) \SHUFFLE \phi_j(\sigma).
-    \end{equation}
-    Then, since $\pi$ appears in $\sigma \SHUFFLE \sigma$, $\phi_j(\pi)$
-    appears in $\phi_j(\sigma \SHUFFLE \sigma)$ and appears also in
-    $\phi_j(\sigma) \SHUFFLE \phi_j(\sigma)$. This shows that
-    $\phi_j(\sigma)$ is a square root of $\phi_j(\pi)$ and
-    implies~\ref{item:square_stability_1}, \ref{item:square_stability_2},
-    and~\ref{item:square_stability_3}.
-    \qed
-\end{proof}
-
-Let us make an observation about Wilf-equivalence classes of
-permutations restrained on square permutations. The
-well-known~\cite{Simion:Schmidt:EJC:1985} fact that there is a single
-Wilf-equivalence class of permutations of size $3$ together with
-Proposition~\ref{prop:square_stability} imply that $123$ and $321$ are
-in the same Wilf-equivalence class of square permutations, and that
-$132$, $213$, $231$, and $312$ are in the same Wilf-equivalence class of
-square permutations. Computer experiments show us that there are two
-Wilf-equivalence classes of square permutations of size $3$. Indeed, the
-number of square permutations avoiding $123$ begins by
-\begin{equation} \label{equ:sequence_square_123}
-    1, 0, 2, 0, 12, 0, 118, 0, 1218, 0, 14272,
-\end{equation}
-while the number of square permutations avoiding $132$ begins by
-\begin{equation} \label{equ:sequence_square_132}
-    1, 0, 2, 0, 11, 0, 84, 0, 743, 0, 7108.
-\end{equation}
-
-Besides, an other consequence of Proposition~\ref{prop:square_stability}
-is that its makes sense to enumerate the sets of square permutations
-quotiented by the operations of mirror image, complementary, and
-inverse. The sequence enumerating these sets begins by
-\begin{equation} \label{equ:sequence_square_classes}
-    1, 0, 1, 0, 6, 0, 81, 0, 2774, 0, 162945.
-\end{equation}
-
-All Sequences~\eqref{equ:sequence_square_123}, \eqref{equ:sequence_square_132},
-and~\eqref{equ:sequence_square_classes} (and their subsequences obtained
-by removing the $0$s) are for the time being not listed in~\cite{Slo}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Algorithmic issues ----
-%%
-\section{Algorithmic issues}
-\label{section:Algorithmic issues}
-
-This section is devoted to proving hardness of recognizing square
-permutations. In the same way as happens with words, we shall use a
-linear graph framework where deciding whether a permutation is a square
-reduces to computing some specific matching in the associated linear
-graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows
-(see Fig.\ref{fig:example containment-free matching}).
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[scale=0.5,inner sep=2pt,node distance=1cm]
-    \draw node (K) {$K_\pi$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=K]   (U01) {$2$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U01] (U02) {$A$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U02] (U03) {$4$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U03] (U04) {$7$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U04] (U05) {$1$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U05] (U06) {$9$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U06] (U07) {$5$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U07] (U08) {$3$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U08] (U09) {$8$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U09] (U10) {$6$};
-    \draw [thick]
-      (U01.north) .. controls ($ (U01.north) + (0,2.25) $) and ($ (U03.north) + (0,2.25) $) .. (U03.north);
-    \draw [thick]
-      (U02.north) .. controls ($ (U02.north) + (0,2.25) $) and ($ (U06.north) + (0,2.25) $) .. (U06.north);
-    \draw [thick]
-      (U04.north) .. controls ($ (U04.north) + (0,2.5) $) and ($ (U07.north) + (0,2.5) $) .. (U07.north);
-    \draw [thick]
-      (U05.north) .. controls ($ (U05.north) + (0,2.5) $) and ($ (U08.north) + (0,2.5) $) .. (U08.north);
-    \draw [thick]
-      (U09.north) .. controls ($ (U09.north) + (0,2.5) $) and ($ (U10.north) + (0,2.5) $) .. (U10.north);
-  \end{tikzpicture}
-  \caption{\label{fig:example containment-free matching}%
-    The linear graph $K_\pi$ associated to the permutation
-    $\pi = 2A47195386$ together with
-    a containment-free perfect matching $\mathcal{M}$
-    such that for any two distinct edges $(a, b)$ and
-    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
-    if $\pi(b) < \pi(b')$.
-    From $\mathcal{M}$, it follows that $\pi$ is square as it is in
-    the shuffle of $2A718$ and $49536$, both being
-    order-isomorphic to $25314$.
-  }% end caption
-\end{figure}
-
-\begin{lemma}
-  \label{lemma:matching}
-  Let $\pi$ be a permutation. The following statements are equivalent:
-  \begin{enumerate}
-    \item The permutation $\pi$ is a square.
-    \item There exists a perfect containment-free matching $\mathcal{M}$
-    of $K_\pi$ such that for any two distinct edges $(a, b)$ and
-    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
-    if $\pi(b) < \pi(b')$.
-  \end{enumerate}
-\end{lemma}
-
-Let $\pi$ be a permutation. For the sake of clarity, we will say that a
-bunch of consecutive positions $P$ of $\pi$ is \emph{above} (resp.
-\emph{below}) above another bunch of consecutive positions $P'$ in $\pi$
-if $\pi(i) > \pi(j)$ (resp. $\pi(i) < \pi(j)$) for every $i \in P$ and
-every $j \in P'$. For example, $\sigma_1$ is above $\sigma_2$ (in an
-equivalent manner, $\sigma_2$ is below $\sigma_1$) in
-Fig.~\ref{subfig:increasing before and above decreasing}, whereas
-$\sigma_1$ is below $\sigma_2$ (in an equivalent manner, $\sigma_2$ is
-above $\sigma_1$) in Fig.~\ref{subfig:decreasing before and below
-increasing}.
-  \begin{figure}[t!]
-    \centering
-    \subfigure[%
-      An increasing pattern before and above a decreasing pattern.
-    ]{%
-      \begin{tikzpicture}
-        [
-          scale=0.325,
-          label/.style={anchor=base}
-        ]
-        % increasing pattern
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
-        \foreach \x/\y in {0/6,1/7,2/8,3/9,4/10,5/11} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % decreasing pattern
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
-        \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % labels
-        \node [label] (a) at (1,5) {$a$};
-        \node [label] (b) at (4,5) {$b$};
-        \node [label] (ap) at (7,-1) {$a'$};
-        \node [label] (bp) at (10,-1) {$b'$};
-        \node (nu1) at (2.5,4) {$\sigma_1$};
-        \node (nu2) at (8.5,-2) {$\sigma_2$};
-        % crossings edges
-        \draw [line width=1pt]
-        (1,7) .. controls +(0,12) and +(0,10) .. (7,4);
-        \draw [line width=1pt]
-        (4,10) .. controls +(0,12) and +(0,10) .. (10,1);
-      \end{tikzpicture}
-      \label{subfig:increasing before and above decreasing}
-    }% end subfigure
-    \qquad\qquad\qquad
-    \subfigure[%
-      A decreasing pattern before and below an increasing pattern.
-    ]{%
-    \begin{tikzpicture}
-      [
-        scale=0.325,
-        label/.style={anchor=base}
-      ]
-      % decreasing pattern
-      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (5.2,5.2);
-      \foreach \x/\y in {0/5,1/4,2/3,3/2,4/1,5/0} {
-          \draw [fill=black] (\x,\y) circle (0.1);
-      }
-      % decreasing pattern
-      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,5.8) grid (11.2,11.2);
-      \foreach \x/\y in {6/6,7/7,8/8,9/9,10/10,11/11} {
-          \draw [fill=black] (\x,\y) circle (0.1);
-      }
-      % labels
-      \node [label] (a) at (1,-1) {$a$};
-      \node [label] (b) at (4,-1) {$b$};
-      \node [label] (ap) at (7,5) {$a'$};
-      \node [label] (bp) at (10,5) {$b'$};
-      \node (nu1) at (2.5,-2) {$\sigma_1$};
-      \node (nu2) at (8.5,4) {$\sigma_2$};
-      % crossings edges
-      \draw [line width=1pt]
-      (1,4) .. controls +(0,10) and +(0,12) .. (7,7);
-      \draw [line width=1pt]
-      (4,1) .. controls +(0,10) and +(0,12) .. (10,10);
-      \end{tikzpicture}
-      \label{subfig:decreasing before and below increasing}
-    }% end subfigure
-    \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
-      Illustration of Lemma~\ref{lemma:at most one edge monotone}.
-    }
-  \end{figure}
-
-Before proving hardness, we give an easy lemma that will prove extremely
-useful for simplifying the proof of upcoming
-Proposition~\ref{proposition:hardness}.
-
-\begin{lemma}
-  \label{lemma:at most one edge monotone}
-  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
-  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
-  and $\mathcal{M}$ be a containment-free perfect matching of $G_\pi$.
-  The following assertions hold:
-  \begin{enumerate}
-    \item
-    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
-    $\sigma_1$ is above $\sigma_2$ (see Fig.~\ref{subfig:increasing before
-    and above decreasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
-    \item
-    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
-    is below $\sigma_2$ (see Fig.~\ref{subfig:decreasing before and below
-    increasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
-  \end{enumerate}
-\end{lemma}
-
-\begin{proposition}
-  \label{proposition:hardness}
-  Deciding whether a permutation is a square is \NPC.
-\end{proposition}
-
-\begin{proof}[of Proposition~\ref{proposition:hardness}]
-  The problem is certainly in $\NPclass$. We propose a reduction from
-  the pattern involvement problem which is known to be
-  \NPC~\cite{Bose:Buss:Lubiw:1998}: Given two permutations $\pi$ and
-  $\sigma$, decide whether $\sigma$ occurs in $\pi$ (as an
-  order-isomorphic pattern).
-
-  \begin{figure}[t!]
-    \centering
-    \begin{tikzpicture}[
-      scale=.9,
-      >=stealth',
-      shorten >=1pt,
-      main node/.style={align=center},
-      cell/.style={draw,ultra thick,fill=black!5},
-      structure link/.style={line width=1.5pt},
-      pattern link/.style={line width=1.5pt,black!20},
-      monotone/.style={->,thick}
-      ]
-      % edges
-      % pi 1 - pi 2
-      \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
-      % pi 2 - sigma 2
-      \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
-      % 1
-      \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
-      \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
-      \node (claim edge 1) at (1.5,13) {Claim~\ref{claim:all (nu_1, nu'_1)-edges}};
-      % 2
-      \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
-      \node (claim edge 2) at (4,4) {Claim~\ref{claim:(nu_2, nu'_2)-edge}};
-      \node (claim edge 2) at (4,3.5) {Claim~\ref{claim:all (nu_2, nu'_2)-edges}};
-      % 3
-      \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
-      \node (claim edge 1) at (5.5,11.5) {Claim~\ref{claim:(nu_3, nu'_3)-edge}};
-      % 4
-      \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
-      \node (claim edge 1) at (8.5,8.5) {Claim~\ref{claim:(nu_4, nu'_4)-edge}};
-      % sigma 1 - pi 1
-      \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
-      % \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
-      % nodes
-      % 1
-      \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
-      \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_1$};
-      %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
-      \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
-      \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_1$};
-      % 2
-      \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
-      \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_2$};
-      \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
-      \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_2$};
-      % 3
-      \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
-      \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_3$};
-      \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
-      \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_3$};
-      % 4
-      \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
-      \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_4$};
-      \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
-      \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_4$};
-      % sigma 1
-      \draw [cell] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
-      \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
-      % pi 1
-      \draw [cell] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
-      \node [main node] (pi1) at (8.5,5.5) {$\pi'$};
-      % pi 2
-      \draw [cell] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
-      \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
-      % sigma 2
-      \draw [cell] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
-      \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};
-      % grid
-      \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
-      % row size
-      \foreach \y/\N in {0.5/N_2,1.5/N_2,2.5/N_4,3.5/k,4.5/n,5.5/n+2,6.5/k+2,7.5/N_4,8.5/N_3,9.5/N_3,10.5/N_1,11.4/N_1} {
-          \node at (-0.5,\y) (R\y) {$\N$};
-      }
-      % column size
-      \foreach \x/\N in
-      {0.5/N_1,1.5/N_2,2.5/N_1,3.5/N_3,4.5/k+2,5.5/N_4,6.5/N_2,7.5/N_3,8.5/n+2,9.5/N_4,10.5/n,11.5/k} {
-          \node at (\x,-0.5) (C\x) {$\N$};
-      }
-    \end{tikzpicture}
-    \caption{\label{fig:reduction}%
-      Schematic representation of the permutation $\mu$ used
-      in Proposition~\ref{proposition:hardness}.
-      Black edges denote the presence of at least one edge between two bunches of
-      positions in $\mu$.
-      Grey edges denote edges that are  only considered in the forward direction of the
-      proof.
-    }%
-  \end{figure}
-
-  Let $\pi \in S_n$ and $\sigma \in S_k$ be two arbitrary permutations.
-  Define
-  \begin{align*}
-  N_4 &= 2(2n + 2k + 4) + 1  = 4n + 4k + 9 \\
-  N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45 \\
-  N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225 \\
-  N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325\text{.}
-  \end{align*}
-  Notice that $N_1, N_2, N_3$ and $N_4$ are polynomial in $n$.
-  The crucial properties are that
-  (i) $N_1, N_2, N_3$ and $N_4$ are odd integers
-  and
-  (ii) $N_i > \left(\sum_{i < j \leq 4} 2N_j\right) + 2n + 2k + k$
-  for every $1 \leq i \leq 4$.
-
-  We now turn to defining various gadgets (sequences of integers)
-  that act as building blocks in our construction of a new permutation $\mu$:
-  \begin{align*}
-  \sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
-  \pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2] \\
-  \sigma'' &= \sigma \; [2N_2 + N_4] \\
-  \pi''    &= \pi \; [2N_2 + N_4 + k] \\
-  \nu_1    &= \nearrow_{N_1} \; [2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
-  \nu'_1   &= \nearrow_{N_1} \; [N_1 + 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
-  \nu_2    &= \nearrow_{N_2} \; [N_2] \\
-  \nu'_2   &= \searrow_{N_2} \\
-  \nu_3    &= \nearrow_{N_3} \; [2N_2 + 2N_4 + 2n + 2k + 4] \\
-  \nu'_3   &= \nearrow_{N_3} \; [2N_2 + N_3 + 2N_4 + 2n + 2k + 4] \\
-  \nu_4    &= \searrow_{N_4} \; [2N_2 + N_4 + 2n + 2k + 4] \\
-  \nu'_4   &= \searrow_{N_4} \; [2N_2]\text{.}
-  \end{align*}
-  We are now in position to define our target permutation $\mu$
-  (see Fig.~\ref{fig:reduction} for an illustration):
-  $$
-  \mu
-  =
-  \nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
-  \text{.}
-  $$
-
-  It is immediate that H can be constructed in polynomial-time in $n$ and $k$.
-  We claim that $\sigma$ occurs in $\pi$ if and only if
-  there exists a containment-free perfect matching
-  $\mathcal{M}$ in $G_\mu$ such that
-  for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
-
-  Suppose first that $\sigma$ occurs in $\pi$ and fix an occurrence.
-  Construct a matching $\mathcal{M}$ as follows:
-  \begin{itemize}
-    \item $\mathcal{M}$ contains $N_1$ pairwise crossing
-    $(\nu_1, \nu'_1)$-edges.
-    \item $\mathcal{M}$ contains $N_2$ pairwise crossing
-    $(\nu_2, \nu'_2)$-edges.
-    \item $\mathcal{M}$ contains $N_3$ pairwise crossing
-    $(\nu_3, \nu'_3)$-edges.
-    \item $\mathcal{M}$ contains $N_4$ pairwise crossing
-    $(\nu_4, \nu'_4)$-edges.
-    \item $\mathcal{M}$ contains $k+2$ pairwise crossing
-    $(\sigma', \pi')$-edges as depicted in
-    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
-    More precisely,
-    (i) the first position of $\sigma'$
-    (\emph{i.e.}, $(2N_1+N_2+N_3) + 1$) is linked
-    to the first position of $\pi'$
-    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + 1$),
-    (ii) the last position of $\sigma'$
-    (\emph{i.e.}, $(2N_1+N_2+N_3) + k+2$) is linked
-    to the last position of $\pi'$
-    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + n+2$),
-    and all other positions in $\sigma'$ are linked by means of $k$ pairwise
-    crossing edges to the positions in
-    $\pi'$ that correspond to the fixed occurence of $\sigma$ in $\pi$.
-    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
-    \item $\mathcal{M}$ contains $n-k$ pairwise crossing
-    $(\pi', \pi'')$-edges as depicted in
-    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
-    More precisely,
-    all positions in $\pi'$ that do not correspond to the fixed occurence of
-    $\sigma$ in $\pi$ are linked by means of $n-k$ pairwise crossing edges
-    to the positions in $\pi''$ that do not correspond to the fixed
-    occurrence of $\sigma$ in $\pi$.
-    \item $\mathcal{M}$ contains $k$ pairwise crossing
-    $(\pi'', \sigma'')$-edges as depicted in
-    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
-    More precisely, the positions in $\pi''$ that correspond to
-    the fixed occurrence of $\sigma$ in $\pi$ are llinked
-    linked by means of $k$ pairwise crossing edges to all positions in
-    $\sigma''$.
-    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
-  \end{itemize}
-  It can be easily checked (probably referring to Figure~\ref{fig:reduction}) that
-  $\mathcal{M}$ is perfect, containment-free and that
-  for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
-
-    \begin{figure}[t!]
-      \centering
-        \begin{tikzpicture}
-          [
-            scale=0.35,
-            label/.style={anchor=base},
-            cell/.style={draw,thick,fill=black!5}
-          ]
-          % sigma'
-          \draw [cell,rounded corners] (1.5,16.5) -- (4.5,16.5) -- (4.5,19.5) -- (1.5,19.5) -- cycle;
-          \draw[step=1cm,black!60,ultra thin,fill=black!10] (0.8,16.8) grid (5.2,21.2);
-          \foreach \x/\y in {1/20,2/19,3/17,4/18,5/21} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          \node (pip) at (3,22) {$\sigma'$};
-          \node[%
-            draw,
-            text width=5.5cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,-1cm)}
-          ] at (3,26)
-          {%
-            $((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$
-          };%
-          %\node at (3,-0.5) {$((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$};
-          \node at (-0.5,20) {$k+1$};
-          \node at (-0.5,21) {$k+2$};
-          % pi'
-          \draw [cell,rounded corners] (6.5,8.5) -- (12.5,8.5) -- (12.5,14.5) -- (6.5,14.5) -- cycle;
-          \draw[step=1cm,black!60,ultra thin,fill=black!10] (5.8,8.8) grid (13.2,16.2);
-          \foreach \x/\y in {6/15,7/12,8/13,9/10,10/9,11/11,12/14,13/16} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          \node (pip) at (9.5,17) {$\pi'$};
-          \node[%
-            draw,
-            text width=5.5cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,3.5cm)}
-          ] at (9.5,-3.5)
-          {%
-            $((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$
-          };%
-          %\node at (9.5,-2) {$((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$};
-          \node at (-0.5,15) {$n+1$};
-          \node at (-0.5,16) {$n+2$};
-          % pi''
-          \draw [cell,rounded corners] (13.5,2.5) -- (19.5,2.5) -- (19.5,8.5) -- (13.5,8.5) -- cycle;
-          \draw[step=1cm,black!60,ultra thin,fill=black!10] (13.8,2.8) grid (19.2,8.2);
-          \foreach \x/\y in {14/6,15/7,16/4,17/3,18/5,19/8} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          \node (pipp) at (16.75,9) {$\pi''$};
-          \node[%
-            draw,
-            text width=2.25cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,-3cm)}
-          ] at (16.5,19)
-          {%
-            $\pi [2N_2 + N_4 + k]$
-          };%
-          %\node at (16.5,-3.5) {$\pi [2N_2 + N_4 + k]$};
-          % sigma''
-          \draw [cell,rounded corners] (19.5,-0.5) -- (22.5,-.5) -- (22.5,2.5) -- (19.5,2.5) -- cycle;
-          \draw[step=1cm,black!60,ultra thin,fill=black!10] (19.8,-0.2) grid (22.2,2.2);
-          \foreach \x/\y in {20/2,21/0,22/1} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          \node (sigmapp) at (21,3) {$\sigma''$};
-          \node[%
-            draw,
-            text width=1.75cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,0.75cm)}
-          ] at (21,-4.5)
-          {%
-            $\sigma [2N_2 + N_4]$
-          };%
-          %\node at (21,-5) {$\sigma [2N_2 + N_4]$};
-          % edges sigma' - pi'
-          \draw [line width=1pt]
-          (1,20) .. controls +(0,9) and +(0,4) .. (6,15);
-          \draw [line width=1pt]
-          (5,21) .. controls +(0,9) and +(0,4) .. (13,16);
-          \draw [line width=1pt]
-          (2,19) .. controls +(0,-17) and +(0,-7) .. (8,13);
-          \draw [line width=1pt]
-          (3,17) .. controls +(0,-17) and +(0,-7) .. (9,10);
-          \draw [line width=1pt]
-          (4,18) .. controls +(0,-17) and +(0,-7) .. (11,11);
-          % edges pi' - pi''
-          \draw [line width=1pt]
-          (7,12) .. controls +(0,9) and +(0,4) .. (14,6);
-          \draw [line width=1pt]
-          (10,9) .. controls +(0,9) and +(0,4) .. (17,3);
-          \draw [line width=1pt]
-          (12,14) .. controls +(0,9) and +(0,4) .. (19,8);
-          % edges pi'' - sigma''
-          \draw [line width=1pt]
-          (15,7) .. controls +(0,-9) and +(0,-4) .. (20,2);
-          \draw [line width=1pt]
-          (16,4) .. controls +(0,-9) and +(0,-4) .. (21,0);
-          \draw [line width=1pt]
-          (18,5) .. controls +(0,-9) and +(0,-4) .. (22,1);
-          % text
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(0.5cm,-.75cm)}
-          % ] at (1,27)
-          % {%
-          % $2$ pairwise crossing $(\sigma',\pi')$-edges that
-          % link the guards
-          % };%
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(-0.1cm,-2.5cm)}
-          % ] at (17,24)
-          % {%
-          % $n-k$ pairwise crossing $(\sigma',\pi')$-edges that
-          % denote positions that do not correspond to
-          % a specific occurence of $\sigma$ in $\pi$
-          % (at positions $2$, $3$ and $5$).
-          % };%
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(0.75cm,1.5cm)}
-          % ] at (2.5,4,5)
-          % {%
-          % $k$ pairwise crossing $(\sigma',\pi')$-edges that
-          % denote a specific occurence of $\sigma$ in $\pi$
-          % (at positions $2$, $3$ and $5$).
-          % };%
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(2cm,.75cm)}
-          % ] at (7,-3)
-          % {%
-          % $k$ pairwise crossing $(\pi''; \sigma'')$-edges that
-          % denote a specific occurence of $\sigma$ in $\pi$
-          % (at positions $2$, $3$ and $5$).
-          % };%
-        \end{tikzpicture}
-        \caption{\label{fig:subfig:sigma' - pi' - pi'' - sigma''}%
-        Illustration of the containment-free perfect matching $\mathcal{M}$
-        between gadgets $\sigma'$, $\pi'$, $\pi''$ and $\sigma''$ assuming two
-        input permutation $\sigma = 312$ and $\pi = 4\mathbf{5}\mathbf{2}1\mathbf{3}6$
-        (where a specific occurrence of $\sigma$ in $\pi$ is depicted in bold).
-        }
-      \end{figure}
-
-  Conversely, suppose that there exists a containment-free perfect matching
-  $\mathcal{M}$ in $G_\mu$ such that for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
-  We show that $\sigma$ occurs as a pattern in $\pi$.
-  Whereas the matching $\mathcal{M}$ may not be as regular as in the forward
-  direction, the main idea is to prove that $\mathcal{M}$ contains enough structure
-  (more precisely, $k+2$ $(\sigma', \pi')$-edges) so that we can conclude
-  that $\sigma$ occurs in $\pi$.
-  We have divided the reverse direction into a serie of basic claims that
-  progressively defines the overall structure of $\mathcal{M}$.
-
-    \begin{figure}[t!]
-      \centering
-      \subfigure[%
-        A $(\nu_1, \nu_2)$-edge $(a, b)$ together with a
-        $(\nu_2, \nu'_1)$-edge $(x, y)$.
-        Notice that we may have $x<b$
-        (as long as $N_1 < x \leq N_1+N_2$ and $N_1 < b \leq N_1+N_2$
-        do hold).
-      ]{%
-        \begin{tikzpicture}
-          [
-            scale=0.25,
-            label/.style={anchor=base}
-          ]
-          % nu_1
-          \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
-          \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          % nu_4
-          \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
-          \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          % nu'_1
-          \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
-          \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          % labels
-          \node [label] (y) at (17,14.5) {$y$};
-          \node [label] (yp) at (10,-1.5) {$y'$};
-          \node [label] (a) at (0,8.5) {$a$};
-          \node [label] (ap) at (8,-1.5) {$a'$};
-          \node (nu1) at (2.5,7.7) {$\nu_1$};
-          \node (nu2) at (8.5,-2.5) {$\nu_2$};
-          \node (nu1') at (14.5,13.5) {$\nu'_1$};
-          % crossings edges
-          \draw [line width=1pt]
-          (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
-          \draw [line width=1pt]
-          (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
-          \draw [line width=1pt]
-          (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
-          \draw [line width=1pt]
-          (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
-          \draw [line width=1pt]
-          (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
-          % data edges
-          \draw [line width=1pt]
-          (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
-          \draw [line width=1pt]
-          (10,1) .. controls +(0,32) and +(0,10) .. (17,21);
-        \end{tikzpicture}
-        \label{subfig:no (nu_1, nu_2)-edge - 1}
-      }% end subfigure
-      \qquad
-      \subfigure[%
-        A $(\nu_1, \nu'_1)$-edge $(x', y')$ together with an
-        edge $(y, z)$ which is neither
-        a $(\nu_1, \nu'_1)$-edge,
-        nor a $(\nu_2, \nu'_1)$-edge,
-        nor a $(\nu'_1, \nu'_1)$-edge.
-        Notice that $y$ may be any position satisfying
-        $N_1+N_2 < y \leq 2N_1+N_2$.
-      ]{%
-      \begin{tikzpicture}
-        [
-          scale=0.25,
-          label/.style={anchor=base}
-        ]
-        % nu_1
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
-        \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu_4
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
-        \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu'_1
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
-        \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % labels
-        \node [label] (y) at (17,14.5) {$y$};
-        \node [label] (ypp) at (23,-4.5) {$y''$};
-        \node [label] (a) at (0,8.5) {$a$};
-        \node [label] (a) at (2,8.5) {$x'$};
-        \node [label] (a) at (14,14.5) {$y'$};
-        \node [label] (ap) at (8,-1.5) {$a'$};
-        \node (nu1) at (2.5,7.7) {$\nu_1$};
-        \node (nu2) at (8.5,-2.5) {$\nu_2$};
-        \node (nu1') at (14.5,13.5) {$\nu'_1$};
-        % crossings edges
-        \draw [line width=1pt]
-        (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
-        \draw [line width=1pt]
-        (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
-        \draw [line width=1pt]
-        (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
-        \draw [line width=1pt]
-        (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
-        \draw [line width=1pt]
-        (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
-        % data edges
-        \draw [line width=1pt]
-        (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
-        \draw [line width=1pt]
-        (17,21) .. controls +(0,10) and +(0,35) .. (23,-3);
-        \end{tikzpicture}
-        \label{subfig:no (nu_1, nu_2)-edge - 2}
-      }% end subfigure
-      \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
-        Claim~\ref{claim:no (nu_1, nu_2)-edge}:
-        Assuming a $(\nu_1,\nu_2)$-edge $(a, b)$ in $\mathcal{M}$.
-      }
-    \end{figure}
-
-  \begin{claim}
-    \label{claim:no (nu_1, nu_2)-edge}
-    There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_2)-edge}]
-    First, according to Lemma~\ref{lemma:at most one edge monotone},
-    there exists at most one $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
-    Suppose now, aiming at a contradiction, that there exists
-    exactly one $(\nu_1, \nu_2)$-edge $e = (a, a')$ in $\mathcal{M}$.
-    Under this assumption, we show that $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
-    Indeed, suppose, aiming at a contradiction,
-    that there exists a $(\nu_1, \nu_1)$-edge $(x, x')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free,
-    we must have $x < a$.
-    Then it follows that $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing) and
-    $\mu(x') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    Hence $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
-    Furthermore, since $\nu_1$ is increasing and
-    there exists only one $(\nu_1, \nu_2)$-edge $(a, a')$ in $\mathcal{M}$,
-    we conclude that
-    $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
-    But $|\nu_1| = |\nu'_1| = N_1$ and $\mathcal{M}$ is perfect,
-    and hence there exists a position, say $y$, in $\nu'_1$ that is
-    not involved in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-    We rule out this configuration by considering two cases:
-    \begin{itemize}
-      \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-edge,
-      say $(y', y)$. Then, we have $\mu(a) > \mu(y')$ (since $\nu_1$ is above $\nu_2$)
-      and $\mu(a') < \mu(y)$ (since $\nu_2$ is below $\nu'_1$),
-      which contradicts our hypothesis about $\mathcal{M}$.
-      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
-      \item Position $y$ is not involved in a $(\nu_2, \nu'_1)$-edge.
-      Since $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges
-      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge $(y, y'')$ in $\mathcal{M}$
-      that is neither a $(\nu_1, \nu'_1)$-edge nor a $(\nu_2, \nu'_1)$-edge
-      nor a $(\nu'_1, \nu'_1)$-edge.
-      In other words, $y'' > 2N_1 + N_2$.
-      Now, for any $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$, we
-      have $\mu(x) < \mu(y)$ (since $\nu_1$ is below $\nu'_1$)
-      and $\mu(x') > \mu(y'')$ (since $\nu'_1$ is above any other gadget),
-      which contradicts our hypothesis about $\mathcal{M}$.
-      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 2}.
-    \end{itemize}
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:(nu_1, nu'_1)-edge}
-    There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:(nu_1, nu'_1)-edge}]
-    Suppose, aiming at a contradiction, that there is no
-    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-    Then it follows that there exists an edge $(x, x')$ in $\mathcal{M}$,
-    $1 \leq x \leq N_1$, that is neither a
-    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
-    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
-    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
-    (In other words, $x \leq N_1$ and $x' > 2N_1 + N_2$.)
-    Therefore, since $\mathcal{M}$ is containment-free,
-    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
-    in $\mathcal{M}$.
-    Then it follows that every position $y$ of $\nu'_1$
-    (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
-    is involved in an edge $(y, y'$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
-    (otherwise $\mathcal{M}$ would not be containment-free).
-    But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
-    > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
-    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
-    Then it follows that $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  The above claim will be complement in upcoming Claim~\ref{claim:all (nu_1, nu'_1)-edges}.
-
-  \begin{claim}
-    \label{claim:no (nu_2, nu_2)-edge}
-    There is no $(\nu_2, \nu_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_2)-edge}]
-  Combine Claim~\ref{claim:(nu_1, nu'_1)-edge} with
-  the fact that $\mathcal{M}$ is containment-free.
-  \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_2, nu'_1)-edge}
-    There is no $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_1)-edge}]
-    First, according to Lemma~\ref{lemma:at most one edge monotone},
-    there exists at most one $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
-    Suppose now, aiming at a contradiction, that there exists
-    exactly one $(\nu_2, \nu'_1)$-edge $e = (a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_1, nu'_1)-edge}, there exists at least
-    one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-    Furtheremore, since
-    there is no $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge}),
-    every position in $\nu_1$ is involved either in a
-    $(\nu_1, \nu_1)$-edge or in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$
-    (otherwise $\mathcal{M}$ would not be containment-free).
-    Then it follows that that there is an odd number of $(\nu_1, \nu'_1)$-edges
-    in $\mathcal{M}$ since $\mathcal{M}$ is perfect and $|\nu_1| = N_1$ is odd.
-    Combining this with the fact that there exists exactly one $(\nu_2, \nu'_1)$-edge
-    in $\mathcal{M}$ and $|\nu'_1| = N_1$ is odd,
-    we conclude that there exists a position, say $x$, in $\nu'_1$
-    that is neither involved in a $(\nu_1, \nu'_1)$-edge nor a $(\nu_1, \nu'_1)$-edge
-    nor a $(\nu'_1, \nu'_1)$-edge.
-    Write $(x, x')$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $x' > 2N_1 + N_2$, this edge.
-    We have $\mu(a) < \mu(x)$ (since $\nu_2$ is below $\nu'_1$)
-    and $\mu(a') > \mu(x')$ (since $\nu'_1$ is above any other gadget),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:(nu_2, nu'_2)-edge}
-    There is at least one $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:(nu_2, nu'_2)-edge}]
-    Suppose, aiming at a contradiction that there is no
-    $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
-    Notice that there is neither
-    a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
-    nor a $(\nu_2, \nu_2)$-edge (Claim~\ref{claim:no (nu_2, nu_2)-edge})
-    nor a $(\nu_2, \nu'_1)$-edge (Claim~\ref{claim:no (nu_2, nu'_1)-edge})
-    in $\mathcal{M}$.
-    But $|\nu_2| = N_2 > 2N_3 + 2N_4 + 2n + 2k +4$.
-    Hence $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no edge below (nu_2, nu'_2)-edge}
-    There is
-    neither a $(\nu'_1, \nu'_1)$-edge,
-    nor a $(\nu'_1, \nu_3)$-edge,
-    nor a $(\nu'_1, \sigma')$-edge,
-    nor a $(\nu'_1, \nu_4)$-edge
-    nor a $(\nu_3, \nu_3)$-edge,
-    nor a $(\nu_3, \sigma')$-edge,
-    nor a $(\nu_3, \nu_4)$-edge,
-    nor a $(\sigma', \sigma')$-edge,
-    nor a $(\sigma', \nu_4)$-edge,
-    nor a $(\nu_4, \nu_4)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
-    Combine Claim~\ref{claim:(nu_2, nu'_2)-edge} with the fact that
-    $\mathcal{M}$ is containment-free.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim: no edge right of nu'_2}
-    There is
-    neither a $(\nu'_2, \nu'_3)$-edge,
-    nor a $(\nu'_2, \pi')$-edge,
-    nor a $(\nu'_2, \nu'_4)$-edge,
-    nor a $(\nu'_2, \pi'')$-edge
-    nor a $(\nu'_2, \sigma'')$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim: no edge right of nu'_2}]
-    Suppose, aiming at a contradiction, that
-    $\mathcal{M}$ contains an edge $(a, a')$ which is
-    either a $(\nu'_2, \nu'_3)$-edge
-    or a $(\nu'_2, \pi')$-edge
-    or a $(\nu'_2, \nu'_4)$-edge
-    or a $(\nu'_2, \pi'')$-edge
-    or a $(\nu'_2, \sigma'')$-edge.
-    According to Claim~\ref{claim:(nu_2, nu'_2)-edge},
-    there exists a $(\nu_2, \nu'_2)$-edge $(b, b')$ in $\mathcal{M}$.
-    Then it follows that we have
-    $\mu(b) > \mu(a)$ (since $\nu_2$ is above $\nu'_2$) and
-    $\mu(b') < \mu(a')$ (since $\nu'_3$, $\pi'$, $\nu'_4$, $\pi''$ and
-    $\sigma''$ are all above $\nu'_2$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  % \begin{claim}
-  %   \label{claim:no (nu_1, nu'_2)-edge}
-  %   There is no $(\nu_1, \nu'_2)$-edge in $\mathcal{M}$.
-  % \end{claim}
-  %
-  % \begin{proof}[of Claim~\ref{claim:no (nu_1, nu'_2)-edge}]
-  %   Suppose, aiming at a contradiction that there is a
-  %   $(\nu_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-  %   Since $\mathcal{M}$ is containment-free, we must have $b < a$.
-  %   Then it follows that we have
-  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
-  %   $\mu(b') > \mu(a')$ (since $\nu'_2$ is above $\nu'_2$),
-  %   which contradicts our hypothesis about $\mathcal{M}$.
-  %   \qed
-  % \end{proof}
-  %
-  % \begin{claim}
-  %   \label{claim:no (nu'_1, nu'_2)-edge}
-  %   There is no $(\nu'_1, \nu'_2)$-edge in $\mathcal{M}$.
-  % \end{claim}
-  %
-  % \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_2)-edge}]
-  %   Suppose, aiming at a contradiction that there is a
-  %   $(\nu'_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-  %   Then it follows that we have
-  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
-  %   $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu'_2$),
-  %   which contradicts our hypothesis about $\mathcal{M}$.
-  %   \qed
-  % \end{proof}
-
-  \begin{figure}[t!]
-    \centering
-      \begin{tikzpicture}
-        [
-          scale=0.35,
-          label/.style={anchor=base}
-        ]
-        % nu_2
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
-        \foreach \x/\y in {0/11,1/10,2/9,3/8,4/7,5/6} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu_3
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,15.8) grid (11.2,21.2);
-        \foreach \x/\y in {6/16,7/17,8/18,9/19,10/20,11/21} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu'_2
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,-0.2) grid (17.2,5.2);
-
-        \foreach \x/\y in {12/5,13/4,14/3,15/2,16/1,17/0} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % labels
-        \node [label] (x) at (2,4.5) {$x$};
-        \node [label] (a) at (8,14.5) {$a$};
-        \node [label] (xp) at (14,-1.5) {$x'$};
-        \node [label] (ap) at (16,-1.5) {$a'$};
-        \node (nu1) at (2.5,2.5) {$\nu_2$};
-        \node (nu2) at (8.5,12.5) {$\nu_3$};
-        \node (nu1') at (14.5,-3.5) {$\nu'_2$};
-        % crossings edges
-        \draw [line width=1pt]
-        (2,9) .. controls +(0,25) and +(0,30) .. (14,3);
-        \draw [line width=1pt]
-        (8,18) .. controls +(0,15) and +(0,20) .. (16,1);
-      \end{tikzpicture}
-    \caption{\label{fig:subfig:no (nu_3, nu'_2)-edge}%
-      Claim~\ref{claim:no (nu_3, nu'_2)-edge}:
-      Assuming a $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
-    }
-  \end{figure}
-
-  \begin{claim}
-    \label{claim:no (nu_3, nu'_2)-edge}
-    There is no $(\nu_3, \nu'_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
-    Suppose, aiming at a contradiction, that there exists
-    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_2, nu'_2)-edge}, there
-    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free (Lemma~\ref{lemma:matching}),
-    we have $x' < a'$.
-    Therefore,
-    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
-    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:at most one (nu_2, nu_3)-edge}
-    There is at most one $(\nu_2, \nu_3)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
-    Apply Lemma~\ref{lemma:at most one edge monotone}.
-    \qed
-  \end{proof}
-
-    We will see soon (upcoming Claim~\ref{claim:all (nu_2, nu'_2)-edges})
-    that there exists actually no $(\nu_2, \nu_3)$-edge
-    in $\mathcal{M}$.
-
-  \begin{claim}
-    \label{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}
-    There is neither a $(\nu_1, \nu_3)$-edge not a $(\nu'_1, \nu_3)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
- \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}]
-    We first prove that there is no $(\nu_1, \nu_3)$-edge in $\mathcal{M}$.
-    Indeed, suppose, aiming at a contradiction, that these exists
-    a $(\nu_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free, we must have
-    $b < a'$.
-    Then it follows that
-    $\mu(b) < \mu(a)$ (since $\nu_1$ is incresing) and
-    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-
-    As for no $(\nu_1, \nu_3)$-edges in $\mathcal{M}$, suppose,
-    aiming at a contradiction, that these exists
-    a $(\nu'_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-    Then it follows that
-    $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
-    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:(nu_3, nu'_3)-edge}
-    There exists a $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:(nu_3, nu'_3)-edge}]
-    Suppose, aiming at a contradiction, that there is no
-    $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
-    Combining
-    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
-    Claim~\ref{claim:no (nu_3, nu'_2)-edge},
-    Claim~\ref{claim:at most one (nu_2, nu_3)-edge}
-    and Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge},
-    we conclude that $N_3-1$ positions in $\nu_3$ are involved
-    in edges of $\mathcal{M}$ that are
-    neither ($\nu_1, \nu_3)$-edges,
-    nor $(\nu_2, \nu_3)$-edges
-    nor $(\nu'_1, \nu_3)$-edges
-    nor $(\nu_3, \nu_3)$-edges.
-    But $N_3 - 1 > 2N_4 + 2n + 2k + 4$ (the positions left) and hence
-    $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no edge below one (nu_3, nu'_3)-edge}
-    There is
-    neither a $(\sigma', \sigma')$-edge,
-    nor a $(\sigma', \nu_4)$-edge,
-    nor a $(\sigma', \nu'_2)$-edge,
-    nor a $(\nu_4, \nu_4)$-edge
-    nor a $(\nu_4, \nu'_2)$-edge,
-    nor a $(\nu'_2, \nu'_2)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge}]
-    Combine Claim~\ref{claim:(nu_3, nu'_3)-edge} with the fact that
-    $\mathcal{M}$ is containment-free.
-    \qed
-  \end{proof}
-
-  The next two claims state that $\mathcal{M}$ actually contains
-  all $(\nu_1, \nu'_1)$-edges and all $(\nu_2, \nu'_2)$-edges.
-
-  \begin{claim}
-    \label{claim:all (nu_1, nu'_1)-edges}
-    $\mathcal{M}$ contains $N_1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:all (nu_1, nu'_1)-edges}]
-    Suppose, aiming at a contradiction, that
-    $\mathcal{M}$ does not contain $N_1$ $(\nu_1, \nu'_1)$-edge.
-    Combing Claim~\ref{claim:no (nu_2, nu'_1)-edge} and
-    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
-    we conclude that there exists a position $a$ in $\nu'_1$ that is
-    involded in an edge $(a, a')$ in $\mathcal{M}$ that is
-    neither a $(\nu_1, \nu'_1)$-edge,
-    nor a $(\nu_2, \nu'_1)$-edge,
-    nor a $(\nu'_1, \nu'_1)$-edge.
-    (In other words, $a' > 2N_1 + N_2$.)
-    Furthermore, according to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$.
-    Then, we have
-    $\mu(x) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
-    $\mu(x') < \mu(a')$ (since $\nu'_1$ is above any other gadget),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:all (nu_2, nu'_2)-edges}
-    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:all (nu_2, nu'_2)-edges}]
-    The key idea is to focus on $\nu'_2$.
-    Combining Claim~\ref{claim:no (nu_3, nu'_2)-edge},
-    Claim~\ref{claim: no edge right of nu'_2} and
-    Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge},
-    we obtain that every position in $\nu'_2$ is involved in
-    either a $(\nu_1, \nu'_2$)-edge
-    or a $(\nu'_1, \nu'_2$)-edge
-    or $(\nu_2, \nu'_2$)-edge.
-    But, according to Claim~\ref{claim:all (nu_1, nu'_1)-edges},
-    all positions in $\nu_1$ and $\nu'_1$ are involved in
-    $(\nu_1, \nu'_1)$-edges.
-    Then it follows that
-    all positions in $\nu'_2$ are involved in a $(\nu_2, \nu'_2)$,
-    and hence
-    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_2, nu_4)-edge}
-    There is no $(\nu_2, \nu_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_4)-edge}]
-    According to Claim~\ref{claim:all (nu_2, nu'_2)-edges},
-    all positions in $\nu_2$ and $\nu'_2$ are involved in
-    $(\nu_2, \nu'_2)$-edges in $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_4, nu'_3)-edge}
-    There is no $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_4, nu'_3)-edge}]
-    Suppose, aiming at a contradiction, that there exists
-    a $(\nu_4, \nu'_3)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_3, nu'_3)-edge},
-    there exists a $(\nu_3, \nu'_3)$-edge $(x, x')$
-    in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free, we must have
-    $x' < a'$.
-    Therefore,
-    we have $\mu(x) > \mu(a)$ (since $\nu_3$ is above $\nu_4$) and
-    $\mu(x') < \nu(a')$ (since $\nu'_3$ is increasing),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:(nu_4, nu'_4)-edge}
-    There is at least one $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:(nu_4, nu'_4)-edge}]
-    Suppose, aiming at a contradiction, that there is no
-    $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
-    First, according to
-    Claim~\ref{claim:all (nu_1, nu'_1)-edges} and
-    Claim~\ref{claim:all (nu_2, nu'_2)-edges},
-    there is neither a $(\nu_1, \nu_4)$-edge
-    nor a $(\nu'_1, \nu_4)$-edge
-    nor a $(\nu_2, \nu_4)$-edge
-    nor a $(\nu_4, \nu'_2)$-edge in $\mathcal{M}$.
-    Furthermore, according to
-    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge} and
-    Claim~\ref{claim:no (nu_4, nu'_3)-edge},
-    there is neither a $(\nu_3, \nu_4)$-edge
-    nor a $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
-    But $N_4 > 2n + 2k + 2$, and
-    hence $\mathcal{M}$ is not perfect,
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no edge below (nu_4, nu'_4)-edge}
-    There is
-    neither a $(\pi', \pi')$-edge
-    nor a $(\sigma', \pi'')$-edge
-    nor a $(\sigma', \sigma'')$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
-    Combine Claim~\ref{claim:(nu_4, nu'_4)-edge} with the
-    fact that $\mathcal{M}$ is containment-free.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (sigma', nu'_3)-edge, no (sigma', nu'_4)-edge}
-    There is neither a $(\sigma', \nu'_3)$-edge
-    nor a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\label{claim:no (sigma', nu'_3)-edge}]
-    According to
-    Claim~\ref{claim: no edge right of nu'_2},
-    Claim~\ref{claim:all (nu_1, nu'_1)-edges}
-    Claim~\ref{claim:all (nu_2, nu'_2)-edges} and
-    Claim~\ref{claim:no (nu_4, nu'_3)-edge}
-    every position in $\sigma'$ is involved in
-    either a $(\sigma', \nu'_3)$-edge
-    or a $(\sigma', \pi')$-edge
-    or a $(\sigma', \nu'_4)$-edge
-    in $\mathcal{M}$.
-    Let $a$ be the first position of $\sigma'$
-    (\emph{i.e.}, $a = 2N_1 + N_2 + N_3 + 1$)
-    and
-    $b$ be the last position of $\sigma'$
-    (\emph{i.e.}, $b = 2N_1 + N_2 + N_3 + k + 2$),
-    and write $(a, a')$ and $(b, b')$ for the two associated edges of $\mathcal{M}$.
-    We consider the presence of a $(\sigma', \nu'_3)$-edge
-    or a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$ separately.
-    \begin{itemize}
-      \item
-      Suppose, aiming at a contradiction, that there exists
-      a $(\sigma', \nu'_3)$-edge in $\mathcal{M}$.
-      Since $\mathcal{M}$ is containment-free, we conclude that
-      $(a, a')$ must be a $(\sigma', \nu'_3)$-edge.
-      But $\mu(a) = 2N_2 + N_4 + 2n + k + 2 + (k+1) <
-      \mu(b) = 2N_2 + N_4 + 2n + k + 2 + (k+2)$, and hence
-      $(b, b')$ is also a $(\sigma', \nu'_3)$-edge
-      (otherwise we would have $\mu(a') > \mu(b')$ since
-      $\nu'_3$ is above $\pi'$ and $\nu'_4$,
-      which contradicts our hypothesis).
-      Consider any position $c$ but $a$ and $b$ of $\sigma'$
-      (\emph{i.e.},
-      $a = 2N_1 + N_2 + N_3 + 1 < c < 2N_1 + N_2 + N_3 + k + 2 = b$),
-      and let $(c, c')$ be the associate edge in $\mathcal{M}$.
-      We first observe that $(c, c')$ is not a $(\sigma', \nu'_3)$-edge,
-      since otherwise we would have
-      $\mu(a) > \mu(c)$ and $\mu(a') < \mu(c')$
-      (since $\mathcal{M}$ is containment-free and $\nu'_3$ is increasing),
-      thereby contradicting our hypothesis about $\mathcal{M}$.
-      Therefore, $(c, c')$ is either a $(\sigma', \pi)$-edge or a $(\sigma', \nu'_4)$-edge.
-      In both cases, $(b, b')$ is contained in $(c, c')$,
-      which contradicts our hypothesis about $\mathcal{M}$ being containment-free.
-      \item
-      Suppose, aiming at a contradiction, that there exists
-      a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
-      Since $\mathcal{M}$ is containment-free, we conclude that
-      $(b, b')$ must be a $(\sigma', \nu'_4)$-edge.
-      Since there is no $(\sigma', \nu'_3)$-edge in $\mathcal{M}$
-      (preceding point),
-      $(a, a')$ is either a
-      $(\sigma', \pi')$-edge or a $(\sigma', \nu'_4)$-edge.
-      In both cases, we have $a' < a$.
-      Then it follows that
-      $\mu(a) < \mu(b)$
-      (since $2N_2 + N_4 + 2n + k + 2 + (k+1) < 2N_2 + N_4 + 2n + k + 2 + (k+2)$)
-      and
-      $\mu(a') > \mu(b')$ (since $\pi'$ is above $\nu'_4$ if $(a, a')$ is
-      a $(\sigma', \pi')$-edge and
-      $\nu'_4$ is decreasing if $(a, a')$ is
-      a $(\sigma', \nu'_4)$-edge),
-      which contradicts our hypothesis about $\mathcal{M}$.
-    \end{itemize}
-    \qed
-  \end{proof}
-
-  Combining the above claims, we conclude that there are
-  $k+2$ $(\sigma', \pi')$-edges in $\mathcal{M}$.
-  Recall that
-  $\sigma' = ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$ and that
-  $\pi' = ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$.
-  Then it follows we have
-  at least $k$ (possibly $k+1$ or $k+2$) independent $(\sigma', \pi')$-edges
-  $(a, a')$ in $\mathcal{M}$ with
-  $2N_1 + N_2 + N_3 + 1 < a < 2N_1 + N_2 + N_3 + k + 2$
-  and
-  $2N_1 + N_2 + N_3 + (k + 2) + 1 < a' < 2N_1 + N_2 + N_3 + (k + 2) + n + 2$.
-  Therefore, by our hypothesis about $\mathcal{M}$,
-  $\sigma$ occurs as a pattern in $\pi$.
-  \qed
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% %%
-% %% ---- Patterns ----
-% %%
-% \section{Patterns}
-% \label{section:Patterns}
-%
-% \begin{proposition}
-%   \label{proposition:patterns:lower bound}
-%   Let $\pi \in S_{n}$.
-%   Then $\pi$ contains a pattern of size at least
-%   $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
-%   is the union of two order-isomorphic patterns.
-% \end{proposition}
-%
-% \begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
-%   The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935}
-%   states that every permutation of
-%   length at least $pq+1$ must contain either the pattern
-%   $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
-%   The result now follows from the fact that
-%   the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
-% \qed
-% \end{proof}
-%
-% \begin{proposition}
-%   \label{proposition:patterns:upper bound}
-%   If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
-%   a permutation of $S_{n}$
-%   that does not contains a pattern
-%   of length $2k$ that is the union of two
-%   order-isomorphic patterns.
-% \end{proposition}
-%
-% \begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
-% The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
-% Let $\pi \in S_n$ be chosen randomly and uniformly.
-% For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
-% of $\pi$,
-% let $X_\sigma$ be the indicator variable that
-% $\sigma$ is the union of two order-ismorphic patterns.
-% According to Boole's inequality, we have
-% $P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
-% Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
-% the probability that $\pi$ contains a pattern of length $2k$
-% is the union of two order-ismorphic patterns is at most
-% $\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
-% Thus, with positive probability, no event $X_\sigma$ occurs and
-% there is a permutation of $S_n$ that does not contain a pattern
-% of length $2k$ that is the union of two order-isomorphic patterns.
-% \qed
-% \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Conclusion ----
-%%
-\section{Conclusion}
-\label{section:Conclusion}
-
-% Unifier (une seule partie)
-
 There are a number of further directions of investigation in this
@@
 deciding whether $\sigma$ is a square root of~$\pi$?
-%Besides, in this work, we investigate some algebraic properties of the
-%shuffle of permutations which have direct combinatorial consequences.
-%However, this work raises several additional classical algebraic
-%questions. Among these,
 Moreover, one can ask for a complete algebraic study of
@@
 free as an associative algebra are worthwhile questions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Bibliography
-%%
-
-\bibliographystyle{plain}
-\bibliography{biblio}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\section*{Appendix (Reviewers version only)}
-
-\begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
-    Let us first show that the image of $\BINTOPERM$ contains only
-    permutations avoiding $213$ and $231$. Let $u$ be a binary word,
-    $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
-    positions of the occurrences of $0$ (resp. $1$) in $u$. By
-    definition of $\BINTOPERM$, from left to right, the subword
-    $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
-    is decreasing, and all letters of $w$ are greater than those
-    of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
-    Then, since $v$ is increasing and $w$ is decreasing, there is an
-    occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
-    occurrence of $21$ (resp. $2$, $1$). All these three cases
-    contradict the fact that all letters of $w$ are greater than
-    those of $v$. A similar argument shows that $\pi$ avoids~$231$
-    as well.
-    \smallskip
-
-    Finally, observe that any permutation $\pi$ avoiding $213$ and
-    $231$ necessarily starts by the smallest possible letter or the
-    greatest possible letter. This property is then true for the
-    suffix of $\pi$ obtained by deleting its first letter,
-    and so on for all of its suffixes. Thus, by
-    replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
-    role of a smallest (resp. greatest) letter, one obtains a binary
-    word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
-    avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
-    Since $u$ is a square binary word, there is a binary word $v$
-    such that $u \in v \shuffle v$. Then, there are two disjoint
-    sets $P$ and $Q$ of positions of letters of $u$ such that
-    $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
-    words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
-    same standarized $\sigma$. Hence, and by definition of
-    the shuffle product of permutations, $\BINTOPERM(u)$ appears in
-    $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
-    square permutation.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
-    Let $\pi$ be a square permutation avoiding $213$ and $231$. By
-    Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
-    the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
-    well-defined binary word. Since $\pi$ is a square permutation,
-    there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
-    of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
-    order-isomorphic. This implies, by the definitions of $\BINTOPERM$
-    and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
-    is a square binary word.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lem:endomorphisms}]
-    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
-    associative algebras, we have to prove that for all permutations
-    $\sigma$ and $\nu$,
-    \begin{equation}
-        \phi_j(\sigma \SHUFFLE \nu) =
-        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
-    \end{equation}
-    We shall proceed here as we have done during the proof of
-    Proposition~\ref{prop:shuffle_associative_commutative} by showing
-    that $\phi_j$ is a morphism of coalgebras, that is
-    \begin{equation}
-        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
-    \end{equation}
-    In the sequel, $\pi$ is a permutation.
-    \smallskip
-
-    If $P$ is a set of indexes of letters of $\pi$, we denote by
-    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
-    the operation $\widetilde{\,}$ defines a bijection on the set of the
-    subsets of $[|\pi|]$, and since the standardization operation
-    commutes with the mirror operation on words without multiple
-    occurrence of a letter, we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_1(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_1(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|P_1}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \widetilde{\STD\left(\pi_{|P_1}\right)}
-        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
-    $\phi_1$ is a morphism of associative algebras.
-    \smallskip
-
-    Next, since by definition of the complementation operation on
-    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
-    we have $\tau(i) < \tau(k)$ if and only if $\bar \tau(i) > \bar \tau(k)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_2(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_2(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\bar \pi_{|P_1}\right)
-        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
-    $\phi_2$ is a morphism of associative algebras.
-    \smallskip
-
-    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
-    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
-    $\{\tau(i) : i \in P\}$. Since the map sending a subset $P$ of
-    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
-    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_3(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_3(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|P_1}\right)
-        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
-        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right)^{-1}
-        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
-    $\phi_3$ is a morphism of associative algebras.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lemma:matching}]
-    The proof is, naturally enough, a slight extension of
-    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
-    for the sake of completeness.
-
-    (\emph{1} $\Rightarrow$ \emph{2}).
-    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
-    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
-    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
-    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
-    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
-    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
-    with vertices
-    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
-    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
-    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
-    matching since $I^1 \cap I^2 = \emptyset$, and
-    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
-    Indeed, if it is not the case then there would exist two edges
-    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
-    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
-    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
-    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
-    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
-    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
-    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
-    elements (at position $j$ and $k$) in two patterns that are
-    order-isomorphic to $\sigma$.
-
-    (\emph{2} $\Rightarrow$ \emph{1}).
-    Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$, and
-    $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$.
-    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
-    endpoints of the edges in $\mathcal{M}$ assuming the positions
-    $1, 2, \dots, 2$ are drawn along an horizontal line.
-    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
-    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
-    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
-    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
-    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
-    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
-    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
-    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
-    contradiction, that this is not the case. Therefore, there exist two
-    positions $j$ and $k$, say $j < k$, such that
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
-    But since $\mathcal{M}$ is containment-free, it follows that both
-    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
-    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
-    since for any two distinct edges $(a, b)$ and $(a', b')$ in
-    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
-    $\pi(b) < \pi(b')$.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
-  We first prove \emph{1}. Suppose, aiming at a contradiction, that
-  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
-  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
-  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
-  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
-  $\sigma_2$ is decreasing), which contradicts our hypothesis about
-  $\mathcal{M}$.
-
-  As for \emph{2}, suppose, still aiming at a contradiction, that there
-  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
-  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
-  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
-  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
-  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
-  \qed
-\end{proof}
-\bigskip
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-09-25 submission.tex
add nex conclusion

--- 2015-09-25 submission.tex
@@
   \sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
-  \pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2] \\
+  \pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k] \\
   \sigma'' &= \sigma \; [2N_2 + N_4] \\
@@
 There are a number of further directions of investigation in this
-general subject. We mention several -~not necessarily new~- open
+general subject. They cover several areas: algorithmic, combinatorics,
+and algebra. Let us mention several -~not necessarily new~- open
 problems that are, in our opinion, the most interesting. How many
@@
 deciding whether $\sigma$ is a square root of~$\pi$?
-
-Besides, in this work, we investigate some algebraic properties of the
-shuffle of permutations which have direct combinatorial consequences.
-However, this work raises several additional classical algebraic
-questions. Among these, one can ask for a complete algebraic study of
+Moreover, one can ask for a complete algebraic study of
 $\QQ[S]$ as a graded associative algebra for the shuffle  product

### 2015-09-26 submission.tex
Typos

@@
 Besides purely combinatorial questions,
-the shuffle product of words leads to the following computational ones:
+the shuffle product of words leads to the following algorithmic ones:
 \begin{enumerate}
@@
 shuffle product of permutations $\SHUFFLE$ both from a combinatorial and
-from a computational point of view by focusing on {\em square}
+from an algorithmic point of view by focusing on {\em square}
 permutations, that are permutations $\pi$ being in the shuffle of a
@@
 
-Let us now provide some definitions about graphs and matching that are used in
+Let us now provide some definitions about graphs and matchings that are used in
 the sequel. For a graph $G$, we denote $\VS{G}$ as the set of vertices
@@
     \multirow{2}{*}{\, $12$, $21$ \,} &
-    $1234$, $1243$, $1423$, $1324$, $1342$, $4132$, $3124$, $3142$,
+    \, $1234$, $1243$, $1423$, $1324$, $1342$, $4132$, $3124$, $3142$,
     $3412$, $4312$, \\
-    & & $2134$, $2143$, $2413$, $4213$, $2314$, $2431$, $4231$, $3241$,
+    & & \, $2134$, $2143$, $2413$, $4213$, $2314$, $2431$, $4231$, $3241$,
     $3421$, $4321$
@@
 and~\eqref{equ:sequence_square_classes} (and their subsequences obtained
-by removing the $0$s) are for the time being not listed in~\cite{Slo}.
+by removing the $0$'s) are for the time being not listed in~\cite{Slo}.
 
@@
   be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
-  and $\mathcal{M}$ be a containment-free perfect matching of $G_\pi$.
+  and $\mathcal{M}$ be a containment-free perfect matching of $K_\pi$.
   The following assertions hold:
@@
   It can be shown (see Appendix) that
-  that $\sigma$ occurs in $\pi$ if and only if
+  $\sigma$ occurs in $\pi$ if and only if
   there exists a containment-free perfect matching
-  $\mathcal{M}$ in $G_\mu$ such that
+  $\mathcal{M}$ in $K_\mu$ such that
   for any two disctinct edges

### 2015-09-26 submission.tex
Typos

@@
   % there exists a containment-free perfect matching
-  % $\mathcal{M}$ in $G_\mu$ such that
+  % $\mathcal{M}$ in $K_\mu$ such that
   % for any two disctinct edges
@@
     occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
-    occurrence of $21$ (resp. $2$, $1$). All these three cases
+    occurrence of $21$ (resp. $2$, $1$) in $w$. All these three cases
     contradict the fact that all letters of $w$ are greater than
@@
     order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
-    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
+    $|\sigma| = n$). Let $K_\pi$ be the linear clique associated to $\pi$
     with vertices
-    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
+    $\VS{K_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
     is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
-    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
+    is a subset of $\ES{K_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
     matching since $I^1 \cap I^2 = \emptyset$, and
@@
   Conversely, suppose that there exists a containment-free perfect matching
-  $\mathcal{M}$ in $G_\mu$ such that for any two disctinct edges
+  $\mathcal{M}$ in $K_\mu$ such that for any two disctinct edges
   $(a, b)$ and $(a', b')$ in $\mathcal{M}$

### 2015-09-27 submission.tex
G_mu to K_mu

--- 2015-09-25 submission.tex
@@
   =
-  \nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
+  \nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_2 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
   \text{.}
@@
   We have divided the reverse direction into a serie of basic claims that
-  progressively defines the overall structure of $\mathcal{M}$.
+  progressively defines the overall structure of $\mathcal{M}$
+  (refer to Fig.~\ref{fig:reduction}).
 

### 2015-09-27 submission.tex
typos

@@
       Black edges denote the presence of at least one edge between two bunches of
-      positions in $\mu$.
+      vertices in $K_\mu$.
       Grey edges denote edges that are only considered in the forward direction of the
@@
   there exists a containment-free perfect matching
-  $\mathcal{M}$ in $G_\mu$ such that
+  $\mathcal{M}$ in $K_\mu$ such that
   for any two disctinct edges

### 2015-09-27 submission.tex
Typos

@@
 \end{proof}
-
 \bigskip
@@
     $(\sigma', \pi')$-edges as depicted in
-    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
+    Fig.~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
     More precisely,
@@
     $(\pi', \pi'')$-edges as depicted in
-    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
+    Fig.~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
     More precisely,
@@
     $(\pi'', \sigma'')$-edges as depicted in
-    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
+    Fig.~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
     More precisely, the positions in $\pi''$ that correspond to
@@
     $\sigma''$.
-    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
+    (Notice that we use here the fact that $\sigma$ occurs in $\pi$.)
   \end{itemize}
-  It can be easily checked (probably referring to Figure~\ref{fig:reduction}) that
-  $\mathcal{M}$ is perfect, containment-free and that
+  It can be easily checked (probably referring to Fig.~\ref{fig:reduction}) that
+  $\mathcal{M}$ is perfect, containment-free, and that
   for any two disctinct edges
@@
         between gadgets $\sigma'$, $\pi'$, $\pi''$ and $\sigma''$ assuming two
-        input permutation $\sigma = 312$ and $\pi = 4\mathbf{5}\mathbf{2}1\mathbf{3}6$
+        input permutations $\sigma = 312$ and $\pi = 4\mathbf{5}\mathbf{2}1\mathbf{3}6$
         (where a specific occurrence of $\sigma$ in $\pi$ is depicted in bold).
@@
   \end{proof}
+  \bigskip
 
@@
     (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
-    is involved in an edge $(y, y'$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
+    is involved in an edge $(y, y')$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
     (otherwise $\mathcal{M}$ would not be containment-free).
@@
   \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
-    Suppose, aiming at a contradiction, that there exists
+    Suppose, aiming at a contradiction, that there exists a
     $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
@@
     $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
-    which contradicts our hypothesis about $\mathcal{M}$.
+    which contradicts our hypothesis about~$\mathcal{M}$.
 
     As for no $(\nu_1, \nu_3)$-edges in $\mathcal{M}$, suppose,
-    aiming at a contradiction, that these exists
+    aiming at a contradiction, that there exists
     a $(\nu'_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
@@
     $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
+    which contradicts our hypothesis about~$\mathcal{M}$.
+    \qed
+  \end{proof}
+  \bigskip
 
@@
     Suppose, aiming at a contradiction, that
-    $\mathcal{M}$ does not contain $N_1$ $(\nu_1, \nu'_1)$-edge.
+    $\mathcal{M}$ does not contain $N_1$ $(\nu_1, \nu'_1)$-edges.
     Combing Claim~\ref{claim:no (nu_2, nu'_1)-edge} and

### 2015-09-27 submission.tex
NP-hard proof
Fig 5

--- 2015-09-27 submission.tex
@@
 
-  It is immediate that H can be constructed in polynomial-time in $n$ and $k$.
+  It is immediate that $\mu$ can be constructed in polynomial-time in $n$ and $k$.
   % We claim that $\sigma$ occurs in $\pi$ if and only if
@@
           % labels
-          \node [label] (y) at (17,14.5) {$y$};
-          \node [label] (yp) at (10,-1.5) {$y'$};
+          \node [label] (y) at (17,14.5) {$b$};
+          \node [label] (yp) at (10,-1.5) {$b'$};
           \node [label] (a) at (0,8.5) {$a$};
@@
         % labels
-        \node [label] (y) at (17,14.5) {$y$};
-        \node [label] (ypp) at (23,-4.5) {$y''$};
+        \node [label] (y) at (17,14.5) {$b$};
+        \node [label] (ypp) at (23,-4.5) {$b''$};
         \node [label] (a) at (0,8.5) {$a$};
-        \node [label] (a) at (2,8.5) {$x'$};
-        \node [label] (a) at (14,14.5) {$y'$};
+        \node [label] (a) at (2,8.5) {$x$};
+        \node [label] (a) at (14,14.5) {$x'$};
         \node [label] (ap) at (8,-1.5) {$a'$};
@@
     Indeed, suppose, aiming at a contradiction,
-    that there exists a $(\nu_1, \nu_1)$-edge $(x, x')$ in $\mathcal{M}$.
+    that there exists a $(\nu_1, \nu_1)$-edge $(b, b')$ in $\mathcal{M}$.
     Since $\mathcal{M}$ is containment-free,
-    we must have $x < a$.
-    Then it follows that $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing) and
-    $\mu(x') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
+    we must have $b < a$.
+    Then it follows that $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
+    $\mu(b') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
     which contradicts our hypothesis about $\mathcal{M}$.
@@
     But $|\nu_1| = |\nu'_1| = N_1$ and $\mathcal{M}$ is perfect,
-    and hence there exists a position, say $y$, in $\nu'_1$ that is
+    and hence there exists a position, say $b$, in $\nu'_1$ that is
     not involved in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
@@
       \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-edge,
-      say $(y', y)$. Then, we have $\mu(a) > \mu(y')$ (since $\nu_1$ is above $\nu_2$)
-      and $\mu(a') < \mu(y)$ (since $\nu_2$ is below $\nu'_1$),
+      say $(b', b)$.
+      Then, we have $\mu(a) > \mu(b')$ (since $\nu_1$ is above $\nu_2$)
+      and $\mu(a') < \mu(b)$ (since $\nu_2$ is below $\nu'_1$),
       which contradicts our hypothesis about $\mathcal{M}$.
       See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
-      \item Position $y$ is not involved in a $(\nu_2, \nu'_1)$-edge.
+      \item Position $b$ is not involved in a $(\nu_2, \nu'_1)$-edge.
       Since $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges
-      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge $(y, y'')$ in $\mathcal{M}$
+      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge
+      $(b, b'')$ in $\mathcal{M}$
       that is neither a $(\nu_1, \nu'_1)$-edge nor a $(\nu_2, \nu'_1)$-edge
@@
       Now, for any $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$, we
-      have $\mu(x) < \mu(y)$ (since $\nu_1$ is below $\nu'_1$)
-      and $\mu(x') > \mu(y'')$ (since $\nu'_1$ is above any other gadget),
+      have $\mu(x) < \mu(b)$ (since $\nu_1$ is below $\nu'_1$)
+      and $\mu(x') > \mu(b'')$ (since $\nu'_1$ is above any other gadget),
       which contradicts our hypothesis about $\mathcal{M}$.

### 2015-09-27 submission.tex
Fig 6

@@
         % labels
-        \node [label] (x) at (2,4.5) {$x$};
+        \node [label] (x) at (2,4.5) {$b$};
         \node [label] (a) at (8,14.5) {$a$};
-        \node [label] (xp) at (14,-1.5) {$x'$};
+        \node [label] (xp) at (14,-1.5) {$b'$};
         \node [label] (ap) at (16,-1.5) {$a'$};
-        \node (nu1) at (2.5,2.5) {$\nu_2$};
-        \node (nu2) at (8.5,12.5) {$\nu_3$};
-        \node (nu1') at (14.5,-3.5) {$\nu'_2$};
+        \node (nu1) at (2.5,3.5) {$\nu_2$};
+        \node (nu2) at (8.5,13.5) {$\nu_3$};
+        \node (nu1') at (14.5,-2.5) {$\nu'_2$};
         % crossings edges
@@
     According to Claim~\ref{claim:(nu_2, nu'_2)-edge}, there
-    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
+    exists at least one $(\nu_2, \nu'_2)$-edge $(b, b')$ in $\mathcal{M}$.
     Since $\mathcal{M}$ is containment-free (Lemma~\ref{lemma:matching}),
-    we have $x' < a'$.
+    we have $b' < a'$.
     Therefore,
-    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
-    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
+    we have $\mu(b) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
+    $\mu(b') > \mu(a')$ (since $\nu'_2$ is decreasing),
     which contradicts our hypothesis about $\mathcal{M}$.

### 2015-09-27 submission.tex
introduction

@@
 Besides purely combinatorial questions,
-the shuffle product of words leads to the following computational ones:
+the shuffle product of words naturally leads to the following computational
+problems:
 \begin{enumerate}
@@
 Even if these problems seem similar, they are very different in terms
-of complexity. Let us now review some facts about these. In what follows,
+of time complexity. Let us now review some facts about these. In what follows,
 $n$ denotes the size of $u$ and $m_i$ denotes the size of each $v_i$.

### 2015-09-27 submission.tex
typos

@@
   It can be shown (see Appendix) that
-  that $\sigma$ occurs in $\pi$ if and only if
+  $\sigma$ occurs in $\pi$ if and only if
   there exists a containment-free perfect matching

### 2015-09-27 submission.tex
typos

@@
     (\emph{1} $\Rightarrow$ \emph{2}).
-    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
-    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
+    Since $\pi$ is a square, it admits a square root $\sigma$.
+    Let $2n = |\pi|$ (and hence
+    $|\sigma| = n$).
+    By definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
     and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
-    letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
-    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
-    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
-    with vertices
+    $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
+    order-isomorphic to $\sigma$.
+    Let $G_\pi$ be the linear clique associated to $\pi$ with vertices
     $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It

### 2015-09-27 submission.tex
typos

@@
     order-isomorphic to $\sigma$.
-    Let $G_\pi$ be the linear clique associated to $\pi$ with vertices
-    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
+    Let $K_\pi$ be the linear clique associated to $\pi$ with vertices
+    $\VS{K_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
     is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
-    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
+    is a subset of $\ES{K_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
     matching since $I^1 \cap I^2 = \emptyset$, and
@@
     More precisely, the positions in $\pi''$ that correspond to
-    the fixed occurrence of $\sigma$ in $\pi$ are llinked
-    linked by means of $k$ pairwise crossing edges to all positions in
+    the fixed occurrence of $\sigma$ in $\pi$ are linked
+    by means of $k$ pairwise crossing edges to all positions in
     $\sigma''$.
@@
     (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
-    is involved in an edge $(y, y'$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
+    is involved in an edge $(y, y')$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
     (otherwise $\mathcal{M}$ would not be containment-free).
-    But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
+    But $|\nu'_1| = N_1 = 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1
     > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one

### 2015-09-27 submission.tex
typos

@@
     we conclude that there exists a position, say $x$, in $\nu'_1$
-    that is neither involved in a $(\nu_1, \nu'_1)$-edge nor a $(\nu_1, \nu'_1)$-edge
-    nor a $(\nu'_1, \nu'_1)$-edge.
+    that is neither involved in a $(\nu_1, \nu'_1)$-edge nor in a $(\nu_1, \nu'_1)$-edge
+    nor in a $(\nu'_1, \nu'_1)$-edge.
     Write $(x, x')$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $x' > 2N_1 + N_2$, this edge.
@@
   Then it follows we have
-  at least $k$ (possibly $k+1$ or $k+2$) independent $(\sigma', \pi')$-edges
+  at least $k$
+  %(possibly $k+1$ or $k+2$)
+  independent $(\sigma', \pi')$-edges
   $(a, a')$ in $\mathcal{M}$ with

### 2015-12-01 submission.tex
last inequality
Debut modifications indiquees par les rapporteurs

@@
     languages. In this paper, we tackle the problem of recognizing square
-    permutations from both the point of view of algebra and algorithmic.
+    permutations from both the point of view of algebra and algorithms.
     On the one hand, we present some algebraic and combinatorial
@@
     between square $(213,231)$-avoiding permutations and square binary
-    words.
-    On the other hand,
-    using a containment-free perfect matching criterion,
-    we prove that recognizing square permutations is \NPC.
+    words. On the other hand, using a containment-free perfect matching 
+    criterion, we prove that recognizing square permutations is \NPC.
 \end{abstract}
@@
 squares (Proposition~\ref{prop:square_stability}). Finally,
-Section~\ref{section:Algorithmic issues} presents the most import result
+Section~\ref{section:Algorithmic issues} presents the most important result
 of this paper: the fact that deciding if a permutation is a square is
@@
 $1 2 \dots n$ (resp. $n (n-1) \dots 1$). If $\pi$ is a permutation of
-$S_n$, we denote by $\bar \pi$ the {\em complementary} of $\pi$, that is
+$S_n$, we denote by $\bar \pi$ the {\em complement} of $\pi$, that is
 the permutation satisfying $\bar \pi(i) = n - \pi(i) + 1$ for all
@@
 and $\ES{G}$ as the set of edges of $G$. Let $u = u_1 u_2 \dots u_n$ be
-a word. The \emph{clique graph associated to $u$}, denoted by $K_u$, is
+a word. The \emph{clique graph associated with $u$}, denoted by $K_u$, is
 the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$. The
@@
 not exist (independent) edges $(u_i, u_j)$ and $(u_k, u_\ell)$ in
-$\mathcal{M}$ such that $u_i < u_k < u_\ell < u_j$ or
-$u_k < u_i < u_j < u_\ell$
+$\mathcal{M}$ such that $u_i < u_k < u_\ell < u_j$ 
+%or $u_k < u_i < u_j < u_\ell$
 (\emph{i.e.}, $\mathcal{M}$ avoids~\MWithContainment).
@@
     \Delta(213) =
-    \epsilon \otimes 213 + 2. 1 \otimes 12 +
-    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
+    \epsilon \otimes 213 + 2 \cdot 1 \otimes 12 +
+    1 \otimes 21 + 2 \cdot 12 \otimes 1 + 21 \otimes 1 
+    + 213 \otimes \epsilon,
 \end{equation}
@@
     \Delta(1234) =
-    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
-    4. 123 \otimes 1 + 1234 \otimes \epsilon,
+    \epsilon \otimes 1234 + 4 \cdot 1 \otimes 123 + 6 \cdot 12 \otimes 12 +
+    4 \cdot 123 \otimes 1 + 1234 \otimes \epsilon,
 \end{equation}
@@
     \Delta(1432) & =
-    \epsilon \otimes 1432 + 3. {\bf 1 \otimes 132} + 1 \otimes 321 +
-    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
-    321 \otimes 1 + 1432 \otimes \epsilon.
+    \epsilon \otimes 1432 + 3 \cdot {\bf 1 \otimes 132} + 1 \otimes 321 +
+    3 \cdot 12 \otimes 21 \\ & + 3 \cdot 21 \otimes 12 
+    + 3 \cdot 132 \otimes 1 + 321 \otimes 1 + 1432 \otimes \epsilon.
 \end{split}\end{equation}
@@
 
-Besides, a permutation $\pi$ is a {\em square} if there is a
-permutation $\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
+A permutation $\pi$ is a {\em square} if there is a permutation 
+$\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
 In this case, we say that $\sigma$ is a {\em square root} of $\pi$.
@@
 Let $u$ be a binary word of length $n$ with $k$ occurrences of $0$.
-We denote by $\BINTOPERM$ the map sending any such word $u$ to the
-permutation obtained by replacing from left to right each occurrence of
-$0$ in $u$ by $1$, $2$, \dots, $k$, and from right to left each
-occurrence of $1$ in $u$ by $k + 1$, $k + 2$, \dots, $n$. For instance,
+We denote by $\BINTOPERM$ (Binary word To Permutation) the map sending 
+any such word $u$ to the permutation obtained by replacing from left to 
+right each occurrence of $0$ in $u$ by $1$, $2$, \dots, $k$, and from 
+right to left each occurrence of $1$ in $u$ by $k + 1$, $k + 2$, \dots, $n$.
+For instance,
 \begin{equation}
@@
 $\BINTOPERM(u0) = \BINTOPERM(u1) = \pi$. In support of this observation,
-when $\pi$ has an even size, we denote by $\PERMTOBIN(\pi)$ the word $ua$
-such that $|ua|_0$ and $|ua|_1$ are both even, where $a \in \{0, 1\}$.
+when $\pi$ has an even size, we denote by $\PERMTOBIN(\pi)$ (Permutation
+To Binary word) the word $ua$ such that $|ua|_0$ and $|ua|_1$ are both 
+even, where $a \in \{0, 1\}$.
 
@@
     \end{equation}
-    where $I$ denote the identity map. This strategy relies on the fact
+    where $I$ denotes the identity map. This strategy relies on the fact
     that a product is associative if and only if its dual coproduct is
@@
 is that its makes sense to enumerate the sets of square permutations
-quotiented by the operations of mirror image, complementary, and
+quotiented by the operations of mirror image, complement, and
 inverse. The sequence enumerating these sets begins by
@@
   \caption{\label{fig:example containment-free matching}%
-    The linear graph $K_\pi$ associated to the permutation
+    The linear graph $K_\pi$ associated with the permutation
     $\pi = 2A47195386$ together with
@@
     order-isomorphic to $\sigma$.
-    Let $K_\pi$ be the linear clique associated to $\pi$ with vertices
+    Let $K_\pi$ be the linear clique associated with $\pi$ with vertices
     $\VS{K_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
@@
     \label{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}
-    There is neither a $(\nu_1, \nu_3)$-edge not a $(\nu'_1, \nu_3)$-edge
+    There is neither a $(\nu_1, \nu_3)$-edge nor a $(\nu'_1, \nu_3)$-edge
     in $\mathcal{M}$.
@@
   and
-  $2N_1 + N_2 + N_3 + (k + 2) + 1 < a' < 2N_1 + N_2 + N_3 + (k + 2) + n + 2$.
+  $2N_1 + 2N_2 + 2N_3 + N_4 + (k + 2) + 1 < a' <2N_1 + 2N_2 + 2N_3 + N_4 + (k + 2) + n + 2$.
   Therefore, by our hypothesis about $\mathcal{M}$,

### 2015-12-02 submission.tex
Suite modifications suggerees par les rapporteurs

@@
 as a word) is a shuffle of two words that are order-isomorphic to
-$\sigma_1$ and $\sigma_2$. Our intention in this paper is to study this
-shuffle product of permutations $\SHUFFLE$ both from a combinatorial and
-from a computational point of view by focusing on {\em square}
-permutations, that are permutations $\pi$ being in the shuffle of a
-permutation $\sigma$ with itself. The shuffle product defined and
-considered in this paper is different from other existing and well-known
-shuffle products on permutations. For instance, in~\cite{DHT:IJAC:2002},
-the authors define the {\em shifted shuffle product}. For this product,
-$\pi$ is a shuffle of $\sigma_1$ and $\sigma_2$ if $\pi$ is in the
-shuffle, as words, of $\sigma_1$ and the word obtained by incrementing
-all the letters of $\sigma_2$ by the size of $\sigma_1$. It is a simple
-exercise to prove that, given three permutations $\pi$, $\sigma_1$, and
-$\sigma_2$, deciding if $\pi$ is a in the shifted shuffle of $\sigma_1$
-and $\sigma_2$ is in~\Pclass.
+$\sigma_1$ and $\sigma_2$. This shuffle product was first introduced by
+Vargas~\cite{Vargas:2014} under the name of {\em supershuffle}. Our 
+intention in this paper is to study this shuffle product of permutations 
+$\SHUFFLE$ both from a combinatorial and from a computational point of 
+view by focusing on {\em square} permutations, that are permutations 
+$\pi$ being in the shuffle of a permutation $\sigma$ with itself. Many 
+other shuffle products on permutations appear in the literature. For 
+instance, in~\cite{DHT:IJAC:2002}, the authors define the {\em convolution 
+product} and the {\em shifted shuffle product}. For this last product, 
+$\pi$ is a shuffle of $\sigma_1$ and $\sigma_2$ if $\pi$ is in the shuffle, 
+as words, of $\sigma_1$ and the word obtained by incrementing all the 
+letters of $\sigma_2$ by the size of $\sigma_1$. It is a simple exercise 
+to prove that, given three permutations $\pi$, $\sigma_1$, and $\sigma_2$,
+deciding if $\pi$ is in the shifted shuffle of $\sigma_1$ and $\sigma_2$ 
+is in~\Pclass.
 \smallskip
@@
 If $u$ is a word of integers without multiple occurrences of a same
-letter, $\STD(u)$ is the {\em standarized} of $u$, that is the unique
+letter, $\STD(u)$ is the {\em standardized} of $u$, that is the unique
 permutation of the same size as $u$ such that for all $i, j \in [|u|]$,
@@
 image of the map $\STD$ is the set $S$ of all permutations. Two words $u$
-and $v$ having the same standarized are {\em order-isomorphic}. If
+and $v$ having the same standardized are {\em order-isomorphic}. If
 $\sigma$ is a permutation, there is an {\em occurrence} of (the
@@
 When $\pi$ does not admit any occurrence of $\sigma$, $\pi$ {\em avoids}
-$\sigma$.
+$\sigma$. The set of permutations of size $n$ avoiding $\sigma$ is 
+denoted by $S_n(\sigma)$.
 
@@
 The purpose of this section is to define a shuffle product $\SHUFFLE$
-on permutations. To define this shuffle product, we shall first define
+on permutations. Recall that a first definition of this product was 
+provided by Vargas~\cite{Vargas:2014}. To present an alternative 
+definition of this product adapted to our study, we shall first define
 a coproduct denoted by $\Delta$, enabling to unshuffle permutations.
@@
 
-Let use denote by $\QQ[S]$ the linear span of all permutations. We
+Let us denote by $\QQ[S]$ the linear span of all permutations. We
 define a linear coproduct $\Delta$ on $\QQ[S]$ in the following way. For
@@
 Intuitively, this product shuffles the values and the positions of the
-letters of the permutations. As first and easy properties of $\SHUFFLE$,
-one can observe that the empty permutation $\epsilon$ is a unit for
-$\SHUFFLE$ and that this product is graded by the sizes of the
-permutations ({\em i.e.}, the product of a permutation of size $n$ with
-a permutation of size $m$ produces a sum of permutations of size $n + m$).
-Notice that there are other shuffle products on
-permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
-{\em convolution product}. Our shuffle product is different from these
-last two.
+letters of the permutations. One can observe that the empty permutation 
+$\epsilon$ is a unit for $\SHUFFLE$ and that this product is graded by 
+the sizes of the permutations ({\em i.e.}, the product of a permutation 
+of size $n$ with a permutation of size $m$ produces a sum of permutations 
+of size $n + m$).
 
@@
 $\sigma \otimes \sigma$ with a nonzero coefficient. In a more
-combinatorial way, this is equivalent to say that there are two sets
+combinatorial way, this is equivalent to saying that there are two sets
 $P_1$ and $P_2$ of disjoints indexes of letters of $\pi$ satisfying
@@
     \BINTOPERM({\bf 1}00{\bf 1}0{\bf 1}{\bf 1}0{\bf 1}000) =
-    {\bf C} 12 {\bf B} 3 {\bf A} {\bf 9} 4 {\bf 8} 567.
+    {\bf C} 12 {\bf B} 3 {\bf A} {\bf 9} 4 {\bf 8} 567,
 \end{equation}
-Observe that for any nonempty permutation $\pi$ in the image of
-$\BINTOPERM$, there is exactly one binary word $u$ such that
-$\BINTOPERM(u0) = \BINTOPERM(u1) = \pi$. In support of this observation,
-when $\pi$ has an even size, we denote by $\PERMTOBIN(\pi)$ (Permutation
-To Binary word) the word $ua$ such that $|ua|_0$ and $|ua|_1$ are both 
-even, where $a \in \{0, 1\}$.
+where $\mathrm{A}$, $\mathrm{B}$, and $\mathrm{C}$ respectively stand 
+for $10$, $11$, and $12$. Observe that for any nonempty permutation 
+$\pi$ in the image of $\BINTOPERM$, there is exactly one binary word $u$ 
+such that $\BINTOPERM(u0) = \BINTOPERM(u1) = \pi$. In support of this
+observation, when $\pi$ has an even size, we denote by $\PERMTOBIN(\pi)$
+(Permutation To Binary word) the word $ua$ such that $|ua|_0$ and $|ua|_1$ 
+are both even, where $a \in \{0, 1\}$.
 
@@
 
-Let us make an observation about Wilf-equivalence classes of
-permutations restrained on square permutations. The
+Let us make an observation about Wilf-equivalence classes of permutations 
+restrained on square permutations. Recall that two permutations $\sigma$ 
+and $\nu$ of the same size are {\em Wilf equivalent} if 
+$\# S_n(\sigma) = \# S_n(\nu)$ for all $n \geq 0$. The
 well-known~\cite{Simion:Schmidt:EJC:1985} fact that there is a single
@@
     words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
-    same standarized $\sigma$. Hence, and by definition of
+    same standardized $\sigma$. Hence, and by definition of
     the shuffle product of permutations, $\BINTOPERM(u)$ appears in

### 2015-12-03 submission.tex
Petites modifs

@@
     12 \SHUFFLE 21 & =
-    1243 + 1324 + 2.1342 + 2.1423 + 3. {\bf 1432} + 2134 + 2.2314 \\
-    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
-    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
+    1243 + 1324 + 2 \cdot 1342 + 2 \cdot 1423 + 3 \cdot {\bf 1432} + 
+    2134 + 2 \cdot 2314 \\
+    & + 3 \cdot 2341 + 2413 + 2 \cdot 2431 + 2 \cdot 3124 + 3142 + 
+    3 \cdot 3214 + 2 \cdot 3241 \\
+    & + 3421 + 3 \cdot 4123 + 2 \cdot 4132 + 2 \cdot 4213 + 4231 + 4312.
 \end{split}\end{equation}

### 2015-12-13 annexes.tex
Creation fichier annexes.tex

--- 2015-12-01 submission.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[a4paper]{llncs}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\usepackage[utf8x]{inputenc}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{llncsdoc}
-\usepackage[english]{babel}
-\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
-\usepackage{algorithmic}
-\usepackage{float}
-\usepackage{todonotes}
-\usepackage{subfigure}
-\usepackage{enumitem}
-\usepackage{shuffle}
-\usepackage{multirow}
-
-\usepackage{tikz}
-\usetikzlibrary{matrix, calc, arrows, shapes.callouts}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{complexity}
-\input{graph}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \renewcommand{\floatpagefraction}{.9}
-% \renewcommand{\textfraction}{.1}
-\renewcommand{\arraystretch}{1.2}
-\newcommand{\rt}{\textcolor{red}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\DeclareMathOperator{\LEFT}{\ell}
-\DeclareMathOperator{\RIGHT}{r}
-\DeclareMathOperator{\QQ}{\mathbb{Q}}
-\DeclareMathOperator{\STD}{\mathrm{s}}
-\DeclareMathOperator{\SHUFFLE}{\bullet}
-\DeclareMathOperator{\BINTOPERM}{\mathrm{btp}}
-\DeclareMathOperator{\PERMTOBIN}{\mathrm{ptb}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% 1 2 2 1
-\newcommand{\MWithContainment}{%
-\begin{tikzpicture}[scale=0.1,inner sep=.8pt,node distance=.25cm]
-\draw node [draw,circle,fill=black] (U1)               {};
-\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
-\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
-\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
-\draw (U2.north) .. controls ($ (U2.north) + (0,1.75) $) and ($ (U3.north) + (0,1.75) $) .. (U3.north);
-\draw (U1.north) .. controls ($ (U1.north) + (0,2.5) $) and ($ (U4.north) + (0,2.5) $) .. (U4.north);
-\end{tikzpicture}
-}% end \MWithContainment
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newtheorem{fact}{Fact}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\makeatletter
-\newcommand{\pushright}[1]{\ifmeasuring@#1\else\omit\hfill$\displaystyle#1$\fi\ignorespaces}
-\newcommand{\pushleft}[1]{\ifmeasuring@#1\else\omit$\displaystyle#1$\hfill\fi\ignorespaces}
-\makeatother
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{%
-Unshuffling Permutations}%
-
-\author{%
-  Samuele Giraudo \and
-  St\'ephane Vialette
-}% end author
-\institute{%
-  Universit\'e Paris-Est, LIGM (UMR 8049), CNRS, UPEM, ESIEE Paris, ENPC,
-  F-77454, Marne-la-Vallée, France\\
-  \email{samuele.giraudo@univ-mlv.fr} \\
-  \email{vialette@univ-mlv.fr}
-}% end institute
-\date{\today}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{abstract}
-    A permutation is said to be a square if it can be obtained by
-    shuffling two order-isomorphic patterns. The definition is intended
-    to be the natural counterpart to the ordinary shuffle of words and
-    languages. In this paper, we tackle the problem of recognizing square
-    permutations from both the point of view of algebra and algorithms.
-    On the one hand, we present some algebraic and combinatorial
-    properties of the shuffle product of permutations. We follow an
-    unusual line consisting in defining the shuffle of permutations by
-    means of an unshuffling operator, known as a coproduct. This
-    strategy allows to obtain easy proofs for algebraic and combinatorial
-    properties of our shuffle product. We besides exhibit a bijection
-    between square $(213,231)$-avoiding permutations and square binary
-    words. On the other hand, using a containment-free perfect matching 
-    criterion, we prove that recognizing square permutations is \NPC.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%%
-%% ---- Introduction ----
-%%
-\section{Introduction}
-\label{section:Introduction}
-The {\em shuffle product}, denoted $\shuffle$, is a very famous operation on words
-first defined by Eilenberg and Mac Lane~\cite{Eilenberg:MacLane:1953}.
-Given three words $u$, $v_1$, and $v_2$, $u$ is said to be a \emph{shuffle}
-of $v_1$ and $v_2$ if it can be formed by interleaving the letters from
-$v_1$ and $v_2$ in a way that maintains the left-to-right ordering of the
-letters from each word.
-Besides purely combinatorial questions,
-the shuffle product of words naturally leads to the following computational
-problems:
-\begin{enumerate}
-    \item \label{item:problem_1}
-    Given two words $v_1$ and $v_2$, compute the set $v_1 \shuffle v_2$.
-    \item \label{item:problem_2}
-    Given three words $u$, $v_1$, and $v_2$, decide if $u$ is a shuffle
-    of $v_1$ and $v_2$.
-    \item \label{item:problem_3}
-    Given words $u$, $v_1$, \dots, $v_k$, decide if $u$ is in
-    $v_1 \shuffle \dots \shuffle v_k$.
-    \item \label{item:problem_4}
-    Given a word $u$, decide if there is a word $v$ such that $u$ is
-    in $v \shuffle v$.
-\end{enumerate}
-Even if these problems seem similar, they are very different in terms
-of time complexity. Let us now review some facts about these. In what follows,
-$n$ denotes the size of $u$ and $m_i$ denotes the size of each $v_i$.
-A solution to Problem~\ref{item:problem_1} can be computed in
-% \begin{equation}
-%     O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)
-% \end{equation}
-$O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)$
-time~\cite{Spehner:TCS:1986}. An improvement and a generalization
-of~Problem~\ref{item:problem_1} has been proposed
-in~\cite{Allauzen:IGM:2000}, where it is proved that given words
-$v_1$, \dots, $v_k$, the iterated shuffle
-$v_1 \shuffle \dots \shuffle v_k$ can be computed in
-% \begin{equation}
-%     O\left(\binom{m_1 + \dots + m_k}{m_1, \dots, m_k}\right)
-% \end{equation}
-$O\left(\binom{m_1 + \dots + m_k}{m_1, \dots, m_k}\right)$
-time. Problem~\ref{item:problem_2} is in \Pclass; it is indeed a
-classical textbook exercise to design an efficient dynamic programming
-algorithm solving it. It can be tested in $O\left(n^2 / \log(n)\right)$
-time~\cite{Leeuwen:Nivat:IPL:1982}. To the best of our knowledge, the
-first $O(n^2)$ time algorithm for this problem appeared
-in~\cite{Mansfield:DAM:1983}. This algorithm can easily be extended to
-check in polynomial-time whether or not a word is in the shuffle of any
-fixed number of given words. Nevertheless, Problem~\ref{item:problem_3}
-is \NPC~\cite{Mansfield:DAM:1983,Warmuth:Haussler:JCSS:1984}. This
-remains true even if the ground alphabet has size
-$3$~\cite{Warmuth:Haussler:JCSS:1984}. Of particular interest,
-it is shown in~\cite{Warmuth:Haussler:JCSS:1984} that
-Problem~\ref{item:problem_3} remains \NPC even if all the words $v_i$,
-$i \in [k]$, are identical, thereby proving that, for two words $u$ and
-$v$, it is \NPC to decide whether or not $u$ is in the iterated shuffle
-of $v$. Again, this remains true even if the ground alphabet has size $3$.
-Let us now finally focus on Problem~\ref{item:problem_4}. It is shown
-in~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013} that it is \NPC to
-decide if a word $u$ is a \emph{square} (w.r.t. the shuffle), that is
-a word $u$ with the property that there exists a word $v$ such that $u$
-is a shuffle of $v$ with itself. Hence, Problem~\ref{item:problem_4}
-is \NPC.
-\smallskip
-
-This paper is intended to study a natural generalization of $\shuffle$,
-denoted by $\SHUFFLE$, as a shuffle of permutations. Roughly speaking,
-given three permutations $\pi$, $\sigma_1$, and $\sigma_2$, $\pi$ is
-said to be a {\em shuffle} of $\sigma_1$ and $\sigma_2$ if $\pi$ (viewed
-as a word) is a shuffle of two words that are order-isomorphic to
-$\sigma_1$ and $\sigma_2$. Our intention in this paper is to study this
-shuffle product of permutations $\SHUFFLE$ both from a combinatorial and
-from a computational point of view by focusing on {\em square}
-permutations, that are permutations $\pi$ being in the shuffle of a
-permutation $\sigma$ with itself. The shuffle product defined and
-considered in this paper is different from other existing and well-known
-shuffle products on permutations. For instance, in~\cite{DHT:IJAC:2002},
-the authors define the {\em shifted shuffle product}. For this product,
-$\pi$ is a shuffle of $\sigma_1$ and $\sigma_2$ if $\pi$ is in the
-shuffle, as words, of $\sigma_1$ and the word obtained by incrementing
-all the letters of $\sigma_2$ by the size of $\sigma_1$. It is a simple
-exercise to prove that, given three permutations $\pi$, $\sigma_1$, and
-$\sigma_2$, deciding if $\pi$ is a in the shifted shuffle of $\sigma_1$
-and $\sigma_2$ is in~\Pclass.
-\smallskip
-
-This paper is organized as follows. In
-Section~\ref{section:Shuffle product on permutations} we provide a
-precise definition of $\SHUFFLE$. This definition passes through the
-preliminary definition of an operator~$\Delta$, allowing to
-{\em unshuffle} permutations. This operator is in fact a coproduct,
-endowing the linear span of all permutations with a coalgebra structure
-(see~\cite{Joni:Rota:1979} or~\cite{Grinberg:Reiner:2014} for the
-definition of these algebraic structures). By duality, the unshuffling
-operator $\Delta$ leads to the definition of our shuffle operation on
-permutations. This approach has many advantages. First, some
-combinatorial properties of $\SHUFFLE$ depend on properties of $\Delta$
-and are more easy to prove on the coproduct side. Second, this way of
-doing allows to obtain a clear description of the multiplicities of the
-elements appearing in the shuffle of two permutations, which are worthy
-of interest from a combinatorial point of view.
-Section~\ref{section:Binary square words and permutations} is devoted to
-show that the problems related to the shuffle of words has links with
-the shuffle of permutations. In particular, we show that binary words
-that are square are in one-to-one correspondence with square
-permutations avoiding some patterns
-(Proposition~\ref{prop:bijection_binary_to_permutations_squares}). Next,
-Section~\ref{section:Algebraic issues} presents some algebraic and
-combinatorial properties of~$\SHUFFLE$. We show that $\SHUFFLE$ is
-associative and commutative
-(Proposition~\ref{prop:shuffle_associative_commutative}), and that if a
-permutation is a square, its mirror, complement, and inverse are also
-squares (Proposition~\ref{prop:square_stability}). Finally,
-Section~\ref{section:Algorithmic issues} presents the most important result
-of this paper: the fact that deciding if a permutation is a square is
-\NPC (Proposition~\ref{proposition:hardness}). This result is obtained
-by exhibiting a reduction from the pattern involvement
-problem~\cite{Bose:Buss:Lubiw:1998} which is \NPC.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Notations ----
-%%
-\section{Notations}
-\label{section:Notations}
-
-If $S$ is a finite set, the cardinality of $S$ is denoted by $|S|$,
-and if $P$ and $Q$ are two disjoint sets, $P \sqcup Q$ denotes the
-disjoint union of $P$ and $Q$. For any nonnegative integer $n$, $[n]$
-is the set $\{1, \dots, n\}$.
-
-We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
-Let us recall here the most important ones. Let $u$ be a word. The
-length of $u$ is denoted by $|u|$. The {\em empty word}, the only word
-of null length, is denoted by $\epsilon$. We denote by $\widetilde{u}$
-the {\em mirror image} of $u$, that is the word
-$u_{|u|} u_{|u| - 1} \dots u_1$. If $P$ is a subset of $[|u|]$, $u_{|P}$
-is the subword of $u$ consisting in the letters of $u$ at the positions
-specified by the elements of $P$. If $u$ is a word of integers and $k$
-is an integer, we denote by $u[k]$ the word obtained by incrementing by
-$k$ all letters of $u$. The {\em shuffle} of two words $u$ and $v$ is
-the set recursively defined by
-$u \shuffle \epsilon = \{u\} = \epsilon \shuffle u$ and
-$ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
-$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
-such that $u$ belongs to $v \shuffle v$.
-
-We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
-set of all permutations. In this paper, permutations of a size $n$ are
-specified by words of length $n$ on the alphabet $[n]$ and without
-multiple occurrence of a letter, so that all above definitions about
-words remain valid on permutations. The only difference lies on the
-fact that we shall denote by $\pi(i)$ (instead of $\pi_i$) the $i$-th
-letter of any permutation $\pi$. For any nonnegative integer $n$, we
-write $\nearrow_{n}$ (resp. $\searrow_{n}$) for the permutation
-$1 2 \dots n$ (resp. $n (n-1) \dots 1$). If $\pi$ is a permutation of
-$S_n$, we denote by $\bar \pi$ the {\em complement} of $\pi$, that is
-the permutation satisfying $\bar \pi(i) = n - \pi(i) + 1$ for all
-$i \in [n]$. The {\em inverse} of $\pi$ is denoted by $\pi^{-1}$.
-
-If $u$ is a word of integers without multiple occurrences of a same
-letter, $\STD(u)$ is the {\em standarized} of $u$, that is the unique
-permutation of the same size as $u$ such that for all $i, j \in [|u|]$,
-$u_i < u_j$ if and only if $\STD(u)(i) < \STD(u)(j)$. In particular, the
-image of the map $\STD$ is the set $S$ of all permutations. Two words $u$
-and $v$ having the same standarized are {\em order-isomorphic}. If
-$\sigma$ is a permutation, there is an {\em occurrence} of (the
-{\em pattern}) $\sigma$ in $\pi$ if there is a set $P$ of indexes of
-letters of $\pi$ such that $\sigma$ and $\pi_{|P}$ are order-isomorphic.
-When $\pi$ does not admit any occurrence of $\sigma$, $\pi$ {\em avoids}
-$\sigma$.
-
-Let us now provide some definitions about graphs and matching that are used in
-the sequel. For a graph $G$, we denote $\VS{G}$ as the set of vertices
-and $\ES{G}$ as the set of edges of $G$. Let $u = u_1 u_2 \dots u_n$ be
-a word. The \emph{clique graph associated with $u$}, denoted by $K_u$, is
-the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$. The
-structure of this underlying graph is linear, \emph{i.e.}, the set of
-its vertices is equipped with a natural total order $<$ defined by
-$u_i < u_j$ if and only if $i < j$. For a linear graph, we write
-$(u_i, u_j)$ instead of $\{u_i, u_j\}$ for an edge if it is clear from
-the context that $i < j$. Recall that two edges of a graph are
-\emph{independent} if they do not share a common vertex, and that a
-\emph{matching} $\mathcal{M}$ in $G$ is a set of pairwise independent
-edges. A matching is \emph{perfect} if it covers all the vertices of the
-graph.
-For an edge $e = \{u,v\}$, we shall occasionally say that $u$ and $v$
-are \emph{involved} in $e$.
-In case the set of vertices is equipped with a total order, a
-matching $\mathcal{M}$ is said to be \emph{containment-free} if there do
-not exist (independent) edges $(u_i, u_j)$ and $(u_k, u_\ell)$ in
-$\mathcal{M}$ such that $u_i < u_k < u_\ell < u_j$ 
-%or $u_k < u_i < u_j < u_\ell$
-(\emph{i.e.}, $\mathcal{M}$ avoids~\MWithContainment).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Shuffle product on permutations ----
-%%
-\section{Shuffle product on permutations}
-\label{section:Shuffle product on permutations}
-
-The purpose of this section is to define a shuffle product $\SHUFFLE$
-on permutations. To define this shuffle product, we shall first define
-a coproduct denoted by $\Delta$, enabling to unshuffle permutations.
-By duality, $\Delta$ implies the definition of $\SHUFFLE$. The reason
-why we need to pass by the definition of $\Delta$ to define $\SHUFFLE$
-is justified by the fact that a lot of properties of $\SHUFFLE$ depend
-of properties of $\Delta$, and that this strategy allows to write concise
-and clear proofs of them. We invite the reader unfamiliar with the
-concepts of coproduct and duality to consult~\cite{Joni:Rota:1979}
-or~\cite{Grinberg:Reiner:2014}.
-
-Let use denote by $\QQ[S]$ the linear span of all permutations. We
-define a linear coproduct $\Delta$ on $\QQ[S]$ in the following way. For
-any permutation $\pi$, we set
-\begin{equation} \label{equ:unshuffling_coproduct}
-    \Delta(\pi) =
-    \sum_{P_1 \sqcup P_2 = [|\pi|]}
-    \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
-\end{equation}
-We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
-instance,
-\begin{equation}
-    \Delta(213) =
-    \epsilon \otimes 213 + 2 \cdot 1 \otimes 12 +
-    1 \otimes 21 + 2 \cdot 12 \otimes 1 + 21 \otimes 1 
-    + 213 \otimes \epsilon,
-\end{equation}
-\begin{equation}
-    \Delta(1234) =
-    \epsilon \otimes 1234 + 4 \cdot 1 \otimes 123 + 6 \cdot 12 \otimes 12 +
-    4 \cdot 123 \otimes 1 + 1234 \otimes \epsilon,
-\end{equation}
-\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
-    \Delta(1432) & =
-    \epsilon \otimes 1432 + 3 \cdot {\bf 1 \otimes 132} + 1 \otimes 321 +
-    3 \cdot 12 \otimes 21 \\ & + 3 \cdot 21 \otimes 12 
-    + 3 \cdot 132 \otimes 1 + 321 \otimes 1 + 1432 \otimes \epsilon.
-\end{split}\end{equation}
-Observe that the coefficient of the tensor $1 \otimes 132$ is $3$
-in~\eqref{equ:example_unshuffling_coproduct} because there are exactly
-three ways to extract from the permutation $1432$ two disjoint subwords
-respectively order-isomorphic to the permutations $1$ and $132$.
-
-As announced, let us now use $\Delta$ to define a shuffle product on
-permutations. As any coproduct, $\Delta$ leads to the definition of a
-product obtained by duality in the following way.
-From~\eqref{equ:unshuffling_coproduct}, for any permutation $\pi$, we
-have
-\begin{equation} \label{equ:schematic_coproduct}
-    \Delta(\pi) =
-    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
-    \sigma \otimes \nu,
-\end{equation}
-where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
-by definition of duality, the dual product of $\Delta$, denoted by
-$\SHUFFLE$, is a linear binary product on $\QQ[S]$. It satisfies, for
-any permutations $\sigma$ and $\nu$,
-\begin{equation}
-    \sigma \SHUFFLE \nu =
-    \sum_{\pi \in S}
-    \lambda_{\sigma, \nu}^\pi \; \pi,
-\end{equation}
-where the coefficients $\lambda_{\sigma, \nu}^\pi$ are the ones
-of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
-{\em shuffle product of permutations}. For instance,
-\begin{equation}\begin{split} \label{equ:example_shuffle_product}
-    12 \SHUFFLE 21 & =
-    1243 + 1324 + 2.1342 + 2.1423 + 3. {\bf 1432} + 2134 + 2.2314 \\
-    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
-    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
-\end{split}\end{equation}
-Observe that the coefficient $3$ of the permutation $1432$
-in~\eqref{equ:example_shuffle_product} comes from the fact that the
-coefficient of the tensor $12 \otimes 21$ is $3$
-in~\eqref{equ:example_unshuffling_coproduct}.
-
-Intuitively, this product shuffles the values and the positions of the
-letters of the permutations. As first and easy properties of $\SHUFFLE$,
-one can observe that the empty permutation $\epsilon$ is a unit for
-$\SHUFFLE$ and that this product is graded by the sizes of the
-permutations ({\em i.e.}, the product of a permutation of size $n$ with
-a permutation of size $m$ produces a sum of permutations of size $n + m$).
-Notice that there are other shuffle products on
-permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
-{\em convolution product}. Our shuffle product is different from these
-last two.
-
-We say that a permutation $\pi$ {\em appears} in the shuffle
-$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
-coefficient $\lambda_{\sigma, \nu}^\pi$ defined above is different from
-zero. In a more combinatorial way, this is equivalent to say that there
-are two sets $P_1$ and $P_2$ of disjoints indexes of letters of $\pi$
-satisfying $P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$
-is order-isomorphic to $\sigma$ and the subword $\pi_{|P_2}$ is
-order-isomorphic to $\nu$.
-
-A permutation $\pi$ is a {\em square} if there is a permutation 
-$\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
-In this case, we say that $\sigma$ is a {\em square root} of $\pi$.
-Equivalently, $\pi$ is a square with $\sigma$ as square root if and only
-if in the expansion of $\Delta(\pi)$, there is a tensor
-$\sigma \otimes \sigma$ with a nonzero coefficient. In a more
-combinatorial way, this is equivalent to say that there are two sets
-$P_1$ and $P_2$ of disjoints indexes of letters of $\pi$ satisfying
-$P_1 \sqcup P_2 = [|\pi|]$ such that the subwords $\pi_{|P_1}$ and
-$\pi_{|P_2}$ are order-isomorphic. Computer experiments give us the
-first numbers of square permutations with respects to their size, which
-are, from size $0$ to $10$,
-\begin{equation}
-    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
-\end{equation}
-This sequence (and its subsequence obtained by removing the $0$'s) is for
-the time being not listed in~\cite{Slo}. The square permutations of
-sizes $0$ to $4$ are
-\smallskip
-
-\begin{tabular}{c|c|c}
-    Size $0$ \, & Size $2$ & Size $4$ \\ \hline
-    \multirow{2}{*}{\, $\epsilon$} &
-    \multirow{2}{*}{\, $12$, $21$ \,} &
-    $1234$, $1243$, $1423$, $1324$, $1342$, $4132$, $3124$, $3142$,
-    $3412$, $4312$, \\
-    & & $2134$, $2143$, $2413$, $4213$, $2314$, $2431$, $4231$, $3241$,
-    $3421$, $4321$
-\end{tabular}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Binary square words and permutations ----
-%%
-\section{Binary square words and permutations}
-\label{section:Binary square words and permutations}
-In this section, we shall show that the square binary words are in
-one-to-one correspondence with square permutations avoiding some
-patterns. This property establishes a link between the shuffle of binary
-words and our shuffle of permutations and allows to obtain a new
-description of square binary words.
-
-Let $u$ be a binary word of length $n$ with $k$ occurrences of $0$.
-We denote by $\BINTOPERM$ (Binary word To Permutation) the map sending 
-any such word $u$ to the permutation obtained by replacing from left to 
-right each occurrence of $0$ in $u$ by $1$, $2$, \dots, $k$, and from 
-right to left each occurrence of $1$ in $u$ by $k + 1$, $k + 2$, \dots, $n$.
-For instance,
-\begin{equation}
-    \BINTOPERM({\bf 1}00{\bf 1}0{\bf 1}{\bf 1}0{\bf 1}000) =
-    {\bf C} 12 {\bf B} 3 {\bf A} {\bf 9} 4 {\bf 8} 567.
-\end{equation}
-Observe that for any nonempty permutation $\pi$ in the image of
-$\BINTOPERM$, there is exactly one binary word $u$ such that
-$\BINTOPERM(u0) = \BINTOPERM(u1) = \pi$. In support of this observation,
-when $\pi$ has an even size, we denote by $\PERMTOBIN(\pi)$ (Permutation
-To Binary word) the word $ua$ such that $|ua|_0$ and $|ua|_1$ are both 
-even, where $a \in \{0, 1\}$.
-
-\begin{proposition} \label{prop:bijection_binary_to_permutations_squares}
-    For any $n \geq 0$, the map $\BINTOPERM$ restricted to the set of
-    square binary words of length $2n$ is a bijection between this last
-    set and the set of square permutations of size $2n$ avoiding the
-    patterns $213$ and $231$.
-\end{proposition}
-\begin{proof}
-    [of Proposition~\ref{prop:bijection_binary_to_permutations_squares}]
-    The statement of the proposition is a consequence of the
-    following claims implying that $\PERMTOBIN$ is the inverse
-    map of $\BINTOPERM$ over the set of square binary words.
-    \begin{claim} \label{claim:binary_to_permutation_avoiding}
-        The image of $\BINTOPERM$ is the set of all permutations
-        avoiding $213$ and $231$.
-    \end{claim}
-
-    \begin{claim} \label{claim:square_binary_to_square_permutation}
-        If $u$ is a square binary word, $\BINTOPERM(u)$ is a square
-        permutation.
-    \end{claim}
-
-    \begin{claim} \label{claim:square_permutation_to_square_binary}
-        If $\pi$ is a square permutation avoiding $213$ and $231$,
-        $\PERMTOBIN(\pi)$ is a square binary word.
-    \end{claim}
-    \qed
-\end{proof}
-
-The number of square binary words is Sequence A191755 of~\cite{Slo}
-beginning by
-\begin{equation}
-1, 0, 2, 0, 6, 0, 22, 0, 82, 0, 320, 0, 1268, 0, 5102, 0, 020632.
-\end{equation}
-By Proposition~\ref{prop:bijection_binary_to_permutations_squares}, this
-is also the sequence enumerating square permutations avoiding $213$
-and $231$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Algebraic issues ----
-%%
-\section{Algebraic issues}
-\label{section:Algebraic issues}
-The aim of this section is to establish some of properties of the
-shuffle product of permutations $\SHUFFLE$. It is worth to note that, as
-we will see, algebraic properties of the unshuffling coproduct $\Delta$
-of permutations defined in
-Section~\ref{section:Shuffle product on permutations} lead to
-combinatorial properties of $\SHUFFLE$.
-
-\begin{proposition} \label{prop:shuffle_associative_commutative}
-    The shuffle product $\SHUFFLE$ of permutations is associative and
-    commutative.
-\end{proposition}
-\begin{proof}[of Proposition~\ref{prop:shuffle_associative_commutative}]
-    To prove the associativity of $\SHUFFLE$, it is convenient to show
-    that its dual coproduct $\Delta$ is coassociative, that is
-    \begin{equation}
-        (\Delta \otimes I) \Delta = (I \otimes \Delta) \Delta,
-    \end{equation}
-    where $I$ denotes the identity map. This strategy relies on the fact
-    that a product is associative if and only if its dual coproduct is
-    coassociative. For any permutation $\pi$, we have
-    \begin{equation} \begin{split}
-    \label{equ:shuffle_associative_commutative}
-        (\Delta \otimes I) \Delta(\pi) & =
-        (\Delta \otimes I)
-        \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
-        & =
-        \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes I\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & =
-        \sum_{P_1 \sqcup P_2 = [|\pi|]} \;
-        \sum_{Q_1 \sqcup Q_2 = [|P_1|]}
-        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
-        \otimes
-        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_2}\right)
-        \otimes \STD\left(\pi_{|P_2}\right) \\
-        & =
-        \sum_{P_1 \sqcup P_2 \sqcup P_3 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right) \otimes
-        \STD\left(\pi_{|P_2}\right) \otimes
-        \STD\left(\pi_{|P_3}\right).
-    \end{split} \end{equation}
-    An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
-    is equal to the last member
-    of~\eqref{equ:shuffle_associative_commutative}, whence the
-    associativity of $\SHUFFLE$.
-    \smallskip
-
-    Finally, to prove the commutativity of $\SHUFFLE$, we shall show
-    that $\Delta$ is cocommutative, that is for any permutation $\pi$,
-    if in the expansion of $\Delta(\pi)$ there is a tensor
-    $\sigma \otimes \nu$ with a coefficient $\lambda$, there is in the
-    same expansion the tensor $\nu \otimes \sigma$ with the same
-    coefficient $\lambda$. Clearly, a product is commutative if and only
-    if its dual coproduct is cocommutative. Now, from the
-    definition~\eqref{equ:unshuffling_coproduct} of $\Delta$, one
-    observes that if the pair $(P_1, P_2)$ of subsets of $[|\pi|]$
-    contributes to the coefficient of
-    $\STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right)$,
-    the pair $(P_2, P_1)$ contributes to the coefficient of
-    $\STD\left(\pi_{|P_2}\right) \otimes \STD\left(\pi_{|P_1}\right)$.
-    This shows that $\Delta$ is cocommutative and hence, that $\SHUFFLE$
-    is commutative.
-    \qed
-\end{proof}
-
-Proposition~\ref{prop:shuffle_associative_commutative} implies that
-$\QQ[S]$ endowed with the unshuffling coproduct $\Delta$ is a
-coassociative cocommutative coalgebra, or in an equivalent way, that
-$\QQ[S]$ endowed with the shuffle product $\SHUFFLE$ is an associative
-commutative algebra.
-
-\begin{lemma} \label{lem:endomorphisms}
-    The three linear maps
-    \begin{equation}
-        \phi_1, \phi_2, \phi_3 : \QQ[S] \to \QQ[S]
-    \end{equation}
-    linearly sending a permutation $\pi$ to, respectively,
-    $\widetilde{\pi}$, $\bar \pi$, and $\pi^{-1}$ are endomorphims of
-    associative algebras.
-\end{lemma}
-
-We now use the algebraic properties of $\SHUFFLE$ exhibited by
-Lemma~\ref{lem:endomorphisms} to obtain combinatorial properties
-of square permutations.
-
-\begin{proposition} \label{prop:square_stability}
-    Let $\pi$ be a square permutation and $\sigma$ be a square root of
-    $\pi$. Then,
-    \begin{enumerate}[label={\it (\roman*)},fullwidth]
-        \item \label{item:square_stability_1}
-        the permutation $\widetilde{\pi}$ is a square and
-        $\widetilde{\sigma}$ is one of its square roots;
-        \item \label{item:square_stability_2}
-        the permutation $\bar \pi$ is a square and $\bar \sigma$ is one of
-        its square roots;
-        \item \label{item:square_stability_3}
-        the permutation $\pi^{-1}$ is a square and $\sigma^{-1}$ is one of
-        its square roots.
-    \end{enumerate}
-\end{proposition}
-\begin{proof}[of Proposition~\ref{prop:square_stability}]
-    All statements~\ref{item:square_stability_1},
-    \ref{item:square_stability_2}, and~\ref{item:square_stability_3} are
-    consequences of Lemma~\ref{lem:endomorphisms}. Indeed,
-    since $\pi$ is a square permutation and $\sigma$ is a square root of
-    $\pi$, by definition, $\pi$ appears in the product
-    $\sigma \SHUFFLE \sigma$. Now, by Lemma~\ref{lem:endomorphisms},
-    for any $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
-    algebras from $\QQ[S]$ to $\QQ[S]$, $\phi_j$ commutes with the
-    shuffle product of permutations $\SHUFFLE$. Hence, in particular,
-    one has
-    \begin{equation}
-        \phi_j(\sigma \SHUFFLE \sigma) =
-        \phi_j(\sigma) \SHUFFLE \phi_j(\sigma).
-    \end{equation}
-    Then, since $\pi$ appears in $\sigma \SHUFFLE \sigma$, $\phi_j(\pi)$
-    appears in $\phi_j(\sigma \SHUFFLE \sigma)$ and appears also in
-    $\phi_j(\sigma) \SHUFFLE \phi_j(\sigma)$. This shows that
-    $\phi_j(\sigma)$ is a square root of $\phi_j(\pi)$ and
-    implies~\ref{item:square_stability_1}, \ref{item:square_stability_2},
-    and~\ref{item:square_stability_3}.
-    \qed
-\end{proof}
-
-Let us make an observation about Wilf-equivalence classes of
-permutations restrained on square permutations. The
-well-known~\cite{Simion:Schmidt:EJC:1985} fact that there is a single
-Wilf-equivalence class of permutations of size $3$ together with
-Proposition~\ref{prop:square_stability} imply that $123$ and $321$ are
-in the same Wilf-equivalence class of square permutations, and that
-$132$, $213$, $231$, and $312$ are in the same Wilf-equivalence class of
-square permutations. Computer experiments show us that there are two
-Wilf-equivalence classes of square permutations of size $3$. Indeed, the
-number of square permutations avoiding $123$ begins by
-\begin{equation} \label{equ:sequence_square_123}
-    1, 0, 2, 0, 12, 0, 118, 0, 1218, 0, 14272,
-\end{equation}
-while the number of square permutations avoiding $132$ begins by
-\begin{equation} \label{equ:sequence_square_132}
-    1, 0, 2, 0, 11, 0, 84, 0, 743, 0, 7108.
-\end{equation}
-
-Besides, an other consequence of Proposition~\ref{prop:square_stability}
-is that its makes sense to enumerate the sets of square permutations
-quotiented by the operations of mirror image, complement, and
-inverse. The sequence enumerating these sets begins by
-\begin{equation} \label{equ:sequence_square_classes}
-    1, 0, 1, 0, 6, 0, 81, 0, 2774, 0, 162945.
-\end{equation}
-
-All Sequences~\eqref{equ:sequence_square_123}, \eqref{equ:sequence_square_132},
-and~\eqref{equ:sequence_square_classes} (and their subsequences obtained
-by removing the $0$s) are for the time being not listed in~\cite{Slo}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Algorithmic issues ----
-%%
-\section{Algorithmic issues}
-\label{section:Algorithmic issues}
-
-This section is devoted to proving hardness of recognizing square
-permutations. In the same way as happens with words, we shall use a
-linear graph framework where deciding whether a permutation is a square
-reduces to computing some specific matching in the associated linear
-graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows
-(see Fig.\ref{fig:example containment-free matching}).
-\begin{figure}[ht!]
-  \centering
-  \begin{tikzpicture}[scale=0.5,inner sep=2pt,node distance=1cm]
-    \draw node (K) {$K_\pi$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=K]   (U01) {$2$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U01] (U02) {$A$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U02] (U03) {$4$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U03] (U04) {$7$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U04] (U05) {$1$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U05] (U06) {$9$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U06] (U07) {$5$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U07] (U08) {$3$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U08] (U09) {$8$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U09] (U10) {$6$};
-    \draw [thick]
-      (U01.north) .. controls ($ (U01.north) + (0,2.25) $) and ($ (U03.north) + (0,2.25) $) .. (U03.north);
-    \draw [thick]
-      (U02.north) .. controls ($ (U02.north) + (0,2.25) $) and ($ (U06.north) + (0,2.25) $) .. (U06.north);
-    \draw [thick]
-      (U04.north) .. controls ($ (U04.north) + (0,2.5) $) and ($ (U07.north) + (0,2.5) $) .. (U07.north);
-    \draw [thick]
-      (U05.north) .. controls ($ (U05.north) + (0,2.5) $) and ($ (U08.north) + (0,2.5) $) .. (U08.north);
-    \draw [thick]
-      (U09.north) .. controls ($ (U09.north) + (0,2.5) $) and ($ (U10.north) + (0,2.5) $) .. (U10.north);
-  \end{tikzpicture}
-  \caption{\label{fig:example containment-free matching}%
-    The linear graph $K_\pi$ associated with the permutation
-    $\pi = 2A47195386$ together with
-    a containment-free perfect matching $\mathcal{M}$
-    such that for any two distinct edges $(a, b)$ and
-    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
-    if $\pi(b) < \pi(b')$.
-    From $\mathcal{M}$, it follows that $\pi$ is square as it is in
-    the shuffle of $2A718$ and $49536$, both being
-    order-isomorphic to $25314$.
-  }% end caption
-\end{figure}
-
-\begin{figure}[t!]
-  \centering
-  \subfigure[%
-    An increasing pattern before and above a decreasing pattern.
-  ]{%
-    \begin{tikzpicture}
-      [
-        scale=0.325,
-        label/.style={anchor=base}
-      ]
-      % increasing pattern
-      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
-      \foreach \x/\y in {0/6,1/7,2/8,3/9,4/10,5/11} {
-          \draw [fill=black] (\x,\y) circle (0.1);
-      }
-      % decreasing pattern
-      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
-      \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
-          \draw [fill=black] (\x,\y) circle (0.1);
-      }
-      % labels
-      \node [label] (a) at (1,5) {$a$};
-      \node [label] (b) at (4,5) {$b$};
-      \node [label] (ap) at (7,-1) {$a'$};
-      \node [label] (bp) at (10,-1) {$b'$};
-      \node (nu1) at (2.5,4) {$\sigma_1$};
-      \node (nu2) at (8.5,-2) {$\sigma_2$};
-      % crossings edges
-      \draw [line width=1pt]
-      (1,7) .. controls +(0,12) and +(0,10) .. (7,4);
-      \draw [line width=1pt]
-      (4,10) .. controls +(0,12) and +(0,10) .. (10,1);
-    \end{tikzpicture}
-    \label{subfig:increasing before and above decreasing}
-  }% end subfigure
-  \qquad\qquad\qquad
-  \subfigure[%
-    A decreasing pattern before and below an increasing pattern.
-  ]{%
-  \begin{tikzpicture}
-    [
-      scale=0.325,
-      label/.style={anchor=base}
-    ]
-    % decreasing pattern
-    \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (5.2,5.2);
-    \foreach \x/\y in {0/5,1/4,2/3,3/2,4/1,5/0} {
-        \draw [fill=black] (\x,\y) circle (0.1);
-    }
-    % decreasing pattern
-    \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,5.8) grid (11.2,11.2);
-    \foreach \x/\y in {6/6,7/7,8/8,9/9,10/10,11/11} {
-        \draw [fill=black] (\x,\y) circle (0.1);
-    }
-    % labels
-    \node [label] (a) at (1,-1) {$a$};
-    \node [label] (b) at (4,-1) {$b$};
-    \node [label] (ap) at (7,5) {$a'$};
-    \node [label] (bp) at (10,5) {$b'$};
-    \node (nu1) at (2.5,-2) {$\sigma_1$};
-    \node (nu2) at (8.5,4) {$\sigma_2$};
-    % crossings edges
-    \draw [line width=1pt]
-    (1,4) .. controls +(0,10) and +(0,12) .. (7,7);
-    \draw [line width=1pt]
-    (4,1) .. controls +(0,10) and +(0,12) .. (10,10);
-    \end{tikzpicture}
-    \label{subfig:decreasing before and below increasing}
-  }% end subfigure
-  \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
-    Illustration of Lemma~\ref{lemma:at most one edge monotone}.
-  }
-\end{figure}
-
-\begin{lemma}
-  \label{lemma:matching}
-  Let $\pi$ be a permutation. The following statements are equivalent:
-  \begin{enumerate}
-    \item The permutation $\pi$ is a square.
-    \item There exists a perfect containment-free matching $\mathcal{M}$
-    of $K_\pi$ such that for any two distinct edges $(a, b)$ and
-    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
-    if $\pi(b) < \pi(b')$.
-  \end{enumerate}
-\end{lemma}
-
-Let $\pi$ be a permutation. For the sake of clarity, we will say that a
-bunch of consecutive positions $P$ of $\pi$ is \emph{above} (resp.
-\emph{below}) above another bunch of consecutive positions $P'$ in $\pi$
-if $\pi(i) > \pi(j)$ (resp. $\pi(i) < \pi(j)$) for every $i \in P$ and
-every $j \in P'$. For example, $\sigma_1$ is above $\sigma_2$ (in an
-equivalent manner, $\sigma_2$ is below $\sigma_1$) in
-Fig.~\ref{subfig:increasing before and above decreasing}, whereas
-$\sigma_1$ is below $\sigma_2$ (in an equivalent manner, $\sigma_2$ is
-above $\sigma_1$) in Fig.~\ref{subfig:decreasing before and below
-increasing}.
-
-Before proving hardness, we give an easy lemma that will prove extremely
-useful for simplifying the proof of upcoming
-Proposition~\ref{proposition:hardness}.
-
-\begin{lemma}
-  \label{lemma:at most one edge monotone}
-  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
-  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
-  and $\mathcal{M}$ be a containment-free perfect matching of $K_\pi$.
-  The following assertions hold:
-  \begin{enumerate}
-    \item
-    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
-    $\sigma_1$ is above $\sigma_2$ (see Fig.~\ref{subfig:increasing before
-    and above decreasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
-    \item
-    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
-    is below $\sigma_2$ (see Fig.~\ref{subfig:decreasing before and below
-    increasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
-  \end{enumerate}
-\end{lemma}
-
-\begin{proposition}
-  \label{proposition:hardness}
-  Deciding whether a permutation is a square is \NPC.
-\end{proposition}
-
-\begin{proof}[of Proposition~\ref{proposition:hardness}]
-  The problem is certainly in $\NPclass$. We propose a reduction from
-  the pattern involvement problem which is known to be
-  \NPC~\cite{Bose:Buss:Lubiw:1998}: Given two permutations $\pi$ and
-  $\sigma$, decide whether $\sigma$ occurs in $\pi$ (as an
-  order-isomorphic pattern).
-
-  \begin{figure}[t!]
-    \centering
-    \begin{tikzpicture}[
-      scale=.8,
-      >=stealth',
-      shorten >=1pt,
-      main node/.style={align=center},
-      cell/.style={draw,ultra thick,fill=black!5},
-      structure link/.style={line width=1.5pt},
-      pattern link/.style={line width=1.5pt,black!20},
-      monotone/.style={->,thick}
-      ]
-      % edges
-      % pi 1 - pi 2
-      \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
-      % pi 2 - sigma 2
-      \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
-      % 1
-      \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
-      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
-      % \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
-      % \node (claim edge 1) at (1.5,13) {Claim~\ref{claim:all (nu_1, nu'_1)-edges}};
-      % 2
-      \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
-      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
-      % \node (claim edge 2) at (4,4) {Claim~\ref{claim:(nu_2, nu'_2)-edge}};
-      % \node (claim edge 2) at (4,3.5) {Claim~\ref{claim:all (nu_2, nu'_2)-edges}};
-      % 3
-      \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
-      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
-      % \node (claim edge 1) at (5.5,11.5) {Claim~\ref{claim:(nu_3, nu'_3)-edge}};
-      % 4
-      \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
-      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
-      % \node (claim edge 1) at (8.5,8.5) {Claim~\ref{claim:(nu_4, nu'_4)-edge}};
-      % sigma 1 - pi 1
-      \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
-      % \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
-      % nodes
-      % 1
-      \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
-      \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_1$};
-      %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
-      \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
-      \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_1$};
-      % 2
-      \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
-      \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_2$};
-      \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
-      \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_2$};
-      % 3
-      \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
-      \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_3$};
-      \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
-      \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_3$};
-      % 4
-      \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
-      \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_4$};
-      \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
-      \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_4$};
-      % sigma 1
-      \draw [cell] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
-      \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
-      % pi 1
-      \draw [cell] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
-      \node [main node] (pi1) at (8.5,5.5) {$\pi'$};
-      % pi 2
-      \draw [cell] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
-      \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
-      % sigma 2
-      \draw [cell] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
-      \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};
-      % grid
-      \draw[step=1cm,black!30,thin,fill=black!10] (0,0) grid (12,12);
-      % row size
-      \foreach \y/\N in {0.5/N_2,1.5/N_2,2.5/N_4,3.5/k,4.5/n,5.5/n+2,6.5/k+2,7.5/N_4,8.5/N_3,9.5/N_3,10.5/N_1,11.4/N_1} {
-          \node at (-0.5,\y) (R\y) {$\N$};
-      }
-      % column size
-      \foreach \x/\N in
-      {0.5/N_1,1.5/N_2,2.5/N_1,3.5/N_3,4.5/k+2,5.5/N_4,6.5/N_2,7.5/N_3,8.5/n+2,9.5/N_4,10.5/n,11.5/k} {
-          \node at (\x,-0.5) (C\x) {$\N$};
-      }
-    \end{tikzpicture}
-    \caption{\label{fig:reduction}%
-      Schematic representation of the permutation $\mu$ used
-      in Proposition~\ref{proposition:hardness}.
-      Black edges denote the presence of at least one edge between two bunches of
-      vertices in $K_\mu$.
-      Grey edges denote edges that are only considered in the forward direction of the
-      proof.
-    }%
-  \end{figure}
-
-  Let $\pi \in S_n$ and $\sigma \in S_k$ be two arbitrary permutations.
-  Define
-  % \begin{align*}
-  % N_4 &= 2(2n + 2k + 4) + 1  = 4n + 4k + 9 \\
-  % N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45 \\
-  % N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225 \\
-  % N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325\text{.}
-  % \end{align*}
-  $N_4 = 2(2n + 2k + 4) + 1  = 4n + 4k + 9$,
-  $N_3 = 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45$,
-  $N_2 = 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225$ and
-  $N_1 = 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325$.
-  Notice that $N_1, N_2, N_3$ and $N_4$ are polynomial in $n$.
-  The crucial properties are that
-  (i) $N_1, N_2, N_3$ and $N_4$ are odd integers
-  and
-  (ii) $N_i > \left(\sum_{i < j \leq 4} 2N_j\right) + 2n + 2k + k$
-  for every $1 \leq i \leq 4$.
-
-  We now turn to defining various gadgets (sequences of integers)
-  that act as building blocks in our construction of a new permutation $\mu$:
-  \begin{align*}
-  \sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
-  \pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k] \\
-  \sigma'' &= \sigma \; [2N_2 + N_4] \\
-  \pi''    &= \pi \; [2N_2 + N_4 + k] \\
-  \nu_1    &= \nearrow_{N_1} \; [2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
-  \nu'_1   &= \nearrow_{N_1} \; [N_1 + 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
-  \nu_2    &= \nearrow_{N_2} \; [N_2] \\
-  \nu'_2   &= \searrow_{N_2} \\
-  \nu_3    &= \nearrow_{N_3} \; [2N_2 + 2N_4 + 2n + 2k + 4] \\
-  \nu'_3   &= \nearrow_{N_3} \; [2N_2 + N_3 + 2N_4 + 2n + 2k + 4] \\
-  \nu_4    &= \searrow_{N_4} \; [2N_2 + N_4 + 2n + 2k + 4] \\
-  \nu'_4   &= \searrow_{N_4} \; [2N_2]\text{.}
-  \end{align*}
-  We are now in position to define our target permutation $\mu$
-  (see Fig.~\ref{fig:reduction} for an illustration):
-  $$
-  \mu
-  =
-  \nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_2 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
-  \text{.}
-  $$
-
-  It is immediate that $\mu$ can be constructed in polynomial-time in $n$ and $k$.
-  % We claim that $\sigma$ occurs in $\pi$ if and only if
-  % there exists a containment-free perfect matching
-  % $\mathcal{M}$ in $K_\mu$ such that
-  % for any two disctinct edges
-  % $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  % we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
-  It can be shown (see Appendix) that
-  $\sigma$ occurs in $\pi$ if and only if
-  there exists a containment-free perfect matching
-  $\mathcal{M}$ in $K_\mu$ such that
-  for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
-  \qed
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% %%
-% %% ---- Patterns ----
-% %%
-% \section{Patterns}
-% \label{section:Patterns}
-%
-% \begin{proposition}
-%   \label{proposition:patterns:lower bound}
-%   Let $\pi \in S_{n}$.
-%   Then $\pi$ contains a pattern of size at least
-%   $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
-%   is the union of two order-isomorphic patterns.
-% \end{proposition}
-%
-% \begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
-%   The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935}
-%   states that every permutation of
-%   length at least $pq+1$ must contain either the pattern
-%   $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
-%   The result now follows from the fact that
-%   the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
-% \qed
-% \end{proof}
-%
-% \begin{proposition}
-%   \label{proposition:patterns:upper bound}
-%   If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
-%   a permutation of $S_{n}$
-%   that does not contains a pattern
-%   of length $2k$ that is the union of two
-%   order-isomorphic patterns.
-% \end{proposition}
-%
-% \begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
-% The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
-% Let $\pi \in S_n$ be chosen randomly and uniformly.
-% For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
-% of $\pi$,
-% let $X_\sigma$ be the indicator variable that
-% $\sigma$ is the union of two order-ismorphic patterns.
-% According to Boole's inequality, we have
-% $P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
-% Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
-% the probability that $\pi$ contains a pattern of length $2k$
-% is the union of two order-ismorphic patterns is at most
-% $\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
-% Thus, with positive probability, no event $X_\sigma$ occurs and
-% there is a permutation of $S_n$ that does not contain a pattern
-% of length $2k$ that is the union of two order-isomorphic patterns.
-% \qed
-% \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% ---- Conclusion ----
-%%
-\section{Conclusion}
-\label{section:Conclusion}
-
-There are a number of further directions of investigation in this
-general subject. They cover several areas: algorithmic, combinatorics,
-and algebra. Let us mention several -~not necessarily new~- open
-problems that are, in our opinion, the most interesting. How many
-permutations of $S_{2n}$ are squares? How many $(213,231)$-avoiding
-permutations of $S_{2n}$ are squares? (Equivalently, by
-Proposition~\ref{prop:bijection_binary_to_permutations_squares},
-how many binary strings of length $2n$ are squares; Problem~4
-in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of
-deciding whether a $(213,231)$-avoiding permutation is a square
-(Problem~4 in \cite{Henshall:Rampersad:Shallit:2011},
-see also \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013})?
-Given two permutations $\pi$ and~$\sigma$, how hard is the problem of
-deciding whether $\sigma$ is a square root of~$\pi$?
-Moreover, one can ask for a complete algebraic study of
-$\QQ[S]$ as a graded associative algebra for the shuffle  product
-$\SHUFFLE$. Describing a generating family for $\QQ[S]$, defining
-multiplicative bases of $\QQ[S]$, and determining whether $\QQ[S]$ is
-free as an associative algebra are worthwhile questions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Bibliography
-%%
-
-\bibliographystyle{plain}
-\bibliography{biblio}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
 \section*{Appendix (Reviewers version only)}
@@
 \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-12-13 submission.tex
Supp. annexes + ajouts demos bijection avec carres binaires

@@
     \end{claim}
+    \begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
+        Let us first show that the image of $\BINTOPERM$ contains only
+        permutations avoiding $213$ and $231$. Let $u$ be a binary word,
+        $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
+        positions of the occurrences of $0$ (resp. $1$) in $u$. By
+        definition of $\BINTOPERM$, from left to right, the subword
+        $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
+        is decreasing, and all letters of $w$ are greater than those
+        of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
+        Then, since $v$ is increasing and $w$ is decreasing, there is an
+        occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
+        occurrence of $21$ (resp. $2$, $1$). All these three cases
+        contradict the fact that all letters of $w$ are greater than
+        those of $v$. A similar argument shows that $\pi$ avoids~$231$
+        as well.
+        \smallskip
+
+        Finally, observe that any permutation $\pi$ avoiding $213$ and
+        $231$ necessarily starts by the smallest possible letter or the
+        greatest possible letter. This property is then true for the
+        suffix of $\pi$ obtained by deleting its first letter,
+        and so on for all of its suffixes. Thus, by
+        replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
+        role of a smallest (resp. greatest) letter, one obtains a binary
+        word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
+        avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
+        \qed
+    \end{proof}
 
@@
     \end{claim}
+    \begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
+        Since $u$ is a square binary word, there is a binary word $v$
+        such that $u \in v \shuffle v$. Then, there are two disjoint
+        sets $P$ and $Q$ of positions of letters of $u$ such that
+        $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
+        words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
+        same standardized $\sigma$. Hence, and by definition of
+        the shuffle product of permutations, $\BINTOPERM(u)$ appears in
+        $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
+        square permutation.
+        \qed
+    \end{proof}
 
@@
     \end{claim}
+    \begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
+        Let $\pi$ be a square permutation avoiding $213$ and $231$. By
+        Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
+        the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
+        well-defined binary word. Since $\pi$ is a square permutation,
+        there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
+        of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
+        order-isomorphic. This implies, by the definitions of $\BINTOPERM$
+        and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
+        is a square binary word.
+        \qed
+    \end{proof}
     \qed
@@
 
-\newpage
-\section*{Appendix (Reviewers version only)}
-
-\begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
-    Let us first show that the image of $\BINTOPERM$ contains only
-    permutations avoiding $213$ and $231$. Let $u$ be a binary word,
-    $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
-    positions of the occurrences of $0$ (resp. $1$) in $u$. By
-    definition of $\BINTOPERM$, from left to right, the subword
-    $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
-    is decreasing, and all letters of $w$ are greater than those
-    of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
-    Then, since $v$ is increasing and $w$ is decreasing, there is an
-    occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
-    occurrence of $21$ (resp. $2$, $1$). All these three cases
-    contradict the fact that all letters of $w$ are greater than
-    those of $v$. A similar argument shows that $\pi$ avoids~$231$
-    as well.
-    \smallskip
-
-    Finally, observe that any permutation $\pi$ avoiding $213$ and
-    $231$ necessarily starts by the smallest possible letter or the
-    greatest possible letter. This property is then true for the
-    suffix of $\pi$ obtained by deleting its first letter,
-    and so on for all of its suffixes. Thus, by
-    replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
-    role of a smallest (resp. greatest) letter, one obtains a binary
-    word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
-    avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
-    Since $u$ is a square binary word, there is a binary word $v$
-    such that $u \in v \shuffle v$. Then, there are two disjoint
-    sets $P$ and $Q$ of positions of letters of $u$ such that
-    $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
-    words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
-    same standardized $\sigma$. Hence, and by definition of
-    the shuffle product of permutations, $\BINTOPERM(u)$ appears in
-    $\sigma \SHUFFLE \sigma$, showing that $\BINTOPERM(u)$ is a
-    square permutation.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
-    Let $\pi$ be a square permutation avoiding $213$ and $231$. By
-    Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
-    the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
-    well-defined binary word. Since $\pi$ is a square permutation,
-    there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
-    of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
-    order-isomorphic. This implies, by the definitions of $\BINTOPERM$
-    and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
-    is a square binary word.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lem:endomorphisms}]
-    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
-    associative algebras, we have to prove that for all permutations
-    $\sigma$ and $\nu$,
-    \begin{equation}
-        \phi_j(\sigma \SHUFFLE \nu) =
-        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
-    \end{equation}
-    We shall proceed here as we have done during the proof of
-    Proposition~\ref{prop:shuffle_associative_commutative} by showing
-    that $\phi_j$ is a morphism of coalgebras, that is
-    \begin{equation}
-        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
-    \end{equation}
-    In the sequel, $\pi$ is a permutation.
-    \smallskip
-
-    If $P$ is a set of indexes of letters of $\pi$, we denote by
-    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
-    the operation $\widetilde{\,}$ defines a bijection on the set of the
-    subsets of $[|\pi|]$, and since the standardization operation
-    commutes with the mirror operation on words without multiple
-    occurrence of a letter, we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_1(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_1(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|P_1}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
-        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \widetilde{\STD\left(\pi_{|P_1}\right)}
-        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
-    $\phi_1$ is a morphism of associative algebras.
-    \smallskip
-
-    Next, since by definition of the complementation operation on
-    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
-    we have $\tau(i) < \tau(k)$ if and only if $\bar \tau(i) > \bar \tau(k)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_2(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_2(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\bar \pi_{|P_1}\right)
-        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
-    $\phi_2$ is a morphism of associative algebras.
-    \smallskip
-
-    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
-    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
-    $\{\tau(i) : i \in P\}$. Since the map sending a subset $P$ of
-    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
-    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
-    we have
-    \begin{equation} \begin{split}
-        \Delta(\phi_3(\pi))
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\phi_3(\pi)_{|P_1}\right)
-        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|P_1}\right)
-        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
-        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \STD\left(\pi_{|P_1}\right)^{-1}
-        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
-        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
-        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
-        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
-        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
-    \end{split} \end{equation}
-    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
-    $\phi_3$ is a morphism of associative algebras.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lemma:matching}]
-    The proof is, naturally enough, a slight extension of
-    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
-    for the sake of completeness.
-
-    (\emph{1} $\Rightarrow$ \emph{2}).
-    Since $\pi$ is a square, it admits a square root $\sigma$.
-    Let $2n = |\pi|$ (and hence
-    $|\sigma| = n$).
-    By definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
-    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
-    $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
-    order-isomorphic to $\sigma$.
-    Let $K_\pi$ be the linear clique associated with $\pi$ with vertices
-    $\VS{K_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
-    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
-    is a subset of $\ES{K_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
-    matching since $I^1 \cap I^2 = \emptyset$, and
-    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
-    Indeed, if it is not the case then there would exist two edges
-    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
-    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
-    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
-    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
-    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
-    we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
-    $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
-    elements (at position $j$ and $k$) in two patterns that are
-    order-isomorphic to $\sigma$.
-
-    (\emph{2} $\Rightarrow$ \emph{1}).
-    Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$, and
-    $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$.
-    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
-    endpoints of the edges in $\mathcal{M}$ assuming the positions
-    $1, 2, \dots, 2$ are drawn along an horizontal line.
-    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
-    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
-    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
-    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
-    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
-    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
-    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
-    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
-    contradiction, that this is not the case. Therefore, there exist two
-    positions $j$ and $k$, say $j < k$, such that
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
-    But since $\mathcal{M}$ is containment-free, it follows that both
-    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
-    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
-    since for any two distinct edges $(a, b)$ and $(a', b')$ in
-    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
-    $\pi(b) < \pi(b')$.
-    \qed
-\end{proof}
-\bigskip
-
-\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
-  We first prove \emph{1}. Suppose, aiming at a contradiction, that
-  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
-  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
-  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
-  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
-  $\sigma_2$ is decreasing), which contradicts our hypothesis about
-  $\mathcal{M}$.
-
-  As for \emph{2}, suppose, still aiming at a contradiction, that there
-  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
-  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
-  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
-  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
-  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
-  \qed
-\end{proof}
-
-\bigskip
-
-\begin{proof}[of Proposition~\ref{proposition:hardness}]
-  Suppose first that $\sigma$ occurs in $\pi$ and fix an occurrence.
-  Construct a matching $\mathcal{M}$ as follows:
-  \begin{itemize}
-    \item $\mathcal{M}$ contains $N_1$ pairwise crossing
-    $(\nu_1, \nu'_1)$-edges.
-    \item $\mathcal{M}$ contains $N_2$ pairwise crossing
-    $(\nu_2, \nu'_2)$-edges.
-    \item $\mathcal{M}$ contains $N_3$ pairwise crossing
-    $(\nu_3, \nu'_3)$-edges.
-    \item $\mathcal{M}$ contains $N_4$ pairwise crossing
-    $(\nu_4, \nu'_4)$-edges.
-    \item $\mathcal{M}$ contains $k+2$ pairwise crossing
-    $(\sigma', \pi')$-edges as depicted in
-    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
-    More precisely,
-    (i) the first position of $\sigma'$
-    (\emph{i.e.}, $(2N_1+N_2+N_3) + 1$) is linked
-    to the first position of $\pi'$
-    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + 1$),
-    (ii) the last position of $\sigma'$
-    (\emph{i.e.}, $(2N_1+N_2+N_3) + k+2$) is linked
-    to the last position of $\pi'$
-    (\emph{i.e.}, $(2N_1 + 2N_2 + 2N_3 + N_4 + k + 2) + n+2$),
-    and all other positions in $\sigma'$ are linked by means of $k$ pairwise
-    crossing edges to the positions in
-    $\pi'$ that correspond to the fixed occurence of $\sigma$ in $\pi$.
-    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
-    \item $\mathcal{M}$ contains $n-k$ pairwise crossing
-    $(\pi', \pi'')$-edges as depicted in
-    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
-    More precisely,
-    all positions in $\pi'$ that do not correspond to the fixed occurence of
-    $\sigma$ in $\pi$ are linked by means of $n-k$ pairwise crossing edges
-    to the positions in $\pi''$ that do not correspond to the fixed
-    occurrence of $\sigma$ in $\pi$.
-    \item $\mathcal{M}$ contains $k$ pairwise crossing
-    $(\pi'', \sigma'')$-edges as depicted in
-    Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
-    More precisely, the positions in $\pi''$ that correspond to
-    the fixed occurrence of $\sigma$ in $\pi$ are linked
-    by means of $k$ pairwise crossing edges to all positions in
-    $\sigma''$.
-    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
-  \end{itemize}
-  It can be easily checked (probably referring to Figure~\ref{fig:reduction}) that
-  $\mathcal{M}$ is perfect, containment-free and that
-  for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
-
-    \begin{figure}[t!]
-      \centering
-        \begin{tikzpicture}
-          [
-            scale=0.35,
-            label/.style={anchor=base},
-            cell/.style={draw,thick,fill=black!5}
-          ]
-          % sigma'
-          \draw [cell,rounded corners] (1.5,16.5) -- (4.5,16.5) -- (4.5,19.5) -- (1.5,19.5) -- cycle;
-          \draw[step=1cm,black!60,ultra thin,fill=black!10] (0.8,16.8) grid (5.2,21.2);
-          \foreach \x/\y in {1/20,2/19,3/17,4/18,5/21} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          \node (pip) at (3,22) {$\sigma'$};
-          \node[%
-            draw,
-            text width=5.5cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,-1cm)}
-          ] at (3,26)
-          {%
-            $((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$
-          };%
-          %\node at (3,-0.5) {$((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$};
-          \node at (-0.5,20) {$k+1$};
-          \node at (-0.5,21) {$k+2$};
-          % pi'
-          \draw [cell,rounded corners] (6.5,8.5) -- (12.5,8.5) -- (12.5,14.5) -- (6.5,14.5) -- cycle;
-          \draw[step=1cm,black!60,ultra thin,fill=black!10] (5.8,8.8) grid (13.2,16.2);
-          \foreach \x/\y in {6/15,7/12,8/13,9/10,10/9,11/11,12/14,13/16} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          \node (pip) at (9.5,17) {$\pi'$};
-          \node[%
-            draw,
-            text width=5.5cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,3.5cm)}
-          ] at (9.5,-3.5)
-          {%
-            $((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$
-          };%
-          %\node at (9.5,-2) {$((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$};
-          \node at (-0.5,15) {$n+1$};
-          \node at (-0.5,16) {$n+2$};
-          % pi''
-          \draw [cell,rounded corners] (13.5,2.5) -- (19.5,2.5) -- (19.5,8.5) -- (13.5,8.5) -- cycle;
-          \draw[step=1cm,black!60,ultra thin,fill=black!10] (13.8,2.8) grid (19.2,8.2);
-          \foreach \x/\y in {14/6,15/7,16/4,17/3,18/5,19/8} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          \node (pipp) at (16.75,9) {$\pi''$};
-          \node[%
-            draw,
-            text width=2.25cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,-3cm)}
-          ] at (16.5,19)
-          {%
-            $\pi [2N_2 + N_4 + k]$
-          };%
-          %\node at (16.5,-3.5) {$\pi [2N_2 + N_4 + k]$};
-          % sigma''
-          \draw [cell,rounded corners] (19.5,-0.5) -- (22.5,-.5) -- (22.5,2.5) -- (19.5,2.5) -- cycle;
-          \draw[step=1cm,black!60,ultra thin,fill=black!10] (19.8,-0.2) grid (22.2,2.2);
-          \foreach \x/\y in {20/2,21/0,22/1} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          \node (sigmapp) at (21,3) {$\sigma''$};
-          \node[%
-            draw,
-            text width=1.75cm,text justified,color=black!50,rounded corners,
-            rectangle callout,callout relative pointer={(0.0cm,0.75cm)}
-          ] at (21,-4.5)
-          {%
-            $\sigma [2N_2 + N_4]$
-          };%
-          %\node at (21,-5) {$\sigma [2N_2 + N_4]$};
-          % edges sigma' - pi'
-          \draw [line width=1pt]
-          (1,20) .. controls +(0,9) and +(0,4) .. (6,15);
-          \draw [line width=1pt]
-          (5,21) .. controls +(0,9) and +(0,4) .. (13,16);
-          \draw [line width=1pt]
-          (2,19) .. controls +(0,-17) and +(0,-7) .. (8,13);
-          \draw [line width=1pt]
-          (3,17) .. controls +(0,-17) and +(0,-7) .. (9,10);
-          \draw [line width=1pt]
-          (4,18) .. controls +(0,-17) and +(0,-7) .. (11,11);
-          % edges pi' - pi''
-          \draw [line width=1pt]
-          (7,12) .. controls +(0,9) and +(0,4) .. (14,6);
-          \draw [line width=1pt]
-          (10,9) .. controls +(0,9) and +(0,4) .. (17,3);
-          \draw [line width=1pt]
-          (12,14) .. controls +(0,9) and +(0,4) .. (19,8);
-          % edges pi'' - sigma''
-          \draw [line width=1pt]
-          (15,7) .. controls +(0,-9) and +(0,-4) .. (20,2);
-          \draw [line width=1pt]
-          (16,4) .. controls +(0,-9) and +(0,-4) .. (21,0);
-          \draw [line width=1pt]
-          (18,5) .. controls +(0,-9) and +(0,-4) .. (22,1);
-          % text
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(0.5cm,-.75cm)}
-          % ] at (1,27)
-          % {%
-          % $2$ pairwise crossing $(\sigma',\pi')$-edges that
-          % link the guards
-          % };%
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(-0.1cm,-2.5cm)}
-          % ] at (17,24)
-          % {%
-          % $n-k$ pairwise crossing $(\sigma',\pi')$-edges that
-          % denote positions that do not correspond to
-          % a specific occurence of $\sigma$ in $\pi$
-          % (at positions $2$, $3$ and $5$).
-          % };%
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(0.75cm,1.5cm)}
-          % ] at (2.5,4,5)
-          % {%
-          % $k$ pairwise crossing $(\sigma',\pi')$-edges that
-          % denote a specific occurence of $\sigma$ in $\pi$
-          % (at positions $2$, $3$ and $5$).
-          % };%
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(2cm,.75cm)}
-          % ] at (7,-3)
-          % {%
-          % $k$ pairwise crossing $(\pi''; \sigma'')$-edges that
-          % denote a specific occurence of $\sigma$ in $\pi$
-          % (at positions $2$, $3$ and $5$).
-          % };%
-        \end{tikzpicture}
-        \caption{\label{fig:subfig:sigma' - pi' - pi'' - sigma''}%
-        Illustration of the containment-free perfect matching $\mathcal{M}$
-        between gadgets $\sigma'$, $\pi'$, $\pi''$ and $\sigma''$ assuming two
-        input permutation $\sigma = 312$ and $\pi = 4\mathbf{5}\mathbf{2}1\mathbf{3}6$
-        (where a specific occurrence of $\sigma$ in $\pi$ is depicted in bold).
-        }
-      \end{figure}
-
-  Conversely, suppose that there exists a containment-free perfect matching
-  $\mathcal{M}$ in $K_\mu$ such that for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
-  We show that $\sigma$ occurs as a pattern in $\pi$.
-  Whereas the matching $\mathcal{M}$ may not be as regular as in the forward
-  direction, the main idea is to prove that $\mathcal{M}$ contains enough structure
-  (more precisely, $k+2$ $(\sigma', \pi')$-edges) so that we can conclude
-  that $\sigma$ occurs in $\pi$.
-  We have divided the reverse direction into a serie of basic claims that
-  progressively defines the overall structure of $\mathcal{M}$
-  (refer to Fig.~\ref{fig:reduction}).
-
-    \begin{figure}[t!]
-      \centering
-      \subfigure[%
-        A $(\nu_1, \nu_2)$-edge $(a, b)$ together with a
-        $(\nu_2, \nu'_1)$-edge $(x, y)$.
-        Notice that we may have $x<b$
-        (as long as $N_1 < x \leq N_1+N_2$ and $N_1 < b \leq N_1+N_2$
-        do hold).
-      ]{%
-        \begin{tikzpicture}
-          [
-            scale=0.25,
-            label/.style={anchor=base}
-          ]
-          % nu_1
-          \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
-          \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          % nu_4
-          \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
-          \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          % nu'_1
-          \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
-          \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
-              \draw [fill=black] (\x,\y) circle (0.1);
-          }
-          % labels
-          \node [label] (y) at (17,14.5) {$b$};
-          \node [label] (yp) at (10,-1.5) {$b'$};
-          \node [label] (a) at (0,8.5) {$a$};
-          \node [label] (ap) at (8,-1.5) {$a'$};
-          \node (nu1) at (2.5,7.7) {$\nu_1$};
-          \node (nu2) at (8.5,-2.5) {$\nu_2$};
-          \node (nu1') at (14.5,13.5) {$\nu'_1$};
-          % crossings edges
-          \draw [line width=1pt]
-          (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
-          \draw [line width=1pt]
-          (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
-          \draw [line width=1pt]
-          (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
-          \draw [line width=1pt]
-          (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
-          \draw [line width=1pt]
-          (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
-          % data edges
-          \draw [line width=1pt]
-          (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
-          \draw [line width=1pt]
-          (10,1) .. controls +(0,32) and +(0,10) .. (17,21);
-        \end{tikzpicture}
-        \label{subfig:no (nu_1, nu_2)-edge - 1}
-      }% end subfigure
-      \qquad
-      \subfigure[%
-        A $(\nu_1, \nu'_1)$-edge $(x', y')$ together with an
-        edge $(y, z)$ which is neither
-        a $(\nu_1, \nu'_1)$-edge,
-        nor a $(\nu_2, \nu'_1)$-edge,
-        nor a $(\nu'_1, \nu'_1)$-edge.
-        Notice that $y$ may be any position satisfying
-        $N_1+N_2 < y \leq 2N_1+N_2$.
-      ]{%
-      \begin{tikzpicture}
-        [
-          scale=0.25,
-          label/.style={anchor=base}
-        ]
-        % nu_1
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
-        \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu_4
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
-        \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu'_1
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
-        \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % labels
-        \node [label] (y) at (17,14.5) {$b$};
-        \node [label] (ypp) at (23,-4.5) {$b''$};
-        \node [label] (a) at (0,8.5) {$a$};
-        \node [label] (a) at (2,8.5) {$x$};
-        \node [label] (a) at (14,14.5) {$x'$};
-        \node [label] (ap) at (8,-1.5) {$a'$};
-        \node (nu1) at (2.5,7.7) {$\nu_1$};
-        \node (nu2) at (8.5,-2.5) {$\nu_2$};
-        \node (nu1') at (14.5,13.5) {$\nu'_1$};
-        % crossings edges
-        \draw [line width=1pt]
-        (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
-        \draw [line width=1pt]
-        (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
-        \draw [line width=1pt]
-        (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
-        \draw [line width=1pt]
-        (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
-        \draw [line width=1pt]
-        (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
-        % data edges
-        \draw [line width=1pt]
-        (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
-        \draw [line width=1pt]
-        (17,21) .. controls +(0,10) and +(0,35) .. (23,-3);
-        \end{tikzpicture}
-        \label{subfig:no (nu_1, nu_2)-edge - 2}
-      }% end subfigure
-      \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
-        Claim~\ref{claim:no (nu_1, nu_2)-edge}:
-        Assuming a $(\nu_1,\nu_2)$-edge $(a, b)$ in $\mathcal{M}$.
-      }
-    \end{figure}
-
-  \begin{claim}
-    \label{claim:no (nu_1, nu_2)-edge}
-    There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_2)-edge}]
-    First, according to Lemma~\ref{lemma:at most one edge monotone},
-    there exists at most one $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
-    Suppose now, aiming at a contradiction, that there exists
-    exactly one $(\nu_1, \nu_2)$-edge $e = (a, a')$ in $\mathcal{M}$.
-    Under this assumption, we show that $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
-    Indeed, suppose, aiming at a contradiction,
-    that there exists a $(\nu_1, \nu_1)$-edge $(b, b')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free,
-    we must have $b < a$.
-    Then it follows that $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
-    $\mu(b') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    Hence $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
-    Furthermore, since $\nu_1$ is increasing and
-    there exists only one $(\nu_1, \nu_2)$-edge $(a, a')$ in $\mathcal{M}$,
-    we conclude that
-    $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
-    But $|\nu_1| = |\nu'_1| = N_1$ and $\mathcal{M}$ is perfect,
-    and hence there exists a position, say $b$, in $\nu'_1$ that is
-    not involved in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-    We rule out this configuration by considering two cases:
-    \begin{itemize}
-      \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-edge,
-      say $(b', b)$.
-      Then, we have $\mu(a) > \mu(b')$ (since $\nu_1$ is above $\nu_2$)
-      and $\mu(a') < \mu(b)$ (since $\nu_2$ is below $\nu'_1$),
-      which contradicts our hypothesis about $\mathcal{M}$.
-      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
-      \item Position $b$ is not involved in a $(\nu_2, \nu'_1)$-edge.
-      Since $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges
-      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge
-      $(b, b'')$ in $\mathcal{M}$
-      that is neither a $(\nu_1, \nu'_1)$-edge nor a $(\nu_2, \nu'_1)$-edge
-      nor a $(\nu'_1, \nu'_1)$-edge.
-      In other words, $y'' > 2N_1 + N_2$.
-      Now, for any $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$, we
-      have $\mu(x) < \mu(b)$ (since $\nu_1$ is below $\nu'_1$)
-      and $\mu(x') > \mu(b'')$ (since $\nu'_1$ is above any other gadget),
-      which contradicts our hypothesis about $\mathcal{M}$.
-      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 2}.
-    \end{itemize}
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:(nu_1, nu'_1)-edge}
-    There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:(nu_1, nu'_1)-edge}]
-    Suppose, aiming at a contradiction, that there is no
-    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-    Then it follows that there exists an edge $(x, x')$ in $\mathcal{M}$,
-    $1 \leq x \leq N_1$, that is neither a
-    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
-    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
-    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
-    (In other words, $x \leq N_1$ and $x' > 2N_1 + N_2$.)
-    Therefore, since $\mathcal{M}$ is containment-free,
-    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
-    in $\mathcal{M}$.
-    Then it follows that every position $y$ of $\nu'_1$
-    (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
-    is involved in an edge $(y, y')$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
-    (otherwise $\mathcal{M}$ would not be containment-free).
-    But $|\nu'_1| = N_1 = 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1
-    > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
-    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
-    Then it follows that $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  The above claim will be complement in upcoming Claim~\ref{claim:all (nu_1, nu'_1)-edges}.
-
-  \begin{claim}
-    \label{claim:no (nu_2, nu_2)-edge}
-    There is no $(\nu_2, \nu_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_2)-edge}]
-  Combine Claim~\ref{claim:(nu_1, nu'_1)-edge} with
-  the fact that $\mathcal{M}$ is containment-free.
-  \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_2, nu'_1)-edge}
-    There is no $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_1)-edge}]
-    First, according to Lemma~\ref{lemma:at most one edge monotone},
-    there exists at most one $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
-    Suppose now, aiming at a contradiction, that there exists
-    exactly one $(\nu_2, \nu'_1)$-edge $e = (a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_1, nu'_1)-edge}, there exists at least
-    one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-    Furtheremore, since
-    there is no $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge}),
-    every position in $\nu_1$ is involved either in a
-    $(\nu_1, \nu_1)$-edge or in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$
-    (otherwise $\mathcal{M}$ would not be containment-free).
-    Then it follows that that there is an odd number of $(\nu_1, \nu'_1)$-edges
-    in $\mathcal{M}$ since $\mathcal{M}$ is perfect and $|\nu_1| = N_1$ is odd.
-    Combining this with the fact that there exists exactly one $(\nu_2, \nu'_1)$-edge
-    in $\mathcal{M}$ and $|\nu'_1| = N_1$ is odd,
-    we conclude that there exists a position, say $x$, in $\nu'_1$
-    that is neither involved in a $(\nu_1, \nu'_1)$-edge nor in a $(\nu_1, \nu'_1)$-edge
-    nor in a $(\nu'_1, \nu'_1)$-edge.
-    Write $(x, x')$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $x' > 2N_1 + N_2$, this edge.
-    We have $\mu(a) < \mu(x)$ (since $\nu_2$ is below $\nu'_1$)
-    and $\mu(a') > \mu(x')$ (since $\nu'_1$ is above any other gadget),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:(nu_2, nu'_2)-edge}
-    There is at least one $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:(nu_2, nu'_2)-edge}]
-    Suppose, aiming at a contradiction that there is no
-    $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
-    Notice that there is neither
-    a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
-    nor a $(\nu_2, \nu_2)$-edge (Claim~\ref{claim:no (nu_2, nu_2)-edge})
-    nor a $(\nu_2, \nu'_1)$-edge (Claim~\ref{claim:no (nu_2, nu'_1)-edge})
-    in $\mathcal{M}$.
-    But $|\nu_2| = N_2 > 2N_3 + 2N_4 + 2n + 2k +4$.
-    Hence $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no edge below (nu_2, nu'_2)-edge}
-    There is
-    neither a $(\nu'_1, \nu'_1)$-edge,
-    nor a $(\nu'_1, \nu_3)$-edge,
-    nor a $(\nu'_1, \sigma')$-edge,
-    nor a $(\nu'_1, \nu_4)$-edge
-    nor a $(\nu_3, \nu_3)$-edge,
-    nor a $(\nu_3, \sigma')$-edge,
-    nor a $(\nu_3, \nu_4)$-edge,
-    nor a $(\sigma', \sigma')$-edge,
-    nor a $(\sigma', \nu_4)$-edge,
-    nor a $(\nu_4, \nu_4)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
-    Combine Claim~\ref{claim:(nu_2, nu'_2)-edge} with the fact that
-    $\mathcal{M}$ is containment-free.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim: no edge right of nu'_2}
-    There is
-    neither a $(\nu'_2, \nu'_3)$-edge,
-    nor a $(\nu'_2, \pi')$-edge,
-    nor a $(\nu'_2, \nu'_4)$-edge,
-    nor a $(\nu'_2, \pi'')$-edge
-    nor a $(\nu'_2, \sigma'')$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim: no edge right of nu'_2}]
-    Suppose, aiming at a contradiction, that
-    $\mathcal{M}$ contains an edge $(a, a')$ which is
-    either a $(\nu'_2, \nu'_3)$-edge
-    or a $(\nu'_2, \pi')$-edge
-    or a $(\nu'_2, \nu'_4)$-edge
-    or a $(\nu'_2, \pi'')$-edge
-    or a $(\nu'_2, \sigma'')$-edge.
-    According to Claim~\ref{claim:(nu_2, nu'_2)-edge},
-    there exists a $(\nu_2, \nu'_2)$-edge $(b, b')$ in $\mathcal{M}$.
-    Then it follows that we have
-    $\mu(b) > \mu(a)$ (since $\nu_2$ is above $\nu'_2$) and
-    $\mu(b') < \mu(a')$ (since $\nu'_3$, $\pi'$, $\nu'_4$, $\pi''$ and
-    $\sigma''$ are all above $\nu'_2$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  % \begin{claim}
-  %   \label{claim:no (nu_1, nu'_2)-edge}
-  %   There is no $(\nu_1, \nu'_2)$-edge in $\mathcal{M}$.
-  % \end{claim}
-  %
-  % \begin{proof}[of Claim~\ref{claim:no (nu_1, nu'_2)-edge}]
-  %   Suppose, aiming at a contradiction that there is a
-  %   $(\nu_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-  %   Since $\mathcal{M}$ is containment-free, we must have $b < a$.
-  %   Then it follows that we have
-  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
-  %   $\mu(b') > \mu(a')$ (since $\nu'_2$ is above $\nu'_2$),
-  %   which contradicts our hypothesis about $\mathcal{M}$.
-  %   \qed
-  % \end{proof}
-  %
-  % \begin{claim}
-  %   \label{claim:no (nu'_1, nu'_2)-edge}
-  %   There is no $(\nu'_1, \nu'_2)$-edge in $\mathcal{M}$.
-  % \end{claim}
-  %
-  % \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_2)-edge}]
-  %   Suppose, aiming at a contradiction that there is a
-  %   $(\nu'_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-  %   Then it follows that we have
-  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
-  %   $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu'_2$),
-  %   which contradicts our hypothesis about $\mathcal{M}$.
-  %   \qed
-  % \end{proof}
-
-  \begin{figure}[t!]
-    \centering
-      \begin{tikzpicture}
-        [
-          scale=0.35,
-          label/.style={anchor=base}
-        ]
-        % nu_2
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
-        \foreach \x/\y in {0/11,1/10,2/9,3/8,4/7,5/6} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu_3
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,15.8) grid (11.2,21.2);
-        \foreach \x/\y in {6/16,7/17,8/18,9/19,10/20,11/21} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % nu'_2
-        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,-0.2) grid (17.2,5.2);
-
-        \foreach \x/\y in {12/5,13/4,14/3,15/2,16/1,17/0} {
-            \draw [fill=black] (\x,\y) circle (0.1);
-        }
-        % labels
-        \node [label] (x) at (2,4.5) {$b$};
-        \node [label] (a) at (8,14.5) {$a$};
-        \node [label] (xp) at (14,-1.5) {$b'$};
-        \node [label] (ap) at (16,-1.5) {$a'$};
-        \node (nu1) at (2.5,3.5) {$\nu_2$};
-        \node (nu2) at (8.5,13.5) {$\nu_3$};
-        \node (nu1') at (14.5,-2.5) {$\nu'_2$};
-        % crossings edges
-        \draw [line width=1pt]
-        (2,9) .. controls +(0,25) and +(0,30) .. (14,3);
-        \draw [line width=1pt]
-        (8,18) .. controls +(0,15) and +(0,20) .. (16,1);
-      \end{tikzpicture}
-    \caption{\label{fig:subfig:no (nu_3, nu'_2)-edge}%
-      Claim~\ref{claim:no (nu_3, nu'_2)-edge}:
-      Assuming a $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
-    }
-  \end{figure}
-
-  \begin{claim}
-    \label{claim:no (nu_3, nu'_2)-edge}
-    There is no $(\nu_3, \nu'_2)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
-    Suppose, aiming at a contradiction, that there exists
-    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_2, nu'_2)-edge}, there
-    exists at least one $(\nu_2, \nu'_2)$-edge $(b, b')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free (Lemma~\ref{lemma:matching}),
-    we have $b' < a'$.
-    Therefore,
-    we have $\mu(b) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
-    $\mu(b') > \mu(a')$ (since $\nu'_2$ is decreasing),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:at most one (nu_2, nu_3)-edge}
-    There is at most one $(\nu_2, \nu_3)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
-    Apply Lemma~\ref{lemma:at most one edge monotone}.
-    \qed
-  \end{proof}
-
-    We will see soon (upcoming Claim~\ref{claim:all (nu_2, nu'_2)-edges})
-    that there exists actually no $(\nu_2, \nu_3)$-edge
-    in $\mathcal{M}$.
-
-  \begin{claim}
-    \label{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}
-    There is neither a $(\nu_1, \nu_3)$-edge nor a $(\nu'_1, \nu_3)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}]
-    We first prove that there is no $(\nu_1, \nu_3)$-edge in $\mathcal{M}$.
-    Indeed, suppose, aiming at a contradiction, that these exists
-    a $(\nu_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free, we must have
-    $b < a'$.
-    Then it follows that
-    $\mu(b) < \mu(a)$ (since $\nu_1$ is incresing) and
-    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-
-    As for no $(\nu_1, \nu_3)$-edges in $\mathcal{M}$, suppose,
-    aiming at a contradiction, that these exists
-    a $(\nu'_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
-    Then it follows that
-    $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
-    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:(nu_3, nu'_3)-edge}
-    There exists a $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:(nu_3, nu'_3)-edge}]
-    Suppose, aiming at a contradiction, that there is no
-    $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
-    Combining
-    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
-    Claim~\ref{claim:no (nu_3, nu'_2)-edge},
-    Claim~\ref{claim:at most one (nu_2, nu_3)-edge}
-    and Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge},
-    we conclude that $N_3-1$ positions in $\nu_3$ are involved
-    in edges of $\mathcal{M}$ that are
-    neither ($\nu_1, \nu_3)$-edges,
-    nor $(\nu_2, \nu_3)$-edges
-    nor $(\nu'_1, \nu_3)$-edges
-    nor $(\nu_3, \nu_3)$-edges.
-    But $N_3 - 1 > 2N_4 + 2n + 2k + 4$ (the positions left) and hence
-    $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no edge below one (nu_3, nu'_3)-edge}
-    There is
-    neither a $(\sigma', \sigma')$-edge,
-    nor a $(\sigma', \nu_4)$-edge,
-    nor a $(\sigma', \nu'_2)$-edge,
-    nor a $(\nu_4, \nu_4)$-edge
-    nor a $(\nu_4, \nu'_2)$-edge,
-    nor a $(\nu'_2, \nu'_2)$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge}]
-    Combine Claim~\ref{claim:(nu_3, nu'_3)-edge} with the fact that
-    $\mathcal{M}$ is containment-free.
-    \qed
-  \end{proof}
-
-  The next two claims state that $\mathcal{M}$ actually contains
-  all $(\nu_1, \nu'_1)$-edges and all $(\nu_2, \nu'_2)$-edges.
-
-  \begin{claim}
-    \label{claim:all (nu_1, nu'_1)-edges}
-    $\mathcal{M}$ contains $N_1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:all (nu_1, nu'_1)-edges}]
-    Suppose, aiming at a contradiction, that
-    $\mathcal{M}$ does not contain $N_1$ $(\nu_1, \nu'_1)$-edge.
-    Combing Claim~\ref{claim:no (nu_2, nu'_1)-edge} and
-    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
-    we conclude that there exists a position $a$ in $\nu'_1$ that is
-    involded in an edge $(a, a')$ in $\mathcal{M}$ that is
-    neither a $(\nu_1, \nu'_1)$-edge,
-    nor a $(\nu_2, \nu'_1)$-edge,
-    nor a $(\nu'_1, \nu'_1)$-edge.
-    (In other words, $a' > 2N_1 + N_2$.)
-    Furthermore, according to Claim~\ref{claim:(nu_1, nu'_1)-edge},
-    there exists a $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$.
-    Then, we have
-    $\mu(x) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
-    $\mu(x') < \mu(a')$ (since $\nu'_1$ is above any other gadget),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:all (nu_2, nu'_2)-edges}
-    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:all (nu_2, nu'_2)-edges}]
-    The key idea is to focus on $\nu'_2$.
-    Combining Claim~\ref{claim:no (nu_3, nu'_2)-edge},
-    Claim~\ref{claim: no edge right of nu'_2} and
-    Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge},
-    we obtain that every position in $\nu'_2$ is involved in
-    either a $(\nu_1, \nu'_2$)-edge
-    or a $(\nu'_1, \nu'_2$)-edge
-    or $(\nu_2, \nu'_2$)-edge.
-    But, according to Claim~\ref{claim:all (nu_1, nu'_1)-edges},
-    all positions in $\nu_1$ and $\nu'_1$ are involved in
-    $(\nu_1, \nu'_1)$-edges.
-    Then it follows that
-    all positions in $\nu'_2$ are involved in a $(\nu_2, \nu'_2)$,
-    and hence
-    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_2, nu_4)-edge}
-    There is no $(\nu_2, \nu_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_4)-edge}]
-    According to Claim~\ref{claim:all (nu_2, nu'_2)-edges},
-    all positions in $\nu_2$ and $\nu'_2$ are involved in
-    $(\nu_2, \nu'_2)$-edges in $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (nu_4, nu'_3)-edge}
-    There is no $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no (nu_4, nu'_3)-edge}]
-    Suppose, aiming at a contradiction, that there exists
-    a $(\nu_4, \nu'_3)$-edge $(a, a')$ in $\mathcal{M}$.
-    According to Claim~\ref{claim:(nu_3, nu'_3)-edge},
-    there exists a $(\nu_3, \nu'_3)$-edge $(x, x')$
-    in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free, we must have
-    $x' < a'$.
-    Therefore,
-    we have $\mu(x) > \mu(a)$ (since $\nu_3$ is above $\nu_4$) and
-    $\mu(x') < \nu(a')$ (since $\nu'_3$ is increasing),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:(nu_4, nu'_4)-edge}
-    There is at least one $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:(nu_4, nu'_4)-edge}]
-    Suppose, aiming at a contradiction, that there is no
-    $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
-    First, according to
-    Claim~\ref{claim:all (nu_1, nu'_1)-edges} and
-    Claim~\ref{claim:all (nu_2, nu'_2)-edges},
-    there is neither a $(\nu_1, \nu_4)$-edge
-    nor a $(\nu'_1, \nu_4)$-edge
-    nor a $(\nu_2, \nu_4)$-edge
-    nor a $(\nu_4, \nu'_2)$-edge in $\mathcal{M}$.
-    Furthermore, according to
-    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge} and
-    Claim~\ref{claim:no (nu_4, nu'_3)-edge},
-    there is neither a $(\nu_3, \nu_4)$-edge
-    nor a $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
-    But $N_4 > 2n + 2k + 2$, and
-    hence $\mathcal{M}$ is not perfect,
-    which contradicts our hypothesis about $\mathcal{M}$.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no edge below (nu_4, nu'_4)-edge}
-    There is
-    neither a $(\pi', \pi')$-edge
-    nor a $(\sigma', \pi'')$-edge
-    nor a $(\sigma', \sigma'')$-edge
-    in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
-    Combine Claim~\ref{claim:(nu_4, nu'_4)-edge} with the
-    fact that $\mathcal{M}$ is containment-free.
-    \qed
-  \end{proof}
-
-  \begin{claim}
-    \label{claim:no (sigma', nu'_3)-edge, no (sigma', nu'_4)-edge}
-    There is neither a $(\sigma', \nu'_3)$-edge
-    nor a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
-  \end{claim}
-
-  \begin{proof}[of Claim~\label{claim:no (sigma', nu'_3)-edge}]
-    According to
-    Claim~\ref{claim: no edge right of nu'_2},
-    Claim~\ref{claim:all (nu_1, nu'_1)-edges}
-    Claim~\ref{claim:all (nu_2, nu'_2)-edges} and
-    Claim~\ref{claim:no (nu_4, nu'_3)-edge}
-    every position in $\sigma'$ is involved in
-    either a $(\sigma', \nu'_3)$-edge
-    or a $(\sigma', \pi')$-edge
-    or a $(\sigma', \nu'_4)$-edge
-    in $\mathcal{M}$.
-    Let $a$ be the first position of $\sigma'$
-    (\emph{i.e.}, $a = 2N_1 + N_2 + N_3 + 1$)
-    and
-    $b$ be the last position of $\sigma'$
-    (\emph{i.e.}, $b = 2N_1 + N_2 + N_3 + k + 2$),
-    and write $(a, a')$ and $(b, b')$ for the two associated edges of $\mathcal{M}$.
-    We consider the presence of a $(\sigma', \nu'_3)$-edge
-    or a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$ separately.
-    \begin{itemize}
-      \item
-      Suppose, aiming at a contradiction, that there exists
-      a $(\sigma', \nu'_3)$-edge in $\mathcal{M}$.
-      Since $\mathcal{M}$ is containment-free, we conclude that
-      $(a, a')$ must be a $(\sigma', \nu'_3)$-edge.
-      But $\mu(a) = 2N_2 + N_4 + 2n + k + 2 + (k+1) <
-      \mu(b) = 2N_2 + N_4 + 2n + k + 2 + (k+2)$, and hence
-      $(b, b')$ is also a $(\sigma', \nu'_3)$-edge
-      (otherwise we would have $\mu(a') > \mu(b')$ since
-      $\nu'_3$ is above $\pi'$ and $\nu'_4$,
-      which contradicts our hypothesis).
-      Consider any position $c$ but $a$ and $b$ of $\sigma'$
-      (\emph{i.e.},
-      $a = 2N_1 + N_2 + N_3 + 1 < c < 2N_1 + N_2 + N_3 + k + 2 = b$),
-      and let $(c, c')$ be the associate edge in $\mathcal{M}$.
-      We first observe that $(c, c')$ is not a $(\sigma', \nu'_3)$-edge,
-      since otherwise we would have
-      $\mu(a) > \mu(c)$ and $\mu(a') < \mu(c')$
-      (since $\mathcal{M}$ is containment-free and $\nu'_3$ is increasing),
-      thereby contradicting our hypothesis about $\mathcal{M}$.
-      Therefore, $(c, c')$ is either a $(\sigma', \pi)$-edge or a $(\sigma', \nu'_4)$-edge.
-      In both cases, $(b, b')$ is contained in $(c, c')$,
-      which contradicts our hypothesis about $\mathcal{M}$ being containment-free.
-      \item
-      Suppose, aiming at a contradiction, that there exists
-      a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
-      Since $\mathcal{M}$ is containment-free, we conclude that
-      $(b, b')$ must be a $(\sigma', \nu'_4)$-edge.
-      Since there is no $(\sigma', \nu'_3)$-edge in $\mathcal{M}$
-      (preceding point),
-      $(a, a')$ is either a
-      $(\sigma', \pi')$-edge or a $(\sigma', \nu'_4)$-edge.
-      In both cases, we have $a' < a$.
-      Then it follows that
-      $\mu(a) < \mu(b)$
-      (since $2N_2 + N_4 + 2n + k + 2 + (k+1) < 2N_2 + N_4 + 2n + k + 2 + (k+2)$)
-      and
-      $\mu(a') > \mu(b')$ (since $\pi'$ is above $\nu'_4$ if $(a, a')$ is
-      a $(\sigma', \pi')$-edge and
-      $\nu'_4$ is decreasing if $(a, a')$ is
-      a $(\sigma', \nu'_4)$-edge),
-      which contradicts our hypothesis about $\mathcal{M}$.
-    \end{itemize}
-    \qed
-  \end{proof}
-
-  Combining the above claims, we conclude that there are
-  $k+2$ $(\sigma', \pi')$-edges in $\mathcal{M}$.
-  Recall that
-  $\sigma' = ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$ and that
-  $\pi' = ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$.
-  Then it follows we have
-  at least $k$
-  %(possibly $k+1$ or $k+2$)
-  independent $(\sigma', \pi')$-edges
-  $(a, a')$ in $\mathcal{M}$ with
-  $2N_1 + N_2 + N_3 + 1 < a < 2N_1 + N_2 + N_3 + k + 2$
-  and
-  $2N_1 + 2N_2 + 2N_3 + N_4 + (k + 2) + 1 < a' <2N_1 + 2N_2 + 2N_3 + N_4 + (k + 2) + n + 2$.
-  Therefore, by our hypothesis about $\mathcal{M}$,
-  $\sigma$ occurs as a pattern in $\pi$.
-  \qed
-\end{proof}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-02-02 article.tex
add arcs

--- 2015-09-18 article.tex
@@
 \usepackage{multirow}
+\usepackage{esvect}
 
@@
 \DeclareMathOperator{\PERMTOBIN}{\mathrm{ptb}}
+
+\newcommand{\CrossingLL}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingLL
+
+\newcommand{\CrossingLR}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingLR
+
+\newcommand{\CrossingRL}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingRL
+
+\newcommand{\CrossingRR}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingRR
+
+\newcommand{\InclusionLL}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionLL
+
+\newcommand{\InclusionLR}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionLR
+
+\newcommand{\InclusionRL}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionRL
+
+\newcommand{\InclusionRR}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionRR
+
+\newcommand{\PrecedenceLL}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLL
+
+\newcommand{\PrecedenceLR}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLR
+
+\newcommand{\PrecedenceRL}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRL
+
+\newcommand{\PrecedenceRR}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRR
 
@@
 graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows.
+
+\begin{lemma}
+  \label{lemma:matching}
+  Let $\pi$ be a permutation. The following statements are equivalent:
+  \begin{enumerate}
+    \item The permutation $\pi$ is a square.
+    \item There exists an oriented perfect matching $\vv{\mathcal{M}}$
+    of $G_\pi$ that avoids \InclusionLL, \InclusionLR,\InclusionRL,\InclusionRR,
+    \CrossingLR and \CrossingRL such that for any two distinct arcs $(a, a)$ and
+    $(b, b')$ in $\vv{\mathcal{M}}$ we have $\pi(a) < \pi(b)$ if and only
+    if $\pi(a') < \pi(b')$.
+  \end{enumerate}
+\end{lemma}
 

### 2016-02-02 submission.tex

@@
 \label{section:Introduction}
-The {\em shuffle product}, denoted $\shuffle$, is a very famous operation on words
+The {\em shuffle product}, denoted $\shuffle$, is a well-known operation on words
 first defined by Eilenberg and Mac Lane~\cite{Eilenberg:MacLane:1953}.
@@
 \end{enumerate}
-Even if these problems seem similar, they are very different in terms
+Even if these problems seem similar, they radically differ in terms
 of time complexity. Let us now review some facts about these. In what follows,
@@
 Section~\ref{section:Binary square words and permutations} is devoted to
-show that the problems related to the shuffle of words has links with
+showing that the problems related to the shuffle of words has links with
 the shuffle of permutations. In particular, we show that binary words
@@
 write $\nearrow_{n}$ (resp. $\searrow_{n}$) for the permutation
-$1 2 \dots n$ (resp. $n (n-1) \dots 1$). If $\pi$ is a permutation of
+$1 2 \dots n$ (resp. $n\,(n-1) \dots 1$). If $\pi$ is a permutation of
 $S_n$, we denote by $\bar \pi$ the {\em complement} of $\pi$, that is
@@
 \end{equation}
-By Proposition~\ref{prop:bijection_binary_to_permutations_squares}, this
+According to
+Proposition~\ref{prop:bijection_binary_to_permutations_squares}, this
 is also the sequence enumerating square permutations avoiding $213$
@@
   % we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
-  It can be shown (see Appendix) that
-  $\sigma$ occurs in $\pi$ if and only if
+  It can be shown
+  %(see Appendix)
+  that $\sigma$ occurs in $\pi$ if and only if
   there exists a containment-free perfect matching
@@
 Proposition~\ref{prop:bijection_binary_to_permutations_squares},
-how many binary strings of length $2n$ are squares; Problem~4
+how many binary strings of length $2n$ are squares; see also Problem~4
 in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of
@@
 deciding whether $\sigma$ is a square root of~$\pi$?
-Moreover, one can ask for a complete algebraic study of
+As for algebra, one can ask for a complete algebraic study of
 $\QQ[S]$ as a graded associative algebra for the shuffle  product

### 2016-02-02 article.tex
add tikz arcs

@@
 
+\newcommand{\Crossing}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \Crossing
+
 \newcommand{\CrossingLL}{%
@@
 
+\newcommand{\Inclusion}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \Inclusion
+
 \newcommand{\InclusionLL}{%
@@
 }% end \InclusionRR
+
+\newcommand{\Precedence}{%
+\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \Precedence
 
@@
     \item There exists an oriented perfect matching $\vv{\mathcal{M}}$
-    of $G_\pi$ that avoids \InclusionLL, \InclusionLR,\InclusionRL,\InclusionRR,
-    \CrossingLR and \CrossingRL such that for any two distinct arcs $(a, a)$ and
+    of the linear graph $G_\pi$ that avoids
+    \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
+    and \CrossingRL, and such that for any two distinct arcs $(a, a)$ and
     $(b, b')$ in $\vv{\mathcal{M}}$ we have $\pi(a) < \pi(b)$ if and only

### 2016-02-02 new-article.tex
add new-article.tex

--- 2016-02-02 article.tex
@@
 
+\newcommand{\LabeledInclusionLL}[4]{%
+\begin{tikzpicture}[scale=0.25,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#1};
+\draw node [] [right of=U1] (U2) {#2};
+\draw node [] [right of=U2] (U3) {#3};
+\draw node [] [right of=U3] (U4) {#4};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionLL
+
 \newcommand{\InclusionLL}{%
@@
 
+\newcommand{\LabeledInclusionLR}[4]{%
+\begin{tikzpicture}[scale=0.25,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#1};
+\draw node [] [right of=U1] (U2) {#2};
+\draw node [] [right of=U2] (U3) {#3};
+\draw node [] [right of=U3] (U4) {#4};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionLR
+
 \newcommand{\InclusionLR}{%
@@
 
+\newcommand{\LabeledInclusionRL}[4]{%
+\begin{tikzpicture}[scale=0.25,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#1};
+\draw node [] [right of=U1] (U2) {#2};
+\draw node [] [right of=U2] (U3) {#3};
+\draw node [] [right of=U3] (U4) {#4};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionRL
+
 \newcommand{\InclusionRL}{%
@@
 }% end \InclusionRL
+
+\newcommand{\LabeledInclusionRR}[4]{%
+\begin{tikzpicture}[scale=0.25,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#1};
+\draw node [] [right of=U1] (U2) {#2};
+\draw node [] [right of=U2] (U3) {#3};
+\draw node [] [right of=U3] (U4) {#4};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionRR
 
@@
 
-\begin{lemma}
-  \label{lemma:matching}
-  Let $\pi$ be a permutation. The following statements are equivalent:
-  \begin{enumerate}
-    \item The permutation $\pi$ is a square.
-    \item There exists a perfect containment-free matching $\mathcal{M}$
-    of $G_\pi$ such that for any two distinct edges $(a, b)$ and
-    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
-    if $\pi(b) < \pi(b')$.
-  \end{enumerate}
-\end{lemma}
-
 \begin{proof}[of Lemma~\ref{lemma:matching}]
-    The proof is, naturally enough, a slight extension of
-    Lemma~1 in \cite{Rizzi:Vialette:CSR:2013} and is added here in full
-    for the sake of completeness.
-
     (\emph{1} $\Rightarrow$ \emph{2}).
-    Since $\pi$ is a square, it admits a square root $\sigma$. Then, by
-    definition, there are two sets $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
+    Since $\pi$ is a square, it admits a square root, say $\sigma$.
+    Let $2n = |\pi|$ (and hence $|\sigma| = n$).
+    Then, by definition, there exist two sets
+    $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
     and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
     letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
-    order-isomorphic to $\sigma$. Let $2n = |\pi|$ (and hence
-    $|\sigma| = n$). Let $G_\pi$ be the linear clique associated to $\pi$
-    with vertices
-    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$. It
-    is easily seen that $\mathcal{M} = \{\{i^1_j, i^2_j\} : j \in [n]\}$
-    is a subset of $\ES{G_\pi}$. Furthermore, $\mathcal{M}$ is a perfect
+    order-isomorphic to $\sigma$.
+    Let $G_\pi$ be the linear clique associated to $\pi$ with vertices
+    $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$.
+    Define $\vv{\mathcal{M}} = \{(i^1_j, i^2_j) : j \in [n]\}$.
+    It is easily seen that $\vv{\mathcal{M}}$ is an oriented perfect
     matching since $I^1 \cap I^2 = \emptyset$, and
-    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$. It is also containment-free.
+    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$.
+    It also avoids \InclusionLL,~\InclusionLR,~\InclusionRL and \InclusionRR.
+    Indeed, suppose, aiming at a contradiction, that such a bad configuration exists
+    for arcs $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$
+    with $j < k$.
+    \begin{itemize}
+      \item
+      \LabeledInclusionLL{$i^1_j$}{$i^1_k$}{$i^2_k$}{$i^2_j$}
+      \item
+      \LabeledInclusionLL{$i^1_j$}{$i^1_k$}{$i^2_k$}{$i^2_j$}
+      \item
+      \LabeledInclusionLL{$i^1_j$}{$i^1_k$}{$i^2_k$}{$i^2_j$}
+      \item
+      \LabeledInclusionLL{$i^1_j$}{$i^1_k$}{$i^2_k$}{$i^2_j$}
+    \end{itemize}
+
     Indeed, if it is not the case then there would exist two edges

### 2016-02-03 new-article.tex
LATIN 2016

@@
 
+\newcommand{\LabeledCrossingLL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingLL
+
 \newcommand{\CrossingLL}{%
@@
 \draw node [draw,circle,fill=black] [right of=U3] (U4) {};
-\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
-\draw [->] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U3.north) + (0,2) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
 \end{tikzpicture}
 }% end \CrossingLL
+
+\newcommand{\LabeledCrossingLR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingLR
 
@@
 
+\newcommand{\LabeledCrossingRL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingRL
+
 \newcommand{\CrossingRL}{%
@@
 
+\newcommand{\LabeledCrossingRR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingRR
+
 \newcommand{\CrossingRR}{%
@@
 
-\newcommand{\LabeledInclusionLL}[4]{%
-\begin{tikzpicture}[scale=0.25,inner sep=.8pt,node distance=.45cm,>=latex']
-\draw node [] (U1)               {#1};
-\draw node [] [right of=U1] (U2) {#2};
-\draw node [] [right of=U2] (U3) {#3};
-\draw node [] [right of=U3] (U4) {#4};
+\newcommand{\LabeledInclusionLL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
 \draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
@@
 
-\newcommand{\LabeledInclusionLR}[4]{%
-\begin{tikzpicture}[scale=0.25,inner sep=.8pt,node distance=.45cm,>=latex']
-\draw node [] (U1)               {#1};
-\draw node [] [right of=U1] (U2) {#2};
-\draw node [] [right of=U2] (U3) {#3};
-\draw node [] [right of=U3] (U4) {#4};
+\newcommand{\LabeledInclusionLR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
 \draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
@@
 
-\newcommand{\LabeledInclusionRL}[4]{%
-\begin{tikzpicture}[scale=0.25,inner sep=.8pt,node distance=.45cm,>=latex']
-\draw node [] (U1)               {#1};
-\draw node [] [right of=U1] (U2) {#2};
-\draw node [] [right of=U2] (U3) {#3};
-\draw node [] [right of=U3] (U4) {#4};
+\newcommand{\LabeledInclusionRL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
 \draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
@@
 
-\newcommand{\LabeledInclusionRR}[4]{%
-\begin{tikzpicture}[scale=0.25,inner sep=.8pt,node distance=.45cm,>=latex']
-\draw node [] (U1)               {#1};
-\draw node [] [right of=U1] (U2) {#2};
-\draw node [] [right of=U2] (U3) {#3};
-\draw node [] [right of=U3] (U4) {#4};
+\newcommand{\LabeledInclusionRR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
 \draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
@@
 
+\newcommand{\LabeledPrecedenceLL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLL
+
 \newcommand{\PrecedenceLL}{%
@@
 
+\newcommand{\LabeledPrecedenceLR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLR
+
 \newcommand{\PrecedenceLR}{%
@@
 
+\newcommand{\LabeledPrecedenceRL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRL
+
 \newcommand{\PrecedenceRL}{%
@@
 }% end \PrecedenceRL
+
+\newcommand{\LabeledPrecedenceRR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRR
 
@@
     Then, by definition, there exist two sets
-    $I^1 = \{i^1_1 < i^1_2 < \dots < i^1_n\}$
-    and $I^2 = \{i^2_1 < i^2_2 < \dots < i^2_n\}$ of disjoint indexes of
+    $I^1 = (i^1_1 < i^1_2 < \dots < i^1_n)$
+    and $I^2 = (i^2_1 < i^2_2 < \dots < i^2_n)$ of disjoint indexes of
     letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
@@
     $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$.
-    It also avoids \InclusionLL,~\InclusionLR,~\InclusionRL and \InclusionRR.
+    We first show that it avoids
+    \InclusionLL,~\InclusionLR,~\InclusionRL and \InclusionRR.
     Indeed, suppose, aiming at a contradiction, that such a bad configuration exists
-    for arcs $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$
-    with $j < k$.
-    \begin{itemize}
-      \item
-      \LabeledInclusionLL{$i^1_j$}{$i^1_k$}{$i^2_k$}{$i^2_j$}
-      \item
-      \LabeledInclusionLL{$i^1_j$}{$i^1_k$}{$i^2_k$}{$i^2_j$}
-      \item
-      \LabeledInclusionLL{$i^1_j$}{$i^1_k$}{$i^2_k$}{$i^2_j$}
-      \item
-      \LabeledInclusionLL{$i^1_j$}{$i^1_k$}{$i^2_k$}{$i^2_j$}
-    \end{itemize}
-
-    Indeed, if it is not the case then there would exist two edges
-    $e = \{i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$, $j < k$, in
-    $\mathcal{M}$ such that $i^1_j < i^1_k < i^2_k < i^2_j$ or
-    $i^1_j < i^2_k < i^1_k < i^2_j$. This is a contradiction since
-    $i^2_k > i^2_j$ if $k > j$. Finally, for any two distinct edges
-    $e = \{\i^1_j, i^2_j\}$ and $e' = \{i^1_k, i^2_k\}$ of $\mathcal{M}$,
+    for, say,  arcs $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$.
+    Assuming $i^1_j < i^1_k$ we are left with the following $4$ configurations:
+    \begin{center}
+      \begin{tabular}{cccc}
+        \LabeledInclusionLL{.3}{$i^1_j$}{$i^1_k$}{$i^2_k$}{$i^2_j$}
+        &\qquad
+        \LabeledInclusionLR{.3}{$i^1_j$}{$i^2_k$}{$i^1_k$}{$i^2_j$}
+        &\qquad
+        \LabeledInclusionRL{.3}{$i^2_k$}{$i^1_j$}{$i^2_j$}{$i^1_k$}
+        &\qquad
+        \LabeledInclusionRR{.3}{$i^2_k$}{$i^2_j$}{$i^1_j$}{$i^1_k$}
+      \end{tabular}
+    \end{center}
+    Then it follows that $i^2_j > i^2_k$.
+    This is a contradiction since
+    $i^1_j < i^1_k \Rightarrow j < k \Rightarrow i^2_j < i^2_k$.
+    We now turn to proving that $\vv{\mathcal{M}}$ also avoids
+    \CrossingLR and \CrossingRL.
+    Indeed, suppose, aiming at a contradiction, that such a bad configuration exists
+    for, say,  arcs $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$.
+    Assuming $i^1_j < i^1_k$ we are left with the following $2$ configurations:
+    \begin{center}
+      \begin{tabular}{cccc}
+        \LabeledCrossingLR{.3}{$i^1_j$}{$i^2_k$}{$i^2_j$}{$i^1_k$}
+        &\qquad
+        \LabeledCrossingRL{.3}{$i^2_k$}{$i^1_j$}{$i^1_k$}{$i^2_j$}
+      \end{tabular}
+    \end{center}
+    Then it follows that $i^2_j > i^2_k$.
+    Again, this is a contradiction since
+    $i^1_j < i^1_k \Rightarrow j < k \Rightarrow i^2_j < i^2_k$.
+    Finally, for any two distinct arcs
+    $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$,
     we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
@@
     (\emph{2} $\Rightarrow$ \emph{1}).
-    Let $I^1 = \{i^1_j : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$, and
-    $I^2 = \{i^2_k : \exists \{i_j, i_k\} \in \mathcal{M}
-    \;\text{with}\; j < k\}$.
-    Informally, $I^1$ (resp. $I^2$) is the set of the left (resp. right)
-    endpoints of the edges in $\mathcal{M}$ assuming the positions
-    $1, 2, \dots, 2$ are drawn along an horizontal line.
-    Without loss of generality, assume $i^1_1 < i^1_2 < \dots < i^1_n$.
-    % and $I^2 = \{i^2_1, i^2_2, \dots, i^2_n\}$.
-    Let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
-    $p^2 = \pi(i^2_1) \pi(i^2_2) \dots \pi(i^2_n)$ be the two patterns
-    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$
-    and $p^2$ are disjoint in $\pi$ (since $\mathcal{M}$ is a matching)
-    and cover $\pi$ (since $\mathcal{M}$ is perfect). We also claim that
-    $p^1$ and $p^2$ are order-isomorphic. Suppose, aiming at a
-    contradiction, that this is not the case. Therefore, there exist two
-    positions $j$ and $k$, say $j < k$, such that
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$, or
-    $\pi(i^1_j) < \pi(i^2_j)$ and $\pi(i^1_k) > \pi(i^2_k)$.
-    But since $\mathcal{M}$ is containment-free, it follows that both
-    $\{i^1_j, i^2_j\}$, $i^1_j < i^2_j$, and $\{i^1_k, i^2_k\}$,
-    $i^1_k, i^2_k$, belong to $\mathcal{M}$. This is a contradiction
-    since for any two distinct edges $(a, b)$ and $(a', b')$ in
-    $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only if
+    Let $I^1 = (i^1_1 < i^1_2 < \dots < i^1_n)$
+    (resp. $I^2 = (i^2_1 < i^2_2 < \dots < i^2_n)$)
+    be the ordered sequence of the origins (resp. sinks) of the arcs of
+    the oriented perfect matching $\vv{\mathcal{M}}$.
+    We first show that, for every $1 \leq j \leq n$,
+    $(i^1_j, i^2_j)$ is an arc of $\vv{\mathcal{M}}$.
+    The proof is by induction on $n$.
+    Since the assertion is certainly valid for $n = 1$,
+    it is enough to show that $(i^1_n, i^2_n)$ is an arc of $\vv{\mathcal{M}}$.
+    Suppose, aiming at a contradiction that $(i^1_n, i^2_n)$
+    is not an arc of $\vv{\mathcal{M}}$.
+    Let $i^2_p$ be the sink of the arc with origin $i^1_n$ in
+    $\vv{\mathcal{M}}$ (\emph{i.e.}, $(i^1_n, i^2_p)$ is an abstractof
+    $\vv{\mathcal{M}}$).
+    We need to consider two cases.
+    \begin{itemize}
+      \item If $i^1_n > i^2_p$ then there exist $i^1_q$ and $i^2_r$ such
+      that
+      \begin{center}
+        \begin{tabular}{ccc}
+          \LabeledInclusionRL{.3}{$i^2_p$}{$i^1_q$}{$i^2_r$}{$i^1_n$}
+          &\qquad
+          \LabeledInclusionRR{.3}{$i^1_p$}{$i^2_r$}{$i^1_q$}{$i^1_n$}
+          &\qquad
+          \LabeledCrossingLR{.3}{$i^1_p$}{$i^2_p$}{$i^2_r$}{$i^1_n$}
+        \end{tabular}
+      \end{center}
+      This is a contradiction since
+      $\vv{\mathcal{M}}$ avoids
+      \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR and
+      \CrossingRL.
+      \item
+      If $i^1_n < i^2_p$ then there exist $i^1_q$ and $i^2_r$ such
+      that
+      \begin{center}
+        \begin{tabular}{ccc}
+          \LabeledInclusionLL{.3}{$i^1_q$}{$i^1_n$}{$i^2_p$}{$i^2_r$}
+          &\qquad
+          \LabeledInclusionRL{.3}{$i^2_r$}{$i^1_n$}{$i^2_p$}{$i^1_q$}
+          &\qquad
+          \LabeledCrossingRL{.3}{$i^2_p$}{$i^1_q$}{$i^1_n$}{$i^2_r$}
+        \end{tabular}
+      \end{center}
+      This is a contradiction since
+      $\vv{\mathcal{M}}$ avoids
+      \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR and
+      \CrossingRL.
+    \end{itemize}
+    Then it follows that $(i^1_n, i^2_n)$ is an arc
+    of $\vv{\mathcal{M}}$.
+    Now, let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
+    $p^2 = \pi(j^2_1) \pi(^2_2) \dots \pi(i^2_n)$ be the two patterns
+    from $\pi$ induced by $I^1$ and $I^2$, respectively.
+    Clearly $p^1$ and $p^2$ are disjoint in $\pi$
+    (since $\vv{\mathcal{M}}$ is a matching)
+    and cover $\pi$ (since $\vv{\mathcal{M}}$ is perfect).
+    We claim that $p^1$ and $p^2$ are order-isomorphic.
+    Indeed, this is a direct consequence of th fact that,
+    for any two distinct edges $(a, b)$ and $(a', b')$ in
+    $\vv{\mathcal{M}}$ we have $\pi(a) < \pi(a')$ if and only if
     $\pi(b) < \pi(b')$.

### 2016-02-03 new-article.tex
LATIN 2016
Typos lemme 2

@@
 \draw node [draw,circle,fill=black] [right of=U3] (U4) {};
-\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
-\draw [->] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\draw [->] (U4.north) .. controls ($ (U4.north) + (0,4) $) and ($ (U1.north) + (0,4) $) .. (U1.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,3) $) and ($ (U2.north) + (0,3) $) .. (U2.north);
 \end{tikzpicture}
@@
     matching since $I^1 \cap I^2 = \emptyset$, and
-    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$.
+    $I^1 \cup I^2 = [2n]$.
     We first show that it avoids
@@
     for, say,  arcs $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$.
-    Assuming $i^1_j < i^1_k$ we are left with the following $4$ configurations:
+    Assuming without loss of generality that $i^1_j < i^1_k$, we are 
+    left with the following four configurations:
     \begin{center}
@@
     for, say,  arcs $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$.
-    Assuming $i^1_j < i^1_k$ we are left with the following $2$ configurations:
+    Assuming without loss of generality that $i^1_j < i^1_k$, we are 
+    left with the following two configurations:
     \begin{center}
@@
     $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
-    elements (at position $j$ and $k$) in two patterns that are
-    order-isomorphic to $\sigma$.
+    elements (at positions $j$ and $k$) in two patterns that are
+    order-isomorphic to~$\sigma$.
+    \smallskip
 
@@
     the oriented perfect matching $\vv{\mathcal{M}}$.
-    We first show that, for every $1 \leq j \leq n$,
+    We first show that, for every $j \in [n]$,
     $(i^1_j, i^2_j)$ is an arc of $\vv{\mathcal{M}}$.
@@
     Let $i^2_p$ be the sink of the arc with origin $i^1_n$ in
-    $\vv{\mathcal{M}}$ (\emph{i.e.}, $(i^1_n, i^2_p)$ is an abstractof
+    $\vv{\mathcal{M}}$ (\emph{i.e.}, $(i^1_n, i^2_p)$ is an arc of
     $\vv{\mathcal{M}}$).
@@
       \begin{center}
-        \begin{tabular}{ccc}
+        \begin{tabular}{cccc}
           \LabeledInclusionRL{.3}{$i^2_p$}{$i^1_q$}{$i^2_r$}{$i^1_n$}
           &\qquad
-          \LabeledInclusionRR{.3}{$i^1_p$}{$i^2_r$}{$i^1_q$}{$i^1_n$}
+          \LabeledInclusionRR{.3}{$i^2_p$}{$i^2_r$}{$i^1_q$}{$i^1_n$}
           &\qquad
           \LabeledCrossingLR{.3}{$i^1_p$}{$i^2_p$}{$i^2_r$}{$i^1_n$}
+          &\qquad
+          \LabeledCrossingRL{.3}{$i^2_p$}{$i^1_q$}{$i^1_n$}{$i^2_r$}
         \end{tabular}
@@
       \begin{center}
-        \begin{tabular}{ccc}
+        \begin{tabular}{cc}
           \LabeledInclusionLL{.3}{$i^1_q$}{$i^1_n$}{$i^2_p$}{$i^2_r$}
@@
           \LabeledInclusionRL{.3}{$i^2_r$}{$i^1_n$}{$i^2_p$}{$i^1_q$}
-          &\qquad
-          \LabeledCrossingRL{.3}{$i^2_p$}{$i^1_q$}{$i^1_n$}{$i^2_r$}
         \end{tabular}

### 2016-02-03 new-article.tex
LATIN 2016

--- 2016-02-03 new-article.tex
@@
     \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
-    and \CrossingRL, and such that for any two distinct arcs $(a, a)$ and
+    and \CrossingRL, and such that for any two distinct arcs $(a, a')$ and
     $(b, b')$ in $\vv{\mathcal{M}}$ we have $\pi(a) < \pi(b)$ if and only
@@
       \begin{center}
-        \begin{tabular}{ccc}
+        \begin{tabular}{cccc}
           \LabeledInclusionRL{.3}{$i^2_p$}{$i^1_q$}{$i^2_r$}{$i^1_n$}
@@
           \LabeledCrossingLR{.3}{$i^1_p$}{$i^2_p$}{$i^2_r$}{$i^1_n$}
+          &\qquad
+          \LabeledCrossingRL{.3}{$i^2_p$}{$i^1_q$}{$i^1_n$}{$i^2_r$}
         \end{tabular}
@@
 Before proving hardness, we give an easy lemma that will prove extremely
-useful for simplifying the proof of upcoming
-Proposition~\ref{proposition:hardness}.
-
-\begin{lemma}
-  \label{lemma:at most one edge monotone}
-  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
-  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
-  and $\mathcal{M}$ be a containment-free perfect matching of $G_\pi$.
-  The following assertions hold:
-  \begin{enumerate}
-    \item
-    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
-    $\sigma_1$ is above $\sigma_2$ (see Fig.~\ref{subfig:increasing before
-    and above decreasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
-    \item
-    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
-    is below $\sigma_2$ (see Fig.~\ref{subfig:decreasing before and below
-    increasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
-  \end{enumerate}
-\end{lemma}
-
-\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
+useful for simplifying the proof of upcoming Proposition~\ref{proposition:hardness}.
+
 \begin{figure}[t!]
@@
       [
-          scale=0.325,
+        scale=0.2,
         label/.style={anchor=base}
@@
       % labels
-        \node [label] (a) at (1,5) {$a$};
-        \node [label] (b) at (4,5) {$b$};
-        \node [label] (ap) at (7,-1) {$a'$};
-        \node [label] (bp) at (10,-1) {$b'$};
-        \node (nu1) at (2.5,4) {$\sigma_1$};
-        \node (nu2) at (8.5,-2) {$\sigma_2$};
+      \node [label] (a) at (1,4) {$a$};
+      \node [label] (b) at (4,4) {$b$};
+      \node [label] (ap) at (7,-2) {$a'$};
+      \node [label] (bp) at (10,-2) {$b'$};
+      \node (nu1) at (2.5,3) {$\sigma_1$};
+      \node (nu2) at (8.5,-3) {$\sigma_2$};
       % crossings edges
-        \draw [line width=1pt]
+      \draw [line width=1pt,->,>=latex']
       (1,7) .. controls +(0,12) and +(0,10) .. (7,4);
-        \draw [line width=1pt]
+      \draw [line width=1pt,->,>=latex']
+      (4,10) .. controls +(0,12) and +(0,10) .. (10,1);
+    \end{tikzpicture}
+    \;
+    \begin{tikzpicture}
+      [
+        scale=0.2,
+        label/.style={anchor=base}
+      ]
+      % increasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
+      \foreach \x/\y in {0/6,1/7,2/8,3/9,4/10,5/11} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % decreasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
+      \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % labels
+      \node [label] (a) at (1,4) {$a'$};
+      \node [label] (b) at (4,4) {$b'$};
+      \node [label] (ap) at (7,-2) {$a$};
+      \node [label] (bp) at (10,-2) {$b$};
+      \node (nu1) at (2.5,3) {$\sigma_1$};
+      \node (nu2) at (8.5,-3) {$\sigma_2$};
+      % crossings edges
+      \draw [line width=1pt,<-,>=latex']
+      (1,7) .. controls +(0,12) and +(0,10) .. (7,4);
+      \draw [line width=1pt,<-,>=latex']
       (4,10) .. controls +(0,12) and +(0,10) .. (10,1);
@@
   }% end subfigure
-    \qquad\qquad\qquad
+  \qquad
   \subfigure[%
@@
     [
-        scale=0.325,
+      scale=0.2,
       label/.style={anchor=base}
@@
     % labels
-      \node [label] (a) at (1,-1) {$a$};
-      \node [label] (b) at (4,-1) {$b$};
-      \node [label] (ap) at (7,5) {$a'$};
-      \node [label] (bp) at (10,5) {$b'$};
-      \node (nu1) at (2.5,-2) {$\sigma_1$};
-      \node (nu2) at (8.5,4) {$\sigma_2$};
+    \node [label] (a) at (1,-2) {$a$};
+    \node [label] (b) at (4,-2) {$b$};
+    \node [label] (ap) at (7,4) {$a'$};
+    \node [label] (bp) at (10,4) {$b'$};
+    \node (nu1) at (2.5,-3) {$\sigma_1$};
+    \node (nu2) at (8.5,3) {$\sigma_2$};
     % crossings edges
-      \draw [line width=1pt]
+    \draw [line width=1pt,->,>=latex']
     (1,4) .. controls +(0,10) and +(0,12) .. (7,7);
-      \draw [line width=1pt]
+    \draw [line width=1pt,->,>=latex']
+    (4,1) .. controls +(0,10) and +(0,12) .. (10,10);
+    \end{tikzpicture}
+    \;
+    \begin{tikzpicture}
+      [
+        scale=0.2,
+        label/.style={anchor=base}
+      ]
+      % decreasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (5.2,5.2);
+      \foreach \x/\y in {0/5,1/4,2/3,3/2,4/1,5/0} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % decreasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,5.8) grid (11.2,11.2);
+      \foreach \x/\y in {6/6,7/7,8/8,9/9,10/10,11/11} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % labels
+      \node [label] (a) at (1,-2) {$a'$};
+      \node [label] (b) at (4,-2) {$b'$};
+      \node [label] (ap) at (7,4) {$a$};
+      \node [label] (bp) at (10,4) {$b$};
+      \node (nu1) at (2.5,-3) {$\sigma_1$};
+      \node (nu2) at (8.5,3) {$\sigma_2$};
+      % crossings edges
+      \draw [line width=1pt,<-,>=latex']
+      (1,4) .. controls +(0,10) and +(0,12) .. (7,7);
+      \draw [line width=1pt,<-,>=latex']
       (4,1) .. controls +(0,10) and +(0,12) .. (10,10);
@@
 \end{figure}
-  We first prove \emph{1}. Suppose, aiming at a contradiction, that
-  there exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$
-  in $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may
-  assume $a < b$ and $a' < b'$. Therefore, we have $\pi(a) < \pi(b)$
-  (since $\sigma_1$ is increasing) and $\pi(a') > \pi(b')$ (since
+
+\begin{lemma}
+  \label{lemma:at most one edge monotone}
+  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
+  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
+  and $\vv{\mathcal{M}}$ be an oriented perfect matching of $G_\pi$
+  that avoids
+  \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
+  and \CrossingRL, and such that for any two distinct arcs $(a, a')$ and
+  $(b, b')$ in $\vv{\mathcal{M}}$ we have $\pi(a) < \pi(b)$ if and only
+  if $\pi(a') < \pi(b')$.
+  The following assertions hold:
+  \begin{enumerate}
+    \item
+    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
+    $\sigma_1$ is above $\sigma_2$
+    (see Fig.~\ref{subfig:increasing before and above decreasing}),
+    then $\vv{\mathcal{M}}$ contains either at most one
+    $(\sigma_1, \sigma_2)$-arc or at most one
+    $(\sigma_2, \sigma_1)$-arc.
+    \item
+    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
+    is below $\sigma_2$
+    (see Fig.~\ref{subfig:decreasing before and below increasing}),
+    then $\mathcal{M}$ contains at most one
+    $(\sigma_1, \sigma_2)$-edge.
+  \end{enumerate}
+\end{lemma}
+
+\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
+  We only prove the first assertion since the proof for the second one
+  is similar.
+  Suppose, aiming at a contradiction, that the assertion does not hold.
+  Therefore, since $\vv{\mathcal{M}}$ avoids
+  \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
+  and \CrossingRL,
+  $\vv{\mathcal{M}}$ contains either two crossing $(\sigma_1, \sigma_2)$-arcs
+  or two crossing $(\sigma_2, \sigma_1)$-arcs.
+  Suppose first that it contains two crossing $(\sigma_1, \sigma_2)$-arcs
+  $(a, a')$ and $(b, b')$ with $a < b$.
+  Therefore, we have $\pi(a) < \pi(b)$ (since $\sigma_1$ is increasing)
+  and $\pi(a') > \pi(b')$ (since
   $\sigma_2$ is decreasing), which contradicts our hypothesis about
-  $\mathcal{M}$.
-
-  As for \emph{2}, suppose, still aiming at a contradiction, that there
-  exist two $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in
-  $\mathcal{M}$. Since $\mathcal{M}$ is containment-free, we may assume
-  $a < b$ and $a' < b'$. Therefore, we have $\pi(a) > \pi(b)$ (since
-  $\sigma_1$ is decreasing) and $\pi(a') < \pi(b')$ (since $\sigma_2$ is
-  increasing), which contradicts our hypothesis about~$\mathcal{M}$.
+  $\vv{\mathcal{M}}$.
+  If $\vv{\mathcal{M}}$ contains two crossing $(\sigma_2, \sigma_1)$-arcs
+  $(a, a')$ and $(b, b')$ with $a < b$, then
+  we have $\pi(a) > \pi(b)$ (since $\sigma_2$ is decreasing)
+  and $\pi(a') < \pi(b')$ (since
+  $\sigma_2$ is increasing), which contradicts our hypothesis about
+  $\vv{\mathcal{M}}$.
   \qed

### 2016-02-03 new-article.tex
LATIN 2016

@@
   the pattern involvement problem which is known to be
-  \NPC~\cite{Bose:Buss:Lubiw:1998}: Given two permutations $\pi$ and
-  $\sigma$, decide whether $\sigma$ occurs in $\pi$ (as an
-  order-isomorphic pattern).
+  \NPC~\cite{Bose:Buss:Lubiw:1998}. Given two permutations $\pi$ and
+  $\sigma$, pattern involvement problem is to decide whether $\sigma$ occurs in
+  $\pi$ (as an order-isomorphic pattern).
 

### 2016-02-03 new-article.tex
Proof Lemma 2

@@
 \newcommand{\Crossing}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\CrossingLL}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\CrossingLR}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\CrossingRL}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\CrossingRR}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\Inclusion}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\InclusionLL}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \draw node [draw,circle,fill=black] [right of=U3] (U4) {};
-\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
-\draw [->] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
++\draw [->] (U4.north) .. controls ($ (U4.north) + (0,4) $) and ($ (U1.north) + (0,4) $) .. (U1.north);
++\draw [->] (U3.north) .. controls ($ (U3.north) + (0,3) $) and ($ (U2.north) + (0,3) $) .. (U2.north);
 \end{tikzpicture}
@@
 \newcommand{\InclusionLR}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\InclusionRL}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\InclusionRR}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\Precedence}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\PrecedenceLL}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\PrecedenceLR}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\PrecedenceRL}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
 \newcommand{\PrecedenceRR}{%
-\begin{tikzpicture}[scale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
 \draw node [draw,circle,fill=black] (U1)               {};
@@
     (\emph{1} $\Rightarrow$ \emph{2}).
-    Since $\pi$ is a square, it admits a square root, say $\sigma$.
-    Let $2n = |\pi|$ (and hence $|\sigma| = n$).
-    Then, by definition, there exist two sets
-    $I^1 = (i^1_1 < i^1_2 < \dots < i^1_n)$
-    and $I^2 = (i^2_1 < i^2_2 < \dots < i^2_n)$ of disjoint indexes of
+    Since $\pi$ is a square, it admits a square root, say~$\sigma$.
+    Let $2n = |\pi|$ (and hence $|\sigma| = n$). Then, by definition, 
+    there exist two sets $I^1 = (i^1_1 < i^1_2 < \dots < i^1_n)$ and 
+    $I^2 = (i^2_1 < i^2_2 < \dots < i^2_n)$ of disjoint indexes of
     letters of $\pi$ such that $\pi_{|I^1}$ and $\pi_{|I^2}$ are both
-    order-isomorphic to $\sigma$.
-    Let $G_\pi$ be the linear clique associated to $\pi$ with vertices
+    order-isomorphic to $\sigma$. Let $G_\pi$ be the linear clique 
+    associated with $\pi$ with vertices
     $\VS{G_\pi} = \{i^1_j : j \in [n]\} \cup \{i^2_j : j \in [n]\}$.
-    Define $\vv{\mathcal{M}} = \{(i^1_j, i^2_j) : j \in [n]\}$.
-    It is easily seen that $\vv{\mathcal{M}}$ is an oriented perfect
-    matching since $I^1 \cap I^2 = \emptyset$, and
-    $I^1 \cup I^2 = \{1, 2,  \dots 2n\}$.
-    We first show that it avoids
-    \InclusionLL,~\InclusionLR,~\InclusionRL and \InclusionRR.
-    Indeed, suppose, aiming at a contradiction, that such a bad configuration exists
-    for, say,  arcs $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$.
-    Assuming $i^1_j < i^1_k$ we are left with the following $4$ configurations:
+    Define $\vv{\mathcal{M}} = \{(i^1_j, i^2_j) : j \in [n]\}$. It is 
+    easily seen that $\vv{\mathcal{M}}$ is an oriented perfect matching 
+    since $I^1 \cap I^2 = \emptyset$, and $I^1 \cup I^2 = [2n]$. We 
+    first show that it avoids \InclusionLL, \InclusionLR, \InclusionRL, 
+    and \InclusionRR. Indeed, suppose, aiming at a contradiction, that 
+    such a configuration exists for, say, arcs $(i^1_j, i^2_j)$ and 
+    $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$. Assuming without loss of 
+    generality $i^1_j < i^1_k$, we are left with the following four 
+    configurations:
     \begin{center}
@@
     \end{center}
-    Then it follows that $i^2_j > i^2_k$.
-    This is a contradiction since
-    $i^1_j < i^1_k \Rightarrow j < k \Rightarrow i^2_j < i^2_k$.
-    We now turn to proving that $\vv{\mathcal{M}}$ also avoids
-    \CrossingLR and \CrossingRL.
-    Indeed, suppose, aiming at a contradiction, that such a bad configuration exists
-    for, say,  arcs $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$.
-    Assuming $i^1_j < i^1_k$ we are left with the following $2$ configurations:
+    Then it follows that $i^2_j > i^2_k$. This is a contradiction since
+    $i^1_j < i^1_k$ implies $j < k$, and hence, $i^2_j < i^2_k$. We now 
+    turn to proving that $\vv{\mathcal{M}}$ also avoids \CrossingLR and 
+    \CrossingRL. Indeed, suppose, aiming at a contradiction, that such a 
+    configuration exists for, say,  arcs $(i^1_j, i^2_j)$ and 
+    $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$. Assuming without loss of 
+    generality $i^1_j < i^1_k$, we are left with the following two 
+    configurations:
     \begin{center}
@@
     \end{center}
-    Then it follows that $i^2_j > i^2_k$.
-    Again, this is a contradiction since
-    $i^1_j < i^1_k \Rightarrow j < k \Rightarrow i^2_j < i^2_k$.
-    Finally, for any two distinct arcs
-    $(i^1_j, i^2_j)$ and $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$,
+    Then it follows that $i^2_j > i^2_k$. Again, this is a contradiction 
+    since $i^1_j < i^1_k$ implies $j < k$, and hence, $i^2_j < i^2_k$.
+    Finally, for any two distinct arcs $(i^1_j, i^2_j)$ and 
+    $(i^1_k, i^2_k)$ of $\vv{\mathcal{M}}$,
     we have $\pi(i^1_j) < \pi(i^2_j)$ if and only if
     $\pi(i^1_k) < \pi(i^2_k)$ since we are comparing in both cases two
-    elements (at position $j$ and $k$) in two patterns that are
-    order-isomorphic to $\sigma$.
+    elements (at positions $j$ and $k$) in two patterns that are
+    order-isomorphic to~$\sigma$.
+    \smallskip
 
     (\emph{2} $\Rightarrow$ \emph{1}).
-    Let $I^1 = (i^1_1 < i^1_2 < \dots < i^1_n)$
-    (resp. $I^2 = (i^2_1 < i^2_2 < \dots < i^2_n)$)
+    Let $I^1 = (i^1_1 < i^1_2 < \dots < i^1_n)$ (resp. 
+    $I^2 = (i^2_1 < i^2_2 < \dots < i^2_n)$)
     be the ordered sequence of the origins (resp. sinks) of the arcs of
-    the oriented perfect matching $\vv{\mathcal{M}}$.
-    We first show that, for every $1 \leq j \leq n$,
-    $(i^1_j, i^2_j)$ is an arc of $\vv{\mathcal{M}}$.
-    The proof is by induction on $n$.
-    Since the assertion is certainly valid for $n = 1$,
-    it is enough to show that $(i^1_n, i^2_n)$ is an arc of $\vv{\mathcal{M}}$.
-    Suppose, aiming at a contradiction that $(i^1_n, i^2_n)$
-    is not an arc of $\vv{\mathcal{M}}$.
-    Let $i^2_p$ be the sink of the arc with origin $i^1_n$ in
-    $\vv{\mathcal{M}}$ (\emph{i.e.}, $(i^1_n, i^2_p)$ is an arc of
-    $\vv{\mathcal{M}}$).
-    We need to consider two cases.
-    \begin{itemize}
-      \item If $i^1_n > i^2_p$ then there exist $i^1_q$ and $i^2_r$ such
-      that
+    the oriented perfect matching $\vv{\mathcal{M}}$. Let us first show 
+    that, for every $j \in [n]$, $(i^1_j, i^2_j)$ is an arc of 
+    $\vv{\mathcal{M}}$. For that, we show that $(i^1_n, i^2_n)$ is an 
+    arc of $\vv{\mathcal{M}}$. Suppose, aiming at a contradiction that 
+    $(i^1_n, i^2_n)$ is not an arc of $\vv{\mathcal{M}}$. Then, there 
+    exist $i^2_p, i^1_q \in \VS{\vv{\mathcal{M}}}$ such that 
+    $(i^1_n, i^2_p)$ and $(i^1_q, i^2_n)$ are arcs of $\vv{\mathcal{M}}$.
+    Since $p < n$ and $q < n$, there is in $\vv{\mathcal{M}}$ one 
+    of the following configurations:
     \begin{center}
         \begin{tabular}{cccc}
-          \LabeledInclusionRL{.3}{$i^2_p$}{$i^1_q$}{$i^2_r$}{$i^1_n$}
+          \LabeledInclusionLR{.3}{$i^1_q$}{$i^2_p$}{$i^1_n$}{$i^2_n$}
           &\qquad
-          \LabeledInclusionRR{.3}{$i^1_p$}{$i^2_r$}{$i^1_q$}{$i^1_n$}
+          \LabeledCrossingRL{.3}{$i^2_p$}{$i^1_q$}{$i^1_n$}{$i^2_n$}
+          & \qquad
+          \LabeledCrossingLR{.3}{$i^1_q$}{$i^2_p$}{$i^2_n$}{$i^1_n$}
           &\qquad
-          \LabeledCrossingLR{.3}{$i^1_p$}{$i^2_p$}{$i^2_r$}{$i^1_n$}
-          &\qquad
-          \LabeledCrossingRL{.3}{$i^2_p$}{$i^1_q$}{$i^1_n$}{$i^2_r$}
+          \LabeledInclusionRL{.3}{$i^2_p$}{$i^1_q$}{$i^2_n$}{$i^1_n$}
         \end{tabular}
       \end{center}
-      This is a contradiction since
-      $\vv{\mathcal{M}}$ avoids
-      \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR and
-      \CrossingRL.
-      \item
-      If $i^1_n < i^2_p$ then there exist $i^1_q$ and $i^2_r$ such
-      that
-      \begin{center}
-        \begin{tabular}{cc}
-          \LabeledInclusionLL{.3}{$i^1_q$}{$i^1_n$}{$i^2_p$}{$i^2_r$}
-          &\qquad
-          \LabeledInclusionRL{.3}{$i^2_r$}{$i^1_n$}{$i^2_p$}{$i^1_q$}
-        \end{tabular}
-      \end{center}
-      This is a contradiction since
-      $\vv{\mathcal{M}}$ avoids
-      \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR and
-      \CrossingRL.
-    \end{itemize}
-    Then it follows that $(i^1_n, i^2_n)$ is an arc
-    of $\vv{\mathcal{M}}$.
+    This is a contradiction since $\vv{\mathcal{M}}$ avoids \InclusionLR,
+    \CrossingRL, \CrossingLR, and \InclusionRL. Therefore, 
+    $(i^1_n, i^2_n)$ is an arc of $\vv{\mathcal{M}}$. By iteratively
+    applying the same reasoning, this also shows that all 
+    $(i^1_j, i^2_j)$, $j \in [n - 1]$, are arcs of $\vv{\mathcal{M}}$.
+%     The proof is by induction on $n$.
+%     Since the assertion is certainly valid for $n = 1$,
+%     it is enough to show that $(i^1_n, i^2_n)$ is an arc of $\vv{\mathcal{M}}$.
+%     Suppose, aiming at a contradiction that $(i^1_n, i^2_n)$
+%     is not an arc of $\vv{\mathcal{M}}$.
+%     Let $i^2_p$ be the sink of the arc with origin $i^1_n$ in
+%     $\vv{\mathcal{M}}$ (\emph{i.e.}, $(i^1_n, i^2_p)$ is an arc of
+%     $\vv{\mathcal{M}}$).
+%     We need to consider two cases.
+%     \begin{itemize}
+%       \item If $i^1_n > i^2_p$ then there exist $i^1_q$ and $i^2_r$ such
+%       that
+%       \begin{center}
+%         \begin{tabular}{cccc}
+%           \LabeledInclusionRL{.3}{$i^2_p$}{$i^1_q$}{$i^2_r$}{$i^1_n$}
+%           &\qquad
+%           \LabeledInclusionRR{.3}{$i^2_p$}{$i^2_r$}{$i^1_q$}{$i^1_n$}
+%           &\qquad
+%           \LabeledCrossingLR{.3}{$i^1_p$}{$i^2_p$}{$i^2_r$}{$i^1_n$}
+%           &\qquad
+%           \LabeledCrossingRL{.3}{$i^2_p$}{$i^1_q$}{$i^1_n$}{$i^2_r$}
+%         \end{tabular}
+%       \end{center}
+%       This is a contradiction since
+%       $\vv{\mathcal{M}}$ avoids
+%       \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR and
+%       \CrossingRL.
+%       \item
+%       If $i^1_n < i^2_p$ then there exist $i^1_q$ and $i^2_r$ such
+%       that
+%       \begin{center}
+%         \begin{tabular}{cc}
+%           \LabeledInclusionLL{.3}{$i^1_q$}{$i^1_n$}{$i^2_p$}{$i^2_r$}
+%           &\qquad
+%           \LabeledInclusionRL{.3}{$i^2_r$}{$i^1_n$}{$i^2_p$}{$i^1_q$}
+%         \end{tabular}
+%       \end{center}
+%       This is a contradiction since
+%       $\vv{\mathcal{M}}$ avoids
+%       \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR and
+%       \CrossingRL.
+%     \end{itemize}
+%     Then it follows that $(i^1_n, i^2_n)$ is an arc
+%     of $\vv{\mathcal{M}}$.
     Now, let $p^1 = \pi(i^1_1) \pi(i^1_2) \dots \pi(i^1_n)$ and
     $p^2 = \pi(j^2_1) \pi(^2_2) \dots \pi(i^2_n)$ be the two patterns
-    from $\pi$ induced by $I^1$ and $I^2$, respectively.
-    Clearly $p^1$ and $p^2$ are disjoint in $\pi$
-    (since $\vv{\mathcal{M}}$ is a matching)
-    and cover $\pi$ (since $\vv{\mathcal{M}}$ is perfect).
-    We claim that $p^1$ and $p^2$ are order-isomorphic.
-    Indeed, this is a direct consequence of th fact that,
-    for any two distinct edges $(a, b)$ and $(a', b')$ in
-    $\vv{\mathcal{M}}$ we have $\pi(a) < \pi(a')$ if and only if
-    $\pi(b) < \pi(b')$.
+    from $\pi$ induced by $I^1$ and $I^2$, respectively. Clearly $p^1$ 
+    and $p^2$ are disjoint in $\pi$ (since $\vv{\mathcal{M}}$ is a 
+    matching) and cover $\pi$ (since $\vv{\mathcal{M}}$ is perfect). We 
+    claim that $p^1$ and $p^2$ are order-isomorphic. Indeed, this is a 
+    direct consequence of th fact that, for any two distinct edges 
+    $(a, b)$ and $(a', b')$ in $\vv{\mathcal{M}}$ we have $\pi(a) < \pi(a')$ 
+    if and only if $\pi(b) < \pi(b')$.
     \qed

### 2016-02-03 new-article.tex
Lemma 3 (statement completed)

@@
     then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
+    $(\sigma_1, \sigma_2)$-arc or at most one $(\sigma_2, \sigma_1)$-arc.
   \end{enumerate}

### 2016-02-04 new-article.tex
LATIN 2016

@@
 graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows.
+
+\begin{definition}[Property $\mathbf{P_1}$]
+  \label{definition:Property P_1}
+  Let $\pi$ be a permutation and $\vv{\mathcal{M}}$ be an oriented perfect
+  matching of the linear graph $G_\pi$.
+  $\vv{\mathcal{M}}$ is said to have property $\mathbf{P_1}$ if it avoids
+  \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
+  and \CrossingRL.
+\end{definition}
+
+\begin{definition}[Property $\mathbf{P_2}$]
+  \label{definition:Property P_2}
+  Let $\pi$ be a permutation and $\vv{\mathcal{M}}$ be an oriented perfect
+  matching of the linear graph $G_\pi$.
+  $\vv{\mathcal{M}}$ is said to have property $\mathbf{P_2}$ if,
+  for any two distinct arcs $(a, a')$ and $(b, b')$ in $\vv{\mathcal{M}}$,
+  we have $\pi(a) < \pi(b)$ if and only if $\pi(a') < \pi(b')$.
+\end{definition}
+
+\begin{lemma}
+  \label{lemma:matching}
+  Let $\pi$ be a permutation. The following statements are equivalent:
+  \begin{enumerate}
+    \item The permutation $\pi$ is a square.
+    \item There exists an oriented perfect matching $\vv{\mathcal{M}}$
+    of the linear graph $G_\pi$ that has properties $\mathbf{P_1}$ and
+    $\mathbf{P_2}$.
+  \end{enumerate}
+\end{lemma}
 

### 2016-02-04 new-article.tex
LATIN 2016

@@
 \usepackage{esvect}
+\usepackage{todonotes}
+\usepackage[normalem]{ulem}
 
@@
     words. On the other hand, we prove that recognizing square
-    permutations is \NPC. This property is proved by establishing a
+    permutations is \NPC. \sout{This property is proved by establishing a
     reduction from the pattern involvement problem and by using a
     containment-free perfect matching criterion to decide if a
-    permutation is a square.
+    permutation is a square.}
+    \uline{This property is proved by establishing a
+    reduction from the pattern involvement problem and by using a
+    contrained oriented perfect matching criterion to decide if a
+    permutation is a square.}
 \end{abstract}
@@
 \label{section:Introduction}
-The {\em shuffle product} $\shuffle$ is a very famous operation on words
+The \emph{shuffle product} $\shuffle$ is a very famous operation on words
 first defined by Eilenberg and Mac Lane~\cite{Eilenberg:MacLane:1953}.
@@
 given three permutations $\pi$, $\sigma_1$, and $\sigma_2$, $\pi$ is
-said to be a {\em shuffle} of $\sigma_1$ and $\sigma_2$ if $\pi$ (viewed
+said to be a \emph{shuffle} of $\sigma_1$ and $\sigma_2$ if $\pi$ (viewed
 as a word) is a shuffle of two words that are order-isomorphic to
@@
 shuffle product of permutations $\SHUFFLE$ both from a combinatorial and
-from a computational point of view by focusing on {\em square}
+from a computational point of view by focusing on \emph{square}
 permutations, that are permutations $\pi$ being in the shuffle of a
@@
 shuffle products on permutations. For instance, in~\cite{DHT:IJAC:2002},
-the authors define the {\em shifted shuffle product}. For this product,
+the authors define the \emph{shifted shuffle product}. For this product,
 $\pi$ is a shuffle of $\sigma_1$ and $\sigma_2$ if $\pi$ is in the
@@
 preliminary definition of an operator~$\Delta$, allowing to
-{\em unshuffle} permutations. This operator is in fact a coproduct,
+\emph{unshuffle} permutations. This operator is in fact a coproduct,
 endowing the linear span of all permutations with a coalgebra structure
@@
 Let us recall here the most important ones. Let $u$ be a word. The
-length of $u$ is denoted by $|u|$. The {\em empty word}, the only word
+length of $u$ is denoted by $|u|$. The \emph{empty word}, the only word
 of null length, is denoted by $\epsilon$. We denote by $\widetilde{u}$
-the {\em mirror image} of $u$ that is the word
+the \emph{mirror image} of $u$ that is the word
 $u_{|u|} u_{|u| - 1} \dots u_1$. If $P$ is a subset of $[|u|]$, $u_{|P}$
@@
 is an integer, we denote by $u[k]$ the word obtained by incrementing by
-$k$ all letters of $u$. The {\em shuffle} of two words $u$ and $v$ is
+$k$ all letters of $u$. The \emph{shuffle} of two words $u$ and $v$ is
 the set recursively defined by
@@
 $ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
-$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
+$b$ are letters. A word $u$ is a \emph{square} if there exists a word $v$
 such that $u$ belongs to $v \shuffle v$.
@@
 $1 2 \dots n$ (resp. $n (n-1) \dots 1$). If $\pi$ is a permutation of
-$S_n$, we denote by $\bar \pi$ the {\em complementary} of $\pi$ that is
+$S_n$, we denote by $\bar \pi$ the \emph{complementary} of $\pi$ that is
 the permutation satisfying $\bar \pi(i) = n - \pi(i) + 1$ for all
-$i \in [n]$. The {\em inverse} of $\pi$ is denoted by $\pi^{-1}$.
+$i \in [n]$. The \emph{inverse} of $\pi$ is denoted by $\pi^{-1}$.
 
 If $u$ is a word of integers without multiple occurrences of a same
-letter, $\STD(u)$ is the {\em standarized} of $u$, that is the unique
+letter, $\STD(u)$ is the \emph{standarized} of $u$, that is the unique
 permutation of the same size as $u$ such that for all $i, j \in [|u|]$,
@@
 image of the map $\STD$ is the set $S$ of all permutations. Two words $u$
-and $v$ having the same standarized are {\em order-isomorphic}. If
-$\sigma$ is a permutation, there is an {\em occurrence} of (the
-{\em pattern}) $\sigma$ in $\pi$ if there is a set $P$ of indexes of
+and $v$ having the same standarized are \emph{order-isomorphic}. If
+$\sigma$ is a permutation, there is an \emph{occurrence} of (the
+\emph{pattern}) $\sigma$ in $\pi$ if there is a set $P$ of indexes of
 letters of $\pi$ such that $\sigma$ and $\pi_{|P}$ are order-isomorphic.
-When $\pi$ does not admit any occurrence of $\sigma$, $\pi$ {\em avoids}
+When $\pi$ does not admit any occurrence of $\sigma$, $\pi$ \emph{avoids}
 $\sigma$.
 
-Let us now provide some definitions about graphs and matching, used in
-the sequel. For a graph $G$, we denote $\VS{G}$ as the set of vertices
+Let us now provide some definitions about graphs and \uline{oriented}
+matching, used in the sequel.
+For a graph $G$, we denote $\VS{G}$ as the set of vertices
 and $\ES{G}$ as the set of edges of $G$. Let $u = u_1 u_2 \dots u_n$ be
@@
 edges. A matching is \emph{perfect} if it covers all the vertices of the
-graph. In case the set of vertices is equipped with a total order, a
+graph.
+\sout{In case the set of vertices is equipped with a total order, a
 matching $\mathcal{M}$ is said to be \emph{containment-free} if there do
@@
 $\mathcal{M}$ such that $u_i < u_k < u_\ell < u_j$ or
-$u_k < u_i < u_j < u_\ell$.
+$u_k < u_i < u_j < u_\ell$.}
+An \emph{oriented matching}
+$\vv{\mathcal{M}}$ of $\mathcal{M}$ is a directed matching obtained
+from $\mathcal{M}$ by orienting its edges.
+\textbf{D\'efinir l'inclusion de motifs. En fait, je pense même qu'il ne faudrait
+parler que de graphes orientés. Qu'en penses-tu?}
 
@@
 \end{equation}
-We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
+We call $\Delta$ the \emph{unshuffling coproduct of permutations}. For
 instance,
@@
 of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
-{\em shuffle product of permutations}. For instance,
+\emph{shuffle product of permutations}. For instance,
 \begin{equation}\begin{split} \label{equ:example_shuffle_product}
@@
 $\SHUFFLE$ and that this product is graded by the sizes of the
-permutations ({\em i.e.}, the product of a permutation of size $n$ with
+permutations (\emph{i.e.}, the product of a permutation of size $n$ with
 a permutation of size $m$ produces a sum of permutations of size $n + m$).
 Notice that there are other shuffle products on
-permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
-{\em convolution product}. Our shuffle product is different from these
+permutations~\cite{DHT:IJAC:2002}: the \emph{shifted shuffle} and the
+\emph{convolution product}. Our shuffle product is different from these
 last two.
@@
 
-We say that a permutation $\pi$ {\em appears} in the shuffle
+We say that a permutation $\pi$ \emph{appears} in the shuffle
 $\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
@@
 
-Besides, a permutation $\pi$ is a {\em square} if there is a
+Besides, a permutation $\pi$ is a \emph{square} if there is a
 permutation $\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
-In this case, we say that $\sigma$ is a {\em square root} of $\pi$.
+In this case, we say that $\sigma$ is a \emph{square root} of $\pi$.
 Equivalently, $\pi$ is a square with $\sigma$ as square root if and only

### 2016-02-09 new-article.tex
LATIN 2016 corrections

@@
 \newcommand{\LabeledCrossingLL}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledCrossingLR}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledCrossingRL}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledCrossingRR}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']%,>=stealth',->]
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledInclusionLL}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledInclusionLR}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledInclusionRL}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledInclusionRR}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledPrecedenceLL}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledPrecedenceLR}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledPrecedenceRL}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 \newcommand{\LabeledPrecedenceRR}[5]{%
-\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex']
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
 \draw node [] (U1)               {#2};
@@
 reduces to computing some specific matching in the associated linear
-graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows.
+graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}.
+We have, however, to deal with oriented perfect matchings.
+The needed properties read as follows.
 
@@
 \end{definition}
+
+The rationale for introducing properties $\mathbf{P_1}$ and $\mathbf{P_2}$
+stems from the following lemma.
 
@@
 
+% Before proving hardness, we give two easy lemmas that will prove extremely
+% useful for simplifying the proof of upcoming Proposition~\ref{proposition:hardness}.
+
 Let $\pi$ be a permutation. For the sake of clarity, we will say that a
@@
 $\sigma_1$ is below $\sigma_2$ (in an equivalent manner, $\sigma_2$ is
-above $\sigma_1$) in Fig.~\ref{subfig:decreasing before and below
-increasing}.
-
-Before proving hardness, we give an easy lemma that will prove extremely
-useful for simplifying the proof of upcoming Proposition~\ref{proposition:hardness}.
+above $\sigma_1$) in
+Fig.~\ref{subfig:decreasing before and below increasing}.
 
@@
   and $\vv{\mathcal{M}}$ be an oriented perfect matching of $G_\pi$
-  that avoids
-  \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
-  and \CrossingRL, and such that for any two distinct arcs $(a, a')$ and
-  $(b, b')$ in $\vv{\mathcal{M}}$ we have $\pi(a) < \pi(b)$ if and only
-  if $\pi(a') < \pi(b')$.
+  that has properties~$\mathbf{P_1}$ and $\mathbf{P_2}$.
   The following assertions hold:
@@
     (see Fig.~\ref{subfig:increasing before and above decreasing}),
-    then $\vv{\mathcal{M}}$ contains either at most one
-    $(\sigma_1, \sigma_2)$-arc or at most one
-    $(\sigma_2, \sigma_1)$-arc.
+    then there is at most one arc between $\sigma_1$ and
+    $\sigma_2$ in $\vv{\mathcal{M}}$ (this arc can be a
+    $(\sigma_1, \sigma_2)$-arc or a $(\sigma_2, \sigma_1)$-arc).
     \item
@@
     (see Fig.~\ref{subfig:decreasing before and below increasing}),
-    then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-arc or at most one $(\sigma_2, \sigma_1)$-arc.
+    then there is at most one arc between $\sigma_1$ and
+    $\sigma_2$ in $\vv{\mathcal{M}}$ (this arc can be a
+    $(\sigma_1, \sigma_2)$-arc or a $(\sigma_2, \sigma_1)$-arc).
   \end{enumerate}
@@
   Suppose, aiming at a contradiction, that the assertion does not hold.
-  Therefore, since $\vv{\mathcal{M}}$ avoids
-  \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
-  and \CrossingRL,
+  Therefore, since $\vv{\mathcal{M}}$ has Property~$\mathbf{P_1}$,
   $\vv{\mathcal{M}}$ contains either two crossing $(\sigma_1, \sigma_2)$-arcs
@@
 
+\begin{figure}[t!]
+  \centering
+  \subfigure[%
+    $\pi(a) < \pi(a')$.
+  ]{%
+    \begin{tikzpicture}
+      [
+        scale=0.3,
+        label/.style={anchor=base}
+      ]
+      \begin{scope}[]
+        % points
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
+        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (a)  at (1,-1.5)  {$a$};
+        \node [label] (ap) at (3,-1.5)  {$a'$};
+        \node [label] (b)  at (5,-1.5)  {$b$};
+        \node [label] (bp) at (7,-1.5) {$b'$};
+        \node (nu1) at (4,-2.5) {$\pi$};
+        % arcs
+        \draw [line width=1pt,->,>=latex']
+        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
+        \draw [line width=1pt,->,>=latex']
+        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
+      \end{scope}
+      \begin{scope}[yshift=12.5cm]
+        % points
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
+        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (a)  at (1,-1.5)  {$a$};
+        \node [label] (ap) at (3,-1.5)  {$a'$};
+        \node [label] (b)  at (5,-1.5)  {$b$};
+        \node [label] (bp) at (7,-1.5) {$b'$};
+        \node (nu1) at (4,-2.5) {$\pi$};
+        % arcs
+        \draw [line width=1pt,->,>=latex']
+        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
+        \draw [line width=1pt,<-,>=latex']
+        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
+      \end{scope}
+      %
+      \begin{scope}[xshift=10cm]
+        % points
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
+        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (a)  at (1,-1.5)  {$a$};
+        \node [label] (ap) at (3,-1.5)  {$a'$};
+        \node [label] (b)  at (5,-1.5)  {$b$};
+        \node [label] (bp) at (7,-1.5) {$b'$};
+        \node (nu1) at (4,-2.5) {$\pi$};
+        % arcs
+        \draw [line width=1pt,<-,>=latex']
+        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
+        \draw [line width=1pt,->,>=latex']
+        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
+      \end{scope}
+    %
+      \begin{scope}[xshift=10cm,yshift=12.5cm]
+        % points
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
+        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (a)  at (1,-1.5)  {$a$};
+        \node [label] (ap) at (3,-1.5)  {$a'$};
+        \node [label] (b)  at (5,-1.5)  {$b$};
+        \node [label] (bp) at (7,-1.5) {$b'$};
+        \node (nu1) at (4,-2.5) {$\pi$};
+        % arcs
+        \draw [line width=1pt,<-,>=latex']
+        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
+        \draw [line width=1pt,<-,>=latex']
+        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
+      \end{scope}
+    \end{tikzpicture}
+    \label{subfig:decreasing before and below increasing}
+  }% end subfigure
+  \qquad\;
+  \subfigure[%
+    $\pi(a) < \pi(a')$.
+  ]{%
+    \begin{tikzpicture}
+      [
+        scale=0.3,
+        label/.style={anchor=base}
+      ]
+      \begin{scope}[]
+        % points
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
+        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (a)  at (1,-1.5)  {$a$};
+        \node [label] (ap) at (3,-1.5)  {$a'$};
+        \node [label] (b)  at (5,-1.5)  {$b$};
+        \node [label] (bp) at (7,-1.5) {$b'$};
+        \node (nu1) at (4,-2.5) {$\pi$};
+        % arcs
+        \draw [line width=1pt,->,>=latex']
+        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
+        \draw [line width=1pt,->,>=latex']
+        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
+      \end{scope}
+      \begin{scope}[yshift=12.5cm]
+        % points
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
+        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (a)  at (1,-1.5)  {$a$};
+        \node [label] (ap) at (3,-1.5)  {$a'$};
+        \node [label] (b)  at (5,-1.5)  {$b$};
+        \node [label] (bp) at (7,-1.5) {$b'$};
+        \node (nu1) at (4,-2.5) {$\pi$};
+        % arcs
+        \draw [line width=1pt,->,>=latex']
+        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
+        \draw [line width=1pt,<-,>=latex']
+        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
+      \end{scope}
+      %
+      \begin{scope}[xshift=10cm]
+        % points
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
+        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (a)  at (1,-1.5)  {$a$};
+        \node [label] (ap) at (3,-1.5)  {$a'$};
+        \node [label] (b)  at (5,-1.5)  {$b$};
+        \node [label] (bp) at (7,-1.5) {$b'$};
+        \node (nu1) at (4,-2.5) {$\pi$};
+        % arcs
+        \draw [line width=1pt,<-,>=latex']
+        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
+        \draw [line width=1pt,->,>=latex']
+        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
+      \end{scope}
+    %
+      \begin{scope}[xshift=10cm,yshift=12.5cm]
+        % points
+        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (8.2,8.2);
+        \foreach \x/\y in {1/1,3/7,5/3,7/5} {
+            \draw [fill=black] (\x,\y) circle (0.1);
+        }
+        % labels
+        \node [label] (a)  at (1,-1.5)  {$a$};
+        \node [label] (ap) at (3,-1.5)  {$a'$};
+        \node [label] (b)  at (5,-1.5)  {$b$};
+        \node [label] (bp) at (7,-1.5) {$b'$};
+        \node (nu1) at (4,-2.5) {$\pi$};
+        % arcs
+        \draw [line width=1pt,<-,>=latex']
+        (1,1) .. controls +(0,8) and +(0,4) .. (3,7);
+        \draw [line width=1pt,<-,>=latex']
+        (5,3) .. controls +(0,4) and +(0,2) .. (7,5);
+      \end{scope}
+    \end{tikzpicture}
+    \label{subfig:decreasing before and below increasing}
+  }% end subfigure
+  \caption{\label{fig:subfig:value inclusion of arcs}%
+    Illustration of Lemma~\ref{lemma:value inclusion of arcs}
+    for $\min\{\pi(a), \pi(a')\} < \min\{\pi(b), \pi(b')\}$
+    and
+    $\max\{\pi(a), \pi(a')\} > \max\{\pi(b), \pi(b')\}$.
+  }
+\end{figure}
+
+\begin{lemma}
+  \label{lemma:value inclusion of arcs}
+  Let $\pi$ be a permutation
+  and $\vv{\mathcal{M}}$ be an oriented perfect matching of
+  the linear graph $G_\pi$.
+  If $\vv{\mathcal{M}}$ has Property $\mathbf{P_2}$ then it avoids patterns
+  $$
+  \LabeledPrecedenceLL{.3}{$a$}{$a'$}{$b$}{$b'$},
+  \qquad
+  \LabeledPrecedenceLR{.3}{$a$}{$a'$}{$b$}{$b'$},
+  \qquad
+  \LabeledPrecedenceRL{.3}{$a$}{$a'$}{$b$}{$b'$},
+  \qquad
+  \LabeledPrecedenceRR{.3}{$a$}{$a'$}{$b$}{$b'$}
+  $$
+  for
+  (i)
+  $\min\{\pi(a), \pi(a')\} < \min\{\pi(b), \pi(b')\}$
+  and
+  $\max\{\pi(a), \pi(a')\} > \max\{\pi(b), \pi(b')\}$
+  and
+  (ii)
+  $\min\{\pi(a), \pi(a')\} > \min\{\pi(b), \pi(b')\}$
+  and
+  $\max\{\pi(a), \pi(a')\} < \max\{\pi(b), \pi(b')\}$.
+\end{lemma}
+
+\begin{proof}[of Lemma~\ref{lemma:value inclusion of arcs}]
+  We only prove the lemma for case \emph{(i)}
+  since the proof for case \emph{(ii)} is similar.
+  Therefore, assume
+  $\min\{\pi(a), \pi(a')\} < \min\{\pi(b), \pi(b')\}$
+  and
+  $\max\{\pi(a), \pi(a')\} > \max\{\pi(b), \pi(b')\}$,
+  and suppose, aiming at a contradiction, that the assertion of the lemma
+  does not hold.
+  We need to consider two cases.
+  For each of the following patterns
+  $$
+  \LabeledPrecedenceLL{.3}{$a$}{$a'$}{$b$}{$b'$},
+  \qquad
+  \LabeledPrecedenceLR{.3}{$a$}{$a'$}{$b$}{$b'$},
+  $$
+  we have
+  $\pi(a) < \pi(b)$ and $\pi(a') > \pi(b')$,
+  and
+  for each of the following patterns
+  $$
+  \LabeledPrecedenceRL{.3}{$a$}{$a'$}{$b$}{$b'$},
+  \qquad
+  \LabeledPrecedenceRR{.3}{$a$}{$a'$}{$b$}{$b'$},
+  $$
+  we have
+  $\pi(a) > \pi(b)$ and $\pi(a') < \pi(b')$.
+  Hence, it follows that $\vv{\mathcal{M}}$ does not have
+  Property $\mathbf{P_2}$. This is a contradiction.
+  \qed
+\end{proof}
+
+\begin{lemma}
+  \label{lemma:for nu_1 and nu_2}
+  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
+  be a permutation where $\sigma_1$ is an increasing pattern of
+  length $k$, $k \geq 3$,  and $\sigma_2$ is below $\sigma_1$,
+  and $\vv{\mathcal{M}}$ be an oriented perfect matching of $G_\pi$
+  that has properties $\mathbf{P_1}$ and $\mathbf{P_2}$.
+  If $\vv{\mathcal{M}}$ contains either a $(\sigma_1, \sigma_2)$-arc
+  or a $(\sigma_2, \sigma_1)$-arc,
+  then it does not contain a $(\sigma_1, \sigma_1)$-arc.
+\end{lemma}
+
+\begin{proof}[of Lemma~\ref{lemma:for nu_1 and nu_2}]
+  Suppose first that $\vv{\mathcal{M}}$ contains a $(\sigma_1, \sigma_2)$-arc,
+  say $(a, a')$.
+  (The proof for the case $\vv{\mathcal{M}}$ contains a
+  $(\sigma_2, \sigma_1)$-arc is similar.)
+  Now, suppose, aiming at a contradiction, that the assertion does not hold and
+  that $\vv{\mathcal{M}}$ contains a $(\sigma_1, \sigma_1)$-arc, say $(b, b')$.
+  Since $\vv{\mathcal{M}}$ has Property $\mathbf{P_1}$, it avoids
+  \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
+  and \CrossingRL,
+  and hence $(a, a')$ and $(b, b')$ are either crossing arcs
+  with $b < a < b'$, or $(b, b')$ precedes $(a, a')$.
+  We consider the two cases.
+  \begin{itemize}
+    \item
+    If $(a, a')$ and $(b, b')$ are crossing arcs with $b < a < b'$,
+    then we have
+    $\pi(b) < \pi(a)$ (since $\sigma_1$ is increasing) and
+    $\pi(b') < \pi(a')$ (since $\sigma_1$ is above $\sigma_2$).
+    Then it follows that $\vv{\mathcal{M}}$ does not have
+    Property $\mathbf{P_2}$. This is a contradiction.
+    \item
+    If $(b, b')$ precedes $(a, a')$, then we have
+    $\min(b,b') > a'$ and $\max(b, b') < a$.
+    But $\sigma_1$ is above $\sigma_2$, and hence, according
+    to Lemma~\ref{lemma:value inclusion of arcs},
+    $\vv{\mathcal{M}}$ does not have
+    Property $\mathbf{P_2}$. This is a contradiction.
+  \end{itemize}
+  \qed
+\end{proof}
+
+\begin{lemma}
+  \label{lemma:reverse oriented matching}
+  Let $\pi$ be a permutation
+  and $\vv{\mathcal{M}}$ be an oriented perfect matching of
+  the linear graph $G_\pi$.
+  If $\vv{\mathcal{M}}$ has properties $\mathbf{P_1}$ and $\mathbf{P_2}$ then
+  so does the oriented perfect matching
+  $\vv{\mathcal{M}^r}$ obtained from $\vv{\mathcal{M}}$ by reversing
+  each arc.
+\end{lemma}
+
+\begin{proof}[of Lemma~\ref{lemma:reverse oriented matching}]
+  This is immediate for Property~$\mathbf{P_2}$,
+  since, for any two arcs $(a, a')$ and $(b,b')$ of $\vv{\mathcal{M}}$,
+  we have $\pi(a) < \pi(b)$ if and only if $\pi(a') < \pi(b')$.
+  As for Property~$\mathbf{P_1}$, it is enough to observe that
+  the set \InclusionLL,~\InclusionLR,~\InclusionRL,~\InclusionRR,~\CrossingLR
+  or \CrossingRL is closed by arc reversals
+  \qed
+\end{proof}
+
+A direct interpretation of Lemma~\ref{lemma:reverse oriented matching} is that,
+if a permutation $\pi$ is a square, one can exchange the roles of the two
+order-isomorphic patterns that cover $\pi$.
+An immediate but useful consequence of Lemma~\ref{lemma:reverse oriented matching}
+reads as follows.
+
+\begin{corollary}
+  \label{corollary:reverse oriented matching}
+  Let $\pi$ be a permutation and $a$ and $a'$ be two vertices of
+  the linear graph $G_\pi$.
+  There exists an oriented perfect matching of $G_\pi$ with
+  properties $\mathbf{P_1}$ and $\mathbf{P_2}$ that contains arc $(a,a')$
+  if and only of
+  there exists an oriented perfect matching of $G_\pi$ with
+  properties $\mathbf{P_1}$ and $\mathbf{P_2}$ that contains arc $(a',a)$
+\end{corollary}
+
+Having disposed of these preliminary observations,
+we now turn to proving hardness of the targeted problem.
+
 \begin{proposition}
@@
       ]
-      % edges
+      % arcs
       % pi 1 - pi 2
-      \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
+      \draw [pattern link,->,>=latex'] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.4,5);
       % pi 2 - sigma 2
-      \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
+      \draw [pattern link,->,>=latex'] (10.6,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
       % 1
-      \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
+      \draw [structure link,->,>=latex'] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
       \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
@@
       % 2
-      \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
+      \draw [structure link,->,>=latex'] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
       \node (claim edge 2) at (4,4) {Claim~\ref{claim:(nu_2, nu'_2)-edge}};
@@
       % 3
-      \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
+      \draw [structure link,->,>=latex'] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
       \node (claim edge 1) at (5.5,11.5) {Claim~\ref{claim:(nu_3, nu'_3)-edge}};
       % 4
-      \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
+      \draw [structure link,->,>=latex'] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
       \node (claim edge 1) at (8.5,8.5) {Claim~\ref{claim:(nu_4, nu'_4)-edge}};
       % sigma 1 - pi 1
-      \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
+      \draw [structure link,->,>=latex'] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
       % \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
@@
       in Proposition~\ref{proposition:hardness}.
-      Black edges denote the presence of at least one edge between two bunches of
+      Black arcs denote the presence of at least one arc between two bunches of
       positions in $\mu$.
-      Grey edges denote edges that are  only considered in the forward direction of the
+      Grey arcs denote edges that are only considered in the forward direction of the
       proof.
@@
   We claim that $\sigma$ occurs in $\pi$ if and only if
-  there exists a containment-free perfect matching
-  $\mathcal{M}$ in $G_\mu$ such that
-  for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+  there exists an oriented perfect matching
+  $\vv{\mathcal{M}}$ of $G_\mu$ that has properties $\mathbf{P_1}$ and
+  $\mathbf{P_2}$.
 
   Suppose first that $\sigma$ occurs in $\pi$ and fix an occurrence.
-  Construct a matching $\mathcal{M}$ as follows:
+  Construct an oriented matching $\vv{\mathcal{M}}$ as follows
+  (all arcs are oriented to the left):
   \begin{itemize}
     \item $\mathcal{M}$ contains $N_1$ pairwise crossing
-    $(\nu_1, \nu'_1)$-edges.
+    $(\nu_1, \nu'_1)$-arcs.
     \item $\mathcal{M}$ contains $N_2$ pairwise crossing
-    $(\nu_2, \nu'_2)$-edges.
+    $(\nu_2, \nu'_2)$-arcs.
     \item $\mathcal{M}$ contains $N_3$ pairwise crossing
-    $(\nu_3, \nu'_3)$-edges.
+    $(\nu_3, \nu'_3)$-arcs.
     \item $\mathcal{M}$ contains $N_4$ pairwise crossing
-    $(\nu_4, \nu'_4)$-edges.
+    $(\nu_4, \nu'_4)$-arcs.
     \item $\mathcal{M}$ contains $k+2$ pairwise crossing
-    $(\sigma', \pi')$-edges as depicted in
+    $(\sigma', \pi')$-arcs as depicted in
     Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
@@
     and all other positions in $\sigma'$ are linked by means of $k$ pairwise
-    crossing edges to the positions in
+    crossing arcs to the positions in
     $\pi'$ that correspond to the fixed occurence of $\sigma$ in $\pi$.
@@
     \item $\mathcal{M}$ contains $n-k$ pairwise crossing
-    $(\pi', \pi'')$-edges as depicted in
+    $(\pi', \pi'')$-arcs as depicted in
     Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
@@
     all positions in $\pi'$ that do not correspond to the fixed occurence of
-    $\sigma$ in $\pi$ are linked by means of $n-k$ pairwise crossing edges
+    $\sigma$ in $\pi$ are linked by means of $n-k$ pairwise crossing arcs
     to the positions in $\pi''$ that do not correspond to the fixed
@@
     \item $\mathcal{M}$ contains $k$ pairwise crossing
-    $(\pi'', \sigma'')$-edges as depicted in
+    $(\pi'', \sigma'')$-arcs as depicted in
     Figure~\ref{fig:subfig:sigma' - pi' - pi'' - sigma''}.
@@
     the fixed occurrence of $\sigma$ in $\pi$ are llinked
-    linked by means of $k$ pairwise crossing edges to all positions in
+    linked by means of $k$ pairwise crossing arcs to all positions in
     $\sigma''$.
-    (Notice that we use here the fact that $\sigma$ occurs in $\pi$).
+    (Notice that, again, we use here the fact that $\sigma$ occurs in $\pi$).
   \end{itemize}
   It can be easily checked (probably referring to Figure~\ref{fig:reduction}) that
-  $\mathcal{M}$ is perfect, containment-free and that
-  for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+  $\vv{\mathcal{M}}$ is perfect and has
+  properties $\mathbf{P_1}$ and $\mathbf{P_2}$.
 
@@
           % edges sigma' - pi'
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (1,20) .. controls +(0,9) and +(0,4) .. (6,15);
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (5,21) .. controls +(0,9) and +(0,4) .. (13,16);
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (2,19) .. controls +(0,-17) and +(0,-7) .. (8,13);
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (3,17) .. controls +(0,-17) and +(0,-7) .. (9,10);
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (4,18) .. controls +(0,-17) and +(0,-7) .. (11,11);
           % edges pi' - pi''
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (7,12) .. controls +(0,9) and +(0,4) .. (14,6);
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (10,9) .. controls +(0,9) and +(0,4) .. (17,3);
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (12,14) .. controls +(0,9) and +(0,4) .. (19,8);
           % edges pi'' - sigma''
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (15,7) .. controls +(0,-9) and +(0,-4) .. (20,2);
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (16,4) .. controls +(0,-9) and +(0,-4) .. (21,0);
-          \draw [line width=1pt]
+          \draw [line width=1pt,->,>=latex']
           (18,5) .. controls +(0,-9) and +(0,-4) .. (22,1);
-          % text
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(0.5cm,-.75cm)}
-          % ] at (1,27)
-          % {%
-          % $2$ pairwise crossing $(\sigma',\pi')$-edges that
-          % link the guards
-          % };%
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(-0.1cm,-2.5cm)}
-          % ] at (17,24)
-          % {%
-          % $n-k$ pairwise crossing $(\sigma',\pi')$-edges that
-          % denote positions that do not correspond to
-          % a specific occurence of $\sigma$ in $\pi$
-          % (at positions $2$, $3$ and $5$).
-          % };%
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(0.75cm,1.5cm)}
-          % ] at (2.5,4,5)
-          % {%
-          % $k$ pairwise crossing $(\sigma',\pi')$-edges that
-          % denote a specific occurence of $\sigma$ in $\pi$
-          % (at positions $2$, $3$ and $5$).
-          % };%
-          % \node[%
-          %   draw,
-          %   text width=5.5cm,text justified,color=black!50,
-          %   rectangle callout,callout relative pointer={(2cm,.75cm)}
-          % ] at (7,-3)
-          % {%
-          % $k$ pairwise crossing $(\pi''; \sigma'')$-edges that
-          % denote a specific occurence of $\sigma$ in $\pi$
-          % (at positions $2$, $3$ and $5$).
-          % };%
         \end{tikzpicture}
@@
 
-  Conversely, suppose that there exists a containment-free perfect matching
-  $\mathcal{M}$ in $G_\mu$ such that for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+  Conversely, suppose that there exists an oriented perfect matching
+  $\vv{\mathcal{M}}$ in $G_\mu$ that has
+  properties $\mathbf{P_1}$ and $\mathbf{P_2}$.
   We show that $\sigma$ occurs as a pattern in $\pi$.
-  Whereas the matching $\mathcal{M}$ may not be as regular as in the forward
-  direction, the main idea is to prove that $\mathcal{M}$ contains enough structure
-  (more precisely, $k+2$ $(\sigma', \pi')$-edges) so that we can conclude
+  Whereas the oriented matching $\vv{\mathcal{M}}$ may not be as regular as in the
+  forward direction, the main idea is to prove that $\vv{\mathcal{M}}$ contains enough
+  structure
+  (more precisely, $k+2$ $(\sigma', \pi')$-arcs) so that we can conclude
   that $\sigma$ occurs in $\pi$.
   We have divided the reverse direction into a serie of basic claims that
-  progressively defines the overall structure of $\mathcal{M}$.
+  progressively defines (and refines)
+  the overall structure of $\vv{\mathcal{M}}$.
 
@@
     \label{claim:no (nu_1, nu_2)-edge}
-    There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
+    There is no $(\nu_1, \nu_2)$-arc nor $(\nu_2, \nu_1)$-arc in
+    $\vv{\mathcal{M}}$.
   \end{claim}
@@
     First, according to Lemma~\ref{lemma:at most one edge monotone},
-    there exists at most one $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
+    there exists at most one arc between $\nu_1$ and $\nu_2$ in $\vv{\mathcal{M}}$
+    (regardless of the orientation).
     Suppose now, aiming at a contradiction, that there exists
-    exactly one $(\nu_1, \nu_2)$-edge $e = (a, a')$ in $\mathcal{M}$.
-    Under this assumption, we show that $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
-    Indeed, suppose, aiming at a contradiction,
-    that there exists a $(\nu_1, \nu_1)$-edge $(x, x')$ in $\mathcal{M}$.
-    Since $\mathcal{M}$ is containment-free,
-    we must have $x < a$.
-    Then it follows that $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing) and
-    $\mu(x') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
-    which contradicts our hypothesis about $\mathcal{M}$.
-    Hence $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
-    Furthermore, since $\nu_1$ is increasing and
-    there exists only one $(\nu_1, \nu_2)$-edge $(a, a')$ in $\mathcal{M}$,
+    exactly one arc between $\nu_1$ and $\nu_2$ in $\vv{\mathcal{M}}$.
+    According to Corollary~\ref{corollary:reverse oriented matching} there is no
+    loss of generality in assuming that this is a
+    $(\nu_1, \nu'_1)$-arc, say $(a, a')$.
+    Furtheremore, according to Lemma~\ref{lemma:for nu_1 and nu_2},
+    there is no $(\nu_1, \nu_1)$-arc in $\vv{\mathcal{M}}$.
+    Now, since $\nu_1$ is increasing and there exists
+    exactly one $(\nu_1, \nu_2)$ in $\vv{\mathcal{M}}$,
     we conclude that
-    $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
-    But $|\nu_1| = |\nu'_1| = N_1$ and $\mathcal{M}$ is perfect,
+    $\vv{\mathcal{M}}$ contains $N_1-1$ pairwise crossing
+    $(\nu_1, \nu'_1)$-arcs
+    (this follows from the fact that $\vv{\mathcal{M}}$ does not contain \CrossingLR
+    nor \CrossingRL).
+    But $|\nu_1| = |\nu'_1| = N_1$ and $\vv{\mathcal{M}}$ is perfect,
     and hence there exists a position, say $y$, in $\nu'_1$ that is
-    not involved in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+    not involved in a $(\nu_1, \nu'_1)$-arc in $\vv{\mathcal{M}}$.
     We rule out this configuration by considering two cases:
     \begin{itemize}
-      \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-edge,
-      say $(y', y)$. Then, we have $\mu(a) > \mu(y')$ (since $\nu_1$ is above $\nu_2$)
+      \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-arc,
+      say $(y', y)$.
+      Then, we have $\mu(a) > \mu(y')$ (since $\nu_1$ is above $\nu_2$)
       and $\mu(a') < \mu(y)$ (since $\nu_2$ is below $\nu'_1$),
-      which contradicts our hypothesis about $\mathcal{M}$.
+      which contradicts Property~$\mathbf{P_2}$.
       See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
-      \item Position $y$ is not involved in a $(\nu_2, \nu'_1)$-edge.
-      Since $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges
-      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge $(y, y'')$ in $\mathcal{M}$
-      that is neither a $(\nu_1, \nu'_1)$-edge nor a $(\nu_2, \nu'_1)$-edge
-      nor a $(\nu'_1, \nu'_1)$-edge.
-      In other words, $y'' > 2N_1 + N_2$.
-      Now, for any $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$, we
-      have $\mu(x) < \mu(y)$ (since $\nu_1$ is below $\nu'_1$)
-      and $\mu(x') > \mu(y'')$ (since $\nu'_1$ is above any other gadget),
-      which contradicts our hypothesis about $\mathcal{M}$.
+      \item
+      Position $y$ is involved in a $(\nu'_1, \nu_2)$-edge,
+      say $(y', y)$.
+      According to Lemma~\ref{lemma:value inclusion of arcs},
+      $(a, a')$ and $(y, y')$ must be crossing
+      (\emph{i.e.}, $y < a'$).
+      But it follows that $\vv{\mathcal{M}}$ contain \CrossingLR
+      which contradicts Property~$\mathbf{P_1}$.
+      \item
+      $(y, y'')$ in an arc of $\vv{\mathcal{M}}$ for some
+      $y'' > 2N_1 + N_2$.
+      But $\nu'_1$ is above any other gadget, and hence
+      applying Lemma~\ref{lemma:value inclusion of arcs} to
+      arcs
+      $(a, a')$ and $(y, y'')$ yields a contrdiction.
       See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 2}.

### 2016-02-10 submission.tex
Modif LATIN16

@@
 }% end \MWithContainment
+
+\newcommand{\Crossing}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \Crossing
+
+\newcommand{\LabeledCrossingLL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingLL
+
+\newcommand{\CrossingLL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U3.north) + (0,2) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingLL
+
+\newcommand{\LabeledCrossingLR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingLR
+
+\newcommand{\CrossingLR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingLR
+
+\newcommand{\LabeledCrossingRL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingRL
+
+\newcommand{\CrossingRL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingRL
+
+\newcommand{\LabeledCrossingRR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingRR
+
+\newcommand{\CrossingRR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingRR
+
+\newcommand{\Inclusion}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \Inclusion
+
+\newcommand{\LabeledInclusionLL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionLL
+
+\newcommand{\InclusionLL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
++\draw [->] (U4.north) .. controls ($ (U4.north) + (0,4) $) and ($ (U1.north) + (0,4) $) .. (U1.north);
++\draw [->] (U3.north) .. controls ($ (U3.north) + (0,3) $) and ($ (U2.north) + (0,3) $) .. (U2.north);
+\end{tikzpicture}
+}% end \InclusionLL
+
+\newcommand{\LabeledInclusionLR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionLR
+
+\newcommand{\InclusionLR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionLR
+
+\newcommand{\LabeledInclusionRL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionRL
+
+\newcommand{\InclusionRL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionRL
+
+\newcommand{\LabeledInclusionRR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionRR
+
+\newcommand{\InclusionRR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionRR
+
+\newcommand{\Precedence}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \Precedence
+
+\newcommand{\LabeledPrecedenceLL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLL
+
+\newcommand{\PrecedenceLL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLL
+
+\newcommand{\LabeledPrecedenceLR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLR
+
+\newcommand{\PrecedenceLR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLR
+
+\newcommand{\LabeledPrecedenceRL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRL
+
+\newcommand{\PrecedenceRL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRL
+
+\newcommand{\LabeledPrecedenceRR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRR
+
+\newcommand{\PrecedenceRR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRR
 
@@
     between square $(213,231)$-avoiding permutations and square binary
-    words. On the other hand, using a containment-free perfect matching
-    criterion, we prove that recognizing square permutations is \NPC.
+    words. On the other hand, by using a pattern avoidance criterion on
+    oriented perfect matchings, we prove that recognizing square 
+    permutations is \NPC.
 \end{abstract}
@@
 
-Let us now provide some definitions about graphs and matching that are used in
-the sequel. For a graph $G$, we denote $\VS{G}$ as the set of vertices
-and $\ES{G}$ as the set of edges of $G$. Let $u = u_1 u_2 \dots u_n$ be
-a word. The \emph{clique graph associated with $u$}, denoted by $K_u$, is
-the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$. The
-structure of this underlying graph is linear, \emph{i.e.}, the set of
-its vertices is equipped with a natural total order $<$ defined by
-$u_i < u_j$ if and only if $i < j$. For a linear graph, we write
-$(u_i, u_j)$ instead of $\{u_i, u_j\}$ for an edge if it is clear from
-the context that $i < j$. Recall that two edges of a graph are
-\emph{independent} if they do not share a common vertex, and that a
-\emph{matching} $\mathcal{M}$ in $G$ is a set of pairwise independent
-edges. A matching is \emph{perfect} if it covers all the vertices of the
-graph.
-For an edge $e = \{u,v\}$, we shall occasionally say that $u$ and $v$
-are \emph{involved} in $e$.
-In case the set of vertices is equipped with a total order, a
-matching $\mathcal{M}$ is said to be \emph{containment-free} if there do
-not exist (independent) edges $(u_i, u_j)$ and $(u_k, u_\ell)$ in
-$\mathcal{M}$ such that $u_i < u_k < u_\ell < u_j$
-%or $u_k < u_i < u_j < u_\ell$
-(\emph{i.e.}, $\mathcal{M}$ avoids~\MWithContainment).
+Let us now provide some definitions about graphs and oriented perfect 
+matchings that are used in the sequel. If $G$ is an oriented graph 
+without loops, two different edges of $G$ are {\em independent} if they 
+do not share any common vertex. We say that $G$ is an 
+{\em oriented matching} if all edges of $G$ are pairwise independent. 
+Moreover, $G$ is {\em perfect} if all vertices of $G$ belong to at least 
+one arc. For any word $u = u_1 u_2 \dots u_n$, an 
+{\em oriented perfect matching on $u$} is an oriented perfect matching 
+$\mathcal{M}$ on the set of vertices $\{u_1, u_2, \dots, u_n\}$. In the 
+sequel, we shall consider a natural notion of pattern avoidance in 
+oriented perfect matchings on words. For instance, an oriented perfect
+matching $\mathcal{M}$ of a word $u$ {\em admits an occurrence} of the 
+pattern $\CrossingRL$ if there are four positions $i < j < k < \ell$ in 
+$u$ such that $(u_k, u_i)$ and $(u_j, u_\ell)$ are arcs of $\mathcal{M}$. 
+When $\mathcal{M}$ does not admit any occurrence of a pattern 
+$\mathcal{P}$, we say that $\mathcal{M}$ {\em avoids}~$\mathcal{P}$.
+
+% Let us now provide some definitions about graphs and oriented matching 
+% that are used in the sequel. For a graph $G$, we denote $\VS{G}$ as the 
+% set of vertices and $\ES{G}$ as the set of edges of $G$. Let 
+% $u = u_1 u_2 \dots u_n$ be a word. The \emph{clique graph associated with $u$},
+% denoted by $K_u$, is the complete graph on 
+% $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$. The structure of this underlying 
+% graph is linear, \emph{i.e.}, the set of
+% its vertices is equipped with a natural total order $<$ defined by
+% $u_i < u_j$ if and only if $i < j$. For a linear graph, we write
+% $(u_i, u_j)$ instead of $\{u_i, u_j\}$ for an edge if it is clear from
+% the context that $i < j$. Recall that two edges of a graph are
+% \emph{independent} if they do not share a common vertex, and that a
+% \emph{matching} $\mathcal{M}$ in $G$ is a set of pairwise independent
+% edges. A matching is \emph{perfect} if it covers all the vertices of the
+% graph.
+% For an edge $e = \{u,v\}$, we shall occasionally say that $u$ and $v$
+% are \emph{involved} in $e$.
+% In case the set of vertices is equipped with a total order, a
+% matching $\mathcal{M}$ is said to be \emph{containment-free} if there do
+% not exist (independent) edges $(u_i, u_j)$ and $(u_k, u_\ell)$ in
+% $\mathcal{M}$ such that $u_i < u_k < u_\ell < u_j$
+% %or $u_k < u_i < u_j < u_\ell$
+% (\emph{i.e.}, $\mathcal{M}$ avoids~\MWithContainment).
 
@@
 reduces to computing some specific matching in the associated linear
-graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. It reads as follows
-(see Fig.\ref{fig:example containment-free matching}).
+graph~\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}. We have, however, 
+to deal with oriented perfect matchings. The needed properties read 
+as follows (see Fig.~\ref{fig:example containment-free matching}).
 \begin{figure}[ht!]
@@
     \begin{tikzpicture}[scale=0.5,inner sep=2pt,node distance=1cm]
-    \draw node (K) {$K_\pi$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=K]   (U01) {$2$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U01] (U02) {$A$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U02] (U03) {$4$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U03] (U04) {$7$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U04] (U05) {$1$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U05] (U06) {$9$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U06] (U07) {$5$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U07] (U08) {$3$};
-    \draw node [draw, circle, minimum size=.55cm, fill=black!10] [right of=U08] (U09) {$8$};
-    \draw node [draw, circle, minimum size=.55cm] [right of=U09] (U10) {$6$};
-    \draw [thick]
-      (U01.north) .. controls ($ (U01.north) + (0,2.25) $) and ($ (U03.north) + (0,2.25) $) .. (U03.north);
-    \draw [thick]
-      (U02.north) .. controls ($ (U02.north) + (0,2.25) $) and ($ (U06.north) + (0,2.25) $) .. (U06.north);
-    \draw [thick]
-      (U04.north) .. controls ($ (U04.north) + (0,2.5) $) and ($ (U07.north) + (0,2.5) $) .. (U07.north);
-    \draw [thick]
-      (U05.north) .. controls ($ (U05.north) + (0,2.5) $) and ($ (U08.north) + (0,2.5) $) .. (U08.north);
-    \draw [thick]
-      (U09.north) .. controls ($ (U09.north) + (0,2.5) $) and ($ (U10.north) + (0,2.5) $) .. (U10.north);
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10](U01){$1$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U01]          
+            (U02) {$8$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U02](U03){$3$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U03] 
+            (U04) {$9$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U04]
+            (U05) {$2$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U05](U06){$7$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U06](U07){$B$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U07](U08){$5$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U08](U09){$C$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U09](U10){$6$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U10]
+            (U11) {$A$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U11]
+            (U12) {$4$};
+        \draw [thick,->]
+        (U01.north) .. controls ($ (U01.north) + (0,2.25) $) 
+            and ($ (U03.north) + (0,2.25) $) .. (U03.north);
+        \draw [thick,->]
+        (U02.north) .. controls ($ (U02.north) + (0,2.25) $) 
+            and ($ (U06.north) + (0,2.25) $) .. (U06.north);
+        \draw [thick,->]
+        (U04.north) .. controls ($ (U04.north) + (0,2.5) $) 
+            and ($ (U07.north) + (0,2.5) $) .. (U07.north);
+        \draw [thick,->]
+        (U05.north) .. controls ($ (U05.north) + (0,2.5) $) 
+            and ($ (U08.north) + (0,2.5) $) .. (U08.north);
+        \draw [thick,->]
+        (U11.north) .. controls ($ (U11.north) + (0,2.5) $) 
+            and ($ (U09.north) + (0,2.5) $) .. (U09.north);
+        \draw [thick,->]
+        (U12.north) .. controls ($ (U12.north) + (0,2.5) $) 
+            and ($ (U10.north) + (0,2.5) $) .. (U10.north);
     \end{tikzpicture}
     \caption{\label{fig:example containment-free matching}%
-    The linear graph $K_\pi$ associated with the permutation
-    $\pi = 2A47195386$ together with
-    a containment-free perfect matching $\mathcal{M}$
-    such that for any two distinct edges $(a, b)$ and
-    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
-    if $\pi(b) < \pi(b')$.
-    From $\mathcal{M}$, it follows that $\pi$ is square as it is in
-    the shuffle of $2A718$ and $49536$, both being
-    order-isomorphic to $25314$.
+        An oriented perfect matching $\mathcal{M}$ on the permutation
+        $\pi = 183927B5C6A4$ satisfying the properties $\mathbf{P_1}$ 
+        and $\mathbf{P_2}$. From $\mathcal{M}$, it follows that $\pi$ is 
+        a square as it appears in the shuffle of $1892A4$ and $37B5C6$, 
+        both being order-isomorphic to $145263$.
     }% end caption
@@
 
+\begin{definition}[Property $\mathbf{P_1}$]
+  \label{definition:Property P_1}
+  Let $\pi$ be a permutation. An oriented perfect matching $\mathcal{M}$
+  on $\pi$ is said to have property $\mathbf{P_1}$ if it avoids the 
+  patterns \InclusionLL, \InclusionLR, \InclusionRL, 
+  \InclusionRR, \CrossingLR, and \CrossingRL.
+\end{definition}
+
+\begin{definition}[Property $\mathbf{P_2}$]
+  \label{definition:Property P_2}
+  Let $\pi$ be a permutation. An oriented perfect matching $\mathcal{M}$
+  on $\pi$ is said to have property $\mathbf{P_2}$ if, for any two 
+  distinct arcs $(a, a')$ and $(b, b')$ in $\mathcal{M}$,
+  we have $\pi(a) < \pi(b)$ if and only if $\pi(a') < \pi(b')$.
+\end{definition}
+
+The rationale for introducing properties $\mathbf{P_1}$ and $\mathbf{P_2}$
+stems from the following lemma.
+
+\begin{lemma}
+  \label{lemma:matching}
+  Let $\pi$ be a permutation. The following statements are equivalent:
+  \begin{enumerate}
+    \item The permutation $\pi$ is a square.
+    \item There exists an oriented perfect matching $\mathcal{M}$
+    on $\pi$ satisfying~$\mathbf{P_1}$ and~$\mathbf{P_2}$.
+  \end{enumerate}
+\end{lemma}
+
+Let $\pi$ be a permutation. For the sake of clarity, we will say that a
+bunch of consecutive positions $P$ of $\pi$ is \emph{above} (resp.
+\emph{below}) above another bunch of consecutive positions $P'$ in $\pi$
+if $\pi(i) > \pi(j)$ (resp. $\pi(i) < \pi(j)$) for every $i \in P$ and
+every $j \in P'$. For example, $\sigma_1$ is above $\sigma_2$ (in an
+equivalent manner, $\sigma_2$ is below $\sigma_1$) in
+Fig.~\ref{subfig:increasing before and above decreasing}, whereas
+$\sigma_1$ is below $\sigma_2$ (in an equivalent manner, $\sigma_2$ is
+above $\sigma_1$) in Fig.~\ref{subfig:decreasing before and below
+increasing}.
 \begin{figure}[t!]
@@
       [
-        scale=0.325,
+        scale=0.2,
         label/.style={anchor=base}
@@
       % labels
-      \node [label] (a) at (1,5) {$a$};
-      \node [label] (b) at (4,5) {$b$};
-      \node [label] (ap) at (7,-1) {$a'$};
-      \node [label] (bp) at (10,-1) {$b'$};
-      \node (nu1) at (2.5,4) {$\sigma_1$};
-      \node (nu2) at (8.5,-2) {$\sigma_2$};
+      \node [label] (a) at (1,4) {$a$};
+      \node [label] (b) at (4,4) {$b$};
+      \node [label] (ap) at (7,-2) {$a'$};
+      \node [label] (bp) at (10,-2) {$b'$};
+      \node (nu1) at (2.5,3) {$\sigma_1$};
+      \node (nu2) at (8.5,-3) {$\sigma_2$};
       % crossings edges
-      \draw [line width=1pt]
+      \draw [line width=1pt,->,>=latex']
       (1,7) .. controls +(0,12) and +(0,10) .. (7,4);
-      \draw [line width=1pt]
+      \draw [line width=1pt,->,>=latex']
+      (4,10) .. controls +(0,12) and +(0,10) .. (10,1);
+    \end{tikzpicture}
+    \;
+    \begin{tikzpicture}
+      [
+        scale=0.2,
+        label/.style={anchor=base}
+      ]
+      % increasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
+      \foreach \x/\y in {0/6,1/7,2/8,3/9,4/10,5/11} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % decreasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
+      \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % labels
+      \node [label] (a) at (1,4) {$a'$};
+      \node [label] (b) at (4,4) {$b'$};
+      \node [label] (ap) at (7,-2) {$a$};
+      \node [label] (bp) at (10,-2) {$b$};
+      \node (nu1) at (2.5,3) {$\sigma_1$};
+      \node (nu2) at (8.5,-3) {$\sigma_2$};
+      % crossings edges
+      \draw [line width=1pt,<-,>=latex']
+      (1,7) .. controls +(0,12) and +(0,10) .. (7,4);
+      \draw [line width=1pt,<-,>=latex']
       (4,10) .. controls +(0,12) and +(0,10) .. (10,1);
@@
   }% end subfigure
-  \qquad\qquad\qquad
+  \qquad
   \subfigure[%
@@
     [
-      scale=0.325,
+      scale=0.2,
       label/.style={anchor=base}
@@
     % labels
-    \node [label] (a) at (1,-1) {$a$};
-    \node [label] (b) at (4,-1) {$b$};
-    \node [label] (ap) at (7,5) {$a'$};
-    \node [label] (bp) at (10,5) {$b'$};
-    \node (nu1) at (2.5,-2) {$\sigma_1$};
-    \node (nu2) at (8.5,4) {$\sigma_2$};
+    \node [label] (a) at (1,-2) {$a$};
+    \node [label] (b) at (4,-2) {$b$};
+    \node [label] (ap) at (7,4) {$a'$};
+    \node [label] (bp) at (10,4) {$b'$};
+    \node (nu1) at (2.5,-3) {$\sigma_1$};
+    \node (nu2) at (8.5,3) {$\sigma_2$};
     % crossings edges
-    \draw [line width=1pt]
+    \draw [line width=1pt,->,>=latex']
     (1,4) .. controls +(0,10) and +(0,12) .. (7,7);
-    \draw [line width=1pt]
+    \draw [line width=1pt,->,>=latex']
+    (4,1) .. controls +(0,10) and +(0,12) .. (10,10);
+    \end{tikzpicture}
+    \;
+    \begin{tikzpicture}
+      [
+        scale=0.2,
+        label/.style={anchor=base}
+      ]
+      % decreasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (5.2,5.2);
+      \foreach \x/\y in {0/5,1/4,2/3,3/2,4/1,5/0} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % decreasing pattern
+      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,5.8) grid (11.2,11.2);
+      \foreach \x/\y in {6/6,7/7,8/8,9/9,10/10,11/11} {
+          \draw [fill=black] (\x,\y) circle (0.1);
+      }
+      % labels
+      \node [label] (a) at (1,-2) {$a'$};
+      \node [label] (b) at (4,-2) {$b'$};
+      \node [label] (ap) at (7,4) {$a$};
+      \node [label] (bp) at (10,4) {$b$};
+      \node (nu1) at (2.5,-3) {$\sigma_1$};
+      \node (nu2) at (8.5,3) {$\sigma_2$};
+      % crossings edges
+      \draw [line width=1pt,<-,>=latex']
+      (1,4) .. controls +(0,10) and +(0,12) .. (7,7);
+      \draw [line width=1pt,<-,>=latex']
       (4,1) .. controls +(0,10) and +(0,12) .. (10,10);
@@
 
-\begin{lemma}
-  \label{lemma:matching}
-  Let $\pi$ be a permutation. The following statements are equivalent:
-  \begin{enumerate}
-    \item The permutation $\pi$ is a square.
-    \item There exists a perfect containment-free matching $\mathcal{M}$
-    of $K_\pi$ such that for any two distinct edges $(a, b)$ and
-    $(a', b')$ in $\mathcal{M}$ we have $\pi(a) < \pi(a')$ if and only
-    if $\pi(b) < \pi(b')$.
-  \end{enumerate}
-\end{lemma}
-
-Let $\pi$ be a permutation. For the sake of clarity, we will say that a
-bunch of consecutive positions $P$ of $\pi$ is \emph{above} (resp.
-\emph{below}) above another bunch of consecutive positions $P'$ in $\pi$
-if $\pi(i) > \pi(j)$ (resp. $\pi(i) < \pi(j)$) for every $i \in P$ and
-every $j \in P'$. For example, $\sigma_1$ is above $\sigma_2$ (in an
-equivalent manner, $\sigma_2$ is below $\sigma_1$) in
-Fig.~\ref{subfig:increasing before and above decreasing}, whereas
-$\sigma_1$ is below $\sigma_2$ (in an equivalent manner, $\sigma_2$ is
-above $\sigma_1$) in Fig.~\ref{subfig:decreasing before and below
-increasing}.
-
 Before proving hardness, we give an easy lemma that will prove extremely
@@
   be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
-  and $\mathcal{M}$ be a containment-free perfect matching of $K_\pi$.
+  and $\mathcal{M}$ be an oriented perfect matching on $\pi$
+  satisfying~$\mathbf{P_1}$ and $\mathbf{P_2}$.
   The following assertions hold:
@@
     If $\sigma_1$ is increasing, $\sigma_2$ is decreasing, and
-    $\sigma_1$ is above $\sigma_2$ (see Fig.~\ref{subfig:increasing before
-    and above decreasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
+    $\sigma_1$ is above $\sigma_2$
+    (see Fig.~\ref{subfig:increasing before and above decreasing}),
+    then there is at most one arc between $\sigma_1$ and
+    $\sigma_2$ in $\mathcal{M}$ (this arc can be a
+    $(\sigma_1, \sigma_2)$-arc or a $(\sigma_2, \sigma_1)$-arc).
     \item
     If $\sigma_1$ is decreasing, $\sigma_2$ is increasing, and $\sigma_1$
-    is below $\sigma_2$ (see Fig.~\ref{subfig:decreasing before and below
-    increasing}), then $\mathcal{M}$ contains at most one
-    $(\sigma_1, \sigma_2)$-edge.
+    is below $\sigma_2$
+    (see Fig.~\ref{subfig:decreasing before and below increasing}),
+    then there is at most one arc between $\sigma_1$ and
+    $\sigma_2$ in $\mathcal{M}$ (this arc can be a
+    $(\sigma_1, \sigma_2)$-arc or a $(\sigma_2, \sigma_1)$-arc).
   \end{enumerate}
@@
     \begin{tikzpicture}[
-      scale=.8,
+      scale=.9,
       >=stealth',
@@
       ]
-      % edges
+      % arcs
       % pi 1 - pi 2
-      \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
+      \draw [pattern link,->,>=latex'] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.4,5);
       % pi 2 - sigma 2
-      \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
+      \draw [pattern link,->,>=latex'] (10.6,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
       % 1
-      \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
-      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
+      \draw [structure link,->,>=latex'] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
 %       \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
@@
       % 2
-      \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
-      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
+      \draw [structure link,->,>=latex'] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
 %       \node (claim edge 2) at (4,4) {Claim~\ref{claim:(nu_2, nu'_2)-edge}};
@@
       % 3
-      \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
-      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
+      \draw [structure link,->,>=latex'] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
 %       \node (claim edge 1) at (5.5,11.5) {Claim~\ref{claim:(nu_3, nu'_3)-edge}};
       % 4
-      \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
-      % SUPPRIMER COMMENTAIRE POUR VERSION LONGUE
+      \draw [structure link,->,>=latex'] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
 %       \node (claim edge 1) at (8.5,8.5) {Claim~\ref{claim:(nu_4, nu'_4)-edge}};
       % sigma 1 - pi 1
-      \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
+      \draw [structure link,->,>=latex'] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
       % \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
@@
       in Proposition~\ref{proposition:hardness}.
-      Black edges denote the presence of at least one edge between two bunches of
-      vertices in $K_\mu$.
-      Grey edges denote edges that are only considered in the forward direction of the
+      Black arcs denote the presence of at least one arc between two bunches of
+      positions in $\mu$.
+      Grey arcs denote edges that are only considered in the forward direction of the
       proof.
@@
   Define
-  % \begin{align*}
-  % N_4 &= 2(2n + 2k + 4) + 1  = 4n + 4k + 9 \\
-  % N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45 \\
-  % N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225 \\
-  % N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325\text{.}
-  % \end{align*}
-  $N_4 = 2(2n + 2k + 4) + 1  = 4n + 4k + 9$,
-  $N_3 = 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45$,
-  $N_2 = 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225$ and
-  $N_1 = 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325$.
+  \begin{align*}
+  N_4 &= 2(2n + 2k + 4) + 1  = 4n + 4k + 9 \\
+  N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45 \\
+  N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225 \\
+  N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325\text{.}
+  \end{align*}
+%   $N_4 = 2(2n + 2k + 4) + 1  = 4n + 4k + 9$,
+%   $N_3 = 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 20k + 45$,
+%   $N_2 = 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 100n + 100k + 225$ and
+%   $N_1 = 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1 = 1000n + 1000k  + 1325$.
   Notice that $N_1, N_2, N_3$ and $N_4$ are polynomial in $n$.
@@
   that $\sigma$ occurs in $\pi$ if and only if
-  there exists a containment-free perfect matching
-  $\mathcal{M}$ in $K_\mu$ such that
-  for any two disctinct edges
-  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
-  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
+  there exists an oriented perfect matching $\mathcal{M}$ on $\mu$
+  satisfying $\mathbf{P_1}$ and $\mathbf{P_2}$.
   \qed

### 2016-02-10 submission.tex
Small chages

@@
 {\em oriented matching} if all edges of $G$ are pairwise independent. 
-Moreover, $G$ is {\em perfect} if all vertices of $G$ belong to at least 
+Moreover, $G$ is {\em perfect} if any vertex of $G$ belongs to at least 
 one arc. For any word $u = u_1 u_2 \dots u_n$, an 
@@
 oriented perfect matchings on words. For instance, an oriented perfect
-matching $\mathcal{M}$ of a word $u$ {\em admits an occurrence} of the 
+matching $\mathcal{M}$ on a word $u$ {\em admits an occurrence} of the 
 pattern $\CrossingRL$ if there are four positions $i < j < k < \ell$ in 
@@
   Let $\pi$ be a permutation. An oriented perfect matching $\mathcal{M}$
-  on $\pi$ is said to have property $\mathbf{P_1}$ if it avoids the 
-  patterns \InclusionLL, \InclusionLR, \InclusionRL, 
+  on $\pi$ is said to have property $\mathbf{P_1}$ if it avoids all the 
+  six patterns \InclusionLL, \InclusionLR, \InclusionRL, 
   \InclusionRR, \CrossingLR, and \CrossingRL.

### 2016-02-10 new-article.tex
Modif LATIN16
LATIN 2016

@@
     exactly one arc between $\nu_1$ and $\nu_2$ in $\vv{\mathcal{M}}$.
-    According to Corollary~\ref{corollary:reverse oriented matching} there is no
+    According to Corollary~\ref{corollary:reverse oriented matching}, there is no
     loss of generality in assuming that this is a
-    $(\nu_1, \nu'_1)$-arc, say $(a, a')$.
+    $(\nu_1, \nu_2)$-arc and denote it by $(a, a')$.
     Furtheremore, according to Lemma~\ref{lemma:for nu_1 and nu_2},
@@
       (\emph{i.e.}, $y < a'$).
-      But it follows that $\vv{\mathcal{M}}$ contain \CrossingLR
+      But it follows that $\vv{\mathcal{M}}$ contains \CrossingLR
       which contradicts Property~$\mathbf{P_1}$.
@@
     \label{claim:(nu_1, nu'_1)-edge}
-    There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
+    There is at least one $(\nu_1, \nu'_1)$-arc in $\vv{\mathcal{M}}$.
   \end{claim}
@@
   \begin{proof}[of Claim~\ref{claim:(nu_1, nu'_1)-edge}]
-    Suppose, aiming at a contradiction, that there is no
-    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
-    Then it follows that there exists an edge $(x, x')$ in $\mathcal{M}$,
+    Suppose, aiming at a contradiction, that there is no arc between
+    $\nu_1$ and $\nu'_1$ in $\vv{\mathcal{M}}$.
+    Then it follows that there exists an arc $(x, x')$ or
+    $(x', x)$ in $\vv{\mathcal{M}}$,
     $1 \leq x \leq N_1$, that is neither a
-    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
-    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
-    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
+    $(\nu_1, \nu_1)$-arc (since $N_1$ is odd)
+    nor a $(\nu_1, \nu_2)$-arc or a $(\nu_2, \nu_1)$-arc
+    (Claim~\ref{claim:no (nu_1, nu_2)-edge})
+    nor a $(\nu_1, \nu'_1)$-arc or $(\nu_1, \nu'_1)$-arc
+    (by our contradiction hypothesis).
     (In other words, $x \leq N_1$ and $x' > 2N_1 + N_2$.)
-    Therefore, since $\mathcal{M}$ is containment-free,
-    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
-    in $\mathcal{M}$.
+    Therefore, since $\mathcal{M}$ avoids
+    \InclusionLL,~\InclusionLR,~\InclusionRL and \InclusionRR
+    (part of Property~$\mathbf{P_1}$),
+    there is neither a $(\nu_2, \nu_2)$-arc nor a
+    $(\nu'_1, \nu'_1)$-arc in $\mathcal{M}$.
     Then it follows that every position $y$ of $\nu'_1$
     (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
-    is involved in an edge $(y, y'$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
-    (otherwise $\mathcal{M}$ would not be containment-free).
+    is involved in an arc $(y, y')$ or $(y', y')$ in
+    $\vv{\mathcal{M}}$ with $y' \geq 2N_1+N_2$
+    (otherwise $\mathcal{M}$ would contain
+    \InclusionLL,~\InclusionLR,~\InclusionRL or \InclusionRR).
     But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
     > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
-    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
-    Then it follows that $\mathcal{M}$ is not a perfect matching,
-    which contradicts our hypothesis about $\mathcal{M}$.
+    position in $\nu'_1$ can not be involved in an arc in $\vv{\mathcal{M}}$.
+    Then it follows that $\vv{\mathcal{M}}$ is not a perfect matching,
+    which contradicts our hypothesis about $\vv{\mathcal{M}}$.
+
+    Combining tha aboce contradiction with
+    Corollary~\ref{corollary:reverse oriented matching}, we may assume that
+    $\vv{\mathcal{M}}$ contain at least one $(\nu_1, \nu'_1)$-arc
+    (and hence no $(\nu'_1, \nu_1)$-arc).
     \qed

### 2016-02-10 submission.tex

@@
 Moreover, $G$ is {\em perfect} if any vertex of $G$ belongs to at least 
-one arc. For any word $u = u_1 u_2 \dots u_n$, an 
-{\em oriented perfect matching on $u$} is an oriented perfect matching 
-$\mathcal{M}$ on the set of vertices $\{u_1, u_2, \dots, u_n\}$. In the 
-sequel, we shall consider a natural notion of pattern avoidance in 
-oriented perfect matchings on words. For instance, an oriented perfect
-matching $\mathcal{M}$ on a word $u$ {\em admits an occurrence} of the 
+one arc. For any permutation $\pi$ of $S_n$, an 
+{\em oriented perfect matching on $\pi$} is an oriented perfect matching 
+$\mathcal{M}$ on the set of vertices $[n]$. In the sequel, we shall 
+consider a natural notion of pattern avoidance in oriented perfect 
+matchings on permutations. For instance, an oriented perfect matching 
+$\mathcal{M}$ on a permutation $\pi$ {\em admits an occurrence} of the 
 pattern $\CrossingRL$ if there are four positions $i < j < k < \ell$ in 
-$u$ such that $(u_k, u_i)$ and $(u_j, u_\ell)$ are arcs of $\mathcal{M}$. 
-When $\mathcal{M}$ does not admit any occurrence of a pattern 
-$\mathcal{P}$, we say that $\mathcal{M}$ {\em avoids}~$\mathcal{P}$.
+$\pi$ such that $(\pi(k), \pi(i))$ and $(\pi(j), \pi(\ell))$ are arcs of 
+$\mathcal{M}$. When $\mathcal{M}$ does not admit any occurrence of a 
+pattern $\mathcal{P}$, we say that $\mathcal{M}$ {\em avoids}~$\mathcal{P}$.
 
@@
   on $\pi$ is said to have property $\mathbf{P_2}$ if, for any two 
-  distinct arcs $(a, a')$ and $(b, b')$ in $\mathcal{M}$,
+  distinct arcs $(\pi(a), \pi(a'))$ and $(\pi(b), \pi(b'))$ in $\mathcal{M}$,
   we have $\pi(a) < \pi(b)$ if and only if $\pi(a') < \pi(b')$.

### 2016-02-10 submission.tex
LATIN 2016

@@
 {\em oriented perfect matching on $\pi$} is an oriented perfect matching
-$\mathcal{M}$ on the set of vertices $[n]$. In the sequel, we shall 
+$\vv{\mathcal{M}}$ on the set of vertices $[n]$. In the sequel, we shall
 consider a natural notion of pattern avoidance in oriented perfect
 matchings on permutations. For instance, an oriented perfect matching
-$\mathcal{M}$ on a permutation $\pi$ {\em admits an occurrence} of the 
+$\vv{\mathcal{M}}$ on a permutation $\pi$ {\em admits an occurrence} of the
 pattern $\CrossingRL$ if there are four positions $i < j < k < \ell$ in
 $\pi$ such that $(\pi(k), \pi(i))$ and $(\pi(j), \pi(\ell))$ are arcs of
-$\mathcal{M}$. When $\mathcal{M}$ does not admit any occurrence of a 
-pattern $\mathcal{P}$, we say that $\mathcal{M}$ {\em avoids}~$\mathcal{P}$.
+$\vv{\mathcal{M}}$. When $\mathcal{M}$ does not admit any occurrence of a
+pattern $\mathcal{P}$,
+we say that $\vv{\mathcal{M}}$ {\em avoids}~$\mathcal{P}$.
 

### 2016-02-10 submission.tex
LATIN 2016

@@
 $\vv{\mathcal{M}}$. When $\mathcal{M}$ does not admit any occurrence of a
-pattern $\mathcal{P}$,
-we say that $\vv{\mathcal{M}}$ {\em avoids}~$\mathcal{P}$.
+pattern $P$,
+we say that $\vv{\mathcal{M}}$ {\em avoids}~$P$.
+The definition naturally extends to sets of patterns:
+$\vv{\mathcal{M}}$ {\em avoids}~$\mathcal{P}=\{P_i : 1 \leq i \leq k\}$
+if it avoids every pattern $P_i$.
 

### 2016-04-05 slides/slides.tex
add slides

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[unknownkeysallowed,11pt]{beamer}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usetheme{Boadilla}
\usefonttheme{serif}
\usecolortheme{default}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{euler,palatino}
\usepackage{amsmath,amssymb}
\usepackage[utf8]{inputenc}
\usepackage{xxcolor}
\usepackage{graphicx}
\usepackage[ruled]{algorithm2e}
\usepackage[english]{babel}
\usepackage{rotate}
\usepackage{shuffle}
\usepackage{minted}
\usepackage{url}
\usepackage{etoolbox}
\usepackage{tikz}
\usepackage{rotating}
\usetikzlibrary{snakes}
\usetikzlibrary{trees}
\usetikzlibrary{arrows}
\usetikzlibrary{matrix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{csquotes}
\usepackage{breakcites}
\usepackage{hyperref}
\usepackage[autocite=superscript,
            sortcites,
            maxcitenames=3,
            doi=false,
            isbn=false,
            note=false,
            url=false,
            bibstyle=authoryear,
            style=authoryear,
            backend=bibtex]{biblatex}

\AtEveryCitekey{%
    \clearfield{title}%
    \clearfield{note}%
    \clearfield{pages}%
    \clearlist{location}%
    \clearlist{publisher}%
    \clearname{editor}%
}% end AtEveryCitekey

\bibliography{biblio.bib}
\renewcommand*{\multicitedelim}{\\}
\renewcommand{\footfullcite}[1]{\footnote[frame]{\fullcite{#1}}}
\def\bibfont{\tiny}
\beamertemplatenavigationsymbolsempty

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\definecolor{Vert}{RGB}{20,100,20}
\setbeamercolor{title}{fg=Vert}
\setbeamercolor{frametitle}{fg=Vert}
\setbeamercolor{structure}{fg=Vert}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setbeamercovered{dynamic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{complexity}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Unshuffling Permutations]{%
  \textbf{Unshuffling Permutations}
}

\subtitle{
    \textbf{Samuele Giraudo}\quad\textbf{St\'ephane~Vialette}\\
    Laboratoire d'Informatique Gaspard-Monge\\
    Université Paris-Est Marne-la-Vallée \\
    UMR CNRS 8049 \\
    ~\\
    \includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/CNRSfr}
    \qquad
    \includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/ecole_ponts_RVB300.jpg}
    \qquad
    \includegraphics[width=1.5cm,height=2cm,keepaspectratio]{logos/LIGM}
    \qquad
    \raisebox{.25cm}{\includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/LogoEsieeParis}}
    \qquad
    \raisebox{.25cm}{\includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/UPEM_LOGO_EDITION}}
}

\date[LATIN 16]{
    {
    \scriptsize Latin American Theoretical Informatics Symposium.\\
    April $11$-$15$ $2016$
    }
}% end date

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\frame{\titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
  \frametitle{Hardness}
  \framesubtitle{Oriented perfect matching}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-04-05 slides/slides.tex
slides LATIN

@@
 \usetikzlibrary{matrix}
+\usetikzlibrary{calc}
 
@@
 \setbeamercovered{dynamic}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newcommand{\MWithContainment}{%
+\begin{tikzpicture}[scale=0.1,inner sep=.8pt,node distance=.25cm]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw (U2.north) .. controls ($ (U2.north) + (0,1.75) $) and ($ (U3.north) + (0,1.75) $) .. (U3.north);
+\draw (U1.north) .. controls ($ (U1.north) + (0,2.5) $) and ($ (U4.north) + (0,2.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \MWithContainment
+
+\newcommand{\Crossing}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \Crossing
+
+\newcommand{\LabeledCrossingLL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingLL
+
+\newcommand{\CrossingLL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U3.north) + (0,2) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingLL
+
+\newcommand{\LabeledCrossingLR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingLR
+
+\newcommand{\CrossingLR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingLR
+
+\newcommand{\LabeledCrossingRL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingRL
+
+\newcommand{\CrossingRL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingRL
+
+\newcommand{\LabeledCrossingRR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U3.north) + (0,1.5) $) .. (U3.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \LabeledCrossingRR
+
+\newcommand{\CrossingRR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']%,>=stealth',->]
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U3.north) + (0,4) $) .. (U3.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \CrossingRR
+
+\newcommand{\Inclusion}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \Inclusion
+
+\newcommand{\LabeledInclusionLL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionLL
+
+\newcommand{\InclusionLL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
++\draw [->] (U4.north) .. controls ($ (U4.north) + (0,4) $) and ($ (U1.north) + (0,4) $) .. (U1.north);
++\draw [->] (U3.north) .. controls ($ (U3.north) + (0,3) $) and ($ (U2.north) + (0,3) $) .. (U2.north);
+\end{tikzpicture}
+}% end \InclusionLL
+
+\newcommand{\LabeledInclusionLR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionLR
+
+\newcommand{\InclusionLR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionLR
+
+\newcommand{\LabeledInclusionRL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionRL
+
+\newcommand{\InclusionRL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionRL
+
+\newcommand{\LabeledInclusionRR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,2) $) and ($ (U4.north) + (0,2) $) .. (U4.north);
+\draw [<-] (U2.north) .. controls ($ (U2.north) + (0,1) $) and ($ (U3.north) + (0,1) $) .. (U3.north);
+\end{tikzpicture}
+}% end \LabeledInclusionRR
+
+\newcommand{\InclusionRR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\draw [->] (U2.north) .. controls ($ (U2.north) + (0,3) $) and ($ (U3.north) + (0,3) $) .. (U3.north);
+\end{tikzpicture}
+}% end \InclusionRR
+
+\newcommand{\Precedence}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \Precedence
+
+\newcommand{\LabeledPrecedenceLL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLL
+
+\newcommand{\PrecedenceLL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLL
+
+\newcommand{\LabeledPrecedenceLR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLR
+
+\newcommand{\PrecedenceLR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [->] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceLR
+
+\newcommand{\LabeledPrecedenceRL}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRL
+
+\newcommand{\PrecedenceRL}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [->] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRL
+
+\newcommand{\LabeledPrecedenceRR}[5]{%
+\begin{tikzpicture}[scale=#1,inner sep=.8pt,node distance=.45cm,>=latex',
+                    text height=1.5ex,text depth=.25ex]
+\draw node [] (U1)               {#2};
+\draw node [] [right of=U1] (U2) {#3};
+\draw node [] [right of=U2] (U3) {#4};
+\draw node [] [right of=U3] (U4) {#5};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,1.5) $) and ($ (U2.north) + (0,1.5) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,1.5) $) and ($ (U4.north) + (0,1.5) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRR
+
+\newcommand{\PrecedenceRR}{%
+\begin{tikzpicture}[yscale=0.08,xscale=0.09,inner sep=.8pt,node distance=.25cm,>=latex']
+\draw node [draw,circle,fill=black] (U1)               {};
+\draw node [draw,circle,fill=black] [right of=U1] (U2) {};
+\draw node [draw,circle,fill=black] [right of=U2] (U3) {};
+\draw node [draw,circle,fill=black] [right of=U3] (U4) {};
+\draw [<-] (U1.north) .. controls ($ (U1.north) + (0,4) $) and ($ (U2.north) + (0,4) $) .. (U2.north);
+\draw [<-] (U3.north) .. controls ($ (U3.north) + (0,4) $) and ($ (U4.north) + (0,4) $) .. (U4.north);
+\end{tikzpicture}
+}% end \PrecedenceRR
 
@@
   \framesubtitle{Oriented perfect matching}
+
+  \begin{definition}[Property $\mathbf{P_1}$]
+    Let $\pi$ be a permutation. An oriented perfect matching $\mathcal{M}$
+    on $\pi$ is said to have property $\mathbf{P_1}$ if it avoids all the
+    six patterns \InclusionLL, \InclusionLR, \InclusionRL,
+    \InclusionRR, \CrossingLR, and \CrossingRL.
+  \end{definition}
+
+  \medskip
+
+  \begin{definition}[Property $\mathbf{P_2}$]
+    Let $\pi$ be a permutation. An oriented perfect matching $\mathcal{M}$
+    on $\pi$ is said to have property $\mathbf{P_2}$ if, for any two
+    distinct arcs $(\pi(a), \pi(a'))$ and $(\pi(b), \pi(b'))$ in $\mathcal{M}$,
+    we have $\pi(a) < \pi(b)$ if and only if $\pi(a') < \pi(b')$.
+  \end{definition}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Hardness}
+  \framesubtitle{Oriented perfect matching}
+
+  \begin{theorem}
+    Let $\pi$ be a permutation. The following statements are equivalent:
+    \begin{enumerate}
+      \item The permutation $\pi$ is a square.
+      \item There exists an oriented perfect matching $\mathcal{M}$
+      on $\pi$ satisfying~$\mathbf{P_1}$ and~$\mathbf{P_2}$.
+    \end{enumerate}
+  \end{theorem}
+
+  \medskip
+
+  \onslide<2->
+  \begin{exampleblock}{Example}
+    \begin{tikzpicture}[scale=0.35,inner sep=2pt,node distance=1cm]
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10](U01){$1$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U01]
+            (U02) {$8$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U02](U03){$3$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U03]
+            (U04) {$9$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U04]
+            (U05) {$2$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U05](U06){$7$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U06](U07){$B$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U07](U08){$5$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U08](U09){$C$};
+        \draw node [draw,circle,minimum size=.55cm,right of=U09](U10){$6$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U10]
+            (U11) {$A$};
+        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U11]
+            (U12) {$4$};
+        \draw [thick,->,>=latex']
+        (U01.north) .. controls ($ (U01.north) + (0,2.25) $)
+            and ($ (U03.north) + (0,2.25) $) .. (U03.north);
+        \draw [thick,->,>=latex']
+        (U02.north) .. controls ($ (U02.north) + (0,2.25) $)
+            and ($ (U06.north) + (0,2.25) $) .. (U06.north);
+        \draw [thick,->,>=latex']
+        (U04.north) .. controls ($ (U04.north) + (0,2.5) $)
+            and ($ (U07.north) + (0,2.5) $) .. (U07.north);
+        \draw [thick,->,>=latex']
+        (U05.north) .. controls ($ (U05.north) + (0,2.5) $)
+            and ($ (U08.north) + (0,2.5) $) .. (U08.north);
+        \draw [thick,->,>=latex']
+        (U11.north) .. controls ($ (U11.north) + (0,2.5) $)
+            and ($ (U09.north) + (0,2.5) $) .. (U09.north);
+        \draw [thick,->,>=latex']
+        (U12.north) .. controls ($ (U12.north) + (0,2.5) $)
+            and ($ (U10.north) + (0,2.5) $) .. (U10.north);
+    \end{tikzpicture}
+
+    {\small%
+    An oriented perfect matching $\mathcal{M}$ on the permutation
+    $\pi = 183927B5C6A4$ satisfying the properties $\mathbf{P_1}$
+    and $\mathbf{P_2}$. From $\mathcal{M}$, it follows that $\pi$ is
+    a square as it appears in the shuffle of $1892A4$ and $37B5C6$,
+    both being order-isomorphic to $145263$.
+    }%end small
+  \end{exampleblock}
 \end{frame}

### 2016-04-05 slides/slides.tex
slides LATIN
pictures

@@
 
+\begin{frame}
+  \frametitle{Hardness}
+
+  \begin{theorem}
+    Deciding whether a permutation is a square is \NPC.
+  \end{theorem}
+
+  \begin{block}{Remarks}
+    \begin{itemize}
+      \item
+      We reduce from the problem of deciding wether a pattern occurs in a
+      permutation.
+
+      \smallskip
+
+      \item
+    \end{itemize}
+  \end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Hardness}
+
+  \centering
+  \begin{tikzpicture}[
+    scale=.6,
+    >=stealth',
+    shorten >=1pt,
+    main node/.style={align=center},
+    cell/.style={draw,ultra thick,fill=black!5},
+    structure link/.style={line width=1.5pt},
+    pattern link/.style={line width=1.5pt,black!20},
+    monotone/.style={->,thick}
+    ]
+    % arcs
+    % pi 1 - pi 2
+    \draw [pattern link,->,>=latex'] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.4,5);
+    % pi 2 - sigma 2
+    \draw [pattern link,->,>=latex'] (10.6,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
+    % 1
+    \draw [structure link,->,>=latex'] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
+  %       \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
+  %       \node (claim edge 1) at (1.5,13) {Claim~\ref{claim:all (nu_1, nu'_1)-edges}};
+    % 2
+    \draw [structure link,->,>=latex'] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
+  %       \node (claim edge 2) at (4,4) {Claim~\ref{claim:(nu_2, nu'_2)-edge}};
+  %       \node (claim edge 2) at (4,3.5) {Claim~\ref{claim:all (nu_2, nu'_2)-edges}};
+    % 3
+    \draw [structure link,->,>=latex'] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
+  %       \node (claim edge 1) at (5.5,11.5) {Claim~\ref{claim:(nu_3, nu'_3)-edge}};
+    % 4
+    \draw [structure link,->,>=latex'] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
+  %       \node (claim edge 1) at (8.5,8.5) {Claim~\ref{claim:(nu_4, nu'_4)-edge}};
+    % sigma 1 - pi 1
+    \draw [structure link,->,>=latex'] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
+    % \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
+    % nodes
+    % 1
+    \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
+    \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {\small  $\nu_1$};
+    %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
+    \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
+    \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {\small  $\nu'_1$};
+    % 2
+    \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
+    \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {\small  $\nu_2$};
+    \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
+    \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {\small  $\nu'_2$};
+    % 3
+    \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
+    \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {\small  $\nu_3$};
+    \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
+    \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {\small  $\nu'_3$};
+    % 4
+    \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
+    \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {\small $\nu_4$};
+    \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
+    \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {\small  $\nu'_4$};
+    % sigma 1
+    \draw [cell] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
+    \node [main node] (sigma1) at (4.5,6.5) {\small $\sigma'$};
+    % pi 1
+    \draw [cell] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
+    \node [main node] (pi1) at (8.5,5.5) {\small $\pi'$};
+    % pi 2
+    \draw [cell] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
+    \node [main node] (sigma2) at (10.5,4.5) {\small $\pi''$};
+    % sigma 2
+    \draw [cell] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
+    \node [main node] (pi2) at (11.5,3.5) {\small $\sigma''$};
+    % grid
+    \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
+    % row size
+    \foreach \y/\N in {0.5/N_2,1.5/N_2,2.5/N_4,3.5/k,4.5/n,5.5/n+2,6.5/k+2,7.5/N_4,8.5/N_3,9.5/N_3,10.5/N_1,11.4/N_1} {
+        \node at (-0.5,\y) (R\y) {\tiny $\N$};
+    }
+    % column size
+    \foreach \x/\N in
+    {0.5/N_1,1.5/N_2,2.5/N_1,3.5/N_3,4.5/k+2,5.5/N_4,6.5/N_2,7.5/N_3,8.5/n+2,9.5/N_4,10.5/n,11.5/k} {
+        \node at (\x,-0.5) (C\x) {\tiny $\N$};
+    }
+  \end{tikzpicture}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2016-04-05 slides/slides.tex
pictures

@@
     pattern link/.style={line width=1.5pt,black!20},
-    monotone/.style={->,thick}
+    monotone/.style={draw,ultra thick,fill=black!5}
     ]
@@
     % 1
-    \draw [structure link,->,>=latex'] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
+    \draw [structure link,->,>=latex'] (0.5,11) .. controls +(0,2.25) and +(0,1.25) .. (2.5,12);
   %       \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
@@
     % sigma 1 - pi 1
-    \draw [structure link,->,>=latex'] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
+    \draw [structure link,->,>=latex',red] (4.5,6) .. controls +(0,-3) and +(0,-2) .. (8.5,5);
     % \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
@@
     \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
-    \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {\small  $\nu_1$};
+    \draw [monotone] (0.5,10.5) node {\small  $\nu_1$};
     %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
     \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
-    \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {\small  $\nu'_1$};
+    \draw [monotone] (2.5,11.5) node {\small  $\nu'_1$};
     % 2
     \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
-    \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {\small  $\nu_2$};
+    \draw [monotone] (1.5,1.5) node {\small  $\nu_2$};
     \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
-    \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {\small  $\nu'_2$};
+    \draw [monotone] (6.5,0.5) node {\small  $\nu'_2$};
     % 3
     \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
-    \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {\small  $\nu_3$};
+    \draw [monotone] (3.5,8.5) node {\small  $\nu_3$};
     \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
-    \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {\small  $\nu'_3$};
+    \draw [monotone] (7.5,9.5) node {\small  $\nu'_3$};
     % 4
     \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
-    \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {\small $\nu_4$};
+    \draw [monotone] (5.5,7.5) node {\small $\nu_4$};
     \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
-    \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {\small  $\nu'_4$};
+    \draw [monotone] (9.5,2.5) node {\small  $\nu'_4$};
     % sigma 1

### 2016-04-05 slides/slides.tex
pictures

@@
 \begin{frame}
+  \frametitle{Permutation pattern}
+
+  \begin{definition}
+    A permutation $\pi$ is said to \emph{contain} the permutation $\sigma$ if there
+    exists a subsequence of (not necessarily consecutive) entries of $\pi$ that
+    has the same relative order as $\sigma$, and in this case $\sigma$ is said
+    to be a pattern of $\pi$, written $\sigma \leq \pi$.
+
+    \medskip
+
+    Otherwise, $\pi$ is said to \emph{avoid} the permutation $\sigma$.
+  \end{definition}
+
+  \medskip
+
+  \begin{exampleblock}{Example}
+    \begin{itemize}
+      \item
+        The permutation $\pi = 391867452$ contains the pattern $\sigma = 51342$,
+        as can be seen in the highlighted subsequence of
+        $\pi = 3\mathbf{91}8\mathbf{674}52$
+        (or $\pi = 3\mathbf{91}8\mathbf{67}4\mathbf{5}2$ or
+        $\pi = 3\mathbf{91}8\mathbf{67}45\mathbf{2}$ or
+        $\pi = 3\mathbf{91}867\mathbf{452}$.)
+
+      \item
+       Since the permutation $\pi = 391867452$ contains no increasing subsequence
+       of length four, $\pi$ avoids $1234$.
+    \end{itemize}
+  \end{exampleblock}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Square permutation}
+
+  \begin{definition}
+    A permutation is said to be a \emph{square} if it can be obtained by
+    shuffling two order-isomorphic patterns
+  \end{definition}
+
+  \medskip
+
+  \begin{exampleblock}{Example}
+    \begin{itemize}
+      \item
+        The permutation $\pi = 25167834$ is a square since it can obtained by
+        shuffling two order-isomorphic (\emph{i.e.} $1342$)
+        patterns $1683$ and $2574$ as can be seen in the highlighted subsequence of
+        $\pi = \mathbf{25}16\mathbf{7}83\mathbf{4}$.
+
+      \medskip
+
+      \item
+       $\pi = 4123$ and $1432$ are not squares.
+    \end{itemize}
+  \end{exampleblock}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Words}
+
+  \begin{definition}
+    The shuffle operator on words is defined inductively as follows:
+    \begin{align*}
+      u \shuffle \epsilon &= \epsilon \shuffle u = \{u\}
+      \\
+      (ua \shuffle vb) &= (ua \shuffle v)b \cup (u \shuffle vb)a
+    \end{align*}
+  \end{definition}
+
+  \medskip
+
+  \begin{block}{Key results}
+    \begin{itemize}
+      \item
+      Given two words $v_1$ and $v_2$, the set $v_1 \shuffle v_2$
+      can be computed in
+      $O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)$
+      time.
+
+      \smallskip
+
+      \item
+      Given three words $u$, $v_1$, and $v_2$, deciding if $u$ is a shuffle
+      of $v_1$ and $v_2$ can be done in
+      $O\left(n^2 / \log(n)\right)$ time.
+
+      \smallskip
+
+      \item
+      Given a word $u$, deciding if there is a word $v$ such that $u$ is
+      in $v \shuffle v$ is \NPC.
+    \end{itemize}
+  \end{block}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
   \frametitle{Hardness}
@@
 
+  \medskip
+
   \begin{block}{Remarks}
@@
       \item
-      We reduce from the problem of deciding wether a pattern occurs in a
-      permutation.
+      We reduce from the problem of deciding wether a pattern $\sigma$ occurs in a
+      permutation $\pi$.
 
@@
       \item
+      Construct a new permutation $\mu$ such that
+      $\sigma$ occurs in $\pi$ if and only if
+      there exists an oriented perfect matching $\mathcal{M}$ on $\mu$
+      satisfying $\mathbf{P_1}$ and $\mathbf{P_2}$.
     \end{itemize}
@@
     \draw [structure link,->,>=latex'] (0.5,11) .. controls +(0,2.25) and +(0,1.25) .. (2.5,12);
-  %       \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
-  %       \node (claim edge 1) at (1.5,13) {Claim~\ref{claim:all (nu_1, nu'_1)-edges}};
     % 2
     \draw [structure link,->,>=latex'] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
-  %       \node (claim edge 2) at (4,4) {Claim~\ref{claim:(nu_2, nu'_2)-edge}};
-  %       \node (claim edge 2) at (4,3.5) {Claim~\ref{claim:all (nu_2, nu'_2)-edges}};
     % 3
     \draw [structure link,->,>=latex'] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
-  %       \node (claim edge 1) at (5.5,11.5) {Claim~\ref{claim:(nu_3, nu'_3)-edge}};
     % 4
     \draw [structure link,->,>=latex'] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
-  %       \node (claim edge 1) at (8.5,8.5) {Claim~\ref{claim:(nu_4, nu'_4)-edge}};
     % sigma 1 - pi 1
     \draw [structure link,->,>=latex',red] (4.5,6) .. controls +(0,-3) and +(0,-2) .. (8.5,5);
-    % \node (claim edge 1) at (6.5,4) {Claim~\ref{}};
     % nodes
@@
     \draw [monotone] (0.5,10.5) node {\small  $\nu_1$};
-    %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
     \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
@@
 
+\begin{frame}
+  \frametitle{Open problems}
+
+  \begin{block}{Combinatorics}
+    How many permutations of length $2n$ are squares?
+
+    \smallskip
+
+    How many $(213,231)$-avoiding
+    permutations of length $2n$ are squares? (Equivalently,
+    how many binary strings of length $2n$ are squares?)
+  \end{block}
+
+  \medskip
+
+  \begin{block}{Combinatorics}
+    Given two permutations $\pi$ and~$\sigma$, how hard is the problem of
+    deciding whether $\sigma$ is a square root of~$\pi$?
+  \end{block}
+
+  \medskip
+
+  \begin{block}{Algebra}
+    \ldots
+  \end{block}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2016-04-06 slides/slides.tex
(co)products

@@
 \usepackage{shuffle}
-\usepackage{minted}
+%\usepackage{minted}
 \usepackage{url}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\DeclareMathOperator{\QQ}{\mathbb{Q}}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Combinatorial products and structure coefficients}
+Let $C$ be a set of combinatorial objects and let $\QQ[C]$ be the
+$\QQ$-linear span of $C$.
+\medskip
+
+A binary linear product
+\begin{equation*}
+    \cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C].
+\end{equation*}
+on $\QQ[C]$ describes a way to \alert{combine} two objects of
+$C$ to obtain \alert{bigger} objects.
+\medskip
+
+In a general way, for any $x, y \in C$,
+\begin{equation*}
+    x \cdot y = \sum_{z \in C} \lambda_{x, y}^z \; z.
+\end{equation*}
+
+A product is wholly encoded by the coefficients
+$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
+\alert{structure coefficients}.
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Combinatorial coproducts}
+In a dual way, a linear coproduct
+\begin{equation*}
+    \Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C].
+\end{equation*}
+on $\QQ[C]$ describe a way to \alert{break} one object of $C$ into
+\alert{two smaller} pieces.
+\medskip
+
+In a general way, for any $z \in C$,
+\begin{equation*}
+    \Delta(x) = \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
+\end{equation*}
+
+A coproduct is wholly encoded by the coefficients
+$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
+\alert{structure coefficients}.
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Equivalence between products and coproducts}
+If $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$ is a product with
+structure coefficients $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$,
+we can define its \alert{dual coproduct} $\Delta$ by setting
+\begin{equation*}
+    \Delta(x) := \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
+\end{equation*}
+\medskip
+
+Conversely, if $\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$ is
+a coproduct with structure coefficients $\lambda_{x, y}^z \in \QQ$,
+$x, y, z \in C$, we can define its \alert{dual product} $\cdot$ by
+setting
+\begin{equation*}
+    x \cdot y := \sum_{z \in C} \lambda_{x, y}^z \; z.
+\end{equation*}
+
+\begin{block}{Idea}
+    Knowing how to break a combinatorial object is equivalent
+    to knowing how to combine these.
+\end{block}
+\end{frame}
+
 \end{document}

### 2016-04-06 slides/slides.tex
melanges

@@
 
-\documentclass[unknownkeysallowed,11pt]{beamer}
+\documentclass[unknownkeysallowed,10pt,xcolor={dvipsnames}]{beamer}
 
@@
 \DeclareMathOperator{\QQ}{\mathbb{Q}}
+\DeclareMathOperator{\STD}{\mathrm{std}}
+\DeclareMathOperator{\SHUFFLE}{\bullet}
+
+\newcommand{\BIB}[2]{{\footnotesize\textcolor{MidnightBlue!85}{[\textbf{#1}, #2]}}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Example: unshuffling binary words}
+Let $B$ the set of all binary words and let us consider the coproduct
+\begin{equation*}
+    \Delta : \QQ[B] \to \QQ[B] \otimes \QQ[B]
+\end{equation*}
+on binary words defined by
+\begin{equation*}
+    \Delta(u) :=
+    \sum_{P \sqcup Q = \{1, \dots, |u|\}}
+    u_{|P} \otimes u_{|Q}.
+\end{equation*}
+
+\begin{block}{Example}
+\begin{equation*}
+    \Delta(100) =
+    \epsilon \otimes 100 + 1 \otimes 00
+    + 2 (0 \otimes 10) + 2 (10 \otimes 0) + 00 \otimes 1
+    + 100 \otimes \epsilon
+\end{equation*}
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%o%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Example: shuffling binary words}
+\begin{block}{Proposition}
+    The shuffle product of binary words is the dual product of the
+    unshuffling coproduct of binary words.
+
+        REF ? SCHUTZ ?
+\end{block}
+\medskip
+
+\begin{block}{Example}
+\begin{equation*}
+    \Delta(100) =
+    \epsilon \otimes 100 + 1 \otimes 00
+    + \textcolor{BrickRed}{2 (0 \otimes 10)} + 2 (10 \otimes 0) + 00 \otimes 1
+    + 100 \otimes \epsilon
+\end{equation*}
+\smallskip
+
+Since $0 \otimes 10$ has multiplicity $2$ in $\Delta(100)$, by duality,
+$100$ has multiplicity $2$ in $0 \shuffle 10$:
+\begin{equation*}
+    0 \shuffle 10 = 010 + \textcolor{BrickRed}{2 (100)}.
+\end{equation*}
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Unshuffling permutations}
+If $u$ is a word of integers without multiple occurrence of a same letter,
+the \alert{standardized} $\STD(u)$ of $u$ is the unique permutation of
+$S_{|u|}$ order-isomorphic to~$u$.
+
+\begin{block}{Example}
+\begin{equation*}
+    \STD(82194) = 42153
+\end{equation*}
+\end{block}
+\medskip
+
+\begin{block}{Definition}
+    The \alert{unshuffling coproduct of permutations} is the
+    coproduct on $\QQ[S]$ defined by
+    \begin{equation*}
+        \Delta(\pi) =
+        \sum_{P \sqcup Q = \{1, \dots, |\pi|\}}
+        \STD\left(u_{|P}\right) \otimes \STD\left(u_{|Q}\right).
+    \end{equation*}
+\end{block}
+
+\begin{block}{Example}
+\begin{equation*}
+    \Delta(213) =
+    \epsilon \otimes 213 + 2 (1 \otimes 12) +
+    1 \otimes 21 + 2 (12 \otimes 1) + 21 \otimes 1
+    + 213 \otimes \epsilon
+\end{equation*}
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Shuffling permutations}
+\begin{block}{Definition}
+    The \alert{shuffling product of permutations} is the product
+    $\SHUFFLE$ on $\QQ[S]$ defined as the dual product of the unshuffling
+    product of permutations.
+\end{block}
+
+\begin{block}{Example}
+\begin{equation*}\begin{split}
+    12 \SHUFFLE 21 & =
+    1243 + 1324 + 2 (1342) + 2 (1423) + 3 (1432) +
+    2134 + 2 (2314) \\
+    & + 3 (2341) + 2413 + 2 (2431) + 2 (3124) + 3142 +
+    3 (3214) + 2 (3241) \\
+    & + 3421 + 3 (4123) + 2 (4132) + 2 (4213) + 4231 + 4312
+\end{split}\end{equation*}
+\end{block}
+\medskip
+
+This shuffle product has been introduced for the study of some
+Hopf algebra \BIB{Vargas}{2014}.
+\medskip
+
+It is a kind of association of the convolution product and the shifted
+shuffle product of permutations \BIB{Duchamp, Hivert, Thibon}{2002}.
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{}
+
+\end{frame}
+
 \end{document}

### 2016-04-06 slides/slides.tex
Suite

@@
 \begin{frame} \frametitle{Combinatorial products and structure coefficients}
-Let $C$ be a set of combinatorial objects and let $\QQ[C]$ be the
+Let $C$ be a set of combinatorial objects and $\QQ[C]$ be the
 $\QQ$-linear span of $C$.
@@
 \end{equation*}
-
-A product is wholly encoded by the coefficients
+\medskip
+
+A product is therefore wholly encoded by the coefficients
 $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
@@
 \end{equation*}
-
-A coproduct is wholly encoded by the coefficients
+\medskip
+
+A coproduct is therefore wholly encoded by the coefficients
 $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
@@
 \end{equation*}
+\medskip
 
@@
 \begin{frame} \frametitle{Example: shuffling binary words}
-\begin{block}{Proposition}
+\begin{block}{Proposition \BIB{Reutenauer}{1993}}
     The shuffle product of binary words is the dual product of the
     unshuffling coproduct of binary words.
-
-        REF ? SCHUTZ ?
 \end{block}
@@
 \begin{block}{Example}
-\begin{equation*}
-    \Delta(100) =
+Since
+\begin{equation*}
+    \textcolor{RoyalBlue}{\Delta(100)} =
     \epsilon \otimes 100 + 1 \otimes 00
-    + \textcolor{BrickRed}{2 (0 \otimes 10)} + 2 (10 \otimes 0) + 00 \otimes 1
-    + 100 \otimes \epsilon
-\end{equation*}
-\smallskip
-
-Since $0 \otimes 10$ has multiplicity $2$ in $\Delta(100)$, by duality,
-$100$ has multiplicity $2$ in $0 \shuffle 10$:
-\begin{equation*}
-    0 \shuffle 10 = 010 + \textcolor{BrickRed}{2 (100)}.
+    + \textcolor{Orange}{2} (\textcolor{BrickRed}{0 \otimes 10})
+    + 2 (10 \otimes 0) + 00 \otimes 1
+    + 100 \otimes \epsilon,
+\end{equation*}
+we observe that $\textcolor{BrickRed}{0 \otimes 10} $ has multiplicity
+$\textcolor{Orange}{2}$ in $\textcolor{RoyalBlue}{\Delta(100)}$.
+\medskip
+
+By duality, this implies that $\textcolor{RoyalBlue}{100}$ has
+multiplicity $\textcolor{Orange}{2}$ in
+$\textcolor{BrickRed}{0 \shuffle 10}$. Indeed,
+\begin{equation*}
+    \textcolor{BrickRed}{0 \shuffle 10} =
+    010 + \textcolor{Orange}{2} (\textcolor{RoyalBlue}{100}).
 \end{equation*}
@@
 
-This shuffle product has been introduced for the study of some
-Hopf algebra \BIB{Vargas}{2014}.
-\medskip
-
-It is a kind of association of the convolution product and the shifted
-shuffle product of permutations \BIB{Duchamp, Hivert, Thibon}{2002}.
+Product first introduced for the study of some Hopf algebras~\BIB{Vargas}{2014}.
+\medskip
+
+Similar to the convolution product and the shifted
+shuffle product \BIB{Duchamp, Hivert, Thibon}{2002}.
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Square permutations}
+\begin{block}{Definition}
+    A permutation $\pi$ is a \alert{square} (w.r.t. the shuffle product
+    $\SHUFFLE$) if there is $\sigma \in S$ such that $\pi$ appears in
+    $\sigma \SHUFFLE \sigma$.
+\end{block}
+\medskip
+
+By duality, $\pi$ is a square if there is $\sigma \in S$ such that
+$\sigma \otimes \sigma$ appears in $\Delta(\pi)$.
+\medskip
+
+\begin{block}{Proposition}
+    A permutation $\pi$ is a square iff $\pi$ can be obtained by
+    shuffling two order-isomorphic words.
+\end{block}
+\medskip
+
+\begin{block}{Example}
+    The permutation
+    \begin{math}
+        \pi :=
+        \textcolor{RoyalBlue}{25}\textcolor{BrickRed}{16}
+        \textcolor{RoyalBlue}{7}\textcolor{BrickRed}{83}
+        \textcolor{RoyalBlue}{4}
+    \end{math}
+    is square since $\pi$ can be obtained by shuffling
+    $\textcolor{BrickRed}{1683}$ and $\textcolor{RoyalBlue}{2574}$.
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Some properties of square permutations}
+The first numbers of square permutations are, size by size,
+\begin{equation*}
+    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
+\end{equation*}
+\medskip
+
+\begin{block}{Proposition}
+    Let $\pi$ be a square permutation and $\sigma$ be a square root of
+    $\pi$. Then,
+    \begin{enumerate}
+        \item $\widetilde{\pi}$ is a square and
+        $\widetilde{\sigma}$ is one of its square roots;
+        \item $\bar \pi$ is a square and $\bar \sigma$ is one of
+        its square roots;
+        \item $\pi^{-1}$ is a square and $\sigma^{-1}$ is one of
+        its square roots.
+    \end{enumerate}
+\end{block}
+\medskip
+
+\begin{block}{Proposition}
+    The set of binary words of length $n$ that are square w.r.t.
+    $\shuffle$ is in one-to-one correspondence with the set of
+    square permutations of length $n$ avoiding the patterns
+    $213$ and~$231$.
+\end{block}
+
 \end{frame}

### 2016-04-06 slides/slides.tex
Restructuration (pas finie)

@@
 
+% Affichage du plan à chaque changement de partie.
+\AtBeginSection[]{
+    \frame<handout:0>{
+        \frametitle{Plan}
+        \tableofcontents[
+            sectionstyle=show/hide,
+            subsectionstyle=show/shaded/hide]}
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \DeclareMathOperator{\SHUFFLE}{\bullet}
+\DeclareMathOperator{\La}{\mathtt{a}}
+\DeclareMathOperator{\Lb}{\mathtt{b}}
 
@@
     {
-    \scriptsize Latin American Theoretical Informatics Symposium.\\
+    \scriptsize Latin American Theoretical Informatics Symposium \\
     April $11$-$15$ $2016$
@@
 \frame{\titlepage}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Contents}
+    \tableofcontents
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Shuffling words}
+\begin{block}{Definition}
+    The \alert{shuffle product} $\shuffle$ on words of $A^*$ is defined
+    inductively by
+    \begin{align*}
+        u \shuffle \epsilon &= \epsilon \shuffle u = \{u\} \\
+        (ua \shuffle vb) &= (ua \shuffle v)b \cup (u \shuffle vb)a
+    \end{align*}
+\end{block}
+\medskip
+
+\uncover<2->{%
+To take into account multiplicities, we consider $\shuffle$ as a linear
+product
+\begin{equation*}
+    \QQ[A^*] \otimes \QQ[A^*] \to \QQ[A^*]
+\end{equation*}
+on $\QQ[A^*]$, the $\QQ$-linear span of words defined by
+\begin{align*}
+    u \shuffle \epsilon & = \epsilon \shuffle u = u \\
+    (ua \shuffle vb) &= (ua \shuffle v)b + (u \shuffle vb)a
+\end{align*}}
+
+\uncover<3->{%
+\begin{block}{Example}
+    \vspace{-1em}
+    \begin{equation*}\begin{split}
+        \La\Lb \shuffle \textcolor{BrickRed}{\Lb\La} & =
+        \La\Lb\textcolor{BrickRed}{\Lb\La}
+        + \La\textcolor{BrickRed}{\Lb}\Lb\textcolor{BrickRed}{\La}
+        + \La\textcolor{BrickRed}{\Lb\La}\Lb
+        + \textcolor{BrickRed}{\Lb}\La\Lb\textcolor{BrickRed}{\La}
+        + \textcolor{BrickRed}{\Lb}\La\textcolor{BrickRed}{\La}\Lb
+        + \textcolor{BrickRed}{\Lb\La}\La\Lb \\
+        &
+        \uncover<4->{=
+        2 \La\Lb\Lb\La + \La\Lb\La\Lb + \Lb\La\Lb\La + 2 \Lb\La\La\Lb}
+    \end{split}\end{equation*}
+    \vspace{-1.5em}
+\end{block}}
+\end{frame}
 
@@
 \end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Algebraic issues}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Combinatorial products and structure coefficients}
+Let $C$ be a set of combinatorial objects and $\QQ[C]$ be the
+$\QQ$-linear span of $C$.
+\medskip
+
+A binary linear product
+\begin{equation*}
+    \cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C].
+\end{equation*}
+on $\QQ[C]$ describes a way to \alert{combine} two objects of
+$C$ to obtain \alert{bigger} objects.
+\medskip
+
+In a general way, for any $x, y \in C$,
+\begin{equation*}
+    x \cdot y = \sum_{z \in C} \lambda_{x, y}^z \; z.
+\end{equation*}
+\medskip
+
+A product is therefore wholly encoded by the coefficients
+$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
+\alert{structure coefficients}.
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Combinatorial coproducts}
+In a dual way, a linear coproduct
+\begin{equation*}
+    \Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C].
+\end{equation*}
+on $\QQ[C]$ describe a way to \alert{break} one object of $C$ into
+\alert{two smaller} pieces.
+\medskip
+
+In a general way, for any $z \in C$,
+\begin{equation*}
+    \Delta(x) = \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
+\end{equation*}
+\medskip
+
+A coproduct is therefore wholly encoded by the coefficients
+$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
+\alert{structure coefficients}.
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Equivalence between products and coproducts}
+If $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$ is a product with
+structure coefficients $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$,
+we can define its \alert{dual coproduct} $\Delta$ by setting
+\begin{equation*}
+    \Delta(x) := \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
+\end{equation*}
+\medskip
+
+Conversely, if $\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$ is
+a coproduct with structure coefficients $\lambda_{x, y}^z \in \QQ$,
+$x, y, z \in C$, we can define its \alert{dual product} $\cdot$ by
+setting
+\begin{equation*}
+    x \cdot y := \sum_{z \in C} \lambda_{x, y}^z \; z.
+\end{equation*}
+\medskip
+
+\begin{block}{Idea}
+    Knowing how to break a combinatorial object is equivalent
+    to knowing how to combine these.
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Example: unshuffling binary words}
+Let $B$ the set of all binary words and let us consider the coproduct
+\begin{equation*}
+    \Delta : \QQ[B] \to \QQ[B] \otimes \QQ[B]
+\end{equation*}
+on binary words defined by
+\begin{equation*}
+    \Delta(u) :=
+    \sum_{P \sqcup Q = \{1, \dots, |u|\}}
+    u_{|P} \otimes u_{|Q}.
+\end{equation*}
+
+\begin{block}{Example}
+\begin{equation*}
+    \Delta(100) =
+    \epsilon \otimes 100 + 1 \otimes 00
+    + 2 (0 \otimes 10) + 2 (10 \otimes 0) + 00 \otimes 1
+    + 100 \otimes \epsilon
+\end{equation*}
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%o%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Example: shuffling binary words}
+\begin{block}{Proposition \BIB{Reutenauer}{1993}}
+    The shuffle product of binary words is the dual product of the
+    unshuffling coproduct of binary words.
+\end{block}
+\medskip
+
+\begin{block}{Example}
+Since
+\begin{equation*}
+    \textcolor{RoyalBlue}{\Delta(100)} =
+    \epsilon \otimes 100 + 1 \otimes 00
+    + \textcolor{Orange}{2} (\textcolor{BrickRed}{0 \otimes 10})
+    + 2 (10 \otimes 0) + 00 \otimes 1
+    + 100 \otimes \epsilon,
+\end{equation*}
+we observe that $\textcolor{BrickRed}{0 \otimes 10} $ has multiplicity
+$\textcolor{Orange}{2}$ in $\textcolor{RoyalBlue}{\Delta(100)}$.
+\medskip
+
+By duality, this implies that $\textcolor{RoyalBlue}{100}$ has
+multiplicity $\textcolor{Orange}{2}$ in
+$\textcolor{BrickRed}{0 \shuffle 10}$. Indeed,
+\begin{equation*}
+    \textcolor{BrickRed}{0 \shuffle 10} =
+    010 + \textcolor{Orange}{2} (\textcolor{RoyalBlue}{100}).
+\end{equation*}
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Unshuffling permutations}
+If $u$ is a word of integers without multiple occurrence of a same letter,
+the \alert{standardized} $\STD(u)$ of $u$ is the unique permutation of
+$S_{|u|}$ order-isomorphic to~$u$.
+
+\begin{block}{Example}
+\begin{equation*}
+    \STD(82194) = 42153
+\end{equation*}
+\end{block}
+\medskip
+
+\begin{block}{Definition}
+    The \alert{unshuffling coproduct of permutations} is the
+    coproduct on $\QQ[S]$ defined by
+    \begin{equation*}
+        \Delta(\pi) =
+        \sum_{P \sqcup Q = \{1, \dots, |\pi|\}}
+        \STD\left(u_{|P}\right) \otimes \STD\left(u_{|Q}\right).
+    \end{equation*}
+\end{block}
+
+\begin{block}{Example}
+\begin{equation*}
+    \Delta(213) =
+    \epsilon \otimes 213 + 2 (1 \otimes 12) +
+    1 \otimes 21 + 2 (12 \otimes 1) + 21 \otimes 1
+    + 213 \otimes \epsilon
+\end{equation*}
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Shuffling permutations}
+\begin{block}{Definition}
+    The \alert{shuffling product of permutations} is the product
+    $\SHUFFLE$ on $\QQ[S]$ defined as the dual product of the unshuffling
+    product of permutations.
+\end{block}
+
+\begin{block}{Example}
+\begin{equation*}\begin{split}
+    12 \SHUFFLE 21 & =
+    1243 + 1324 + 2 (1342) + 2 (1423) + 3 (1432) +
+    2134 + 2 (2314) \\
+    & + 3 (2341) + 2413 + 2 (2431) + 2 (3124) + 3142 +
+    3 (3214) + 2 (3241) \\
+    & + 3421 + 3 (4123) + 2 (4132) + 2 (4213) + 4231 + 4312
+\end{split}\end{equation*}
+\end{block}
+\medskip
+
+Product first introduced for the study of some Hopf algebras~\BIB{Vargas}{2014}.
+\medskip
+
+Similar to the convolution product and the shifted
+shuffle product \BIB{Duchamp, Hivert, Thibon}{2002}.
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Square permutations}
+\begin{block}{Definition}
+    A permutation $\pi$ is a \alert{square} (w.r.t. the shuffle product
+    $\SHUFFLE$) if there is $\sigma \in S$ such that $\pi$ appears in
+    $\sigma \SHUFFLE \sigma$.
+\end{block}
+\medskip
+
+By duality, $\pi$ is a square if there is $\sigma \in S$ such that
+$\sigma \otimes \sigma$ appears in $\Delta(\pi)$.
+\medskip
+
+\begin{block}{Proposition}
+    A permutation $\pi$ is a square iff $\pi$ can be obtained by
+    shuffling two order-isomorphic words.
+\end{block}
+\medskip
+
+\begin{block}{Example}
+    The permutation
+    \begin{math}
+        \pi :=
+        \textcolor{RoyalBlue}{25}\textcolor{BrickRed}{16}
+        \textcolor{RoyalBlue}{7}\textcolor{BrickRed}{83}
+        \textcolor{RoyalBlue}{4}
+    \end{math}
+    is square since $\pi$ can be obtained by shuffling
+    $\textcolor{BrickRed}{1683}$ and $\textcolor{RoyalBlue}{2574}$.
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Some properties of square permutations}
+The first numbers of square permutations are, size by size,
+\begin{equation*}
+    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
+\end{equation*}
+\medskip
+
+\begin{block}{Proposition}
+    Let $\pi$ be a square permutation and $\sigma$ be a square root of
+    $\pi$. Then,
+    \begin{enumerate}
+        \item $\widetilde{\pi}$ is a square and
+        $\widetilde{\sigma}$ is one of its square roots;
+        \item $\bar \pi$ is a square and $\bar \sigma$ is one of
+        its square roots;
+        \item $\pi^{-1}$ is a square and $\sigma^{-1}$ is one of
+        its square roots.
+    \end{enumerate}
+\end{block}
+\medskip
+
+\begin{block}{Proposition}
+    The set of binary words of length $n$ that are square w.r.t.
+    $\shuffle$ is in one-to-one correspondence with the set of
+    square permutations of length $n$ avoiding the patterns
+    $213$ and~$231$.
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Algorithmic issues}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Combinatorial products and structure coefficients}
-Let $C$ be a set of combinatorial objects and $\QQ[C]$ be the
-$\QQ$-linear span of $C$.
-\medskip
-
-A binary linear product
-\begin{equation*}
-    \cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C].
-\end{equation*}
-on $\QQ[C]$ describes a way to \alert{combine} two objects of
-$C$ to obtain \alert{bigger} objects.
-\medskip
-
-In a general way, for any $x, y \in C$,
-\begin{equation*}
-    x \cdot y = \sum_{z \in C} \lambda_{x, y}^z \; z.
-\end{equation*}
-\medskip
-
-A product is therefore wholly encoded by the coefficients
-$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
-\alert{structure coefficients}.
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Combinatorial coproducts}
-In a dual way, a linear coproduct
-\begin{equation*}
-    \Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C].
-\end{equation*}
-on $\QQ[C]$ describe a way to \alert{break} one object of $C$ into
-\alert{two smaller} pieces.
-\medskip
-
-In a general way, for any $z \in C$,
-\begin{equation*}
-    \Delta(x) = \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
-\end{equation*}
-\medskip
-
-A coproduct is therefore wholly encoded by the coefficients
-$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
-\alert{structure coefficients}.
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Equivalence between products and coproducts}
-If $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$ is a product with
-structure coefficients $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$,
-we can define its \alert{dual coproduct} $\Delta$ by setting
-\begin{equation*}
-    \Delta(x) := \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
-\end{equation*}
-\medskip
-
-Conversely, if $\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$ is
-a coproduct with structure coefficients $\lambda_{x, y}^z \in \QQ$,
-$x, y, z \in C$, we can define its \alert{dual product} $\cdot$ by
-setting
-\begin{equation*}
-    x \cdot y := \sum_{z \in C} \lambda_{x, y}^z \; z.
-\end{equation*}
-\medskip
-
-\begin{block}{Idea}
-    Knowing how to break a combinatorial object is equivalent
-    to knowing how to combine these.
-\end{block}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Example: unshuffling binary words}
-Let $B$ the set of all binary words and let us consider the coproduct
-\begin{equation*}
-    \Delta : \QQ[B] \to \QQ[B] \otimes \QQ[B]
-\end{equation*}
-on binary words defined by
-\begin{equation*}
-    \Delta(u) :=
-    \sum_{P \sqcup Q = \{1, \dots, |u|\}}
-    u_{|P} \otimes u_{|Q}.
-\end{equation*}
-
-\begin{block}{Example}
-\begin{equation*}
-    \Delta(100) =
-    \epsilon \otimes 100 + 1 \otimes 00
-    + 2 (0 \otimes 10) + 2 (10 \otimes 0) + 00 \otimes 1
-    + 100 \otimes \epsilon
-\end{equation*}
-\end{block}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%o%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Example: shuffling binary words}
-\begin{block}{Proposition \BIB{Reutenauer}{1993}}
-    The shuffle product of binary words is the dual product of the
-    unshuffling coproduct of binary words.
-\end{block}
-\medskip
-
-\begin{block}{Example}
-Since
-\begin{equation*}
-    \textcolor{RoyalBlue}{\Delta(100)} =
-    \epsilon \otimes 100 + 1 \otimes 00
-    + \textcolor{Orange}{2} (\textcolor{BrickRed}{0 \otimes 10})
-    + 2 (10 \otimes 0) + 00 \otimes 1
-    + 100 \otimes \epsilon,
-\end{equation*}
-we observe that $\textcolor{BrickRed}{0 \otimes 10} $ has multiplicity
-$\textcolor{Orange}{2}$ in $\textcolor{RoyalBlue}{\Delta(100)}$.
-\medskip
-
-By duality, this implies that $\textcolor{RoyalBlue}{100}$ has
-multiplicity $\textcolor{Orange}{2}$ in
-$\textcolor{BrickRed}{0 \shuffle 10}$. Indeed,
-\begin{equation*}
-    \textcolor{BrickRed}{0 \shuffle 10} =
-    010 + \textcolor{Orange}{2} (\textcolor{RoyalBlue}{100}).
-\end{equation*}
-\end{block}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Unshuffling permutations}
-If $u$ is a word of integers without multiple occurrence of a same letter,
-the \alert{standardized} $\STD(u)$ of $u$ is the unique permutation of
-$S_{|u|}$ order-isomorphic to~$u$.
-
-\begin{block}{Example}
-\begin{equation*}
-    \STD(82194) = 42153
-\end{equation*}
-\end{block}
-\medskip
-
-\begin{block}{Definition}
-    The \alert{unshuffling coproduct of permutations} is the
-    coproduct on $\QQ[S]$ defined by
-    \begin{equation*}
-        \Delta(\pi) =
-        \sum_{P \sqcup Q = \{1, \dots, |\pi|\}}
-        \STD\left(u_{|P}\right) \otimes \STD\left(u_{|Q}\right).
-    \end{equation*}
-\end{block}
-
-\begin{block}{Example}
-\begin{equation*}
-    \Delta(213) =
-    \epsilon \otimes 213 + 2 (1 \otimes 12) +
-    1 \otimes 21 + 2 (12 \otimes 1) + 21 \otimes 1
-    + 213 \otimes \epsilon
-\end{equation*}
-\end{block}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Shuffling permutations}
-\begin{block}{Definition}
-    The \alert{shuffling product of permutations} is the product
-    $\SHUFFLE$ on $\QQ[S]$ defined as the dual product of the unshuffling
-    product of permutations.
-\end{block}
-
-\begin{block}{Example}
-\begin{equation*}\begin{split}
-    12 \SHUFFLE 21 & =
-    1243 + 1324 + 2 (1342) + 2 (1423) + 3 (1432) +
-    2134 + 2 (2314) \\
-    & + 3 (2341) + 2413 + 2 (2431) + 2 (3124) + 3142 +
-    3 (3214) + 2 (3241) \\
-    & + 3421 + 3 (4123) + 2 (4132) + 2 (4213) + 4231 + 4312
-\end{split}\end{equation*}
-\end{block}
-\medskip
-
-Product first introduced for the study of some Hopf algebras~\BIB{Vargas}{2014}.
-\medskip
-
-Similar to the convolution product and the shifted
-shuffle product \BIB{Duchamp, Hivert, Thibon}{2002}.
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Square permutations}
-\begin{block}{Definition}
-    A permutation $\pi$ is a \alert{square} (w.r.t. the shuffle product
-    $\SHUFFLE$) if there is $\sigma \in S$ such that $\pi$ appears in
-    $\sigma \SHUFFLE \sigma$.
-\end{block}
-\medskip
-
-By duality, $\pi$ is a square if there is $\sigma \in S$ such that
-$\sigma \otimes \sigma$ appears in $\Delta(\pi)$.
-\medskip
-
-\begin{block}{Proposition}
-    A permutation $\pi$ is a square iff $\pi$ can be obtained by
-    shuffling two order-isomorphic words.
-\end{block}
-\medskip
-
-\begin{block}{Example}
-    The permutation
-    \begin{math}
-        \pi :=
-        \textcolor{RoyalBlue}{25}\textcolor{BrickRed}{16}
-        \textcolor{RoyalBlue}{7}\textcolor{BrickRed}{83}
-        \textcolor{RoyalBlue}{4}
-    \end{math}
-    is square since $\pi$ can be obtained by shuffling
-    $\textcolor{BrickRed}{1683}$ and $\textcolor{RoyalBlue}{2574}$.
-\end{block}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Some properties of square permutations}
-The first numbers of square permutations are, size by size,
-\begin{equation*}
-    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
-\end{equation*}
-\medskip
-
-\begin{block}{Proposition}
-    Let $\pi$ be a square permutation and $\sigma$ be a square root of
-    $\pi$. Then,
-    \begin{enumerate}
-        \item $\widetilde{\pi}$ is a square and
-        $\widetilde{\sigma}$ is one of its square roots;
-        \item $\bar \pi$ is a square and $\bar \sigma$ is one of
-        its square roots;
-        \item $\pi^{-1}$ is a square and $\sigma^{-1}$ is one of
-        its square roots.
-    \end{enumerate}
-\end{block}
-\medskip
-
-\begin{block}{Proposition}
-    The set of binary words of length $n$ that are square w.r.t.
-    $\shuffle$ is in one-to-one correspondence with the set of
-    square permutations of length $n$ avoiding the patterns
-    $213$ and~$231$.
-\end{block}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{}
-
-\end{frame}
-
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-04-06 slides/slides.tex
Restructuration

@@
 \begin{frame}
-  \frametitle{Permutation pattern}
-
-  \begin{definition}
-    A permutation $\pi$ is said to \emph{contain} the permutation $\sigma$ if there
-    exists a subsequence of (not necessarily consecutive) entries of $\pi$ that
-    has the same relative order as $\sigma$, and in this case $\sigma$ is said
-    to be a pattern of $\pi$, written $\sigma \leq \pi$.
-
-    \medskip
-
-    Otherwise, $\pi$ is said to \emph{avoid} the permutation $\sigma$.
-  \end{definition}
-
-  \medskip
-
-  \begin{exampleblock}{Example}
-    \begin{itemize}
-      \item
-        The permutation $\pi = 391867452$ contains the pattern $\sigma = 51342$,
-        as can be seen in the highlighted subsequence of
-        $\pi = 3\mathbf{91}8\mathbf{674}52$
-        (or $\pi = 3\mathbf{91}8\mathbf{67}4\mathbf{5}2$ or
-        $\pi = 3\mathbf{91}8\mathbf{67}45\mathbf{2}$ or
-        $\pi = 3\mathbf{91}867\mathbf{452}$.)
-
-      \item
-       Since the permutation $\pi = 391867452$ contains no increasing subsequence
-       of length four, $\pi$ avoids $1234$.
-    \end{itemize}
-  \end{exampleblock}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}
-  \frametitle{Square permutation}
-
-  \begin{definition}
-    A permutation is said to be a \emph{square} if it can be obtained by
-    shuffling two order-isomorphic patterns
-  \end{definition}
-
-  \medskip
-
-  \begin{exampleblock}{Example}
-    \begin{itemize}
-      \item
-        The permutation $\pi = 25167834$ is a square since it can obtained by
-        shuffling two order-isomorphic (\emph{i.e.} $1342$)
-        patterns $1683$ and $2574$ as can be seen in the highlighted subsequence of
-        $\pi = \mathbf{25}16\mathbf{7}83\mathbf{4}$.
-
-      \medskip
-
-      \item
-       $\pi = 4123$ and $1432$ are not squares.
-    \end{itemize}
-  \end{exampleblock}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}
-  \frametitle{Words}
-
-  \begin{definition}
-    The shuffle operator on words is defined inductively as follows:
-    \begin{align*}
-      u \shuffle \epsilon &= \epsilon \shuffle u = \{u\}
-      \\
-      (ua \shuffle vb) &= (ua \shuffle v)b \cup (u \shuffle vb)a
-    \end{align*}
-  \end{definition}
-
-  \medskip
-
-  \begin{block}{Key results}
-    \begin{itemize}
-      \item
+  \frametitle{Key results}
+
 Given two words $v_1$ and $v_2$, the set $v_1 \shuffle v_2$
 can be computed in
-      $O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)$
-      time.
-
-      \smallskip
-
-      \item
-      Given three words $u$, $v_1$, and $v_2$, deciding if $u$ is a shuffle
-      of $v_1$ and $v_2$ can be done in
-      $O\left(n^2 / \log(n)\right)$ time.
-
-      \smallskip
-
-      \item
+\begin{equation*}
+    O\left((|v_1| + |v_2|) \; \binom{|v_1| + |v_2|}{|v_1|}\right)
+\end{equation*}
+time \BIB{Spehner}{1986}.
+\bigskip
+
+Given three words $u$, $v_1$, and $v_2$, deciding if $u$ is a
+shuffle of $v_1$ and $v_2$ can be done in
+\begin{equation*}
+O\left(\frac{|u|^2}{\log(|u|)}\right)
+\end{equation*}
+time \BIB{van Leeuwen, Nivat}{1982}.
+\bigskip
+
 Given a word $u$, deciding if there is a word $v$ such that $u$ is
-      in $v \shuffle v$ is \NPC.
+in $v \shuffle v$ is \NPC{} \BIB{Rizzi, Vialette}{2013}
+\BIB{Buss, Soltys}{2014}.
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{???}
+JE PENSE QU'IL FAUDRAIT PARLER ICI DES COUPLAGES PARFAITS POUR LES MOTS.
+\medskip
+
+ÇA PERMET DE FIXER QUELQUES DEFINITIONS POUR PREPARER LE TERRAIN POUR
+NOTRE EXTENSION POUR LES PERMU DANS LA PARTIE ALGO.
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%o%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Main motivations}
+Let $S_n$ be the set of permutations of size $n$.
+\medskip
+
+There are several products on permutations, analogs of the shuffle
+product of words. Among these, there are
+\begin{itemize}
+    \item the shifted shuffle product \BIB{Duchamp, Hivert, Thibon}{2002};
+    \medskip
+
+    \item the convolution product \BIB{Duchamp, Hivert, Thibon}{2002};
+    \medskip
+
+    \item the supershuffle \BIB{Vargas}{2014}.
 \end{itemize}
-  \end{block}
-
+\medskip
+
+\begin{block}{Main question}
+    How the previous key results about the shuffle of words translate
+    for the supershuffle of permutations?
+\end{block}
 \end{frame}
@@
 \begin{equation*}
-    \Delta(x) = \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
+    \Delta(z) = \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
 \end{equation*}
@@
 \begin{frame} \frametitle{Equivalence between products and coproducts}
-If $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$ is a product with
+\begin{block}{Key idea}
+    Knowing how to break a combinatorial object is equivalent
+    to knowing how to combine these.
+\end{block}
+\medskip
+
+Indeed, if $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$ is a product with
 structure coefficients $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$,
@@
 \begin{equation*}
-    \Delta(x) := \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
+    \Delta(z) := \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
 \end{equation*}
@@
 \end{equation*}
-\medskip
-
-\begin{block}{Idea}
-    Knowing how to break a combinatorial object is equivalent
-    to knowing how to combine these.
-\end{block}
 \end{frame}
@@
 \begin{block}{Definition}
-    The \alert{shuffling product of permutations} is the product
-    $\SHUFFLE$ on $\QQ[S]$ defined as the dual product of the unshuffling
-    product of permutations.
-\end{block}
+    The \alert{shuffling product of permutations} (supershuffle) is the
+    product $\SHUFFLE$ on $\QQ[S]$ defined as the dual product of the
+    unshuffling product of permutations.
+\end{block}
+\medskip
+
+\begin{block}{Proposition}
+The permutations appearing in $\sigma \SHUFFLE \nu$ are exactly the
+one obtained by shuffling two words $u$ and $v$ respectively
+order-isomorphic to $\sigma$ and $\nu$.
+\end{block}
+\medskip
 
@@
     1243 + 1324 + 2 (1342) + 2 (1423) + 3 (1432) +
-    2134 + 2 (2314) \\
+    2134 + 2 (2314) \\[.5em]
     & + 3 (2341) + 2413 + 2 (2431) + 2 (3124) + 3142 +
-    3 (3214) + 2 (3241) \\
+    3 (3214) + 2 (3241) \\[.5em]
     & + 3421 + 3 (4123) + 2 (4132) + 2 (4213) + 4231 + 4312
@@
 \end{block}
-\medskip
-
-Product first introduced for the study of some Hopf algebras~\BIB{Vargas}{2014}.
-\medskip
-
-Similar to the convolution product and the shifted
-shuffle product \BIB{Duchamp, Hivert, Thibon}{2002}.
 \end{frame}
@@
 \end{equation*}
-\medskip
+%\medskip
 
@@
         $\widetilde{\sigma}$ is one of its square roots;
+        \smallskip
+
         \item $\bar \pi$ is a square and $\bar \sigma$ is one of
         its square roots;
+        \smallskip
+
         \item $\pi^{-1}$ is a square and $\sigma^{-1}$ is one of
@@
 \section{Algorithmic issues}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Permutation pattern}
+
+  \begin{definition}
+    A permutation $\pi$ is said to \emph{contain} the permutation $\sigma$ if there
+    exists a subsequence of (not necessarily consecutive) entries of $\pi$ that
+    has the same relative order as $\sigma$, and in this case $\sigma$ is said
+    to be a pattern of $\pi$, written $\sigma \leq \pi$.
+
+    \medskip
+
+    Otherwise, $\pi$ is said to \emph{avoid} the permutation $\sigma$.
+  \end{definition}
+
+  \medskip
+
+  \begin{exampleblock}{Example}
+    \begin{itemize}
+      \item
+        The permutation $\pi = 391867452$ contains the pattern $\sigma = 51342$,
+        as can be seen in the highlighted subsequence of
+        $\pi = 3\mathbf{91}8\mathbf{674}52$
+        (or $\pi = 3\mathbf{91}8\mathbf{67}4\mathbf{5}2$ or
+        $\pi = 3\mathbf{91}8\mathbf{67}45\mathbf{2}$ or
+        $\pi = 3\mathbf{91}867\mathbf{452}$.)
+
+      \item
+       Since the permutation $\pi = 391867452$ contains no increasing subsequence
+       of length four, $\pi$ avoids $1234$.
+    \end{itemize}
+  \end{exampleblock}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Square permutation}
+
+  \begin{definition}
+    A permutation is said to be a \emph{square} if it can be obtained by
+    shuffling two order-isomorphic patterns
+  \end{definition}
+
+  \medskip
+
+  \begin{exampleblock}{Example}
+    \begin{itemize}
+      \item
+        The permutation $\pi = 25167834$ is a square since it can obtained by
+        shuffling two order-isomorphic (\emph{i.e.} $1342$)
+        patterns $1683$ and $2574$ as can be seen in the highlighted subsequence of
+        $\pi = \mathbf{25}16\mathbf{7}83\mathbf{4}$.
+
+      \medskip
+
+      \item
+       $\pi = 4123$ and $1432$ are not squares.
+    \end{itemize}
+  \end{exampleblock}
+
+\end{frame}
 

### 2016-04-07 slides/slides.tex
Typo
Suite

@@
 \usetheme{Boadilla}
+%\usetheme{default}
 \usefonttheme{serif}
@@
 
-\usepackage{euler,palatino}
+%\usepackage{euler,palatino}
+\usepackage{lmodern}
 \usepackage{amsmath,amssymb}
@@
     \frame<handout:0>{
-        \frametitle{Plan}
+        \frametitle{Contents}
         \tableofcontents[
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
+\section{Background}
 
@@
     \begin{equation*}\begin{split}
-        \La\Lb \shuffle \textcolor{BrickRed}{\Lb\La} & =
-        \La\Lb\textcolor{BrickRed}{\Lb\La}
-        + \La\textcolor{BrickRed}{\Lb}\Lb\textcolor{BrickRed}{\La}
-        + \La\textcolor{BrickRed}{\Lb\La}\Lb
-        + \textcolor{BrickRed}{\Lb}\La\Lb\textcolor{BrickRed}{\La}
-        + \textcolor{BrickRed}{\Lb}\La\textcolor{BrickRed}{\La}\Lb
-        + \textcolor{BrickRed}{\Lb\La}\La\Lb \\
+        \textcolor{RoyalBlue}{\La\Lb} \shuffle \textcolor{BrickRed}{\Lb\La}
+        & =
+        \textcolor{RoyalBlue}{\La\Lb}\textcolor{BrickRed}{\Lb\La}
+        + \textcolor{RoyalBlue}{\La}\textcolor{BrickRed}{\Lb}%
+        \textcolor{RoyalBlue}{\Lb}\textcolor{BrickRed}{\La}
+        + \textcolor{RoyalBlue}{\La}\textcolor{BrickRed}{\Lb\La}%
+        \textcolor{RoyalBlue}{\Lb}
+        + \textcolor{BrickRed}{\Lb}\textcolor{RoyalBlue}{\La\Lb}%
+        \textcolor{BrickRed}{\La}
+        + \textcolor{BrickRed}{\Lb}\textcolor{RoyalBlue}{\La}%
+        \textcolor{BrickRed}{\La}\textcolor{RoyalBlue}{\Lb}
+        + \textcolor{BrickRed}{\Lb\La}\textcolor{RoyalBlue}{\La\Lb} \\
         &
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{???}
-JE PENSE QU'IL FAUDRAIT PARLER ICI DES COUPLAGES PARFAITS POUR LES MOTS.
-\medskip
-
-ÇA PERMET DE FIXER QUELQUES DEFINITIONS POUR PREPARER LE TERRAIN POUR
-NOTRE EXTENSION POUR LES PERMU DANS LA PARTIE ALGO.
+\begin{frame} \frametitle{Square words}
+\begin{block}{Definition}
+    A word $u$ is a \alert{square} (w.r.t. the shuffle product
+    $\shuffle$) if there is a word $v$ such that $u$ appears
+    in~$v \shuffle v$.
+\end{block}
+\medskip
+
+\begin{block}{Example}
+    The word
+    \begin{math}
+        u := \textcolor{RoyalBlue}{3}\textcolor{BrickRed}{3}
+        \textcolor{RoyalBlue}{12}\textcolor{BrickRed}{1}
+        \textcolor{RoyalBlue}{2}\textcolor{BrickRed}{22}
+    \end{math}
+    is a square since $u$ can be obtained by shuffling $3122$ with itself.
+\end{block}
+\medskip
+
+The first numbers of square binary words of length $2n$ are
+\begin{equation*}
+    1, 2, 6, 22, 82, 320, 1268, 5102, 20632, 83972.
+\end{equation*}
+
+\begin{block}{Open problem}
+    Enumeration of square (binary) words.
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Recognizing square words}
+\begin{block}{Definition}
+    A \alert{perfect matching} of a word $u$ is a graph $(V, E)$ such
+    that
+    \begin{itemize}
+        \item $V := \{(u_i, i) : i \in \{1, \dots, |u|\}$;
+        \item $(u_i, i) \-- (u_j, j) \in E$ implies $u_i = u_j$;
+        \item every vertex of $V$ belongs to exactly one edge of $E$.
+    \end{itemize}
+\end{block}
+
+\begin{block}{Example}
+    The word $21212233$ admits the perfect matching \vspace{-1em}
+    \begin{equation*}
+    \begin{split}
+    \begin{tikzpicture}[xscale=.5,yscale=.3,inner sep=1pt,node distance=1.5cm]
+        \tikzstyle{Vertex}=[draw,circle,minimum size=.35cm,font=\footnotesize]
+        \draw node[Vertex,fill=RoyalBlue!30](U01){$(2, 1)$};
+        \draw node[Vertex,fill=RoyalBlue!30,right of=U01](U02) {$(1, 2)$};
+        \draw node[Vertex,fill=RoyalBlue!30,right of=U02](U03){$(2, 3)$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U03](U04) {$(1, 4)$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U04](U05) {$(2, 5)$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U05](U06){$(2, 6)$};
+        \draw node [Vertex,fill=RoyalBlue!30,right of=U06](U07){$(3, 7)$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U07](U08){$(3, 8)$};
+        \draw [thick,>=latex']
+        (U01.north) .. controls ($ (U01.north) + (0,2.5) $)
+            and ($ (U06.north) + (0,2.5) $) .. (U06.north);
+        \draw [thick,>=latex']
+        (U02.north) .. controls ($ (U02.north) + (0,1.5) $)
+            and ($ (U04.north) + (0,1.5) $) .. (U04.north);
+        \draw [thick,>=latex']
+        (U03.south) .. controls ($ (U03.south) + (0,-1.5) $)
+            and ($ (U05.south) + (0,-1.5) $) .. (U05.south);
+        \draw [thick,>=latex']
+        (U07.north) .. controls ($ (U07.north) + (0,1.5) $)
+            and ($ (U08.north) + (0,1.5) $) .. (U08.north);
+    \end{tikzpicture}
+    \end{split}\,.
+    \end{equation*}
+    \vspace{-1.5em}
+\end{block}
+
+\begin{block}{Definition}
+    A perfect matching $(V, E)$ of a word $u$ is \alert{inclusion-free}
+    if there are no edges $(u_i, i) \-- (u_j, j)$ and
+    $(u_k, k) \-- (u_\ell, \ell)$ of $E$ such that $i < k < \ell < j$.
+\end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame} \frametitle{Recognizing square words}
+
+\begin{block}{Proposition \BIB{Rizzi, Vialette}{2013}}
+    A word $u$ is a square iff $u$ admits an inclusion-free perfect
+    matching.
+\end{block}
+%\medskip
+
+\begin{block}{Example}
+    The square word
+    \begin{math}
+        \textcolor{RoyalBlue}{3}\textcolor{BrickRed}{3}
+        \textcolor{RoyalBlue}{12}\textcolor{BrickRed}{1}
+        \textcolor{RoyalBlue}{2}\textcolor{BrickRed}{22}
+    \end{math}
+    admits the perfect matching \vspace{-1.5em}
+    \begin{equation*}
+        \begin{split}
+        \begin{tikzpicture}[xscale=.5,yscale=.2,inner sep=1pt,node distance=1.5cm]
+            \tikzstyle{Vertex}=[draw,circle,minimum size=.35cm,font=\footnotesize]
+            \draw node[Vertex,fill=RoyalBlue!30](U01){$(3, 1)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U01](U02) {$(3, 2)$};
+            \draw node[Vertex,fill=RoyalBlue!30,right of=U02](U03){$(1, 3)$};
+            \draw node[Vertex,fill=RoyalBlue!30,right of=U03](U04) {$(2, 4)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U04](U05) {$(1, 5)$};
+            \draw node[Vertex,fill=RoyalBlue!30,right of=U05](U06){$(2, 6)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U06](U07){$(2, 7)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U07](U08){$(2, 8)$};
+            \draw [thick,>=latex']
+            (U01.north) .. controls ($ (U01.north) + (0,2.25) $)
+                and ($ (U02.north) + (0,2.25) $) .. (U02.north);
+            \draw [thick,>=latex']
+            (U03.north) .. controls ($ (U03.north) + (0,2.25) $)
+                and ($ (U05.north) + (0,2.25) $) .. (U05.north);
+            \draw [thick,>=latex']
+            (U04.north) .. controls ($ (U04.north) + (0,4.5) $)
+                and ($ (U07.north) + (0,4.5) $) .. (U07.north);
+            \draw [thick,>=latex']
+            (U05.south) .. controls ($ (U05.south) + (0,-2.5) $)
+                and ($ (U08.south) + (0,-2.5) $) .. (U08.south);
+            \draw [thick,>=latex']
+            (U06.north) .. controls ($ (U06.north) + (0,2.5) $)
+                and ($ (U08.north) + (0,2.5) $) .. (U08.north);
+        \end{tikzpicture}
+        \end{split}
+    \end{equation*}
+
+    \vspace{-1.5em}
+    which is inclusion-free.
+\end{block}
+%\medskip
+
+\begin{block}{Example}
+    The word $1221$ is not a square. Its only perfect matching\vspace{-1.5em}
+    \begin{equation*}
+        \begin{split}
+        \begin{tikzpicture}[xscale=.5,yscale=.3,inner sep=1pt,node distance=1.5cm]
+            \tikzstyle{Vertex}=[draw,circle,minimum size=.35cm,font=\footnotesize]
+            \draw node[Vertex,fill=RoyalBlue!30](U01){$(1, 1)$};
+            \draw node[Vertex,fill=RoyalBlue!30,right of=U01](U02) {$(2, 2)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U02](U03){$(2, 3)$};
+            \draw node [Vertex,fill=BrickRed!30,right of=U03](U04) {$(1, 4)$};
+            \draw [thick,>=latex']
+            (U01.north) .. controls ($ (U01.north) + (0,2.5) $)
+                and ($ (U04.north) + (0,2.5) $) .. (U04.north);
+            \draw [thick,>=latex']
+            (U02.north) .. controls ($ (U02.north) + (0,1.5) $)
+                and ($ (U03.north) + (0,1.5) $) .. (U03.north);
+        \end{tikzpicture}\vspace{-1.5em}
+        \end{split}
+    \end{equation*}
+
+    \vspace{-1.5em}
+    is not inclusion-free.
+\end{block}
 \end{frame}
@@
 product of words. Among these, there are
+\smallskip
+
 \begin{itemize}
@@
 
-\begin{block}{Main question}
-    How the previous key results about the shuffle of words translate
-    for the supershuffle of permutations?
+\begin{block}{Main questions}
+    \begin{itemize}
+        \item How the previous key results about the shuffle of words
+        translate for the supershuffle of permutations?
+        \medskip
+
+        \item
+    \end{itemize}
 \end{block}
@@
         \sum_{P \sqcup Q = \{1, \dots, |\pi|\}}
-        \STD\left(u_{|P}\right) \otimes \STD\left(u_{|Q}\right).
+        \STD\left(\pi_{|P}\right) \otimes \STD\left(\pi_{|Q}\right).
     \end{equation*}
@@
     is square since $\pi$ can be obtained by shuffling
-    $\textcolor{BrickRed}{1683}$ and $\textcolor{RoyalBlue}{2574}$.
+    $\textcolor{BrickRed}{1683}$ and $\textcolor{RoyalBlue}{2574}$
+    and
+    \begin{math}
+        \STD(\textcolor{BrickRed}{1683})
+        = \STD(\textcolor{RoyalBlue}{2574})
+        = 1342
+    \end{math}.
 \end{block}
@@
 \begin{frame} \frametitle{Some properties of square permutations}
-The first numbers of square permutations are, size by size,
-\begin{equation*}
-    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
+The first numbers of square permutations of size $2$ are
+\begin{equation*}
+    1, 2, 20, 504, 21032, 1293418.
 \end{equation*}

### 2016-04-07 slides/slides.tex
Suite

@@
         \item $V := \{(u_i, i) : i \in \{1, \dots, |u|\}$;
-        \item $(u_i, i) \-- (u_j, j) \in E$ implies $u_i = u_j$;
-        \item every vertex of $V$ belongs to exactly one edge of $E$.
+        \item every vertex of $V$ belongs to exactly one edge of $E$;
+        \item $(u_i, i) \-- (u_j, j) \in E$ implies $u_i = u_j$.
     \end{itemize}
@@
     \begin{itemize}
-        \item How the previous key results about the shuffle of words
-        translate for the supershuffle of permutations?
-        \medskip
-
-        \item
+        \item Combinatorial properties of ``square permutations''
+        w.r.t. the supershuffle?
+        \medskip
+
+        \item Complexity of the recognition of ``square permutations''?
     \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Algebraic issues}
+\section{Algebraic and combinatorial issues}
 
@@
 \begin{frame}
-  \frametitle{Permutation pattern}
+\frametitle{Permutation patterns}
 
   \begin{definition}
-    A permutation $\pi$ is said to \emph{contain} the permutation $\sigma$ if there
-    exists a subsequence of (not necessarily consecutive) entries of $\pi$ that
+    A permutation $\pi$ \alert{contains} a permutation $\sigma$ if there
+    exists a subsequence of (not necessarily consecutive) letters of $\pi$ that
     has the same relative order as $\sigma$, and in this case $\sigma$ is said
-    to be a pattern of $\pi$, written $\sigma \leq \pi$.
-
-    \medskip
-
-    Otherwise, $\pi$ is said to \emph{avoid} the permutation $\sigma$.
+    to be a \alert{pattern} of $\pi$, written $\sigma \leq \pi$.
+    \smallskip
+
+    Otherwise, $\pi$ \alert{avoids} $\sigma$.
   \end{definition}
@@
   \begin{exampleblock}{Example}
+
+        The permutation
+        $\pi = 3\textcolor{BrickRed}{91}8\textcolor{BrickRed}{674}52$
+        contains the pattern $\sigma := 51342$ since
+        $\STD(\textcolor{BrickRed}{91674}) = 51342$.
+  \end{exampleblock}
+  \medskip
+
+  \begin{block}{Definition}
+    The \alert{pattern involvement problem} is the decision problem
+    consisting in, given $\pi, \sigma \in S$, decide whether $\sigma \leq \pi$.
+  \end{block}
+  \medskip
+
+  \begin{block}{Proposition \BIB{Bose, Buss, Lubiw}{1998}}
+    The pattern involvement problem is $\NPC$.
+  \end{block}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{frame}
+  %\frametitle{Square permutation}
+
+  %\begin{definition}
+    %A permutation is said to be a \emph{square} if it can be obtained by
+    %shuffling two order-isomorphic patterns
+  %\end{definition}
+
+  %\medskip
+
+  %\begin{exampleblock}{Example}
+    %\begin{itemize}
+      %\item
+        %The permutation $\pi = 25167834$ is a square since it can obtained by
+        %shuffling two order-isomorphic (\emph{i.e.} $1342$)
+        %patterns $1683$ and $2574$ as can be seen in the highlighted subsequence of
+        %$\pi = \mathbf{25}16\mathbf{7}83\mathbf{4}$.
+
+      %\medskip
+
+      %\item
+       %$\pi = 4123$ and $1432$ are not squares.
+    %\end{itemize}
+  %\end{exampleblock}
+
+%\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+  \frametitle{Recognizing square permutations}
+
+    \begin{block}{Definition}
+        A \alert{perfect matching} of a permutation $\pi \in S_n$ is a
+        directed graph $(V, A)$ such that
         \begin{itemize}
-      \item
-        The permutation $\pi = 391867452$ contains the pattern $\sigma = 51342$,
-        as can be seen in the highlighted subsequence of
-        $\pi = 3\mathbf{91}8\mathbf{674}52$
-        (or $\pi = 3\mathbf{91}8\mathbf{67}4\mathbf{5}2$ or
-        $\pi = 3\mathbf{91}8\mathbf{67}45\mathbf{2}$ or
-        $\pi = 3\mathbf{91}867\mathbf{452}$.)
-
-      \item
-       Since the permutation $\pi = 391867452$ contains no increasing subsequence
-       of length four, $\pi$ avoids $1234$.
+            \item $V := \{(\pi_i, i) : i \in \{1, \dots, n\}\}$;
+            \item every vertex of $V$ belongs to exactly one arc of $A$.
         \end{itemize}
-  \end{exampleblock}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}
-  \frametitle{Square permutation}
-
-  \begin{definition}
-    A permutation is said to be a \emph{square} if it can be obtained by
-    shuffling two order-isomorphic patterns
-  \end{definition}
-
-  \medskip
-
-  \begin{exampleblock}{Example}
-    \begin{itemize}
-      \item
-        The permutation $\pi = 25167834$ is a square since it can obtained by
-        shuffling two order-isomorphic (\emph{i.e.} $1342$)
-        patterns $1683$ and $2574$ as can be seen in the highlighted subsequence of
-        $\pi = \mathbf{25}16\mathbf{7}83\mathbf{4}$.
-
-      \medskip
-
-      \item
-       $\pi = 4123$ and $1432$ are not squares.
-    \end{itemize}
-  \end{exampleblock}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{frame}
-  \frametitle{Hardness}
-  \framesubtitle{Oriented perfect matching}
+    \end{block}
+    \medskip
 
   \begin{definition}[Property $\mathbf{P_1}$]
-    Let $\pi$ be a permutation. An oriented perfect matching $\mathcal{M}$
-    on $\pi$ is said to have property $\mathbf{P_1}$ if it avoids all the
-    six patterns \InclusionLL, \InclusionLR, \InclusionRL,
+    Let $\pi \in S$. A perfect matching $(V, A)$ of $\pi$ has property
+    $\mathbf{P_1}$ if it avoids all the six patterns
+    \InclusionLL, \InclusionLR, \InclusionRL,
     \InclusionRR, \CrossingLR, and \CrossingRL.
@@
   \begin{definition}[Property $\mathbf{P_2}$]
-    Let $\pi$ be a permutation. An oriented perfect matching $\mathcal{M}$
-    on $\pi$ is said to have property $\mathbf{P_2}$ if, for any two
-    distinct arcs $(\pi(a), \pi(a'))$ and $(\pi(b), \pi(b'))$ in $\mathcal{M}$,
-    we have $\pi(a) < \pi(b)$ if and only if $\pi(a') < \pi(b')$.
+    Let $\pi \in S$. A perfect matching $(V, A)$ of $\pi$ has property
+    $\mathbf{P_2}$ if, for all arcs $(\pi_i, i) \to (\pi_j, j)$ and
+    $(\pi_k, k) \to (\pi_\ell, \ell)$ of $A$, $\pi_i < \pi_k$ iff
+    $\pi_j < \pi_\ell$.
   \end{definition}
@@
 \begin{frame}
-  \frametitle{Hardness}
-  \framesubtitle{Oriented perfect matching}
-
-  \begin{theorem}
-    Let $\pi$ be a permutation. The following statements are equivalent:
-    \begin{enumerate}
-      \item The permutation $\pi$ is a square.
-      \item There exists an oriented perfect matching $\mathcal{M}$
-      on $\pi$ satisfying~$\mathbf{P_1}$ and~$\mathbf{P_2}$.
-    \end{enumerate}
-  \end{theorem}
+  \frametitle{Recognizing square permutations}
+
+  \begin{block}{Proposition}
+    A permutation $\pi$ is a square (w.r.t. the shuffle product $\SHUFFLE$)
+    iff $\pi$ admits a perfect matching satisfying~$\mathbf{P_1}$
+    and~$\mathbf{P_2}$.
+  \end{block}
 
@@
   \begin{exampleblock}{Example}
+    The permutation $\pi := 183927B5C6A4$ is a square since it admits the
+    perfect matching
+    \begin{equation*}
+    \begin{split}
     \begin{tikzpicture}[scale=0.35,inner sep=2pt,node distance=1cm]
-        \draw node [draw,circle,minimum size=.55cm,fill=black!10](U01){$1$};
-        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U01]
-            (U02) {$8$};
-        \draw node [draw,circle,minimum size=.55cm,right of=U02](U03){$3$};
-        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U03]
-            (U04) {$9$};
-        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U04]
-            (U05) {$2$};
-        \draw node [draw,circle,minimum size=.55cm,right of=U05](U06){$7$};
-        \draw node [draw,circle,minimum size=.55cm,right of=U06](U07){$B$};
-        \draw node [draw,circle,minimum size=.55cm,right of=U07](U08){$5$};
-        \draw node [draw,circle,minimum size=.55cm,right of=U08](U09){$C$};
-        \draw node [draw,circle,minimum size=.55cm,right of=U09](U10){$6$};
-        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U10]
-            (U11) {$A$};
-        \draw node [draw,circle,minimum size=.55cm,fill=black!10,right of=U11]
-            (U12) {$4$};
+        \tikzstyle{Vertex}=[draw,circle,minimum size=.45cm,font=\footnotesize]
+        \draw node[Vertex,fill=RoyalBlue!30](U01){$1$};
+        \draw node[Vertex,fill=RoyalBlue!30,right of=U01](U02){$8$};
+        \draw node[Vertex,fill=BrickRed!30,right of=U02](U03){$3$};
+        \draw node[Vertex,fill=RoyalBlue!30,right of=U03](U04){$9$};
+        \draw node[Vertex,fill=RoyalBlue!30,right of=U04](U05){$2$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U05](U06){$7$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U06](U07){$B$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U07](U08){$5$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U08](U09){$C$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U09](U10){$6$};
+        \draw node [Vertex,fill=RoyalBlue!30,right of=U10](U11){$A$};
+        \draw node [Vertex,fill=RoyalBlue!30,right of=U11](U12){$4$};
         \draw [thick,->,>=latex']
@@
         \draw [thick,->,>=latex']
-        (U02.north) .. controls ($ (U02.north) + (0,2.25) $)
-            and ($ (U06.north) + (0,2.25) $) .. (U06.north);
+        (U02.south) .. controls ($ (U02.south) + (0,-2.25) $)
+            and ($ (U06.south) + (0,-2.25) $) .. (U06.south);
         \draw [thick,->,>=latex']
@@
         \draw [thick,->,>=latex']
-        (U05.north) .. controls ($ (U05.north) + (0,2.5) $)
-            and ($ (U08.north) + (0,2.5) $) .. (U08.north);
+        (U05.south) .. controls ($ (U05.south) + (0,-2.5) $)
+            and ($ (U08.south) + (0,-2.5) $) .. (U08.south);
         \draw [thick,->,>=latex']
@@
     \end{tikzpicture}
-
-    {\small%
-    An oriented perfect matching $\mathcal{M}$ on the permutation
-    $\pi = 183927B5C6A4$ satisfying the properties $\mathbf{P_1}$
-    and $\mathbf{P_2}$. From $\mathcal{M}$, it follows that $\pi$ is
-    a square as it appears in the shuffle of $1892A4$ and $37B5C6$,
-    both being order-isomorphic to $145263$.
-    }%end small
+    \end{split}
+    \end{equation*}
+    satisfying $\mathbf{P_1}$ and~$\mathbf{P_2}$.
+    \smallskip
+
+    Moreover, this shows that $\pi$ appears in the shuffle of the
+    words $\textcolor{RoyalBlue}{1892A4}$ and
+    $\textcolor{BrickRed}{37B5C6}$, both order isomorphic to
+    $145263$.
   \end{exampleblock}
@@
 \begin{frame}
-  \frametitle{Hardness}
-
-  \begin{theorem}
+  \frametitle{Hardness of recognizing square permutations}
+
+  \begin{block}{Proposition}
     Deciding whether a permutation is a square is \NPC.
-  \end{theorem}
-
-  \medskip
-
-  \begin{block}{Remarks}
+  \end{block}
+
+  \medskip
+
+  \begin{block}{Main ideas of the proof}
     \begin{itemize}
-      \item
-      We reduce from the problem of deciding wether a pattern $\sigma$ occurs in a
-      permutation $\pi$.
-
-      \smallskip
-
-      \item
-      Construct a new permutation $\mu$ such that
-      $\sigma$ occurs in $\pi$ if and only if
-      there exists an oriented perfect matching $\mathcal{M}$ on $\mu$
-      satisfying $\mathbf{P_1}$ and $\mathbf{P_2}$.
+        \item We show that the pattern involvement problem is reducible
+        in polynomial time to our decision problem.
+        \medskip
+
+        \item For this, given two permutations $\sigma$ and $\pi$, we
+        show that $\sigma \leq \pi$ iff $\mu_{\sigma, \pi}$ is a square,
+        where $\mu_{\sigma, \pi}$ is a permutation constructed in polynomial
+        time from $\sigma$ and $\pi$.
+        \medskip
+
+        \item To prove that $\mu_{\sigma, \pi}$ is a square, we show
+        that $\mu_{\sigma, \pi}$ admits a perfect matching satisfying
+        $\mathbf{P_1}$ and~$\mathbf{P_2}$.
+
     \end{itemize}
@@
 \begin{frame}
-  \frametitle{Hardness}
+  \frametitle{Hardness of recognizing square permutations}
 
@@
   \begin{block}{Combinatorics}
-    How many permutations of length $2n$ are squares?
-
+    \begin{itemize}
+        \item Enumeration of square permutations?
         \smallskip
 
-    How many $(213,231)$-avoiding
-    permutations of length $2n$ are squares? (Equivalently,
-    how many binary strings of length $2n$ are squares?)
-  \end{block}
-
-  \medskip
-
-  \begin{block}{Combinatorics}
+        \item Enumeration of square permutations avoiding $213$ and $231$?
+    \end{itemize}
+  \end{block}
+  \medskip
+
+  \begin{block}{Algorithms}
+
     Given two permutations $\pi$ and~$\sigma$, how hard is the problem of
@@
   \end{block}
-
   \medskip
@@
   \begin{block}{Algebra}
-    \ldots
+        How algebraic properties of the associative algebra $(\QQ[S], \SHUFFLE)$
+        or of the coalgebra $(\QQ[S], \Delta)$ are connected to
+        combinatorial and algorithmic properties of square permutations?
   \end{block}

### 2016-04-07 slides/slides.tex
New example

@@
   \begin{exampleblock}{Example}
-    The permutation $\pi := 183927B5C6A4$ is a square since it admits the
+    The permutation $\pi := 35867241$ is a square since it admits the
     perfect matching
@@
         \tikzstyle{Vertex}=[draw,circle,minimum size=.45cm,font=\footnotesize]
-        \draw node[Vertex,fill=RoyalBlue!30](U01){$1$};
-        \draw node[Vertex,fill=RoyalBlue!30,right of=U01](U02){$8$};
-        \draw node[Vertex,fill=BrickRed!30,right of=U02](U03){$3$};
-        \draw node[Vertex,fill=RoyalBlue!30,right of=U03](U04){$9$};
-        \draw node[Vertex,fill=RoyalBlue!30,right of=U04](U05){$2$};
-        \draw node [Vertex,fill=BrickRed!30,right of=U05](U06){$7$};
-        \draw node [Vertex,fill=BrickRed!30,right of=U06](U07){$B$};
-        \draw node [Vertex,fill=BrickRed!30,right of=U07](U08){$5$};
-        \draw node [Vertex,fill=BrickRed!30,right of=U08](U09){$C$};
-        \draw node [Vertex,fill=BrickRed!30,right of=U09](U10){$6$};
-        \draw node [Vertex,fill=RoyalBlue!30,right of=U10](U11){$A$};
-        \draw node [Vertex,fill=RoyalBlue!30,right of=U11](U12){$4$};
+        \draw node[Vertex,fill=RoyalBlue!30](U01){$(3, 1)$};
+        \draw node[Vertex,fill=BrickRed!30,right of=U01](U02){$(5, 2)$};
+        \draw node[Vertex,fill=BrickRed!30,right of=U02](U03){$(8, 3)$};
+        \draw node[Vertex,fill=RoyalBlue!30,right of=U03](U04){$(6, 4)$};
+        \draw node[Vertex,fill=BrickRed!30,right of=U04](U05){$(7, 5)$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U05](U06){$(2, 6)$};
+        \draw node [Vertex,fill=RoyalBlue!30,right of=U06](U07){$(4, 7)$};
+        \draw node [Vertex,fill=RoyalBlue!30,right of=U07](U08){$(1, 8)$};
         \draw [thick,->,>=latex']
         (U01.north) .. controls ($ (U01.north) + (0,2.25) $)
+            and ($ (U02.north) + (0,2.25) $) .. (U02.north);
+        \draw [thick,->,>=latex']
+        (U04.north) .. controls ($ (U04.north) + (0,2.25) $)
             and ($ (U03.north) + (0,2.25) $) .. (U03.north);
         \draw [thick,->,>=latex']
-        (U02.south) .. controls ($ (U02.south) + (0,-2.25) $)
-            and ($ (U06.south) + (0,-2.25) $) .. (U06.south);
+        (U07.north) .. controls ($ (U07.north) + (0,2.5) $)
+            and ($ (U05.north) + (0,2.5) $) .. (U05.north);
         \draw [thick,->,>=latex']
-        (U04.north) .. controls ($ (U04.north) + (0,2.5) $)
-            and ($ (U07.north) + (0,2.5) $) .. (U07.north);
-        \draw [thick,->,>=latex']
-        (U05.south) .. controls ($ (U05.south) + (0,-2.5) $)
-            and ($ (U08.south) + (0,-2.5) $) .. (U08.south);
-        \draw [thick,->,>=latex']
-        (U11.north) .. controls ($ (U11.north) + (0,2.5) $)
-            and ($ (U09.north) + (0,2.5) $) .. (U09.north);
-        \draw [thick,->,>=latex']
-        (U12.north) .. controls ($ (U12.north) + (0,2.5) $)
-            and ($ (U10.north) + (0,2.5) $) .. (U10.north);
+        (U08.north) .. controls ($ (U08.north) + (0,2.5) $)
+            and ($ (U06.north) + (0,2.5) $) .. (U06.north);
     \end{tikzpicture}
@@
     Moreover, this shows that $\pi$ appears in the shuffle of the
-    words $\textcolor{RoyalBlue}{1892A4}$ and
-    $\textcolor{BrickRed}{37B5C6}$, both order isomorphic to
-    $145263$.
+    words $\textcolor{RoyalBlue}{3641}$ and
+    $\textcolor{BrickRed}{5872}$, both order-isomorphic to
+    $2431$.
   \end{exampleblock}

### 2016-04-08 slides/slides.tex
Modifs

@@
 \DeclareMathOperator{\Lb}{\mathtt{b}}
+\DeclareMathOperator{\Lc}{\mathtt{c}}
+\DeclareMathOperator{\Ld}{\mathtt{d}}
 
@@
 \title[Unshuffling Permutations]{%
-  \textbf{Unshuffling Permutations}
+  \huge \textbf{Unshuffling Permutations}
+  \vspace{1.5em}
 }
@@
 \subtitle{
-    \textbf{Samuele Giraudo}\quad\textbf{St\'ephane~Vialette}\\
-    Laboratoire d'Informatique Gaspard-Monge\\
+    \textbf{\underline{Samuele Giraudo}}\qquad\textbf{St\'ephane~Vialette}\\[1em]
+    {\footnotesize Laboratoire d'Informatique Gaspard-Monge\\
     Université Paris-Est Marne-la-Vallée \\
-    UMR CNRS 8049 \\
-    ~\\
-    \includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/CNRSfr}
+    UMR CNRS 8049 \\[2em]}
+    \includegraphics[scale=.16]{logos/CNRSfr}
     \qquad
-    \includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/ecole_ponts_RVB300.jpg}
+    \includegraphics[scale=.10]{logos/ecole_ponts_RVB300.jpg}
     \qquad
-    \includegraphics[width=1.5cm,height=2cm,keepaspectratio]{logos/LIGM}
+    \includegraphics[scale=.26]{logos/LIGM}
     \qquad
@@
     The \alert{shuffle product} $\shuffle$ on words of $A^*$ is defined
-    inductively by
+    recursively by \vspace{-.5em}
     \begin{align*}
-        u \shuffle \epsilon &= \epsilon \shuffle u = \{u\} \\
-        (ua \shuffle vb) &= (ua \shuffle v)b \cup (u \shuffle vb)a
+        u \shuffle \epsilon & := \{u\} =: \epsilon \shuffle u, \\
+        (ua \shuffle vb) & := (ua \shuffle v)b \cup (u \shuffle vb)a.
     \end{align*}
@@
 \begin{align*}
-    u \shuffle \epsilon & = \epsilon \shuffle u = u \\
-    (ua \shuffle vb) &= (ua \shuffle v)b + (u \shuffle vb)a
+    u \shuffle \epsilon & := u =: \epsilon \shuffle u, \\
+    (ua \shuffle vb) & := (ua \shuffle v)b + (u \shuffle vb)a.
 \end{align*}}
-
+%
 \uncover<3->{%
@@
 
-Given three words $u$, $v_1$, and $v_2$, deciding if $u$ is a
-shuffle of $v_1$ and $v_2$ can be done in
+\uncover<2->{%
+Given three words $u$, $v_1$, and $v_2$, deciding if $u$ is in
+$v_1 \shuffle v_2$ can be done in
 \begin{equation*}
@@
 time \BIB{van Leeuwen, Nivat}{1982}.
-\bigskip
-
+\bigskip}
+
+\uncover<3->{%
 Given a word $u$, deciding if there is a word $v$ such that $u$ is
 in $v \shuffle v$ is \NPC{} \BIB{Rizzi, Vialette}{2013}
-\BIB{Buss, Soltys}{2014}.
+\BIB{Buss, Soltys}{2014}.}
 \end{frame}
@@
 
+\uncover<2->{%
 \begin{block}{Example}
@@
     \begin{math}
-        u := \textcolor{RoyalBlue}{3}\textcolor{BrickRed}{3}
-        \textcolor{RoyalBlue}{12}\textcolor{BrickRed}{1}
-        \textcolor{RoyalBlue}{2}\textcolor{BrickRed}{22}
+        u := \textcolor{RoyalBlue}{\Lc}\textcolor{BrickRed}{\Lc}
+        \textcolor{RoyalBlue}{\La\Lb}\textcolor{BrickRed}{\La}
+        \textcolor{RoyalBlue}{\Lb}\textcolor{BrickRed}{\Lb\Lb}
     \end{math}
-    is a square since $u$ can be obtained by shuffling $3122$ with itself.
-\end{block}
-\medskip
-
+    is a square since $u$ can be obtained by shuffling $\Lc\La\Lb\Lb$
+    with itself.
+\end{block}
+\medskip}
+
+\uncover<3->{%
 The first numbers of square binary words of length $2n$ are
@@
     1, 2, 6, 22, 82, 320, 1268, 5102, 20632, 83972.
-\end{equation*}
-
+\end{equation*}}
+
+\uncover<4->{%
 \begin{block}{Open problem}
     Enumeration of square (binary) words.
-\end{block}
+\end{block}}
 \end{frame}
@@
 
+\uncover<2->{%
 \begin{block}{Example}
-    The word $21212233$ admits the perfect matching \vspace{-1em}
+    The word $\Lb\La\Lb\La\Lb\Lb\Lc\Lc$ admits the perfect matching \vspace{-1em}
     \begin{equation*}
@@
         \tikzstyle{Vertex}=[draw,circle,minimum size=.35cm,font=\footnotesize]
-        \draw node[Vertex,fill=RoyalBlue!30](U01){$(2, 1)$};
-        \draw node[Vertex,fill=RoyalBlue!30,right of=U01](U02) {$(1, 2)$};
-        \draw node[Vertex,fill=RoyalBlue!30,right of=U02](U03){$(2, 3)$};
-        \draw node [Vertex,fill=BrickRed!30,right of=U03](U04) {$(1, 4)$};
-        \draw node [Vertex,fill=BrickRed!30,right of=U04](U05) {$(2, 5)$};
-        \draw node [Vertex,fill=BrickRed!30,right of=U05](U06){$(2, 6)$};
-        \draw node [Vertex,fill=RoyalBlue!30,right of=U06](U07){$(3, 7)$};
-        \draw node [Vertex,fill=BrickRed!30,right of=U07](U08){$(3, 8)$};
+        \draw node[Vertex,fill=RoyalBlue!30](U01){$(\Lb, 1)$};
+        \draw node[Vertex,fill=RoyalBlue!30,right of=U01](U02) {$(\La, 2)$};
+        \draw node[Vertex,fill=RoyalBlue!30,right of=U02](U03){$(\Lb, 3)$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U03](U04) {$(\La, 4)$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U04](U05) {$(\Lb, 5)$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U05](U06){$(\Lb, 6)$};
+        \draw node [Vertex,fill=RoyalBlue!30,right of=U06](U07){$(\Lc, 7)$};
+        \draw node [Vertex,fill=BrickRed!30,right of=U07](U08){$(\Lc, 8)$};
         \draw [thick,>=latex']
@@
     \vspace{-1.5em}
-\end{block}
-
+\end{block}}
+
+\uncover<3->{%
 \begin{block}{Definition}
@@
     $(u_k, k) \-- (u_\ell, \ell)$ of $E$ such that $i < k < \ell < j$.
-\end{block}
+\end{block}}
 \end{frame}
@@
 \end{block}
-%\medskip
-
+
+\uncover<2->{%
 \begin{block}{Example}
-    The square word
+    The word
     \begin{math}
-        \textcolor{RoyalBlue}{3}\textcolor{BrickRed}{3}
-        \textcolor{RoyalBlue}{12}\textcolor{BrickRed}{1}
-        \textcolor{RoyalBlue}{2}\textcolor{BrickRed}{22}
+        \textcolor{RoyalBlue}{\Lc}\textcolor{BrickRed}{\Lc}
+        \textcolor{RoyalBlue}{\La\Lb}\textcolor{BrickRed}{\La}
+        \textcolor{RoyalBlue}{\Lb}\textcolor{BrickRed}{\Lb\Lb}
     \end{math}
-    admits the perfect matching \vspace{-1.5em}
+    admits the perfect matching \vspace{-1.25em}
     \begin{equation*}
         \begin{split}
-        \begin{tikzpicture}[xscale=.5,yscale=.2,inner sep=1pt,node distance=1.5cm]
+        \begin{tikzpicture}[xscale=.5,yscale=.16,inner sep=1pt,node distance=1.5cm]
             \tikzstyle{Vertex}=[draw,circle,minimum size=.35cm,font=\footnotesize]
-            \draw node[Vertex,fill=RoyalBlue!30](U01){$(3, 1)$};
-            \draw node[Vertex,fill=BrickRed!30,right of=U01](U02) {$(3, 2)$};
-            \draw node[Vertex,fill=RoyalBlue!30,right of=U02](U03){$(1, 3)$};
-            \draw node[Vertex,fill=RoyalBlue!30,right of=U03](U04) {$(2, 4)$};
-            \draw node[Vertex,fill=BrickRed!30,right of=U04](U05) {$(1, 5)$};
-            \draw node[Vertex,fill=RoyalBlue!30,right of=U05](U06){$(2, 6)$};
-            \draw node[Vertex,fill=BrickRed!30,right of=U06](U07){$(2, 7)$};
-            \draw node[Vertex,fill=BrickRed!30,right of=U07](U08){$(2, 8)$};
+            \draw node[Vertex,fill=RoyalBlue!30](U01){$(\Lc, 1)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U01](U02) {$(\Lc, 2)$};
+            \draw node[Vertex,fill=RoyalBlue!30,right of=U02](U03){$(\La, 3)$};
+            \draw node[Vertex,fill=RoyalBlue!30,right of=U03](U04) {$(\Lb, 4)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U04](U05) {$(\La, 5)$};
+            \draw node[Vertex,fill=RoyalBlue!30,right of=U05](U06){$(\Lb, 6)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U06](U07){$(\Lb, 7)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U07](U08){$(\Lb, 8)$};
             \draw [thick,>=latex']
@@
 
-    \vspace{-1.5em}
-    which is inclusion-free.
-\end{block}
-%\medskip
-
+    \vspace{-1.25em}
+    which is inclusion-free and hence is a square.
+\end{block}}
+
+\uncover<3->{%
 \begin{block}{Example}
-    The word $1221$ is not a square. Its only perfect matching\vspace{-1.5em}
+    The word $\La\Lb\Lb\La$ is not a square. Its only perfect matching\vspace{-1.25em}
     \begin{equation*}
         \begin{split}
-        \begin{tikzpicture}[xscale=.5,yscale=.3,inner sep=1pt,node distance=1.5cm]
+        \begin{tikzpicture}[xscale=.5,yscale=.2,inner sep=1pt,node distance=1.5cm]
             \tikzstyle{Vertex}=[draw,circle,minimum size=.35cm,font=\footnotesize]
-            \draw node[Vertex,fill=RoyalBlue!30](U01){$(1, 1)$};
-            \draw node[Vertex,fill=RoyalBlue!30,right of=U01](U02) {$(2, 2)$};
-            \draw node[Vertex,fill=BrickRed!30,right of=U02](U03){$(2, 3)$};
-            \draw node [Vertex,fill=BrickRed!30,right of=U03](U04) {$(1, 4)$};
+            \draw node[Vertex,fill=RoyalBlue!30](U01){$(\La, 1)$};
+            \draw node[Vertex,fill=RoyalBlue!30,right of=U01](U02) {$(\Lb, 2)$};
+            \draw node[Vertex,fill=BrickRed!30,right of=U02](U03){$(\Lb, 3)$};
+            \draw node [Vertex,fill=BrickRed!30,right of=U03](U04) {$(\La, 4)$};
             \draw [thick,>=latex']
@@
 
-    \vspace{-1.5em}
+    \vspace{-1.25em}
     is not inclusion-free.
-\end{block}
+\end{block}}
 \end{frame}
@@
 Let $S_n$ be the set of permutations of size $n$.
-\medskip
-
+\bigskip
+
+\uncover<2->{%
 There are several products on permutations, analogs of the shuffle
@@
 \end{itemize}
-\medskip
-
+\bigskip}
+
+\uncover<3->{%
 \begin{block}{Main questions}
@@
     \end{itemize}
-\end{block}
+\end{block}}
 \end{frame}
@@
 \begin{frame} \frametitle{Example: unshuffling binary words}
-Let $B$ the set of all binary words and let us consider the coproduct
+Let $B := \{\La, \Lb\}^*$ be the set of all binary words and consider
+the coproduct
 \begin{equation*}
@@
 
-\begin{block}{Example}
-\begin{equation*}
-    \Delta(100) =
-    \epsilon \otimes 100 + 1 \otimes 00
-    + 2 (0 \otimes 10) + 2 (10 \otimes 0) + 00 \otimes 1
-    + 100 \otimes \epsilon
-\end{equation*}
-\end{block}
+\uncover<2->{%
+\begin{block}{Example}\vspace{-1em}
+\begin{equation*}
+    \Delta(\Lb\La\La) =
+    \epsilon \otimes \Lb\La\La + \Lb \otimes \La\La
+    + 2 (\La \otimes \Lb\La) + 2 (\Lb\La \otimes \La) + \La\La \otimes \Lb
+    + \Lb\La\La \otimes \epsilon
+\end{equation*}
+\end{block}}
 \end{frame}
@@
 \begin{equation*}
-    \textcolor{RoyalBlue}{\Delta(100)} =
-    \epsilon \otimes 100 + 1 \otimes 00
-    + \textcolor{Orange}{2} (\textcolor{BrickRed}{0 \otimes 10})
-    + 2 (10 \otimes 0) + 00 \otimes 1
-    + 100 \otimes \epsilon,
-\end{equation*}
-we observe that $\textcolor{BrickRed}{0 \otimes 10} $ has multiplicity
-$\textcolor{Orange}{2}$ in $\textcolor{RoyalBlue}{\Delta(100)}$.
-\medskip
-
-By duality, this implies that $\textcolor{RoyalBlue}{100}$ has
+    \textcolor{RoyalBlue}{\Delta(\Lb\La\La)} =
+    \epsilon \otimes \Lb\La\La + \Lb \otimes \La\La
+    + \textcolor{Orange}{2} (\textcolor{BrickRed}{\La \otimes \Lb\La})
+    + 2 (\Lb\La \otimes \La) + \La\La \otimes \Lb
+    + \Lb\La\La \otimes \epsilon,
+\end{equation*}
+we observe that $\textcolor{BrickRed}{\La \otimes \Lb\La}$ has
 multiplicity $\textcolor{Orange}{2}$ in
-$\textcolor{BrickRed}{0 \shuffle 10}$. Indeed,
-\begin{equation*}
-    \textcolor{BrickRed}{0 \shuffle 10} =
-    010 + \textcolor{Orange}{2} (\textcolor{RoyalBlue}{100}).
-\end{equation*}
-\end{block}
+$\textcolor{RoyalBlue}{\Delta(\Lb\La\La)}$.
+\bigskip
+
+\uncover<3->{%
+By duality, this implies that $\textcolor{RoyalBlue}{\Lb\La\La}$ has
+multiplicity $\textcolor{Orange}{2}$ in
+$\textcolor{BrickRed}{\La \shuffle \Lb\La}$. Indeed,
+\begin{equation*}
+    \textcolor{BrickRed}{\La \shuffle \Lb\La} =
+    \La\Lb\La + \textcolor{Orange}{2}\, \textcolor{RoyalBlue}{\Lb\La\La}.
+\end{equation*}}
+\vspace{-1.5em}
+\end{block}}
 \end{frame}
@@
 \end{block}
-\medskip
-
+\medskip}
+
+\uncover<3->{%
 \begin{block}{Definition}
@@
     \begin{equation*}
-        \Delta(\pi) =
+        \Delta(\pi) :=
         \sum_{P \sqcup Q = \{1, \dots, |\pi|\}}
@@
     \end{equation*}
-\end{block}
-
+\end{block}}
+
+\uncover<4->{%
 \begin{block}{Example}
@@
 \end{equation*}
-\end{block}
+\end{block}}
 \end{frame}
@@
 
+\uncover<2->{%
 \begin{block}{Proposition}
@@
 \end{block}
-\medskip
-
-\begin{block}{Example}
+\medskip}
+
+\uncover<3->{%
+\begin{block}{Example}\vspace{-1em}
 \begin{equation*}\begin{split}
@@
 \end{split}\end{equation*}
-\end{block}
+\end{block}}
 \end{frame}
@@
 \end{block}
-\medskip
-
+\medskip}
+
+\uncover<3->{%
 \begin{block}{Example}
@@
     \end{math}.
-\end{block}
+\end{block}}
 \end{frame}
@@
 \begin{frame} \frametitle{Some properties of square permutations}
-The first numbers of square permutations of size $2$ are
+The first numbers of square permutations of size $2n$ are
 \begin{equation*}
@@
 \end{equation*}
-%\medskip
-
+
+\uncover<2->{%
 \begin{block}{Proposition}
@@
 \end{block}
-\medskip
-
+\medskip}
+
+\uncover<3->{%
 \begin{block}{Proposition}
@@
     $213$ and~$231$.
-\end{block}
+\end{block}}
 \end{frame}
@@
   \end{definition}
-
-  \medskip
-
-  \begin{exampleblock}{Example}
-
+  \medskip
+
+\uncover<2->{%
+  \begin{block}{Example}
         The permutation
@@
         $\STD(\textcolor{BrickRed}{91674}) = 51342$.
-  \end{exampleblock}
-  \medskip
-
+  \end{block}
+  \medskip}
+
+\uncover<3->{%
   \begin{block}{Definition}
@@
   \end{block}
-  \medskip
-
+  \medskip}
+
+\uncover<4->{%
   \begin{block}{Proposition \BIB{Bose, Buss, Lubiw}{1998}}
     The pattern involvement problem is $\NPC$.
-  \end{block}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\begin{frame}
-  %\frametitle{Square permutation}
-
-  %\begin{definition}
-    %A permutation is said to be a \emph{square} if it can be obtained by
-    %shuffling two order-isomorphic patterns
-  %\end{definition}
-
-  %\medskip
-
-  %\begin{exampleblock}{Example}
-    %\begin{itemize}
-      %\item
-        %The permutation $\pi = 25167834$ is a square since it can obtained by
-        %shuffling two order-isomorphic (\emph{i.e.} $1342$)
-        %patterns $1683$ and $2574$ as can be seen in the highlighted subsequence of
-        %$\pi = \mathbf{25}16\mathbf{7}83\mathbf{4}$.
-
-      %\medskip
-
-      %\item
-       %$\pi = 4123$ and $1432$ are not squares.
-    %\end{itemize}
-  %\end{exampleblock}
-
-%\end{frame}
+  \end{block}}
+\end{frame}
 
@@
 
+\uncover<2->{%
   \begin{definition}[Property $\mathbf{P_1}$]
@@
     \InclusionRR, \CrossingLR, and \CrossingRL.
-  \end{definition}
-
-  \medskip
-
+  \end{definition}}
+
+  \medskip
+
+\uncover<3->{%
   \begin{definition}[Property $\mathbf{P_2}$]
@@
     $\pi_j < \pi_\ell$.
-  \end{definition}
+  \end{definition}}
 \end{frame}
@@
   \onslide<2->
-  \begin{exampleblock}{Example}
+  \begin{block}{Example}
     The permutation $\pi := 35867241$ is a square since it admits the
@@
     $2431$.
-  \end{exampleblock}
+  \end{block}
 \end{frame}
@@
 
+\uncover<2->{%
   \begin{block}{Main ideas of the proof}
@@
 
+        \uncover<3->{%
         \item For this, given two permutations $\sigma$ and $\pi$, we
@@
         time from $\sigma$ and $\pi$.
-        \medskip
-
+        \medskip}
+
+        \uncover<4->{%
         \item To prove that $\mu_{\sigma, \pi}$ is a square, we show
         that $\mu_{\sigma, \pi}$ admits a perfect matching satisfying
-        $\mathbf{P_1}$ and~$\mathbf{P_2}$.
-
+        $\mathbf{P_1}$ and~$\mathbf{P_2}$.}
     \end{itemize}
-  \end{block}
+  \end{block}}
 \end{frame}
@@
 
+    \uncover<2->{%
   \begin{block}{Algorithms}
-
     Given two permutations $\pi$ and~$\sigma$, how hard is the problem of
@@
   \end{block}
-  \medskip
-
+  \medskip}
+
+    \uncover<3->{%
   \begin{block}{Algebra}
@@
         combinatorial and algorithmic properties of square permutations?
-  \end{block}
-
+  \end{block}}
 \end{frame}

### 2016-04-08 slides/slides.tex
Simplifications

@@
 \begin{equation*}
-    \QQ[A^*] \otimes \QQ[A^*] \to \QQ[A^*]
+    \shuffle : \QQ[A^*] \otimes \QQ[A^*] \to \QQ[A^*]
 \end{equation*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Combinatorial products and structure coefficients}
+\begin{frame} \frametitle{Combinatorial products and coproducts}
 Let $C$ be a set of combinatorial objects and $\QQ[C]$ be the
 $\QQ$-linear span of $C$.
-\medskip
-
-A binary linear product
-\begin{equation*}
-    \cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C].
-\end{equation*}
-on $\QQ[C]$ describes a way to \alert{combine} two objects of
-$C$ to obtain \alert{bigger} objects.
-\medskip
-
-In a general way, for any $x, y \in C$,
-\begin{equation*}
-    x \cdot y = \sum_{z \in C} \lambda_{x, y}^z \; z.
-\end{equation*}
-\medskip
-
-A product is therefore wholly encoded by the coefficients
-$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
-\alert{structure coefficients}.
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Combinatorial coproducts}
-In a dual way, a linear coproduct
-\begin{equation*}
-    \Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C].
-\end{equation*}
-on $\QQ[C]$ describe a way to \alert{break} one object of $C$ into
-\alert{two smaller} pieces.
-\medskip
-
-In a general way, for any $z \in C$,
-\begin{equation*}
-    \Delta(z) = \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
-\end{equation*}
-\medskip
-
-A coproduct is therefore wholly encoded by the coefficients
-$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
-\alert{structure coefficients}.
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Equivalence between products and coproducts}
+
+\uncover<2->{%
 \begin{block}{Key idea}
-    Knowing how to break a combinatorial object is equivalent
-    to knowing how to combine these.
-\end{block}
-\medskip
-
-Indeed, if $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$ is a product with
-structure coefficients $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$,
-we can define its \alert{dual coproduct} $\Delta$ by setting
-\begin{equation*}
-    \Delta(z) := \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
-\end{equation*}
-\medskip
-
-Conversely, if $\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$ is
-a coproduct with structure coefficients $\lambda_{x, y}^z \in \QQ$,
-$x, y, z \in C$, we can define its \alert{dual product} $\cdot$ by
-setting
-\begin{equation*}
-    x \cdot y := \sum_{z \in C} \lambda_{x, y}^z \; z.
-\end{equation*}
+    Knowing how to \alert{break} combinatorial objects explains
+    how to \alert{combine} these.
+\end{block}
+\medskip}
+
+\uncover<3->{%
+To define a product $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$,
+it is some cases more convenient to start by defining a \alert{coproduct}
+$\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$.
+\medskip}
+
+\uncover<4->{%
+Every coproduct fits into the general form
+\begin{equation*}
+    \Delta(z)
+    = \sum_{x, y \in C} \textcolor{RoyalBlue}{\lambda_{x, y}^z} \;
+    (x \otimes y)
+\end{equation*}
+where the $\textcolor{RoyalBlue}{\lambda_{x, y}^z}$ are structure
+coefficients of $\QQ$.
+\medskip}
+
+\uncover<5->{%
+Then, $\Delta$ leads by duality to the definition of $\cdot$ by
+\begin{equation*}
+    x \cdot y
+    := \sum_{z \in C} \textcolor{RoyalBlue}{\lambda_{x, y}^z} \; z.
+\end{equation*}}
 \end{frame}

### 2016-04-09 slides/slides.tex
Ameliorations slides

@@
 \usepackage{etoolbox}
+\usepackage[usestackEOL]{stackengine}
 \usepackage{tikz}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title[Unshuffling Permutations]{%
-  \huge \textbf{Unshuffling Permutations}
-  \vspace{1.5em}
-}
-
-\subtitle{
-    \textbf{\underline{Samuele Giraudo}}\qquad\textbf{St\'ephane~Vialette}\\[1em]
-    {\footnotesize Laboratoire d'Informatique Gaspard-Monge\\
+\author[Samuele Giraudo]{
+    \textbf{\underline{Samuele~Giraudo}} \qquad
+    \textbf{St\'ephane~Vialette}}
+\title[Unshuffling Permutations]{
+    {\huge \textbf{Unshuffling Permutations}}}
+\institute[LIGM]{
+    {\footnotesize Laboratoire d'Informatique Gaspard-Monge \\
     Université Paris-Est Marne-la-Vallée \\
-    UMR CNRS 8049 \\[2em]}
-    \includegraphics[scale=.16]{logos/CNRSfr}
-    \qquad
-    \includegraphics[scale=.10]{logos/ecole_ponts_RVB300.jpg}
-    \qquad
-    \includegraphics[scale=.26]{logos/LIGM}
-    \qquad
-    \raisebox{.25cm}{\includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/LogoEsieeParis}}
-    \qquad
-    \raisebox{.25cm}{\includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/UPEM_LOGO_EDITION}}
-}
-
-\date[LATIN 16]{
-    {
+    UMR CNRS 8049}}
+\date[LATIN 16]{{
     \scriptsize Latin American Theoretical Informatics Symposium \\
-    April $11$-$15$ $2016$
-    }
-}% end date
+    April $11$-$15$ $2016$}}
+
+%\subtitle{
+    %\includegraphics[scale=.16]{logos/CNRSfr}
+    %\qquad
+    %\includegraphics[scale=.10]{logos/ecole_ponts_RVB300.jpg}
+    %\qquad
+    %\includegraphics[scale=.26]{logos/LIGM}
+    %\qquad
+    %\raisebox{.25cm}{\includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/LogoEsieeParis}}
+    %\qquad
+    %\raisebox{.25cm}{\includegraphics[width=1.cm,height=1.25cm,keepaspectratio]{logos/UPEM_LOGO_EDITION}}
+%}
 
@@
 \begin{frame} \frametitle{Shuffling words}
-\begin{block}{Definition}
+\begin{block}{{\bf Definition} (shuffle product of words)}
     The \alert{shuffle product} $\shuffle$ on words of $A^*$ is defined
@@
 \uncover<3->{%
-\begin{block}{Example}
+\begin{block}{\bf Example}
     \vspace{-1em}
@@
 can be computed in
-\begin{equation*}
-    O\left((|v_1| + |v_2|) \; \binom{|v_1| + |v_2|}{|v_1|}\right)
-\end{equation*}
+\begin{footnotesize}
+\begin{equation*}
+    O\left((|v_1| + |v_2|) \; \parenVectorstack{|v_1| + |v_2| \\ |v_1|}\right)
+\end{equation*}
+\end{footnotesize}
 time \BIB{Spehner}{1986}.
 \bigskip
+\bigskip
+\bigskip
 
@@
 $v_1 \shuffle v_2$ can be done in
+\begin{footnotesize}
 \begin{equation*}
@@
 \end{equation*}
+\end{footnotesize}
 time \BIB{van Leeuwen, Nivat}{1982}.
+\bigskip
+\bigskip
 \bigskip}
@@
 \begin{frame} \frametitle{Square words}
-\begin{block}{Definition}
+\begin{block}{{\bf Definition} (square words)}
     A word $u$ is a \alert{square} (w.r.t. the shuffle product
@@
 \uncover<2->{%
-\begin{block}{Example}
+\begin{block}{\bf Example}
     The word
@@
 \begin{frame} \frametitle{Recognizing square words}
-\begin{block}{Definition}
+\begin{block}{{\bf Definition} (perfect matchings for words)}
     A \alert{perfect matching} of a word $u$ is a graph $(V, E)$ such
@@
 \uncover<2->{%
-\begin{block}{Example}
+\begin{block}{\bf Example}
     The word $\Lb\La\Lb\La\Lb\Lb\Lc\Lc$ admits the perfect matching \vspace{-1em}
@@
     \begin{split}
-    \begin{tikzpicture}[xscale=.5,yscale=.3,inner sep=1pt,node distance=1.5cm]
+    \begin{tikzpicture}[xscale=.5,yscale=.25,inner sep=1pt,node distance=1.5cm]
         \tikzstyle{Vertex}=[draw,circle,minimum size=.35cm,font=\footnotesize]
@@
 \uncover<3->{%
-\begin{block}{Definition}
+\begin{block}{{\bf Definition} (inclusion-free perfect matchings)}
     A perfect matching $(V, E)$ of a word $u$ is \alert{inclusion-free}
@@
 
-\begin{block}{Proposition \BIB{Rizzi, Vialette}{2013}}
+\begin{block}{{\bf Proposition} \BIB{Rizzi, Vialette}{2013}}
     A word $u$ is a square iff $u$ admits an inclusion-free perfect
@@
             \draw [thick,>=latex']
-            (U04.north) .. controls ($ (U04.north) + (0,4.5) $)
-                and ($ (U07.north) + (0,4.5) $) .. (U07.north);
-            \draw [thick,>=latex']
-            (U05.south) .. controls ($ (U05.south) + (0,-2.5) $)
-                and ($ (U08.south) + (0,-2.5) $) .. (U08.south);
+            (U04.south) .. controls ($ (U04.south) + (0,-2.25) $)
+                and ($ (U07.south) + (0,-2.25) $) .. (U07.south);
             \draw [thick,>=latex']
@@
 \uncover<3->{%
-\begin{block}{Example}
-    The word $\La\Lb\Lb\La$ is not a square. Its only perfect matching\vspace{-1.25em}
+\begin{block}{\bf Example}
+    The word $\La\Lb\Lb\La$ is not a square. Its only perfect matching\vspace{-1em}
     \begin{equation*}
@@
             \draw [thick,>=latex']
-            (U01.north) .. controls ($ (U01.north) + (0,2.5) $)
-                and ($ (U04.north) + (0,2.5) $) .. (U04.north);
+            (U01.north) .. controls ($ (U01.north) + (0,2) $)
+                and ($ (U04.north) + (0,2) $) .. (U04.north);
             \draw [thick,>=latex']
-            (U02.north) .. controls ($ (U02.north) + (0,1.5) $)
-                and ($ (U03.north) + (0,1.5) $) .. (U03.north);
+            (U02.south) .. controls ($ (U02.south) + (0,-1.5) $)
+                and ($ (U03.south) + (0,-1.5) $) .. (U03.south);
         \end{tikzpicture}\vspace{-1.5em}
@@
 \begin{frame} \frametitle{Main motivations}
-Let $S_n$ be the set of permutations of size $n$.
+Let $S_n$ be the set of \alert{permutations} of size $n$.
 \bigskip
@@
 To define a product $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$,
-it is some cases more convenient to start by defining a \alert{coproduct}
+it is in some cases more convenient to start by defining a \alert{coproduct}
 $\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$.
@@
 \end{equation*}
-where the $\textcolor{RoyalBlue}{\lambda_{x, y}^z}$ are structure
-coefficients of $\QQ$.
+where the $\textcolor{RoyalBlue}{\lambda_{x, y}^z} \in \QQ$ are structure
+coefficients.
 \medskip}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Example: unshuffling binary words}
-Let $B := \{\La, \Lb\}^*$ be the set of all binary words and consider
-the coproduct
-\begin{equation*}
-    \Delta : \QQ[B] \to \QQ[B] \otimes \QQ[B]
-\end{equation*}
-on binary words defined by
+\begin{frame} \frametitle{Example: unshuffling words}
+Consider the coproduct
+\begin{equation*}
+    \Delta : \QQ[A^*] \to \QQ[A^*] \otimes \QQ[A^*]
+\end{equation*}
+on words defined by
 \begin{equation*}
@@
 \uncover<2->{%
-\begin{block}{Example}\vspace{-1em}
-\begin{equation*}
-    \Delta(\Lb\La\La) =
-    \epsilon \otimes \Lb\La\La + \Lb \otimes \La\La
-    + 2 (\La \otimes \Lb\La) + 2 (\Lb\La \otimes \La) + \La\La \otimes \Lb
-    + \Lb\La\La \otimes \epsilon
-\end{equation*}
-\end{block}}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%o%%%%%%%%%%%%%%%%%%
-\begin{frame} \frametitle{Example: shuffling binary words}
-\begin{block}{Proposition \BIB{Reutenauer}{1993}}
-    The shuffle product of binary words is the dual product of the
-    unshuffling coproduct of binary words.
-\end{block}
-\medskip
-
-\uncover<2->{%
-\begin{block}{Example}
-Since
+\begin{block}{\bf Example}\vspace{-1.5em}
 \begin{equation*}
@@
     + 2 (\Lb\La \otimes \La) + \La\La \otimes \Lb
-    + \Lb\La\La \otimes \epsilon,
-\end{equation*}
-we observe that $\textcolor{BrickRed}{\La \otimes \Lb\La}$ has
+    + \Lb\La\La \otimes \epsilon
+\end{equation*}
+\end{block}}
+
+\uncover<3->{%
+\begin{block}{{\bf Proposition} \BIB{Reutenauer}{1993}}
+    The shuffle product of words is the dual product of the
+    unshuffling coproduct of words.
+\end{block}}
+
+\uncover<4->{%
+\begin{block}{\bf Example}
+Since $\textcolor{BrickRed}{\La \otimes \Lb\La}$ has
 multiplicity $\textcolor{Orange}{2}$ in
-$\textcolor{RoyalBlue}{\Delta(\Lb\La\La)}$.
-\bigskip
-
-\uncover<3->{%
-By duality, this implies that $\textcolor{RoyalBlue}{\Lb\La\La}$ has
-multiplicity $\textcolor{Orange}{2}$ in
-$\textcolor{BrickRed}{\La \shuffle \Lb\La}$. Indeed,
+$\textcolor{RoyalBlue}{\Delta(\Lb\La\La)}$, the coeff. of
+$\textcolor{RoyalBlue}{\Lb\La\La}$ is
+$\textcolor{Orange}{2}$ in
+$\textcolor{BrickRed}{\La \shuffle \Lb\La}$:\vspace{-.9em}
 \begin{equation*}
@@
     \La\Lb\La + \textcolor{Orange}{2}\, \textcolor{RoyalBlue}{\Lb\La\La}.
-\end{equation*}}
-\vspace{-1.5em}
+\end{equation*}
+\vspace{-1.8em}
 \end{block}}
@@
 \begin{frame} \frametitle{Unshuffling permutations}
-If $u$ is a word of integers without multiple occurrence of a same letter,
-the \alert{standardized} $\STD(u)$ of $u$ is the unique permutation of
-$S_{|u|}$ order-isomorphic to~$u$.
-
-\uncover<2->{%
-\begin{block}{Example}
+\begin{block}{{\bf Definition} (standardization of words)}
+    If $u$ is a word of integers without multiple occurrence of a same
+    letter, the \alert{standardized} $\STD(u)$ of $u$ is the unique
+    permutation of $S_{|u|}$ order-isomorphic to~$u$.
+\end{block}
+
+\uncover<2->{%
+\begin{block}{\bf Example}
+\vspace{-.5em}
 \begin{equation*}
@@
 \uncover<3->{%
-\begin{block}{Definition}
+\begin{block}{{\bf Definition} (unshuffling coproduct of permutations)}
     The \alert{unshuffling coproduct of permutations} is the
@@
 \uncover<4->{%
-\begin{block}{Example}
+\begin{block}{\bf Example}
+\vspace{-.5em}
 \begin{equation*}
@@
 \begin{frame} \frametitle{Shuffling permutations}
-\begin{block}{Definition}
+\begin{block}{{\bf Definition} (shuffling product of permutations)}
     The \alert{shuffling product of permutations} (supershuffle) is the
@@
 \uncover<2->{%
-\begin{block}{Proposition}
+\begin{block}{\bf Proposition}
     The permutations appearing in $\sigma \SHUFFLE \nu$ are exactly the
@@
 \uncover<3->{%
-\begin{block}{Example}\vspace{-1em}
+\begin{block}{\bf Example}\vspace{-1em}
 \begin{equation*}\begin{split}
-    12 \SHUFFLE 21 & =
-    1243 + 1324 + 2 (1342) + 2 (1423) + 3 (1432) +
-    2134 + 2 (2314) \\[.5em]
+    \textcolor{RoyalBlue}{12} \SHUFFLE \textcolor{BrickRed}{21} & =
+    \textcolor{RoyalBlue}{12}\textcolor{BrickRed}{43} +
+    \textcolor{RoyalBlue}{1}\textcolor{BrickRed}{32}\textcolor{RoyalBlue}{4}
+    + 2 (\textcolor{RoyalBlue}{1}34\textcolor{BrickRed}{2}) +
+    2 (1423) + 3 (1432) +
+    2134 + 2 (2314) \\[.4em]
     & + 3 (2341) + 2413 + 2 (2431) + 2 (3124) + 3142 +
-    3 (3214) + 2 (3241) \\[.5em]
+    3 (3214) + 2 (3241) \\[.4em]
     & + 3421 + 3 (4123) + 2 (4132) + 2 (4213) + 4231 + 4312
@@
 \begin{frame} \frametitle{Square permutations}
-\begin{block}{Definition}
+\begin{block}{{\bf Definition} (square permutations)}
     A permutation $\pi$ is a \alert{square} (w.r.t. the shuffle product
@@
 \uncover<2->{%
-\begin{block}{Proposition}
+\begin{block}{\bf Proposition}
     A permutation $\pi$ is a square iff $\pi$ can be obtained by
@@
 \uncover<3->{%
-\begin{block}{Example}
+\begin{block}{\bf Example}
     The permutation
@@
 \uncover<2->{%
-\begin{block}{Proposition}
+Square permutations are compatible with some involutions on permutations:
+\begin{block}{\bf Proposition}
     Let $\pi$ be a square permutation and $\sigma$ be a square root of
@@
 \uncover<3->{%
-\begin{block}{Proposition}
+There is a link with square binary words:
+\begin{block}{\bf Proposition}
     The set of binary words of length $n$ that are square w.r.t.
@@
 \begin{frame}
-\frametitle{Permutation patterns}
-
-  \begin{definition}
+\frametitle{Pattern involvement problem}
+
+  \begin{block}{{\bf Definition} (permutation patterns)}
     A permutation $\pi$ \alert{contains} a permutation $\sigma$ if there
@@
     Otherwise, $\pi$ \alert{avoids} $\sigma$.
-  \end{definition}
-  \medskip
-
-\uncover<2->{%
-  \begin{block}{Example}
+  \end{block}
+
+\uncover<2->{%
+  \begin{block}{\bf Example}
         The permutation
-        $\pi = 3\textcolor{BrickRed}{91}8\textcolor{BrickRed}{674}52$
+        $\pi = 3\textcolor{RoyalBlue}{91}8\textcolor{RoyalBlue}{674}52$
         contains the pattern $\sigma := 51342$ since
-        $\STD(\textcolor{BrickRed}{91674}) = 51342$.
+        $\STD(\textcolor{RoyalBlue}{91674}) = 51342$.
   \end{block}
@@
 \uncover<3->{%
-  \begin{block}{Definition}
+  \begin{block}{{\bf Definition} (pattern involvement problem)}
     The \alert{pattern involvement problem} is the decision problem
-    consisting in, given $\pi, \sigma \in S$, decide whether $\sigma \leq \pi$.
-  \end{block}
-  \medskip}
+    consisting in, given $\pi, \sigma \in S$, decide
+    whether~$\sigma \leq \pi$.
+  \end{block}}
 
 \uncover<4->{%
-  \begin{block}{Proposition \BIB{Bose, Buss, Lubiw}{1998}}
+  \begin{block}{{\bf Proposition} \BIB{Bose, Buss, Lubiw}{1998}}
     The pattern involvement problem is $\NPC$.
@@
 
-    \begin{block}{Definition}
+    \begin{block}{{\bf Definition} (perfect matchings for permutations)}
         A \alert{perfect matching} of a permutation $\pi \in S_n$ is a
@@
 \uncover<2->{%
-  \begin{definition}[Property $\mathbf{P_1}$]
+  \begin{block}{{\bf Definition} (property $\mathbf{P_1}$)}
     Let $\pi \in S$. A perfect matching $(V, A)$ of $\pi$ has property
@@
     \InclusionRR, \CrossingLR, and \CrossingRL.
-  \end{definition}}
-
-  \medskip
-
-\uncover<3->{%
-  \begin{definition}[Property $\mathbf{P_2}$]
+  \end{block}}
+  \medskip
+
+\uncover<3->{%
+  \begin{block}{{\bf Definition} (property $\mathbf{P_2}$)}
     Let $\pi \in S$. A perfect matching $(V, A)$ of $\pi$ has property
@@
     $\pi_j < \pi_\ell$.
-  \end{definition}}
+  \end{block}}
 \end{frame}
@@
 
-  \begin{block}{Proposition}
+  \begin{block}{\bf Proposition}
     A permutation $\pi$ is a square (w.r.t. the shuffle product $\SHUFFLE$)
@@
   \onslide<2->
-  \begin{block}{Example}
+  \begin{block}{\bf Example}
     The permutation $\pi := 35867241$ is a square since it admits the
@@
     \begin{split}
-    \begin{tikzpicture}[scale=0.35,inner sep=2pt,node distance=1cm]
+    \begin{tikzpicture}[xscale=0.35,yscale=.3,inner sep=2pt,node distance=1.1cm]
         \tikzstyle{Vertex}=[draw,circle,minimum size=.45cm,font=\footnotesize]
@@
 
-  \begin{block}{Proposition}
+  \begin{block}{\bf Proposition}
     Deciding whether a permutation is a square is \NPC.
@@
         \item We show that the pattern involvement problem is reducible
-        in polynomial time to our decision problem.
+        in polynomial time to the problem of deciding if a permutation
+        is a square.
         \medskip
@@
   \begin{block}{Combinatorics}
-    \begin{itemize}
-        \item Enumeration of square permutations?
-        \smallskip
-
-        \item Enumeration of square permutations avoiding $213$ and $231$?
-    \end{itemize}
+        Enumeration of square permutations?
   \end{block}

### 2016-04-09 slides/slides.tex
Ameliorations slides

@@
 \uncover<4->{%
-\begin{block}{Open problem}
+\begin{block}{Open problem \BIB{Henshall, Rampersad, Shallit}{2012}}
     Enumeration of square (binary) words.
@@
         \uncover<4->{%
-        \item To prove that $\mu_{\sigma, \pi}$ is a square, we show
-        that $\mu_{\sigma, \pi}$ admits a perfect matching satisfying
+        \item We use the fact that $\mu_{\sigma, \pi}$ is a square iff
+        $\mu_{\sigma, \pi}$ admits a perfect matching satisfying
         $\mathbf{P_1}$ and~$\mathbf{P_2}$.}
